### 2016-09-26 latex/article.Rnw
Rebirth

\documentclass[article]{jss}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% almost as usual
\author{Achim Zeileis\\Universit\"at Innsbruck \And
        Second Author\\Plus Affiliation}
\title{A Capitalized Title: Something about a Package \pkg{cleanR}}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Achim Zeileis, Second Author} %% comma-separated
\Plaintitle{A Capitalized Title: Something about a Package cleanR} %% without formatting
\Shorttitle{\pkg{cleanR}: A Capitalized Title} %% a short title (if necessary)

%% an abstract and keywords
\Abstract{
  The abstract of the article.
}
\Keywords{keywords, comma-separated, not capitalized, \proglang{Java}}
\Plainkeywords{keywords, comma-separated, not capitalized, Java} %% without formatting
%% at least one keyword must be supplied

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}
%% \Acceptdate{2012-06-04}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Achim Zeileis\\
  Department of Statistics and Mathematics\\
  Faculty of Economics and Statistics\\
  Universit\"at Innsbruck\\
  6020 Innsbruck, Austria\\
  E-mail: \email{Achim.Zeileis@uibk.ac.at}\\
  URL: \url{http://eeecon.uibk.ac.at/~zeileis/}
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/512/507-7103
%% Fax: +43/512/507-2851

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\section{Introduction}
Statisticians and data analysts spend a large portion of their time on
data cleaning and on data wrangling. Packages such as \pkg{}, \pkg{},
and \pkg{} make data wrangling a lot easier in R, but there are only a
few options available for data cleaning.

Data cleaning is a time consuming endeavour and it inherently requires
human interaction since every dataset is different and the variables
in the dataset can only be understood in the proper context of the
experiment. While each dataset is different and require
... (``simple questions''

... there are a number of common occurrences that .... This is particulary
Automated

... partly be done by ... However ... enough ... similar tasks that
pop

However it is not uncommon for ... to run a set of ....

\section{Checking a dataset for errors}

%% include your article here, just as usual
%% Note that you should use the \pkg{}, \proglang{} and \code{}
%% commands.

<<>>=
library(cleanR)
data(testData)
testData
@

\section[About Java]{About \proglang{Java}}
%% Note: If there is markup in \(sub)section, then it has to be escape as above.

\end{document}

### 2016-09-26 R/check.R

@@ -1,0 +1,448 @@
+#Functions called from clean.R for performing cleaning steps/checks
+#TO DO: structure: functions that are to check whether a problem is present
+#             for each variable return a checkRes object, consisting of a
+#             boolean and a message text (possibly empty) (=> define class checkRes)
+#TO DO: Consider: Should the "v <- na.omit(v)" command at the beginning of
+#       some (all?) checking functions be moved to the check-functions
+#       themselves? I.e. do we ever need to consider missing values when checking?
+#   And similarly: Should the labelled-vector unpacking also just be done once?
+#TO DO: What should check() output? Maybe a list of checkRes-objects, and then
+#       I should also implement a print() method for checkRes-lists to be used in clean()
+
+#General comments:
+# arguments: v is a variable (column) from a data.frame, i.e. a vector
+# note: labelled is Wickham's class from the haven/hmisc packages
+# note: Function names are suffixed with an indicator of the data types it
+#       is to be called on. N is numerical, I is integer, F is factor,
+#       L is labeled (Wickham), C is character. The ordering of suffix
+#       letters is arbitrary. All suffixed functions are purely internal
+#       and their calling is controlled using s3 methods
+
+
+##########################################
+
+####Check#####
+
+#' Perform a check of potential errors in a data frame
+#'
+#' Runs a set of validation checks to check a vector for potential errors.  performs checking steps according to user input and/or data type of the inputted variable.
+#'
+#' @param v the vector to be checked
+#' @param \dots other arguments that are passed on the to checking functions
+#' @return An list object of class "checked" summarizing the result from the check of the check. As a minimum the returned class should contain the following elements
+#' \itemize{
+#'   \item{"name"}{The name of the check}
+#'   \item{"description"}{Slightly more information about the check}
+#'   \item{"problem"}{An integer giving an error code. 0 means no potential errors were identified}
+#'   \item{"message"}{A string giving summary information in R markdown format about the results}
+#' }
+#' @author Anne Helby Petersen \email{ahpe@@sund.ku.dk} and Claus Thorn Ekstrom \email{ekstrom@@sund.ku.dk}
[…]

### 2016-09-26 R/clean.R

@@ -1,0 +1,493 @@
+#' Perform a check of potential errors in a data frame
+#'
+#' Runs a set of validation checks to check the variables in a data frame for potential errors.
+#' Performs checking steps according to user input and/or data type of the inputted variable.
+#'
+#' @param standAlone If TRUE, the document begins with a markdown preamble such that it
+#' can be rendered as is.
+#' @param brag If TRUE, a note about cleanR is appended at the end of the output document.
+#' @param ordering Choose the ordering of the variables in the data presentation. The options
+#' are "asIs" (ordering as in the dataset) and "alpha" (alphabetical order).
+#' @param cleanUp - not done yet -.
+#' @param output Output format, options are pdf and html
+#' @param finish "render" (makes pdf/html), "markdown" (makes markdown file), "print" (prints to screen).
+#' @param twoCol Should the results be presented in two columns (if finish is "render" or "markdown")?
+#' @param o the data frame to be checked
+#' @param characterChecks a list of error-checking functions to apply to character vectors
+#' @param integerChecks a list of error-checking functions to apply to integer vectors
+#' @param silent Should clean() run completely silently? Note that this option overrules the settings for
+#' "quiet": A silent session is always quiet.
+#' @param openResult If TRUE, the file produced by clean() is automatically opened by the end of
+#' the function run.
+#' @param mode Vector of tasks to perform among the three categories "summarize", "visualize" and "check".
+#' Note that... SOMETHING ABOUT HOW THE FUNCTIONS CALLED IN EACH PART ARE CONTROLLED.
+#' @param useVar Variables to clean and present results for. Either a list of variable names (as used in
+#' the data.frame o) or one of two special options; "all" (every variable is cleaned and presented) or
+#' "problematic" (only variables yielding problems from the checking function are presented)
+#' @param nagUser Remove at some point
+#' @param smartNum If TRUE, numeric and integer variables with less than maxLevels (defaults to 5) unique
+#' values are treated as factor variables in the checking, visualization and summary functions. A
+#' message is printed in the data summary as well.
+#' @param preChecks Variable checks that are performed before the summary/visualization/checking step. If
[…]

### 2016-09-26 R/makeData.R

@@ -1,0 +1,35 @@
+#make test dataset
+
+vC <- c("a", "b", "c", "a", "b", "d", "a")
+vF <- as.factor(vC)
+vN <- c(1:10, 1, 1, 1, 5, 5)
+vI <- as.integer(vN)
+vB <- c(T, F, T, T, T, F) 
+
+vCPR <- rep(c("010101-1111", "020102-2929", "121201-1902", "030729-2222",
+              "080909-1212"), 3)
+
+vCPRkey <- c(vCPR[1:5], "020202-0101", "030303-0101", "040404-0101", "050505-0101",
+             "060606-0101", "070707-0101", "080808-0101", "090909-0101", 
+             "020202-0202", "030303-0202")
+vMiss <- c(".", "", "nan", "NaN", "NAN", "na", "NA", "Na", "Inf", "inf",
+             "-Inf", "-inf", "-", "9", "9") #there are more missing strings, add them 
+                                            #when data is expanded to have more obs.
+
+data <- data.frame(charVar=c(rep(vC, 2), NA),
+                   factorVar=factor(c(rep(vC, 2), "999")),
+                   numVar=vN, intVar=vI, boolVar=c(rep(vB, 2), rep(NA, 3)),
+                   keyVar=as.character(1:15),
+                   emptyVar=rep(1, 15),
+                   joeVar = 1:15,
+                   jack__var = 1:15,
+                   numOutlierVar = c(1:14,100),
+                   smartNumVar = c(rep(0, 7), rep(1, 8)),
+                   cprVar=vCPR,
+                   cprKeyVar=vCPRkey,
+                   miscodedMissingVar=vMiss)
+names(data)[names(data)=="joeVar"] <- "_joeVar"
+
+testData <- data
+
+save(testData, file="testData.Rda")

### 2016-10-02 latex/article.tex
Rebirth
Small changes

--- 2016-09-26 latex/article.Rnw
@@
 
-\section{Checking a dataset for errors}
-
 %% include your article here, just as usual
@@
 
-<<>>=
-library(cleanR)
-data(testData)
-testData
-@
+\section{Checking a dataset for errors}
+
+\begin{Schunk}
+\begin{Sinput}
+> library(cleanR)
+> data(testData)
+> testData
+\end{Sinput}
+\begin{Soutput}
+   charVar factorVar numVar intVar boolVar keyVar emptyVar _joeVar jack__var
+1        a         a      1      1    TRUE      1        1       1         1
+2        b         b      2      2   FALSE      2        1       2         2
+3        c         c      3      3    TRUE      3        1       3         3
+4        a         a      4      4    TRUE      4        1       4         4
+5        b         b      5      5    TRUE      5        1       5         5
+6        d         d      6      6   FALSE      6        1       6         6
+7        a         a      7      7    TRUE      7        1       7         7
+8        a         a      8      8   FALSE      8        1       8         8
+9        b         b      9      9    TRUE      9        1       9         9
+10       c         c     10     10    TRUE     10        1      10        10
+11       a         a      1      1    TRUE     11        1      11        11
+12       b         b      1      1   FALSE     12        1      12        12
+13       d         d      1      1      NA     13        1      13        13
+14       a         a      5      5      NA     14        1      14        14
+15    <NA>       999      5      5      NA     15        1      15        15
+   numOutlierVar smartNumVar      cprVar   cprKeyVar miscodedMissingVar
+1              1           0 010101-1111 010101-1111                  .
+2              2           0 020102-2929 020102-2929                   
+3              3           0 121201-1902 121201-1902                nan
+4              4           0 030729-2222 030729-2222                NaN
+5              5           0 080909-1212 080909-1212                NAN
+6              6           0 010101-1111 020202-0101                 na
+7              7           0 020102-2929 030303-0101                 NA
+8              8           1 121201-1902 040404-0101                 Na
+9              9           1 030729-2222 050505-0101                Inf
+10            10           1 080909-1212 060606-0101                inf
+11            11           1 010101-1111 070707-0101               -Inf
+12            12           1 020102-2929 080808-0101               -inf
+13            13           1 121201-1902 090909-0101                  -
+14            14           1 030729-2222 020202-0202                  9
+15           100           1 080909-1212 030303-0202                  9
+\end{Soutput}
+\end{Schunk}
+
+\begin{Schunk}
+\begin{Sinput}
+> clean(testData)
+\end{Sinput}
+\end{Schunk}
+
+Arguments
+
+\section{Extending the error checks}
 

### 2016-10-02 R/clean.R

@@ -3,6 +3,9 @@
 #' Runs a set of validation checks to check the variables in a data frame for potential errors.
 #' Performs checking steps according to user input and/or data type of the inputted variable.
 #'
+#' @param file XXX
+#' @param removeExisting XXX
+#' @param maxnum XXX
 #' @param standAlone If TRUE, the document begins with a markdown preamble such that it
 #' can be rendered as is.
 #' @param brag If TRUE, a note about cleanR is appended at the end of the output document.
@@ -30,12 +33,12 @@
 #' message is printed in the data summary as well.
 #' @param preChecks Variable checks that are performed before the summary/visualization/checking step. If
 #' any of these checks find problems, the variable will not be summarized nor visualized nor checked.
-#' @param replace If "never", an error is thrown if one of the files that we are about to write to 
-#' already exist. If "onlyCleanR", an error is thrown if one of the files that we are about to write to was 
+#' @param replace If "never", an error is thrown if one of the files that we are about to write to
+#' already exist. If "onlyCleanR", an error is thrown if one of the files that we are about to write to was
 #' not produced by cleanR (NOT DONE). If "always", no checks are performed.
-#' @param listChecks If TRUE, the document contains an overview of what checks were performed for 
+#' @param listChecks If TRUE, the document contains an overview of what checks were performed for
 #' each variable data type.
-#' @param checkDetails MAYBE ALSO IMPLEMENT THIS?: If TRUE, details about each check function are added 
+#' @param checkDetails MAYBE ALSO IMPLEMENT THIS?: If TRUE, details about each check function are added
 #' to the document (if available)
 #' @param vol Extra text string that is appended on the end of the output file name(s). For example, if the data
 #' set is called "myData", no file argument is supplied and vol="2", the output file will be called
@@ -59,7 +62,7 @@
 #' data(testData)
 #' clean(testData)
 #' }
-#' 
+#'
 #' \dontrun{
 #' characterFoo <- function(v) {
 #'  if (substr(substitute(a), 1, 1) == "_") {
@@ -82,7 +85,7 @@ clean <- function(o, file=NULL, removeExisting=TRUE, maxnum=NULL,
                   useVar="all", nagUser=TRUE,
                   smartNum=TRUE, preChecks=c("isSpecial", "isCPR"),
[…]

### 2016-10-04 R/clean.R
Tweaks to documentation

@@ -5,7 +5,6 @@
 #'
 #' @param file XXX
 #' @param removeExisting XXX
-#' @param maxnum XXX
 #' @param standAlone If TRUE, the document begins with a markdown preamble such that it
 #' can be rendered as is.
 #' @param brag If TRUE, a note about cleanR is appended at the end of the output document.
@@ -33,9 +32,8 @@
 #' message is printed in the data summary as well.
 #' @param preChecks Variable checks that are performed before the summary/visualization/checking step. If
 #' any of these checks find problems, the variable will not be summarized nor visualized nor checked.
-#' @param replace If "never", an error is thrown if one of the files that we are about to write to
-#' already exist. If "onlyCleanR", an error is thrown if one of the files that we are about to write to was
-#' not produced by cleanR (NOT DONE). If "always", no checks are performed.
+#' @param replace If FALSE (the default) an error is thrown if one of the files that we are about to write to
+#' already exist. If TRUE no checks are performed.
 #' @param listChecks If TRUE, the document contains an overview of what checks were performed for
 #' each variable data type.
 #' @param checkDetails MAYBE ALSO IMPLEMENT THIS?: If TRUE, details about each check function are added
@@ -45,12 +43,6 @@
 #' "cleanR_myData2.Rmd"
 #' @param \dots other arguments that are passed on the to checking, summary and visualization functions
 #' @return ???
-#' \itemize{
-#'   \item{"name"}{The name of the check}
-#'   \item{"description"}{Slightly more information about the check}
-#'   \item{"problem"}{An integer giving an error code. 0 means no potential errors were identified}
-#'   \item{"message"}{A string giving summary information in R markdown format about the results}
-#' }
 #' @author Anne H. Petersen \email{ahpe@@sund.ku.dk} and Claus Thorn Ekstrom \email{ekstrom@@sund.ku.dk}
 #' @seealso \code{\link{clean}}
 #' @keywords misc
@@ -76,7 +68,7 @@
 #' }
 #'
 #' @export
-clean <- function(o, file=NULL, removeExisting=TRUE, maxnum=NULL,
+clean <- function(o, file=NULL, removeExisting=TRUE,
                   standAlone=TRUE, brag=FALSE, ordering=c("asIs", "alpha"),
                   cleanUp="deletethisoption?",
                   quiet=TRUE, output="pdf", finish = "markdown",
[…]

### 2016-10-05 R/clean.R
tweaks

@@ -3,7 +3,8 @@
 #' Runs a set of validation checks to check the variables in a data frame for potential errors.
 #' Performs checking steps according to user input and/or data type of the inputted variable.
 #'
-#' @param file XXX
+#' @param o the data frame object to be checked
+#' @param file The filename of output file. If set to NULL (the default) then the filename will be the name of the data frame prefixed with "cleanR-".
 #' @param removeExisting XXX
 #' @param standAlone If TRUE, the document begins with a markdown preamble such that it
 #' can be rendered as is.
@@ -14,7 +15,7 @@
 #' @param output Output format, options are pdf and html
 #' @param finish "render" (makes pdf/html), "markdown" (makes markdown file), "print" (prints to screen).
 #' @param twoCol Should the results be presented in two columns (if finish is "render" or "markdown")?
-#' @param o the data frame or tibble(???) object to be checked
+
 #' @param characterChecks a list of error-checking functions to apply to character vectors
 #' @param integerChecks a list of error-checking functions to apply to integer vectors
 #' @param silent Should clean() run completely silently? Note that this option overrules the settings for

### 2016-10-09 latex/article.Rnw
Small changes
tweaks
Abstract + figure

@@
 \documentclass[article]{jss}
+\usepackage[utf8]{inputenc}
+\usepackage[nogin]{Sweave}
+\usepackage{pdfpages}
+\usepackage{tikz}
+\usetikzlibrary{shapes.geometric, arrows}
+
+\tikzstyle{io} = [trapezium, trapezium left angle=70, trapezium right angle=110, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=blue!30]
+\tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=orange!30]
+\tikzstyle{decision} = [diamond, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=green!30]
+\tikzstyle{arrow} = [thick,->,>=stealth]
 
@@
 %% almost as usual
-\author{Achim Zeileis\\Universit\"at Innsbruck \And
-        Second Author\\Plus Affiliation}
-\title{A Capitalized Title: Something about a Package \pkg{cleanR}}
+\author{Claus Thorn Ekstr\o m\\Biostatistics\\Department of Public
+  Health\\University of Copenhagen \And
+        Anne Helby Petersen\\Biostatistics\\Department of Public
+  Health\\University of Copenhagen}
+\title{\pkg{cleanR}: Your Personal Maid for Cleaning Datasets in \proglang{R}}
 
 %% for pretty printing and a nice hypersummary also set:
-\Plainauthor{Achim Zeileis, Second Author} %% comma-separated
-\Plaintitle{A Capitalized Title: Something about a Package cleanR} %% without formatting
-\Shorttitle{\pkg{cleanR}: A Capitalized Title} %% a short title (if necessary)
+\Plainauthor{Claus Thorn Ekstr\o m, Anne Helby Petersen} %% comma-separated
+\Plaintitle{{cleanR}: Your maid for cleaning datasets in R} %% without formatting
+\Shorttitle{\pkg{cleanR}: maid for cleaning datasets in R} %% a short title (if necessary)
 
 %% an abstract and keywords
-\Abstract{
-  The abstract of the article.
+\Abstract{Data cleaning and validation is the first step in any data
+  analysis since the validity of the conclusions from the analysis
+  hinged on the quality of the input data. Ideally, a human investigator
+  should go through each variable in the dataset and look for
+  potential errors --- both in input values and coding.
+
+  We describe an \proglang{R} package which implements an extensive
+  and customizeable suite of checks to be applied to the variables in
+  a dataset in order to identify potential problems in the
+  corresponding variables. The typical output is a stand-alone
+  document that summarizes the variables and lists potential
+  errors. The results are typically presented in a stand-alone
+  document that could be perused by an investigator with an
+  understanding of the variables in the data and the experimental design.
+
+  The \pkg{cleanR} package is designed to be easily extended with custom
+  user-created checks that are relevant in particular situations.
 }
-\Keywords{keywords, comma-separated, not capitalized, \proglang{Java}}
-\Plainkeywords{keywords, comma-separated, not capitalized, Java} %% without formatting
+\Keywords{data cleaning, quality control, \proglang{R}}
+\Plainkeywords{data cleaning, quality control, R} %% without formatting
 %% at least one keyword must be supplied
@@
 \Address{
-  Achim Zeileis\\
-  Department of Statistics and Mathematics\\
-  Faculty of Economics and Statistics\\
-  Universit\"at Innsbruck\\
-  6020 Innsbruck, Austria\\
-  E-mail: \email{Achim.Zeileis@uibk.ac.at}\\
+  Claus Thorn Ekstr\o m\\
+  Biostatistics, Department of Public Health\\
+  University of Copenhagen\\
+  Denmark\\
+  E-mail: \email{ekstrom@sund.ku.dk}\\
   URL: \url{http://eeecon.uibk.ac.at/~zeileis/}
@@
 Statisticians and data analysts spend a large portion of their time on
-data cleaning and on data wrangling. Packages such as \pkg{}, \pkg{},
+data cleaning and on data wrangling. Packages such as \pkg{data.table}, \pkg{plyr},
 and \pkg{} make data wrangling a lot easier in R, but there are only a
-few options available for data cleaning.
+few options available for automated data cleaning.
 
@@
 
-\section{Checking a dataset for errors}
+In this paper, we will briefly present some properties of the simplex
+distribution, which provides a foundation for the inference and
+computation in the package simplexreg (Zhang, Qiu, and Shi 2016). The
+package is implemented in the R system and available from the
+Comprehensive R Archive Network (CRAN) at https:
+//CRAN.R-project.org/package=simplexreg. There are three major lines
+of functions:
+1. Calculation of the density, cumulative distribution, quantile and
+random number gen- erating functions for the simplex distribution.
+2. Statistical inference in the simplex generalized linear model
+(SGLM, Zhang and Qiu 2014) via maximum likelihood (ML) for
+cross-sectional proportional data set.
+3. Analysis of longitudinal proportional data set via an extended
+version of generalized estimating equations (GEE) using the simplex
+distribution.
+
+The manusccript is organized as follows. Section \ref{sec:example1}
+presents a worked example and tutorial on how to use the \pkg{cleanR}
+package to create a summary of potential errors. Section
+\ref{sec:internals} illustrates the methodology of modeling
+proportional data using simplex generalized regression. The
+generalized estimating equations for longitudinal proportional
+outcomes are given in Section 4. Then we address model diagnostics in
+Section 5. Section 6 presents the details of the simplexreg
+package. Section 7 further conducts analyses based on the simplex
+distribution in R with real data sets. Finally, plans for extending
+the package are described in Section 8.
 
@@
 
+\begin{figure}
+\begin{center}
+\begin{tikzpicture}[node distance=2cm]
+\node (start) [io] {data.frame};
+\node (pro1) [process, below of=start] {clean};
+\node (screen) [io, right of=pro1, xshift=2cm] {Screen};
+\node (pro2) [process, below of=pro1] {Rmd};
+\node (html) [io, right of=pro2, xshift=2cm] {html};
+\node (pdf) [io, below of=html] {pdf};
+\draw [arrow] (start) -- (pro1);
+\draw [arrow] (pro1) -- (screen);
+\draw [arrow] (pro1) -- (pro2);
+\draw [arrow] (pro2) -- (html);
+\draw [arrow] (pro2) -- (pdf);
+\end{tikzpicture}
+\end{center}
+\label{fig:flowchart1}
+\caption{The process of cleaning data frames using the \pkg{cleanR}
+  package. A series of check is applied to each variable in a
+  \texttt{data.frame} and a summary of the result is either printed to
+  the screen, or an \proglang{R} markdown file is produced which is
+  subsequently rendered.}
+\end{figure}
+
+\section{Checking a dataset for errors} \label{sec:example1}
+
+In \pkg{cleanR} the clean function
+
 <<>>=
@@
 data(testData)
-testData
+head(testData)
 @
 
-\section[About Java]{About \proglang{Java}}
-%% Note: If there is markup in \(sub)section, then it has to be escape as above.
+<<>>=
+                                        # clean(testData)
+2+3
+@
+
+% \includepdf[fitpaper=true, pages=-]{test.pdf}
+
+Arguments
+
+\section{The structure of \pkg{cleanR}} \label{sec:internals}
+
+asd
+asd
+
+\section{Extending \pkg{cleanR} by adding custom error checks} \label{sec:extending}
+
+Lav en situation svarende til eksemplet
+
+<<>>=
+characterFoo <- function(v) {
+    if (substr(substitute(v), 1, 1) == "_") {
+        out <- list(problem=TRUE, message="Note that the variable name begins with \\_")
+    } else out <- list(problem=FALSE, message="")
+    out
+}
+class(characterFoo) <- "checkFunction"
+attr(characterFoo, "description") <- "I really hate underscores"
+#clean(testData, characterChecks=c(defaultCharacterChecks(), "characterFoo"))
+
+@
+
+Lav også et eksempel med rangecheck.
+
+\section{Using the online web-app} \label{sec:web-app}
+
+asd
+asd
 

### 2016-10-09 latex/article.tex
Tweaks to documentation
Abstract + figure

--- 2016-09-26 latex/article.Rnw
@@
 
-<<>>=
-library(cleanR)
-data(testData)
-testData
-@
+\begin{figure}
+\begin{center}
+\begin{tikzpicture}[node distance=2cm]
+\node (start) [io] {data.frame};
+\node (pro1) [process, below of=start] {clean};
+\node (screen) [io, right of=pro1, xshift=2cm] {Screen};
+\node (pro2) [process, below of=pro1] {Rmd};
+\node (html) [io, right of=pro2, xshift=2cm] {html};
+\node (pdf) [io, below of=html] {pdf};
+\draw [arrow] (start) -- (pro1);
+\draw [arrow] (pro1) -- (screen);
+\draw [arrow] (pro1) -- (pro2);
+\draw [arrow] (pro2) -- (html);
+\draw [arrow] (pro2) -- (pdf);
+\end{tikzpicture}
+\end{center}
+\label{fig:flowchart1}
+\caption{The process of cleaning data frames using the \pkg{cleanR}
+  package. A series of check is applied to each variable in a
+  \texttt{data.frame} and a summary of the result is either printed to
+  the screen, or an \proglang{R} markdown file is produced which is
+  subsequently rendered.}
+\end{figure}
 
-\section[About Java]{About \proglang{Java}}
-%% Note: If there is markup in \(sub)section, then it has to be escape as above.
+\section{Checking a dataset for errors} \label{sec:example1}
+
+In \pkg{cleanR} the clean function
+
+\begin{Schunk}
+\begin{Sinput}
+> library(cleanR)
+> data(testData)
+> head(testData)
+\end{Sinput}
+\begin{Soutput}
+  charVar factorVar numVar intVar boolVar keyVar emptyVar _joeVar jack__var
+1       a         a      1      1    TRUE      1        1       1         1
+2       b         b      2      2   FALSE      2        1       2         2
+3       c         c      3      3    TRUE      3        1       3         3
+4       a         a      4      4    TRUE      4        1       4         4
+5       b         b      5      5    TRUE      5        1       5         5
+6       d         d      6      6   FALSE      6        1       6         6
+  numOutlierVar smartNumVar      cprVar   cprKeyVar miscodedMissingVar
+1             1           0 010101-1111 010101-1111                  .
+2             2           0 020102-2929 020102-2929                   
+3             3           0 121201-1902 121201-1902                nan
+4             4           0 030729-2222 030729-2222                NaN
+5             5           0 080909-1212 080909-1212                NAN
+6             6           0 010101-1111 020202-0101                 na
+\end{Soutput}
+\end{Schunk}
+
+\begin{Schunk}
+\begin{Sinput}
+>                                         # clean(testData)
+> 2+3
+\end{Sinput}
+\begin{Soutput}
+[1] 5
+\end{Soutput}
+\end{Schunk}
+
+% \includepdf[fitpaper=true, pages=-]{test.pdf}
+
+Arguments
+
+\section{The structure of \pkg{cleanR}} \label{sec:internals}
+
+asd
+asd
+
+\section{Extending \pkg{cleanR} by adding custom error checks} \label{sec:extending}
+
+Lav en situation svarende til eksemplet
+
+\begin{Schunk}
+\begin{Sinput}
+> characterFoo <- function(v) {
++     if (substr(substitute(v), 1, 1) == "_") {
++         out <- list(problem=TRUE, message="Note that the variable name begins with \\_")
++     } else out <- list(problem=FALSE, message="")
++     out
++ }
+> class(characterFoo) <- "checkFunction"
+> attr(characterFoo, "description") <- "I really hate underscores"
+> #clean(testData, characterChecks=c(defaultCharacterChecks(), "characterFoo"))
+> 
+\end{Sinput}
+\end{Schunk}
+
+Lav også et eksempel med rangecheck.
+
+\section{Using the online web-app} \label{sec:web-app}
+
+asd
+asd
 

### 2016-10-09 R/clean.R

@@ -57,7 +57,7 @@
 #'#Add user defined check-function to the checks performed on character variables:
 #' \dontrun{
 #' characterFoo <- function(v) {
-#'  if (substr(substitute(a), 1, 1) == "_") {
+#'  if (substr(substitute(v), 1, 1) == "_") {
 #'    out <- list(problem=TRUE, message="Note that the variable name begins with \\_")
 #'  } else out <- list(problem=FALSE, message="")
 #'  out

### 2016-10-10 latex/article.Rnw
Rnw-fix

@@
 Statisticians and data analysts spend a large portion of their time on
-data cleaning and on data wrangling. Packages such as \pkg{data.table}, \pkg{plyr},
-and \pkg{} make data wrangling a lot easier in R, but there are only a
-few options available for automated data cleaning.
+data cleaning and on data wrangling. Packages such as
+\pkg{data.table}, and \pkg{plyr} have made data wrangling a lot easier in
+\proglang{R}, but there are only a few options available for automated data
+cleaning.
 

### 2016-10-11 latex/article.Rnw
Fixed paper

@@
 in the dataset can only be understood in the proper context of the
-experiment. While each dataset is different and require
-... (``simple questions''
+experiment. While each dataset is different and requires unique
+attention there are often a number of similar tasks that are
+undertaken as part of the data cleaning and quality control
+process. This is especially true when data are received repeatedly
+from the same source as data providers tend to use a stable setup (and
+hence introduce the same set of potential mistakes that need to be
+identified and corrected).
 
-... there are a number of common occurrences that .... This is particulary
-Automated
+In many situations these errors are discovered in the process of the
+data analysis (e.g., a categorical variable with numeric labels is
+wrongly classified as a numeric variable), but in other cases a human
+with knowledge about the data context area is needed to identify
+possible mistakes in the data (e.g., if there are 4 categories for a
+variable that should only have 3).  It is necessary to summarize
+information --- both numerically and graphically --- about each
+variable in order for an investigator to detect possible errors, and
+help raise flags of warning to draw the attention of investigator for
+the situations where there \emph{might} be a error.
 
-... partly be done by ... However ... enough ... similar tasks that
-pop
-
-However it is not uncommon for ... to run a set of ....
-
-In this paper, we will briefly present some properties of the simplex
-distribution, which provides a foundation for the inference and
-computation in the package simplexreg (Zhang, Qiu, and Shi 2016). The
-package is implemented in the R system and available from the
-Comprehensive R Archive Network (CRAN) at https:
-//CRAN.R-project.org/package=simplexreg. There are three major lines
-of functions:
-1. Calculation of the density, cumulative distribution, quantile and
-random number gen- erating functions for the simplex distribution.
-2. Statistical inference in the simplex generalized linear model
-(SGLM, Zhang and Qiu 2014) via maximum likelihood (ML) for
-cross-sectional proportional data set.
-3. Analysis of longitudinal proportional data set via an extended
-version of generalized estimating equations (GEE) using the simplex
-distribution.
-
-The manusccript is organized as follows. Section \ref{sec:example1}
-presents a worked example and tutorial on how to use the \pkg{cleanR}
-package to create a summary of potential errors. Section
-\ref{sec:internals} illustrates the methodology of modeling
-proportional data using simplex generalized regression. The
-generalized estimating equations for longitudinal proportional
-outcomes are given in Section 4. Then we address model diagnostics in
-Section 5. Section 6 presents the details of the simplexreg
-package. Section 7 further conducts analyses based on the simplex
-distribution in R with real data sets. Finally, plans for extending
-the package are described in Section 8.
-
-%% include your article here, just as usual
-%% Note that you should use the \pkg{}, \proglang{} and \code{}
-%% commands.
+The \pkg{cleanR} package supports the automated checking of errors in
+a dataset and produces an output document with detailed information
+about each variable that can be
+The process is illustrated in Figure~\ref{fig:flowchart1}. A dataset
+is cleaned and an \proglang{R} markdown file is produced and possibly
+rendered into html or pdf.
 
@@
 
+The manusccript is organized as follows. Section \ref{sec:example1}
+presents a worked example and tutorial on how to use the \pkg{cleanR}
+package to create a summary of potential errors. Section
+\ref{sec:internals} illustrates the methodology of modeling
+proportional data using simplex generalized regression. The
+generalized estimating equations for longitudinal proportional
+outcomes are given in Section 4. Then we address model diagnostics in
+Section 5. Section 6 presents the details of the simplexreg
+package. Section 7 further conducts analyses based on the simplex
+distribution in R with real data sets. Finally, plans for extending
+the package are described in Section 8.
+
+%% include your article here, just as usual
+%% Note that you should use the \pkg{}, \proglang{} and \code{}
+%% commands.
+
 \section{Checking a dataset for errors} \label{sec:example1}
@@
 
-\section{The structure of \pkg{cleanR}} \label{sec:internals}
+\section{The structure of cleanR} \label{sec:internals}
 
-asd
-asd
+\begin{figure}
+\begin{center}
+\begin{tikzpicture}[node distance=2cm]
+\node (start) [io] {data.frame};
+\node (pro1) [process, below of=start] {prechecks};
+\node (screen) [io, right of=pro1, xshift=2cm] {Screen};
+\node (pro2) [process, below of=pro1] {Rmd};
+\node (html) [io, right of=pro2, xshift=2cm] {html};
+\node (pdf) [io, below of=html] {pdf};
+\draw [arrow] (start) -- (pro1);
+\draw [arrow] (pro1) -- (screen);
+\draw [arrow] (pro1) -- (pro2);
+\draw [arrow] (pro2) -- (html);
+\draw [arrow] (pro2) -- (pdf);
+\end{tikzpicture}
+\end{center}
+\label{fig:flowchart1}
+\caption{The process of cleaning data frames using the \pkg{cleanR}
+  package. A series of check is applied to each variable in a
+  \texttt{data.frame} and a summary of the result is either printed to
+  the screen, or an \proglang{R} markdown file is produced which is
+  subsequently rendered.}
+\end{figure}
 
-\section{Extending \pkg{cleanR} by adding custom error checks} \label{sec:extending}
+Bør vise DF -> prechecks -> for each variable do this ...
+
+\section{Extending cleanR by adding custom error checks} \label{sec:extending}
 

### 2016-10-11 latex/article.tex
Fixed paper

@@
 \end{figure}
+
+The manusccript is organized as follows. Section \ref{sec:example1}
+presents a worked example and tutorial on how to use the \pkg{cleanR}
+package to create a summary of potential errors. Section
+\ref{sec:internals} illustrates the methodology of modeling
+proportional data using simplex generalized regression. The
+generalized estimating equations for longitudinal proportional
+outcomes are given in Section 4. Then we address model diagnostics in
+Section 5. Section 6 presents the details of the simplexreg
+package. Section 7 further conducts analyses based on the simplex
+distribution in R with real data sets. Finally, plans for extending
+the package are described in Section 8.
+
+%% include your article here, just as usual
+%% Note that you should use the \pkg{}, \proglang{} and \code{}
+%% commands.
 

### 2016-10-14 latex/draft_cleanR_advancedUsers.tex

\documentclass[article]{jss}
\usepackage[utf8]{inputenc}
%\usepackage[nogin]{Sweave}
\usepackage{pdfpages}
%\usepackage{tikz}
%\usetikzlibrary{shapes.geometric, arrows}

%\tikzstyle{io} = [trapezium, trapezium left angle=70, trapezium right angle=110, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=blue!30]
%\tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=orange!30]
%\tikzstyle{decision} = [diamond, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=green!30]
%\tikzstyle{arrow} = [thick,->,>=stealth]

\newcommand{\bofr}{\mathcal{B}(\mathbb{R})}
\newcommand{\RR}{\mathbb{R}}
\newcommand{\bs}{\backslash}
\newcommand{\NN}{\mathbb{N}}
\newcommand{\ZZ}{\mathbb{Z}}
\newcommand{\QQ}{\mathbb{Q}}
\newcommand{\BB}{\mathbb{B}}
\newcommand{\CC}{\mathbb{C}}
\newcommand{\FF}{\mathbb{F}}
\newcommand{\la}{\lambda}
\newcommand{\coola}{\mathcal{A}}
\newcommand{\coolm}{\mathcal{M}}
\newcommand{\cooll}{\mathcal{L}}
\newcommand{\ipl}{\langle}
\newcommand{\ipr}{\rangle}
\newcommand{\limn}{\lim_{n \to \infty}}
\newcommand{\lr}{\Leftrightarrow}
\newcommand{\set}[1]{\left\lbrace #1 \right\rbrace}
\renewcommand{\d}{\,\text{d}}
\newcommand{\pt}[1]{\left(#1\right)}
\newcommand{\fork}[1]{\left\{\begin{array}{lr}#1\end{array}\right.} % gaffel-funktion
\newcommand{\tuad}[1]{\qquad\ \text{#1}\ }
\newcommand{\ND}{\mathcal{N}}
\newcommand{\Perp}{\perp \! \! \! \perp}
\newcommand{\ddu}{\frac{\partial}{\partial u}}
\newcommand{\ddv}{\frac{\partial}{\partial v}}
\newcommand{\ddx}{\frac{\partial}{\partial x}}
\newcommand{\dds}{\frac{\partial}{\partial s}}
\newcommand{\ddt}{\frac{\partial}{\partial t}}
\newcommand{\dd}{\partial}
\newcommand{\g}{\cellcolor[gray]{0.5}}
\newcommand{\var}[1]{\textsc{\lowercase{#1}}}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\newcommand{\blue}[1]{\textcolor{blue}{#1}}
\newcommand{\rhomax}{\rho_{\text{max}}}
\newcommand{\D}{\text{D}}
\newcommand{\nperp}{\not\!\perp\!\!\!\perp}
\newcommand*\rot{\rotatebox{90}}
\newcommand{\green}[1]{\textcolor{OliveGreen}{#1}}
\newcommand{\s}[1]{\bar{#1}}
\newcommand{\us}[1]{\underbar{#1}}
\newcommand{\argmax}{\text{argmax}}

\newcommand{\plim}{\plimo_{n \to \infty}}
\newcommand{\asim}{\stackrel{as.}{\sim}}

\newcommand{\hl}[1]{\textcolor{magenta}{#1}}
\newcommand{\R}[1]{\texttt{#1}}

%\begin{center}
%\includegraphics[scale=0.3]{residualplot_logengel.jpeg} 
%\end{center}

\begin{document}

\section{Introduction}

\section{Checking a dataset for errors}

\section{Customizing the cleanR output}
Though the discussion in the above paints a picture of cleanR as a user-friendly package which requires practically no knowledge of R, one should not be mistaken to think that it is not customizable. In fact, the main function of cleanR, \R{clean}, is mainly a tool for formatting the results from various checking-, summary- and visualizaion functions \hl{as described... blabla}. Thus, the actual work underlying a cleanR output file can be anything or nothing - depending on the arguments given to \R{clean}. This section consists of three parts. We commence this section with an overview of how contents are controlled in \R{cleanR} in terms of what cleaning steps are performed. Secondly, we turn to the possibilities of customizing what the final cleanR file looks like. Lastly, we go through an example of how to put all these customization options together in practice.

\subsection{Controlling contents}
By the \textit{contents} of the cleanR output, we refer to every step that actually involves a function being called on variables from the dataset. There are three stages in which this occurs \hl{(with ref. to figure/flowchart?)}:
\begin{enumerate}
\item In the precheck functions
\item In the summarize/visualize/check (SVC) step
\item In the multivariate visuzalizations \hl{Or whatever, in the stuff that we have not yet implemented. I'm pretending like this stage doesn't exist below.}
\end{enumerate}
Each of these stages are controllable using appropriate function parameters in \R{clean}. In the above, we presented the default \R{cleanR} settings and how to tweak them into providing a slightly different data cleaning outputs. However, if for instance the dataset at hand requires completely different visualizations, more control is needed. \R{cleanR} uses three different types of functions for performing all stages in the above, namely \R{summaryFunction}s, \R{visualFunction}s and \R{checkFunction}s. They each have distinct input-output structures and most instances of the functions are build as \R{S3} generics with methods for different data classes. By understanding how to construct proper \R{summaryFunction}s, \R{visualFunction}s and \R{checkFunction}s, the entire contents of the \R{cleanR} output is at your hands. Therefore, this section is dedicated to introducing each in turn.\\

\hl{Something like a figure that gives an overview of all summaryFunctions, ... available in standard cleanR, including default settings and some kind of description of where they are called (in precheck or SVC step). Also something about what datatypes they can each be called on. Also, more text here about the relationship between the two (three) stages and the three function types.}

\subsubsection{Writing a summaryFunction}
Though \R{cleanR} provides a special class for \R{summaryFunction}s, there really is nothing special about these functions: They are nothing but regular functions with a certain input/output-structure. Specifically, they all follow the template below:
\begin{Verbatim}
mySummaryFunction <- function(v) {
	res <- [result of whatever summary we are doing]
	list(feature = "Feature name", result = res)
}
\end{Verbatim}
and we recommend furthermore adding them to the overview of all summary functions by converting them to proper \R{summaryFunction} objects:
\begin{Verbatim}
mySummaryFunction <- summaryFunction(mySummaryFunction,
	description = "Some text describing what your summaryFunction does")
\end{Verbatim}
which adds the new function to the output of \R{allSummaryFunctions()}. Note that \R{v} is a vector and that \R{res} should be either a character string or something that will be printed as one. In other words, e.g. integers are allowed, but matrices are not. Though a lot of different things can go into the \R{summaryFunction} template, we recommend only using them for summarizing the features of a variable, and leaving tests and checks for the \R{checkFunction}s (presented below).

\subsubsection{Writing a visualFunction}
\R{visualFunction}s are the functions that produce the figures of a \R{cleanR} output document. Writing a visual function is slightly more complicated than writing a summary function. This follows from the fact that visualFunctions need to be able to output standalone code for plots in order for \R{clean} to build standalone rmarkdown files. We recommend using the following structure:
\begin{Verbatim}
myVisualFunction <- function(v, vnam, doEval) {
	thisCall <- call("[the name of the function used to produce the plot]", 
		v, [additional arguments to the plotting function])
	if (doEval) {
		return(eval(thisCall))
	} else return(deparse(thisCall)
}

myVisualFunction <- visualFunction(myVisualFunction, 
	description = "Some text describing your visualFunction")
\end{Verbatim}
In this function, \R{v} is the variable to be visualized, \R{vnam} is its name (which should generally be passed to \R{title} or \R{main} arguments in plotting functions) and \R{doEval} controls whether the output is a plot (if \R{TRUE}) or a character string of standalone code for producing a plot (if \R{FALSE}). The latter \R{doEval} setting is not strictly necessary for its use in \R{clean}, but it makes it easier to assess what visualization options are available. In either case, it should be noted that all the parameters listed above, \R{v}, \R{vnam} and \R{doEval}, are mandatory, so they should be left as is (\hl{as are?}), even if you do not want to use them. As with \R{summaryFunction}s, an overview of all available \R{visualFunction}s in the environment can be obtained by calling
\begin{Verbatim}
allVisualFunctions()
\end{Verbatim}
and by calling 
\begin{Verbatim}
allVisual(v, vnam, output = "html")
\end{Verbatim}
an overview of all plotting options applied on \R{v} is produced and opened as a html document for easy comparison. \hl{Should we mention the side effect of producing .rmd and .html files on disc?}

\subsubsection{Writing a checkFunction}
The last, but also most important, \R{cleanR} function type is the \R{checkFunction}. These are the functions that flag issues in the data and control the flow of the overall data cleaning process in the precheck stage. A \R{checkFunction} can be written using the following template:
\begin{Verbatim}
myCheckFunction <- function(v) {
	[do your check]
	problem <- [is there a problem? TRUE/FALSE]
	problemValues <- [vector of values in v that are problematic]
	problemStatus <- list(problem = problem, problemValues = problemValues)

	problemMessage <- "[The message that should be printed prior to listing 
			problem values in the cleanR output]"

	outMessage <- messageGenerator(list(problem = problem, 
		problemValues = problemValues, message = problemMessage))

	list(problem = problem, message = message) #problem is TRUE/FALSE,
		# message is a text string 
}
myCheckFunction <- checkFunction(myCheckFunction, 
	description = "[A description of your checkFunction]")
\end{Verbatim}
Only the input parameter (\R{v}) and the output format strictly has to follow this structure. However, we recommend using \R{messageGenerator} for consistent styling of all \R{checkFunction} messages. This function simply pastes together the \R{problemMessage} and the \R{problemValues}, with the latter being quoted and sorted alphabetically. Note that printing quotes in rmarkdown requires an extensive amount of character escaping, so opting for \R{messageGenerator} really is the easiest solution.

While the descriptions of \R{summaryFunction}s and \R{visualFunction}s are only for internal use in the \R{allSummaryFunctions()} and \R{allVisualFunctions()} outputs, respectively, \R{checkFunction} descriptions are actually visible in the \R{cleanR} output document. These are the brief descriptions presented in Part 1 \hl{eh? Is this clear? Whatever we call this section in the above} in the output document. If a \R{checkFunction} does not have a description (for instance, if it is just a regular \R{function} using the \R{checkFunction} input/output-structure), the function name will be printed instead of the description.

\subsection{Controlling formatting}
\hl{Something about how to control what is printed where. Should be much briefer than the stuff in the above. This section could maybe include the following points:
\begin{enumerate}
\item Controlling "part 1", i.e. the stuff that is printed before the loop is started. 
\item The effect of prechecks.
\item A table containing other parameters that control formatting/stuff like this, e.g. twoCol, checkDetails (if we ever implement it), listChecks... 
\end{enumerate}
}

\subsection{Controlling flow?}
\hl{I feel like we maybe need yet another section about customization. We have not yet described the following (rather important(?)) features:
\begin{itemize}
\item The mode argument - controlling which SVC steps are performed
\item smartNum - chooosing whether or not numeric/integer variables with only a few levels are treated as factors
\item standAlone - Describing how it is possible to produce .Rmd files that can be included in other .Rmd files using the child option and standAlone = F (removing the YAML preamble (and maybe also the cleanR commerical? Are we that nice?)). 
\end{itemize}
}

\subsection{A worked example}
\hl{Maybe use the examples from the documentation to construct new summaryFunction etc. and add them to the default options. }

\section{To-dos in the code}
\hl{ ...  introduced by discrepancy between what I say we have done and what we have actually done:
\begin{itemize}
\item summaryFunction stuff:
	\begin{itemize}
		\item Make constructor function
		\item Make allSummaryFunctions()
		\item Change class of all summary (and description) functions into summaryFunction (possibly letting description functions be a subclass?)
		\item Maybe: Make clean check if summary functions are really summaryFunctions?
	\end{itemize}
\item visualFunction stuff:
	\begin{itemize}
		\item Make constructor function
		\item Make allVisualFunctions()
		\item Change class of all visual functions into visualFunction
		\item Maybe: Make clean check if visual functions are really visual functions?
	\end{itemize}
\item checkFunction stuff:
	\begin{itemize}
		\item Make allCheckFunctions()
		\item Work on messageGenerator() to make it better suited for being an exported function + export it
		\item Make allVisual()
	\end{itemize}
\end{itemize}
}

\end{document}

### 2016-10-16 latex/article.Rnw
Minor tweaks. TeX merge

@@
 
+\section{Customizing the cleanR output}
+Though the discussion in the above paints a picture of cleanR as a user-friendly package which requires practically no knowledge of R, one should not be mistaken to think that it is not customizable.
+In fact, the main function of cleanR, \R{clean}, is mainly a tool for formatting the results from various checking-, summary- and visualizaion functions \hl{as described...
+blabla}.
+Thus, the actual work underlying a cleanR output file can be anything or nothing - depending on the arguments given to \R{clean}.
+This section consists of three parts.
+We commence this section with an overview of how contents are controlled in \R{cleanR} in terms of what cleaning steps are performed.
+Secondly, we turn to the possibilities of customizing what the final cleanR file looks like.
+Lastly, we go through an example of how to put all these customization options together in practice.
+
+\subsection{Controlling contents}
+By the \textit{contents} of the cleanR output, we refer to every step that actually involves a function being called on variables from the dataset.
+There are three stages in which this occurs \hl{(with ref.
+to figure/flowchart?)}:
+\begin{enumerate}
+\item In the precheck functions
+\item In the summarize/visualize/check (SVC) step
+\item In the multivariate visuzalizations \hl{Or whatever, in the stuff that we have not yet implemented. I'm pretending like this stage doesn't exist below.}
+\end{enumerate}
+Each of these stages are controllable using appropriate function parameters in \R{clean}.
+In the above, we presented the default \R{cleanR} settings and how to tweak them into providing a slightly different data cleaning outputs.
+However, if for instance the dataset at hand requires completely different visualizations, more control is needed.
+\R{cleanR} uses three different types of functions for performing all stages in the above, namely \R{summaryFunction}s, \R{visualFunction}s and \R{checkFunction}s.
+They each have distinct input-output structures and most instances of the functions are build as \R{S3} generics with methods for different data classes.
+By understanding how to construct proper \R{summaryFunction}s, \R{visualFunction}s and \R{checkFunction}s, the entire contents of the \R{cleanR} output is at your hands.
+Therefore, this section is dedicated to introducing each in turn.\\
+
+\hl{Something like a figure that gives an overview of all summaryFunctions, ...
+available in standard cleanR, including default settings and some kind of description of where they are called (in precheck or SVC step).
+Also something about what datatypes they can each be called on.
+Also, more text here about the relationship between the two (three) stages and the three function types.}
+
+\subsubsection{Writing a summaryFunction}
+Though \R{cleanR} provides a special class for \R{summaryFunction}s, there really is nothing special about these functions: They are nothing but regular functions with a certain input/output-structure.
+Specifically, they all follow the template below:
+\begin{Verbatim}
+mySummaryFunction <- function(v) {
+	res <- [result of whatever summary we are doing]
+	list(feature = "Feature name", result = res)
+}
+\end{Verbatim}
+and we recommend furthermore adding them to the overview of all summary functions by converting them to proper \R{summaryFunction} objects:
+\begin{Verbatim}
+mySummaryFunction <- summaryFunction(mySummaryFunction,
+	description = "Some text describing what your summaryFunction does")
+\end{Verbatim}
+which adds the new function to the output of \R{allSummaryFunctions()}.
+Note that \R{v} is a vector and that \R{res} should be either a character string or something that will be printed as one.
+In other words, e.g.
+integers are allowed, but matrices are not.
+Though a lot of different things can go into the \R{summaryFunction} template, we recommend only using them for summarizing the features of a variable, and leaving tests and checks for the \R{checkFunction}s (presented below).
+
+\subsubsection{Writing a visualFunction}
+\R{visualFunction}s are the functions that produce the figures of a \R{cleanR} output document.
+Writing a visual function is slightly more complicated than writing a summary function.
+This follows from the fact that visualFunctions need to be able to output standalone code for plots in order for \R{clean} to build standalone rmarkdown files.
+We recommend using the following structure:
+\begin{Verbatim}
+myVisualFunction <- function(v, vnam, doEval) {
+	thisCall <- call("[the name of the function used to produce the plot]",
+		v, [additional arguments to the plotting function])
+	if (doEval) {
+		return(eval(thisCall))
+	} else return(deparse(thisCall)
+}
+
+myVisualFunction <- visualFunction(myVisualFunction,
+	description = "Some text describing your visualFunction")
+\end{Verbatim}
+In this function, \R{v} is the variable to be visualized, \R{vnam} is its name (which should generally be passed to \R{title} or \R{main} arguments in plotting functions) and \R{doEval} controls whether the output is a plot (if \R{TRUE}) or a character string of standalone code for producing a plot (if \R{FALSE}).
+The latter \R{doEval} setting is not strictly necessary for its use in \R{clean}, but it makes it easier to assess what visualization options are available.
+In either case, it should be noted that all the parameters listed above, \R{v}, \R{vnam} and \R{doEval}, are mandatory, so they should be left as is (\hl{as are?}), even if you do not want to use them.
+As with \R{summaryFunction}s, an overview of all available \R{visualFunction}s in the environment can be obtained by calling
+\begin{Verbatim}
+allVisualFunctions()
+\end{Verbatim}
+and by calling
+\begin{Verbatim}
+allVisual(v, vnam, output = "html")
+\end{Verbatim}
+an overview of all plotting options applied on \R{v} is produced and opened as a html document for easy comparison. \hl{Should we mention the side effect of producing .rmd and .html files on disc?}
+
+\subsubsection{Writing a checkFunction}
+The last, but also most important, \R{cleanR} function type is the \R{checkFunction}.
+These are the functions that flag issues in the data and control the flow of the overall data cleaning process in the precheck stage.
+A \R{checkFunction} can be written using the following template:
+\begin{Verbatim}
+myCheckFunction <- function(v) {
+	[do your check]
+	problem <- [is there a problem? TRUE/FALSE]
+	problemValues <- [vector of values in v that are problematic]
+	problemStatus <- list(problem = problem, problemValues = problemValues)
+
+	problemMessage <- "[The message that should be printed prior to listing
+			problem values in the cleanR output]"
+
+	outMessage <- messageGenerator(list(problem = problem,
+		problemValues = problemValues, message = problemMessage))
+
+	list(problem = problem, message = message) #problem is TRUE/FALSE,
+		# message is a text string
+}
+myCheckFunction <- checkFunction(myCheckFunction,
+	description = "[A description of your checkFunction]")
+\end{Verbatim}
+Only the input parameter (\R{v}) and the output format strictly has to follow this structure.
+However, we recommend using \R{messageGenerator} for consistent styling of all \R{checkFunction} messages.
+This function simply pastes together the \R{problemMessage} and the \R{problemValues}, with the latter being quoted and sorted alphabetically.
+Note that printing quotes in rmarkdown requires an extensive amount of character escaping, so opting for \R{messageGenerator} really is the easiest solution.
+
+While the descriptions of \R{summaryFunction}s and \R{visualFunction}s are only for internal use in the \R{allSummaryFunctions()} and \R{allVisualFunctions()} outputs, respectively, \R{checkFunction} descriptions are actually visible in the \R{cleanR} output document.
+These are the brief descriptions presented in Part 1 \hl{eh?
+Is this clear?
+Whatever we call this section in the above} in the output document.
+If a \R{checkFunction} does not have a description (for instance, if it is just a regular \R{function} using the \R{checkFunction} input/output-structure), the function name will be printed instead of the description.
+
+\subsection{Controlling formatting}
+\hl{Something about how to control what is printed where. Should be much briefer than the stuff in the above. This section could maybe include the following points:
+\begin{enumerate}
+\item Controlling "part 1", i.e. the stuff that is printed before the loop is started.
+\item The effect of prechecks.
+\item A table containing other parameters that control formatting/stuff like this, e.g. twoCol, checkDetails (if we ever implement it), listChecks...
+\end{enumerate}
+}
+
+\subsection{Controlling flow?}
+\hl{I feel like we maybe need yet another section about customization. We have not yet described the following (rather important(?)) features:
+\begin{itemize}
+\item The mode argument - controlling which SVC steps are performed
+\item smartNum - chooosing whether or not numeric/integer variables with only a few levels are treated as factors
+\item standAlone - Describing how it is possible to produce .Rmd files that can be included in other .Rmd files using the child option and standAlone = F (removing the YAML preamble (and maybe also the cleanR commerical?
+Are we that nice?)).
+\end{itemize}
+}
+
+\subsection{A worked example}
+\hl{Maybe use the examples from the documentation to construct new summaryFunction etc. and add them to the default options. }
+
+\section{To-dos in the code}
+\hl{ ...  introduced by discrepancy between what I say we have done and what we have actually done:
+\begin{itemize}
+\item summaryFunction stuff:
+	\begin{itemize}
+		\item Make constructor function
+		\item Make allSummaryFunctions()
+		\item Change class of all summary (and description) functions into summaryFunction (possibly letting description functions be a subclass?)
+		\item Maybe: Make clean check if summary functions are really summaryFunctions?
+	\end{itemize}
+\item visualFunction stuff:
+	\begin{itemize}
+		\item Make constructor function
+		\item Make allVisualFunctions()
+		\item Change class of all visual functions into visualFunction
+		\item Maybe: Make clean check if visual functions are really visual functions?
+	\end{itemize}
+\item checkFunction stuff:
+	\begin{itemize}
+		\item Make allCheckFunctions()
+		\item Work on messageGenerator() to make it better suited for being an exported function + export it
+		\item Make allVisual()
+	\end{itemize}
+\end{itemize}
+}
+
 \section{Using the online web-app} \label{sec:web-app}

### 2016-10-16 R/basicVisual.R

@@ -4,13 +4,13 @@
 #' Plot the distribution of a variable, depending on its data class, using the base R
 #' plotting functions.
 #'
-#' @inheritParams standardVisual
-#'
 #' For character, factor, logical and labelled variables, a barplot is produced. For numeric
 #' or integer variables, \code{basicVisual} produces a histogram instead. Note that for
 #' integer and numeric variables, all non-finite (i.e. \code{NA}, \code{NaN}, \code{Inf}) values are
 #' removed prior to plotting. For character, factor, labelled and logical variables, only \code{NA}
 #' values are removed.
+#'
+#' @inheritParams standardVisual
 #'
 #' @examples
 #'  \dontrun{
@@ -25,7 +25,6 @@
 #' @seealso \code{\link{visualize}}, \code{\link{standardVisual}}
 #'
 #' @inheritParams standardVisual
-#'
 #' @importFrom stats na.omit
 #' @importFrom graphics plot hist
 #' @export

### 2016-10-16 R/standardVisual.R

@@ -8,7 +8,6 @@
 #' removed prior to plotting. For character, factor, labelled and logical variables, only \code{NA}
 #' values are removed.
 #'
-#'
 #' @param v The variable (vector) to be plotted.
 #' @param vnam The name of the variable which will appear as the title of the plot.
 #' @param doEval If TRUE, the plot itself is returned. Otherwise, the function returns

### 2016-10-16 latex/article.Rnw
Man pages + TeX

@@
 \tikzstyle{arrow} = [thick,->,>=stealth]
+
+\newcommand{\hl}[1]{\textcolor{magenta}{#1}}
+\newcommand{\R}{\proglang{R}}
 
@@
 
-In \pkg{cleanR} the clean function
+In \pkg{cleanR} the clean function is the primary workhorse and it
+should be an ideal starting place
+
+First we
+<<>>=
+library(cleanR)
+data(diamonds)
+head(diamonds)
+data(testData)
+#head(testData)
+@
+
+To produce summary data-cleaning document
 
 <<>>=
-library(cleanR)
-data(testData)
-head(testData)
+clean(testData, replace=TRUE, openResult=FALSE)
 @
 
-<<>>=
-                                        # clean(testData)
-2+3
-@
+By default, a ... is produced and the result is shown in
+Figure~\ref{fig:example1}.
 
-% \includepdf[fitpaper=true, pages=-]{test.pdf}
+\begin{figure}[!htp]
+\begin{center}
+\includegraphics[page=2]{cleanR_testData.pdf}
+%\includepdf[pages={2}, pagecommand={}]{cleanR_testData.pdf}
+\end{center}
+\label{fig:example1}
+\caption{Example output from running XXXX on the testData
+  dataset. First a summary of the full dataset, }
+\end{figure}
 

### 2016-10-16 latex/article.tex
Man pages + TeX

@@
 
-In \pkg{cleanR} the clean function
+In \pkg{cleanR} the clean function is the primary workhorse and it
+should be an ideal starting place
+
+First we
+\begin{Schunk}
+\begin{Sinput}
+> library(cleanR)
+> data(diamonds)
+> head(diamonds)
+\end{Sinput}
+\begin{Soutput}
+  carat       cut color clarity depth table price    x    y    z
+1  0.23     Ideal     E     SI2  61.5    55   326 3.95 3.98 2.43
+2  0.21   Premium     E     SI1  59.8    61   326 3.89 3.84 2.31
+3  0.23      Good     E     VS1  56.9    65   327 4.05 4.07 2.31
+4  0.29   Premium     I     VS2  62.4    58   334 4.20 4.23 2.63
+5  0.31      Good     J     SI2  63.3    58   335 4.34 4.35 2.75
+6  0.24 Very Good     J    VVS2  62.8    57   336 3.94 3.96 2.48
+\end{Soutput}
+\begin{Sinput}
+> data(testData)
+> #head(testData)
+\end{Sinput}
+\end{Schunk}
+
+To produce summary data-cleaning document
 
@@
 \begin{Sinput}
-> library(cleanR)
-> data(testData)
-> head(testData)
+> clean(testData, replace=TRUE, openResult=FALSE)
 \end{Sinput}
-\begin{Soutput}
-  charVar factorVar numVar intVar boolVar keyVar emptyVar _joeVar jack__var
-1       a         a      1      1    TRUE      1        1       1         1
-2       b         b      2      2   FALSE      2        1       2         2
-3       c         c      3      3    TRUE      3        1       3         3
-4       a         a      4      4    TRUE      4        1       4         4
-5       b         b      5      5    TRUE      5        1       5         5
-6       d         d      6      6   FALSE      6        1       6         6
-  numOutlierVar smartNumVar      cprVar   cprKeyVar miscodedMissingVar
-1             1           0 010101-1111 010101-1111                  .
-2             2           0 020102-2929 020102-2929                   
-3             3           0 121201-1902 121201-1902                nan
-4             4           0 030729-2222 030729-2222                NaN
-5             5           0 080909-1212 080909-1212                NAN
-6             6           0 010101-1111 020202-0101                 na
-\end{Soutput}
 \end{Schunk}
 
-\begin{Schunk}
-\begin{Sinput}
->                                         # clean(testData)
-> 2+3
-\end{Sinput}
-\begin{Soutput}
-[1] 5
-\end{Soutput}
-\end{Schunk}
+By default, a ... is produced and the result is shown in
+Figure~\ref{fig:example1}.
 
-% \includepdf[fitpaper=true, pages=-]{test.pdf}
+\begin{figure}[!htp]
+\begin{center}
+\includegraphics[page=2]{cleanR_testData.pdf}
+%\includepdf[pages={2}, pagecommand={}]{cleanR_testData.pdf}
+\end{center}
+\label{fig:example1}
+\caption{Example output from running XXXX on the testData
+  dataset. First a summary of the full dataset, }
+\end{figure}
 

### 2016-10-18 latex/article.Rnw

@@
 \usepackage{pdfpages}
+\usepackage{xspace}
 \usepackage{tikz}
@@
 \newcommand{\hl}[1]{\textcolor{magenta}{#1}}
-\newcommand{\R}{\proglang{R}}
+\newcommand{\R}{\proglang{R}\xspace}
 
@@
 %% almost as usual
-\author{Claus Thorn Ekstr\o m\\Biostatistics\\Department of Public
-  Health\\University of Copenhagen \And
-        Anne Helby Petersen\\Biostatistics\\Department of Public
+\author{Anne Helby Petersen\\Biostatistics\\Department of Public
+  Health\\University of Copenhagen \And Claus Thorn Ekstr\o m\\Biostatistics\\Department of Public
   Health\\University of Copenhagen}
-\title{\pkg{cleanR}: Your Personal Maid for Cleaning Datasets in \proglang{R}}
+\title{\pkg{cleanR}: Maid for Cleaning Datasets in \proglang{R}}
 
 %% for pretty printing and a nice hypersummary also set:
-\Plainauthor{Claus Thorn Ekstr\o m, Anne Helby Petersen} %% comma-separated
-\Plaintitle{{cleanR}: Your maid for cleaning datasets in R} %% without formatting
+\Plainauthor{Anne Helby Petersen, Claus Thorn Ekstr\o m} %% comma-separated
+\Plaintitle{{cleanR}: maid for cleaning datasets in R} %% without formatting
 \Shorttitle{\pkg{cleanR}: maid for cleaning datasets in R} %% a short title (if necessary)
@@
   E-mail: \email{ekstrom@sund.ku.dk}\\
-  URL: \url{http://eeecon.uibk.ac.at/~zeileis/}
+  URL: \url{http://staff.pubhealth.ku.dk/~ekstrom/}
 }
@@
 package to create a summary of potential errors. Section
-\ref{sec:internals} illustrates the methodology of modeling
-proportional data using simplex generalized regression. The
-generalized estimating equations for longitudinal proportional
-outcomes are given in Section 4. Then we address model diagnostics in
-Section 5. Section 6 presents the details of the simplexreg
-package. Section 7 further conducts analyses based on the simplex
-distribution in R with real data sets. Finally, plans for extending
-the package are described in Section 8.
+\ref{sec:internals} presents the steps that comprise the creation of
+the automated data cleaning that is part of the \code{clean} function,
+while the \ref{sec:extending} Section explains how the set of checks
+and graphical visualisations that are part of the cleaning suite are
+extended by custom, used-written additions. Finally, plans for
+extending the package are described in Section \ref{sec:conclusion}.
 
@@
 
-In \pkg{cleanR} the clean function is the primary workhorse and it
-should be an ideal starting place
+In \pkg{cleanR} the \code{clean} function is the primary workhorse and it
+should be the primary entry point for data cleaning.
 

### 2016-10-19 latex/article.Rnw
Search for strange speced in tables done

@@
 Though the discussion in the above paints a picture of cleanR as a user-friendly package which requires practically no knowledge of R, one should not be mistaken to think that it is not customizable.
-In fact, the main function of cleanR, \R{clean}, is mainly a tool for formatting the results from various checking-, summary- and visualizaion functions \hl{as described...
+In fact, the main function of cleanR, \code{clean}, is mainly a tool for formatting the results from various checking-, summary- and visualizaion functions \hl{as described...
 blabla}.
-Thus, the actual work underlying a cleanR output file can be anything or nothing - depending on the arguments given to \R{clean}.
+Thus, the actual work underlying a cleanR output file can be anything or nothing - depending on the arguments given to \code{clean}.
 This section consists of three parts.
-We commence this section with an overview of how contents are controlled in \R{cleanR} in terms of what cleaning steps are performed.
+We commence this section with an overview of how contents are controlled in \pkg{cleanR} in terms of what cleaning steps are performed.
 Secondly, we turn to the possibilities of customizing what the final cleanR file looks like.
@@
 \end{enumerate}
-Each of these stages are controllable using appropriate function parameters in \R{clean}.
-In the above, we presented the default \R{cleanR} settings and how to tweak them into providing a slightly different data cleaning outputs.
+Each of these stages are controllable using appropriate function parameters in \code{clean}.
+In the above, we presented the default \pkg{cleanR} settings and how to tweak them into providing a slightly different data cleaning outputs.
 However, if for instance the dataset at hand requires completely different visualizations, more control is needed.
-\R{cleanR} uses three different types of functions for performing all stages in the above, namely \R{summaryFunction}s, \R{visualFunction}s and \R{checkFunction}s.
-They each have distinct input-output structures and most instances of the functions are build as \R{S3} generics with methods for different data classes.
-By understanding how to construct proper \R{summaryFunction}s, \R{visualFunction}s and \R{checkFunction}s, the entire contents of the \R{cleanR} output is at your hands.
+\pkg{cleanR} uses three different types of functions for performing all stages in the above, namely \code{summaryFunction}s, \code{visualFunction}s and \code{checkFunction}s.
+They each have distinct input-output structures and most instances of the functions are build as \code{S3} generics with methods for different data classes.
+By understanding how to construct proper \code{summaryFunction}s, \code{visualFunction}s and \code{checkFunction}s, the entire contents of the \pkg{cleanR} output is at your hands.
 Therefore, this section is dedicated to introducing each in turn.\\
@@
 \subsubsection{Writing a summaryFunction}
-Though \R{cleanR} provides a special class for \R{summaryFunction}s, there really is nothing special about these functions: They are nothing but regular functions with a certain input/output-structure.
+Though \pkg{cleanR} provides a special class for \code{summaryFunction}s, there really is nothing special about these functions: They are nothing but regular functions with a certain input/output-structure.
 Specifically, they all follow the template below:
@@
 \end{Verbatim}
-and we recommend furthermore adding them to the overview of all summary functions by converting them to proper \R{summaryFunction} objects:
+and we recommend furthermore adding them to the overview of all summary functions by converting them to proper \code{summaryFunction} objects:
 \begin{Verbatim}
@@
 \end{Verbatim}
-which adds the new function to the output of \R{allSummaryFunctions()}.
-Note that \R{v} is a vector and that \R{res} should be either a character string or something that will be printed as one.
+which adds the new function to the output of \code{allSummaryFunctions()}.
+Note that \code{v} is a vector and that \code{res} should be either a character string or something that will be printed as one.
 In other words, e.g.
 integers are allowed, but matrices are not.
-Though a lot of different things can go into the \R{summaryFunction} template, we recommend only using them for summarizing the features of a variable, and leaving tests and checks for the \R{checkFunction}s (presented below).
+Though a lot of different things can go into the \code{summaryFunction} template, we recommend only using them for summarizing the features of a variable, and leaving tests and checks for the \code{checkFunction}s (presented below).
 
 \subsubsection{Writing a visualFunction}
-\R{visualFunction}s are the functions that produce the figures of a \R{cleanR} output document.
+\code{visualFunction}s are the functions that produce the figures of a \pkg{cleanR} output document.
 Writing a visual function is slightly more complicated than writing a summary function.
-This follows from the fact that visualFunctions need to be able to output standalone code for plots in order for \R{clean} to build standalone rmarkdown files.
+This follows from the fact that visualFunctions need to be able to output standalone code for plots in order for \code{clean} to build standalone rmarkdown files.
 We recommend using the following structure:
@@
 \end{Verbatim}
-In this function, \R{v} is the variable to be visualized, \R{vnam} is its name (which should generally be passed to \R{title} or \R{main} arguments in plotting functions) and \R{doEval} controls whether the output is a plot (if \R{TRUE}) or a character string of standalone code for producing a plot (if \R{FALSE}).
-The latter \R{doEval} setting is not strictly necessary for its use in \R{clean}, but it makes it easier to assess what visualization options are available.
-In either case, it should be noted that all the parameters listed above, \R{v}, \R{vnam} and \R{doEval}, are mandatory, so they should be left as is (\hl{as are?}), even if you do not want to use them.
-As with \R{summaryFunction}s, an overview of all available \R{visualFunction}s in the environment can be obtained by calling
+In this function, \code{v} is the variable to be visualized, \code{vnam} is its name (which should generally be passed to \code{title} or \code{main} arguments in plotting functions) and \code{doEval} controls whether the output is a plot (if \code{TRUE}) or a character string of standalone code for producing a plot (if \code{FALSE}).
+The latter \code{doEval} setting is not strictly necessary for its use in \code{clean}, but it makes it easier to assess what visualization options are available.
+In either case, it should be noted that all the parameters listed above, \code{v}, \code{vnam} and \code{doEval}, are mandatory, so they should be left as is (\hl{as are?}), even if you do not want to use them.
+As with \code{summaryFunction}s, an overview of all available \code{visualFunction}s in the environment can be obtained by calling
 \begin{Verbatim}
@@
 \end{Verbatim}
-an overview of all plotting options applied on \R{v} is produced and opened as a html document for easy comparison. \hl{Should we mention the side effect of producing .rmd and .html files on disc?}
+an overview of all plotting options applied on \code{v} is produced and opened as a html document for easy comparison. \hl{Should we mention the side effect of producing .rmd and .html files on disc?}
 
 \subsubsection{Writing a checkFunction}
-The last, but also most important, \R{cleanR} function type is the \R{checkFunction}.
+The last, but also most important, \pkg{cleanR} function type is the \code{checkFunction}.
 These are the functions that flag issues in the data and control the flow of the overall data cleaning process in the precheck stage.
-A \R{checkFunction} can be written using the following template:
+A \code{checkFunction} can be written using the following template:
 \begin{Verbatim}
@@
 \end{Verbatim}
-Only the input parameter (\R{v}) and the output format strictly has to follow this structure.
-However, we recommend using \R{messageGenerator} for consistent styling of all \R{checkFunction} messages.
-This function simply pastes together the \R{problemMessage} and the \R{problemValues}, with the latter being quoted and sorted alphabetically.
-Note that printing quotes in rmarkdown requires an extensive amount of character escaping, so opting for \R{messageGenerator} really is the easiest solution.
+Only the input parameter (\code{v}) and the output format strictly has to follow this structure.
+However, we recommend using \code{messageGenerator} for consistent styling of all \code{checkFunction} messages.
+This function simply pastes together the \code{problemMessage} and the \code{problemValues}, with the latter being quoted and sorted alphabetically.
+Note that printing quotes in rmarkdown requires an extensive amount of character escaping, so opting for \code{messageGenerator} really is the easiest solution.
 
-While the descriptions of \R{summaryFunction}s and \R{visualFunction}s are only for internal use in the \R{allSummaryFunctions()} and \R{allVisualFunctions()} outputs, respectively, \R{checkFunction} descriptions are actually visible in the \R{cleanR} output document.
+While the descriptions of \code{summaryFunction}s and \code{visualFunction}s are only for internal use in the \code{allSummaryFunctions()} and \code{allVisualFunctions()} outputs, respectively, \code{checkFunction} descriptions are actually visible in the \pkg{cleanR} output document.
 These are the brief descriptions presented in Part 1 \hl{eh?
@@
 Whatever we call this section in the above} in the output document.
-If a \R{checkFunction} does not have a description (for instance, if it is just a regular \R{function} using the \R{checkFunction} input/output-structure), the function name will be printed instead of the description.
+If a \code{checkFunction} does not have a description (for instance, if it is just a regular function using the \code{checkFunction} input/output-structure), the function name will be printed instead of the description.
 

### 2016-10-19 R/clean.R

@@ -293,6 +293,7 @@ clean <- function(o,
     ## panderOptions("table.alignment.default", "left")
     panderOptions('table.alignment.default', 'center')  ## XXX CE only one of these two
     panderOptions("table.split.table", Inf)
+    panderOptions("table.split.cells", Inf)
     panderOptions('table.alignment.rownames', 'left')
 
     ##
@@ -505,8 +506,10 @@ clean <- function(o,
                 if (extraMessages$do) writer(paste("* ", extraMessages$messages, "\n", collapse=" \n ",
                                                    sep=""))
 
-            ## make Summary table
-            if (doSummarize) sumTable <- pander_return(summarize(v, characterSummaries = characterSummaries,
+
+
+                ## make Summary table
+                if (doSummarize) sumTable <- pander_return(summarize(v, characterSummaries = characterSummaries,
                                                                  factorSummaries = factorSummaries,
                                                                  labelledSummaries = labelledSummaries,
                                                                  numericSummaries = numericSummaries,
@@ -514,7 +517,7 @@ clean <- function(o,
                                                                  logicalSummaries = logicalSummaries,
                                                                  ...), justify="lr")
               #if (doSummarize) sumTable <- pandoc.table.return(summarize(v, ...))
-                #exactly the same result as with pander_return()
+                                        #exactly the same result as with pander_return()
 
                 ## Label information
                 ## Right now we are not doing anything besides wirint the label above

### 2016-10-19 R/identifyMissing.R

@@ -86,29 +86,30 @@ identifyMissNine <- function(v) {
 
 #factor and character variables
 identifyMissingCF <- function(v) {
-  v <- na.omit(v)
-  problem <- FALSE
-  problemValues <- NULL
+    v <- na.omit(v)
+    problem <- FALSE
+    problemValues <- NULL
 
-  missStrs <- c("", "nan", "NaN", "NAN", "na", "NA", "Na", "Inf", "inf",
-                "-Inf", "-inf", "-", " ", "9")
-  missStrsOcc <- intersect(v, missStrs) #what potential missing value strings occur?
+    missStrs <- c("", "nan", "NaN", "NAN", "na", "NA", "Na", "Inf", "inf",
+                  "-Inf", "-inf", "-", " ", "9")
+    missStrsOcc <- intersect(v, missStrs) #what potential missing value strings occur?
 
-  missNinesOcc <- identifyMissRepChar(v, "9") #what 99, 999, ... strings occur?
-  missSpaceOcc <- identifyMissRepChar(v, " ") #what "  ", "   ", ... strings occur?
+    missNinesOcc <- identifyMissRepChar(v, "9") #what 99, 999, ... strings occur?
+    missSpaceOcc <- identifyMissRepChar(v, " ") #what "  ", "   ", ... strings occur?
 
-  #STATA-style: .something describes a "something" type of missing value
-  missDotPrefixOcc <- unique(v[substr(v, 1, 1) == "."])
+    ## STATA-style: .something describes a "something" type of missing value
+    missDotPrefixOcc <- unique(v[substr(v, 1, 1) == "."])
 
-  allProblemOcc <- c(missStrsOcc, missNinesOcc, missSpaceOcc, missDotPrefixOcc)
+    allProblemOcc <- c(missStrsOcc, missNinesOcc, missSpaceOcc, missDotPrefixOcc)
 
-  if (length(allProblemOcc) > 0) {
-    problemValues <- allProblemOcc
-    problem <- TRUE
-  }
-  outMessage <- messageGenerator(list(problem = problem, problemValues = problemValues),
-                                 check = "identifyMiss")
-  list(problem = problem, message = outMessage)
+    if (length(allProblemOcc) > 0) {
+        problemValues <- allProblemOcc
+        problem <- TRUE
+    }
+
+    outMessage <- messageGenerator(list(problem = problem, problemValues = problemValues),
+                                   check = "identifyMiss")
+    list(problem = problem, message = outMessage)
 }
 
 #labbeled variables

### 2016-10-19 R/clean.R
documentation

@@ -746,6 +746,10 @@ description.default <- function(x) deparse(substitute(x))
 description.checkFunction <- function(x) attr(x, "description")
 description.summaryFunction <- function(x) attr(x, "description")
 description.visualFunction <- function(x) attr(x, "description")
+`description<-` <- function(x, value) {
+  attr(x, "description") <- value
+  x
+}
 
 #' @export
 classes <- function(x) UseMethod("classes")
@@ -753,7 +757,10 @@ classes.defualt <- function(x) NULL
 classes.checkFunction <- function(x) attr(x, "classes")
 classes.summaryFunction <- function(x) attr(x, "classes")
 classes.visualFunction <- function(x) attr(x, "classes")
-
+`classes<-` <- function(x, value) {
+  attr(x, "classes") <- value
+  x
+}
 
 #' @export
 print.functionSummary <- function(x) {

### 2016-10-19 R/summaryFunction.R

@@ -1,6 +1,67 @@
+#' @title Construct a summaryFunction  
 #'
-#'document me!
+#' @description Convert a function into class \code{summaryFunction} used for creating variable 
+#' summaries in \code{\link{summarize}}, typically called from \code{\link{clean}}.
+#' 
+#' @param f A function. In order to be a valid \code{summaryFunction}, \code{f} should 
+#' follow the structure outlined in details below.
+#' 
+#' @param description A character string with a (brief!) description of \code{f}.
+#' 
+#' @param classes A vector of classes that \code{f} is designed to be called on. If NULL (the default) 
+#' one of two standard solutions are tried: If \code{f} is a S3 generic function, the \code{classes}
+#' argument will automatically be filled out by checking what methods are available in the global
+#' environment. If \code{f} is not a S3 generic function, the \code{classes} attribute is set to an
+#' empty character object (i.e. \code{character(0)}). 
+#' 
+#' @return A function of class \code{summaryFunction}, i.e. with two attributes: A description 
+#' ("description") and a vector of classes which the function can be called upon ("classes"). A 
+#' summary function always takes a vector (\code{v}) as its sole argument and returns a list with two
+#' slots: \code{$feature} (a character string with the name of the summary) and \code{$results} 
+#' (a character string with the result of the summary operation). See e.g. \code{\link{variableType}} or
+#' \code{\link{countMissing}} for examples of \code{summaryFunction}s.
+#' 
+#' @details The structure of a valid \code{summaryFunction} follows the following structure:
+#' \code{
+#'    mySummaryFunction <- function(v) {
+#'      res <- [result of whatever summary we are doing]
+#'      list(feature = "Feature name", result = res)
+#'    }
+#'  }
+#'  followed by a class change:
+#'  \code{
+#'    mySummaryFunction <- summaryFunction(mySummaryFunction, description = "[describe function]",
+#'                                         classes = c([classes]))
+#'  }
+#'  Note that it is not necessary to formally change functions to be used by \code{\link{summarize}}
+#'  into \code{summaryFunction}s, but it is recommended, as the functions are then added to the output 
+#'  of \code{allSummaryFunctions()}, making it easier to
+#'  gain an overview of the summary functions available for a \code{clean()} run.
+#'        
 #'
+#' @seealso \code{\link{allSummaryFunctions}} \code{\link{summarize}}
[…]

### 2016-10-19 R/visualFunction.R

@@ -1,8 +1,71 @@
+#' @title Construct a visualFunction  
 #'
-#' Visual funtction
+#' @description Convert a function into class \code{visualFunction} used for creating variable 
+#' visualizations \code{\link{visualize}}, typically called from \code{\link{clean}}.
+#' 
+#' @param f A function. In order to be a valid \code{visualFunction}, \code{f} should 
+#' follow the structure outlined in details below.
+#' 
+#' @inheritParams summaryFunction
+#' 
+#' @return A function of class \code{visualFunction}, i.e. with two attributes: A description 
+#' ("description") and a vector of classes which the function can be called upon ("classes"). A 
+#' visualFunction always 
+#' 
+#' 
+#'  
+#' @details We recommend using the following structure for defining new \code{visualFunction}s:
+#' #' \code{
+#'   myVisualFunction <- function(v, vnam, doEval) {
+#'    thisCall <- call("[the name of the function used to produce the plot]", 
+#'                     v, [additional arguments to the plotting function])
+#'    if (doEval) {
+#'     return(eval(thisCall))
+#'    } else return(deparse(thisCall)
+#'   }
+#'  }
+#'  followed by a class change:
+#'  \code{
+#'    myVisualFunction <- visualFunction(myVisualFunction, description = "[describe function]",
+#'                                         classes = c([classes]))
+#'  }
+#'  It is not strictly necessary to follow this exact template, but the input/output structure 
+#'  should be respected - in other words, the arguments are mandatory and the function needs
+#'  to return a character string containing code (if \code{doEval = FALSE}) or have the 
+#'  side effect of producing a plot (if \code{doEval = TRUE}). See examples below for an example
+#'  of how the template can be used in practice.  
+#'  
+#'  Note that it is not necessary to formally change functions to be used by \code{\link{visualize}}
+#'  into \code{visualFunction}s, but it is recommended, as the functions are then added to the output 
+#'  of \code{allVisualFunctions()}, making it easier to gain an overview of the visualization 
+#'  functions available for a \code{clean()} run.
+#'        
 #'
-#' document me!
+#' @seealso \code{link{allVisualFunctions}} \code{\link{visualize}} 
+#' \code{\link{description}} \code{\link{classes}} 
+#'   
+#' @examples 
+#' #Defining a new visualFunction:
+#'  mosaicVisual <- function(v, vnam, doEval) {
[…]

### 2016-12-13 latex/article.tex
Småopdateringer af artikel

@@
 \newcommand{\hl}[1]{\textcolor{magenta}{#1}}
-\newcommand{\R}{\proglang{R}\xspace}
+
+%Virker ikke:
+%\newcommand{\R}{\proglang{R}\xspace}
+
+\newcommand{\R}[1]{\code{#1}}
 
@@
 %% almost as usual
-\author{Anne Helby Petersen\\Biostatistics\\Department of Public
+\author{Anne H. Petersen\\Biostatistics\\Department of Public
   Health\\University of Copenhagen \And Claus Thorn Ekstr\o m\\Biostatistics\\Department of Public
@@
 \proglang{R}, but there are only a few options available for automated data
-cleaning.
+cleaning. \hl{Hov! Er det ikke et job for autocleanR?}
 
@@
 
-In \pkg{cleanR} the \code{clean} function is the primary workhorse and it
-should be the primary entry point for data cleaning.
+\begin{table}
+\begin{tabular}{p{0.2\linewidth}p{0.5\linewidth}p{0.2\linewidth}}
+Argument & Description & Default value \\
+\hline
+\R{output} & Type of output file to be produced (html, pdf or no file) & \R{"pdf"} \\
+\R{ordering} & Ordering of the variables in the data summary & \R{"asIs"} \\
+\R{onlyProblematic} & Should only variables flagged as problematic be included in the output? & \R{FALSE} \\
+\R{maxProbVals} & Maximum number of problematic values to print, if any are found in data checks & \R{Inf} \\
+\R{maxDecimals} & Maximum number of decimals to print for numeric problematic values & \R{2}
+\end{tabular}
+\caption{A selection of commonly used arguments of \R{clean()}.}
+\label{table.cleanFormals}
+\end{table}
+
+In \pkg{cleanR} the \code{clean} function is the primary workhorse and knowledge of nothing more than this function is required to produce data cleaning outputs in \pkg{cleanR}.
+The data cleaning output itself is an overview document, intended for reading by humans, in either pdf or html format.
 
@@
 
-\section{The structure of cleanR} \label{sec:internals}
+\section{Behind the scenes} \label{sec:internals}
 
@@
 \hl{ ...  introduced by discrepancy between what I say we have done and what we have actually done:
-\begin{itemize}
-\item summaryFunction stuff:
-	\begin{itemize}
-		\item Make constructor function
-		\item Make allSummaryFunctions()
-		\item Change class of all summary (and description) functions into summaryFunction (possibly letting description functions be a subclass?)
-		\item Maybe: Make clean check if summary functions are really summaryFunctions?
-	\end{itemize}
-\item visualFunction stuff:
-	\begin{itemize}
-		\item Make constructor function
-		\item Make allVisualFunctions()
-		\item Change class of all visual functions into visualFunction
-		\item Maybe: Make clean check if visual functions are really visual functions?
-	\end{itemize}
-\item checkFunction stuff:
-	\begin{itemize}
-		\item Make allCheckFunctions()
-		\item Work on messageGenerator() to make it better suited for being an exported function + export it
-		\item Make allVisual()
-	\end{itemize}
-\end{itemize}
+(nothing right now)
 }

### 2016-12-14 latex/article_vol2.tex
Added examples in section 5

\documentclass[article]{jss}
\usepackage[utf8]{inputenc}
\usepackage[nogin]{Sweave}
\usepackage{pdfpages}
\usepackage{xspace}
\usepackage{tikz}
\usetikzlibrary{shapes.geometric, arrows}

\tikzstyle{io} = [trapezium, trapezium left angle=70, trapezium right angle=110, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=blue!30]
\tikzstyle{process} = [rectangle, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=orange!30]
\tikzstyle{decision} = [diamond, minimum width=3cm, minimum height=1cm, text centered, draw=black, fill=green!30]
\tikzstyle{arrow} = [thick,->,>=stealth]

\newcommand{\hl}[1]{\textcolor{magenta}{#1}}

%Virker ikke:
%\newcommand{\R}{\proglang{R}\xspace}

\newcommand{\R}[1]{\code{#1}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% declarations for jss.cls %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% almost as usual
\author{Anne H. Petersen\\Biostatistics\\Department of Public
  Health\\University of Copenhagen \And Claus Thorn Ekstr\o m\\Biostatistics\\Department of Public
  Health\\University of Copenhagen}
\title{\pkg{cleanR}: Maid for Cleaning Datasets in \proglang{R}}

%% for pretty printing and a nice hypersummary also set:
\Plainauthor{Anne H. Petersen, Claus Thorn Ekstr\o m} %% comma-separated
\Plaintitle{{cleanR}: maid for cleaning datasets in R} %% without formatting
\Shorttitle{\pkg{cleanR}: maid for cleaning datasets in R} %% a short title (if necessary)

%% an abstract and keywords
\Abstract{Data cleaning and -validation are the first steps in any data
  analysis, as the validity of the conclusions from the analysis are
  hinged on the quality of the input data. Mistakes in the data can arise for any number of reasons, including erroneous codings, malfunctioning measurement equipment, inconsistent data generation manuals and many more.  Ideally, a human investigator
  should go through each variable in the dataset and look for
  potential errors --- both in input values and coding.  \hl{segway}

  We describe an \proglang{R} package which implements an extensive
  and customizeable suite of checks to be applied to the variables in
  a dataset in order to identify potential problems in the
  corresponding variables. The results can be presented in an auto-generated, non-technical, stand-alone
  overview document, intended to be perused by an investigator with an
  understanding of the variables in the data and the experimental design, but not necessarily knowledge of \R{R}. Thereby, \R{cleanR} aids the dialogue between data analysist and field experts, while also providing easy documentation of data cleaning steps and data quality control. Moreover, the \R{cleanR} solution changes the data cleaning process from the usual ad hoc approach to a systematic, well-documented endeavor. 
  \R{cleanR} also provides a suite of more typical \R{R} tools for interactive data quality assessment and -cleaning.

  The \pkg{cleanR} package is designed to be easily extended with custom
  user-created checks that are relevant in particular situations.
}
\Keywords{data cleaning, quality control, \proglang{R}}
\Plainkeywords{data cleaning, quality control, R} %% without formatting
%% at least one keyword must be supplied

%% publication information
%% NOTE: Typically, this can be left commented and will be filled out by the technical editor
%% \Volume{50}
%% \Issue{9}
%% \Month{June}
%% \Year{2012}
%% \Submitdate{2012-06-04}
%% \Acceptdate{2012-06-04}

%% The address of (at least) one author should be given
%% in the following format:
\Address{
  Claus Thorn Ekstr\o m\\
  Biostatistics, Department of Public Health\\
  University of Copenhagen\\
  Denmark\\
  E-mail: \email{ekstrom@sund.ku.dk}\\
  URL: \url{http://staff.pubhealth.ku.dk/~ekstrom/}
}
%% It is also possible to add a telephone and fax number
%% before the e-mail in the following format:
%% Telephone: +43/512/507-7103
%% Fax: +43/512/507-2851

%% for those who use Sweave please include the following line (with % symbols):
%% need no \usepackage{Sweave.sty}

%% end of declarations %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\section{Introduction}
Statisticians and data analysts spend a large portion of their time on
data cleaning and on data wrangling. Packages such as
\pkg{data.table}, and \pkg{plyr} have made data wrangling a lot easier in
\proglang{R}, but there are only a few tools available for data cleaning. \hl{What are these tools? More here about the solutions already available.}

\hl{A text here (with offset in what tools are already available) that
\begin{itemize}
\item Makes the point that data cleaning is usually done in a very ad hoc manner
\item Maybe hints to the discussion later about poor documentation of data cleaning/quality assessment as the status quo for most analysts
\item Gives a specification of what we mean by data errors/mistakes/whatever. I'm not completely sure this will be clear to everyone.
\end{itemize}
}

Data cleaning is a time consuming endeavor, as it inherently requires
human interaction since every dataset is different and the variables
in the dataset can only be understood in the proper context of their origin. 
\hl{Segway: Something like maybe this is why proper data cleaning is not always done...}
In many situations these errors are discovered in the process of the
data analysis (e.g., \hl{a categorical variable with numeric labels is
wrongly classified as a numeric variable\footnote{Is that something we can fix in clean()?}} or a variable where all values have erroneously been coded to the same value), but in other cases a human
with knowledge about the data context area is needed to identify
possible mistakes in the data (e.g., if there are 4 categories for a
variable that should only have 3).  \hl{segway} \\

The \R{cleanR} approach to data cleaning and -quality assessment is characterized by two principles. First of all, there is no need for data cleaning to be an ad hoc procedure. Often, we have a very clear idea of what flags are raisable in a given dataset before we look at it, as we were the ones to produce it in the first place. This means that data cleaning can easily be a well-documented, well-specified procedure. This angle on data cleaning also adds to the transparency of this first step of data analysis: Though data cleaning might be regarded as a somewhat dull activity, adequate data cleaning is crucial in any data analysis. In order to aid this principle, \R{cleanR} provides easy-to-use, automated tools for data quality assessment in \R{R} on which data cleaning decisions can be build. This quality assessment is presented in a auto-generated overview document, readable by data analysts and field experts alike, thereby also contributing to a inter-field dialogue about the data at hand. Oftentimes, e.g. separating faulty codings of a numeric value from unusual, but correct, values requires problem-specific expertise that might not be held by the data analyst. Hopefully, having easy access to data descriptions through \R{cleanR} will help this necessary knowledge sharing.

While \R{cleanR} is build for auto-generating data quality assessment overview documents, we also wish to emphasize that it is \textit{not} a tool for automatic data cleaning. This qualifies as the second principle of \R{cleanR}: Data cleaning decisions should always be made by humans. Therefore, \R{cleanR} does not provide any tools for "fixing" errors in the data. However, we do provide interactive functions that can be used to identify potentially erroneous entries in a dataset and that can make it easier to solve data issues.\\

This manuscript is structured as follows: First, we introduce the representative of the first principle, namely the \R{clean()} function, which generates data cleaning overview documents. In the \R{cleanR} package, we have provided a number of default cleaning steps that cover the data cleaning challenges, we find to be most common. However, every dataset is different, and some datasets might include problems that cannot be detected by our data checking functions. In Section 3 \hl{proper reference here}, we turn to the question of how such \R{cleanR} extensions can be made such that they are integrable with the \R{clean()} function and with other tools available in \R{cleanR}. In this section, we also present the interactive mode of \R{cleanR}, as motivated by the second principle above. At last, in Section 4 \hl{proper ref}, we discuss a number of examples of specific data cleaning challenges and how \R{cleanR} can be used to solve them.

%% include your article here, just as usual
%% Note that you should use the \pkg{}, \proglang{} and \code{}
%% commands.

\section{Something like "producing a data cleaning overview"} \label{sec:example1}

In \pkg{cleanR}, the \code{clean} function is the primary workhorse and knowledge of nothing more than this function is required to produce data cleaning outputs. The data cleaning output itself is an overview document, intended for reading by humans, in either pdf or html format. Appendix \hl{something} provides an example of a data cleaning output document, produced by calling \R{clean} on the dataset \R{toyData} available in  \R{cleanR}. The first two pages of this data cleaning output are also available in Figure~\ref{fig:example1}. \R{toyData} is a very small ($15$ by $6$), artificial dataset, whose only purpose is to illustrate the main capabilities of \R{cleanR}. The following commands load the dataset and produce the cleaning output:

\begin{Schunk}
\begin{Sinput}
> library(cleanR)
> data(toyData)
> toydata
\end{Sinput}
\begin{Soutput}
   var1 var2  var3        var4 var5       var6
1   red    1     a -0.65959383    1 Irrelevant
2   red    1     a  0.08671649    2 Irrelevant
3   red    1     a -0.10951326    3 Irrelevant
4   red    2     a  0.08630221    4 Irrelevant
5   red    2     a -1.84311184    5 Irrelevant
6   red    6     b  0.92210680    6 Irrelevant
7   red    6     b  1.01921086    7 Irrelevant
8   red    6     b -0.92428326    8 Irrelevant
9   red  999     c -0.65340163    9 Irrelevant
10  red   NA     c  0.21133941   10 Irrelevant
11 blue    4     c  0.91783009   11 Irrelevant
12 blue   82     .  0.10313983   12 Irrelevant
13 blue   NA        0.16954218   13 Irrelevant
14 <NA>  NaN other  0.41967230   14 Irrelevant
15 <NA>    5 OTHER  0.77143836   15 Irrelevant
\end{Soutput}
\begin{Sinput}
> clean(toyData)
\end{Sinput}
\end{Schunk}

By default, a pdf overview document is produced, saved to the disc and opened for inspection. Turning to Figure~\ref{fig:example1}, we see that such a data cleaning output document consists of two parts. First, an overview of what was done is presented under the title \textit{Data cleaning summary}. Secondly, each variable in the dataset is presented in turn using (up to) three tools in the \textit{Variable list}: A table summarizing key features of the variable, a figure visualizing its distribution and potentially also a list of flagged issues. For instance, in the \R{numeric}-type variable \R{var2} from \R{toyData}, \R{clean()} has identified two values that are suspected to be miscoded missing values (\R{999} and \R{NaN}), while two values were also flagged as potential outliers that should be investigated more carefully. We can then return to Part 1, the data cleaning summary, and inspect what sorts of checks were performed on this variable. \hl{hm, delete the last part? or use different variable? Only these exact two checks were performed}.

Of course, 
\\

\begin{figure}[!htp]
\begin{center}
\includegraphics[width=8cm,page=2]{cleanR_toyData.pdf} 
\includegraphics[width=8cm, page=3]{cleanR_toyData.pdf} 
%\includepdf[pages={2}, pagecommand={}]{cleanR_testData.pdf}

\end{center}
\label{fig:example1}
\caption{Example output from running XXXX on the testData
  dataset. First a summary of the full dataset,.. 
  \hl{Maybe fix these pdfs such that they are one file with no blankspace? I.e. basically we want the .html-version of this file printed as pdf.}}
\end{figure}

\subsection{Controlling contents}
\begin{table}
\small
\begin{tabular}{p{0.2\linewidth}p{0.5\linewidth}p{0.2\linewidth}}
\hline
Argument & Description & Default value \\
\hline

\smallskip Control \hl{...eh?}\\
\quad \R{useVar} & What variables should be used? & \R{NULL} (corresponding to all variables) \\
\quad \R{ordering} & Ordering of the variables in the data summary (as is or alphabetical) & \R{"asIs"} \\
\quad \R{onlyProblematic} & Should only variables flagged as problematic be included in the variable list? & \R{FALSE} \\
\quad \R{listChecks} & Should an overview of what checks were performed by listed in the \textit{Data cleaning summary}? &  \R{TRUE} \\

\smallskip Control SVC-step \\
\quad \R{mode} & What steps should be performed for each variable (out of the three possibilities \textit{summarize}, \textit{visualize}, \textit{check})? & \R{c("summarize", "visualize", "check")} \\
\quad \R{labelled\_as} & How should variables of class \R{labelled} be handled (as factors, is missing values or by ignoring labels)? & \R{"factor"} \\
\quad \R{smartNum} & Should numerical values with only a few unique levels be flagged an treated as factor variables? & \R{TRUE} \\
\quad \R{maxProbVals} & Maximum number of problematic values to print, if any are found in data checks & \R{Inf} \\
\quad \R{maxDecimals} & Maximum number of decimals to print for numeric values in the variable list & \R{2} \\
\quad \R{twoCol} & Should the summary table and visualizations be placed side-by-side (in two columns)? & \R{TRUE} \\

\smallskip Control output \\
\quad \R{output} & Type of output file to be produced (html, pdf or \hl{?? what does screen mean here??}) & \R{"pdf"} \\
\quad \R{render} & Should the output file be rendered from markdown? & \R{TRUE} \\
\quad \R{openResult} & If a  pdf/html file is rendered, should it automatically open afterwards? & \R{TRUE} \\

\hline
\end{tabular}
\caption{A selection of commonly used arguments of \R{clean()}.  \hl{I have everything here except \R{file} and \R{quiet}, replace, vol, standAlone + stuff related to selecting SVC-functions}}
\label{table.cleanFormals}
\end{table}

By the \textit{contents} of the cleanR output, we refer to every step that actually involves a function being called on variables from the dataset. There are three stages in which this occurs \hl{(with ref. to figure/flowchart?)}:
\begin{enumerate}
\item In the precheck functions
\item In the summarize/visualize/check (SVC) step
\item In the multivariate visuzalizations \hl{Or whatever, in the stuff that we have not yet implemented. I'm pretending like this stage doesn't exist below.}
\end{enumerate}
Each of these stages are controllable using appropriate function parameters in \code{clean}. In the above, we presented the default \pkg{cleanR} settings and how to tweak them into providing a slightly different data cleaning outputs. However, if for instance the dataset at hand requires completely different visualizations, more control is needed. \pkg{cleanR} uses three different types of functions for performing all stages in the above, namely \code{summaryFunction}s, \code{visualFunction}s and \code{checkFunction}s. They each have distinct input-output structures and most instances of the functions are build as \code{S3} generics with methods for different data classes. By understanding how to construct proper \code{summaryFunction}s, \code{visualFunction}s and \code{checkFunction}s, the entire contents of the \pkg{cleanR} output is at your hands. Therefore, this section is dedicated to introducing each in turn.\\

\hl{Something like a figure that gives an overview of all summaryFunctions, ... available in standard cleanR, including default settings and some kind of description of where they are called (in precheck or SVC step). Also something about what datatypes they can each be called on. Also, more text here about the relationship between the two (three) stages and the three function types.}

\subsection{Something about what check, visual and summary functions are available}

\begin{table}[h]
\small
\centering
\begin{tabular}{p{0.3\linewidth} p{0.35\linewidth} p{0.01\linewidth} p{0.01\linewidth} p{0.01\linewidth} p{0.01\linewidth} p{0.01\linewidth}
 p{0.01\linewidth} p{0.01\linewidth}}
  \hline
& Description &  \multicolumn{7}{c}{Relevant variable classes} \\ \smallskip
 & &  C & F & I & L & B & N & D\\ 
  \hline \smallskip
  \textbf{\R{summaryFunction}s}  \smallskip \\
  \quad \R{centralValue} & Compute median or mode &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\ 
  \quad \R{countMissing} & Compute ratio of missing observations &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\ 
  \quad \R{minMax} & Find minimum and maximum values &   &  & $\times$ & &  & $\times$ & $\times$  \\ 
  \quad \R{quartiles} & Compute 1st and 3rd quartiles &    &  & $\times$ & &  & $\times$ &  \\ 
  \quad \R{uniqueValue} & Count number of unique values &   $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\ 
  \quad \R{variableType} & Data class of variable & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\ 
  \smallskip \\
 \textbf{\R{visualFunction}s} \smallskip \\
  \quad \R{basicVisual} & Histograms and barplots using graphics &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\ 
  \quad \R{standardVisual} & Histograms and barplots using ggplot2 &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\ 
  \smallskip \\
 \textbf{\R{checkFunction}s} \smallskip \\
 \quad \R{identifyCaseIssues} & Identify case issues &  $\times$ & $\times$ & & & & &  \\ 
 \quad \R{identifyLoners} & Identify levels with $<$ 6 obs. & $\times$ & $\times$ & & & & &  \\ 
 \quad \R{identifyMissing} & Identify miscoded missing values &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &  \\ 
 \quad \R{identifyNums} & Identify misclassified numeric or integer variables & $\times$ & $\times$ & & & & &  \\ 
 \quad \R{identifyOutliers} & Identify outliers &  & & $\times$ & & $\times$ & \\ 
 \quad \R{identifyOutliersTBStyle} & Identify outliers (Turkish Boxplot style) &  & & $\times$ & & $\times$ &  \\ 
 \quad \R{identifyWhitespace} & Identify prefixed and suffixed whitespace &  $\times$ & $\times$ & & $\times$ & & &  \\ 
 \quad \R{isCPR} & Identify Danish CPR numbers & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &   \\ 
 \quad \R{isEmpty} & Check if the variable contains only a single value & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &   \\ 
 \quad \R{isKey} & Check if the variable is a key & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & \smallskip   \\ 
 \hline
\end{tabular}
\caption{\hl{Blabla, mention that C is character, F is factor, I is integer, L is labelled, B is logical (boolean), N is numeric and D is Date. \newline Also: Check that everything in here is correct (i.e. corresponds to the output of allSummaryFunctions() etc}}
\label{table.SVCfunctions}
\end{table}

\hl{
\begin{itemize}
\item Introduce the relevant \R{clean}-arguments and the \R{defaultWhateverSummaries} etc.-functions
\item Introduce \R{allSummaryFunctions()} etc. and present a table corresponding to the output of this call
\item Small example
\end{itemize}
}

\section{Something about the interactive cleanR mode}
While overview documents are great for documenting our work, they might not feel very natural for actually performing data cleaning or data wrangling based on the results we found. Therefore, \R{cleanR} also provides more standard \R{R} interactive tools, such as functions that print results to the console or saves it in a variable for later use. Or rather, the functions used by \R{clean()} to do actual data summaries, visualizations and checks, as described above, can also be used interactively. \hl{segway}

\hl{Maybe write here that this section introduces the three functions: \R{summarize()}, \R{check()} and \R{visualize()}?}

\subsection{Blabla: An example}
Let's say we wish to look further into a certain variable, namely \R{var2}, from \R{toyData}. The data cleaning summary found some issues, and we would like to recall what these issues were. This is done using the command
\begin{Schunk} 
\begin{Sinput} 
> check(toyData$var2)
\end{Sinput}
\begin{Soutput}
$identifyMissing
The following suspected missing value codes enter as regular values: 999, NaN.
$identifyOutliers
Note that the following possible outlier values were detected: 82, 999. $
\end{Soutput}
\end{Schunk}
\hl{delete that last \$, it's only there to make my TeX editor stop highlighting everything. Note: \$s cannot be escaped here, as Schunk is verbatim.}

Note that the arguments specifying which checks to perform, as described in \hl{XXXX} above, are in fact passed to \R{check()}, and thus they can also be used here. For instance, if we only want the result of the check for miscoded missing values, we write
\begin{Schunk}
\begin{Sinput}
> check(toyData$var2, numericChecks = "identifyMissing")
\end{Sinput}
\begin{Soutput}
$identifyMissing
The following suspected missing value codes enter as regular values: 999, NaN.
\end{Soutput}
\end{Schunk}
An equivalent way to call only a single, specific \R{checkFunction} is by using it directly on the variable, i.e. 
\begin{Schunk}
\begin{Sinput}
> identifyMissing(toyData$var2)
\end{Sinput}
\begin{Soutput}
The following suspected missing value codes enter as regular values: 999, NaN. $
\end{Soutput}
\end{Schunk}
\hl{delete that last \$}

The result of a \R{checkFunction} is an object of class \R{checkResult}. By using the structure function, \R{str()}, we can look further into its components:

\begin{Schunk}
\begin{Sinput}
> missVar2 <- identifyMissing(toyData$var2)
> str(missVar2)
\end{Sinput}
\begin{Soutput}
List of 3
 $ problem      : logi TRUE
 $ message      : chr "The following suspected missing value codes enter as 
 		 regular values: \\\"999\\\", \\\"NaN\\\"."
 $ problemValues: num [1:2] 999 NaN
 - attr(*, "class")= chr "checkResult"
\end{Soutput}
\end{Schunk}
The most important thing to note here is that while the printed message is made for human reading, the actual values of the variable causing the issue are still attainable. If we for instance decide that the values \R{999} and \R{NaN} in \R{var2} are in fact miscoded missing values, we can replace them easily:
\begin{Schunk}
\begin{Sinput}
> toyData$var2[toyData$var2 %in% missVar2$problemValues] <- NA
> identifyMissing(toyData$var2}
\end{Sinput}
\begin{Soutput}
No problems found. $
\end{Soutput}
\end{Schunk}
\hl{remove that last \$}. \hl{This is quite a practical example, but it essentially also gives the basic structure for misusing cleanR by generating a loop that does autocleaning... What do you think? Perhaps those who know how to make loops also know that they shouldn't do automatic data cleaning?}

\hl{
\begin{itemize}
\item  
\end{itemize}
}

\section{Extending cleanR} 
\label{sec:extending}

Though the discussion in the above paints a picture of cleanR as a user-friendly package which requires practically no knowledge of R, one should not be mistaken to think that it is not customizable. In fact, the main function of cleanR, \code{clean}, is mainly a tool for formatting the results from various checking-, summary- and visualizaion functions \hl{as described... blabla}. Thus, the actual work underlying a cleanR output file can be anything or nothing - depending on the arguments given to \code{clean}. This section consists of three parts. We commence this section with an overview of how contents are controlled in \pkg{cleanR} in terms of what cleaning steps are performed. Secondly, we turn to the possibilities of customizing what the final cleanR file looks like. Lastly, we go through an example of how to put all these customization options together in practice.

\subsubsection{Writing a summaryFunction}
Though \pkg{cleanR} provides a special class for \code{summaryFunction}s, there really is nothing special about these functions: They are nothing but regular functions with a certain input/output-structure. Specifically, they all follow the template below:
\begin{Verbatim}
mySummaryFunction <- function(v) {
	res <- [result of whatever summary we are doing]
	list(feature = "Feature name", result = res)
}
\end{Verbatim}
and we recommend furthermore adding them to the overview of all summary functions by converting them to proper \code{summaryFunction} objects:
\begin{Verbatim}
mySummaryFunction <- summaryFunction(mySummaryFunction,
	description = "Some text describing what your summaryFunction does")
\end{Verbatim}
which adds the new function to the output of \code{allSummaryFunctions()}. Note that \code{v} is a vector and that \code{res} should be either a character string or something that will be printed as one. In other words, e.g. integers are allowed, but matrices are not. Though a lot of different things can go into the \code{summaryFunction} template, we recommend only using them for summarizing the features of a variable, and leaving tests and checks for the \code{checkFunction}s (presented below).

\subsubsection{Writing a visualFunction}
\code{visualFunction}s are the functions that produce the figures of a \pkg{cleanR} output document. Writing a visual function is slightly more complicated than writing a summary function. This follows from the fact that visualFunctions need to be able to output standalone code for plots in order for \code{clean} to build standalone rmarkdown files. We recommend using the following structure:
\begin{Verbatim}
myVisualFunction <- function(v, vnam, doEval) {
	thisCall <- call("[the name of the function used to produce the plot]",
		v, [additional arguments to the plotting function])
	if (doEval) {
		return(eval(thisCall))
	} else return(deparse(thisCall)
}

myVisualFunction <- visualFunction(myVisualFunction,
	description = "Some text describing your visualFunction")
\end{Verbatim}
In this function, \code{v} is the variable to be visualized, \code{vnam} is its name (which should generally be passed to \code{title} or \code{main} arguments in plotting functions) and \code{doEval} controls whether the output is a plot (if \code{TRUE}) or a character string of standalone code for producing a plot (if \code{FALSE}). The latter \code{doEval} setting is not strictly necessary for its use in \code{clean}, but it makes it easier to assess what visualization options are available. In either case, it should be noted that all the parameters listed above, \code{v}, \code{vnam} and \code{doEval}, are mandatory, so they should be left as is (\hl{as are?}), even if you do not want to use them. As with \code{summaryFunction}s, an overview of all available \code{visualFunction}s in the environment can be obtained by calling
\begin{Verbatim}
allVisualFunctions()
\end{Verbatim}
and by calling
\begin{Verbatim}
allVisual(v, vnam, output = "html")
\end{Verbatim}
an overview of all plotting options applied on \code{v} is produced and opened as a html document for easy comparison. \hl{Should we mention the side effect of producing .rmd and .html files on disc?}

\subsubsection{Writing a checkFunction}
The last, but also most important, \pkg{cleanR} function type is the \code{checkFunction}. These are the functions that flag issues in the data and control the flow of the overall data cleaning process in the precheck stage. A \code{checkFunction} can be written using the following template:
\begin{Verbatim}
myCheckFunction <- function(v) {
	[do your check]
	problem <- [is there a problem? TRUE/FALSE]
	problemValues <- [vector of values in v that are problematic]
	problemStatus <- list(problem = problem, problemValues = problemValues)

	problemMessage <- "[The message that should be printed prior to listing
			problem values in the cleanR output]"

	outMessage <- messageGenerator(list(problem = problem,
		problemValues = problemValues, message = problemMessage))

	list(problem = problem, message = message) #problem is TRUE/FALSE,
		# message is a text string
}
myCheckFunction <- checkFunction(myCheckFunction,
	description = "[A description of your checkFunction]")
\end{Verbatim}
Only the input parameter (\code{v}) and the output format strictly has to follow this structure. However, we recommend using \code{messageGenerator} for consistent styling of all \code{checkFunction} messages. This function simply pastes together the \code{problemMessage} and the \code{problemValues}, with the latter being quoted and sorted alphabetically. Note that printing quotes in rmarkdown requires an extensive amount of character escaping, so opting for \code{messageGenerator} really is the easiest solution.

While the descriptions of \code{summaryFunction}s and \code{visualFunction}s are only for internal use in the \code{allSummaryFunctions()} and \code{allVisualFunctions()} outputs, respectively, \code{checkFunction} descriptions are actually visible in the \pkg{cleanR} output document. These are the brief descriptions presented in Part 1 \hl{eh? Is this clear? Whatever we call this section in the above} in the output document. If a \code{checkFunction} does not have a description (for instance, if it is just a regular function using the \code{checkFunction} input/output-structure), the function name will be printed instead of the description.

\begin{table}
\scriptsize
\begin{tabular}{p{0.2\linewidth}p{0.2\linewidth}p{0.2\linewidth}p{0.2\linewidth}}
& \R{summaryFunction} & \R{visualFunction} & \R{checkFunction} \\
\hline
Input (required) & \R{v} - a variable vector \newline \R{...}  &  \R{v} - a variable vector \newline \R{vnam} - the variable name (as character string) \newline \R{doEval} - a logical (\R{TRUE}/\R{FALSE}) controlling the output type of the function & \R{v} - a variable vector \newline \R{nMax} - an integer (or \R{Inf}), controlling how many problematic values are printed, if relevant \newline \R{...}  \\
Input (optional) &  \R{maxDecimals} - number of decimals printed in outputted numerical values  & - &  \R{maxDecimals}  - number of decimals printed in outputted numerical values  \\
Purpose & Describe some aspect of the variable, e.g. a central value, its dispersion or missingness. & Produce a distribution plot. & Check a variable for a specific issue and, if relevant, identify the values in the variable that cause the issue. \\
Output (required) & A list with entries \newline \R{\$feature} Label for the summary value (as character string) \newline \R{\$result} The result of the summary (as character string) & A character string with \R{R} code for producing a plot. This code should be standalone, i.e. should include the data if necessary. & A list with entries \newline \R{\$problem} - a logical identifying whether an issue was found \newline \R{\$message} - character string (possibly empty) decribing the issue that was found, properly escaped and ready for use in \R{rmarkdown} \\
Output (recommended) & A \R{summaryResult} object (i.e. an attributed list with entries \R{\$feature}, \R{\$result} and \R{\$value}, the latter being the values from \R{\$result} in their original format). & \begin{tabular}{p{0.4\linewidth} | p{0.4\linewidth}} If \R{doEval} is \R{TRUE}: \newline A plot that will be opened by the graphic device in \R{R}. & If \R{doEval} is \R{FALSE}: \newline   A text string with \R{R} code, as described above. \end{tabular} & A \R{checkResult} object (an attributed list with entries \R{\$problem}, \R{\$message} and \R{\$problemValues}, the latter being either \R{NULL} or the problem causing values, as they were found in \R{v}, whichever is relevant. \\
Tools available for producing the function & \R{summaryResult()} & - & \R{messageGenerator()} \newline \R{checkResult()} 
\end{tabular}
\caption{\hl{blabla. Fix formatting! Maybe use multirow and raggedRight stuff?}}
\label{table.functionTypes}
\end{table}

\hl{Eksempel}
\begin{Schunk}
\begin{Sinput}
> characterFoo <- function(v) {
+     if (substr(substitute(v), 1, 1) == "_") {
+         out <- list(problem=TRUE, message="Note that the variable name begins with \\_")
+     } else out <- list(problem=FALSE, message="")
+     out
+ }
> class(characterFoo) <- "checkFunction"
> attr(characterFoo, "description") <- "I really hate underscores"
> #clean(testData, characterChecks=c(defaultCharacterChecks(), "characterFoo"))
> 
\end{Sinput}
\end{Schunk}

\section{Something like examples}

\subsection{Cleaning large datasets}
If the dataset becomes very large, the standard use of \R{clean()} outlined above might not be ideal. If there is a large number of variables, production of the \R{rmarkdown} document might be very slow, while an extensive amount of observations generally makes rendering of this document into pdf or html very slow. In this section, we give a few practical examples of ways to deal with large data, while wishing to still produce (potentially very long) data cleaning overview documents. Note that the interactive tools of \R{cleanR} can be used as usual or sequentially in small subsets of the large dataset. \\

\subsubsection{Dropping the figures}
Though figures give a nice overview of each variable, they are also quite heavy objects in terms of memory allocation. Therefore, it might be beneficial to not include figures in the \R{cleanR} outputs for very large datasets. This is controlled via the \R{mode} argument:
\begin{Schunk}
\begin{Sinput}
> clean(toyData, mode = c("summarize", "check")
\end{Sinput}
\end{Schunk}

\subsubsection{Economic memory use} 
Another solution, which is especially relevant to Windows users due to the unfortunate combination of memory control in this operating system and RStudio \hl{And also just R, right? There's got to be a nice reference on this..?}, is simply splitting the two steps performed by \R{clean}, namely producing the \R{rmarkdown} file and rendering it afterwards. If the \R{rmarkdown} file is very long, as it will typically be in very large datasets, having this file opened in memory waists precious memory capacities. Therefore, we advice users to instead split the two steps. This can be done in the following manner:
\begin{Schunk}
\begin{Sinput}
> clean(toyData, render = FALSE}
> render("cleanR_toyData.Rmd", quiet = FALSE)
\end{Sinput}
\end{Schunk}
This also deals with the fact that \R{cleanR} can produce \R{rmarkdown} files that supersedes the upper size limit of RStudio, which is currently \hl{find number} GBs (using RStudio version 1.0.44). \hl{Is this maybe too editor specific? A lot of people do use RStudio...}.

\subsection{Use cleanR for problem flagging}
If the data is large, but memory issues and computation time are less of an issue than the human time it takes to look through the data cleaning document, a viable solution might be to not include all information about all variables. Or even for more reasonably sized datasets, sometimes a brief overview of the most pressing issues can be useful. This can be achieved by using the \R{onlyProblematic} argument in \R{clean()}. By specifying \R{onlyProblematic = TRUE}, only variables that raise a flag in the checking steps will be summarized. But perhaps we are not even interested in obtaining general information about these variables, but only in getting a quick overview of the problems, they might have. This can be done by also controlling the \R{mode} argument:
\begin{Schunk}
\begin{Sinput}
> clean(toyData, onlyProblematic = TRUE, mode = c("check"))
\end{Sinput}
\end{Schunk}
Now only the checking results are printed, and only for variables where problems were identified. An even more minimal output can be generated by also leaving out the checking results - then \R{clean()} essentially just produces a list of the variable names that should be investigated further:
\begin{Schunk}
\begin{Sinput}
> clean(toyData, onlyProblematic = TRUE, mode = NULL)
\end{Sinput}
\end{Schunk}
Of course, this can also be done without generating an overview document, by use of the \R{check()} function: 
\begin{Schunk}
\begin{Sinput}
> toyChecks <- check(toyData)
> foo <- function(x) {
>   any(sapply(x, function(y) y[["problem"]]))
> }
> sapply(toyChecks, foo)
\end{Sinput}
\begin{Soutput}
 var1  var2  var3  var4  var5  var6 
 TRUE  TRUE  TRUE  TRUE  TRUE FALSE 
\end{Soutput}
\end{Schunk}
\hl{drop this last bit? too technical?}

\subsection{Include cleanR document in other files}

\section{Appendix Something}
\includepdf[pages=2-4, pagecommand={}, frame=true, noautoscale=true, scale=0.7]{cleanR_toyData.pdf}

\end{document}
\R{characterChecks} & What checks should be performed on character variables? & \hl{what to write here?} \\
\R{factorChecks} & What checks should be performed on character variables? & \hl{what to write here?} \\
\R{numericChecks} & What checks should be performed on factor variables? & \hl{what to write here?} \\
\R{integerChecks} & What checks should be performed on integer variables? & \hl{what to write here?} \\
\R{logicalChecks} & What checks should be performed on logical variables? & \hl{what to write here?} \\ 
\R{dateChecks} & What checks should be performed on Date variables? & \hl{what to write here?} \\
\R{allChecks} & What checks should be performed on all variables (overwriting variable type specific settings if not \R{NULL})? & \R{NULL} \\

### 2016-12-15 latex/article_vol2.tex

@@
 By default, a pdf overview document is produced, saved to the disc and opened for inspection.
-Turning to Figure~\ref{fig:example1}, we see that such a data cleaning output document consists of two parts.
+Turning to Figure~\ref{fig:example1}, we see that a data cleaning output document consists of two parts.
 First, an overview of what was done is presented under the title \textit{Data cleaning summary}.
@@
 \end{figure}
+
+Arguments
 
@@
 
-\begin{table}[h]
+\begin{table}[ht]
 \small
@@
 \subsection{Cleaning large datasets}
-If the dataset becomes very large, the standard use of \R{clean()} outlined above might not be ideal.
-If there is a large number of variables, production of the \R{rmarkdown} document might be very slow, while an extensive amount of observations generally makes rendering of this document into pdf or html very slow.
-In this section, we give a few practical examples of ways to deal with large data, while wishing to still produce (potentially very long) data cleaning overview documents.
-Note that the interactive tools of \R{cleanR} can be used as usual or sequentially in small subsets of the large dataset.
-\\
-
-\subsubsection{Dropping the figures}
-Though figures give a nice overview of each variable, they are also quite heavy objects in terms of memory allocation.
-Therefore, it might be beneficial to not include figures in the \R{cleanR} outputs for very large datasets.
-This is controlled via the \R{mode} argument:
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, mode = c("summarize", "check")
-\end{Sinput}
-\end{Schunk}
-
-\subsubsection{Economic memory use} 
-Another solution, which is especially relevant to Windows users due to the unfortunate combination of memory control in this operating system and RStudio \hl{And also just R, right?
-There's got to be a nice reference on this..?}, is simply splitting the two steps performed by \R{clean}, namely producing the \R{rmarkdown} file and rendering it afterwards.
-If the \R{rmarkdown} file is very long, as it will typically be in very large datasets, having this file opened in memory waists precious memory capacities.
-Therefore, we advice users to instead split the two steps.
-This can be done in the following manner:
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, render = FALSE}
-> render("cleanR_toyData.Rmd", quiet = FALSE)
-\end{Sinput}
-\end{Schunk}
-This also deals with the fact that \R{cleanR} can produce \R{rmarkdown} files that supersedes the upper size limit of RStudio, which is currently \hl{find number} GBs (using RStudio version 1.0.44).
-\hl{Is this maybe too editor specific?
-A lot of people do use RStudio...}.
 
 \subsection{Use cleanR for problem flagging}
-If the data is large, but memory issues and computation time are less of an issue than the human time it takes to look through the data cleaning document, a viable solution might be to not include all information about all variables.
-Or even for more reasonably sized datasets, sometimes a brief overview of the most pressing issues can be useful.
-This can be achieved by using the \R{onlyProblematic} argument in \R{clean()}.
-By specifying \R{onlyProblematic = TRUE}, only variables that raise a flag in the checking steps will be summarized.
-But perhaps we are not even interested in obtaining general information about these variables, but only in getting a quick overview of the problems, they might have.
-This can be done by also controlling the \R{mode} argument:
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, onlyProblematic = TRUE, mode = c("check"))
-\end{Sinput}
-\end{Schunk}
-Now only the checking results are printed, and only for variables where problems were identified.
-An even more minimal output can be generated by also leaving out the checking results - then \R{clean()} essentially just produces a list of the variable names that should be investigated further:
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, onlyProblematic = TRUE, mode = NULL)
-\end{Sinput}
-\end{Schunk}
-Of course, this can also be done without generating an overview document, by use of the \R{check()} function: 
-\begin{Schunk}
-\begin{Sinput}
-> toyChecks <- check(toyData)
-> foo <- function(x) {
->   any(sapply(x, function(y) y[["problem"]]))
-> }
-> sapply(toyChecks, foo)
-\end{Sinput}
-\begin{Soutput}
- var1  var2  var3  var4  var5  var6 
- TRUE  TRUE  TRUE  TRUE  TRUE FALSE 
-\end{Soutput}
-\end{Schunk}
-\hl{drop this last bit? too technical?}
 
@@
 \section{Appendix Something}
-\includepdf[pages=2-4, pagecommand={}, frame=true, noautoscale=true, scale=0.7]{cleanR_toyData.pdf}
+\includepdf[pages=-, pagecommand={}, frame=true, noautoscale=true, scale=0.7]{cleanR_toyData.pdf}
 

### 2016-12-15 latex/article_vol2.tex
More structure, finished examples.

--- 2016-12-14 latex/article_vol2.tex
@@
 
+\begin{figure}
+\begin{center}
+\frame{\includegraphics[width=7.5cm,page=2]{cleanR_toyData.pdf}} 
+\frame{\includegraphics[width=7.5cm, page=3]{cleanR_toyData.pdf}}
+%\includepdf[pages={2}, pagecommand={}]{cleanR_testData.pdf}
+\end{center}
+\label{fig:example1}
+\caption{Example output from running \R{clean()} on the \R{toyData} dataset. First a summary of the full dataset,.. 
+  \hl{Maybe fix these pdfs such that they are one file with no blankspace? I.e. basically the .html-version of this file printed as pdf. This version is too small.}}
+\end{figure}
+
 In \pkg{cleanR}, the \code{clean} function is the primary workhorse and knowledge of nothing more than this function is required to produce data cleaning outputs.
@@
 
-By default, a pdf overview document is produced, saved to the disc and opened for inspection.
+By default, a pdf overview document is produced, saved to the disc (in the working directory) and opened for inspection.
 Turning to Figure~\ref{fig:example1}, we see that such a data cleaning output document consists of two parts.
@@
 Only these exact two checks were performed}.
-
-Of course, 
 \\
 
-\begin{figure}[!htp]
-\begin{center}
-\includegraphics[width=8cm,page=2]{cleanR_toyData.pdf} 
-\includegraphics[width=8cm, page=3]{cleanR_toyData.pdf} 
-%\includepdf[pages={2}, pagecommand={}]{cleanR_testData.pdf}
+\hl{metatext here: blabla more can be done blabla, in this section we present...}
 
-\end{center}
-\label{fig:example1}
-\caption{Example output from running XXXX on the testData
-  dataset. First a summary of the full dataset,.. 
-  \hl{Maybe fix these pdfs such that they are one file with no blankspace? I.e. basically we want the .html-version of this file printed as pdf.}}
-\end{figure}
+\subsection{Controlling [something]}
 
-\subsection{Controlling contents}
 \begin{table}
@@
 \end{tabular}
-\caption{A selection of commonly used arguments of \R{clean()}.  \hl{I have everything here except \R{file} and \R{quiet}, replace, vol, standAlone + stuff related to selecting SVC-functions}}
+\caption{A selection of commonly used arguments of \R{clean()}.
+\hl{Note: I have everything here except \R{file} and \R{quiet}, replace, vol, standAlone + stuff related to selecting SVC-functions (dealt with in next subsection)}.
+\hl{Fix formatting for \R{onlyProblematic}.}}
 \label{table.cleanFormals}
@@
 
-By the \textit{contents} of the cleanR output, we refer to every step that actually involves a function being called on variables from the dataset.
-There are three stages in which this occurs \hl{(with ref.
-to figure/flowchart?)}:
-\begin{enumerate}
-\item In the precheck functions
-\item In the summarize/visualize/check (SVC) step
-\item In the multivariate visuzalizations \hl{Or whatever, in the stuff that we have not yet implemented. I'm pretending like this stage doesn't exist below.}
-\end{enumerate}
-Each of these stages are controllable using appropriate function parameters in \code{clean}.
-In the above, we presented the default \pkg{cleanR} settings and how to tweak them into providing a slightly different data cleaning outputs.
-However, if for instance the dataset at hand requires completely different visualizations, more control is needed.
-\pkg{cleanR} uses three different types of functions for performing all stages in the above, namely \code{summaryFunction}s, \code{visualFunction}s and \code{checkFunction}s.
-They each have distinct input-output structures and most instances of the functions are build as \code{S3} generics with methods for different data classes.
-By understanding how to construct proper \code{summaryFunction}s, \code{visualFunction}s and \code{checkFunction}s, the entire contents of the \pkg{cleanR} output is at your hands.
-Therefore, this section is dedicated to introducing each in turn.\\
-
-\hl{Something like a figure that gives an overview of all summaryFunctions, ...
-available in standard cleanR, including default settings and some kind of description of where they are called (in precheck or SVC step).
-Also something about what datatypes they can each be called on.
-Also, more text here about the relationship between the two (three) stages and the three function types.}
+\hl{Describe arguments in the table...}
 
@@
 
-\begin{table}[h]
-\small
+\begin{table}
 \centering
@@
 }
+
+By the \textit{contents} of the cleanR output, we refer to every step that actually involves a function being called on variables from the dataset.
+There are two stages in which this occurs \hl{(with ref.
+to figure/flowchart?)}:
+\begin{enumerate}
+\item In the precheck functions
+\item In the summarize/visualize/check (SVC) step
+\end{enumerate}
+Each of these stages are controllable using appropriate function parameters in \code{clean}.
+In the above, we presented the default \pkg{cleanR} settings and how to tweak them into providing a slightly different data cleaning outputs.
+However, if for instance the dataset at hand requires completely different visualizations, more control is needed.
+\pkg{cleanR} uses three different types of functions for performing all stages in the above, namely \code{summaryFunction}s, \code{visualFunction}s and \code{checkFunction}s.
+They each have distinct input-output structures and most instances of the functions are build as \code{S3} generics with methods for different data classes.
+By understanding how to construct proper \code{summaryFunction}s, \code{visualFunction}s and \code{checkFunction}s, the entire contents of the \pkg{cleanR} output is at your hands.
+Therefore, this section is dedicated to introducing each in turn.\\
 
@@
 \begin{itemize}
-\item  
+\item  Mention allCheckFunctions() etc. again here
+\item Mention check(), visualize() and summarize() modes for data.frames
 \end{itemize}
@@
 
-Though the discussion in the above paints a picture of cleanR as a user-friendly package which requires practically no knowledge of R, one should not be mistaken to think that it is not customizable.
-In fact, the main function of cleanR, \code{clean}, is mainly a tool for formatting the results from various checking-, summary- and visualizaion functions \hl{as described...
-blabla}.
-Thus, the actual work underlying a cleanR output file can be anything or nothing - depending on the arguments given to \code{clean}.
-This section consists of three parts.
-We commence this section with an overview of how contents are controlled in \pkg{cleanR} in terms of what cleaning steps are performed.
-Secondly, we turn to the possibilities of customizing what the final cleanR file looks like.
-Lastly, we go through an example of how to put all these customization options together in practice.
-
-\subsubsection{Writing a summaryFunction}
-Though \pkg{cleanR} provides a special class for \code{summaryFunction}s, there really is nothing special about these functions: They are nothing but regular functions with a certain input/output-structure.
-Specifically, they all follow the template below:
-\begin{Verbatim}
-mySummaryFunction <- function(v) {
-	res <- [result of whatever summary we are doing]
-	list(feature = "Feature name", result = res)
-}
-\end{Verbatim}
-and we recommend furthermore adding them to the overview of all summary functions by converting them to proper \code{summaryFunction} objects:
-\begin{Verbatim}
-mySummaryFunction <- summaryFunction(mySummaryFunction,
-	description = "Some text describing what your summaryFunction does")
-\end{Verbatim}
-which adds the new function to the output of \code{allSummaryFunctions()}.
-Note that \code{v} is a vector and that \code{res} should be either a character string or something that will be printed as one.
-In other words, e.g.
-integers are allowed, but matrices are not.
-Though a lot of different things can go into the \code{summaryFunction} template, we recommend only using them for summarizing the features of a variable, and leaving tests and checks for the \code{checkFunction}s (presented below).
-
-\subsubsection{Writing a visualFunction}
-\code{visualFunction}s are the functions that produce the figures of a \pkg{cleanR} output document.
-Writing a visual function is slightly more complicated than writing a summary function.
-This follows from the fact that visualFunctions need to be able to output standalone code for plots in order for \code{clean} to build standalone rmarkdown files.
-We recommend using the following structure:
-\begin{Verbatim}
-myVisualFunction <- function(v, vnam, doEval) {
-	thisCall <- call("[the name of the function used to produce the plot]",
-		v, [additional arguments to the plotting function])
-	if (doEval) {
-		return(eval(thisCall))
-	} else return(deparse(thisCall)
-}
-
-myVisualFunction <- visualFunction(myVisualFunction,
-	description = "Some text describing your visualFunction")
-\end{Verbatim}
-In this function, \code{v} is the variable to be visualized, \code{vnam} is its name (which should generally be passed to \code{title} or \code{main} arguments in plotting functions) and \code{doEval} controls whether the output is a plot (if \code{TRUE}) or a character string of standalone code for producing a plot (if \code{FALSE}).
-The latter \code{doEval} setting is not strictly necessary for its use in \code{clean}, but it makes it easier to assess what visualization options are available.
-In either case, it should be noted that all the parameters listed above, \code{v}, \code{vnam} and \code{doEval}, are mandatory, so they should be left as is (\hl{as are?}), even if you do not want to use them.
-As with \code{summaryFunction}s, an overview of all available \code{visualFunction}s in the environment can be obtained by calling
-\begin{Verbatim}
-allVisualFunctions()
-\end{Verbatim}
-and by calling
-\begin{Verbatim}
-allVisual(v, vnam, output = "html")
-\end{Verbatim}
-an overview of all plotting options applied on \code{v} is produced and opened as a html document for easy comparison. \hl{Should we mention the side effect of producing .rmd and .html files on disc?}
-
-\subsubsection{Writing a checkFunction}
-The last, but also most important, \pkg{cleanR} function type is the \code{checkFunction}.
-These are the functions that flag issues in the data and control the flow of the overall data cleaning process in the precheck stage.
-A \code{checkFunction} can be written using the following template:
-\begin{Verbatim}
-myCheckFunction <- function(v) {
-	[do your check]
-	problem <- [is there a problem? TRUE/FALSE]
-	problemValues <- [vector of values in v that are problematic]
-	problemStatus <- list(problem = problem, problemValues = problemValues)
-
-	problemMessage <- "[The message that should be printed prior to listing
-			problem values in the cleanR output]"
-
-	outMessage <- messageGenerator(list(problem = problem,
-		problemValues = problemValues, message = problemMessage))
-
-	list(problem = problem, message = message) #problem is TRUE/FALSE,
-		# message is a text string
-}
-myCheckFunction <- checkFunction(myCheckFunction,
-	description = "[A description of your checkFunction]")
-\end{Verbatim}
-Only the input parameter (\code{v}) and the output format strictly has to follow this structure.
-However, we recommend using \code{messageGenerator} for consistent styling of all \code{checkFunction} messages.
-This function simply pastes together the \code{problemMessage} and the \code{problemValues}, with the latter being quoted and sorted alphabetically.
-Note that printing quotes in rmarkdown requires an extensive amount of character escaping, so opting for \code{messageGenerator} really is the easiest solution.
-
-While the descriptions of \code{summaryFunction}s and \code{visualFunction}s are only for internal use in the \code{allSummaryFunctions()} and \code{allVisualFunctions()} outputs, respectively, \code{checkFunction} descriptions are actually visible in the \pkg{cleanR} output document.
-These are the brief descriptions presented in Part 1 \hl{eh?
-Is this clear?
-Whatever we call this section in the above} in the output document.
-If a \code{checkFunction} does not have a description (for instance, if it is just a regular function using the \code{checkFunction} input/output-structure), the function name will be printed instead of the description.
-
-\begin{table}
-\scriptsize
+\begin{table}[ht]
+\small
 \begin{tabular}{p{0.2\linewidth}p{0.2\linewidth}p{0.2\linewidth}p{0.2\linewidth}}
@@
 
-\hl{Eksempel}
-\begin{Schunk}
-\begin{Sinput}
-> characterFoo <- function(v) {
-+     if (substr(substitute(v), 1, 1) == "_") {
-+         out <- list(problem=TRUE, message="Note that the variable name begins with \\_")
-+     } else out <- list(problem=FALSE, message="")
-+     out
-+ }
-> class(characterFoo) <- "checkFunction"
-> attr(characterFoo, "description") <- "I really hate underscores"
-> #clean(testData, characterChecks=c(defaultCharacterChecks(), "characterFoo"))
-> 
-\end{Sinput}
-\end{Schunk}
+Though the discussion in the above paints a picture of cleanR as a user-friendly package which requires practically no knowledge of R, one should not be mistaken to think that it is not customizable.
+In fact, the main function of cleanR, \code{clean}, is mainly a tool for formatting the results from various checking-, summary- and visualization functions \hl{as described...
+blabla}.
+Thus, the actual work underlying a cleanR output file can be anything or nothing - depending on the arguments given to \code{clean}.
+\hl{something about interactive mode.
+}
+
+This section consists of three parts.
+First, we describe how the \R{clean} function can be extended by adding custom prechecks, summaries, checks and visualizations.
+In order to do this, one needs to produce small functions that obey to the syntax of \R{cleanR}, and therefore, the minimal requirements of this syntax are introduced.
+However, if one wishes to also use these custom functions in the interactive \R{cleanR} mode and e.g.
+have them appear in a \R{allSummaryFunctions()} call, have nice printing functions etc., one needs a few extra steps when producing the functions.
+These steps are presented in the second part of this section.
+Lastly, we turn to an example of how either type of custom cleaning functions can be produced, used in \R{clean()} and used interactively.
+
+Throughout this section, Table~\ref{table.functionTypes} serves as an overview of the requirements and contents of each of the three function types used in \R{cleanR}, namely \R{summaryFunction}s (used for creating summaries), \R{visualFunction}s (used for creating plots) and \R{checkFunction}s (used for performing variable checks).
+
+\subsection{Writing cleanR extensions: The minimal approach}
+
+\subsection{Getting the full cleanR workhorse on your side}
+
+\subsection{A worked example}
 
 \section{Something like examples}
+Finally, we will present a few examples of how to make \R{cleanR} solve certain issues related to data cleaning.
+First, we discuss the challenges related to cleaning large datasets, particularly in terms of memory use and computation speed.
+Next, we show how \R{cleanR} can be used for problem-flagging.
+Lastly, we discuss how the \R{cleanR} output document can be included in other \R{rmarkdown} documents as a mean to produce clear and concise documentation of a dataset.
 
@@
 \end{Schunk}
-Of course, this can also be done without generating an overview document, by use of the \R{check()} function: 
+Of course, this can also be done without generating an overview document, by use of the \R{check()} function.
+When called on a \R{data.frame}, this function produces a list (of variables) of lists (of checks) of lists (or rather, \R{checkResult}s).
+Thus, the overall problem status of each variable can easily be unravelled using the list manipulation function \R{sapply()}:
 \begin{Schunk}
@@
 \end{Schunk}
-\hl{drop this last bit? too technical?}
+and we find that only the final variable, \R{var6}, for which all observations have the value \R{"Irrelevant"}, is problem-free. \hl{drop this last bit? too technical?}
 
 \subsection{Include cleanR document in other files}
+Sometimes, a \R{cleanR} document might be a useful addition to a more general overview document, including also e.g.
+pairwise association plots, time series plots or exploratory analysis results.
+To this end, it is possible to produce a \R{cleanR} document that can readily be included in other \R{rmarkdown} files.
+This is done by using the \R{standAlone} argument in \R{clean}, which removes the preamble from the outputted \R{rmarkdown} file.
+Please note, that it is still necessary to indicate which \R{rmarkdown} type is being created; the pdf and html \R{rmarkdown} styles are unfortunately not identical.
+
+If it is important that the embedded \R{cleanR} document can be rendered to either of these two file types, we recommend setting \R{twoCols = FALSE} and \R{output = html} in \R{clean()}, thereby essentially removing almost all output type specific formatting code from the generated \R{rmarkdown} file.
+
+On the other hand, if a pdf document is to be produced, the a few extra lines need to be added to the preamble of the master \R{rmarkdown} document - otherwise, the two-column layout code will produce an error.
+The following is an example of how such a master document preamble might look like and how the \R{cleanR\_toyData.Rmd} can be included:
+\begin{Verbatim}
+---
+output: pdf_document
+documentclass: report
+header-includes:
+  - \renewcommand{\chaptername}{Part}
+  - \newcommand{\fullline}{\noindent\makebox[\linewidth]{\rule{\textwidth}{0.4pt}}}
+  - \newcommand{\bminione}{\begin{minipage}{0.75 \textwidth}}
+  - \newcommand{\bminitwo}{\begin{minipage}{0.25 \textwidth}}
+  - \newcommand{\emini}{\end{minipage}}
+---
+
+```{r, child = 'cleanR_toyData.Rmd'}
+```
+\end{Verbatim}
+\hl{Use proper formatting here.}
+
+In the this example, the \R{cleanR\_toyData.Rmd} file could have been created as follows:
+\begin{Schunk}
+\begin{Sinput}
+> clean(toyData, standAlone = FALSE)
+\end{Sinput}
+\end{Schunk}
+and the more minimal, html-style \R{rmarkdown} file described above can be produced using
+\begin{Schunk}
+\begin{Sinput}
+> clean(toyData, standAlone = FALSE, output = "html", twoCols = FALSE)
+\end{Sinput}
+\end{Schunk}
+
+\section{Conclusion/Concluding remarks/summary/?}
+\hl{Is this a thing in this journal? Otherwise, we might want to make some final remarks in the previous sections. Feels awkward to end with a bunch of code....}
 

### 2016-12-15 latex/article_vol2.tex

@@
 
-\subsection{Getting the full cleanR workhorse on your side}
+\subsection{Getting the full cleanR machinery on your side}
 
 \subsection{A worked example}
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.2]{mosaicPlotExample.pdf}
+\end{center}
+\label{fig:mosaicPlot}
+\caption{\hl{something, note that code for producing this plot is available in ./latex/codeForArticle.R}}
+\end{figure}
+
+Now, let's say we wish to perform quality control on a slightly more complicated dataset than \R{toyData}.
+\hl{something about which dataset}.
+We will build four new functions and show both how they can be used interactively and how they can be integrated with the \R{clean()} function.
+These four new functions are:
+\begin{description}
+\item[\R{isID}] A new \R{checkFunction} intended for use in the precheck-step.
+This function checks whether a variable consists exclusively long ($> 10$) entries that are all of equal length, as this might be personal identification codes that we do not wish to print out in the data summary.
+\item[\R{mosaicVisual}] A new \R{visualFunction} that produces so-called mosaic plots. This function will be used in the \textit{visualize} step of \R{clean()}.
+\item[\R{countZeros}] A new \R{summaryFunction} that counts the number of occurrences of the value \R{0} in a variable. This function will be used in the \textit{summarize} step of \R{clean()}.
+\item[\R{identifyColon}] A new \R{checkFunction} that flags variables that colons that appear before and after alphanumerical characters, e.g.
+practical for identifying autogenerated interaction effects.
+This function will be used in the \textit{check} step of \R{clean()}.
+\end{description}
+First, let's define the \R{isID} function.
+We do not particularly wish to use this function interactively, so we will stick to the minimal requirements of a \R{checkFunction} used in \R{check()}.
+The function might then be defined by
+\begin{Schunk}
+\begin{Sinput}
+isID <- function(v, nMax = NULL, ...) {
+	out <- list(problem = FALSE, message = "")
+	if (class(v) %in% c("character", "factor", "labelled", "numeric", "integer")) {
+		v <- as.character(v)
+		lengths <- c(nchar(v))
+		if (all(lengths > 10) & length(unique(lengths)) == 1) {
+			out$problem <- TRUE
+			out$message <- "Warning: This variable seems to contain ID codes."
+		}
+	}
+	out
+}
+\end{Sinput}
+\end{Schunk} 
+This is essentially all we need to do in order to include this function as precheck-function in \R{clean()}, so we will leave this function as is and move on to the next function, namely \R{mosaicVisual}.
+We will define this function such that it gets the full \R{cleanR} functionality.
+This can be done using the code
+\begin{Schunk}
+\begin{Sinput}
+ mosaicVisual <- function(v, vnam, doEval) {
+   thisCall <- call("mosaicplot", table(v), main = vnam, xlab = "")
+   if (doEval) {
+    return(eval(thisCall))
+   } else return(deparse(thisCall))
+ }
+\end{Sinput}
+\end{Schunk}
+This function can now be called directly or used in \R{clean()}.
+We will return to its usage in \R{clean()} below.
+Depending on the argument \R{doEval}, either a text string with code or a plot is produced.
+The plot resulting from the following call is found in Figure~\ref{fig:mosaicPlot}:
+\begin{Schunk}
+\begin{Sinput}
+mosaicVisual(toyData$var1, "variable 1", doEval = TRUE)   $
+\end{Sinput}
+\end{Schunk}
+\hl{remove \$.}
+Even though \R{mosaicVisual}, as written above, follows the style of a \R{visualFunction}, it is not yet truly one and therefore, it will not appear in a \R{allVisualFunctions()} call.
+In order to get this functionality, we need to change its object class.
+This can be done by writing
+\begin{Schunk}
+\begin{Sinput}
+ mosaicVisual <- visualFunction(mosaicVisual, 
+ 				description = "Mosaic plots using graphics",
+                 classes = allClasses())
+\end{Sinput}
+\end{Schunk}
+Here, we use the function \R{allClasses()} to quickly obtain a vector of all the seven variable classes addressed in \R{cleanR}.
+Note that if R{mosaicVisual} has been an S3 generic function, this argument could have been left as \R{NULL} and then the classes for which methods are available would be added automatically.
+
+As \R{mosaicVisual} is now a full blood \R{visualFunction}, it will also be included in the \R{allVisualFunctions()} output table:
+\begin{Schunk}
+\begin{Sinput}
+> allVisualFunctions()
+\end{Sinput}
+\begin{Soutput}
+------------------------------------------------------------------------
+name           description                   classes                    
+-------------- ----------------------------- ---------------------------
+mosaicVisual   Mosaic plots using graphics   character, Date, factor,   
+                                             integer, labelled, logical,
+                                             numeric
+
+basicVisual    Histograms and barplots using                            
+               graphics
+
+standardVisual Histograms and barplots using                            
+               ggplot2                                                  
+------------------------------------------------------------------------
+\end{Soutput}
+\end{Schunk}
+\hl{Redo this output table when makeXFunction issues are fixed.}
+
+Now, we are done with the definition of \R{mosaicVisual} and we can turn to the next function in line, \R{countZeros}. This \R{summaryFunction} in spe is defined in the following lines of code:
+\begin{Schunk}
+\begin{Sinput}
+countZeros <- function(v, ...) {
+	res <- length(which(v == 0))
+	summaryResult(list(feature = "No. zeros", result = res, value = res))
+}
+\end{Sinput}
+\end{Schunk}
+Note that as this function computes an integer (the number of zeros), there is no difference between the entires \R{\$result} and \R{\$value}.
+If, on the other hand, the result had been a character string, extra formatting might be required in the \R{\$result} entry (such as escaping of quotation marks), and in this scenario, the two entries would then differ.
+As the result is returned as a \R{summaryResult} object, a printing method is automatically called when \R{countZeros} is used interactively:
+\begin{Schunk}
+\begin{Sinput}
+> countZeros(c(rep(0, 5), 1:100))
+\end{Sinput}
+\begin{Soutput}
+No. zeros: 5
+\end{Soutput}
+\end{Schunk}
+As in the above, we change the class of this function in order to make it appear in \R{allSummaryFunctions()} calls.
+But now we wish to emphasize that the function is not intended to be called on all variable types, as zeros have different roles in \R{Date}s and in \R{logical} variables:
+\begin{Schunk}
+\begin{Sinput}
+> countZeros <- summaryFunction(countZeros,
+	description = "Count number of zeros",
+	classes = c("character", "factor", "integer", 
+	  "labelled", "numeric"))
+\end{Sinput}
+\end{Schunk}
+
+The last function mentioned above is \R{identifyColon()}.
+We define it using the helper function \R{messageGenerator} to obtain a properly escaped message, and we use \R{checkResult} to make it appear in \R{allCheckFunction()} calls:
+\begin{Schunk}
+\begin{Sinput}
+identifyColon <- function(v, nMax = Inf, ... ) {
+  v <- unique(na.omit(v))
+  problemMessage <- "Note: The following values include colons:"
+  problem <- FALSE
+  problemValues <- NULL
+
+  problemValues <- v[sapply(gregexpr("[[:xdigit:]]:[[:xdigit:]]", v),
+                            function(x) all(x != -1))]
+
+  if (length(problemValues) > 0) {
+    problem <- TRUE 
+  }
+
+  problemStatus <- list(problem = problem, 
+                        problemValues = problemValues)
+  outMessage <- messageGenerator(problemStatus, problemMessage, nMax)
+
+  checkResult(list(problem = problem, 
+                   message = outMessage,
+                   problemValues = problemValues))
+}
+
+identifyColon <- checkFunction(identifyColon, 
+    description = "Identify non-suffixed nor -prefixed colons",
+    classes = c("character", "factor", "labelled"))
+
+\end{Sinput}
+\end{Schunk}
+As with the previous two functions, we also change its class.
+Note, however, that for \R{checkFunction}s, the function description will appear in the document produced by \R{clean()} (in the \textit{Data cleaning summary} section), so now this is not only done for the sake of the \R{allCheckFunctions()} output.\\
+
+Now, we are ready to use these new functions in a \R{clean()} call.
+\hl{do this...
+what dataset should we use here?
+Include output in appendix, maybe.}.
+The extended \R{cleanR} output document should have the following modifications, relative to the standard \R{cleanR} output:
+\begin{itemize}
+\item We want to add the new pre-check function, \R{isID}, to the already existing pre-checks.
+\item We wish to change the plot type for all variables to the new mosaic plot.
+\item We want the new summary function, \R{countZeros}, to be added to the summaries performed on all variable types but \R{Date} and \R{logical}.
+\item We want the new check function, \R{identifyColon}, to be added to the checks performed on \R{character}, \R{factor} and \R{labelled} variables.
+\end{itemize}
+These options are specified as follows:
+\begin{Schunk}
+\begin{Sinput}
+> clean(dataSet, 
+      preChecks = c("isKey", "isEmpty", "isID"),
+      allVisuals = "mosaicVisual",
+      characterSummaries = c(defaultCharacterSummaries(), "countZeros"),
+      factorSummaries = c(defaultFactorSummaries(), "countZeros"),
+      labelledSummaries = c(defaultLabelledSummaries(), "countZeros"),
+      numericSummaries = c(defaultNumericSummaries(), "countZeros"),
+      integerSummaries = c(defaultIntegerSummaries(), "countZeros"),
+      characterChecks = c(defaultCharacterChecks(), "identifyColon"),
+      factorChecks = c(defaultFactorChecks(), "identifyColon"),
+      labelledCheck = c(defaultLabelledChecks(), "identifyColon"))
+\end{Sinput}
+\end{Schunk}
+The outputted document is found in Appendix \hl{NUMTWO}. \hl{Comment. Make appendix. Remember to change dataSet to something else, when we decide what data to use here.}
 
@@
 \begin{Sinput}
-> clean(toyData, mode = c("summarize", "check")
+> clean(toyData, mode = c("summarize", "check"))
 \end{Sinput}
 \end{Schunk}
+\hl{blabla, end?}
 
@@
 
+\section{Appendix NUMTWO}
+\hl{Data cleaning with extension here}
+
 \end{document}

### 2016-12-15 latex/article_vol2.tex
Added function writing templates + minor structure stuff.

@@
 
-\hl{metatext here: blabla more can be done blabla, in this section we present...}
+\hl{metatext here: blabla more can be done blabla...}
+
+Though the \R{clean()} function is very easy to use, it should not be mistaken to be inflexible.
+A large number of function arguments allows for the cleaning overview document to be molded as the user sees best.
+Below, we present a selection of these arguments.
+We do this under two distinct headlines to emphasize that two levels of control are available when using \R{clean()}.
+Either, we can control the \hl{[something].
+Essentially, this is everything but SVC-functions.} \\
+
+\hl{Note that the terminology "SVC"-step (referring to the summarize/visualize/check-step) needs to be introduced in this part, as it is used in Table \ref{table.cleanFormals}.}
 
 \subsection{Controlling [something]}
+\label{subsection:controlSomething}
 
@@
   \hline
-& Description &  \multicolumn{7}{c}{Relevant variable classes} \\ \smallskip
+& Description &  \multicolumn{7}{c}{Variable classes} \\ \smallskip
  & &  C & F & I & L & B & N & D\\ 
@@
 Thus, the actual work underlying a cleanR output file can be anything or nothing - depending on the arguments given to \code{clean}.
-\hl{something about interactive mode.
+Specfically, user made functions can be added to the SVC-function arguments discussed above, e.g.
+\R{factorSummaries}, \R{allVisuals} or \R{numericChecks}: All that is needed is to specify their names, just like the names of the build-in SVC functions are specified.
+However, not just any function can be called from these three steps, and therefore, we will now present how \R{summaryFunction}s, \R{visualFunction}s and \R{checkFunction}s are made.
+\hl{Mention something about the interactive mode here as well?} \\
+
+This section consists of two parts.
+First, we describe how the \R{clean} function can be extended by adding custom prechecks, summaries, checks and visualizations.
+In order to do this, one needs to produce small functions that obey to the syntax of \R{cleanR}.
+This can be done with different levels of strictness.
+If the custom functions are only to be used by \R{clean()}, only the input/output structures of the functions need to be restricted.
+However, by adding a few extra tweaks to these functions, the full machinery of \R{cleanR} becomes available for the new, user-made functions, just as it is for the build-in functions.
+The presentation below is given in the format of pseudo-code function templates, designed for getting the full functionality, but please note that Table~\ref{table.functionTypes} serves as a reference to the minimal requirements, while also presenting the "full" versions of the function types.
+
+After this abstract, templatic overview of the internals of the three SVC functions is given, we turn to a worked example of how to use custom made functions in practice in \hl{ref to worked example section}.
+Here, four new SVC functions are defined and used, both interactively and in \R{clean()}.
+
+\subsection{Writing cleanR extensions: Function templates}
+\hl{metatext}
+
+\subsubsection{Writing a summaryFunction}
+Though \pkg{cleanR} provides a dedicated class for \code{summaryFunction}s, there really is nothing special about these functions: They are nothing but regular functions with a certain input/output-structure.
+Specifically, they all follow the template below:
+\begin{Verbatim}
+mySummaryFunction <- function(v, ...) {
+  res <- [result of whatever summary we are doing]
+  summaryResult(list(feature = "Feature name", result = res))
+}
+\end{Verbatim}
+The last function called, \R{summaryResult()}, changes the class of the output and thereby makes a \R{print()} functin available.
+Note that \code{v} is a vector and that \code{res} should be either a character string or something that will be printed as one.
+In other words, e.g.
+integers are allowed, but matrices are not.
+Though a lot of different things can go into the \code{summaryFunction} template, we recommend only using them for summarizing the features of a variable, and leaving tests and checks for the \code{checkFunction}s (presented below).
+
+Though following the template above is sufficient for using \R{mySummaryFunction} in \R{clean()}, we recommend furthermore adding it to the overview of all summary functions by converting it to a proper \code{summaryFunction} object.
+This is done by writing
+\begin{Verbatim}
+mySummaryFunction <- summaryFunction(mySummaryFunction,
+	description = "Some text describing what your summaryFunction does",
+	classes = c([the data types that this function is intended to be used for]))
+\end{Verbatim}
+which adds the new function to the output of \code{allSummaryFunctions()}.
+One comment should be devoted to the two attributes of a \R{summaryFunction}.
+If \R{description} is left unspecified, the name of the function (in this case, \R{mySummaryFunction}) will be filled in.
+What happens if the \R{classes} argument depends on the type of \R{mySummaryFunction}.
+If \R{mySummaryFunction} is a \R{S3} generic function with associated methods, the call to \R{summaryFunction()} will automatically produce a vector of the names of the classes for which the function can be called.
+If \R{mySummaryFunction} is not an \R{S3} generic and \R{classes} is left unspecified, the attribute will simply be empty.
+Note that the helper function \R{allClasses()} might be useful for filling out the \R{classes} argument, as it simply lists all available classes in \R{cleanR}:
+\begin{Schunk}
+\begin{Sinput}
+> allClasses()
+\end{Sinput}
+\begin{Soutput}
+[1] "character" "Date"      "factor"   "integer"   "labelled"  
+[6] "logical"  "numeric"  
+\end{Soutput}
+\end{Schunk}
+\hl{write something here, don't end paragraph with code? Also, maybe move the allClasses() stuff somewhere else, it doesn't really belong under this header.}
+
+\subsubsection{Writing a visualFunction}
+\code{visualFunction}s are the functions that produce the figures of a \pkg{cleanR} output document.
+Writing a \R{visualFunction} is slightly more complicated than writing a \R{summaryFunction}.
+This follows from the fact that \R{visualFunctions} need to be able to output standalone code for plots in order for \code{clean()} to build standalone \R{rmarkdown} files.
+We recommend using the following structure:
+\begin{Verbatim}
+myVisualFunction <- function(v, vnam, doEval) {
+  thisCall <- call("[the name of the function used to produce the plot]",
+    v, [additional arguments to the plotting function])
+  if (doEval) {
+    return(eval(thisCall))
+  } else return(deparse(thisCall))
 }
 
-This section consists of three parts.
-First, we describe how the \R{clean} function can be extended by adding custom prechecks, summaries, checks and visualizations.
-In order to do this, one needs to produce small functions that obey to the syntax of \R{cleanR}, and therefore, the minimal requirements of this syntax are introduced.
-However, if one wishes to also use these custom functions in the interactive \R{cleanR} mode and e.g.
-have them appear in a \R{allSummaryFunctions()} call, have nice printing functions etc., one needs a few extra steps when producing the functions.
-These steps are presented in the second part of this section.
-Lastly, we turn to an example of how either type of custom cleaning functions can be produced, used in \R{clean()} and used interactively.
+myVisualFunction <- visualFunction(myVisualFunction,
+	description = "Some text describing your visualFunction",
+	classes = c([the data types that this function is intended to be used for]))
+	)
+\end{Verbatim}
+In this function, \code{v} is the variable to be visualized, \code{vnam} is its name (which should generally be passed to \code{title} or \code{main} arguments in plotting functions) and \code{doEval} controls whether the output is a plot (if \code{TRUE}) or a character string of standalone code for producing a plot (if \code{FALSE}).
+The \code{doEval = TRUE} setting is not strictly necessary for a \R{visualFunction}'s use in \code{clean}, but it makes it easier to assess what visualization options are available, and obviously, it also allows for interactive usage of \R{myVisualFunction}.
+In either case, it should be noted that all the parameters listed above, \code{v}, \code{vnam} and \code{doEval}, are mandatory, so they should be left as is, even if they are not in use.
 
-Throughout this section, Table~\ref{table.functionTypes} serves as an overview of the requirements and contents of each of the three function types used in \R{cleanR}, namely \R{summaryFunction}s (used for creating summaries), \R{visualFunction}s (used for creating plots) and \R{checkFunction}s (used for performing variable checks).
+\subsubsection{Writing a checkFunction}
+The last, but perhaps most important, \pkg{cleanR} function type is the \code{checkFunction}.
+These are the functions that flag issues in the data in the check step and control the overall flow of the data cleaning process in the precheck stage.
+A \code{checkFunction} can follow two overall structures, depending on the type of check.
+Either, it tries to identify problematic values in the variable (as e.g.
+\R{identifyMissing()} does) or it performs a check concerning the variable as a whole (e.g.
+the functions used for prechecks and the function \R{identifyNums()}).
+We present templates for both types of \R{checkFunction}s below separately, but it should be emphasized that formally, they belong to the same class.
 
-\subsection{Writing cleanR extensions: The minimal approach}
+First, a template for the full-variable check function type:
+\begin{Schunk}
+\begin{Sinput}
+myFullVarCheckFunction <- function(v, ...) {
+  [do your check]
+  problem <- [is there a problem? TRUE/FALSE]
+  message <- "[message describing the problem, if any]"
+  checkResult(list(problem = problem,
+    message = message,
+    problemValues = NULL))
+}
 
-\subsection{Getting the full cleanR machinery on your side}
+myFullVarCheckFunction <- checkFunction(myFullVarCheckFunction, 
+  description = "Some text describing your checkFunction",
+  classes = c([the data types that this function is intended to be used for])
+  )
+\end{Sinput}
+\end{Schunk}
+Again, as with \R{summaryFunction}s and \R{visualFunction}s, the change of function class by use of \R{checkFunction()} is not strictly necessary.
+Note however, that if \R{myFullVarCheckFunction} is to be used in the SVC step in \R{clean()}, the description attribute will be printed in the overview table in \textit{Data cleaning summary}.
+
+If problematic values are to be identified, the template from above should be expanded to follow a slightly more complicated structure:
+\begin{Verbatim}
+myProbValCheckFunction <- function(v, nMax, maxDecimals, ...) {
+	[do your check]
+	problem <- [is there a problem? TRUE/FALSE]
+	problemValues <- [vector of values in v that are problematic]
+	problemStatus <- list(problem = problem, 
+	 problemValues = problemValues)
+
+	problemMessage <- "[The message that should be printed prior to listing
+			problem values in the cleanR output, ending with a colon]"
+
+	outMessage <- messageGenerator(problemStatus, problemMessage, nMax)
+
+	checkResult(list(problem = problem,
+	  message = outMessage,
+	  problemValues = problemValues)) 
+}
+
+myProbValCheckFunction <- checkFunction(myProbValCheckFunction,
+  description = "[Some text describing your checkFunction]",
+  classes = c([the data types that this function is intended to be used for])
+  )
+\end{Verbatim}
+In this template, the argument \R{maxDecimals} is not used.
+This argument should be used to round off the \R{problemValues} passed to \R{messageGenerator()}, if they are numerical.
+This is done by substituting the \R{problemStatus} assignment above with the following code:
+\begin{Schunk}
+\begin{Sinput}
+problemStatus <- list(problem = problem, 
+  problemValues = round(problemValues, maxDecimals))
+\end{Sinput}
+\end{Schunk}
+Another noteworthy component of the template is the usage of the helper function \R{messageGenerator()}, which aids consistent styling of all \code{checkFunction} messages.
+This function simply pastes together the \code{problemMessage} and the \code{problemValues}, with the latter being quoted and sorted alphabetically.
+If \R{nMax} is not \R{Inf}, only the first \R{nMax} problem values will be pasted onto the message, accompanied by a comment about how many problem values were left out.
+Note that printing quotes in rmarkdown requires an extensive amount of character escaping, so opting for \code{messageGenerator()} really is the easiest solution.
 
@@
 
-Now, let's say we wish to perform quality control on a slightly more complicated dataset than \R{toyData}.
-\hl{something about which dataset}.
-We will build four new functions and show both how they can be used interactively and how they can be integrated with the \R{clean()} function.
-These four new functions are:
+We will now build four new functions and show both how they can be used interactively and how they can be integrated with the \R{clean()} function. These four new functions are:
 \begin{description}
@@
 \item[\R{countZeros}] A new \R{summaryFunction} that counts the number of occurrences of the value \R{0} in a variable. This function will be used in the \textit{summarize} step of \R{clean()}.
-\item[\R{identifyColon}] A new \R{checkFunction} that flags variables that colons that appear before and after alphanumerical characters, e.g.
+\item[\R{identifyColons}] A new \R{checkFunction} that flags variables that colons that appear before and after alphanumerical characters, e.g.
 practical for identifying autogenerated interaction effects.
@@
 \end{description}
+These functions are defined in turn below, and afterwards, an example of how they can be called from \R{clean()} is provided.
+
+\subsubsection{isID - a new checkFunction without problem values}
 First, let's define the \R{isID} function.
-We do not particularly wish to use this function interactively, so we will stick to the minimal requirements of a \R{checkFunction} used in \R{check()}.
-The function might then be defined by
+As this function is not supposed to list problematic values in the variable, it falls within the category of \R{checkFunctions} represented by \R{myFullVarCheckFunction()} in the above.
+We do not particularly wish to use this function interactively, so we will stick to the minimal requirements of a \R{checkFunction} used in \R{check()} (see Table~\ref{table.functionTypes}).
+The function can then be defined by
 \begin{Schunk}
@@
   out <- list(problem = FALSE, message = "")
-	if (class(v) %in% c("character", "factor", "labelled", "numeric", "integer")) {
+  if (class(v) %in% setdiff(allClasses(), c("logical", "Date"))) {
     v <- as.character(v)
@@
 \end{Schunk} 
-This is essentially all we need to do in order to include this function as precheck-function in \R{clean()}, so we will leave this function as is and move on to the next function, namely \R{mosaicVisual}.
-We will define this function such that it gets the full \R{cleanR} functionality.
-This can be done using the code
+This is essentially all we need to do in order to include this function as a precheck-function in \R{clean()}, so we will leave it as is and move on to the next function, namely \R{mosaicVisual}.
+
+\subsubsection{mosiacVisual - a new visualFunction}
+We will define this function such that it gets the full \R{cleanR} functionality. This can be done using the code 
 \begin{Schunk}
@@
 
-Now, we are done with the definition of \R{mosaicVisual} and we can turn to the next function in line, \R{countZeros}. This \R{summaryFunction} in spe is defined in the following lines of code:
+Now, we are done with the definition of \R{mosaicVisual} and we can turn to the next function in line, \R{countZeros}.
+
+\subsubsection{countZeros - a new summaryFunction}
+This \R{summaryFunction} in spe is defined in the following lines of code:
 \begin{Schunk}
@@
 \end{Schunk}
+\hl{more? don't end on code.}
 
+\subsubsection{identifyColons - a new checkFunction with problem values}
 The last function mentioned above is \R{identifyColon()}.
@@
 \begin{Sinput}
-identifyColon <- function(v, nMax = Inf, ... ) {
+identifyColons <- function(v, nMax = Inf, ... ) {
   v <- unique(na.omit(v))
@@
 
-identifyColon <- checkFunction(identifyColon, 
+identifyColons <- checkFunction(identifyColons, 
     description = "Identify non-suffixed nor -prefixed colons",
@@
 
+\subsubsection{Calling the new SVC function from clean()}
 Now, we are ready to use these new functions in a \R{clean()} call.
@@
       integerSummaries = c(defaultIntegerSummaries(), "countZeros"),
-      characterChecks = c(defaultCharacterChecks(), "identifyColon"),
-      factorChecks = c(defaultFactorChecks(), "identifyColon"),
-      labelledCheck = c(defaultLabelledChecks(), "identifyColon"))
+      characterChecks = c(defaultCharacterChecks(), "identifyColons"),
+      factorChecks = c(defaultFactorChecks(), "identifyColons"),
+      labelledCheck = c(defaultLabelledChecks(), "identifyColons"))
 \end{Sinput}
@@
 Lastly, we discuss how the \R{cleanR} output document can be included in other \R{rmarkdown} documents as a mean to produce clear and concise documentation of a dataset.
+\hl{I feel like there should be more topics here, but I'm all out of ideas...}
 
@@
 
-\subsubsection{Dropping the figures}
+\subsubsection{Omitting figures}
 Though figures give a nice overview of each variable, they are also quite heavy objects in terms of memory allocation.
@@
 \hl{Is this maybe too editor specific?
-A lot of people do use RStudio...}.
+On the other hand, a lot of people do use RStudio...}.
 
@@
 \end{Schunk}
+\hl{don't end on code.}
 
 \section{Conclusion/Concluding remarks/summary/?}
-\hl{Is this a thing in this journal? Otherwise, we might want to make some final remarks in the previous sections. Feels awkward to end with a bunch of code....}
+\hl{Is this a thing in this journal? Otherwise, we might want to make some final remarks in the previous sections. Feels awkward to end with a bunch of code and some super specific examples...}
+
+\appendix
 
@@
 \section{Appendix NUMTWO}
-\hl{Data cleaning with extension here}
+\hl{Data cleaning with user supplied extensions here}
 

### 2016-12-15 latex/codeForArticle.R

@@ -101,8 +101,8 @@ identifyColon <- checkFunction(identifyColon,
                                description = "Identify non-suffixed nor -prefixed colons",
                                classes = c("character", "factor", "labelled"))
 
-
-clean(toyData, replace = T, 
+data(testData)
+clean(testData, replace = T, 
       preChecks = c("isKey", "isEmpty", "isID"),
       allVisuals = "mosaicVisual",
       characterSummaries = c(defaultCharacterSummaries(), "countZeros"),
@@ -112,7 +112,7 @@ clean(toyData, replace = T,
       integerSummaries = c(defaultIntegerSummaries(), "countZeros"),
       characterChecks = c(defaultCharacterChecks(), "identifyColon"),
       factorChecks = c(defaultFactorChecks(), "identifyColon"),
-      lablledCheck = c(defaultLabelledChecks(), "identifyColon"))
+      labelledCheck = c(defaultLabelledChecks(), "identifyColon"))
       
       
       

### 2016-12-15 latex/article_vol2.tex
Added figure to describe the structure of clean() + example mosaic plot pdf.

@@
 
+\begin{figure}
+\begin{description}
+\item[Input] A dataset
+	\begin{itemize} 
+		\item This should be of type \R{data.frame} or \R{tibble}
+	\end{itemize}
+\item[Create contents] For each variable in the dataset, do the following:
+	\begin{description}
+		\item[Stage 1:] Pre-checks 
+			\begin{itemize}
+				\item Is the variable suitable for summarization, visualization and checks?
+				\begin{description}
+					\item[Yes:] Go to stage 2 
+					\item[No:] Move on to the next variable
+				\end{description}
+			\end{itemize}
+		\item[Stage 2:] SVC-steps
+			\begin{description}
+				\item[Summarize] Call \R{summarize()} to produce a summary table describing the variable. What features enter this table depends on the data class of the variable.
+				\item[Visualize]  Call \R{visualize()} to produce a plot visualizing the distribution of the variable. 
+				\item[Check]  Call \R{check()} to apply quality- and error checks to the variable. What checks are used depends on the data class of the variable.
+			\end{description}
+	\end{description}
+\item[Output] Files for a overview document 
+	\begin{itemize}
+		\item Always a \R{rmarkdown} (.Rmd) file
+		\item Possible also a html or pdf file
+	\end{itemize}
+\end{description}
+\caption{\hl{quite odd having a figure like this? maybe do a flowchart?}}
+\label{figure:cleanStructure}
+\end{figure}
+
 In \pkg{cleanR}, the \code{clean} function is the primary workhorse and knowledge of nothing more than this function is required to produce data cleaning outputs.
@@
 
-\hl{metatext here: blabla more can be done blabla...}
+\hl{metatext here: blabla more can be done blabla...} \\
 
@@
 A large number of function arguments allows for the cleaning overview document to be molded as the user sees best.
-Below, we present a selection of these arguments.
+Below, we present a selection of these arguments in sections \hl{xx} and \hl{xxx}.
 We do this under two distinct headlines to emphasize that two levels of control are available when using \R{clean()}.
-Either, we can control the \hl{[something].
-Essentially, this is everything but SVC-functions.} \\
-
-\hl{Note that the terminology "SVC"-step (referring to the summarize/visualize/check-step) needs to be introduced in this part, as it is used in Table \ref{table.cleanFormals}.}
+Either, we can control the \hl{[something] - essentially, this is everything but SVC-functions - } or we can control what quality assessments and descriptions each variable are subjected to.
+In order to understand this distinction, a brief glimpse of the inner structure of \R{clean()} might be useful.
+This can be obtained from Figure~\ref{figure:cleanStructure}.
+\hl{segway to next bit or just something to end the paragraph}
 

### 2016-12-16 latex/article_vol2.tex
Proof-reading, minor stuff, structure.

@@
   should go through each variable in the dataset and look for
-  potential errors --- both in input values and coding.  \hl{segway}
+  potential errors --- both in input values and coding.  \hl{segway: But, ...}
 
   We describe an \proglang{R} package which implements an extensive
-  and customizeable suite of checks to be applied to the variables in
-  a dataset in order to identify potential problems in the
-  corresponding variables. The results can be presented in an auto-generated, non-technical, stand-alone
+  and customizeable suite of quality assessment tools to be applied to
+  a dataset in order to identify potential problems in its variables. The results can be presented in an auto-generated, non-technical, stand-alone
   overview document, intended to be perused by an investigator with an
@@
   The \pkg{cleanR} package is designed to be easily extended with custom
-  user-created checks that are relevant in particular situations.
+  user-created checks that are relevant in particular situations. \hl{Already said that above. Either expand on it or delete this.}
 }
@@
 \section{Introduction}
-Statisticians and data analysts spend a large portion of their time on
-data cleaning and on data wrangling. Packages such as
-\pkg{data.table}, and \pkg{plyr} have made data wrangling a lot easier in
-\proglang{R}, but there are only a few tools available for data cleaning. \hl{What are these tools? More here about the solutions already available.}
+Though data cleaning might be regarded as a somewhat dull activity, adequate data cleaning is crucial in any data analysis.
+With ever-growing dataset sizes and complexities, statisticians and data analysts find themselves spending a large portion of their time on
+data cleaning and on data wrangling. \proglang{R}-packages such as
+\pkg{data.table}, and \pkg{plyr} have made data wrangling a lot easier, but there are only a few tools available for data cleaning.
+\hl{What are these tools?
+More here about the solutions already available.}
 
@@
 \item Maybe hints to the discussion later about poor documentation of data cleaning/quality assessment as the status quo for most analysts
-\item Gives a specification of what we mean by data errors/mistakes/whatever. I'm not completely sure this will be clear to everyone.
+\item Gives a specification of what we mean by data errors/mistakes/whatever.
+I'm not completely sure this will be clear to everyone and we should probably mention it early on to not lose readers/users.
 \end{itemize}
@@
 data analysis (e.g., \hl{a categorical variable with numeric labels is
-wrongly classified as a numeric variable\footnote{Is that something we can fix in clean()?}} or a variable where all values have erroneously been coded to the same value), but in other cases a human
+wrongly classified as a numeric variable\footnote{\hl{Is that something we can fix in clean()?
+Not quite sure what "labels" mean in this context.}}} or a variable where all values have erroneously been coded to the same value), but in other cases a human
 with knowledge about the data context area is needed to identify
@@
 
-The \R{cleanR} approach to data cleaning and -quality assessment is characterized by two principles.
+The \R{cleanR} approach to data cleaning and -quality assessment is characterized by two principles \hl{use a different term than "principles" here and below, when we refer to this sentence}.
 First of all, there is no need for data cleaning to be an ad hoc procedure.
@@
 This means that data cleaning can easily be a well-documented, well-specified procedure.
-This angle on data cleaning also adds to the transparency of this first step of data analysis: Though data cleaning might be regarded as a somewhat dull activity, adequate data cleaning is crucial in any data analysis.
 In order to aid this principle, \R{cleanR} provides easy-to-use, automated tools for data quality assessment in \R{R} on which data cleaning decisions can be build.
@@
 Oftentimes, e.g.
-separating faulty codings of a numeric value from unusual, but correct, values requires problem-specific expertise that might not be held by the data analyst.
+distinguishing between faulty codings of a numeric value and unusual, but correct, values requires problem-specific expertise that might not be held by the data analyst.
 Hopefully, having easy access to data descriptions through \R{cleanR} will help this necessary knowledge sharing.
 
-While \R{cleanR} is build for auto-generating data quality assessment overview documents, we also wish to emphasize that it is \textit{not} a tool for automatic data cleaning.
+While \R{cleanR}s primary raison d'être is auto-generating data quality assessment overview documents, we still wish to emphasize that it is \textit{not} a tool for automatic data cleaning.
 This qualifies as the second principle of \R{cleanR}: Data cleaning decisions should always be made by humans.
-Therefore, \R{cleanR} does not provide any tools for "fixing" errors in the data.
-However, we do provide interactive functions that can be used to identify potentially erroneous entries in a dataset and that can make it easier to solve data issues.\\
+Therefore, \R{cleanR} does not supply any tools for "fixing" errors in the data.
+However, we do provide interactive functions that can be used to identify potentially erroneous entries in a dataset and that can make it easier to solve data issues, one variable at a time.\\
 
-This manuscript is structured as follows: First, we introduce the representative of the first principle, namely the \R{clean()} function, which generates data cleaning overview documents.
+This manuscript is structured as follows: First, in section \hl{2}, we introduce the representative of the first \hl{principle}, namely the \R{clean()} function, which generates data cleaning overview documents.
 In the \R{cleanR} package, we have provided a number of default cleaning steps that cover the data cleaning challenges, we find to be most common.
-However, every dataset is different, and some datasets might include problems that cannot be detected by our data checking functions.
-In Section 3 \hl{proper reference here}, we turn to the question of how such \R{cleanR} extensions can be made such that they are integrable with the \R{clean()} function and with other tools available in \R{cleanR}.
-In this section, we also present the interactive mode of \R{cleanR}, as motivated by the second principle above.
-At last, in Section 4 \hl{proper ref}, we discuss a number of examples of specific data cleaning challenges and how \R{cleanR} can be used to solve them.
+Next, in section \hl{3}, we present the interactive mode of \R{cleanR}, as motivated by the second \hl{principle} above.
+But, as any data analyst knows, every dataset is different, and some datasets might include problems that cannot be detected by our data checking functions.
+In Section \hl{4}, we turn to the question of how such \R{cleanR} extensions can be made, such that they are integrable with the \R{clean()} function and with the other tools available in \R{cleanR}.
+At last, in Section 5 \hl{proper ref}, we discuss a number of examples of specific data cleaning challenges and how \R{cleanR} can be used to solve them.
+\hl{Is there a Section 6?}\\
+
+\hl{Do they typically discuss notation in articles like this?
+I try (somewhat inconsistently) to refer to functions as \R{function()} while other \R{R} objects are referred to merely as \R{object}.
+This is also the style adopted by Wickham in his books.}
 
@@
 \caption{Example output from running \R{clean()} on the \R{toyData} dataset. First a summary of the full dataset,.. 
-  \hl{Maybe fix these pdfs such that they are one file with no blankspace? I.e. basically the .html-version of this file printed as pdf. This version is too small.}}
+  \hl{Maybe fix these pdfs such that they are one file with no blankspace? I.e. basically the .html-version of this file printed as pdf. This version is too small, right?}}
 \end{figure}
@@
 	\begin{itemize} 
-		\item This should be of type \R{data.frame} or \R{tibble}
+		\item This should be of type \R{data.frame} or \R{tibble} \hl{or data.table?}
 	\end{itemize}
@@
 				\begin{description}
-					\item[Yes:] Go to stage 2 
-					\item[No:] Move on to the next variable
+					\item[Yes:] Go to stage 2. 
+					\item[No:] Move on to the next variable.
 				\end{description}
@@
 	\end{description}
-\item[Output] Files for a overview document 
+\item[Output] Files for a overview document are saved to the disc and possibly also opened.
 	\begin{itemize}
@@
 
-In \pkg{cleanR}, the \code{clean} function is the primary workhorse and knowledge of nothing more than this function is required to produce data cleaning outputs.
+In \pkg{cleanR}, the function \R{clean()} is the primary workhorse and knowledge of nothing more than this tool is required to produce data cleaning outputs.
 The data cleaning output itself is an overview document, intended for reading by humans, in either pdf or html format.
@@
 We can then return to Part 1, the data cleaning summary, and inspect what sorts of checks were performed on this variable.
-\hl{hm, delete the last part?
+\hl{hm, delete the last sentence?
 or use different variable?
@@
 \\
-
-\hl{metatext here: blabla more can be done blabla...} \\
 
@@
 \small
-\begin{tabular}{p{0.2\linewidth}p{0.5\linewidth}p{0.2\linewidth}}
+\begin{tabular}{p{0.25\linewidth}p{0.45\linewidth}p{0.2\linewidth}}
 \hline
@@
 \quad \R{ordering} & Ordering of the variables in the data summary (as is or alphabetical) & \R{"asIs"} \\
-\quad \R{onlyProblematic} & Should only variables flagged as problematic be included in the variable list? & \R{FALSE} \\
+\quad \R{onlyProblematic} & Should only variables flagged as problematic be included in the \textit{Variable list}? & \R{FALSE} \\
 \quad \R{listChecks} & Should an overview of what checks were performed by listed in the \textit{Data cleaning summary}? &  \R{TRUE} \\
@@
 \smallskip Control output \\
-\quad \R{output} & Type of output file to be produced (html, pdf or \hl{?? what does screen mean here??}) & \R{"pdf"} \\
+\quad \R{output} & Type of output file to be produced (html, pdf or \hl{?? what does screen mean in the documentation??}) & \R{"pdf"} \\
 \quad \R{render} & Should the output file be rendered from markdown? & \R{TRUE} \\
-\quad \R{openResult} & If a  pdf/html file is rendered, should it automatically open afterwards? & \R{TRUE} \\
+\quad \R{openResult} & If a  pdf/html file is rendered, should it automatically open afterwards, and if not, should the \R{rmarkdown} file? & \R{TRUE} \\
 
@@
 \end{tabular}
-\caption{A selection of commonly used arguments of \R{clean()}.
+\caption{A selection of commonly used\hl{/useful?} arguments of \R{clean()}.
 \hl{Note: I have everything here except \R{file} and \R{quiet}, replace, vol, standAlone + stuff related to selecting SVC-functions (dealt with in next subsection)}.
@@
 
-\hl{Describe arguments in the table...}
+\hl{Describe arguments in Table \ref{table.cleanFormals}. Maybe do a few examples of interesting argument combinations?}
 
@@
 \centering
-\begin{tabular}{p{0.3\linewidth} p{0.35\linewidth} p{0.01\linewidth} p{0.01\linewidth} p{0.01\linewidth} p{0.01\linewidth} p{0.01\linewidth}
+\begin{tabular}{p{0.35\linewidth} p{0.3\linewidth} p{0.01\linewidth} p{0.01\linewidth} p{0.01\linewidth} p{0.01\linewidth} p{0.01\linewidth}
  p{0.01\linewidth} p{0.01\linewidth}}
@@
 \newline Also: Check that everything in here is correct (i.e.
-corresponds to the output of allSummaryFunctions() etc}}
+corresponds to the output of allSummaryFunctions() etc when makeXfunction has been fixed.}}
 \label{table.SVCfunctions}
@@
 \item Introduce \R{allSummaryFunctions()} etc. and present a table corresponding to the output of this call
-\item Small example
+\item Small examples: 
+\begin{itemize}
+\item Add a function to one of the XXXSummaries/XXXVisuals/XXXChecks-arguments (still calling default options)
+\item Remove all but a single function from one of these arguments
+\item Describe what happens if the argument is NULL
+\end{itemize}
 \end{itemize}
@@
 
-By the \textit{contents} of the cleanR output, we refer to every step that actually involves a function being called on variables from the dataset.
-There are two stages in which this occurs \hl{(with ref.
-to figure/flowchart?)}:
+In this section, we discuss how to control every step in \R{clean()} that actually involves a function being called on variables from the dataset.
+There are two stages in which this occurs, as mentioned in Figure~\ref{figure:cleanStructure}, namely:
 \begin{enumerate}
@@
 \pkg{cleanR} uses three different types of functions for performing all stages in the above, namely \code{summaryFunction}s, \code{visualFunction}s and \code{checkFunction}s.
-They each have distinct input-output structures and most instances of the functions are build as \code{S3} generics with methods for different data classes.
-By understanding how to construct proper \code{summaryFunction}s, \code{visualFunction}s and \code{checkFunction}s, the entire contents of the \pkg{cleanR} output is at your hands.
-Therefore, this section is dedicated to introducing each in turn.\\
+\hl{Something like "the avaialable options for functions used in the precheck- and check steps are obtainable by calling \R{allCheckFunction()}, blablabla, similarly with visualFunctions and summaryFunctions.
+Mention Section 4 and the fact that one can expand the possibilities of e.g.
+allVisualFunctions() by producing new functions quite easily.}
 
@@
 Therefore, \R{cleanR} also provides more standard \R{R} interactive tools, such as functions that print results to the console or saves it in a variable for later use.
-Or rather, the functions used by \R{clean()} to do actual data summaries, visualizations and checks, as described above, can also be used interactively.
+Or, more precisely, the functions used by \R{clean()} to do actual data summaries, visualizations and checks, as described above, can also be used interactively.
 \hl{segway}
 
-\hl{Maybe write here that this section introduces the three functions: \R{summarize()}, \R{check()} and \R{visualize()}?}
+\hl{Maybe write here explicitly that this section introduces the three functions: \R{summarize()}, \R{check()} and \R{visualize()}?}
 
-\subsection{Blabla: An example}
-Let's say we wish to look further into a certain variable, namely \R{var2}, from \R{toyData}.
-The data cleaning summary found some issues, and we would like to recall what these issues were.
-This is done using the command
+\subsection{Cleaning by hand: An example}
+Let's say we wish to look further into a certain variable from \R{toyData}, namely \R{var2}.
+The data cleaning summary found some issues in this variable, and we would like to recall what these issues were.
+This can be done using the command
 \begin{Schunk} 
@@
 \end{Schunk}
-The most important thing to note here is that while the printed message is made for human reading, the actual values of the variable causing the issue are still attainable.
-If we for instance decide that the values \R{999} and \R{NaN} in \R{var2} are in fact miscoded missing values, we can replace them easily:
+The most important thing to note here is that while the printed message is made for easy reading, the actual values of the variable causing the issue are still obtainable.
+If we for instance decide that the values \R{999} and \R{NaN} in \R{var2} are in fact miscoded missing values, we can easily replace them with \R{NA}s:
 \begin{Schunk}
@@
 \begin{itemize}
+\item Do an example with visualize() and summarize(), like the one with check(). Especially visualize and to doEval = T thing needs a bit of special attention.
 \item  Mention allCheckFunctions() etc. again here
-\item Mention check(), visualize() and summarize() modes for data.frames
+\item Mention check(), visualize() and summarize() modes for data.frames.
+Maybe also advice against it, at it will often produce a lot of information at once, and such large amounts of information really should be documented.
 \end{itemize}
@@
 Input (required) & \R{v} - a variable vector \newline \R{...}  &  \R{v} - a variable vector \newline \R{vnam} - the variable name (as character string) \newline \R{doEval} - a logical (\R{TRUE}/\R{FALSE}) controlling the output type of the function & \R{v} - a variable vector \newline \R{nMax} - an integer (or \R{Inf}), controlling how many problematic values are printed, if relevant \newline \R{...}  \\
-Input (optional) &  \R{maxDecimals} - number of decimals printed in outputted numerical values  & - &  \R{maxDecimals}  - number of decimals printed in outputted numerical values  \\
+Input (optional) &  \R{maxDecimals} - number of decimals printed in outputted numerical values.  & - &  \R{maxDecimals}  - number of decimals printed in outputted numerical values.  \\
 Purpose & Describe some aspect of the variable, e.g.
-a central value, its dispersion or missingness.
+a central value, its dispersion or level of missingness.
 & Produce a distribution plot.
@@
 \\
-Output (required) & A list with entries \newline \R{\$feature} Label for the summary value (as character string) \newline \R{\$result} The result of the summary (as character string) & A character string with \R{R} code for producing a plot.
+Output (required) & A list with entries \newline \R{\$feature} - a label for the summary value (as character string) \newline \R{\$result} - the result of the summary (as character string) & A character string with \R{R} code for producing a plot.
 This code should be standalone, i.e.
 should include the data if necessary.
-& A list with entries \newline \R{\$problem} - a logical identifying whether an issue was found \newline \R{\$message} - character string (possibly empty) decribing the issue that was found, properly escaped and ready for use in \R{rmarkdown} \\
+& A list with entries \newline \R{\$problem} - a logical identifying whether an issue was found \newline \R{\$message} - a character string (possibly empty) decribing the issue that was found, properly escaped and ready for use in \R{rmarkdown} \\
 Output (recommended) & A \R{summaryResult} object (i.e.
@@
 
-Though the discussion in the above paints a picture of cleanR as a user-friendly package which requires practically no knowledge of R, one should not be mistaken to think that it is not customizable.
-In fact, the main function of cleanR, \code{clean}, is mainly a tool for formatting the results from various checking-, summary- and visualization functions \hl{as described...
-blabla}.
-Thus, the actual work underlying a cleanR output file can be anything or nothing - depending on the arguments given to \code{clean}.
+Though the discussion in the above paints a picture of cleanR as a user-friendly package which requires practically no knowledge of \R{R}, one should not be mistaken to think that it is not customizable.
+In fact, the main function of \R{cleanR}, \code{clean}, is mainly a tool for formatting the results from various checking-, summary- and visualization functions.
+Thus, the actual work underlying a \R{cleanR} output file can be anything or nothing - depending on the arguments given to \code{clean}.
 Specfically, user made functions can be added to the SVC-function arguments discussed above, e.g.
@@
 If the custom functions are only to be used by \R{clean()}, only the input/output structures of the functions need to be restricted.
-However, by adding a few extra tweaks to these functions, the full machinery of \R{cleanR} becomes available for the new, user-made functions, just as it is for the build-in functions.
-The presentation below is given in the format of pseudo-code function templates, designed for getting the full functionality, but please note that Table~\ref{table.functionTypes} serves as a reference to the minimal requirements, while also presenting the "full" versions of the function types.
+However, by adding a few extra \hl{word?} to these functions, the full machinery of \R{cleanR} becomes available for the new, user-made functions, just as it is for the build-in functions.
+The presentation below is given in the format of function templates, written in pseudo-code.
+These templates are designed for getting the full functionality, but please note that Table~\ref{table.functionTypes} serves as a reference to the minimal requirements, while also presenting the "full" versions of the function types.
 
@@
 
-\subsection{Writing cleanR extensions: Function templates}
-\hl{metatext}
+\subsection{Function templates}
+\label{sec:functionTemplates}
+\hl{metatext, mention Table \ref{table.functionTypes} again.}
 
 \subsubsection{Writing a summaryFunction}
-Though \pkg{cleanR} provides a dedicated class for \code{summaryFunction}s, there really is nothing special about these functions: They are nothing but regular functions with a certain input/output-structure.
+As mentioned above, \pkg{cleanR} provides a dedicated class for \code{summaryFunction}s.
+However, this does not imply that they are particularly advanced or complicated to create; in fact, they are nothing but regular functions with a certain input/output-structure.
 Specifically, they all follow the template below:
@@
   res <- [result of whatever summary we are doing]
-  summaryResult(list(feature = "Feature name", result = res))
+  summaryResult(list(feature = "[Feature name]", result = res))
 }
 \end{Verbatim}
-The last function called, \R{summaryResult()}, changes the class of the output and thereby makes a \R{print()} functin available.
+The last function called here, \R{summaryResult()}, changes the class of the output, thereby making a \R{print()} method available for it.
 Note that \code{v} is a vector and that \code{res} should be either a character string or something that will be printed as one.
@@
 integers are allowed, but matrices are not.
-Though a lot of different things can go into the \code{summaryFunction} template, we recommend only using them for summarizing the features of a variable, and leaving tests and checks for the \code{checkFunction}s (presented below).
+Though a lot of different things can go into the \code{summaryFunction} template, we recommend only using it for summarizing the features of a variable, and leaving tests and checks for the \code{checkFunction}s (presented below).
 
-Though following the template above is sufficient for using \R{mySummaryFunction} in \R{clean()}, we recommend furthermore adding it to the overview of all summary functions by converting it to a proper \code{summaryFunction} object.
+Though adhering to the template above is sufficient for using the freshly made \R{mySummaryFunction()} in \R{clean()}, we recommend furthermore adding it to the overview of all summary functions by converting it to a proper \code{summaryFunction} object.
 This is done by writing
@@
 mySummaryFunction <- summaryFunction(mySummaryFunction,
-	description = "Some text describing what your summaryFunction does",
+	description = "[Some text describing what the summaryFunction does]",
 	classes = c([the data types that this function is intended to be used for]))
 \end{Verbatim}
-which adds the new function to the output of \code{allSummaryFunctions()}.
+which adds the new function to the output of an \code{allSummaryFunctions()} call.
 One comment should be devoted to the two attributes of a \R{summaryFunction}.
-If \R{description} is left unspecified, the name of the function (in this case, \R{mySummaryFunction}) will be filled in.
-What happens if the \R{classes} argument depends on the type of \R{mySummaryFunction}.
+If the \R{description} argument is left unspecified, the name of the function (in this case, \R{"mySummaryFunction"}) will be filled in.
+What happens if the \R{classes} argument is not specified depends on the type of \R{mySummaryFunction}.
 If \R{mySummaryFunction} is a \R{S3} generic function with associated methods, the call to \R{summaryFunction()} will automatically produce a vector of the names of the classes for which the function can be called.
@@
 \end{Schunk}
-\hl{write something here, don't end paragraph with code? Also, maybe move the allClasses() stuff somewhere else, it doesn't really belong under this header.}
+\hl{Write something here, don't end paragraph with code. Also, maybe move the allClasses() stuff somewhere else, it doesn't really belong under this header. Not sure where to, though.}
 
@@
 myVisualFunction <- visualFunction(myVisualFunction,
-	description = "Some text describing your visualFunction",
+	description = "[Some text describing the visualFunction]",
 	classes = c([the data types that this function is intended to be used for]))
@@
 In this function, \code{v} is the variable to be visualized, \code{vnam} is its name (which should generally be passed to \code{title} or \code{main} arguments in plotting functions) and \code{doEval} controls whether the output is a plot (if \code{TRUE}) or a character string of standalone code for producing a plot (if \code{FALSE}).
-The \code{doEval = TRUE} setting is not strictly necessary for a \R{visualFunction}'s use in \code{clean}, but it makes it easier to assess what visualization options are available, and obviously, it also allows for interactive usage of \R{myVisualFunction}.
-In either case, it should be noted that all the parameters listed above, \code{v}, \code{vnam} and \code{doEval}, are mandatory, so they should be left as is, even if they are not in use.
+Implementing the \code{doEval = TRUE} setting is not strictly necessary for a \R{visualFunction}'s use in \code{clean}, but it makes it easier to assess what visualization options are available, and obviously, it is crucial for interactive usage of \R{myVisualFunction()}.
+In either case, it should be noted that all the parameters listed above, \code{v}, \code{vnam} and \code{doEval}, are mandatory, so they must be left as is, even if they are not in use.
 
@@
 myFullVarCheckFunction <- checkFunction(myFullVarCheckFunction, 
-  description = "Some text describing your checkFunction",
+  description = "[Some text describing the checkFunction]",
   classes = c([the data types that this function is intended to be used for])
@@
 Again, as with \R{summaryFunction}s and \R{visualFunction}s, the change of function class by use of \R{checkFunction()} is not strictly necessary.
-Note however, that if \R{myFullVarCheckFunction} is to be used in the SVC step in \R{clean()}, the description attribute will be printed in the overview table in \textit{Data cleaning summary}.
+Note however, that if \R{myFullVarCheckFunction} is to be used in the SVC step in \R{clean()}, the description attribute will be printed in the overview table in the \textit{Data cleaning summary}.
 
@@
 myProbValCheckFunction <- checkFunction(myProbValCheckFunction,
-  description = "[Some text describing your checkFunction]",
+  description = "[Some text describing the checkFunction]",
   classes = c([the data types that this function is intended to be used for])
@@
 \end{Verbatim}
-In this template, the argument \R{maxDecimals} is not used.
+In this template, the argument \R{maxDecimals} is not in use.
 This argument should be used to round off the \R{problemValues} passed to \R{messageGenerator()}, if they are numerical.
@@
 This function simply pastes together the \code{problemMessage} and the \code{problemValues}, with the latter being quoted and sorted alphabetically.
-If \R{nMax} is not \R{Inf}, only the first \R{nMax} problem values will be pasted onto the message, accompanied by a comment about how many problem values were left out.
+If \R{nMax} is not \R{Inf}, only the first \R{nMax} problem values will be pasted onto the message, accompanied by a comment about how many problem values were left out (if any).
 Note that printing quotes in rmarkdown requires an extensive amount of character escaping, so opting for \code{messageGenerator()} really is the easiest solution.
@@
 \begin{description}
-\item[\R{isID}] A new \R{checkFunction} intended for use in the precheck-step.
-This function checks whether a variable consists exclusively long ($> 10$) entries that are all of equal length, as this might be personal identification codes that we do not wish to print out in the data summary.
+\item[\R{isID}] A new \R{checkFunction} intended for use in the precheck-stage.
+This function checks whether a variable consists exclusively of long ($> 10$ characters/digits) entries that are all of equal length, as this might be personal identification codes that we do not wish to print out in the data summary.
 \item[\R{mosaicVisual}] A new \R{visualFunction} that produces so-called mosaic plots. This function will be used in the \textit{visualize} step of \R{clean()}.
 \item[\R{countZeros}] A new \R{summaryFunction} that counts the number of occurrences of the value \R{0} in a variable. This function will be used in the \textit{summarize} step of \R{clean()}.
-\item[\R{identifyColons}] A new \R{checkFunction} that flags variables that colons that appear before and after alphanumerical characters, e.g.
+\item[\R{identifyColons}] A new \R{checkFunction} that flags variables in which values have colons that appear before and after alphanumerical characters.
+This is e.g.
 practical for identifying autogenerated interaction effects.
@@
 \end{Schunk} 
-This is essentially all we need to do in order to include this function as a precheck-function in \R{clean()}, so we will leave it as is and move on to the next function, namely \R{mosaicVisual}.
+\hl{Mention somewhere that  \R{mosiacplot()} is a base-R (\R{graphics}) function?
+Otherwise, it might not be completely clear, that we are passing a function name in the above...} This is essentially all we need to do in order to include this function as a precheck-function in \R{clean()}, so we will leave it as is and move on to the next function, namely \R{mosaicVisual}.
 
@@
 Here, we use the function \R{allClasses()} to quickly obtain a vector of all the seven variable classes addressed in \R{cleanR}.
-Note that if R{mosaicVisual} has been an S3 generic function, this argument could have been left as \R{NULL} and then the classes for which methods are available would be added automatically.
+Note that if \R{mosaicVisual} were an S3 generic function, this argument could have been left as \R{NULL} and then the classes for which methods are available would be added automatically.
+\hl{I'm repeating myself here, but I think it is quite a neat feature, so maybe that's okay?}
 
-As \R{mosaicVisual} is now a full blood \R{visualFunction}, it will also be included in the \R{allVisualFunctions()} output table:
+As \R{mosaicVisual} is now a full-blooded \R{visualFunction}, it will also be included in the \R{allVisualFunctions()} output table:
 \begin{Schunk}
@@
 \end{Schunk}
-As in the above, we change the class of this function in order to make it appear in \R{allSummaryFunctions()} calls.
+As with \R{mosaicVisual()}, we change the class of this function in order to make it appear in \R{allSummaryFunctions()} calls.
 But now we wish to emphasize that the function is not intended to be called on all variable types, as zeros have different roles in \R{Date}s and in \R{logical} variables:
@@
 \subsubsection{identifyColons - a new checkFunction with problem values}
-The last function mentioned above is \R{identifyColon()}.
-We define it using the helper function \R{messageGenerator} to obtain a properly escaped message, and we use \R{checkResult} to make it appear in \R{allCheckFunction()} calls:
+The last function mentioned above is \R{identifyColons()}.
+We define it using the helper function \R{messageGenerator} to obtain a properly escaped message, and we use \R{checkResult} to make its output print neatly:
 \begin{Schunk}
@@
 identifyColons <- checkFunction(identifyColons, 
-    description = "Identify non-suffixed nor -prefixed colons",
+    description = "Identify colons surrounded by alphanumeric characters",
     classes = c("character", "factor", "labelled"))
@@
 
-\subsubsection{Calling the new SVC function from clean()}
+\subsubsection{Calling the new SVC functions from clean()}
 Now, we are ready to use these new functions in a \R{clean()} call.
@@
 \section{Something like examples}
-Finally, we will present a few examples of how to make \R{cleanR} solve certain issues related to data cleaning.
+Finally, we will present a few examples of how to make \R{cleanR} solve specific issues related to data cleaning.
 First, we discuss the challenges related to cleaning large datasets, particularly in terms of memory use and computation speed.
@@
 If the dataset becomes very large, the standard use of \R{clean()} outlined above might not be ideal.
-If there is a large number of variables, production of the \R{rmarkdown} document might be very slow, while an extensive amount of observations generally makes rendering of this document into pdf or html very slow.
+If there is a vast number of variables, production of the \R{rmarkdown} document might be quite slow, while an extensive amount of observations generally affects the rendering time of this document.
 In this section, we give a few practical examples of ways to deal with large data, while wishing to still produce (potentially very long) data cleaning overview documents.
-Note that the interactive tools of \R{cleanR} can be used as usual or sequentially in small subsets of the large dataset.
-\\
+Note that the interactive tools of \R{cleanR} can be used as usual or sequentially in small subsets of the large dataset, if no such overview documents are needed.
 
-\subsubsection{Omitting figures}
+\subsubsection{Attacking the figures}
 Though figures give a nice overview of each variable, they are also quite heavy objects in terms of memory allocation.
@@
 \end{Schunk}
-\hl{blabla, end?}
+If figures are indeed needed, a different approach is to choose the less memory heavy standard \R{R} figure style instead of the \R{ggplot2}\hl{reference?} figures that are the default option in \R{clean()}.
+This can be done using the \R{allVisuals} argument:
+\begin{Schunk}
+\begin{Sinput}
+> clean(toyData, allVisuals = "basicVisual")
+\end{Sinput}
+\end{Schunk}
+Of course, even less heavy plots might be achieved by writing new \R{visualFunction}s, using the guidelines from section \ref{sec:functionTemplates}.
+For instance, a future extension of \R{cleanR} might be the inclusion of ASCII plots, as e.g.
+represented in the \R{R} package \R{txtplot}\hl{reference here?
+It's on CRAN.}.
+\\
+
+\hl{I really feel like we should do some benchmarking here, maybe just on toyData, both in terms of speed and memory use. I would make the recommendations more trustworthy and serious.}
 
@@
 \begin{Sinput}
-> clean(toyData, render = FALSE}
+> clean(toyData, render = FALSE, openResult = FALSE}
 > render("cleanR_toyData.Rmd", quiet = FALSE)
@@
 
-\subsection{Use cleanR for problem flagging}
-If the data is large, but memory issues and computation time are less of an issue than the human time it takes to look through the data cleaning document, a viable solution might be to not include all information about all variables.
+\subsection{Using cleanR for problem flagging}
+If the data is large, but memory issues and computation time are less of an issue than the human time it takes to look through the data cleaning document, a viable solution might be not to include all information about all variables.
 Or even for more reasonably sized datasets, sometimes a brief overview of the most pressing issues can be useful.
 This can be achieved by using the \R{onlyProblematic} argument in \R{clean()}.
-By specifying \R{onlyProblematic = TRUE}, only variables that raise a flag in the checking steps will be summarized.
-But perhaps we are not even interested in obtaining general information about these variables, but only in getting a quick overview of the problems, they might have.
+By specifying \R{onlyProblematic = TRUE}, only variables that raise a flag in the checking steps will be summarized and visualized.
+But perhaps we are not even interested in obtaining general information about these variables, but only in getting a quick overview of the problems they might have.
 This can be done by also controlling the \R{mode} argument:
@@
 \end{Schunk}
-Of course, this can also be done without generating an overview document, by use of the \R{check()} function.
+Of course, this can also be done without generating an overview document, by direct use of the \R{check()} function.
 When called on a \R{data.frame}, this function produces a list (of variables) of lists (of checks) of lists (or rather, \R{checkResult}s).
@@
 
-On the other hand, if a pdf document is to be produced, the a few extra lines need to be added to the preamble of the master \R{rmarkdown} document - otherwise, the two-column layout code will produce an error.
-The following is an example of how such a master document preamble might look like and how the \R{cleanR\_toyData.Rmd} can be included:
+On the other hand, if a pdf document is to be produced, a few extra lines need to be added to the preamble of the master \R{rmarkdown} document - otherwise, the two-column layout code will produce an error.
+The following is an example of how such a master document preamble might look like and how the \R{cleanR\_toyData.Rmd} file can then be included:
 \begin{Verbatim}
@@
 \end{Verbatim}
-\hl{Use proper formatting here.}
+\hl{Use proper formatting here. How do we do non-R code?}
 

### 2016-12-16 latex/codeForArticle.R

@@ -75,7 +75,7 @@ mosaicVisual(toyData$var1, "Variable 1", doEval = TRUE)
 dev.off()
 
 
-identifyColon <- function(v, nMax = Inf, ... ) {
+identifyColons <- function(v, nMax = Inf, ... ) {
   v <- unique(na.omit(v))
   problemMessage <- "Note: The following values include colons:"
   problem <- FALSE
@@ -97,7 +97,7 @@ identifyColon <- function(v, nMax = Inf, ... ) {
                    problemValues = problemValues))
 }
 
-identifyColon <- checkFunction(identifyColon, 
+identifyColons <- checkFunction(identifyColons,
                                description = "Identify non-suffixed nor -prefixed colons",
                                classes = c("character", "factor", "labelled"))
 
@@ -110,10 +110,12 @@ clean(testData, replace = T,
       labelledSummaries = c(defaultLabelledSummaries(), "countZeros"),
       numericSummaries = c(defaultNumericSummaries(), "countZeros"),
       integerSummaries = c(defaultIntegerSummaries(), "countZeros"),
-      characterChecks = c(defaultCharacterChecks(), "identifyColon"),
-      factorChecks = c(defaultFactorChecks(), "identifyColon"),
-      labelledCheck = c(defaultLabelledChecks(), "identifyColon"))
+      characterChecks = c(defaultCharacterChecks(), "identifyColons"),
+      factorChecks = c(defaultFactorChecks(), "identifyColons"),
+      labelledCheck = c(defaultLabelledChecks(), "identifyColons"))
       
       
+
+#benchmark CleanR after figure
       
       

### 2016-12-18 R/basicVisual.R
Finished documentation + fixed a bunch of minor bugs.

@@ -2,10 +2,11 @@
 #' \code{\link{barplot}}
 #'
 #' Plot the distribution of a variable, depending on its data class, using the base R
-#' plotting functions.
+#' plotting functions. Note that \code{basicVisual} is a \code{\link{visualFunction}}, compatible with the 
+#' \code{\link{visualize}} and \code{\link{clean}} functions. 
 #'
-#' For character, factor, logical and labelled variables, a barplot is produced. For numeric
-#' or integer variables, \code{basicVisual} produces a histogram instead. Note that for
+#' For character, factor, logical and labelled variables, a barplot is produced. For numeric, 
+#' integer or Date variables, \code{basicVisual} produces a histogram instead. Note that for
 #' integer and numeric variables, all non-finite (i.e. \code{NA}, \code{NaN}, \code{Inf}) values are
 #' removed prior to plotting. For character, factor, labelled and logical variables, only \code{NA}
 #' values are removed.
@@ -56,7 +57,8 @@ basicVisual.Date <- function(v, vnam, doEval = TRUE) basicVisualD(v, vnam, doEva
 
 #Make it a visualFunction
 #' @include visualFunction.R 
-basicVisual <- visualFunction(basicVisual, "Histograms and barplots using graphics")
+basicVisual <- visualFunction(basicVisual, "Histograms and barplots using graphics",
+                              allClasses())
 
 
 

### 2016-12-18 R/centralValue.R

@@ -16,12 +16,16 @@
 #' only NA values are ignored for factor, character, Date and labelled variables. No values are
 #' ignored for logical variables.
 #'
-#' @return A list with \code{$feature}: [Mode/median] and \code{$result}: [the central value of \code{v}].
+#' @return An object of class \code{summaryResult} with the following entries: \code{$feature} 
+#' (the mode/median),\code{$result} (the central value of \code{v}) and \code{$value} (identical 
+#' to \code{$result}).
+#' 
 #' If the mode is returned and it is not uniquely determined, the first value qualifying as a mode is
 #' returned, when the variable is sorted according to \code{\link{sort}}. 
 #'
-#' @seealso \code{\link{summaryFunction}}, \code{\link{summarize}}
-#'
+#' @seealso \code{\link{summaryFunction}}, \code{\link{summarize}}, \code{\link{summaryResult}},
+#' \code{\link{allSummaryFunctions}}
+#' 
 #' @examples
 #'  #central value of an integer variable:
 #'    centralValue(c(rep(1, 25), rep(2, 10), rep(3, 20)))
@@ -33,8 +37,6 @@
 #' @export
 centralValue <- function(v, ...) UseMethod("centralValue")
 
-#' @include summaryFunction.R 
-centralValue <- summaryFunction(centralValue, "Compute median or mode")
 
 #assign methods to generic centralValue function
 
@@ -59,6 +61,9 @@ centralValue.logical <- function(v, ...) centralValueB(v)
 #' @export
 centralValue.Date <- function(v, ...) centralValueCF(v)
 
+
+#' @include summaryFunction.R 
+centralValue <- summaryFunction(centralValue, "Compute median or mode", allClasses())
 
 
 ##########################################Not exported below#########################################

### 2016-12-18 R/checkFunction.R

@@ -1,6 +1,6 @@
 #' @title Create an object of class checkFunction
 #' 
-#' @description convert a function, \code{f}, into an S3 
+#' @description Convert a function, \code{f}, into an S3 
 #' \code{checkFunction} object. This adds \code{f} to the
 #' overview list returned by an \code{allCheckFunctions()} 
 #' call.
@@ -28,11 +28,12 @@
 #' @details \code{checkFunction} represents the functions used in 
 #' \code{\link{check}} and \code{\link{clean}} for performing
 #' error checks and quality control on variables in dataset. 
+#' 
 #' An example of defining a new \code{checkFunction} is given below. 
 #' Note that the minimal requirements for such a function (in order for it to be 
 #' compatible with \code{check()} and \code{clean()}) is the following 
 #' input/output-structure: It must input at least two arguments, namely 
-#' \code{v} (a vector variable) and \code{...}. Addtional implemented 
+#' \code{v} (a vector variable) and \code{...}. Additional implemented 
 #' arguments from \code{check()} and \code{clean()} include \code{nMax} and 
 #' \code{maxDecimals}, see e.g. the pre-defined \code{checkFunction} 
 #' \code{\link{identifyMissing}} for more details about how these arguments should
@@ -52,8 +53,8 @@
 #' 
 #' @include makeXFunction.R messageGenerator.R
 #' 
-#' @seealso \code{\link{allCheckFunctions}} \code{\link{check}} \code{\link{clean}}
-#' \code{\link{messageGenerator}} \code{\link{checkResult}}
+#' @seealso \code{\link{allCheckFunctions}}, \code{\link{check}}, \code{\link{clean}},
+#' \code{\link{messageGenerator}}, \code{\link{checkResult}}
 #' 
 #' @examples 
 #' 

### 2016-12-18 R/checkResult.R

@@ -1,6 +1,6 @@
 #' @title Create object of class checkResult
 #' 
-#' @description Covert a list resulting from the checks performed in a 
+#' @description Convert a list resulting from the checks performed in a 
 #' \code{\link{checkFunction}} into a \code{checkResult} object, thereby
 #' supplying it with a \code{print()} method.
 #' 

### 2016-12-18 R/countMissing.R

@@ -7,9 +7,9 @@
 #' @param v A variable (vector).
 #' @param ... Not in use.
 #'
-#' @return A \code{\link{summaryResult}} object with \code{$feature}:
-#' "No. missing obs.", \code{$result}: [the number and percentage missing observations].
-#' and \code{$value}: [the number of missing observations].
+#' @return A \code{\link{summaryResult}} object with the following entries: 
+#' \code{$feature} ("No. missing obs."), \code{$result} (the number and percentage 
+#' missing observations) and \code{$value} (the number of missing observations).
 #'
 #' @seealso \code{\link{summarize}}, \code{\link{allSummaryFunctions}},
 #' \code{\link{summaryFunction}}, \code{\link{summaryResult}}

### 2016-12-18 R/fancyVisual.R

@@ -1,94 +1,0 @@
-##########################################Not exported below#########################################
-
-
-########fancyVisual###########
-
-#NOT DONE - SHOULD NOT BE EXPORTED. NEED TO IMPLIMENT DOEVAL=F STUFF FIRST...
-
-#THINK ABOUT: use myDistrPlots-ish function from my thesis?
-#       - adds density smoothers to histograms
-#       - for barplots: move x-axis breaks into the plot for more
-#           control wrt. the (actual) size of the plot
-
-#TO DO: If we are to keep "fancy"-option: Rewrite myDistrPlots
-#       in s3 style and such that it is only designed to produce one
-#       plot
-
-#NOTE: the use of substitute() requires that v is a named vector.
-#       better way to do this? And will this be a problem in our
-#       setting? I don't think so.
-
-#fancyVisual <- function(v, vnam) {
-#  data <- data.frame(v)
-#  names(data) <- vnam
-#  myDistrPlots(data, vnam)[[1]]
-#}
-
-
-# @importFrom ggplot2 ggplot aes aes_string geom_bar geom_text scale_fill_discrete scale_x_continuous xlab ylab ggtitle theme_bw geom_histogram geom_line scale_x_discrete
-#myDistrPlots <- function(data, var, labels=NULL,
-#                         makeCont=NULL,
-#                         makeCat=NULL,
-#                         catStyle="bars",
-#                         adjustFunction=function(x){1}) {
-#  if (is.null(labels)) labels <- var
-#  n <- length(var)
-#  outLst <- NULL
-#  if (class(data[, var]) %in% c("logical", "character", "labelled")) {
-#    data[, var] <- factor(data[, var])
-#  }
-#  for (i in 1:n) {
-#    thisVar <- var[i]
-#    if ((is.factor(data[, thisVar]) & !(thisVar %in% makeCont)) |
-#        thisVar %in% makeCat) {
-#      if (thisVar %in% makeCat) {
-#        data[, thisVar] <- factor(data[, thisVar])
-#      }
-#      thisTab <- table(data[, thisVar])
-#      thisCount <- as.numeric(c(thisTab))
-#      ## Possibly define freq and lab as NULL so the devtools::check function does not complain
-#      tpDat <- data.frame(freq=thisCount/sum(thisCount),
-#                          lab=factor(names(thisTab), levels=levels(data[, thisVar])))
-#      if (catStyle == "stacked") {
[…]

### 2016-12-18 R/identifyCaseIssues.R

@@ -38,7 +38,8 @@ identifyCaseIssues.factor <- function(v, nMax = Inf) identifyCaseIssuesF(v, nMax
 
 #make it a checkFunction
 #' @include checkFunction.R
-identifyCaseIssues <- checkFunction(identifyCaseIssues, "Identify case issues")
+identifyCaseIssues <- checkFunction(identifyCaseIssues, "Identify case issues",
+                                    c("character", "factor"))
 
 
 ##########################################Not exported below#########################################

### 2016-12-18 R/identifyLoners.R

@@ -42,7 +42,8 @@ identifyLoners.character <- function(v, nMax = Inf) identifyLonersC(v, nMax = nM
 
 #make it a checkFunction
 #' @include checkFunction.R
-identifyLoners <- checkFunction(identifyLoners, "Identify levels with < 6 obs.")
+identifyLoners <- checkFunction(identifyLoners, "Identify levels with < 6 obs.",
+                                c("character", "factor"))
 
 ##########################################Not exported below#########################################
 

### 2016-12-18 R/identifyMissing.R

@@ -69,7 +69,8 @@ identifyMissing.logical <- function(v, nMax = Inf, ...) identifyMissingB(v, nMax
 
 #make it a checkFunction
 #' @include checkFunction.R
-identifyMissing <- checkFunction(identifyMissing, "Identify miscoded missing values")
+identifyMissing <- checkFunction(identifyMissing, "Identify miscoded missing values",
+                                 setdiff(allClasses(), "Date"))
 
 ##########################################Not exported below#########################################
 

### 2016-12-18 R/identifyOutliers.R

@@ -59,7 +59,8 @@ identifyOutliers.integer <- function(v, nMax = Inf, maxDecimals = 2) {
 
 #make it a checkFunction
 #' @include checkFunction.R
-identifyOutliers <- checkFunction(identifyOutliers, "Identify outliers")
+identifyOutliers <- checkFunction(identifyOutliers, "Identify outliers",
+                                  c("integer", "numeric"))
 
 
 ##########################################Not exported below#########################################

### 2016-12-18 R/identifyOutliersTBStyle.R

@@ -46,7 +46,8 @@ identifyOutliersTBStyle.integer <- function(v, nMax = Inf, maxDecimals = 2) {
 #make it a checkFunction
 #' @include checkFunction.R
 identifyOutliersTBStyle <- checkFunction(identifyOutliersTBStyle, 
-                                         "Identify outliers (Turkish Boxplot style)")
+                                         "Identify outliers (Turkish Boxplot style)",
+                                         c("integer", "numeric"))
 
 
 

### 2016-12-18 R/identifyWhitespace.R

@@ -43,7 +43,8 @@ identifyWhitespace.labelled <- function(v, nMax = Inf) identifyWhitespaceL(v, nM
 
 #make it a checkFunction
 #' @include checkFunction.R
-identifyWhitespace <- checkFunction(identifyWhitespace, "Identify prefixed and suffixed whitespace")
+identifyWhitespace <- checkFunction(identifyWhitespace, "Identify prefixed and suffixed whitespace",
+                                    c("character", "factor", "labelled"))
 
 
 ##########################################Not exported below#########################################

### 2016-12-18 R/makeXFunction.R

@@ -1,20 +1,24 @@
 
-# '   @e xport
-#'@importFrom utils .S3methods methods
+#Make a function of type X (among visualFunction, summaryFunction and
+#checkFunction). If classes is null, and fName is a S3 generic, 
+#the function will make a look-up for all methods available in the
+#global enviroment and fill out the slot accordingly.
+#Note: This does not work for functions defined wihtin cleanR. As 
+#of now, they must have their classes specified explicitly!
+#'@importFrom utils methods
 makeXFunction <- function(fName, description, classes, X) {
   f <- get(fName)
+  
+  if (is.null(description)) description <- fName
   if (is.null(classes)) {
-    #browser()
-    theseMethods <- as.character(utils::methods(fName)) #methods() needs the name in order
-    if (length(theseMethods) == 0) {
-      callEnv <- parent.env(as.environment(-1L))
-      theseMethods <- as.character(utils::.S3methods(fName, envir = callEnv))
-    }
+    theseMethods <- as.character(methods(fName)) #methods() needs the name in order
     #to work inside the function
-    #print(standardVisual)
-    #browser()
-    # print(fName)
-    #print(methods)
+    
+    #if (length(theseMethods) == 0) {
+    #  callEnv <- parent.env(as.environment(-1L))
+    #  theseMethods <- as.character(utils::.S3methods(fName, envir = callEnv))
+    #}
+    
     if (length(theseMethods) > 0) {
       classes <- sub(paste(fName, ".", sep=""),
                      "", theseMethods)
@@ -25,6 +29,8 @@ makeXFunction <- function(fName, description, classes, X) {
   attr(f, "classes") <- classes
   f
 }
+
+
 
 
 #foo <- function(x) UseMethod("foo")

### 2016-12-18 R/messageGenerator.R

@@ -1,34 +1,38 @@
 #' @title Produce a message for the output of a checkFunction
-#' @description Internal function for producing output messages in \code{\link{checkFunction}}
-#' type functions called by \code{\link{clean}}.
+#' @description Helper function for producing output messages for 
+#' \code{\link{checkFunction}} type functions. 
 #'
-#' @param problemStatus A list consisting of two entries: $problem - TRUE/FALSE indicating
-#' whether a problem was found by the \code{checkFunction}, $problemValues - a vector of values
-#' from the variable that were deemed problematic. See details below.
+#' @param problemStatus A list consisting of two entries: 
+#' 
+#' \code{$problem} - logical indicating whether a problem was found by the 
+#' \code{checkFunction} responsible for the making the \code{messageGenerator()} call, 
+#' 
+#' \code{$problemValues} - a vector of values from the variable that were 
+#' deemed problematic (see details below).
 #'
 #' @param message Optional, but recommended. A message describing what problem the
 #' problem values are related to. If \code{NULL} a standard message is added using the name
 #' of the function that called \code{messageGenerator}.
 #'
 #' @param nMax Maximum number of problem values to be printed in the message. If the total
-#' number of problem values exceeds nMax, the number of remaining (non-printed) problem
-#' values are added to the message. Defaults to \code{Inf} in which case all problem
+#' number of problem values exceeds nMax, the number of omitted problem
+#' values are added to the message. Defaults to \code{Inf}, in which case all problem
 #' values are printed.
 #'
 #' @details This function is a tool for building \code{\link{checkFunction}}s for the
-#' cleanR \code{\link{clean}} function. \code{checkFunction}s will often identify a number
+#' \code{cleanR} \code{\link{clean}} function. \code{checkFunction}s will often identify a number
 #' of values in a variable that are somehow problematic. \code{messageGenerator} takes
 #' these values, pastes them together with a problem description and makes sure that the
-#' formatting is appropriate for being rendered in a rmarkdown document. We recommend writing
-#' short at precise problem descriptions (see examples), but if no message is supplied,
-#' the following message is generated:
+#' formatting is appropriate for being rendered in a \code{rmarkdown} document. 
+#' We recommend writing short and precise problem descriptions (see examples),
[…]

### 2016-12-18 R/minMax.R

@@ -3,22 +3,28 @@
 #' @description A description type function to be called from \code{\link{summarize}}, which returns
 #' the minimum and maximum value of a variable. NA, NaN and Inf values are removed prior to  the
 #' computations.
-#'
+#' @description A description type \code{summaryFunction}, intended to be called from 
+#' \code{\link{summarize}}, which returns the minimum and maximum values of a variable.
+#' NA, NaN and Inf values are removed prior to  the computations.
+#' 
 #' @param v A variable (vector) of type numeric or integer.
 #'
 #' @inheritParams clean
+#' 
+#' @return An object of class \code{summaryResult} with the following entries: \code{$feature} 
+#' ("Min. and max."), \code{$result} (the minimum and maximum of \code{v}), and \code{$value}
+#' (minimum and maximum in their orignial format).
 #'
-#' @return A list with $feature: "Min. and max." and $result: [the minimum and maximum of \code{v}].
-#'
-#' @seealso \code{\link{summarize}}
-#'
+#' @seealso \code{\link{summaryFunction}}, \code{\link{summarize}}, \code{\link{summaryResult}},
+#' \code{\link{allSummaryFunctions}}
+#' 
 #' @examples
 #' minMax(c(1:100))
 #'
 #' @importFrom stats na.omit
 #' @export
 minMax <- function(v, maxDecimals = 2) {
-  v <- na.omit(v) #maybe keep Inf's?
+  v <- na.omit(v) #maybe keep Infs instead?
   minV <- min(v)
   maxV <- max(v)
   summaryResult(list(feature="Min. and max.",
@@ -27,5 +33,8 @@ minMax <- function(v, maxDecimals = 2) {
                                   sep="; "),
                      value = c(minV, maxV)))
 }
+
+#Make it a summaryFunction
+#' @include summaryFunction.R
 minMax <- summaryFunction(minMax, "Find minimum and maximum values",
                           c("integer", "numeric", "Date"))

### 2016-12-18 R/oClass.R

@@ -1,4 +1,0 @@
-#' @export
-oClass <- function(v) UseMethod("oClass")
-oClass.default <- function(v) class(v)
-oClass.smartNum <- function(v) attr(v, "originalClass")

### 2016-12-18 R/quartiles.R

@@ -3,6 +3,10 @@
 #' @description A description type function to be called from \code{\link{summarize}}, which calculates
 #' the 1st and 3rd quartiles of a variable. NA, NaN and Inf values are removed prior to  the
 #' computations.
+#' 
+#' @description A description type \code{summaryFunction}, intended to be called from 
+#' \code{\link{summarize}}, which calculates the 1st and 3rd quartiles of a variable. 
+#' NA, NaN and Inf values are removed prior to  the computations.
 #'
 #' @param v A variable (vector) of type numeric or integer.
 #' 
@@ -10,10 +14,18 @@
 #'
 #' @return A list with $feature: "1st and 3rd quartiles" and $result: [the quartiles of \code{v}].
 #'
-#' @seealso \code{\link{summarize}}
+#'
+#' @return An object of class \code{summaryResult} with the following entries: \code{$feature} 
+#' ("1st and 3rd quartiles"), \code{$result} (the 1st and 3rd quartiles of \code{v}) and
+#' \code{$value} (the quartiles in their original format).
+#'
+#' @seealso \code{\link{summaryFunction}}, \code{\link{summarize}}, \code{\link{summaryResult}},
+#' \code{\link{allSummaryFunctions}}
 #'
 #' @examples
 #' quartiles(c(1:100))
+#' 
+#' quartiles(rnorm(1000), maxDecimals = 4)
 #'
 #' @importFrom stats na.omit quantile
 #' @export
@@ -25,6 +37,8 @@ quartiles <- function(v, maxDecimals = 2) {
                                     collapse="; "),
                      value = as.numeric(quants)))
 }
+
+#' @include summaryFunction.R
 quartiles <- summaryFunction(quartiles, "Compute 1st and 3rd quartiles",
                              classes = c("integer", "numeric"))
 

### 2016-12-18 R/render.R

@@ -1,7 +1,23 @@
-
-# make pdf/html from a .Rmd file
-# CE: made this an internal function  # ' @export
-# This function just calls the corresponding function from rmarkdown
+#' @title Simplified Rmarkdown rendering
+#' 
+#' @description Render a Rmarkdown (.Rmd) file, \code{file}, to the output
+#' format specified in its preamble. If no output format is specified,
+#' it will be rendered to html.
+#' 
+#' @param file A character string path to the file that is to be rendered. 
+#' This file must be of type Rmarkdown (.Rmd)
+#' 
+#' @param quiet A logical. Should messages during rendering be surpressed?
+#' 
+#' @details This function is merely a simplified version (in terms of 
+#' possible arguments) of the rendering function from the \code{rmarkdown} package. 
+#' Therefore, we refer to this functions for more details:  
+#' \code{\link[rmarkdown]{render}}. We have included this simplified version in 
+#' \code{cleanR} in order to help new R users with rendering their output 
+#' documents as generated by \code{\link{clean}}.
+#' 
+#' @seealso \code{\link[rmarkdown]{render}}.
+#' 
 #' @export
 render <- function(file, quiet) rmarkdown::render(file, quiet=quiet)
 

### 2016-12-18 R/smartNum.R

@@ -1,4 +1,8 @@
-#' @export
+#S3 class meant for representing numeric variables that act like 
+#factor variables by taking only a few different values. This class
+#is used in clean() in order to get appropriate summaries, visualizations
+#and checks for such variables. In other words, such variables will be
+#treated like factor variables instead of numerics.
 smartNum <- function(v) {
   oriClass <- class(v)
   v <- factor(v)
@@ -6,3 +10,9 @@ smartNum <- function(v) {
   class(v) <- c("smartNum", "factor")
   v
 }
+
+
+#Get the original class of a smartNum object. Used in clean().
+oClass <- function(v) UseMethod("oClass")
+oClass.default <- function(v) class(v)
+oClass.smartNum <- function(v) attr(v, "originalClass")

### 2016-12-18 R/standardVisual.R

@@ -1,12 +1,14 @@
 #' Produce distribution plots using ggplot from ggplot2.
 #'
-#' Plot the distribution of a variable, depending on its data class, by use of ggplot2.
+#' Plot the distribution of a variable, depending on its data class, by use of ggplot2. 
+#' Note that \code{standardVisual} is a \code{\link{visualFunction}}, compatible with the 
+#' \code{\link{visualize}} and \code{\link{clean}} functions. 
 #'
-#' For character, factor, logical and labelled variables, a barplot is produced. For numeric
-#' or integer variables, \code{standardVisual} produces a histogram instead. Note that for
+#' For character, factor, logical and labelled variables, a barplot is produced. For numeric,
+#' integer or Date variables, \code{standardVisual} produces a histogram instead. Note that for
 #' integer and numeric variables, all non-finite (i.e. \code{NA}, \code{NaN}, \code{Inf}) values are
-#' removed prior to plotting. For character, factor, labelled and logical variables, only \code{NA}
-#' values are removed.
+#' removed prior to plotting. For character, Date, factor, labelled and logical variables, 
+#' only \code{NA} values are removed.
 #'
 #' @param v The variable (vector) to be plotted.
 #' @param vnam The name of the variable which will appear as the title of the plot.
@@ -16,9 +18,11 @@
 #' @examples
 #' \dontrun{
 #' #Save a variable
-#' myVar <- c(1:10)
+#' myVar <- c(1:10) 
+#' 
 #' #Plot a variable
 #' standardVisual(myVar, "MyVar")
+#' 
 #' #Produce code for plotting a variable
 #' standardVisual(myVar, "MyVar", doEval = FALSE)
 #' }
@@ -26,7 +30,7 @@
 #'
 #' @importFrom ggplot2 qplot
 #' @importFrom stats na.omit
-## #' @export
+#' @export
 standardVisual <- function(v, vnam, doEval = TRUE) UseMethod("standardVisual")
 
 #assign methods to generic standardVisual function
@@ -57,8 +61,9 @@ standardVisual.Date <- function(v, vnam, doEval = TRUE) standardVisualD(v, vnam,
 
 
 
-#' @export
-standardVisual <- visualFunction(standardVisual, "Histograms and barplots using ggplot2")
+#' @include visualFunction.R
+standardVisual <- visualFunction(standardVisual, "Histograms and barplots using ggplot2",
+                                 allClasses())
 
 
[…]

### 2016-12-18 R/summarize.R

@@ -1,36 +1,49 @@
-#' Make summary matrix
+#' @title Make summary matrix
 #'
-#' Generic shell function that produces a summary matrix for a variable \code{v},
-#' given a number of summary functions and depending on its data class.
+#' @description Generic shell function that produces a summary matrix for a variable (or for each
+#' variable in an entire dataset), given a number of summary functions and 
+#' depending on its data class.
 #'
-#' @param v The variable (vector) which is to be summarized.
-#' @param descriptive If TRUE (the default) then descriptive functions are added to the summary matrix
-#' (see details). They are not included if set to FALSE.
+#' @param v The variable (vector) or dataset (data.frame) to be summarized.
+#' 
+#' @param descriptive A logical. If \code{TRUE} (the default), descriptive functions 
+#' are added to the summary matrix (see details). They are not included if 
+#' \code{descriptive} is set to \code{FALSE}.
+#' 
 #' @param ... Additional argument passed to data class specific methods. First and foremost,
 #' this is where to supply the summary functions and description functions (see details).
 #'
 #' @details Summary functions are supplied using their
 #' names (in character strings) in the class-specific argument, e.g.
 #' \code{characterSummaries = c("countMissing", "uniqueValues")} for character variables and
-#' similarly for the remaining 5 data classes (factor, labelled, numeric, integer, logical).
-#' Secondly, additional functions can be added using the "descriptions"-arguments, e.g.
+#' similarly for the remaining 6 data classes (factor, Date, labelled, numeric, integer, logical).
+#' Note that an overview of all available \code{summaryFunction}s can be obtained by calling
+#' \code{\link{allSummaryFunctions}}. 
+#' 
+#' Additional functions can be added using the "descriptions"-arguments, e.g.
 #' \code{characterDescriptions = "centralValue"}. If \code{descriptive = TRUE} the results of
 #' these functions are added to the outputted summary matrix. Note that this is nothing
 #' more than a convenient way to add an extra category of summary functions to the output.
-#' Summary functions and description functions are treated in the exact same way.
+#' Summary functions and description functions are treated in the exact same way and they
+#' should both ideally be of class \code{\link{summaryFunction}}. 
 #'
 #' The default summary- and character functions are available in data class specific functions, e.g.
[…]

### 2016-12-18 R/summaryFunction.R

@@ -1,65 +1,75 @@
-#' Construct a summaryFunction
+#' @title Create an object of class summaryFunction
+#' 
+#' @description Convert a function, \code{f}, into an S3 
+#' \code{summaryFunction} object. This adds \code{f} to the
+#' overview list returned by an \code{allSummaryFunctions()} 
+#' call.
 #'
-#' Convert a function into class \code{summaryFunction} used for creating variable
-#' summaries in \code{\link{summarize}}, typically called from \code{\link{clean}}.
-#'
-#' The structure of a valid \code{summaryFunction} follows the following structure:
-#' \code{
-#'    mySummaryFunction <- function(v) {
-#'      res <- [result of whatever summary we are doing]
-#'      list(feature = "Feature name", result = res)
-#'    }
+#' @inheritParams checkFunction
+#' 
+#' @param description A character string describing the summary
+#' returned by \code{f}. If \code{NULL} (the default), the
+#' name of \code{f} will be used instead.
+#' 
+#' @return A function of class \code{summaryFunction} which has to attributes, 
+#' namely \code{classes} and \code{description}. 
+#' 
+#' @details \code{summaryFunction} represents the functions used in 
+#' \code{\link{summarize}} and \code{\link{clean}} for summarizing the
+#' features of variables in a dataset.
+#' 
+#' An example of defining a new \code{summaryFunction} is given below. 
+#' Note that the minimal requirements for such a function (in order for it to be 
+#' compatible with \code{summarize()} and \code{clean()}) is the following 
+#' input/output-structure: It must input at least two arguments, namely 
+#' \code{v} (a vector variable) and \code{...}. Additional implemented 
+#' arguments from \code{summarize()} and \code{clean()} include
+#' \code{maxDecimals}, see e.g. the pre-defined \code{summaryFunction} 
+#' \code{\link{minMax}} for more details about how this arguments should
+#' be used. 
+#' The output must be a list with at least the two entries \code{$feature} 
+#' (a short character string describing what was summarized) and \code{$result} 
+#' (a value or a character string with the result of the summarization). 
+#' However, if the result of a \code{summaryFunction} is furthermore 
+#' converted to a \code{\link{summaryResult}} object, a \code{print()} 
+#' method also becomes available for consistent formatting of 
+#' \code{summaryFunction} results.
+#' 
+#' Note that all available \code{summaryFunction}s are listed by the call
[…]

### 2016-12-18 R/summaryResult.R

@@ -1,15 +1,14 @@
-#' Create object of class summaryResult
+#' @title Create object of class summaryResult
 #'
-#' Covert a list resulting from the summaries performed in a
+#' @description Convert a list resulting from the summaries performed in a
 #' \code{\link{summaryFunction}} into a \code{summaryResult} object, thereby
 #' supplying it with a \code{print()} method.
 #'
-#' @param ls A list with entries \code{$problem} (logical indicating whether
-#' a problem was found), \code{$message} (a character string containing a
-#' message describing the problem) and \code{$problemValues} (the values
-#' in the checked variables that were marked as problematic). Note that
-#' \code{$message} and \code{$problemValues} can be left empty (i.e.
-#' \code{""} and \code{NULL}, respectively), if they are not relevant.  XXX
+#' @param ls A list with entries \code{$feature} (a character string describing
+#' what summary was obtained), \code{$result} (the result of the summary, either
+#' a value from the variable, a numeric or a character string) and 
+#' \code{$value} (the result in its most raw format, often identical to the 
+#' \code{$result} input). 
 #'
 #' @return A S3 object of class \code{summaryResult}, identical to the inputted
 #' list, \code{ls}, except for its class attribute.
@@ -19,7 +18,7 @@
 #' @export
 summaryResult <- function(ls) {
   if (length(setdiff(names(ls), c("feature", "result", "value"))) != 0) {
-    stop("Hmm, what's informative here? Describe issue...")
+    stop("The inputted list does not qualify as a summaryResult")
   }
   class(ls) <- "summaryResult"
   ls

### 2016-12-18 R/uniqueValues.R

@@ -1,16 +1,19 @@
-#' @title Summary function for unique values
+#' @title summaryFunction for unique values
 #'
-#' @description A summary type function to be called from \code{\link{summarize}}, which counts the
-#' number of unique (excluding NA) values in a variable.
+#' @description A \code{\link{summaryFunction}} type function, intended to be called from 
+#' \code{\link{summarize}} to be called from \code{\link{summarize}}, which counts the
+#' number of unique (excluding \code{NA}s) values in a variable.
 #'
 #' @param v A variable (vector).
 #'
 #' @param ... Not in use.
 #'
-#' @return A list with $feature: "No. unique values" and
-#' $result: [the number of unique values].
+#' @return An object of class \code{summaryResult} with the following entries: 
+#' \code{$feature} ("No. unique values") and \code{$result} (the number of unique 
+#' values in \code{v}).
 #'
-#' @seealso \code{\link{summarize}}
+#' @seealso \code{\link{summaryFunction}}, \code{\link{summarize}}, \code{\link{summaryResult}},
+#' \code{\link{allSummaryFunctions}}
 #'
 #' @examples
 #' uniqueValues(c(1:3, rep(NA, 10), Inf, NaN))
@@ -44,8 +47,9 @@ uniqueValues.Date <- function(v, ...) uniqueValuesCFLBI(v)
 
 
 #Make it a summaryFunction
+#' @include summaryFunction.R
 #' @export
-uniqueValues <- summaryFunction(uniqueValues, "Count number of unique values")
+uniqueValues <- summaryFunction(uniqueValues, "Count number of unique values", allClasses())
 
 
 ##########################################Not exported below#########################################

### 2016-12-18 R/unpackLabelled.R

@@ -1,7 +1,6 @@
 #Maybe delete? Deal when labelled stuff is implemented.
 #this function is used in identifyMissing.R and
 #identifyWhitespace.R
-#' @export
 unpackLabelled <- function(v) {
   c(as.character(v), attributes(v)$labels)
 }

### 2016-12-18 R/variableType.R

@@ -1,15 +1,17 @@
 #' @title Summary function for original class
 #'
-#' @description A summary type function to be called from \code{\link{summarize}}, which finds the
+#' @description A \code{\link{summaryFunction}} type function, intended to be called from 
+#' \code{\link{summarize}} to be called from \code{\link{summarize}}, which finds the
 #' orignial class of a variable. This is just the class for all objects but those of class
-#' smartNum.
+#' \code{smartNum}.
 #'
 #' @param v A variable (vector).
 #'
 #' @param ... Not in use.
 #'
-#' @return A list with $feature: "Variable type" and
-#' $result: [the (original) class of the variable].
+#' @return An object of class \code{summaryResult} with the following entries: 
+#' \code{$feature} ("Variable type"), \code{$result} (the (original) class of 
+#' \code{v}) and \code{$value} (identical to \code{$result}).
 #'
 #' @seealso \code{\link{summarize}}
 #'
@@ -20,18 +22,16 @@
 #'    variableType(varX)
 #'
 #'  #For smartNum variables:
-#'    smartX <- smartNum(varX)
+#'    smartX <- cleanR:::smartNum(varX)
 #'    class(smartX)
 #'    variableType(smartX)
 #'
-#'  #Something with multiple classes, like ordered factor...
-#'
+#' @include smartNum.R
 #' @export
 variableType <- function(v, ...) {
   vClass <- oClass(v)[1]
   summaryResult(list(feature="Variable type", result = vClass, value = vClass))
 }
 
-variableType <- summaryFunction(variableType, "Data class of variable",
-                                c("character", "factor", "integer", "labelled", "logical",
-                                  "numeric", "Date"))
+#' @include variableType.R
+variableType <- summaryFunction(variableType, "Data class of variable", allClasses())

### 2016-12-18 R/visualFunction.R

@@ -1,49 +1,49 @@
-#' @title Construct a visualFunction
+#' @title Create an object of class visualFunction
+#' 
+#' @description Convert a function, \code{f}, into an S3 
+#' \code{visualFunction} object. This adds \code{f} to the
+#' overview list returned by an \code{allVisualFunctions()} 
+#' call.
 #'
-#' @description Convert a function into class \code{visualFunction} used for creating variable
-#' visualizations \code{\link{visualize}}, typically called from \code{\link{clean}}.
-#'
-#' @param f A function. In order to be a valid \code{visualFunction}, \code{f} should
-#' follow the structure outlined in details below.
-#'
-#' @inheritParams summaryFunction
-#'
-#' @return A function of class \code{visualFunction}, i.e. with two attributes: A description
-#' ("description") and a vector of classes which the function can be called upon ("classes"). A
-#' visualFunction always
-#'
-#'
-#'
-#' @details We recommend using the following structure for defining new \code{visualFunction}s:
-#' FIND A WAY TO FIX FORMATTING HERE. GOTTA BE POSSIBLE.
-#' #' \code{
-#'   myVisualFunction <- function(v, vnam, doEval) {
-#'    thisCall <- call("[the name of the function used to produce the plot]",
-#'                     v, [additional arguments to the plotting function])
-#'    if (doEval) {
-#'     return(eval(thisCall))
-#'    } else return(deparse(thisCall)
-#'   }
-#'  }
-#'  followed by a class change:
-#'  \code{
-#'    myVisualFunction <- visualFunction(myVisualFunction, description = "[describe function]",
-#'                                         classes = c([classes]))
-#'  }
-#'  It is not strictly necessary to follow this exact template, but the input/output structure
-#'  should be respected - in other words, the arguments are mandatory and the function needs
-#'  to return a character string containing code (if \code{doEval = FALSE}) or have the
-#'  side effect of producing a plot (if \code{doEval = TRUE}). See examples below for an example
-#'  of how the template can be used in practice.
-#'
-#'  Note that it is not necessary to formally change functions to be used by \code{\link{visualize}}
-#'  into \code{visualFunction}s, but it is recommended, as the functions are then added to the output
-#'  of \code{allVisualFunctions()}, making it easier to gain an overview of the visualization
[…]

### 2016-12-18 R/visualize.R

@@ -1,65 +1,85 @@
-#'@title Produce marginal distribution plots
+#' @title Produce distribution plots
 #'
-#'@description Generic shell function that calls a plotting function in order to produce a marginal
-#'distribution plot for a variable, depending on the class of the variable.
+#' @description Generic shell function that calls a plotting function in order to produce a marginal
+#' distribution plot for a variable (or for each variable in a dataset). What type of plot is made 
+#' might depend on the data class of the variable.
 #'
-#'@param v The variable (vector) which is to be plotted.
-#'@param vnam The name of the variable. This name might be printed on the plots, depending on the
-#'choice of plotting function. If not supplied, it will default to the name of \code{v}.
-#'@param allVisuals The name (as a character string) of the actual plotting function to be used, no
-#'matter the class of \code{v}. See \code{details} for more details about the structure of such plotting
-#'functions. Note that this option is overwritten if a non-null class specific
-#'plotting function is supplied in \code{...} (see \code{details}). Two options are readily available,
-#'\code{\link{standardVisual}} and \code{\link{basicVisual}}.
-#'@param doEval If TRUE (the default), \code{visualize} returns a plot (IS THIS THE CORRECT WAY TO SAY IT?). Otherwise,
-#'visualize returns a character string containing R-code for producing a plot.
-#'@param ... ALLOW FOR ARGUMENTS TO BE PASSED ON TO E.G. STANDARDVISUAL AS WELL? If the plotting
-#'function to be used should differ by variable type, this can be specified by using additional arguments
-#'on the form \code{characterVisual = "standardVisual"}, ..., \code{logicalVisual = "standardVisual"} for
-#'each of the 6 data classes \code{character}, \code{factor}, \code{labelled}, \code{numeric},
-#'\code{integer} and \code{logical}.
+#' @param v The variable (vector) or dataset (data.frame) which is to be plotted.
 #'
-#'@details The function supplied in \code{allVisuals} should take a variable, \code{v}, and its name,
-#'\code{vnam}, and an evaluation indicator, \code{doEval}, and return a character string containing
-#'(standalone) code for producing a plot if \code{doEval = FALSE} and a plot otherwise.
[…]

### 2016-12-18 latex/article_vol2.tex

@@
 %% an abstract and keywords
-\Abstract{Data cleaning and -validation are the first steps in any data
-  analysis, as the validity of the conclusions from the analysis are
-  hinged on the quality of the input data.
-Mistakes in the data can arise for any number of reasons, including erroneous codings, malfunctioning measurement equipment, inconsistent data generation manuals and many more.
-Ideally, a human investigator
-  should go through each variable in the dataset and look for
-  potential errors --- both in input values and coding.  \hl{segway: But, ...}
+\Abstract{Data cleaning and -validation are the first steps in any
+  data analysis, as the validity of the conclusions from the analysis
+  hinges on the quality of the input data. Mistakes in the data can
+  arise for any number of reasons, including erroneous codings,
+  malfunctioning measurement equipment, inconsistent data generation
+  manuals and many more.  Ideally, a human investigator should go
+  through each variable in the dataset and look for potential errors
+  --- both in input values and coding --- but that process can be very
+  time-consuming, expensive and error-prone by itself.
 
@@
   and customizeable suite of quality assessment tools to be applied to
-  a dataset in order to identify potential problems in its variables. The results can be presented in an auto-generated, non-technical, stand-alone
-  overview document, intended to be perused by an investigator with an
-  understanding of the variables in the data and the experimental design, but not necessarily knowledge of \R{R}.
-Thereby, \R{cleanR} aids the dialogue between data analysist and field experts, while also providing easy documentation of data cleaning steps and data quality control.
-Moreover, the \R{cleanR} solution changes the data cleaning process from the usual ad hoc approach to a systematic, well-documented endeavor.
-  \R{cleanR} also provides a suite of more typical \R{R} tools for interactive data quality assessment and -cleaning.
+  a dataset in order to identify potential problems in its
+  variables. The results can be presented in an auto-generated,
+  non-technical, stand-alone overview document, intended to be perused
+  by an investigator with an understanding of the variables in the
+  data and the experimental design, but not necessarily knowledge of
+  \R{R}. Thereby, \R{cleanR} aids the dialogue between data analysist
+  and field experts, while also providing easy documentation of data
+  cleaning steps and data quality control. Moreover, the \R{cleanR}
+  solution changes the data cleaning process from the usual ad hoc
+  approach to a systematic, well-documented endeavor.  \R{cleanR} also
+  provides a suite of more typical \R{R} tools for interactive data
+  quality assessment and -cleaning. \hl{Hvad menes der/henvises der
+    til  med denne sætning?}
 
-  The \pkg{cleanR} package is designed to be easily extended with custom
-  user-created checks that are relevant in particular situations. \hl{Already said that above. Either expand on it or delete this.}
+  % The \pkg{cleanR} package is designed to be easily extended with
+  % custom user-created checks that are relevant in particular
+  % situations. \hl{Already said that above. Either expand on it or
+  % delete this.}
 }
@@
 \section{Introduction}
-Though data cleaning might be regarded as a somewhat dull activity, adequate data cleaning is crucial in any data analysis.
-With ever-growing dataset sizes and complexities, statisticians and data analysts find themselves spending a large portion of their time on
+Though data cleaning might be regarded as a somewhat tedious activity,
+adequate data cleaning is crucial in any data analysis. With
+ever-growing dataset sizes and complexities, statisticians and data
+analysts find themselves spending a large portion of their time on
 data cleaning and on data wrangling. \proglang{R}-packages such as
-\pkg{data.table}, and \pkg{plyr} have made data wrangling a lot easier, but there are only a few tools available for data cleaning.
-\hl{What are these tools?
-More here about the solutions already available.}
+\pkg{data.table}, and \pkg{plyr} have made data wrangling a lot
+easier, but there are only a few tools available for data
+cleaning. \hl{What are these tools? More here about the solutions
+  already available.}
 

### 2016-12-19 latex/article_vol2.tex
Andre værker

@@
   data and the experimental design, but not necessarily knowledge of
-  \R{R}. Thereby, \R{cleanR} aids the dialogue between data analysist
+  \R{R}. Thereby, \pkg{cleanR} aids the dialogue between data analysist
   and field experts, while also providing easy documentation of data
-  cleaning steps and data quality control. Moreover, the \R{cleanR}
+  cleaning steps and data quality control. Moreover, the \pkg{cleanR}
   solution changes the data cleaning process from the usual ad hoc
-  approach to a systematic, well-documented endeavor.  \R{cleanR} also
-  provides a suite of more typical \R{R} tools for interactive data
-  quality assessment and -cleaning. \hl{Hvad menes der/henvises der
+  approach to a systematic, well-documented endeavor.  \pkg{cleanR} also
+  provides a suite of more typical \proglang{R} tools for interactive data
+  quality assessment and -cleaning. \hl{Hvad henvises der
     til  med denne sætning?}
@@
 analysts find themselves spending a large portion of their time on
-data cleaning and on data wrangling. \proglang{R}-packages such as
+data cleaning and on data wrangling. Online tools such as OpenRefine
+(\url{http://openrefine.org/}) and \proglang{R}-packages such as
 \pkg{data.table}, and \pkg{plyr} have made data wrangling a lot
-easier, but there are only a few tools available for data
-cleaning. \hl{What are these tools? More here about the solutions
+easier, but only a handful of packages such as \pkg{editrules},
+\pkg{validate}, \pkg{DataCombine}, and \pkg{janitor} attempt to
+implement systematic, reproducible data cleaning.  These packages use
+different approaches for data cleaning: \pkg{editrules} and
+\pkg{validate} provide frameworks for setting up and checking
+constraints on the variables, while \pkg{DataCombine} and
+\pkg{janitor} both provide a few functions for identifying problems
+(e.g, duplicates, dates coded as numbers) in data.
+
+While some of these tools ...
+
+data cleaning is usually done in a very ad hoc manner
+
+ checking your data easy, maintainable and reproducibl
+
+\hl{What are these tools? More here about the solutions
   already available.}
@@
 human interaction since every dataset is different and the variables
-in the dataset can only be understood in the proper context of their origin. 
-\hl{Segway: Something like maybe this is why proper data cleaning is not always done...}
-In many situations these errors are discovered in the process of the
-data analysis (e.g., \hl{a categorical variable with numeric labels is
-wrongly classified as a numeric variable\footnote{\hl{Is that something we can fix in clean()?
-Not quite sure what "labels" mean in this context.}}} or a variable where all values have erroneously been coded to the same value), but in other cases a human
-with knowledge about the data context area is needed to identify
-possible mistakes in the data (e.g., if there are 4 categories for a
+in the dataset can only be understood in the proper context of their
+origin.  \hl{Segway: Something like maybe this is why proper data
+  cleaning is not always done...}  In many situations these errors are
+discovered in the process of the data analysis (e.g., \hl{a
+  categorical variable with numeric labels for each category may be 
+  wrongly classified as a quantitative variable\footnote{\hl{Is that
+      something we can fix in clean()? Not quite sure what "labels"
+      mean in this context. --- Kun i den forstand, at den skriver
+      klassen ud, så ja. Har prøvet at opdatere teksten}}} or a
+variable where all values have erroneously been coded to the same
+value), but in other cases a human with knowledge about the data
+context area is needed to identify possible mistakes in the data
+(e.g., if there are 4 categories for a
 variable that should only have 3).  \hl{segway} \\
 
-The \R{cleanR} approach to data cleaning and -quality assessment is characterized by two principles \hl{use a different term than "principles" here and below, when we refer to this sentence}.
+The \pkg{cleanR} approach to data cleaning and -quality assessment is characterized by two principles \hl{use a different term than "principles" here and below, when we refer to this sentence}.
 First of all, there is no need for data cleaning to be an ad hoc procedure.
@@
 This means that data cleaning can easily be a well-documented, well-specified procedure.
-In order to aid this principle, \R{cleanR} provides easy-to-use, automated tools for data quality assessment in \R{R} on which data cleaning decisions can be build.
+In order to aid this principle, \pkg{cleanR} provides easy-to-use, automated tools for data quality assessment in \R{R} on which data cleaning decisions can be build.
 This quality assessment is presented in a auto-generated overview document, readable by data analysts and field experts alike, thereby also contributing to a inter-field dialogue about the data at hand.
@@
 distinguishing between faulty codings of a numeric value and unusual, but correct, values requires problem-specific expertise that might not be held by the data analyst.
-Hopefully, having easy access to data descriptions through \R{cleanR} will help this necessary knowledge sharing.
-
-While \R{cleanR}s primary raison d'être is auto-generating data quality assessment overview documents, we still wish to emphasize that it is \textit{not} a tool for automatic data cleaning.
-This qualifies as the second principle of \R{cleanR}: Data cleaning decisions should always be made by humans.
-Therefore, \R{cleanR} does not supply any tools for "fixing" errors in the data.
+Hopefully, having easy access to data descriptions through \pkg{cleanR} will help this necessary knowledge sharing.
+
+While \pkg{cleanR}s primary raison d'être is auto-generating data quality assessment overview documents, we still wish to emphasize that it is \textit{not} a tool for automatic data cleaning.
+This qualifies as the second principle of \pkg{cleanR}: Data cleaning decisions should always be made by humans.
+Therefore, \pkg{cleanR} does not supply any tools for "fixing" errors in the data.
 However, we do provide interactive functions that can be used to identify potentially erroneous entries in a dataset and that can make it easier to solve data issues, one variable at a time.\\
@@
 This manuscript is structured as follows: First, in section \hl{2}, we introduce the representative of the first \hl{principle}, namely the \R{clean()} function, which generates data cleaning overview documents.
-In the \R{cleanR} package, we have provided a number of default cleaning steps that cover the data cleaning challenges, we find to be most common.
-Next, in section \hl{3}, we present the interactive mode of \R{cleanR}, as motivated by the second \hl{principle} above.
+In the \pkg{cleanR} package, we have provided a number of default cleaning steps that cover the data cleaning challenges, we find to be most common.
+Next, in section \hl{3}, we present the interactive mode of \pkg{cleanR}, as motivated by the second \hl{principle} above.
 But, as any data analyst knows, every dataset is different, and some datasets might include problems that cannot be detected by our data checking functions.
-In Section \hl{4}, we turn to the question of how such \R{cleanR} extensions can be made, such that they are integrable with the \R{clean()} function and with the other tools available in \R{cleanR}.
-At last, in Section 5 \hl{proper ref}, we discuss a number of examples of specific data cleaning challenges and how \R{cleanR} can be used to solve them.
+In Section \hl{4}, we turn to the question of how such \pkg{cleanR} extensions can be made, such that they are integrable with the \R{clean()} function and with the other tools available in \pkg{cleanR}.
+At last, in Section 5 \hl{proper ref}, we discuss a number of examples of specific data cleaning challenges and how \pkg{cleanR} can be used to solve them.
 \hl{Is there a Section 6?}\\
@@
 
-In \pkg{cleanR}, the function \R{clean()} is the primary workhorse and knowledge of nothing more than this tool is required to produce data cleaning outputs.
-The data cleaning output itself is an overview document, intended for reading by humans, in either pdf or html format.
-Appendix \hl{something} provides an example of a data cleaning output document, produced by calling \R{clean} on the dataset \R{toyData} available in  \R{cleanR}.
-The first two pages of this data cleaning output are also available in Figure~\ref{fig:example1}.
-\R{toyData} is a very small ($15$ by $6$), artificial dataset, whose only purpose is to illustrate the main capabilities of \R{cleanR}.
-The following commands load the dataset and produce the cleaning output:
+The \R{clean()} function is the primary workhorse of \pkg{cleanR} and
+this is the only function needed by the user if the standard battery
+of tests are used to output the data cleaning summaries. The data
+cleaning output itself is an overview document, intended for reading
+by humans, in either pdf or html format. Appendix \hl{something}
+provides an example of a data cleaning output document, produced by
+calling \R{clean()} on the dataset \R{toyData} available in
+\pkg{cleanR}. The first two pages of this data cleaning output are
+also available in Figure~\ref{fig:example1}. \R{toyData} is a very
+small ($15$ rows and $6$ variables), artificial dataset, whose only purpose is to
+illustrate the main capabilities of \pkg{cleanR}. The following
+commands load the dataset and produce the cleaning output:
 
@@
 
-By default, a pdf overview document is produced, saved to the disc (in the working directory) and opened for inspection.
-Turning to Figure~\ref{fig:example1}, we see that such a data cleaning output document consists of two parts.
-First, an overview of what was done is presented under the title \textit{Data cleaning summary}.
-Secondly, each variable in the dataset is presented in turn using (up to) three tools in the \textit{Variable list}: A table summarizing key features of the variable, a figure visualizing its distribution and potentially also a list of flagged issues.
-For instance, in the \R{numeric}-type variable \R{var2} from \R{toyData}, \R{clean()} has identified two values that are suspected to be miscoded missing values (\R{999} and \R{NaN}), while two values were also flagged as potential outliers that should be investigated more carefully.
-We can then return to Part 1, the data cleaning summary, and inspect what sorts of checks were performed on this variable.
-\hl{hm, delete the last sentence?
-or use different variable?
-Only these exact two checks were performed}.
-\\
+By default, a pdf overview document is produced, saved to the disc (in
+the working directory) and opened for immediate inspection. Turning to
+Figure~\ref{fig:example1}, we see that such a data cleaning output
+document consists of two parts. First, an overview of what was done is
+presented under the title \textit{Data cleaning summary}. Secondly,
+each variable in the dataset is presented in turn using (up to) three
+tools in the \textit{Variable list}: A table summarizing key features
+of the variable, a figure visualizing its distribution and potentially
+also a list of flagged issues. For instance, in the \R{numeric}-type
+variable \R{var2} from \R{toyData}, \R{clean()} has identified two
+values that are suspected to be miscoded missing values (\R{999} and
+\R{NaN}), while two values were also flagged as potential outliers
+that should be investigated more carefully. We can then return to Part
+1, the data cleaning summary, and inspect what sorts of checks were
+performed on this variable. \hl{hm, delete the last sentence? or use
+  different variable? Only these exact two checks were performed}.
 
@@
 While overview documents are great for documenting our work, they might not feel very natural for actually performing data cleaning or data wrangling based on the results we found.
-Therefore, \R{cleanR} also provides more standard \R{R} interactive tools, such as functions that print results to the console or saves it in a variable for later use.
+Therefore, \pkg{cleanR} also provides more standard \R{R} interactive tools, such as functions that print results to the console or saves it in a variable for later use.
 Or, more precisely, the functions used by \R{clean()} to do actual data summaries, visualizations and checks, as described above, can also be used interactively.
@@
 Though the discussion in the above paints a picture of cleanR as a user-friendly package which requires practically no knowledge of \R{R}, one should not be mistaken to think that it is not customizable.
-In fact, the main function of \R{cleanR}, \code{clean}, is mainly a tool for formatting the results from various checking-, summary- and visualization functions.
-Thus, the actual work underlying a \R{cleanR} output file can be anything or nothing - depending on the arguments given to \code{clean}.
+In fact, the main function of \pkg{cleanR}, \code{clean}, is mainly a tool for formatting the results from various checking-, summary- and visualization functions.
+Thus, the actual work underlying a \pkg{cleanR} output file can be anything or nothing - depending on the arguments given to \code{clean}.
 Specfically, user made functions can be added to the SVC-function arguments discussed above, e.g.
@@
 First, we describe how the \R{clean} function can be extended by adding custom prechecks, summaries, checks and visualizations.
-In order to do this, one needs to produce small functions that obey to the syntax of \R{cleanR}.
+In order to do this, one needs to produce small functions that obey to the syntax of \pkg{cleanR}.
 This can be done with different levels of strictness.
 If the custom functions are only to be used by \R{clean()}, only the input/output structures of the functions need to be restricted.
-However, by adding a few extra \hl{word?} to these functions, the full machinery of \R{cleanR} becomes available for the new, user-made functions, just as it is for the build-in functions.
+However, by adding a few extra \hl{word?} to these functions, the full machinery of \pkg{cleanR} becomes available for the new, user-made functions, just as it is for the build-in functions.
 The presentation below is given in the format of function templates, written in pseudo-code.
@@
 If \R{mySummaryFunction} is not an \R{S3} generic and \R{classes} is left unspecified, the attribute will simply be empty.
-Note that the helper function \R{allClasses()} might be useful for filling out the \R{classes} argument, as it simply lists all available classes in \R{cleanR}:
+Note that the helper function \R{allClasses()} might be useful for filling out the \R{classes} argument, as it simply lists all available classes in \pkg{cleanR}:
 \begin{Schunk}
@@
 \subsubsection{mosiacVisual - a new visualFunction}
-We will define this function such that it gets the full \R{cleanR} functionality. This can be done using the code 
+We will define this function such that it gets the full \pkg{cleanR} functionality. This can be done using the code 
 \begin{Schunk}
@@
 \end{Schunk}
-Here, we use the function \R{allClasses()} to quickly obtain a vector of all the seven variable classes addressed in \R{cleanR}.
+Here, we use the function \R{allClasses()} to quickly obtain a vector of all the seven variable classes addressed in \pkg{cleanR}.
 Note that if \R{mosaicVisual} were an S3 generic function, this argument could have been left as \R{NULL} and then the classes for which methods are available would be added automatically.
@@
 Include output in appendix, maybe.}.
-The extended \R{cleanR} output document should have the following modifications, relative to the standard \R{cleanR} output:
+The extended \pkg{cleanR} output document should have the following modifications, relative to the standard \pkg{cleanR} output:
 \begin{itemize}
@@
 \section{Something like examples}
-Finally, we will present a few examples of how to make \R{cleanR} solve specific issues related to data cleaning.
+Finally, we will present a few examples of how to make \pkg{cleanR} solve specific issues related to data cleaning.
 First, we discuss the challenges related to cleaning large datasets, particularly in terms of memory use and computation speed.
-Next, we show how \R{cleanR} can be used for problem-flagging.
-Lastly, we discuss how the \R{cleanR} output document can be included in other \R{rmarkdown} documents as a mean to produce clear and concise documentation of a dataset.
+Next, we show how \pkg{cleanR} can be used for problem-flagging.
+Lastly, we discuss how the \pkg{cleanR} output document can be included in other \R{rmarkdown} documents as a mean to produce clear and concise documentation of a dataset.
 \hl{I feel like there should be more topics here, but I'm all out of ideas...}
@@
 In this section, we give a few practical examples of ways to deal with large data, while wishing to still produce (potentially very long) data cleaning overview documents.
-Note that the interactive tools of \R{cleanR} can be used as usual or sequentially in small subsets of the large dataset, if no such overview documents are needed.
+Note that the interactive tools of \pkg{cleanR} can be used as usual or sequentially in small subsets of the large dataset, if no such overview documents are needed.
 
@@
 Though figures give a nice overview of each variable, they are also quite heavy objects in terms of memory allocation.
-Therefore, it might be beneficial to not include figures in the \R{cleanR} outputs for very large datasets.
+Therefore, it might be beneficial to not include figures in the \pkg{cleanR} outputs for very large datasets.
 This is controlled via the \R{mode} argument:
@@
 Of course, even less heavy plots might be achieved by writing new \R{visualFunction}s, using the guidelines from section \ref{sec:functionTemplates}.
-For instance, a future extension of \R{cleanR} might be the inclusion of ASCII plots, as e.g.
+For instance, a future extension of \pkg{cleanR} might be the inclusion of ASCII plots, as e.g.
 represented in the \R{R} package \R{txtplot}\hl{reference here?
@@
 \end{Schunk}
-This also deals with the fact that \R{cleanR} can produce \R{rmarkdown} files that supersedes the upper size limit of RStudio, which is currently \hl{find number} GBs (using RStudio version 1.0.44).
+This also deals with the fact that \pkg{cleanR} can produce \R{rmarkdown} files that supersedes the upper size limit of RStudio, which is currently \hl{find number} GBs (using RStudio version 1.0.44).
 \hl{Is this maybe too editor specific?
@@
 \subsection{Include cleanR document in other files}
-Sometimes, a \R{cleanR} document might be a useful addition to a more general overview document, including also e.g.
+Sometimes, a \pkg{cleanR} document might be a useful addition to a more general overview document, including also e.g.
 pairwise association plots, time series plots or exploratory analysis results.
-To this end, it is possible to produce a \R{cleanR} document that can readily be included in other \R{rmarkdown} files.
+To this end, it is possible to produce a \pkg{cleanR} document that can readily be included in other \R{rmarkdown} files.
 This is done by using the \R{standAlone} argument in \R{clean}, which removes the preamble from the outputted \R{rmarkdown} file.
@@
 
-If it is important that the embedded \R{cleanR} document can be rendered to either of these two file types, we recommend setting \R{twoCols = FALSE} and \R{output = html} in \R{clean()}, thereby essentially removing almost all output type specific formatting code from the generated \R{rmarkdown} file.
+If it is important that the embedded \pkg{cleanR} document can be rendered to either of these two file types, we recommend setting \R{twoCols = FALSE} and \R{output = html} in \R{clean()}, thereby essentially removing almost all output type specific formatting code from the generated \R{rmarkdown} file.
 

### 2016-12-20 latex/article_vol2.tex
Text wrangling

@@
   data and the experimental design, but not necessarily knowledge of
-  \R{R}. Thereby, \pkg{cleanR} aids the dialogue between data analysist
-  and field experts, while also providing easy documentation of data
-  cleaning steps and data quality control. Moreover, the \pkg{cleanR}
-  solution changes the data cleaning process from the usual ad hoc
-  approach to a systematic, well-documented endeavor.  \pkg{cleanR} also
-  provides a suite of more typical \proglang{R} tools for interactive data
-  quality assessment and -cleaning. \hl{Hvad henvises der
-    til  med denne sætning?}
+  \R{R}. Thereby, \pkg{cleanR} aids the dialogue between data
+  analysist and field experts, while also providing easy
+  documentation of reproducible data cleaning steps and data quality
+  control. Moreover, the \pkg{cleanR} solution changes the data
+  cleaning process from the usual ad hoc approach to a systematic,
+  well-documented endeavor.  \pkg{cleanR} also provides a suite of
+  more typical \proglang{R} tools for interactive data quality
+  assessment and -cleaning. \hl{Hvad henvises der til med denne
+    sætning?}
 
@@
 analysts find themselves spending a large portion of their time on
-data cleaning and on data wrangling. Online tools such as OpenRefine
-(\url{http://openrefine.org/}) and \proglang{R}-packages such as
-\pkg{data.table}, and \pkg{plyr} have made data wrangling a lot
-easier, but only a handful of packages such as \pkg{editrules},
-\pkg{validate}, \pkg{DataCombine}, and \pkg{janitor} attempt to
-implement systematic, reproducible data cleaning.  These packages use
-different approaches for data cleaning: \pkg{editrules} and
-\pkg{validate} provide frameworks for setting up and checking
-constraints on the variables, while \pkg{DataCombine} and
+data cleaning and on data wrangling. While a computer should never
+provide an unsupervised decision on what should be done to potential
+errors in the dataset, it can be an extremely useful tool in tracking
+down and flagging potential erroneous data and in providing
+information for humans to easier identify errors based on the context.
+
+Online tools such as OpenRefine (\url{http://openrefine.org/}) and
+\proglang{R}-packages such as \pkg{data.table}, and \pkg{plyr} have
+made data wrangling a lot easier, but only a handful of packages such
+as \pkg{editrules}, \pkg{validate}, \pkg{DataCombine}, and
+\pkg{janitor} attempt to implement systematic, reproducible data
+cleaning.  These packages use different approaches for data cleaning:
+\pkg{editrules} and \pkg{validate} provide frameworks for setting up
+and checking constraints on the variables, while \pkg{DataCombine} and
 \pkg{janitor} both provide a few functions for identifying problems
-(e.g, duplicates, dates coded as numbers) in data.
-
-While some of these tools ...
-
-data cleaning is usually done in a very ad hoc manner
-
- checking your data easy, maintainable and reproducibl
-
-\hl{What are these tools? More here about the solutions
-  already available.}
-
-\hl{A text here (with offset in what tools are already available) that
-\begin{itemize}
-\item Makes the point that data cleaning is usually done in a very ad hoc manner
-\item Maybe hints to the discussion later about poor documentation of data cleaning/quality assessment as the status quo for most analysts
-\item Gives a specification of what we mean by data errors/mistakes/whatever.
+(e.g, duplicates, dates coded as numbers, etc.) in data.
+
+While these tools attempt to alleviate the ubiquitous ad hoc approach
+to data cleaning they are primarily intended for the data savvy users
+and less on the general researcher with a knowledge about the specific
+field and context of the available data. The \pkg{cleanR} package
+tries to address this by providing a framework that both allows for
+extendable, systematic, reproducible data cleaning, and summarizing
+findings for researchers from other fields such that they can act as
+human experts when tracking down potential errors.
+
+% \begin{itemize}
+% \item Makes the point that data cleaning is usually done in a very ad hoc manner
+% \item Maybe hints to the discussion later about poor documentation of data cleaning/quality assessment as the status quo for most analysts
+% \item Gives a specification of what we mean by data errors/mistakes/whatever.
 I'm not completely sure this will be clear to everyone and we should probably mention it early on to not lose readers/users.
-\end{itemize}
-}
+% \end{itemize}
+% }
 
@@
 in the dataset can only be understood in the proper context of their
-origin.  \hl{Segway: Something like maybe this is why proper data
-  cleaning is not always done...}  In many situations these errors are
-discovered in the process of the data analysis (e.g., \hl{a
-  categorical variable with numeric labels for each category may be 
-  wrongly classified as a quantitative variable\footnote{\hl{Is that
-      something we can fix in clean()? Not quite sure what "labels"
-      mean in this context. --- Kun i den forstand, at den skriver
-      klassen ud, så ja. Har prøvet at opdatere teksten}}} or a
+origin. This often requires a collaborative effort between an expert
+in the field and a statistician or data scientist, which may be why
+the process of proper data cleaning is not always undertaken. In many
+situations, these errors are discovered in the process of the data
+analysis (e.g., a categorical variable with numeric labels for each
+category may be wrongly classified as a quantitative variable or a
 variable where all values have erroneously been coded to the same
@@
 context area is needed to identify possible mistakes in the data
-(e.g., if there are 4 categories for a
-variable that should only have 3).  \hl{segway} \\
-
-The \pkg{cleanR} approach to data cleaning and -quality assessment is characterized by two principles \hl{use a different term than "principles" here and below, when we refer to this sentence}.
-First of all, there is no need for data cleaning to be an ad hoc procedure.
-Often, we have a very clear idea of what flags are raisable in a given dataset before we look at it, as we were the ones to produce it in the first place.
-This means that data cleaning can easily be a well-documented, well-specified procedure.
-In order to aid this principle, \pkg{cleanR} provides easy-to-use, automated tools for data quality assessment in \R{R} on which data cleaning decisions can be build.
-This quality assessment is presented in a auto-generated overview document, readable by data analysts and field experts alike, thereby also contributing to a inter-field dialogue about the data at hand.
-Oftentimes, e.g.
-distinguishing between faulty codings of a numeric value and unusual, but correct, values requires problem-specific expertise that might not be held by the data analyst.
-Hopefully, having easy access to data descriptions through \pkg{cleanR} will help this necessary knowledge sharing.
-
-While \pkg{cleanR}s primary raison d'être is auto-generating data quality assessment overview documents, we still wish to emphasize that it is \textit{not} a tool for automatic data cleaning.
-This qualifies as the second principle of \pkg{cleanR}: Data cleaning decisions should always be made by humans.
-Therefore, \pkg{cleanR} does not supply any tools for "fixing" errors in the data.
-However, we do provide interactive functions that can be used to identify potentially erroneous entries in a dataset and that can make it easier to solve data issues, one variable at a time.\\
-
-This manuscript is structured as follows: First, in section \hl{2}, we introduce the representative of the first \hl{principle}, namely the \R{clean()} function, which generates data cleaning overview documents.
-In the \pkg{cleanR} package, we have provided a number of default cleaning steps that cover the data cleaning challenges, we find to be most common.
-Next, in section \hl{3}, we present the interactive mode of \pkg{cleanR}, as motivated by the second \hl{principle} above.
-But, as any data analyst knows, every dataset is different, and some datasets might include problems that cannot be detected by our data checking functions.
-In Section \hl{4}, we turn to the question of how such \pkg{cleanR} extensions can be made, such that they are integrable with the \R{clean()} function and with the other tools available in \pkg{cleanR}.
-At last, in Section 5 \hl{proper ref}, we discuss a number of examples of specific data cleaning challenges and how \pkg{cleanR} can be used to solve them.
-\hl{Is there a Section 6?}\\
-
-\hl{Do they typically discuss notation in articles like this?
+(e.g., if there are 4 categories for a variable that should only have
+3).
+
+The \pkg{cleanR} approach to data cleaning and -quality assessment is
+governed by two fundamental paradigms. First of all, there is no need
+for data cleaning to be an ad hoc procedure. Often, we have a very
+clear idea of what flags are raisable in a given dataset before we
+look at it, as we were the ones to produce it in the first place. This
+means that data cleaning can easily be a well-documented,
+well-specified procedure. In order to aid this paradigm, \pkg{cleanR}
+provides easy-to-use, automated tools for data quality assessment in
+\proglang{R} on which data cleaning decisions can be build. This quality
+assessment is presented in an auto-generated overview document,
+readable by data analysts and field experts alike, thereby also
+contributing to a inter-field dialogue about the data at
+hand. Oftentimes, e.g. distinguishing between faulty codings of a
+numeric value and unusual, but correct, values requires
+problem-specific expertise that might not be held by the data
+analyst. Hopefully, having easy access to data descriptions through
+\pkg{cleanR} will help this necessary knowledge sharing.
+
+While \pkg{cleanR}s primary raison d'être is auto-generating data
+quality assessment overview documents, we still wish to emphasize that
+it is \emph{not} a tool for unsupervised data cleaning. This qualifies
+as the second paradigm of \pkg{cleanR}: Data cleaning decisions
+should always be made by humans. Therefore, \pkg{cleanR} does not
+supply any tools for ``fixing'' errors in the data. However, we do
+provide interactive functions that can be used to identify potentially
+erroneous entries in a dataset and that can make it easier to solve
+data issues, one variable at a time.
+
+This manuscript is structured as follows: First, in Section
+\ref{sec:usingcleanR}, we introduce the representative of the first
+paradigm, namely the \R{clean()} function, which generates data
+cleaning overview documents. In the \pkg{cleanR} package, we have
+provided a number of default cleaning steps that cover the data
+cleaning challenges, we find to be most common. Next, in Section
+\hl{3}, we present the interactive mode of \pkg{cleanR}, as motivated
+by the second paradigm above. But, as any data analyst knows,
+every dataset is different, and some datasets might include problems
+that cannot be detected by our data checking functions. In Section
+\hl{4}, we turn to the question of how such \pkg{cleanR} extensions
+can be made, such that they are integrable with the \R{clean()}
+function and with the other tools available in \pkg{cleanR}.  At last,
+in Section 5 \hl{proper ref}, we discuss a number of examples of
+specific data cleaning challenges and how \pkg{cleanR} can be used to
+solve them.
+
+% \hl{Is there a Section 6?}
+
+%\hl{Do they typically discuss notation in articles like this?
 I try (somewhat inconsistently) to refer to functions as \R{function()} while other \R{R} objects are referred to merely as \R{object}.
@@
 
-\section{Something like "producing a data cleaning overview"} \label{sec:example1}
+\section{Creating a data cleaning overview} \label{sec:example1}\label{sec:usingcleanR}
+
+The \R{clean()} function is the primary workhorse of \pkg{cleanR} and
+this is the only function needed by the user if the standard battery
+of tests are used to output the data cleaning summaries. The data
+cleaning output itself is an overview document, intended for reading
+by humans, in either pdf or html format. Appendix \ref{sec:appendix1}
+provides an example of a data cleaning output document, produced by
+calling \R{clean()} on the dataset \R{toyData} available in
+\pkg{cleanR}. The first two pages of this data cleaning output are
+shown in Figure~\ref{fig:example1}. \R{toyData} is a very
+small ($15$ rows and $6$ variables), artificial dataset, whose only purpose is to
+illustrate the main capabilities of \pkg{cleanR}. The following
+commands load the dataset and produce the cleaning output:
 
@@
 \frame{\includegraphics[width=7.5cm,page=2]{cleanR_toyData.pdf}} 
-\frame{\includegraphics[width=7.5cm, page=3]{cleanR_toyData.pdf}}
+\frame{\includegraphics[width=7.5cm,page=3]{cleanR_toyData.pdf}}
 %\includepdf[pages={2}, pagecommand={}]{cleanR_testData.pdf}
@@
 \label{fig:example1}
-\caption{Example output from running \R{clean()} on the \R{toyData} dataset. First a summary of the full dataset,.. 
-  \hl{Maybe fix these pdfs such that they are one file with no blankspace? I.e. basically the .html-version of this file printed as pdf. This version is too small, right?}}
+\caption{Example output from running \R{clean()} on the \R{toyData}
+  dataset. First a summary of the full dataset is given and then
+  type-dependent information on each variable is given in a table and
+  a graph. Larger versions of the pages can be seen in
+  Appendix~\ref{sec:appendix1}.}
 \end{figure}
+
+\begin{Schunk}
+\begin{Sinput}
+> library(cleanR)
+> data(toyData)
+> toydata
+\end{Sinput}
+\begin{Soutput}
+   var1 var2  var3        var4 var5       var6
+1   red    1     a -0.65959383    1 Irrelevant
+2   red    1     a  0.08671649    2 Irrelevant
+3   red    1     a -0.10951326    3 Irrelevant
+4   red    2     a  0.08630221    4 Irrelevant
+5   red    2     a -1.84311184    5 Irrelevant
+6   red    6     b  0.92210680    6 Irrelevant
+7   red    6     b  1.01921086    7 Irrelevant
+8   red    6     b -0.92428326    8 Irrelevant
+9   red  999     c -0.65340163    9 Irrelevant
+10  red   NA     c  0.21133941   10 Irrelevant
+11 blue    4     c  0.91783009   11 Irrelevant
+12 blue   82     .  0.10313983   12 Irrelevant
+13 blue   NA        0.16954218   13 Irrelevant
+14 <NA>  NaN other  0.41967230   14 Irrelevant
+15 <NA>    5 OTHER  0.77143836   15 Irrelevant
+\end{Soutput}
+\begin{Sinput}
+> clean(toyData)
+\end{Sinput}
+\end{Schunk}
+
+By default, a pdf overview document is produced, saved to the disc (in
+the working directory) and opened for immediate inspection. Turning to
+Figure~\ref{fig:example1}, we see that such a data cleaning output
+document consists of two parts. First, an overview of what was done is
+presented under the title \textit{Data cleaning summary}. Secondly,
+each variable in the dataset is presented in turn using (up to) three
+tools in the \textit{Variable list}: A table summarizing key features
+of the variable, a figure visualizing its distribution and potentially
+also a list of flagged issues. For instance, in the \R{numeric}-type
+variable \R{var2} from \R{toyData}, \R{clean()} has identified two
+values that are suspected to be miscoded missing values (\R{999} and
+\R{NaN}), while two values were also flagged as potential outliers
+that should be investigated more carefully. 
+%We can then return to Part
+%1, the data cleaning summary, and inspect what sorts of checks were
+%performed on this variable. \hl{hm, delete the last sentence? or use
+%  different variable? Only these exact two checks were performed}.
+
+Though the \R{clean()} function is very easy to use, it should not be
+mistaken to be inflexible. A large number of function arguments allows
+for the cleaning overview document to be molded according to the
+user's needs. The most common arguments are summarized in Table~\ref{}
+but in order to understand this distinction, a brief glimpse of the
+inner structure of \R{clean()} might be useful. This can be obtained
+from Figure~\ref{figure:cleanStructure}.
+
+Below, we present a selection of these arguments in
+sections \hl{xx} and \hl{xxx}. We do this under two distinct headlines
+to emphasize that two levels of control are available when using
+\R{clean()}. Either, we can control the \hl{[something] - essentially,
+  this is everything but SVC-functions - } or we can control what
+quality assessments and descriptions each variable are subjected
+to.  \hl{segway to next bit or
+  just something to end the paragraph}
 
@@
 \end{figure}
-
-The \R{clean()} function is the primary workhorse of \pkg{cleanR} and
-this is the only function needed by the user if the standard battery
-of tests are used to output the data cleaning summaries. The data
-cleaning output itself is an overview document, intended for reading
-by humans, in either pdf or html format. Appendix \hl{something}
-provides an example of a data cleaning output document, produced by
-calling \R{clean()} on the dataset \R{toyData} available in
-\pkg{cleanR}. The first two pages of this data cleaning output are
-also available in Figure~\ref{fig:example1}. \R{toyData} is a very
-small ($15$ rows and $6$ variables), artificial dataset, whose only purpose is to
-illustrate the main capabilities of \pkg{cleanR}. The following
-commands load the dataset and produce the cleaning output:
-
-\begin{Schunk}
-\begin{Sinput}
-> library(cleanR)
-> data(toyData)
-> toydata
-\end{Sinput}
-\begin{Soutput}
-   var1 var2  var3        var4 var5       var6
-1   red    1     a -0.65959383    1 Irrelevant
-2   red    1     a  0.08671649    2 Irrelevant
-3   red    1     a -0.10951326    3 Irrelevant
-4   red    2     a  0.08630221    4 Irrelevant
-5   red    2     a -1.84311184    5 Irrelevant
-6   red    6     b  0.92210680    6 Irrelevant
-7   red    6     b  1.01921086    7 Irrelevant
-8   red    6     b -0.92428326    8 Irrelevant
-9   red  999     c -0.65340163    9 Irrelevant
-10  red   NA     c  0.21133941   10 Irrelevant
-11 blue    4     c  0.91783009   11 Irrelevant
-12 blue   82     .  0.10313983   12 Irrelevant
-13 blue   NA        0.16954218   13 Irrelevant
-14 <NA>  NaN other  0.41967230   14 Irrelevant
-15 <NA>    5 OTHER  0.77143836   15 Irrelevant
-\end{Soutput}
-\begin{Sinput}
-> clean(toyData)
-\end{Sinput}
-\end{Schunk}
-
-By default, a pdf overview document is produced, saved to the disc (in
-the working directory) and opened for immediate inspection. Turning to
-Figure~\ref{fig:example1}, we see that such a data cleaning output
-document consists of two parts. First, an overview of what was done is
-presented under the title \textit{Data cleaning summary}. Secondly,
-each variable in the dataset is presented in turn using (up to) three
-tools in the \textit{Variable list}: A table summarizing key features
-of the variable, a figure visualizing its distribution and potentially
-also a list of flagged issues. For instance, in the \R{numeric}-type
-variable \R{var2} from \R{toyData}, \R{clean()} has identified two
-values that are suspected to be miscoded missing values (\R{999} and
-\R{NaN}), while two values were also flagged as potential outliers
-that should be investigated more carefully. We can then return to Part
-1, the data cleaning summary, and inspect what sorts of checks were
-performed on this variable. \hl{hm, delete the last sentence? or use
-  different variable? Only these exact two checks were performed}.
-
-Though the \R{clean()} function is very easy to use, it should not be mistaken to be inflexible.
-A large number of function arguments allows for the cleaning overview document to be molded as the user sees best.
-Below, we present a selection of these arguments in sections \hl{xx} and \hl{xxx}.
-We do this under two distinct headlines to emphasize that two levels of control are available when using \R{clean()}.
-Either, we can control the \hl{[something] - essentially, this is everything but SVC-functions - } or we can control what quality assessments and descriptions each variable are subjected to.
-In order to understand this distinction, a brief glimpse of the inner structure of \R{clean()} might be useful.
-This can be obtained from Figure~\ref{figure:cleanStructure}.
-\hl{segway to next bit or just something to end the paragraph}
 
@@
 
-\section{Appendix Something}
+\section{Appendix Something} \label{sec:appendix1}
 \includepdf[pages=2-4, pagecommand={}, frame=true, noautoscale=true, scale=0.7]{cleanR_toyData.pdf}

### 2016-12-20 latex/article_vol2.tex
Fixed typo

--- 2016-12-18 latex/article_vol2.tex
@@
   data and the experimental design, but not necessarily knowledge of
-  \R{R}. Thereby, \R{cleanR} aids the dialogue between data analysist
+  \R{R}. Thereby, \R{cleanR} aids the dialogue between data analysists
   and field experts, while also providing easy documentation of data

### 2016-12-20 latex/article_vol2.tex

--- 2016-12-20 latex/article_vol2.tex
@@
   data and the experimental design, but not necessarily knowledge of
+
   \R{R}. Thereby, \pkg{cleanR} aids the dialogue between data
-  analysist and field experts, while also providing easy
+  analysists and field experts, while also providing easy
   documentation of reproducible data cleaning steps and data quality

### 2016-12-21 latex/article_vol2.tex
Fixed data

@@
 Online tools such as OpenRefine (\url{http://openrefine.org/}) and
-\proglang{R}-packages such as \pkg{data.table}, and \pkg{plyr} have
+\proglang{R}-packages such as \pkg{plyr}, and \pkg{data.table} have
 made data wrangling a lot easier, but only a handful of packages such
@@
 
-\begin{figure}
+\begin{figure}[tb]
 \begin{center}
@@
 for the cleaning overview document to be molded according to the
-user's needs. The most common arguments are summarized in Table~\ref{}
-but in order to understand this distinction, a brief glimpse of the
-inner structure of \R{clean()} might be useful. This can be obtained
-from Figure~\ref{figure:cleanStructure}.
-
-Below, we present a selection of these arguments in
-sections \hl{xx} and \hl{xxx}. We do this under two distinct headlines
-to emphasize that two levels of control are available when using
-\R{clean()}. Either, we can control the \hl{[something] - essentially,
-  this is everything but SVC-functions - } or we can control what
-quality assessments and descriptions each variable are subjected
-to.  \hl{segway to next bit or
-  just something to end the paragraph}
-
-\begin{figure}
+user's needs.
+
+The most commonly used arguments are summarized in
+Table~\ref{table.cleanFormals} and they are grouped according to the
+part of the cleaning process they influence. In order to understand
+this distinction, a glimpse of the inner structure of \R{clean()} is
+shown in Figure~\ref{figure:cleanStructure}. In the following we give
+examples on how to use these parameters to influence the output of
+\R{clean() }.
+
+\begin{table}
+\small
+\begin{tabular}{p{0.25\linewidth}p{0.45\linewidth}p{0.2\linewidth}}
+\hline
+Argument & Description & Default value \\
+\hline
+
+\smallskip Control input variable and summary\\
+\quad \R{useVar} & What variables should be used? & \R{NULL} (corresponding to all variables) \\
+\quad \R{ordering} & Ordering of the variables in the data summary (as is or alphabetical) & \R{"asIs"} \\
+\quad \R{onlyProblematic} & Should only variables flagged as problematic be included in the \textit{Variable list}? & \R{FALSE} \\
+\quad \R{listChecks} & Should an overview of what checks were performed by listed in the \textit{Data cleaning summary}? &  \R{TRUE} \\
+
+\smallskip Control summarize, visualize, and check steps \\
+\quad \R{mode} & What steps should be performed for each variable (out of the three possibilities \textit{summarize}, \textit{visualize}, \textit{check})? & \R{c("summarize", "visualize", "check")} \\
+\quad \R{labelled\_as} & How should variables of class \R{labelled} be handled (as factors, is missing values or by ignoring labels)? & \R{"factor"} \\
+\quad \R{smartNum} & Should numerical values with only a few unique
+                     levels be flagged and treated as a factor variable? & \R{TRUE} \\
+\quad \R{maxProbVals} & Maximum number of problematic values to print, if any are found in data checks & \R{Inf} \\
+\quad \R{maxDecimals} & Maximum number of decimals to print for numeric values in the variable list & \R{2} \\
+\quad \R{twoCol} & Should the summary table and visualizations be placed side-by-side (in two columns)? & \R{TRUE} \\
+
+\smallskip Control output and post-processing \\
+\quad \R{output} & Type of output file to be produced (html, or pdf) & \R{"pdf"} \\
+\quad \R{render} & Should the output file be rendered from markdown? & \R{TRUE} \\
+\quad \R{openResult} & If a  pdf/html file is rendered, should it automatically open afterwards, and if not, should the \R{rmarkdown} file? & \R{TRUE} \\
+
+\hline
+\end{tabular}
+\caption{A selection of commonly used arguments to \R{clean()} separated into the parts they control.}
+\label{table.cleanFormals}
+\end{table}
+
+% Below, we present a selection of these arguments in
+% sections \hl{xx} and \hl{xxx}. We do this under two distinct headlines
+% to emphasize that two levels of control are available when using
+% \R{clean()}. Either, we can control the \hl{[something] - essentially,
+%   this is everything but SVC-functions - } or we can control what
+% quality assessments and descriptions each variable are subjected
+% to.  \hl{segway to next bit or
+%   just something to end the paragraph}
+
+\begin{figure}[tb]
 \begin{description}
@@
 
-\subsection{Controlling [something]}
-\label{subsection:controlSomething}
-
-\begin{table}
-\small
-\begin{tabular}{p{0.25\linewidth}p{0.45\linewidth}p{0.2\linewidth}}
-\hline
-Argument & Description & Default value \\
-\hline
-
-\smallskip Control \hl{...eh?}\\
-\quad \R{useVar} & What variables should be used? & \R{NULL} (corresponding to all variables) \\
-\quad \R{ordering} & Ordering of the variables in the data summary (as is or alphabetical) & \R{"asIs"} \\
-\quad \R{onlyProblematic} & Should only variables flagged as problematic be included in the \textit{Variable list}? & \R{FALSE} \\
-\quad \R{listChecks} & Should an overview of what checks were performed by listed in the \textit{Data cleaning summary}? &  \R{TRUE} \\
-
-\smallskip Control SVC-step \\
-\quad \R{mode} & What steps should be performed for each variable (out of the three possibilities \textit{summarize}, \textit{visualize}, \textit{check})? & \R{c("summarize", "visualize", "check")} \\
-\quad \R{labelled\_as} & How should variables of class \R{labelled} be handled (as factors, is missing values or by ignoring labels)? & \R{"factor"} \\
-\quad \R{smartNum} & Should numerical values with only a few unique levels be flagged an treated as factor variables? & \R{TRUE} \\
-\quad \R{maxProbVals} & Maximum number of problematic values to print, if any are found in data checks & \R{Inf} \\
-\quad \R{maxDecimals} & Maximum number of decimals to print for numeric values in the variable list & \R{2} \\
-\quad \R{twoCol} & Should the summary table and visualizations be placed side-by-side (in two columns)? & \R{TRUE} \\
-
-\smallskip Control output \\
-\quad \R{output} & Type of output file to be produced (html, pdf or \hl{?? what does screen mean in the documentation??}) & \R{"pdf"} \\
-\quad \R{render} & Should the output file be rendered from markdown? & \R{TRUE} \\
-\quad \R{openResult} & If a  pdf/html file is rendered, should it automatically open afterwards, and if not, should the \R{rmarkdown} file? & \R{TRUE} \\
-
-\hline
-\end{tabular}
-\caption{A selection of commonly used\hl{/useful?} arguments of \R{clean()}.
-\hl{Note: I have everything here except \R{file} and \R{quiet}, replace, vol, standAlone + stuff related to selecting SVC-functions (dealt with in next subsection)}.
-\hl{Fix formatting for \R{onlyProblematic}.}}
-\label{table.cleanFormals}
-\end{table}
-
-\hl{Describe arguments in Table \ref{table.cleanFormals}. Maybe do a few examples of interesting argument combinations?}
-
-\subsection{Something about what check, visual and summary functions are available}
+% \subsection{Controlling [something]}
+% \label{subsection:controlSomething}
+
+For example, to get a summary document that only contains the
+variables with potential problems, and with a limit of maximum 10
+printed potential errors, we can write (output not shown)
+
+\begin{Schunk}
+\begin{Sinput}
+> clean(toyData, onlyProblematic=TRUE, maxProbVals=10)
+\end{Sinput}
+\end{Schunk}
+
+The final rendering of the generated
+
+By default, \R{clean()} runs all the summary,
+visualization, and check functions that are
+implemented. Table~\ref{table.SVCfunctions} lists these functions but
+we can also use the \R{allSummaryFunctions()},
+\R{allVisualFunctions()}, and \R{allCheckFunctions()} function in
+\proglang{R} to obtain a list. For example the implemented summary
+functions are:
+
+%\subsection{Something about what check, visual and summary functions are available}
 
@@
 
+\begin{Schunk}
+\begin{Sinput}
+> allSummaryFunctions()
+\end{Sinput}
+\begin{Soutput}
+----------------------------------------------------------------------------
+name           description                     classes                      
+-------------- ------------------------------- -----------------------------
+centralValue   Compute median or mode          character, Date,
+factor,     
+                                               integer, labelled,
+                                               logical,  
+                                               numeric
+
+countMissing   Compute ratio of missing        character, Date,
+factor,     
+               observations                    integer, labelled,
+               logical,  
+                                               numeric
+
+minMax         Find minimum and maximum        integer, numeric, Date       
+               values
+
+quartiles      Compute 1st and 3rd quartiles   integer, numeric
+
+uniqueValues   Count number of unique values   character, Date,
+factor,     
+                                               integer, labelled,
+                                               logical,  
+                                               numeric
+
+variableType   Data class of variable          character, Date,
+factor,     
+                                               integer, labelled,
+                                               logical,  
+                                               numeric                      
+----------------------------------------------------------------------------
+\end{Soutput}
+\end{Schunk}
+
 In this section, we discuss how to control every step in \R{clean()} that actually involves a function being called on variables from the dataset.
@@
 
-\section{Something about the interactive cleanR mode}
-While overview documents are great for documenting our work, they might not feel very natural for actually performing data cleaning or data wrangling based on the results we found.
-Therefore, \pkg{cleanR} also provides more standard \R{R} interactive tools, such as functions that print results to the console or saves it in a variable for later use.
-Or, more precisely, the functions used by \R{clean()} to do actual data summaries, visualizations and checks, as described above, can also be used interactively.
-\hl{segway}
-
-\hl{Maybe write here explicitly that this section introduces the three functions: \R{summarize()}, \R{check()} and \R{visualize()}?}
-
-\subsection{Cleaning by hand: An example}
-Let's say we wish to look further into a certain variable from \R{toyData}, namely \R{var2}.
-The data cleaning summary found some issues in this variable, and we would like to recall what these issues were.
-This can be done using the command
+\section{Using cleanR interactively}
+While overview documents are great for presenting and documenting the
+data cleaning checks, it may be natural to work more interactively
+through the data cleaning process. \pkg{cleanR} also
+provides more standard \proglang{R} interactive tools, such as
+functions that print results to the console or returns the information
+as an object for later use. This section describes how to use the
+functions \R{check()}, \R{summarize()} and \R{visualize()} to work
+interactively with \pkg{cleanR}.
+
+\subsection{Data cleaning by hand: An example}
+Let's say we wish to look further into a certain variable from
+\R{toyData}, namely \R{var2}. The data cleaning summary found some
+issues in this variable, and we would like to recall what these issues
+were. This can be done using the \R{check()} command
 \begin{Schunk} 
@@
 $identifyOutliers
-Note that the following possible outlier values were detected: 82, 999. $
+Note that the following possible outlier values were detected: 82, 999. 
 \end{Soutput}
 \end{Schunk}
-\hl{delete that last \$, it's only there to make my TeX editor stop highlighting everything. Note: \$s cannot be escaped here, as Schunk is verbatim.}
-
-Note that the arguments specifying which checks to perform, as described in \hl{XXXX} above, are in fact passed to \R{check()}, and thus they can also be used here.
-For instance, if we only want the result of the check for miscoded missing values, we write
+% $
+
+Note that the arguments specifying which checks to perform, as
+described in the previous section, are in fact passed to \R{check()}, and
+thus they can also be used here. For instance, if we only want the
+result of the check for miscoded missing values, we write
 \begin{Schunk}
@@
 \end{Schunk}
-An equivalent way to call only a single, specific \R{checkFunction} is by using it directly on the variable, i.e. 
+The \R{numericChecks} argument XXX.. ... not defined
+
+An equivalent way to call only a single, specific \R{checkFunction}
+such as \R{identifyMissing} (see \ref{table.SVCfunctions} for a list
+of check functions) is by using it directly on the variable, i.e.
 \begin{Schunk}
@@
 \begin{Soutput}
-The following suspected missing value codes enter as regular values: 999, NaN. $
+The following suspected missing value codes enter as regular values: 999, NaN. 
 \end{Soutput}
 \end{Schunk}
-\hl{delete that last \$}
-
-The result of a \R{checkFunction} is an object of class \R{checkResult}. By using the structure function, \R{str()}, we can look further into its components:
+% $
+
+The result of a \R{checkFunction} is an object of class
+\R{checkResult}. By using the structure function, \R{str()}, we can
+look further into its components:
 
@@
 \end{Schunk}
-The most important thing to note here is that while the printed message is made for easy reading, the actual values of the variable causing the issue are still obtainable.
-If we for instance decide that the values \R{999} and \R{NaN} in \R{var2} are in fact miscoded missing values, we can easily replace them with \R{NA}s:
+The most important thing to note here is that while the printed
+message is made for easy reading, the actual values of the variable
+causing the issue are still obtainable in the element
+\R{problemValues}. If we for instance decide that the values \R{999}
+and \R{NaN} in \R{var2} are in fact miscoded missing values, we can
+easily replace them with \R{NA}s:
 \begin{Schunk}
@@
 \begin{Soutput}
-No problems found. $
+No problems found.
 \end{Soutput}
 \end{Schunk}
-\hl{remove that last \$}.
-\hl{This is quite a practical example, but it essentially also gives the basic structure for misusing cleanR by generating a loop that does autocleaning...
-What do you think?
-Perhaps those who know how to make loops also know that they shouldn't do automatic data cleaning?}
+% $
 
@@
 \begin{table}[ht]
-\small
+\scriptsize
 \begin{tabular}{p{0.2\linewidth}p{0.2\linewidth}p{0.2\linewidth}p{0.2\linewidth}}

### 2016-12-21 R/cleanR-package.R

@@ -2,44 +2,42 @@
 #'
 #' A dataset of constructed data used to as test bed when identifying potential errors in a data frame.
 #'
-#' @name testData
-#' @docType data
 #' @format A data frame with 15 rows and 14 variables.
 #' \describe{
-#'    \item{charVar}{a simple factor}
-#'    \item{factorVar}{a factor with a "special" level 999}
-#'    \item{numVar}{a numeric vector}
-#'    \item{intVar}{an integer vector}
-#'    \item{boolVar}{a logical vector}
-#'    \item{keyVar}{a factor}
-#'    \item{emptyVar}{a numeric vector where all the elements are identical}
-#'    \item{_joeVar}{a numeric vector}
-#'    \item{jack__var}{a numeric vector}
-#'    \item{numOutlierVar}{a numeric vector with a possible outlier}
-#'    \item{smartNumVar}{a numeric vector}
-#'    \item{cprVar}{a factor with levels in the format of the Danish CPR number (Social security number)}
-#'    \item{cprKeyVar}{a factor with levels in the format of the Danish CPR number}
-#'    \item{miscodedMissingVar}{a factor with levels corresponding to various misssing codes}
+#'    \item{charVar}{A character vector with a single missing observation.}
+#'    \item{factorVar}{A factor vector with a miscoded missing observation, \code{999}.}
+#'    \item{numVar}{A numeric vector}
+#'    \item{intVar}{An integer vector}
+#'    \item{boolVar}{A logical vector with three missing observations.}
+#'    \item{keyVar}{A character vector with unique codes for each observation.}
+#'    \item{emptyVar}{A numeric vector where all entries are identical.}
+#'    \item{numOutlierVar}{A numeric vector with a possible outlier (\code{100}).}
+#'    \item{smartNumVar}{A numeric vector that takes only two different values.}
+#'    \item{cprVar}{A character vector with levels in the format of Danish CPR numbers
+#'    (social security numbers).}
+#'    \item{cprKeyVar}{A character vector with levels in the format of Danish CPR numbers
+#'    (social security numbers) with unique levels for each observation.}
+#'    \item{miscodedMissingVar}{A character vector with levels corresponding to
+#'    various miscoded (non-\code{NA}) misssing codes.}
+#'    \item{misclassifiedNumVar}{A misclassified factor variable, where every level
+#'    is a number and a many (12) different levels are in use.}
[…]

### 2016-12-21 latex/cleanR_toydata.Rmd

---
cleanR: yes
title: toyData
subtitle: "Autogenerated data summary from cleanR"
date: "`r Sys.Date()`"
output: pdf_document
documentclass: report
header-includes:
  - \renewcommand{\chaptername}{Part}
  - \newcommand{\fullline}{\noindent\makebox[\linewidth]{\rule{\textwidth}{0.4pt}}}
  - \newcommand{\bminione}{\begin{minipage}{0.75 \textwidth}}
  - \newcommand{\bminitwo}{\begin{minipage}{0.25 \textwidth}}
  - \newcommand{\emini}{\end{minipage}}
---
```{r , echo=FALSE, include=FALSE, warning=FALSE, message=FALSE, error=FALSE }
library(ggplot2)
library(pander)
```
# Data cleaning summary
The dataset examined has the following dimensions:

------------------------------
Feature                 Result
--------------------- --------
Number of rows              15

Number of variables          6
------------------------------

### Checks performed
The following variable checks were performed, depending on the data type of each variable:

------------------------------------------------------------------------------------------------------------------------------
&nbsp;                                                 character    factor    labelled   numeric    integer    logical   Date 
----------------------------------------------------- ----------- ---------- ---------- ---------- ---------- --------- ------
Identify miscoded missing values                       $\times$    $\times$   $\times$   $\times$   $\times$

Identify prefixed and suffixed whitespace              $\times$    $\times$   $\times$

Identify levels with < 6 obs.                          $\times$    $\times$

Identify case issues                                   $\times$    $\times$

Identify misclassified numeric or integer variables    $\times$    $\times$

Identify outliers                                                                        $\times$   $\times$                  
------------------------------------------------------------------------------------------------------------------------------

Please note that all numerical values in the following have been rounded to 2 decimals.

# Variable list
## **var1**

\bminione

---------------------------------------
Feature                          Result
------------------------- -------------
Variable type                    factor

Number of missing obs.      2 (13.33 %)

Number of unique values               2

Mode                              "red"
---------------------------------------

\emini
\bminitwo
```{r Var-1-var1 , echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE }
qplot(x = structure(c(2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 
1L, 1L, 1L), .Label = c("blue", "red"), class = "factor", na.action = structure(14:15, class = "omit")), 
    geom = "bar", xlab = "", main = "var1")
```
\emini

- Note that the following levels have at most five observations: \"blue\".

\fullline

## **var2**

\bminione

------------------------------------
Feature                       Result
------------------------- ----------
Variable type                numeric

Number of missing obs.      3 (20 %)

Number of unique values            8

Median                           4.5

1st and 3rd quartiles        1.75; 6

Min. and max.                 1; 999
------------------------------------

\emini
\bminitwo
```{r Var-2-var2 , echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE }
qplot(x = c(1, 1, 1, 2, 2, 6, 6, 6, 999, 4, 82, 5), geom = "histogram", 
    xlab = "", main = "var2", bins = 20)
```
\emini

- The following suspected missing value codes enter as regular values: \"999\", \"NaN\".

- Note that the following possible outlier values were detected: \"82\", \"999\".

\fullline

## **var3**

\bminione

-----------------------------------
Feature                      Result
------------------------- ---------
Variable type                factor

Number of missing obs.      0 (0 %)

Number of unique values           7

Mode                            "a"
-----------------------------------

\emini
\bminitwo
```{r Var-3-var3 , echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE }
qplot(x = structure(c(3L, 3L, 3L, 3L, 3L, 4L, 4L, 4L, 5L, 5L, 
5L, 2L, 1L, 6L, 7L), .Label = c(" ", ".", "a", "b", "c", "other", 
"OTHER"), class = "factor"), geom = "bar", xlab = "", main = "var3")
```
\emini

- The following suspected missing value codes enter as regular values: \" \", \".\".

- The following values appear with prefixed or suffixed white space: \" \".

- Note that the following levels have at most five observations: \" \", \".\", \"a\", \"b\", \"c\", \"other\", \"OTHER\".

- Note that there might be case problems with the following levels: \"other\", \"OTHER\".

\fullline

## **var4**

\bminione

---------------------------------------
Feature                          Result
------------------------- -------------
Variable type                   numeric

Number of missing obs.          0 (0 %)

Number of unique values              15

Median                             0.33

1st and 3rd quartiles       -0.62; 0.66

Min. and max.                -2.21; 1.6
---------------------------------------

\emini
\bminitwo
```{r Var-4-var4 , echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE }
qplot(x = c(-0.626453810742332, 0.183643324222082, -0.835628612410047, 
1.59528080213779, 0.329507771815361, -0.820468384118015, 0.487429052428485, 
0.738324705129217, 0.575781351653492, -0.305388387156356, 1.51178116845085, 
0.389843236411431, -0.621240580541804, -2.2146998871775, 1.12493091814311
), geom = "histogram", xlab = "", main = "var4", bins = 20)
```
\emini

- Note that the following possible outlier values were detected: \"1.12\", \"1.51\", \"1.6\".

\fullline

## **var5**

* The variable is a key (distinct values for each observation).

\fullline

## **var6**

* The variable only takes one (non-missing) value: \"Irrelevant\". The variable contains 0 \% missing observations.

\fullline

This report was created by cleanR v0.5.1.

### 2016-12-28 latex/article_vol2.tex
First bunch of commits

@@
 \usepackage{xspace}
+\usepackage{array}
 \usepackage{tikz}
@@
 \newcommand{\R}[1]{\code{#1}}
+
+\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
 
@@
 
-By default, a pdf overview document is produced, saved to the disc (in
-the working directory) and opened for immediate inspection. Turning to
-Figure~\ref{fig:example1}, we see that such a data cleaning output
-document consists of two parts. First, an overview of what was done is
-presented under the title \textit{Data cleaning summary}. Secondly,
-each variable in the dataset is presented in turn using (up to) three
-tools in the \textit{Variable list}: A table summarizing key features
-of the variable, a figure visualizing its distribution and potentially
-also a list of flagged issues. For instance, in the \R{numeric}-type
-variable \R{var2} from \R{toyData}, \R{clean()} has identified two
-values that are suspected to be miscoded missing values (\R{999} and
-\R{NaN}), while two values were also flagged as potential outliers
-that should be investigated more carefully. 
+By default, an \proglang{R} markdown file and rendered pdf overview
+document is produced, saved to the disc (in the working directory) and
+opened for immediate inspection. Turning to Figure~\ref{fig:example1},
+we see that such a data cleaning output document consists of two
+parts. First, an overview of what was done is presented under the
+title \textit{Data cleaning summary}. Secondly, each variable in the
+dataset is presented in turn using (up to) three tools in the
+\textit{Variable list}: A table summarizing key features of the
+variable, a figure visualizing its distribution and potentially also a
+list of flagged issues. For instance, in the \R{numeric}-type variable
+\R{var2} from \R{toyData}, \R{clean()} has identified two values that
+are suspected to be miscoded missing values (\R{999} and \R{NaN}),
+while two values were also flagged as potential outliers that should
+be investigated more carefully.
 %We can then return to Part
@@
 \quad \R{render} & Should the output file be rendered from markdown? & \R{TRUE} \\
-\quad \R{openResult} & If a  pdf/html file is rendered, should it automatically open afterwards, and if not, should the \R{rmarkdown} file? & \R{TRUE} \\
+\quad \R{openResult} & If a  pdf/html file is rendered, should it
+                       automatically open afterwards, and if not,
+                       should the \R{rmarkdown} file be opened? & \R{TRUE} \\
+\quad \R{replace} & Overwrite an existing markdown file with the same name? & \R{FALSE} \\
 
@@
 \begin{figure}[tb]
-\begin{description}
-\item[Input] A dataset
-	\begin{itemize} 
-		\item This should be of type \R{data.frame} or \R{tibble} \hl{or data.table?}
-	\end{itemize}
-\item[Create contents] For each variable in the dataset, do the following:
-	\begin{description}
-		\item[Stage 1:] Pre-checks 
-			\begin{itemize}
-				\item Is the variable suitable for summarization, visualization and checks?
-				\begin{description}
-					\item[Yes:] Go to stage 2. 
-					\item[No:] Move on to the next variable.
-				\end{description}
-			\end{itemize}
-		\item[Stage 2:] SVC-steps
-			\begin{description}
-				\item[Summarize] Call \R{summarize()} to produce a summary table describing the variable. What features enter this table depends on the data class of the variable.
-				\item[Visualize]  Call \R{visualize()} to produce a plot visualizing the distribution of the variable. 
-				\item[Check]  Call \R{check()} to apply quality- and error checks to the variable. What checks are used depends on the data class of the variable.
-			\end{description}
-	\end{description}
-\item[Output] Files for a overview document are saved to the disc and possibly also opened.
-	\begin{itemize}
-		\item Always a \R{rmarkdown} (.Rmd) file
-		\item Possible also a html or pdf file
-	\end{itemize}
-\end{description}
-\caption{\hl{quite odd having a figure like this? maybe do a flowchart?}}
+% \begin{description}
+% \item[Input] A dataset
+% 	\begin{itemize} 
+% 		\item This should be of type \R{data.frame} or \R{tibble} \hl{or data.table?}
+% 	\end{itemize}
+% \item[Create contents] For each variable in the dataset, do the following:
+% 	\begin{description}
+% 		\item[Stage 1:] Pre-checks 
+% 			\begin{itemize}
+% 				\item Is the variable suitable for summarization, visualization and checks?
+% 				\begin{description}
+% 					\item[Yes:] Go to stage 2. 
+% 					\item[No:] Move on to the next variable.
+% 				\end{description}
+% 			\end{itemize}
+% 		\item[Stage 2:] SVC-steps
+% 			\begin{description}
+% 				\item[Summarize] Call \R{summarize()} to produce a summary table describing the variable. What features enter this table depends on the data class of the variable.
+% 				\item[Visualize]  Call \R{visualize()} to produce a plot visualizing the distribution of the variable. 
+% 				\item[Check]  Call \R{check()} to apply quality- and error checks to the variable. What checks are used depends on the data class of the variable.
+% 			\end{description}
+% 	\end{description}
+% \item[Output] Files for a overview document are saved to the disc and possibly also opened.
+% 	\begin{itemize}
+% 		\item Always a \R{rmarkdown} (.Rmd) file
+% 		\item Possible also a html or pdf file
+% 	\end{itemize}
+% \end{description}
+
+% Define block styles
+\tikzstyle{decision} = [diamond, draw, fill=blue!20, 
+    text width=5.5em, text badly centered, node distance=3cm, inner
+    sep=0pt]
+\tikzstyle{block} = [rectangle, draw, fill=blue!20, 
+    text width=6em, text centered, rounded corners, minimum
+    height=4em]
+\tikzstyle{line} = [draw, -latex']
+\tikzstyle{cloud} = [draw, ellipse,fill=red!20, node distance=3cm,
+    text width=5em,
+    minimum height=2em]
+
+\begin{center}    
+\begin{tikzpicture}[node distance = 2cm, auto,thick,scale=0.75, every node/.style={transform shape}]
+    % Place nodes
+    \node [block] (init) {Get next variable and run pre checks};
+    \node [cloud, above of=init] (input) {Input \texttt{data.frame} or \texttt{tibble}};
+%    \node [cloud, right of=init] (system) {system};
+    \node [decision, right of=init, node distance=4cm] (precheck) {Is variable suitable for inclusion};
+    \node [block, right of=precheck, node distance=4cm] (summarize)
+    {Run \texttt{summarize()} to produce summary table};
+    \node [block, below of=summarize, node distance=3cm] (visualize)
+    {Run \texttt{visualize()} to plot variable};
+    \node [block, below of=visualize, node distance=2.5cm] (check)
+    {Call \texttt{check()} to run error checks};
+    \node [decision, below of=check, node distance=2.7cm] (done) {More
+      variables?};
+    \node [block, right of=done, node distance=3.5cm] (stop) {Write
+      \proglang{R} markdown file};
+    \node [cloud, below of=stop, node distance=3.5cm] (render) {Render
+      markdown and possiby open};
+    % Draw edges
+    \path [line] (summarize) -- (visualize);
+    \path [line] (visualize) -- (check);
+    \path [line] (check) -- (done);
+    \path [line] (done.south) -- +(0,-10pt) -| node [near start] {yes} (init);
+%    \path [line] (identify) -- (evaluate);
+%    \path [line] (evaluate) -- (decide);
+%    \path [line] (init) -| node [near start] {yes} (precheck);
+    \path [line] (init) -- (precheck);
+    \path [line] (precheck) -- node [near start] {yes} (summarize);
+    \path [line] (precheck) |- node [near start] {no} (done);
+    \path [line] (done) -- node [near start] {no} (stop);
+%    \path [line] (update) |- (identify);
+%    \path [line] (decide) -- node {no}(stop);
+    \path [line,dashed] (input) -- (init);
+%    \path [line,dashed] (system) -- (init);
+%    \path [line,dashed] (system) |- (evaluate);
+\end{tikzpicture}
+\end{center}
+
+\caption{Overall illustration of the stages undertaken when running
+  \R{clean()}. Each variable is checked for eligibility before
+  running \R{summarize()}, \R{visualize()}, and \R{check()}, and the
+  resulting \proglang{R} markdown file may be rendered and opened.}
 \label{figure:cleanStructure}
@@
 
-For example, to get a summary document that only contains the
-variables with potential problems, and with a limit of maximum 10
-printed potential errors, we can write (output not shown)
-
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, onlyProblematic=TRUE, maxProbVals=10)
-\end{Sinput}
-\end{Schunk}
-
-The final rendering of the generated
-
-By default, \R{clean()} runs all the summary,
-visualization, and check functions that are
-implemented. Table~\ref{table.SVCfunctions} lists these functions but
-we can also use the \R{allSummaryFunctions()},
-\R{allVisualFunctions()}, and \R{allCheckFunctions()} function in
-\proglang{R} to obtain a list. For example the implemented summary
-functions are:
-
-%\subsection{Something about what check, visual and summary functions are available}
+For example, to get a summary document in html format that only
+contains the variables with potential problems, and with a limit of
+maximum 10 printed potential errors for each variable we can write
+(output not shown). Also, we can add the argument \R{replace=TRUE} in
+order to force \R{clean()} to overwrite any existing files produced by
+\R{clean()}.
+
+\begin{Schunk}
+\begin{Sinput}
+> clean(toyData, output="html", onlyProblematic=TRUE, maxProbVals=10,
+        replace=TRUE)
+\end{Sinput}
+\end{Schunk}
+
+The final rendering of the generated markdown file is controlled with
+the \R{render} and \R{openResult} arguments which both default to
+\R{TRUE} \hl{Bliver det ved med at være sandt?}. \R{render} determines
+if the \proglang{R} markdown file produced should be rendered using
+the \pkg{rmarkdown} package and \R{openResult} decides whether the
+output html or pdf file should be opened. The following command
+produces an \proglang{R} markdown file containing the data cleaning
+process information but the markdown file is not rendered into html
+and it is not opened.
+
+\begin{Schunk}
+\begin{Sinput}
+> clean(toyData, output="html", render=FALSE, openResult=FALSE)
+\end{Sinput}
+\end{Schunk}
 
@@
   \quad \R{minMax} & Find minimum and maximum values &   &  & $\times$ & &  & $\times$ & $\times$  \\ 
-  \quad \R{quartiles} & Compute 1st and 3rd quartiles &    &  & $\times$ & &  & $\times$ &  \\ 
+  \quad \R{quartiles} & Compute 1st and 3rd quartiles &    &  & $\times$ & &  & $\times$ & $\times$ \\ 
   \quad \R{uniqueValue} & Count number of unique values &   $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\ 
@@
  \textbf{\R{visualFunction}s} \smallskip \\
-  \quad \R{basicVisual} & Histograms and barplots using graphics &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\ 
+  \quad \R{basicVisual} & Histograms and barplots using base \proglang{R} graphics &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\ 
   \quad \R{standardVisual} & Histograms and barplots using ggplot2 &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\ 
@@
  \quad \R{identifyNums} & Identify misclassified numeric or integer variables & $\times$ & $\times$ & & & & &  \\ 
- \quad \R{identifyOutliers} & Identify outliers &  & & $\times$ & & $\times$ & \\ 
- \quad \R{identifyOutliersTBStyle} & Identify outliers (Turkish Boxplot style) &  & & $\times$ & & $\times$ &  \\ 
+ \quad \R{identifyOutliers} & Identify outliers &  & & $\times$ & & $\times$ & $\times$ \\ 
+ \quad \R{identifyOutliersTBStyle} & Identify outliers (Turkish Boxplot style) &  & & $\times$ & & $\times$ & $\times$ \\ 
  \quad \R{identifyWhitespace} & Identify prefixed and suffixed whitespace &  $\times$ & $\times$ & & $\times$ & & &  \\ 
- \quad \R{isCPR} & Identify Danish CPR numbers & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &   \\ 
- \quad \R{isEmpty} & Check if the variable contains only a single value & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &   \\ 
- \quad \R{isKey} & Check if the variable is a key & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & \smallskip   \\ 
+ \quad \R{isCPR} & Identify Danish CPR numbers & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &$\times$   \\ 
+ \quad \R{isEmpty} & Check if the variable contains only a single value & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\ 
+ \quad \R{isKey} & Check if the variable is a key & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$\smallskip   \\ 
  \hline
 \end{tabular}
-\caption{\hl{Blabla, mention that C is character, F is factor, I is integer, L is labelled, B is logical (boolean), N is numeric and D is Date.
-\newline Also: Check that everything in here is correct (i.e.
-corresponds to the output of allSummaryFunctions() etc when makeXfunction has been fixed.}}
+\caption{List of all summary functions (used in the summary table for
+  each variable in the output), visual functions (used for  visualization of each variable), and
+  check functions (used for data checks for each variable) currently implemented in \pkg{cleanR}. The variable
+  classes C, F, I, L, B, N, and D, refer to \R{character}, \R{factor},
+  \R{integer}, \R{labelled}, \R{logical} (boolean), \R{numeric}, and \R{Date}, respectively.}
 \label{table.SVCfunctions}
@@
 
-\hl{
-\begin{itemize}
-\item Introduce the relevant \R{clean}-arguments and the \R{defaultWhateverSummaries} etc.-functions
-\item Introduce \R{allSummaryFunctions()} etc. and present a table corresponding to the output of this call
-\item Small examples: 
-\begin{itemize}
-\item Add a function to one of the XXXSummaries/XXXVisuals/XXXChecks-arguments (still calling default options)
-\item Remove all but a single function from one of these arguments
-\item Describe what happens if the argument is NULL
-\end{itemize}
-\end{itemize}
-}
+\pkg{cleanR} uses three
+different types of functions for performing all steps in the above,
+namely \code{summaryFunction}s, \code{visualFunction}s and
+\code{checkFunction}s.
+By default, \R{clean()} runs all the summary, visualization, and check
+functions currently implemented for each variable type as part of the
+algorithm (see
+Figure~\ref{figure:cleanStructure}). Table~\ref{table.SVCfunctions}
+lists all these functions but we can also use the
+\R{allSummaryFunctions()}, \R{allVisualFunctions()}, and
+\R{allCheckFunctions()} function in \proglang{R} to print an overview
+list. For example the implemented summary functions are:
+
+%\subsection{Something about what check, visual and summary functions are available}
 
@@
 -------------- ------------------------------- -----------------------------
-centralValue   Compute median or mode          character, Date,
-factor,     
+centralValue   Compute median or mode          character, Date, factor,     
                                                integer, labelled,
-                                               logical,  
-                                               numeric
-
-countMissing   Compute ratio of missing        character, Date,
-factor,     
-               observations                    integer, labelled,
-               logical,  
+                                               logical,  numeric
+
+countMissing   Compute ratio of missing        character, Date, factor,     
+               observations                    integer, labelled, logical,  
                                                numeric
@@
 
-uniqueValues   Count number of unique values   character, Date,
-factor,     
+uniqueValues   Count number of unique values   character, Date, factor,     
                                                integer, labelled,
-                                               logical,  
-                                               numeric
-
-variableType   Data class of variable          character, Date,
-factor,     
+                                               logical, numeric
+
+variableType   Data class of variable          character, Date, factor,     
                                                integer, labelled,
-                                               logical,  
-                                               numeric                      
+                                               logical, numeric                      
 ----------------------------------------------------------------------------
@@
 
-In this section, we discuss how to control every step in \R{clean()} that actually involves a function being called on variables from the dataset.
-There are two stages in which this occurs, as mentioned in Figure~\ref{figure:cleanStructure}, namely:
+Thus we can see, for example, that the summary table for a
+\R{numeric}, \R{integer}, or \R{Date} variable will include
+information about the minimum and maximum values, while all variable
+classes will produce some kind of information about the central
+tendency (i.e., mode or median).
+
+We can control the default summaries and checks for each variable type
+through the \R{XXXSummaries} and \R{XXXChecks} arguments where \R{XXX}
+represents a variable type, e.g., \R{factorChecks} for factors,
+\R{numericChecks} for numeric variables, etc. These arguments accept a
+vector of summary or check functions that should be run for a
+particular variable type and the default values for each of these
+arguments (i.e., the functions from Table~\ref{table.SVCfunctions}
+that are applied) can be obtained through the \R{defaultXXXChecks()}
+and \R{defaultXXXSummaries()} functions, where \R{XXX} again refers to
+the variable type.
+
+For example, the default summaries being used for a factor variable is 
+\begin{Schunk}
+\begin{Sinput}
+> defaultFactorSummaries()
+\end{Sinput}
+\begin{Soutput}
+[1] "variableType" "countMissing" "uniqueValues" "centralValue"
+\end{Soutput}
+\end{Schunk}
+
+We can change the summaries (and similarly the checks) by setting the
+corresponding arguments when calling \R{clean()}. For example, to get
+only the variable type and the central tendency listed in the summary
+table of each variable we write
+
+\begin{Schunk}
+\begin{Sinput}
+> clean(toyData, output="html",  
+        characterSummaries = c("variableType", "centralValue"),
+        factorSummaries = c("variableType", "centralValue"),
+        labelledSummaries = c("variableType", "centralValue"),
+        numericSummaries = c("variableType", "centralValue"),
+        integerSummaries = c("variableType", "centralValue"),
+        logicalSummaries = c("variableType", "centralValue"),
+        dateSummaries = c("variableType", "centralValue"))
+\end{Sinput}
+\end{Schunk}
+
+In this particular case, where we specify the same set of summary
+functions for each variable type, we can use the simpler argument
+\R{allSummaries} which overrides the summary functions for all
+variable types. Thus, the same result as above could be obtained with
+
+\begin{Schunk}
+\begin{Sinput}
+> clean(toyData, output="html",  
+        allSummaries = c("variableType", "centralValue"))
+\end{Sinput}
+\end{Schunk}
+
+Similarly, the checks applied are set with the \R{XXXChecks}
+arguments. The default checks being applied to a factor is
+
+\begin{Schunk}
+\begin{Sinput}
+> defaultFactorChecks()
+\end{Sinput}
+\begin{Soutput}
+[1] "identifyMissing"    "identifyWhitespace" "identifyLoners"    
+[4] "identifyCaseIssues" "identifyNums" 
+\end{Soutput}
+\end{Schunk}
+
+Now, if we only wanted to apply the identify whitespace function for
+factor variables then we would need to set the \R{factorChecks} accordingly
+
+\begin{Schunk}
+\begin{Sinput}
+> clean(toyData, output="html",  
+        factorChecks = c("identifyWhitespace"))
+\end{Sinput}
+\end{Schunk}
+
+or we could remove checks for factor altogether by setting the
+corresponding argument to \R{NULL} in which case factor variables will
+not be checked for any potential errors.
+
+\begin{Schunk}
+\begin{Sinput}
+> clean(toyData, output="html", factorChecks = NULL)
+\end{Sinput}
+\end{Schunk}
+
+%\hl{
+%\begin{itemize}
+%\item Introduce the relevant \R{clean}-arguments and the \R{defaultWhateverSummaries} etc.-functions
+%\item Introduce \R{allSummaryFunctions()} etc. and present a table corresponding to the output of this call
+%\item Small examples: 
+%\begin{itemize}
+%\item Add a function to one of the XXXSummaries/XXXVisuals/XXXChecks-arguments (still calling default options)
+%\item Remove all but a single function from one of these arguments
+%\item Describe what happens if the argument is NULL
+%\end{itemize}
+%\end{itemize}
+%}
+
+As indicated in Figure~\ref{figure:cleanStructure} there are two stages
+where \R{clean()} applies function(s) to each of the variables: 
 \begin{enumerate}
 \item In the precheck functions
-\item In the summarize/visualize/check (SVC) step
+\item As part of the summarize/visualize/check (SVC) steps
 \end{enumerate}
-Each of these stages are controllable using appropriate function parameters in \code{clean}.
-In the above, we presented the default \pkg{cleanR} settings and how to tweak them into providing a slightly different data cleaning outputs.
-However, if for instance the dataset at hand requires completely different visualizations, more control is needed.
-\pkg{cleanR} uses three different types of functions for performing all stages in the above, namely \code{summaryFunction}s, \code{visualFunction}s and \code{checkFunction}s.
-\hl{Something like "the avaialable options for functions used in the precheck- and check steps are obtainable by calling \R{allCheckFunction()}, blablabla, similarly with visualFunctions and summaryFunctions.
-Mention Section 4 and the fact that one can expand the possibilities of e.g.
-allVisualFunctions() by producing new functions quite easily.}
+Each of these stages are controllable using appropriate function
+arguments in \R{clean()}, and above we have shown examples of how to
+tweak them to modify the data cleaning outputs. However, if for
+instance the dataset at hand requires completely different
+visualizations then more control is needed, and
+Section~\ref{sec:extending} explains how to modify, and expand the
+possibilities by producing new functions quite easily.
 
@@
 Note that the arguments specifying which checks to perform, as
-described in the previous section, are in fact passed to \R{check()}, and
-thus they can also be used here. For instance, if we only want the
-result of the check for miscoded missing values, we write
+described in the previous section, are in fact passed to \R{check()},
+and thus they can also be used here. For instance, if we only want to
+check for potentially miscoded missing values, we can use the relevant
+\R{XXXChecks} argument (e.g., \R{numericChecks}, \R{factorChecks},
+etc. as described in Section~\ref{sec:usingcleanR}) to provide a
+vector of the check functions that should be applied. Recall that
+Table~\ref{table.SVCfunctions} or the \R{allCheckFunctions()} provides
+overviews of the available check functions.
 \begin{Schunk}
@@
 \end{Schunk}
-The \R{numericChecks} argument XXX.. ... not defined
 
@@
 such as \R{identifyMissing} (see \ref{table.SVCfunctions} for a list
-of check functions) is by using it directly on the variable, i.e.
+of check functions) is by using it directly on the variable, e.g.,
 \begin{Schunk}
@@
 
-\hl{
-\begin{itemize}
-\item Do an example with visualize() and summarize(), like the one with check(). Especially visualize and to doEval = T thing needs a bit of special attention.
-\item  Mention allCheckFunctions() etc. again here
-\item Mention check(), visualize() and summarize() modes for data.frames.
+Similarly, the \R{visualize()} and \R{summarize()} functions can be
+used to run the corresponding visualizations and summaries for each
+variable. See Figure~\ref{fig:example2} for the visualization output.
+
+\begin{Schunk}
+\begin{Sinput}
+> visualize(toyData$var2)
+> summarize(toyData$var2)
+\end{Sinput}
+\begin{Soutput}
+     Feature                   Result    
+[1,] "Variable type"           "numeric" 
+[2,] "Number of missing obs."  "3 (20 %)"
+[3,] "Number of unique values" "8"       
+[4,] "Median"                  "4.5"     
+[5,] "1st and 3rd quartiles"   "1.75; 6" 
+[6,] "Min. and max."           "1; 999"  
+\end{Soutput}
+\end{Schunk}
+% $
+
+\begin{figure}[tb]
+\begin{center}
+\includegraphics[width=7.5cm]{toyData-var2.pdf}
+\end{center}
+\label{fig:example2}
+\caption{Output from running \R{visualize()} on the variable \texttt{var2} from the
+\R{toyData} dataset.}
+\end{figure}
+
+As we saw with the \R{check()} function, the summary can be modified
+by providing the relevant \R{XXXSummaries} arguments. Setting the
+\R{numericSummaries} argument we can control the summary output by
+providing a vector of functions to run for a particular summary. To
+only get the minimum and maximum we set
+\R{numericSummaries=c("minMax")}.
+
+\begin{Schunk}
+\begin{Sinput}
+> summarize(toyData$var2, numericSummaries=c("minMax"))
+\end{Sinput}
+\begin{Soutput}
+     Feature         Result  
+[1,] "Min. and max." "1; 999"
+\end{Soutput}
+\end{Schunk}
+% $
+
+%\hl{
+%\begin{itemize}
+%\item Do an example with visualize() and summarize(), like the one with check(). Especially visualize and to doEval = T thing needs a bit of special attention.
+% \item  Mention allCheckFunctions() etc. again here
+% \item Mention check(), visualize() and summarize() modes for data.frames.
 Maybe also advice against it, at it will often produce a lot of information at once, and such large amounts of information really should be documented.
-\end{itemize}
-}
+%\end{itemize}
+%}
 
@@
 \scriptsize
-\begin{tabular}{p{0.2\linewidth}p{0.2\linewidth}p{0.2\linewidth}p{0.2\linewidth}}
+\begin{tabular}{L{0.2\linewidth}L{0.2\linewidth}L{0.2\linewidth}L{0.2\linewidth}}
 & \R{summaryFunction} & \R{visualFunction} & \R{checkFunction} \\
@@
 \end{tabular}
-\caption{\hl{blabla. Fix formatting! Maybe use multirow and raggedRight stuff?}}
+\caption{Reference information for creating new functions to be used
+  as part of the summarize, visualize, and check steps. The three
+  columns correspond to each of the three function types.}
 \label{table.functionTypes}
@@
 
-Though the discussion in the above paints a picture of cleanR as a user-friendly package which requires practically no knowledge of \R{R}, one should not be mistaken to think that it is not customizable.
-In fact, the main function of \pkg{cleanR}, \code{clean}, is mainly a tool for formatting the results from various checking-, summary- and visualization functions.
-Thus, the actual work underlying a \pkg{cleanR} output file can be anything or nothing - depending on the arguments given to \code{clean}.
-Specfically, user made functions can be added to the SVC-function arguments discussed above, e.g.
-\R{factorSummaries}, \R{allVisuals} or \R{numericChecks}: All that is needed is to specify their names, just like the names of the build-in SVC functions are specified.
-However, not just any function can be called from these three steps, and therefore, we will now present how \R{summaryFunction}s, \R{visualFunction}s and \R{checkFunction}s are made.
-\hl{Mention something about the interactive mode here as well?} \\
-
-This section consists of two parts.
-First, we describe how the \R{clean} function can be extended by adding custom prechecks, summaries, checks and visualizations.
-In order to do this, one needs to produce small functions that obey to the syntax of \pkg{cleanR}.
-This can be done with different levels of strictness.
-If the custom functions are only to be used by \R{clean()}, only the input/output structures of the functions need to be restricted.
-However, by adding a few extra \hl{word?} to these functions, the full machinery of \pkg{cleanR} becomes available for the new, user-made functions, just as it is for the build-in functions.
-The presentation below is given in the format of function templates, written in pseudo-code.
-These templates are designed for getting the full functionality, but please note that Table~\ref{table.functionTypes} serves as a reference to the minimal requirements, while also presenting the "full" versions of the function types.
-
-After this abstract, templatic overview of the internals of the three SVC functions is given, we turn to a worked example of how to use custom made functions in practice in \hl{ref to worked example section}.
-Here, four new SVC functions are defined and used, both interactively and in \R{clean()}.
+\pkg{cleanR} can be used as a user-friendly, self-contained package as
+shown in the previous section. However, \pkg{cleanR} is fully
+customizable and \R{clean()} is mainly a tool for formatting the
+results from various checking-, summary- and visualization
+functions. Thus, the actual work underlying a \pkg{cleanR} output file
+can be anything --- depending on the arguments given to \R{clean()}
+--- and user made functions are easily added to the
+summarize/visualize/check-function arguments as discussed previously
+(e.g., \R{factorSummaries}, or \R{numericChecks}). However, the
+functions used in the summarize/visualize/check steps need to adhere
+to a specific structure to be called from these three steps, and
+therefore, we will now present how \R{summaryFunction}s,
+\R{visualFunction}s and \R{checkFunction}s are made.
+%\hl{Mention something about the interactive mode here as well?}
+
+This section consists of two parts. First, we describe how the
+\R{clean()} function can be extended by adding custom prechecks,
+summaries, checks and visualizations. In order to do this, one needs
+to create functions with a specific structure required by
+\pkg{cleanR}. This can be done with different levels of strictness. If
+the custom functions are only to be used part of the corresponding
+arguments of \R{clean()} then only the input/output structures of the
+functions need to be fixed.
+
+However, new user-defined functions can also be registered locally to
+be part of the full machinery of \pkg{cleanR}, and these function will
+be recognized and behave in the same way as the build-in functionson
+\pkg{cleanR}. The presentation below is given in the format of
+function templates, written in pseudo-code. These templates are
+designed for getting the full functionality, but note that
+Table~\ref{table.functionTypes} serves as a reference to the minimal
+requirements, while also presenting the "full" versions of the
+function types. We end this section with a 
+%After this abstract, templatic overview of the internals of the three
+%SVC functions is given, we turn to a 
+worked example of how to use custom made functions in practice, where
+four new summarize/visualize/check functions are defined and used,
+both interactively and in \R{clean()}.
 
@@
 \label{sec:functionTemplates}
-\hl{metatext, mention Table \ref{table.functionTypes} again.}
+% \hl{metatext, mention Table \ref{table.functionTypes} again.}
 
 \subsubsection{Writing a summaryFunction}
-As mentioned above, \pkg{cleanR} provides a dedicated class for \code{summaryFunction}s.
-However, this does not imply that they are particularly advanced or complicated to create; in fact, they are nothing but regular functions with a certain input/output-structure.
-Specifically, they all follow the template below:
+As mentioned above, \pkg{cleanR} provides a dedicated class for
+\code{summaryFunction}s. However, this does not imply that they are
+particularly advanced or complicated to create; in fact, they are
+nothing but regular functions with a particular
+input/output-structure. Specifically, they all follow the template
+below:
 \begin{Verbatim}
@@
 \end{Verbatim}
-The last function called here, \R{summaryResult()}, changes the class of the output, thereby making a \R{print()} method available for it.
-Note that \code{v} is a vector and that \code{res} should be either a character string or something that will be printed as one.
-In other words, e.g.
-integers are allowed, but matrices are not.
-Though a lot of different things can go into the \code{summaryFunction} template, we recommend only using it for summarizing the features of a variable, and leaving tests and checks for the \code{checkFunction}s (presented below).
+The last function called here, \R{summaryResult()}, changes the class
+of the output, thereby making a \R{print()} method available for it.
+Note that \code{v} is a vector and that \code{res} should be either a
+character string or something that will be printed as one. In other
+words, e.g. integers are allowed, but matrices are not. Though a lot
+of different things can go into the \code{summaryFunction} template,
+we recommend only using it for summarizing the features of a variable,
+and leaving tests and checks for the \code{checkFunction}s (presented
+below).
 

### 2016-12-28 R/clean.R

@@ -1,11 +1,18 @@
 #' Produce a data cleaning overview document
 #'
-#' Run a set of validation checks to check the variables in a dataset for potential errors.
-#' Performs checking steps according to user input and/or data type of the inputted variable.
-#' The checks are saved to an R markdown file which can rendered into an easy-to-read document.
-#' This document also includes summaries and visualizations of each variable in the dataset.
+#' Run a set of class-specific validation checks to check the
+#' variables in a dataset for potential errors.  Performs checking
+#' steps according to user input and/or data type of the inputted
+#' variable.  The checks are saved to an R markdown file which can
+#' rendered into an easy-to-read document.  This document also
+#' includes summaries and visualizations of each variable in the
+#' dataset.
 #'
-#' @param data The data frame to be checked. This dataset should be of class \code{data.frame},
+#' The checks done for each class of variable is chosen by the
+#' SVC ...
+#'
+#'
+#' @param data The dataset to be checked. This dataset should be of class \code{data.frame},
 #' \code{tibble} or \code{matrix}. If it is of classs \code{matrix}, it will be converted to a
 #' \code{data.frame}.
 #'
@@ -368,7 +375,7 @@ clean <- function(data, output=c("pdf", "html"), render=TRUE,
                    "- rename your cleanR output file using the \"file\" option \n",
                    "- Add a volume number to your file name using the \"vol\" option \n",
                    "- check that you do not want to keep the original file and if so,",
-                   "use cleanR with replace = TRUE"))
+                   "use clean() with argument replace = TRUE"))
     }
 
       #if (replace=="onlyCleanR") {
@@ -452,7 +459,7 @@ clean <- function(data, output=c("pdf", "html"), render=TRUE,
     }
 
     chunk.wrapper <- function(x, ..., outfile=fileConn, options=c("echo=FALSE", "warning=FALSE"), label=NULL) {
-        writer(paste("```{r", ifelse(is.null(label),"," , paste(label, ",")) , paste(options, collapse=", "), "}"))
+        writer(paste0("```{r ", ifelse(is.null(label), ", ", paste0(label, ", ")), paste0(options, collapse=", "), "}"))
         writer(x, ..., outfile=outfile)
         writer("```")
     }

### 2016-12-28 R/identifyWhitespace.R

@@ -1,23 +1,23 @@
 #' @title A checkFunction for identifying whitespace
 #'
-#' @description A checkFunction to be called from \code{\link{check}} 
-#' that identifies prefixed and suffixed whitespace(s) in character, 
+#' @description A checkFunction to be called from \code{\link{check}}
+#' that identifies prefixed and suffixed whitespace(s) in character,
 #' labelled or factor variables.
 #'
 #' @param v A character, labelled or factor variable to check.
-#' 
+#'
 #' @param nMax The maximum number of problematic values to report.
-#'  Default is \code{Inf}, in which case all problematic values are 
+#'  Default is \code{Inf}, in which case all problematic values are
 #'  included in the outputted message.
 #'
-#' @return  A \code{\link{checkResult}} with three entires: 
+#' @return  A \code{\link{checkResult}} with three entires:
 #' \code{$problem} (a logical indicating whether any whitespaces were
 #' fount), \code{$message} (a message describing which values were prefixed
 #' or suffixed with whitespace) and \code{$problemValues} (the problematic
-#' values). Note that only unique values are printed in the message, and that 
+#' values). Note that only unique values are printed in the message, and that
 #' they are sorted alphabetically.
 #'
-#' @seealso \code{\link{check}}, \code{\link{allCheckFunctions}}, 
+#' @seealso \code{\link{check}}, \code{\link{allCheckFunctions}},
 #' \code{\link{checkFunction}}, \code{\link{checkResult}}
 #'
 #' @examples
@@ -69,7 +69,7 @@ identifyWhitespaceC <- function(v, nMax) {
   outMessage <- messageGenerator(list(problem=problem,
                                       problemValues=problemValues),
                                  nMax = nMax)
-  checkResult(list(problem = problem, message = outMessage, 
+  checkResult(list(problem = problem, message = outMessage,
                    problemValues = problemValues))
 }
 
@@ -81,7 +81,7 @@ identifyWhitespaceF <- function(v, nMax) {
 #labelled variables
 identifyWhitespaceL <- function(v, nMax) {
   v <- na.omit(v)
-  v <- unpackLabelled(v)
+  v <- haven::as_factor(v)
   identifyWhitespaceC(v, nMax = nMax)
 }
 

### 2016-12-29 latex/article_vol2.tex
Second bunch of commits

@@
 
-Though adhering to the template above is sufficient for using the freshly made \R{mySummaryFunction()} in \R{clean()}, we recommend furthermore adding it to the overview of all summary functions by converting it to a proper \code{summaryFunction} object.
-This is done by writing
+Adhering to the template above is sufficient for using the freshly
+made \R{mySummaryFunction()} in \R{clean()}, but we recommend
+furthermore adding the new function to the overview of all summary
+functions by converting it to a proper \code{summaryFunction}
+object. This is done by calling the \R{summaryFunction} function with
+the user-defined function as first argument, and additional arguments
+\R{description} which adds explanatory flavour text as an attribute of
+the function, and \R{classes} which determines which classes the
+user-defined function should be applied to (also stored as an
+attribute).
 \begin{Verbatim}
 mySummaryFunction <- summaryFunction(mySummaryFunction,
-	description = "[Some text describing what the summaryFunction does]",
-	classes = c([the data types that this function is intended to be used for]))
+	description = "[Text describing what the summaryFunction does]",
+	classes = c([vector of data types that the function is intended for]))
 \end{Verbatim}
-which adds the new function to the output of an \code{allSummaryFunctions()} call.
-One comment should be devoted to the two attributes of a \R{summaryFunction}.
-If the \R{description} argument is left unspecified, the name of the function (in this case, \R{"mySummaryFunction"}) will be filled in.
-What happens if the \R{classes} argument is not specified depends on the type of \R{mySummaryFunction}.
-If \R{mySummaryFunction} is a \R{S3} generic function with associated methods, the call to \R{summaryFunction()} will automatically produce a vector of the names of the classes for which the function can be called.
-If \R{mySummaryFunction} is not an \R{S3} generic and \R{classes} is left unspecified, the attribute will simply be empty.
-Note that the helper function \R{allClasses()} might be useful for filling out the \R{classes} argument, as it simply lists all available classes in \pkg{cleanR}:
+which adds the new function to the output of an
+\code{allSummaryFunctions()} call. 
+% One comment should be devoted to
+% the two attributes of a \R{summaryFunction}. 
+If the \R{description} argument is left unspecified, the name of the
+function (which in this case is \R{"mySummaryFunction"}) will be filled in and
+stored as an attribute. What happens if the \R{classes} argument is
+not specified depends on the type of \R{mySummaryFunction}. If
+\R{mySummaryFunction} is a \R{S3} generic function with associated
+methods, the call to \R{summaryFunction()} will automatically produce
+a vector of the names of the classes for which the function can be
+called. If \R{mySummaryFunction} is not an \R{S3} generic and
+\R{classes} is left unspecified, the attribute will simply be
+empty. Note that the helper function \R{allClasses()} might be useful
+for filling out the \R{classes} argument, as it simply lists all
+available classes in \pkg{cleanR}:
 \begin{Schunk}
@@
 \end{Schunk}
-\hl{Write something here, don't end paragraph with code. Also, maybe move the allClasses() stuff somewhere else, it doesn't really belong under this header. Not sure where to, though.}
+% \hl{Write something here, don't end paragraph with code. Also, maybe move the allClasses() stuff somewhere else, it doesn't really belong under this header. Not sure where to, though.}
+
+If we run the \R{allSummaryFunctions()} function we can see that the
+new function has been registered (output slightly appreviated):
+
+\begin{Schunk}
+\begin{Sinput}
+> allSummaryFunctions()
+\end{Sinput}
+\begin{Soutput}
+----------------------------------------------------------------------------
+name                description                                      classes
+------------------- --------------------------------------------------------
+mySummaryFunction   [Text describing what the summaryFunction does]  numeric
+.
+.
+.
+\end{Soutput}
+\end{Schunk}
 
 \subsubsection{Writing a visualFunction}
-\code{visualFunction}s are the functions that produce the figures of a \pkg{cleanR} output document.
-Writing a \R{visualFunction} is slightly more complicated than writing a \R{summaryFunction}.
-This follows from the fact that \R{visualFunctions} need to be able to output standalone code for plots in order for \code{clean()} to build standalone \R{rmarkdown} files.
-We recommend using the following structure:
+\code{visualFunction}s are the functions that produce the figures in a
+\pkg{cleanR} output document. Writing a \R{visualFunction} is slightly
+more complicated than writing a \R{summaryFunction}. This follows from
+the fact that \R{visualFunctions} need to be able to output standalone
+code for plots in order for \code{clean()} to build standalone
+\pkg{rmarkdown} files. We recommend using the following structure:
 \begin{Verbatim}
@@
 }
-
+\end{Verbatim}
+
+In this function, \code{v} is the variable to be visualized,
+\code{vnam} is its name (which should generally be passed to
+\code{title} or \code{main} arguments in plotting functions) and
+\code{doEval} controls whether the output is a plot (if \code{TRUE})
+or a character string of standalone code for producing a plot (if
+\code{FALSE}). Implementing the \code{doEval = TRUE} setting is not
+strictly necessary for a \R{visualFunction}'s use in \code{clean()}, but
+it makes it easier to assess what visualization options are available,
+and obviously, it is crucial for interactive usage of
+\R{myVisualFunction()}. In either case, it should be noted that all
+the parameters listed above, \code{v}, \code{vnam} and \code{doEval},
+are mandatory, so they must be left as is, even if they are not in
+use. See Table~\ref{table.functionTypes}.
+
+Similarly to the summary function, we call \R{visualFunction()} to
+register our newly created function. Like before, it accepts the
+user-defined function as the first argument, and two additional
+arguments description which adds explanatory flavour text as an
+attribute of the function, and a vector which determines which classes
+the user-defined function should be applied to (which is also stored
+as an attribute).
+
+\begin{Verbatim}
 myVisualFunction <- visualFunction(myVisualFunction,
 	description = "[Some text describing the visualFunction]",
-	classes = c([the data types that this function is intended to be used for]))
+	classes = c([data types that this function is intended for]))
 	)
 \end{Verbatim}
-In this function, \code{v} is the variable to be visualized, \code{vnam} is its name (which should generally be passed to \code{title} or \code{main} arguments in plotting functions) and \code{doEval} controls whether the output is a plot (if \code{TRUE}) or a character string of standalone code for producing a plot (if \code{FALSE}).
-Implementing the \code{doEval = TRUE} setting is not strictly necessary for a \R{visualFunction}'s use in \code{clean}, but it makes it easier to assess what visualization options are available, and obviously, it is crucial for interactive usage of \R{myVisualFunction()}.
-In either case, it should be noted that all the parameters listed above, \code{v}, \code{vnam} and \code{doEval}, are mandatory, so they must be left as is, even if they are not in use.
+
+Again, registering the visualization function allows it to be used
+without manually adding it to the list of visualizations.
 
 \subsubsection{Writing a checkFunction}
-The last, but perhaps most important, \pkg{cleanR} function type is the \code{checkFunction}.
-These are the functions that flag issues in the data in the check step and control the overall flow of the data cleaning process in the precheck stage.
-A \code{checkFunction} can follow two overall structures, depending on the type of check.
-Either, it tries to identify problematic values in the variable (as e.g.
-\R{identifyMissing()} does) or it performs a check concerning the variable as a whole (e.g.
-the functions used for prechecks and the function \R{identifyNums()}).
-We present templates for both types of \R{checkFunction}s below separately, but it should be emphasized that formally, they belong to the same class.
-
-First, a template for the full-variable check function type:
+The last, but perhaps most important, \pkg{cleanR} function type is
+the \code{checkFunction}. These are the functions that flag issues in
+the data in the check step and control the overall flow of the data
+cleaning process in the precheck stage. A \code{checkFunction} follows
+one of two overall structures, depending on the type of check. Either,
+it tries to identify problematic values in the variable (as
+e.g., \R{identifyMissing()} does) or it performs a check concerning the
+variable as a whole (e.g. the functions used for prechecks and the
+function \R{identifyNums()}). We present templates for both types of
+\R{checkFunction}s below separately, but it should be emphasized that
+formally, they belong to the same class.
+
+First, a template for the full-variable check function type, where we
+first define the function and subsequently register it as a check
+function using \R{checkFunction()}:
 \begin{Schunk}
@@
 \end{Schunk}
-Again, as with \R{summaryFunction}s and \R{visualFunction}s, the change of function class by use of \R{checkFunction()} is not strictly necessary.
-Note however, that if \R{myFullVarCheckFunction} is to be used in the SVC step in \R{clean()}, the description attribute will be printed in the overview table in the \textit{Data cleaning summary}.
-
-If problematic values are to be identified, the template from above should be expanded to follow a slightly more complicated structure:
+
+Again, as with \R{summaryFunction}s and \R{visualFunction}s, the
+change of function class by use of \R{checkFunction()} is not strictly
+necessary. Note however, that if \R{myFullVarCheckFunction} is to be
+used in the summarize/visualize/check steps in \R{clean()}, the
+description attribute will be printed in the overview table in the
+\textit{Data cleaning summary} part of the output document.
+
+If problematic values are to be identified, the template from above
+should be expanded to follow a slightly more complicated structure:
 \begin{Verbatim}
@@
 
-	problemMessage <- "[The message that should be printed prior to listing
-			problem values in the cleanR output, ending with a colon]"
+	problemMessage <- "[Message that is printed prior to listing
+			    problem values in the cleanR output, 
+                            ending with a colon]"
 
@@
 \end{Verbatim}
-In this template, the argument \R{maxDecimals} is not in use.
-This argument should be used to round off the \R{problemValues} passed to \R{messageGenerator()}, if they are numerical.
-This is done by substituting the \R{problemStatus} assignment above with the following code:
+For this template, the argument \R{maxDecimals} is not in use. This
+argument is used to round off the \R{problemValues} passed to
+\R{messageGenerator()}, if they are numerical.  This is done by
+substituting the \R{problemStatus} assignment above with the following
+code: \hl{Denne foregående sætning forstår jeg ikke helt. Måske skal
+  messagegenerator have en sætning eller to med på vejen? (Og det spiller lidt sammen med nedenstående)}
 \begin{Schunk}
@@
 \end{Schunk}
-Another noteworthy component of the template is the usage of the helper function \R{messageGenerator()}, which aids consistent styling of all \code{checkFunction} messages.
-This function simply pastes together the \code{problemMessage} and the \code{problemValues}, with the latter being quoted and sorted alphabetically.
-If \R{nMax} is not \R{Inf}, only the first \R{nMax} problem values will be pasted onto the message, accompanied by a comment about how many problem values were left out (if any).
-Note that printing quotes in rmarkdown requires an extensive amount of character escaping, so opting for \code{messageGenerator()} really is the easiest solution.
+
+It is worth mentioning the 
+%Another noteworthy component of the template is the usage of the
+helper function \R{messageGenerator()}, which aids in consistent
+styling of all \code{checkFunction} messages. This function simply
+pastes together the \code{problemMessage} and the
+\code{problemValues}, with the latter being quoted and sorted
+alphabetically. If \R{nMax} is not \R{Inf}, only the first \R{nMax}
+problem values will be pasted onto the message, accompanied by a
+comment about how many problem values were left out (if any).  Note
+that printing quotes in \pkg{rmarkdown} requires an extensive amount
+of character escaping, so opting for \code{messageGenerator()} really
+is the easiest solution.
 
 \subsection{A worked example}
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.2]{mosaicPlotExample.pdf}
-\end{center}
-\label{fig:mosaicPlot}
-\caption{\hl{something, note that code for producing this plot is available in ./latex/codeForArticle.R}}
-\end{figure}
-
-We will now build four new functions and show both how they can be used interactively and how they can be integrated with the \R{clean()} function. These four new functions are:
+
+We will now build four new functions and show both how they can be
+used interactively and how they can be integrated with the \R{clean()}
+function. These four new functions are:
 \begin{description}
@@
 This function checks whether a variable consists exclusively of long ($> 10$ characters/digits) entries that are all of equal length, as this might be personal identification codes that we do not wish to print out in the data summary.
-\item[\R{mosaicVisual}] A new \R{visualFunction} that produces so-called mosaic plots. This function will be used in the \textit{visualize} step of \R{clean()}.
+\item[\R{mosaicVisual}] A new \R{visualFunction} that produces mosaic plots. This function will be used in the \textit{visualize} step of \R{clean()}.
 \item[\R{countZeros}] A new \R{summaryFunction} that counts the number of occurrences of the value \R{0} in a variable. This function will be used in the \textit{summarize} step of \R{clean()}.
 \item[\R{identifyColons}] A new \R{checkFunction} that flags variables in which values have colons that appear before and after alphanumerical characters.
-This is e.g.
-practical for identifying autogenerated interaction effects.
+This is practical for identifying autogenerated interaction effects.
 This function will be used in the \textit{check} step of \R{clean()}.
@@
 
-\subsubsection{isID - a new checkFunction without problem values}
-First, let's define the \R{isID} function.
-As this function is not supposed to list problematic values in the variable, it falls within the category of \R{checkFunctions} represented by \R{myFullVarCheckFunction()} in the above.
-We do not particularly wish to use this function interactively, so we will stick to the minimal requirements of a \R{checkFunction} used in \R{check()} (see Table~\ref{table.functionTypes}).
-The function can then be defined by
+\subsubsection{isID --- a new checkFunction without problem values}
+First, let's define the \R{isID()} function. As this function is not
+supposed to list problematic values in the variable, it falls within
+the category of \R{checkFunctions} represented by
+\R{myFullVarCheckFunction()} in the above. We do not particularly wish
+to use this function interactively, so we will stick to the minimal
+requirements of a \R{checkFunction} used in \R{check()} (see
+Table~\ref{table.functionTypes}). The function can then be defined by
 \begin{Schunk}
@@
   out <- list(problem = FALSE, message = "")
+  # Should work for all but logical and Date variables
   if (class(v) %in% setdiff(allClasses(), c("logical", "Date"))) {
@@
     lengths <- c(nchar(v))
+    # Check that all lengths are the same and greater than 10 characters
     if (all(lengths > 10) & length(unique(lengths)) == 1) {
@@
 \end{Schunk}
-\hl{Mention somewhere that  \R{mosiacplot()} is a base-R (\R{graphics}) function?
-Otherwise, it might not be completely clear, that we are passing a function name in the above...} This is essentially all we need to do in order to include this function as a precheck-function in \R{clean()}, so we will leave it as is and move on to the next function, namely \R{mosaicVisual}.
-
-\subsubsection{mosiacVisual - a new visualFunction}
-We will define this function such that it gets the full \pkg{cleanR} functionality. This can be done using the code 
+
+This is essentially all we need to do in order to include this
+function as a precheck-function in \R{clean()}, so we will leave it as
+is and move on to the next function, namely \R{mosaicVisual}.
+
+\subsubsection{mosiacVisual --- a new visualFunction}
+\proglang{R} contains a function, \R{mosiacplot}, which produces
+mosaic plots. We intend to use this existing high-level plotting
+function as part of the new visualization function.  We will define
+the new function such that it gets the full \pkg{cleanR}
+functionality. This can be done using the following code.
 \begin{Schunk}
@@
  mosaicVisual <- function(v, vnam, doEval) {
+   # Setup the call using the existing function mosaicplot
    thisCall <- call("mosaicplot", table(v), main = vnam, xlab = "")
-   if (doEval) {
+   if (doEval) {                     # Return the graphics
     return(eval(thisCall))
-   } else return(deparse(thisCall))
+   } else return(deparse(thisCall))  # Else return the code for the call
  }
@@
 \end{Schunk}
-This function can now be called directly or used in \R{clean()}.
-We will return to its usage in \R{clean()} below.
-Depending on the argument \R{doEval}, either a text string with code or a plot is produced.
-The plot resulting from the following call is found in Figure~\ref{fig:mosaicPlot}:
-\begin{Schunk}
-\begin{Sinput}
-mosaicVisual(toyData$var1, "variable 1", doEval = TRUE)   $
-\end{Sinput}
-\end{Schunk}
-\hl{remove \$.}
-Even though \R{mosaicVisual}, as written above, follows the style of a \R{visualFunction}, it is not yet truly one and therefore, it will not appear in a \R{allVisualFunctions()} call.
-In order to get this functionality, we need to change its object class.
-This can be done by writing
+This function can now be called directly or used in \R{clean()} and we
+will show an example below. Depending on the \R{doEval} argument
+\hl{skal doEval have en default value ovenfor?}, either a text string
+with code or a plot is produced. The plot resulting from the following
+call is found in Figure~\ref{fig:mosaicPlot}:
+\begin{Schunk}
+\begin{Sinput}
+> mosaicVisual(toyData$var1, "variable 1", doEval = TRUE)   
+\end{Sinput}
+\end{Schunk}
+% $
+
+\begin{figure}
+\begin{center}
+\includegraphics[scale=0.2]{mosaicPlotExample.pdf}
+\end{center}
+
+\caption{Mosaic plot generated by the user-defined visualization
+  function \texttt{mosaicVisual}.}
+\label{fig:mosaicPlot}
+\end{figure}
+
+Even though \R{mosaicVisual()}, as written above, follows the style of a
+\R{visualFunction}, it is not yet truly one and therefore, it will not
+appear in a \R{allVisualFunctions()} call. In order to get this
+functionality, we need to change its object class. This can be done by
+writing
 \begin{Schunk}
@@
 \end{Schunk}
-Here, we use the function \R{allClasses()} to quickly obtain a vector of all the seven variable classes addressed in \pkg{cleanR}.
-Note that if \R{mosaicVisual} were an S3 generic function, this argument could have been left as \R{NULL} and then the classes for which methods are available would be added automatically.
-\hl{I'm repeating myself here, but I think it is quite a neat feature, so maybe that's okay?}
+Here, we use the function \R{allClasses()} to quickly obtain a vector
+of all the seven variable classes addressed in \pkg{cleanR}. Note that
+if \R{mosaicVisual} were an S3 generic function, this argument could
+have been left as \R{NULL} and then the classes for which methods are
+available would be added automatically. 
+%\hl{I'm repeating myself here,
+%  but I think it is quite a neat feature, so maybe that's okay?}
 
@@
 
-basicVisual    Histograms and barplots using                            
-               graphics
-
-standardVisual Histograms and barplots using                            
-               ggplot2                                                  
+basicVisual    Histograms and barplots using character, Date, factor,   
+               graphics                      integer, labelled, logical,                           
+                                             numeric                    
+standardVisual Histograms and barplots using character, Date, factor,                              
+               ggplot2                       integer, labelled, logical,               
+                                             numeric                    
 ------------------------------------------------------------------------
@@
 \end{Schunk}
-\hl{Redo this output table when makeXFunction issues are fixed.}
-
-Now, we are done with the definition of \R{mosaicVisual} and we can turn to the next function in line, \R{countZeros}.
-
-\subsubsection{countZeros - a new summaryFunction}
-This \R{summaryFunction} in spe is defined in the following lines of code:
+
+Now that we are done with the definition of \R{mosaicVisual()} we can
+turn to the next function in line, \R{countZeros}.
+
+\subsubsection{countZeros --- a new summaryFunction}
+This \R{summaryFunction} is defined in the following lines of code:
 \begin{Schunk}
@@
 \end{Schunk}
-Note that as this function computes an integer (the number of zeros), there is no difference between the entires \R{\$result} and \R{\$value}.
-If, on the other hand, the result had been a character string, extra formatting might be required in the \R{\$result} entry (such as escaping of quotation marks), and in this scenario, the two entries would then differ.
-As the result is returned as a \R{summaryResult} object, a printing method is automatically called when \R{countZeros} is used interactively:
+As this function computes an integer (the number of zeros),
+there is no difference between the entries \R{\$result} and
+\R{\$value}. If, on the other hand, the result had been a character
+string, extra formatting might be required in the \R{\$result} entry
+(such as escaping of quotation marks), and in this scenario, the two
+entries would have differed. As the result is returned as a
+\R{summaryResult} object, a printing method is automatically called
+when \R{countZeros} is used interactively:
 \begin{Schunk}
@@
 \end{Schunk}
-As with \R{mosaicVisual()}, we change the class of this function in order to make it appear in \R{allSummaryFunctions()} calls.
-But now we wish to emphasize that the function is not intended to be called on all variable types, as zeros have different roles in \R{Date}s and in \R{logical} variables:
+As with \R{mosaicVisual()}, we change the class of this function in
+order to make it appear in \R{allSummaryFunctions()} calls. But now we
+wish to emphasize that the function is not intended to be called on
+all variable types, as zeros have different roles in \R{Date}s and in
+\R{logical} variables:
 \begin{Schunk}
@@
 \end{Schunk}
-\hl{more? don't end on code.}
+% \hl{more? don't end on code.}
 
 \subsubsection{identifyColons - a new checkFunction with problem values}
-The last function mentioned above is \R{identifyColons()}.
-We define it using the helper function \R{messageGenerator} to obtain a properly escaped message, and we use \R{checkResult} to make its output print neatly:
+The last function mentioned above is \R{identifyColons()}. We define
+it using the helper function \R{messageGenerator()} to obtain a properly
+escaped message, and we use \R{checkResult} to make its output print
+neatly:
 \begin{Schunk}
@@
 
+  # Use regular expressions to identify colons between two words  
   problemValues <- v[sapply(gregexpr("[[:xdigit:]]:[[:xdigit:]]", v),
@@
                         problemValues = problemValues)
+  # Use the messagegenerator function to produce the output 
+  # message from the problemStatus and problemMessage
   outMessage <- messageGenerator(problemStatus, problemMessage, nMax)
@@
     classes = c("character", "factor", "labelled"))
-
-\end{Sinput}
-\end{Schunk}
-As with the previous two functions, we also change its class.
-Note, however, that for \R{checkFunction}s, the function description will appear in the document produced by \R{clean()} (in the \textit{Data cleaning summary} section), so now this is not only done for the sake of the \R{allCheckFunctions()} output.\\
-
-\subsubsection{Calling the new SVC functions from clean()}
-Now, we are ready to use these new functions in a \R{clean()} call.
-\hl{do this...
-what dataset should we use here?
-Include output in appendix, maybe.}.
-The extended \pkg{cleanR} output document should have the following modifications, relative to the standard \pkg{cleanR} output:
+\end{Sinput}
+\end{Schunk}
+As with the previous two functions, we also change its class. Note,
+however, that for \R{checkFunction}s, the function description will
+appear in the document produced by \R{clean()} (in the \textit{Data
+  cleaning summary} section of the output), so now this is not only
+done for the sake of the \R{allCheckFunctions()} output.
+
+\subsubsection{Calling the new summarize/visualize/check functions from clean()}
+Now, we are ready to use these new functions in a \R{clean()}
+call. The extended \pkg{cleanR} output document should have the
+following modifications, relative to the standard \pkg{cleanR} output:
 \begin{itemize}
@@
 \begin{Sinput}
-> clean(dataSet, 
+> data(testData) 
+> clean(testData, 
+      # Add the new pre-check function
       preChecks = c("isKey", "isEmpty", "isID"),
+      # Change the visual
       allVisuals = "mosaicVisual",
+      # Add the new checks for specific data types
       characterSummaries = c(defaultCharacterSummaries(), "countZeros"),
@@
 \end{Schunk}
-The outputted document is found in Appendix \hl{NUMTWO}. \hl{Comment. Make appendix. Remember to change dataSet to something else, when we decide what data to use here.}
-
-\section{Something like examples}
-Finally, we will present a few examples of how to make \pkg{cleanR} solve specific issues related to data cleaning.
-First, we discuss the challenges related to cleaning large datasets, particularly in terms of memory use and computation speed.
-Next, we show how \pkg{cleanR} can be used for problem-flagging.
-Lastly, we discuss how the \pkg{cleanR} output document can be included in other \R{rmarkdown} documents as a mean to produce clear and concise documentation of a dataset.
-\hl{I feel like there should be more topics here, but I'm all out of ideas...}
+The outputted document is found in Appendix \ref{sec:appendix2}.
+
+\section{Using cleanR in specific situations}
+Finally, we will present a few examples of how to make \pkg{cleanR}
+solve specific issues related to data cleaning. First, we discuss the
+challenges related to cleaning large datasets, particularly in terms
+of memory use and computation speed. Next, we show how \pkg{cleanR}
+can be used for problem-flagging. Lastly, we discuss how the
+\pkg{cleanR} output document can be included in other \pkg{rmarkdown}
+documents as a way to produce clear and concise documentation of a
+dataset. %\hl{I feel like there should be more topics here, but I'm all
+%  out of ideas...}
 
 \subsection{Cleaning large datasets}
-If the dataset becomes very large, the standard use of \R{clean()} outlined above might not be ideal.
-If there is a vast number of variables, production of the \R{rmarkdown} document might be quite slow, while an extensive amount of observations generally affects the rendering time of this document.
-In this section, we give a few practical examples of ways to deal with large data, while wishing to still produce (potentially very long) data cleaning overview documents.
-Note that the interactive tools of \pkg{cleanR} can be used as usual or sequentially in small subsets of the large dataset, if no such overview documents are needed.
-
-\subsubsection{Attacking the figures}
-Though figures give a nice overview of each variable, they are also quite heavy objects in terms of memory allocation.
-Therefore, it might be beneficial to not include figures in the \pkg{cleanR} outputs for very large datasets.
-This is controlled via the \R{mode} argument:
+If the dataset becomes very large, the standard use of \R{clean()}
+outlined above might not be ideal. If there is a vast number of
+variables, creation of the \pkg{rmarkdown} document might be quite
+slow, while a large number of observations generally will affect the
+rendering time of the document. In this section, we give a few
+practical examples of ways to deal with large data, while wishing to
+still produce (potentially very long) data cleaning overview
+documents. Note that the interactive tools of \pkg{cleanR} can be used
+as usual or sequentially in small subsets of the large dataset, if no
+such overview documents are needed.
+
+\subsubsection{Handling the figures}
+Though figures give a nice overview of each variable, they are also
+quite heavy objects in terms of memory allocation. Therefore, it might
+be beneficial to not include figures in the \pkg{cleanR} outputs for
+very large datasets. This is controlled via the \R{mode} argument:
 \begin{Schunk}
@@
 \end{Schunk}
-If figures are indeed needed, a different approach is to choose the less memory heavy standard \R{R} figure style instead of the \R{ggplot2}\hl{reference?} figures that are the default option in \R{clean()}.
-This can be done using the \R{allVisuals} argument:
+If figures are indeed needed, a different approach is to choose the
+less memory heavy standard \proglang{R} figure style instead of the
+\pkg{ggplot2} figures that are the default option in \R{clean()}. This
+can be done by setting the \R{allVisuals = "basicVisual"} argument:
 \begin{Schunk}
@@
 \end{Schunk}
-Of course, even less heavy plots might be achieved by writing new \R{visualFunction}s, using the guidelines from section \ref{sec:functionTemplates}.
-For instance, a future extension of \pkg{cleanR} might be the inclusion of ASCII plots, as e.g.
-represented in the \R{R} package \R{txtplot}\hl{reference here?
-It's on CRAN.}.
-\\
-
-\hl{I really feel like we should do some benchmarking here, maybe just on toyData, both in terms of speed and memory use. I would make the recommendations more trustworthy and serious.}
+Of course, even less heavy plots might be achieved by writing new
+\R{visualFunction}s, using the guidelines from section
+\ref{sec:functionTemplates}. For instance, a future extension of
+\pkg{cleanR} might be the inclusion of ASCII plots, as
+e.g. represented in the \proglang{R} package \pkg{txtplot}.
+
+%\hl{I really feel like we should do some benchmarking here, maybe just on toyData, both in terms of speed and memory use. I would make the recommendations more trustworthy and serious.}
 
 \subsubsection{Economic memory use} 
-Another solution, which is especially relevant to Windows users due to the unfortunate combination of memory control in this operating system and RStudio \hl{And also just R, right?
-There's got to be a nice reference on this..?}, is simply splitting the two steps performed by \R{clean}, namely producing the \R{rmarkdown} file and rendering it afterwards.
-If the \R{rmarkdown} file is very long, as it will typically be in very large datasets, having this file opened in memory waists precious memory capacities.
-Therefore, we advice users to instead split the two steps.
-This can be done in the following manner:
+Another solution, which is especially relevant to Windows users due to
+the unfortunate combination of memory control in this operating system
+and RStudio %\hl{And also just R, right? There's got to be a nice
+%  reference on this..?}, 
+is simply splitting the two steps performed
+by \R{clean()}, namely producing the \pkg{rmarkdown} file and rendering it
+afterwards. If the \R{rmarkdown} file is very long, as it will
+typically be for very large datasets, having this file opened in memory
+wastes precious memory capacities. Therefore, we advice users to
+instead split the two steps. This can be done in the following manner:
+
 \begin{Schunk}
@@
 \end{Schunk}
-This also deals with the fact that \pkg{cleanR} can produce \R{rmarkdown} files that supersedes the upper size limit of RStudio, which is currently \hl{find number} GBs (using RStudio version 1.0.44).
-\hl{Is this maybe too editor specific?
-On the other hand, a lot of people do use RStudio...}.
+This also deals with the fact that \pkg{cleanR} can produce
+\pkg{rmarkdown} files that supersedes the upper size limit.
+% of RStudio,
+%which is currently \hl{find number} GBs (using RStudio version
+%1.0.44). 
+%\hl{Is this maybe too editor specific? On the other hand, a
+%  lot of people do use RStudio...}.
 
 \subsection{Using cleanR for problem flagging}
-If the data is large, but memory issues and computation time are less of an issue than the human time it takes to look through the data cleaning document, a viable solution might be not to include all information about all variables.
-Or even for more reasonably sized datasets, sometimes a brief overview of the most pressing issues can be useful.
-This can be achieved by using the \R{onlyProblematic} argument in \R{clean()}.
-By specifying \R{onlyProblematic = TRUE}, only variables that raise a flag in the checking steps will be summarized and visualized.
-But perhaps we are not even interested in obtaining general information about these variables, but only in getting a quick overview of the problems they might have.
-This can be done by also controlling the \R{mode} argument:
+If the data is large, but memory issues and computation time are less
+of an issue than the human time it takes to look through the data
+cleaning document, a viable solution might be not to include all
+information about all variables. Or even for more reasonably sized
+datasets, sometimes a brief overview of the most pressing issues can
+be useful. This can be achieved by using the \R{onlyProblematic}
+argument in \R{clean()}. By specifying \R{onlyProblematic = TRUE},
+only variables that raise a flag in the checking steps will be
+summarized and visualized. But perhaps we are not even interested in
+obtaining general information about these variables, but only in
+getting a quick overview of the problems they might have. This can be
+done by also controlling the \R{mode} argument:
 \begin{Schunk}
@@
 
-\section{Appendix NUMTWO}
+\section{Appendix NUMTWO} \label{sec:appendix2}
 \hl{Data cleaning with user supplied extensions here}

### 2016-12-29 latex/article_vol2.tex
Part 3

@@
-\documentclass[article]{jss}
+\documentclass[article,shortnames]{jss}
 \usepackage[utf8]{inputenc}
 \usepackage[nogin]{Sweave}
+\usepackage{natbib}
 \usepackage{pdfpages}
@@
 Now only the checking results are printed, and only for variables where problems were identified.
-An even more minimal output can be generated by also leaving out the checking results - then \R{clean()} essentially just produces a list of the variable names that should be investigated further:
+An even more minimal output can be generated by also leaving out the checking results --- then \R{clean()} essentially just produces a list of the variable names that should be investigated further:
 \begin{Schunk}
@@
 \end{Schunk}
-Of course, this can also be done without generating an overview document, by direct use of the \R{check()} function.
-When called on a \R{data.frame}, this function produces a list (of variables) of lists (of checks) of lists (or rather, \R{checkResult}s).
-Thus, the overall problem status of each variable can easily be unravelled using the list manipulation function \R{sapply()}:
+Of course, this can also be done without generating an overview
+document, by direct, interactive use of the \R{check()} function. When
+called on a \R{data.frame}, this function produces a list (of
+variables) of lists (of checks) of lists (or rather,
+\R{checkResult}s). Thus, the overall problem status of each variable
+can easily be unravelled using the list manipulation function
+\R{sapply()}:
 \begin{Schunk}
@@
 \end{Schunk}
-and we find that only the final variable, \R{var6}, for which all observations have the value \R{"Irrelevant"}, is problem-free. \hl{drop this last bit? too technical?}
+and we find that only the final variable, \R{var6}, for which all
+observations have the value \R{"Irrelevant"}, is problem-free.
 
 \subsection{Include cleanR document in other files}
-Sometimes, a \pkg{cleanR} document might be a useful addition to a more general overview document, including also e.g.
-pairwise association plots, time series plots or exploratory analysis results.
-To this end, it is possible to produce a \pkg{cleanR} document that can readily be included in other \R{rmarkdown} files.
-This is done by using the \R{standAlone} argument in \R{clean}, which removes the preamble from the outputted \R{rmarkdown} file.
-Please note, that it is still necessary to indicate which \R{rmarkdown} type is being created; the pdf and html \R{rmarkdown} styles are unfortunately not identical.
-
-If it is important that the embedded \pkg{cleanR} document can be rendered to either of these two file types, we recommend setting \R{twoCols = FALSE} and \R{output = html} in \R{clean()}, thereby essentially removing almost all output type specific formatting code from the generated \R{rmarkdown} file.
-
-On the other hand, if a pdf document is to be produced, a few extra lines need to be added to the preamble of the master \R{rmarkdown} document - otherwise, the two-column layout code will produce an error.
-The following is an example of how such a master document preamble might look like and how the \R{cleanR\_toyData.Rmd} file can then be included:
+Sometimes, a \pkg{cleanR} document might be a useful addition to a
+more general overview document, including also for example pairwise
+association plots, time series plots, or exploratory analysis
+results. To this end, it is possible to produce a \pkg{cleanR}
+document that can readily be included in other \pkg{rmarkdown}
+files. This is done by using the \R{standAlone} argument in \R{clean()},
+which removes the preamble from the outputted \pkg{rmarkdown}
+file. Please note, that it is still necessary to indicate which
+\R{rmarkdown} type is being created; the pdf and html \pkg{rmarkdown}
+styles are unfortunately not identical.
+
+If it is important that the embedded \pkg{cleanR} document can be
+rendered to either of these two file types, we recommend setting
+\R{twoCols = FALSE} and \R{output = html} in \R{clean()}, thereby
+essentially removing almost all output type specific formatting code
+from the generated \pkg{rmarkdown} file.
+
+On the other hand, if a pdf document is to be produced, a few extra
+lines need to be added to the preamble of the master \pkg{rmarkdown}
+document --- otherwise, the two-column layout code will produce an
+error. The following is an example of how such a master document
+preamble YAML might look like and how the \R{cleanR\_toyData.Rmd} file can
+then be included:
 \begin{Verbatim}
@@
 \end{Verbatim}
-\hl{Use proper formatting here. How do we do non-R code?}
+%\hl{Use proper formatting here. How do we do non-R code?}
 
@@
 \end{Schunk}
-and the more minimal, html-style \R{rmarkdown} file described above can be produced using
+and the more minimal, html-style \pkg{rmarkdown} file described above can be produced using
 \begin{Schunk}
@@
 \end{Schunk}
-\hl{don't end on code.}
-
-\section{Conclusion/Concluding remarks/summary/?}
-\hl{Is this a thing in this journal? Otherwise, we might want to make some final remarks in the previous sections. Feels awkward to end with a bunch of code and some super specific examples...}
+% \hl{don't end on code.}
+
+\section{Concluding remarks}
+%\hl{Is this a thing in this journal? Otherwise, we might want to make some final remarks in the previous sections. Feels awkward to end with a bunch of code and some super specific examples...}
+
+In this paper we have introduced the \proglang{R} package \pkg{cleanR}
+for performing reproducible error detection and data cleaning
+summaries. The package provides a general and extendable framework for
+identifying potential errors and for creating human-readable summary
+documents that will help investigators to identify possible errors in
+the data.
+
+\hl{Noget mere ind her?}
+
+While the the current release is stable, the authors have an interest
+in further developing the functionality of \pkg{cleanR} by providing
+more summary, visual, and check function as part of the default
+package. We are also currently considering adding options to handle
+repeated measurement, where the visualizations --- in particular ---
+might be improved by visualizing measurements over time. In addition,
+an online \pkg{shiny} application where user that are not
+\proglang{R}-savvy can upload their data and get a data cleaning
+document is currently planned.
+
+\nocite{R}
+\nocite{shiny}
+\nocite{cleanR}
+\nocite{rmarkdown}
+\nocite{ggplot2}
+\nocite{plyr}
+\nocite{data.table}
+\nocite{validate}
+\nocite{editrules}
+\nocite{janitor}
+\nocite{DataCombine}
+\nocite{txtplot}
+
+% \bibliographystyle{jss}
+\bibliography{foo}
+
+\hl{Alle pakker. Andet?}
 

### 2016-12-29 latex/codeForArticle.R

@@ -1,25 +1,25 @@
 library(xtable)
 
 #Make summaryFunction overview table
-sumTab <- data.frame(name=allSummaryFunctions()$name, 
-                     description = allSummaryFunctions()$description, 
-                     classes = sapply(allSummaryFunctions()$classes, 
+sumTab <- data.frame(name=allSummaryFunctions()$name,
+                     description = allSummaryFunctions()$description,
+                     classes = sapply(allSummaryFunctions()$classes,
                                       function (x) paste(x, collapse=", ")))
 sumTab <- as.matrix(sumTab)[,-1]
 xtable(sumTab)
 
 #Make visualFunction overview table
-visualTab <- data.frame(name=allVisualFunctions()$name, 
-                     description = allVisualFunctions()$description, 
-                     classes = sapply(allVisualFunctions()$classes, 
+visualTab <- data.frame(name=allVisualFunctions()$name,
+                     description = allVisualFunctions()$description,
+                     classes = sapply(allVisualFunctions()$classes,
                                       function (x) paste(x, collapse=", ")))
 visualTab <- as.matrix(visualTab)[,-1]
 xtable(visualTab)
 
 #Make checkFunction overview table
-checkTab <- data.frame(name=allCheckFunctions()$name, 
-                        description = allCheckFunctions()$description, 
-                        classes = sapply(allCheckFunctions()$classes, 
+checkTab <- data.frame(name=allCheckFunctions()$name,
+                        description = allCheckFunctions()$description,
+                        classes = sapply(allCheckFunctions()$classes,
                                          function (x) paste(x, collapse=", ")))
 checkTab <- as.matrix(checkTab)[,-1]
 xtable(checkTab)
@@ -55,7 +55,7 @@ countZeros <- function(v, ...) {
 }
 countZeros <- summaryFunction(countZeros, description = "Count number of zeros",
                               classes = allClasses())
-summarize(toyData, numericSummaries = c(defaultNumericSummaries))
+summarize(toyData, numericSummaries = c(defaultNumericSummaries()))
 
 
 
@@ -65,7 +65,7 @@ mosaicVisual <- function(v, vnam, doEval) {
     return(eval(thisCall))
   } else return(deparse(thisCall))
 }
-mosaicVisual <- visualFunction(mosaicVisual, 
+mosaicVisual <- visualFunction(mosaicVisual,
[…]

### 2016-12-29 latex/article_vol2.tex
Drafting done

@@
 
-\section{Appendix Something} \label{sec:appendix1}
+\section{Appendix A: cleaning the toyData data frame} \label{sec:appendix1}
 \includepdf[pages=2-4, pagecommand={}, frame=true, noautoscale=true, scale=0.7]{cleanR_toyData.pdf}
 
-\section{Appendix NUMTWO} \label{sec:appendix2}
-\hl{Data cleaning with user supplied extensions here}
+\section{Appendix B: User-extended cleaning of
+  testData} \label{sec:appendix2}
+\includepdf[pages=2-, pagecommand={}, frame=true, noautoscale=true, scale=0.7]{cleanR_testData.pdf}
+%\hl{Data cleaning with user supplied extensions here}
 

### 2016-12-29 R/identifyWhitespace.R

@@ -82,7 +82,7 @@ identifyWhitespaceF <- function(v, nMax) {
 identifyWhitespaceL <- function(v, nMax) {
   v <- na.omit(v)
   v <- haven::as_factor(v)
-  identifyWhitespaceC(v, nMax = nMax)
+  identifyWhitespaceF(v, nMax = nMax)
 }
 
 

### 2016-12-29 latex/codeForArticle.R

@@ -118,4 +118,8 @@ clean(testData, replace = T,
 
 #benchmark CleanR after figure
 
+clean(testData, replace = T,
+      preChecks = c("isKey", "isEmpty", "isID"),
+      allVisuals = "mosaicVisual",
+      labelledCheck = c(defaultLabelledChecks(), "identifyColons"))
 

### 2016-12-29 latex/article_vol2.tex
Updated article

@@
   through each variable in the dataset and look for potential errors
-  --- both in input values and coding --- but that process can be very
-  time-consuming, expensive and error-prone by itself.
+  --- both in input values and codings --- but that process can be very
+  time-consuming, expensive and error-prone in itself.
 
   We describe an \proglang{R} package which implements an extensive
-  and customizeable suite of quality assessment tools to be applied to
+  and customizeable suite of quality assessment tools that can be applied to
   a dataset in order to identify potential problems in its
@@
   by an investigator with an understanding of the variables in the
-  data and the experimental design, but not necessarily knowledge of
-
-  \R{R}. Thereby, \pkg{cleanR} aids the dialogue between data
-  analysists and field experts, while also providing easy
+  data, but not necessarily knowledge of \R{R}. Thereby, \pkg{cleanR} 
+  aids the dialogue between data
+  analysts and field experts, while also providing easy
   documentation of reproducible data cleaning steps and data quality
@@
   more typical \proglang{R} tools for interactive data quality
-  assessment and -cleaning. \hl{Hvad henvises der til med denne
-    sætning?}
+  assessment and -cleaning, where the data inspections all live and die
+  within the \proglang{R} console.
 
@@
 data cleaning and on data wrangling. While a computer should never
-provide an unsupervised decision on what should be done to potential
-errors in the dataset, it can be an extremely useful tool in tracking
-down and flagging potential erroneous data and in providing
-information for humans to easier identify errors based on the context.
+make unsupervised decisions on what should be done to potential
+errors in the dataset, it can still be an extremely useful tool in the 
+data cleaning process. Some errors can be tracked down and flagged by a 
+computer without further ado, while other types of errors need an analytic
+context in order to shine through. But even in the latter case, well-designed
+software can aid the process tremendously by giving the human investigator the 
+information needed for identifying issues.
 
@@
 While these tools attempt to alleviate the ubiquitous ad hoc approach
-to data cleaning they are primarily intended for the data savvy users
-and less on the general researcher with a knowledge about the specific
-field and context of the available data. The \pkg{cleanR} package
+to data cleaning, they are primarily intended for the data savvy users
+and less so for the general researcher with a knowledge about a specific
+field and the context of the available data. The \pkg{cleanR} package
 tries to address this by providing a framework that both allows for
@@
 provides easy-to-use, automated tools for data quality assessment in
-\proglang{R} on which data cleaning decisions can be build. This quality
+\proglang{R} on which data cleaning decisions can be made. This quality
 assessment is presented in an auto-generated overview document,
 readable by data analysts and field experts alike, thereby also
-contributing to a inter-field dialogue about the data at
+contributing to an inter-field dialogue about the data at
 hand. Oftentimes, e.g. distinguishing between faulty codings of a
@@
 This manuscript is structured as follows: First, in Section
-\ref{sec:usingcleanR}, we introduce the representative of the first
+\ref{sec:usingcleanR}, we introduce the main representative of the first
 paradigm, namely the \R{clean()} function, which generates data
@@
 cleaning challenges, we find to be most common. Next, in Section
-\hl{3}, we present the interactive mode of \pkg{cleanR}, as motivated
+\ref{sec:interactiveCleanR}, we present the interactive mode of \pkg{cleanR}, as motivated
 by the second paradigm above. But, as any data analyst knows,
@@
 that cannot be detected by our data checking functions. In Section
-\hl{4}, we turn to the question of how such \pkg{cleanR} extensions
+\ref{sec:extending}, we turn to the question of how \pkg{cleanR} extensions
 can be made, such that they are integrable with the \R{clean()}
 function and with the other tools available in \pkg{cleanR}.  At last,
-in Section 5 \hl{proper ref}, we discuss a number of examples of
+in Section \ref{sec:specificExamples}, we discuss a number of examples of
 specific data cleaning challenges and how \pkg{cleanR} can be used to
@@
 
-\section{Creating a data cleaning overview} \label{sec:example1}\label{sec:usingcleanR}
+\section{Creating a data cleaning overview} 
+\label{sec:usingcleanR}
 
 The \R{clean()} function is the primary workhorse of \pkg{cleanR} and
-this is the only function needed by the user if the standard battery
-of tests are used to output the data cleaning summaries. The data
-cleaning output itself is an overview document, intended for reading
+this is the only function needed if a standard battery
+of tests are used to generate data cleaning summaries. The data
+cleaning summary itself is an overview document, intended for reading
 by humans, in either pdf or html format. Appendix \ref{sec:appendix1}
@@
 shown in Figure~\ref{fig:example1}. \R{toyData} is a very
-small ($15$ rows and $6$ variables), artificial dataset, whose only purpose is to
+small ($15$ rows of $6$ variables), artificial dataset which was created to
 illustrate the main capabilities of \pkg{cleanR}. The following
@@
 
-By default, an \proglang{R} markdown file and rendered pdf overview
+By default, an \proglang{R} markdown file and a rendered pdf overview
 document is produced, saved to the disc (in the working directory) and
@@
 \textit{Variable list}: A table summarizing key features of the
-variable, a figure visualizing its distribution and potentially also a
-list of flagged issues. For instance, in the \R{numeric}-type variable
+variable, a figure visualizing its distribution and a
+list of flagged issues, if any. For instance, in the \R{numeric}-type variable
 \R{var2} from \R{toyData}, \R{clean()} has identified two values that
@@
 for the cleaning overview document to be molded according to the
-user's needs.
-
-The most commonly used arguments are summarized in
+user's needs. The most commonly used arguments are summarized in
 Table~\ref{table.cleanFormals} and they are grouped according to the
@@
 this distinction, a glimpse of the inner structure of \R{clean()} is
-shown in Figure~\ref{figure:cleanStructure}. In the following we give
-examples on how to use these parameters to influence the output of
-\R{clean() }.
+shown in Figure~\ref{figure:cleanStructure}. Below, we present a few
+examples on how to use the arguments from Table \ref{table.cleanFormals}
+ to influence the output of a \R{clean()} call.
 
@@
 
-\smallskip Control input variable and summary\\
+\smallskip Control input variables and summary\\
 \quad \R{useVar} & What variables should be used? & \R{NULL} (corresponding to all variables) \\
@@
 \quad \R{onlyProblematic} & Should only variables flagged as problematic be included in the \textit{Variable list}? & \R{FALSE} \\
-\quad \R{listChecks} & Should an overview of what checks were performed by listed in the \textit{Data cleaning summary}? &  \R{TRUE} \\
+\quad \R{listChecks} & Should an overview of what checks were performed be listed in the \textit{Data cleaning summary}? &  \R{TRUE} \\
+\quad \R{preChecks} & What check functions should be called to determine whether a variable is suitable for summarization, visualization and checking? & \R{c("isKey", "isEmpty")}
 
 \smallskip Control summarize, visualize, and check steps \\
-\quad \R{mode} & What steps should be performed for each variable (out of the three possibilities \textit{summarize}, \textit{visualize}, \textit{check})? & \R{c("summarize", "visualize", "check")} \\
-\quad \R{labelled\_as} & How should variables of class \R{labelled} be handled (as factors, is missing values or by ignoring labels)? & \R{"factor"} \\
+\quad \R{mode} & What steps should be performed for each variable (out of the three possibilities \textit{summarize}, \textit{visualize}, \textit{check})?
+& \R{c("summarize", "visualize", "check")} (corresponding to all three) \\
 \quad \R{smartNum} & Should numerical values with only a few unique
@@
                        should the \R{rmarkdown} file be opened? & \R{TRUE} \\
-\quad \R{replace} & Overwrite an existing markdown file with the same name? & \R{FALSE} \\
+\quad \R{replace} & Overwrite an existing files with the same name? & \R{FALSE} \\
 
@@
     % Place nodes
-    \node [block] (init) {Get next variable and run pre checks};
+    \node [block] (init) {Get next variable and run pre-checks};
     \node [cloud, above of=init] (input) {Input \texttt{data.frame} or \texttt{tibble}};
@@
 
-\caption{Overall illustration of the stages undertaken when running
+\caption{Schematic illustration of the stages undertaken when running
   \R{clean()}. Each variable is checked for eligibility before
@@
 
-For example, to get a summary document in html format that only
-contains the variables with potential problems, and with a limit of
+\subsection{Rolling out the arguments}
+We begin with an example that is intended as an illustration of how \R{clean()} might
+be used in the very first stages of data cleaning, where we are still uncertain
+about exactly what is needed and how much times should be allocated to data
+cleaning. At this stage, what is really needed, if a very rough idea of the extends of
+errors in the dataset. In this scenario, we might wish to obtain a summary document in html format 
+that only contains the variables with potential problems, and with a limit of
 maximum 10 printed potential errors for each variable we can write
@@
 order to force \R{clean()} to overwrite any existing files produced by
-\R{clean()}.
-
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, output="html", onlyProblematic=TRUE, maxProbVals=10,
-        replace=TRUE)
-\end{Sinput}
-\end{Schunk}
-
-The final rendering of the generated markdown file is controlled with
-the \R{render} and \R{openResult} arguments which both default to
+\R{clean()}, as we might run this function a few times in a row before deciding
+exactly which cleaning overview we want. Using the \R{toyData} dataset as a guinea pig, we type:
+
+\begin{Schunk}
+\begin{Sinput}
+> clean(toyData, output = "html", onlyProblematic = TRUE, maxProbVals = 10,
+        replace = TRUE)
+\end{Sinput}
+\end{Schunk}
+
+The final rendering of the generated markdown file is controlled by
+the \R{render} and \R{openResult} arguments, which both default to
 \R{TRUE} \hl{Bliver det ved med at være sandt?}. \R{render} determines
@@
 produces an \proglang{R} markdown file containing the data cleaning
-process information but the markdown file is not rendered into html
-and it is not opened.
+summary information, but without rendering nor opening the markdown file:
 
@@
 
-\pkg{cleanR} uses three
-different types of functions for performing all steps in the above,
+We will now move on to discuss how not only the structure of the cleaning summary is manipulated, but also its very contents.
+This is done through the summarize-visualize-check (SVC) step, as illustrated in Figure \ref{figure:cleanStructure}.
+\pkg{cleanR} uses three different types of functions for performing these steps,
 namely \code{summaryFunction}s, \code{visualFunction}s and
 \code{checkFunction}s.
-By default, \R{clean()} runs all the summary, visualization, and check
-functions currently implemented for each variable type as part of the
-algorithm (see
-Figure~\ref{figure:cleanStructure}). Table~\ref{table.SVCfunctions}
-lists all these functions but we can also use the
+By default, \R{clean()} runs selected summary, visualization and check functions on each variable in the dataset, and the exact choice of these functions depends on the classes of the variables.
+For instance, though detection of outlier values might be interesting for numerical variables, it holds little meaning for factor variables, and therefore, numerical and factor variables need different checks.
+Table~\ref{table.SVCfunctions}
+lists all available SVC functions, but we can also use the
 \R{allSummaryFunctions()}, \R{allVisualFunctions()}, and
-\R{allCheckFunctions()} function in \proglang{R} to print an overview
-list. For example the implemented summary functions are:
+\R{allCheckFunctions()} functions in \pkg{cleanR} to print overview
+lists in \proglang{R}. For example, the implemented \R{summaryFunction}s are:
 
@@
 centralValue   Compute median or mode          character, Date, factor,     
-                                               integer, labelled,
-                                               logical,  numeric
+                                               integer, labelled, logical,  
+                                               numeric
 
@@
 
-minMax         Find minimum and maximum        integer, numeric, Date       
+minMax         Find minimum and maximum        Date, integer, numeric        
                values
 
-quartiles      Compute 1st and 3rd quartiles   integer, numeric
+quartiles      Compute 1st and 3rd quartiles   Date, integer, numeric
 
 uniqueValues   Count number of unique values   character, Date, factor,     
-                                               integer, labelled,
-                                               logical, numeric
+                                               integer, labelled, logical,  
+                                               numeric
 
 variableType   Data class of variable          character, Date, factor,     
-                                               integer, labelled,
-                                               logical, numeric                      
+                                               integer, labelled, logical,  
+                                               numeric                      
 ----------------------------------------------------------------------------
@@
 
-Thus we can see, for example, that the summary table for a
-\R{numeric}, \R{integer}, or \R{Date} variable will include
-information about the minimum and maximum values, while all variable
-classes will produce some kind of information about the central
-tendency (i.e., mode or median).
-
-We can control the default summaries and checks for each variable type
-through the \R{XXXSummaries} and \R{XXXChecks} arguments where \R{XXX}
+Thus we can see, for example, that for \R{numeric}, \R{integer}, or \R{Date} 
+variables, \pkg{cleanR} provides functions for adding summary information about 
+the minimum and maximum values, while all seven variable
+classes dealt with in \pkg{cleanR} have functions for central
+tendency summaries (i.e., mode or median).
+
+We can control what summaries and checks are applied for each variable type
+through the \R{XXXSummaries} and \R{XXXChecks} arguments, where \R{XXX}
 represents a variable type, e.g., \R{factorChecks} for factors,
 \R{numericChecks} for numeric variables, etc. These arguments accept a
-vector of summary or check functions that should be run for a
-particular variable type and the default values for each of these
-arguments (i.e., the functions from Table~\ref{table.SVCfunctions}
-that are applied) can be obtained through the \R{defaultXXXChecks()}
+vector of summary- or check function names that should be run for a
+particular variable type. The default values for each of these
+arguments can be obtained through the \R{defaultXXXChecks()}
 and \R{defaultXXXSummaries()} functions, where \R{XXX} again refers to
@@
 \begin{Sinput}
-> clean(toyData, output="html",  
-        characterSummaries = c("variableType", "centralValue"),
+> clean(toyData, characterSummaries = c("variableType", "centralValue"),
         factorSummaries = c("variableType", "centralValue"),
@@
 \begin{Sinput}
-> clean(toyData, output="html",  
-        allSummaries = c("variableType", "centralValue"))
+> clean(toyData, allSummaries = c("variableType", "centralValue"))
 \end{Sinput}
@@
 Now, if we only wanted to apply the identify whitespace function for
-factor variables then we would need to set the \R{factorChecks} accordingly
-
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, output="html",  
-        factorChecks = c("identifyWhitespace"))
-\end{Sinput}
-\end{Schunk}
-
-or we could remove checks for factor altogether by setting the
-corresponding argument to \R{NULL} in which case factor variables will
-not be checked for any potential errors.
-
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, output="html", factorChecks = NULL)
+factor variables, then we would need to set the \R{factorChecks} accordingly
+
+\begin{Schunk}
+\begin{Sinput}
+> clean(toyData, factorChecks = c("identifyWhitespace"))
+\end{Sinput}
+\end{Schunk}
+
+or we could remove checks for factors altogether by setting the
+corresponding argument to \R{NULL}, in which case factor variables will
+not be checked for any potential errors:
+
+\begin{Schunk}
+\begin{Sinput}
+> clean(toyData, factorChecks = NULL)
+\end{Sinput}
+\end{Schunk}
+
+As with \R{summaryFunction}s, a complete list of available \R{checkFunction}s is obtained by calling 
+\R{allCheckFunctions()}. Note however, that \R{checkFunction}s have a usage beyond the 
+\R{XXXChecks} arguments, namely in the \textit{pre-check} stage. In this stage, it is determined 
+whether or not each variable is suitable for the SVC-steps. The functions used in the pre-check stage
+should be \R{checkFunction}s that are applicable to all variable classes. The default pre-checks, the functions
+\R{isKey} and \R{isEmpty}, check whether a variable has unique values for all observations or only a single value for 
+all observations, respectively. If one of these statements are true, the variable will not be subjected to the SVC steps. 
+We can allow empty variables to move on to the SVC step by only checking for keys in the pre-check step:
+
+\begin{Schunk}
+\begin{Sinput}
+> clean(toyData, preChecks = "isKey")
+\end{Sinput}
+\end{Schunk}
+
+Note that the data visualizations in the cleaning summary are also controllable, though only a single function can be provided for all variable types.
+If for instance, we wish to change the visualizations from the default \pkg{ggplot2} style histograms and barplots to base \proglang{R} histograms and barplots, we type
+\begin{Schunk}
+\begin{Sinput}
+> clean(toyData, allVisuals = "basicVisual")
 \end{Sinput}
@@
 
-As indicated in Figure~\ref{figure:cleanStructure} there are two stages
-where \R{clean()} applies function(s) to each of the variables: 
+In summary, and as indicated in Figure~\ref{figure:cleanStructure}, there are two stages
+where \R{clean()} applies functions to each of the variables: 
 \begin{enumerate}
-\item In the precheck functions
+\item In the pre-check stage
 \item As part of the summarize/visualize/check (SVC) steps
@@
 tweak them to modify the data cleaning outputs. However, if for
-instance the dataset at hand requires completely different
-visualizations then more control is needed, and
-Section~\ref{sec:extending} explains how to modify, and expand the
-possibilities by producing new functions quite easily.
+instance the dataset at hand requires new, additional checks, then more control is needed, and
+Section~\ref{sec:extending} explains how to modify and expand the
+possibilities by producing new summary-, visual- and check functions.
 
 \section{Using cleanR interactively}
+\label{sec:interactiveCleanR}
+
 While overview documents are great for presenting and documenting the
-data cleaning checks, it may be natural to work more interactively
-through the data cleaning process. \pkg{cleanR} also
-provides more standard \proglang{R} interactive tools, such as
+data cleaning checks, it may be useful to be able to work more interactively
+through the data cleaning process as well. Aside from the \R{clean()} function 
+presented above, \pkg{cleanR} also provides more standard \proglang{R} interactive tools, such as
 functions that print results to the console or returns the information
@@
 vector of the check functions that should be applied. Recall that
-Table~\ref{table.SVCfunctions} or the \R{allCheckFunctions()} provides
+Table~\ref{table.SVCfunctions} or an \R{allCheckFunctions()} call provide
 overviews of the available check functions. 
+Moving forward, we limit the numeric checks to only identify miscoded 
+missing values:
+
 \begin{Schunk}
@@
 
-An equivalent way to call only a single, specific \R{checkFunction}
-such as \R{identifyMissing} (see \ref{table.SVCfunctions} for a list
-of check functions) is by using it directly on the variable, e.g.,
+An equivalent way to call only a single, specific \R{checkFunction},
+such as \R{identifyMissing}, is by using it directly on the variable, e.g.,
 \begin{Schunk}
@@
 by providing the relevant \R{XXXSummaries} arguments. Setting the
-\R{numericSummaries} argument we can control the summary output by
-providing a vector of functions to run for a particular summary. To
-only get the minimum and maximum we set
-\R{numericSummaries=c("minMax")}.
-
-\begin{Schunk}
-\begin{Sinput}
-> summarize(toyData$var2, numericSummaries=c("minMax"))
+\R{numericSummaries} argument, we can control the summary output by
+providing a vector of function names to run for a particular summary. To
+only get the median, the minimum and the maximum we set
+\R{numericSummaries=c("centralValue", "minMax")}:
+
+\begin{Schunk}
+\begin{Sinput}
+> summarize(toyData$var2, numericSummaries = c("centralValue", "minMax"))
 \end{Sinput}
@@
      Feature         Result  
-[1,] "Min. and max." "1; 999"
+[1,] "Median"        "4.5"   
+[2,] "Min. and max." "1; 999"
 \end{Soutput}
@@
 % $
+
+Note that the \R{summarize()}, \R{check()} and \R{visualize()} functions are also available interactively for full datasets, by calling e.g.
+\R{summarize(toyData)}.
+However, this produces an extensive amount of output in the console, and therefore, we generally do not recommend it, unless working with very small datasets or subsets of datasets.
 
@@
 \\
-Output (required) & A list with entries \newline \R{\$feature} - a label for the summary value (as character string) \newline \R{\$result} - the result of the summary (as character string) & A character string with \R{R} code for producing a plot.
+Output (required) & A list with entries: \newline \R{\$feature} - a label for the summary value (as character string); \newline \R{\$result} - the result of the summary (as character string) & A character string with \R{R} code for producing a plot.
 This code should be standalone, i.e.
 should include the data if necessary.
-& A list with entries \newline \R{\$problem} - a logical identifying whether an issue was found \newline \R{\$message} - a character string (possibly empty) decribing the issue that was found, properly escaped and ready for use in \R{rmarkdown} \\
-Output (recommended) & A \R{summaryResult} object (i.e.
+& A list with entries: \newline \R{\$problem} - a logical identifying whether an issue was found; \newline \R{\$message} - a character string (possibly empty) decribing the issue that was found, properly escaped and ready for use in \R{rmarkdown} \\
+Output (recommended) & A \R{summaryResult} object, i.e.
 an attributed list with entries \R{\$feature}, \R{\$result} and \R{\$value}, the latter being the values from \R{\$result} in their original format).
@@
 & If \R{doEval} is \R{FALSE}: \newline   A text string with \R{R} code, as described above.
-\end{tabular} & A \R{checkResult} object (an attributed list with entries \R{\$problem}, \R{\$message} and \R{\$problemValues}, the latter being either \R{NULL} or the problem causing values, as they were found in \R{v}, whichever is relevant.
+\end{tabular} & A \R{checkResult} object, i.e.
+an attributed list with entries \R{\$problem}, \R{\$message} and \R{\$problemValues}, the latter being either \R{NULL} or the problem causing values, as they were found in \R{v}, whichever is relevant.
 \\
@@
 \caption{Reference information for creating new functions to be used
-  as part of the summarize, visualize, and check steps. The three
-  columns correspond to each of the three function types.}
+  as part of the summarize-, visualize-, and check steps. The three
+  columns correspond to each of the three function types.
+\hl{I think the formatting here is still sort of chaotic.
+Maybe include horizontal lines everywhere?
+Is it allowed in JSS?
+It's impossible to read what's in which row right now.}}
 \label{table.functionTypes}
@@
 
-\pkg{cleanR} can be used as a user-friendly, self-contained package as
-shown in the previous section. However, \pkg{cleanR} is fully
+\pkg{cleanR} can be used as a user-friendly, self-contained package, as
+shown in the previous sections. However, \pkg{cleanR} is fully
 customizable and \R{clean()} is mainly a tool for formatting the
@@
 --- and user made functions are easily added to the
-summarize/visualize/check-function arguments as discussed previously
-(e.g., \R{factorSummaries}, or \R{numericChecks}). However, the
-functions used in the summarize/visualize/check steps need to adhere
-to a specific structure to be called from these three steps, and
+summarize/visualize/check-function arguments, as discussed previously. 
+However, the functions used in the SVC steps must adhere
+to specific structures in order to be called from these three steps, and
 therefore, we will now present how \R{summaryFunction}s,
 \R{visualFunction}s and \R{checkFunction}s are made.
-%\hl{Mention something about the interactive mode here as well?}
-
-This section consists of two parts. First, we describe how the
-\R{clean()} function can be extended by adding custom prechecks,
-summaries, checks and visualizations. In order to do this, one needs
-to create functions with a specific structure required by
-\pkg{cleanR}. This can be done with different levels of strictness. If
-the custom functions are only to be used part of the corresponding
-arguments of \R{clean()} then only the input/output structures of the
-functions need to be fixed.
-
+
+This section consists of two parts. First, we describe how custom \R{summaryFunction}s, 
+\R{visualFunction}s and \R{checkFunctions} can be made, including an introduction to a 
+few convenient tools available in \pkg{cleanR} for aiding SVC-function construction. 
+Table \ref{table.functionTypes} serves as an overview of the internals of these three
+central function types. Secondly, we turn to a 
+worked example of how to use custom made functions in practice. Here,
+four new SVC functions are defined and used, both interactively and in \R{clean()}.
+
+\subsection{Function templates}
+\label{sec:functionTemplates}
+In order to construct a summary-, visual or check function, one needs
+to create a new function with a specific structure.  This can be done with 
+different levels of strictness. If
+the new custom function is only to be used as part of the SVC steps in \R{clean()},
+then only the input/output structure of the function needs special attention.
 However, new user-defined functions can also be registered locally to
 be part of the full machinery of \pkg{cleanR}, and these function will
-be recognized and behave in the same way as the build-in functionson
+be recognized and behave in the same way as the build-in functions in
 \pkg{cleanR}. The presentation below is given in the format of
@@
 requirements, while also presenting the "full" versions of the
-function types. We end this section with a 
-%After this abstract, templatic overview of the internals of the three
-%SVC functions is given, we turn to a 
-worked example of how to use custom made functions in practice, where
-four new summarize/visualize/check functions are defined and used,
-both interactively and in \R{clean()}.
-
-\subsection{Function templates}
-\label{sec:functionTemplates}
+function types.
+
 % \hl{metatext, mention Table \ref{table.functionTypes} again.}
@@
 functions by converting it to a proper \code{summaryFunction}
-object. This is done by calling the \R{summaryFunction} function with
-the user-defined function as first argument, and additional arguments
-\R{description} which adds explanatory flavour text as an attribute of
-the function, and \R{classes} which determines which classes the
-user-defined function should be applied to (also stored as an
-attribute).
-\begin{Verbatim}
+object. This is done by calling the \R{summaryFunction()} creator with
+the user-defined function as the first argument, and additional arguments
+\R{description} (an explanatory text which will be added to the attributes of
+the function), and \R{classes} (a vector of variable classes the
+user-defined function is intended to be applied to, also stored as an
+attribute). In other words, a call on the following form should be made:
+\begin{Schunk}
+\begin{Sinput}
 mySummaryFunction <- summaryFunction(mySummaryFunction,
@@
   classes = c([vector of data types that the function is intended for]))
-\end{Verbatim}
+\end{Sinput}
+\end{Schunk}
 which adds the new function to the output of an
@@
 function (which in this case is \R{"mySummaryFunction"}) will be filled in and
-stored as an attribute. What happens if the \R{classes} argument is
+stored under the \R{description} attribute. What happens if the \R{classes} argument is
 not specified depends on the type of \R{mySummaryFunction}. If
@@
 \R{classes} is left unspecified, the attribute will simply be
-empty. Note that the helper function \R{allClasses()} might be useful
+left empty. Note that the helper function \R{allClasses()} might be useful
 for filling out the \R{classes} argument, as it simply lists all
@@
 
-If we run the \R{allSummaryFunctions()} function we can see that the
-new function has been registered (output slightly appreviated):
-
-\begin{Schunk}
-\begin{Sinput}
-> allSummaryFunctions()
-\end{Sinput}
-\begin{Soutput}
-----------------------------------------------------------------------------
-name                description                                      classes
-------------------- --------------------------------------------------------
-mySummaryFunction   [Text describing what the summaryFunction does]  numeric
-.
-.
-.
-\end{Soutput}
-\end{Schunk}
+%Jeg foreslår at droppe nedenstående, da jeg synes det bliver lidt søgt i forhold 
+%til at det kun er beskrevet i pseudokode. (fx. er classes = numeric meget arbitrært)
+%If we run the \R{allSummaryFunctions()} function we can see that the
+%new function has been registered (output slightly appreviated):
+
+%\begin{Schunk}
+%\begin{Sinput}
+%> allSummaryFunctions()
+%\end{Sinput}
+%\begin{Soutput}
+%----------------------------------------------------------------------------
+%name                description                                      classes
+%------------------- --------------------------------------------------------
+%mySummaryFunction   [Text describing what the summaryFunction does]  numeric
+%.
+%.
+%.
+%\end{Soutput}
+%\end{Schunk}
 
@@
 \pkg{rmarkdown} files. We recommend using the following structure:
-\begin{Verbatim}
+\begin{Schunk}
+\begin{Sinput}
 myVisualFunction <- function(v, vnam, doEval) {
   thisCall <- call("[the name of the function used to produce the plot]",
-    v, [additional arguments to the plotting function])
+    v, [additional arguments for the plotting function])
   if (doEval) {
@@
 }
-\end{Verbatim}
+\end{Sinput}
+\end{Schunk}
 
@@
 are mandatory, so they must be left as is, even if they are not in
-use. See Table~\ref{table.functionTypes}.
-
-Similarly to the summary function, we call \R{visualFunction()} to
-register our newly created function. Like before, it accepts the
-user-defined function as the first argument, and two additional
-arguments description which adds explanatory flavour text as an
-attribute of the function, and a vector which determines which classes
-the user-defined function should be applied to (which is also stored
-as an attribute).
-
-\begin{Verbatim}
+use (c.f. Table~\ref{table.functionTypes}).
+
+As with the summary function, we call \R{visualFunction()} to
+register our newly created function:
+%. Like before, it accepts the
+%user-defined function as the first argument, and two additional
+%arguments description which adds explanatory flavour text as an
+%attribute of the function, and a vector which determines which classes
+%the user-defined function should be applied to (which is also stored
+%as an attribute).
+\begin{Schunk}
+\begin{Sinput}
 myVisualFunction <- visualFunction(myVisualFunction,
@@
   )
-\end{Verbatim}
-
-Again, registering the visualization function allows it to be used
-without manually adding it to the list of visualizations.
+\end{Sinput}
+\end{Schunk}
+
+Now, \R{myVisualFunction()} will be available in a \R{allVisualFunctions()}
+call, just like the two build-in \R{visualFunction}s, \R{standardVisual}
+and \R{basicVisual}.
 
@@
 it tries to identify problematic values in the variable (as
-e.g., \R{identifyMissing()} does) or it performs a check concerning the
+e.g., \R{identifyMissing()} does), or it performs a check concerning the
 variable as a whole (e.g. the functions used for prechecks and the
@@
 should be expanded to follow a slightly more complicated structure:
-\begin{Verbatim}
+\begin{Schunk}
+\begin{Sinput}
 myProbValCheckFunction <- function(v, nMax, maxDecimals, ...) {
@@
 )
-\end{Verbatim}
-For this template, the argument \R{maxDecimals} is not in use. This
-argument is used to round off the \R{problemValues} passed to
-\R{messageGenerator()}, if they are numerical.  This is done by
-substituting the \R{problemStatus} assignment above with the following
-code: \hl{Denne foregående sætning forstår jeg ikke helt. Måske skal
-  messagegenerator have en sætning eller to med på vejen? (Og det spiller lidt sammen med nedenstående)}
-\begin{Schunk}
-\begin{Sinput}
-problemStatus <- list(problem = problem, 
-   problemValues = round(problemValues, maxDecimals))
-\end{Sinput}
-\end{Schunk}
-
-It is worth mentioning the 
-%Another noteworthy component of the template is the usage of the
-helper function \R{messageGenerator()}, which aids in consistent
-styling of all \code{checkFunction} messages. This function simply
+\end{Sinput}
+\end{Schunk}
+One comment should be devoted to the helper function, \R{messageGenerator}. 
+This function's sole purpose is aiding consistent
+styling of all \code{checkFunction} messages. The function simply
 pastes together the \code{problemMessage} and the
@@
 is the easiest solution.
+
+In the template above, the argument \R{maxDecimals} is not in use. This
+argument should be used to round off the \R{problemValues} passed to
+\R{messageGenerator()}, if they are numerical.  This can be done by
+adding an extra line of code after defining \R{problemStatus} in the template 
+above:
+\begin{Schunk}
+\begin{Sinput}
+myProbValCheckFunction <- function(v, nMax, maxDecimals, ...) {
+ ...
+  problemStatus <- list(problem = problem, 
+    problemValues = problemValues)
+
+  if (!is.null(problemValues)) {
+    problemStatus$problemValues <- round(problemValues, maxDecimals)
+  }
+ ...  
+}
+\end{Sinput}
+\end{Schunk}
+%$
+Now, problematic values will be rounded in the outputted message, while they
+will still appear in their original format under the entry \R{\$problemValues} %$
+in the outputted \R{checkResult}.
+
+\hl{Jeg forstår ikke helt, hvad der var uklart her ang. messageGenerator. Jeg har nu
+stort set bare byttet om på rækkefølgen af de to sidste paragraphs her, hjælper
+det?}
 
@@
 First, let's define the \R{isID()} function. As this function is not
-supposed to list problematic values in the variable, it falls within
+supposed to list problematic values, it falls within
 the category of \R{checkFunctions} represented by
@@
 function as a precheck-function in \R{clean()}, so we will leave it as
-is and move on to the next function, namely \R{mosaicVisual}.
+is and move on to the next function, namely \R{mosaicVisual()}.
 
 \subsubsection{mosiacVisual --- a new visualFunction}
-\proglang{R} contains a function, \R{mosiacplot}, which produces
+\proglang{R} contains a function, \R{mosiacplot()}, which produces
 mosaic plots. We intend to use this existing high-level plotting
@@
 \end{Schunk}
-This function can now be called directly or used in \R{clean()} and we
-will show an example below. Depending on the \R{doEval} argument
-\hl{skal doEval have en default value ovenfor?}, either a text string
+This function can now be called directly or used in \R{clean()}, as will
+presented in an example below. Depending on the \R{doEval} argument, either a text string
 with code or a plot is produced. The plot resulting from the following
@@
 \R{visualFunction}, it is not yet truly one and therefore, it will not
-appear in a \R{allVisualFunctions()} call. In order to get this
+appear in an \R{allVisualFunctions()} call. In order to get this
 functionality, we need to change its object class. This can be done by
@@
 of all the seven variable classes addressed in \pkg{cleanR}. Note that
-if \R{mosaicVisual} were an S3 generic function, this argument could
+if \R{mosaicVisual()} were an S3 generic function, this argument could
 have been left as \R{NULL} and then the classes for which methods are
@@
 \end{itemize}
-These options are specified as follows:
+These options are specified as follows: \hl{Jeg synes ikke testData egner sig her.
+Der er ingen koloner og CPRvar er et mærkeligt, internt navn, hvis ikke man kender danske CPR-numre.
+Langt de fleste variable har desuden ingen nuller.
+Måske skal der laves et nyt datasæt?}
 \begin{Schunk}
@@
       allVisuals = "mosaicVisual",
-      # Add the new checks for specific data types
+      # Add the new summary for specified data types
       characterSummaries = c(defaultCharacterSummaries(), "countZeros"),
@@
       integerSummaries = c(defaultIntegerSummaries(), "countZeros"),
+      # Add the new check for specific data types
       characterChecks = c(defaultCharacterChecks(), "identifyColons"),
@@
 
-\section{Using cleanR in specific situations}
-Finally, we will present a few examples of how to make \pkg{cleanR}
+\section{Zooming in on cleaning challenges}
+\label{sec:specificExamples}
+\hl{Lidt poppet titel, men den gamle ("Using cleanR in specific situations")
+fik mig til flabel at tænke "i modsætning til uspecifikke situationer?". Jeg 
+er dog meget åben for en bedre titel.}
+
+Finally, we present a few examples of how to make \pkg{cleanR}
 solve specific issues related to data cleaning. First, we discuss the
@@
 variables, creation of the \pkg{rmarkdown} document might be quite
-slow, while a large number of observations generally will affect the
+slow, while a large number of observations will generally affect the
 rendering time of the document. In this section, we give a few
@@
 the unfortunate combination of memory control in this operating system
-and RStudio %\hl{And also just R, right? There's got to be a nice
-%  reference on this..?}, 
+and RStudio, \hl{And also just R, right? Du har udkommenterert denne kommentar,
+men jeg mener stadig at det er tilfældet, og jeg mener desuden ikke vi kan skrive 
+det uden en præcision af hvad det går ud på. Hvad tænker du?}
 is simply splitting the two steps performed
@@
 afterwards. If the \R{rmarkdown} file is very long, as it will
-typically be for very large datasets, having this file opened in memory
+typically be for very large datasets, keeping this file open in memory
 wastes precious memory capacities. Therefore, we advice users to
@@
 This also deals with the fact that \pkg{cleanR} can produce
-\pkg{rmarkdown} files that supersedes the upper size limit.
+\pkg{rmarkdown} files that supersedes the upper size limit. \hl{dette giver
+ingen mening uden en præcision? "upper size limit" af hvad? Se udkommenteret 
+kommentar.}
 % of RStudio,
@@
 \end{Schunk}
-and we find that only the final variable, \R{var6}, for which all
-observations have the value \R{"Irrelevant"}, is problem-free.
-
-\subsection{Include cleanR document in other files}
+and we find that only a single variable in \R{toyData}, \R{var6} (for which all
+observations have the value \R{"Irrelevant"}), is problem-free.
+
+\subsection{Including cleanR documents in other files}
 Sometimes, a \pkg{cleanR} document might be a useful addition to a
@@
 \end{Schunk}
+Note that with the latter option, no special YAML preamble is needed in the \R{rmarkdown} document.
 % \hl{don't end on code.}
@@
 \section{Concluding remarks}
+\label{conclusion}
 %\hl{Is this a thing in this journal? Otherwise, we might want to make some final remarks in the previous sections. Feels awkward to end with a bunch of code and some super specific examples...}
@@
 
-\hl{Noget mere ind her?}
-
 While the the current release is stable, the authors have an interest
@@
 might be improved by visualizing measurements over time. In addition,
-an online \pkg{shiny} application where user that are not
+an online \pkg{shiny} application where users that are not
 \proglang{R}-savvy can upload their data and get a data cleaning
@@
 
-\section{Appendix B: User-extended cleaning of
-  testData} \label{sec:appendix2}
+\section{Appendix B: User-extended cleaning of  testData} \label{sec:appendix2}
 \includepdf[pages=2-, pagecommand={}, frame=true, noautoscale=true, scale=0.7]{cleanR_testData.pdf}
@@
 \end{document}
-\R{characterChecks} & What checks should be performed on character variables? & \hl{what to write here?} \\
-\R{factorChecks} & What checks should be performed on character variables? & \hl{what to write here?} \\
-\R{numericChecks} & What checks should be performed on factor variables? & \hl{what to write here?} \\
-\R{integerChecks} & What checks should be performed on integer variables? & \hl{what to write here?} \\
-\R{logicalChecks} & What checks should be performed on logical variables? & \hl{what to write here?} \\ 
-\R{dateChecks} & What checks should be performed on Date variables? & \hl{what to write here?} \\
-\R{allChecks} & What checks should be performed on all variables (overwriting variable type specific settings if not \R{NULL})? & \R{NULL} \\

### 2016-12-30 latex/article_vol2.tex
Halfway done with TeX fixes

@@
 
+\renewcommand\topfraction{.9}
+\renewcommand\textfraction{.1}
+\renewcommand{\floatpagefraction}{.6}
+
+\renewcommand{\topfraction}{0.85}
+\renewcommand{\bottomfraction}{0.85}
+\renewcommand{\textfraction}{0.15}
+\renewcommand{\floatpagefraction}{0.7}
+
 %Virker ikke:
@@
 
-\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
-\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}m{#1}}
+\newcolumntype{L}[1]{>{\raggedright\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}
+\newcolumntype{C}[1]{>{\centering\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}
+\newcolumntype{R}[1]{>{\raggedleft\let\newline\\\arraybackslash\hspace{0pt}}p{#1}}
 
@@
     \path [line,dashed] (input) -- (init);
+    \path [line] (stop) -- (render);
 %    \path [line,dashed] (system) -- (init);
@@
 
-\begin{table}[ht]
-\scriptsize
+\pkg{cleanR} can be used as a user-friendly, self-contained package, as
+shown in the previous sections. However, \pkg{cleanR} is fully
+customizable and \R{clean()} is mainly a tool for formatting the
+results from various checking-, summary- and visualization
+functions. Thus, the actual work underlying a \pkg{cleanR} output file
+can be anything --- depending on the arguments given to \R{clean()}
+--- and user made functions are easily added to the
+summarize/visualize/check-function arguments, as discussed previously. 
+However, the functions used in the SVC steps must adhere
+to specific structures in order to be called from these three steps, and
+therefore, we will now present how \R{summaryFunction}s,
+\R{visualFunction}s and \R{checkFunction}s are made.
+
+This section consists of two parts. First, we describe how custom \R{summaryFunction}s, 
+\R{visualFunction}s and \R{checkFunctions} can be made, including an introduction to a 
+few convenient tools available in \pkg{cleanR} for aiding SVC-function construction. 
+Table \ref{table.functionTypes} serves as an overview of the internals of these three
+central function types. Secondly, we turn to a 
+worked example of how to use custom made functions in practice. Here,
+four new SVC functions are defined and used, both interactively and in \R{clean()}.
+
+\begin{table}[htbp]
+\footnotesize
+\bgroup
+\def\arraystretch{1.8}% 
 \begin{tabular}{L{0.2\linewidth}L{0.2\linewidth}L{0.2\linewidth}L{0.2\linewidth}}
@@
 \hline
-Input (required) & \R{v} - a variable vector \newline \R{...}  &  \R{v} - a variable vector \newline \R{vnam} - the variable name (as character string) \newline \R{doEval} - a logical (\R{TRUE}/\R{FALSE}) controlling the output type of the function & \R{v} - a variable vector \newline \R{nMax} - an integer (or \R{Inf}), controlling how many problematic values are printed, if relevant \newline \R{...}  \\
+\vspace{0pt} Input (required) & \R{v} - a variable vector \newline \R{...}  &  \R{v} - a variable vector \newline \R{vnam} - the variable name (as character string) \newline \R{doEval} - a logical (\R{TRUE}/\R{FALSE}) controlling the output type of the function & \R{v} - a variable vector \newline \R{nMax} - an integer (or \R{Inf}), controlling how many problematic values are printed, if relevant \newline \R{...}  \\
 Input (optional) &  \R{maxDecimals} - number of decimals printed in outputted numerical values.  & - &  \R{maxDecimals}  - number of decimals printed in outputted numerical values.  \\
@@
 an attributed list with entries \R{\$feature}, \R{\$result} and \R{\$value}, the latter being the values from \R{\$result} in their original format).
-& \begin{tabular}{p{0.4\linewidth} | p{0.4\linewidth}} If \R{doEval} is \R{TRUE}: \newline A plot that will be opened by the graphic device in \R{R}.
-& If \R{doEval} is \R{FALSE}: \newline   A text string with \R{R} code, as described above.
-\end{tabular} & A \R{checkResult} object, i.e.
+& \bgroup
+\def\arraystretch{1.8}% 
+\hspace*{-.2cm}\begin{tabular}[t]{p{0.45\linewidth} | p{0.45\linewidth}} \vspace*{-.9cm}If \R{doEval} is \R{TRUE}: \newline A plot that is opened by the graphic device in \R{R}.
+& \vspace*{-.9cm}If \R{doEval} is \R{FALSE}: \newline   A text string with \R{R} code, as described above.
+\end{tabular} \egroup  & A \R{checkResult} object, i.e.
 an attributed list with entries \R{\$problem}, \R{\$message} and \R{\$problemValues}, the latter being either \R{NULL} or the problem causing values, as they were found in \R{v}, whichever is relevant.
 \\
-Tools available for producing the function & \R{summaryResult()} & - & \R{messageGenerator()} \newline \R{checkResult()} 
+Tools available for producing the function & \R{summaryResult()} & --- & \R{messageGenerator()} \newline \R{checkResult()} 
 \end{tabular}
+\egroup
 \caption{Reference information for creating new functions to be used
-  as part of the summarize-, visualize-, and check steps. The three
+  as part of the \mbox{summarize-,} visualize-, and check steps. The three
   columns correspond to each of the three function types.
-\hl{I think the formatting here is still sort of chaotic.
+%\hl{I think the formatting here is still sort of chaotic.
 Maybe include horizontal lines everywhere?
 Is it allowed in JSS?
-It's impossible to read what's in which row right now.}}
+It's impossible to read what's in which row right now.}
+}
 \label{table.functionTypes}
 \end{table}
-
-\pkg{cleanR} can be used as a user-friendly, self-contained package, as
-shown in the previous sections. However, \pkg{cleanR} is fully
-customizable and \R{clean()} is mainly a tool for formatting the
-results from various checking-, summary- and visualization
-functions. Thus, the actual work underlying a \pkg{cleanR} output file
-can be anything --- depending on the arguments given to \R{clean()}
---- and user made functions are easily added to the
-summarize/visualize/check-function arguments, as discussed previously. 
-However, the functions used in the SVC steps must adhere
-to specific structures in order to be called from these three steps, and
-therefore, we will now present how \R{summaryFunction}s,
-\R{visualFunction}s and \R{checkFunction}s are made.
-
-This section consists of two parts. First, we describe how custom \R{summaryFunction}s, 
-\R{visualFunction}s and \R{checkFunctions} can be made, including an introduction to a 
-few convenient tools available in \pkg{cleanR} for aiding SVC-function construction. 
-Table \ref{table.functionTypes} serves as an overview of the internals of these three
-central function types. Secondly, we turn to a 
-worked example of how to use custom made functions in practice. Here,
-four new SVC functions are defined and used, both interactively and in \R{clean()}.
 

### 2016-12-30 latex/article_vol2.tex
Added example data + documentation

@@
 \end{itemize}
-These options are specified as follows: \hl{Jeg synes ikke testData egner sig her.
+These options are specified as follows: %\hl{Jeg synes ikke testData egner sig her.
 Der er ingen koloner og CPRvar er et mærkeligt, internt navn, hvis ikke man kender danske CPR-numre.
@@
 \begin{Sinput}
-> data(testData) 
-> clean(testData, 
+> data(exampleData) 
+> clean(exampleData, 
       # Add the new pre-check function
@@
 \section{Appendix B: User-extended cleaning of  testData} \label{sec:appendix2}
-\includepdf[pages=2-, pagecommand={}, frame=true, noautoscale=true, scale=0.7]{cleanR_testData.pdf}
+\includepdf[pages=2-, pagecommand={}, frame=true, noautoscale=true, scale=0.7]{cleanR_exampleData.pdf}
 %\hl{Data cleaning with user supplied extensions here}

### 2016-12-30 R/clean.R

@@ -8,9 +8,11 @@
 #' includes summaries and visualizations of each variable in the
 #' dataset.
 #'
-#' The checks done for each class of variable is chosen by the
-#' summarize/visualize/check functions ... XXX
-#'
+#' For each variable, a set of pre-check (controlled by the
+#' \code{preChecks} argument) is first run and then then a battery of
+#' functions are applied depending on the variable class.  For each
+#' variable type the summarize/visualize/check functions are applied
+#' and and the results are written to an R markdown file.
 #'
 #' @param data The dataset to be checked. This dataset should be of class \code{data.frame},
 #' \code{tibble} or \code{matrix}. If it is of classs \code{matrix}, it will be converted to a
@@ -33,7 +35,8 @@
 #' problematic in the check step will be included in the variable list.
 #'
 #' @param labelled_as A string explaining the way to handle labelled vectors.
-#' Currently \code{"factor"} (the default) is the only possibility.
+#' Currently \code{"factor"} (the default) is the only possibility and the argument
+#' is not currently in use.
 #'
 #' @param mode Vector of tasks to perform among the three categories "summarize", "visualize" and "check".
 #' The default, \code{c("summarize", "visualize", "check")}, implies that all three steps are
@@ -162,8 +165,6 @@
 #' of a data cleaning overview document) is the reason for running the function.
 #'
 #' @examples
-#'
-#'
 #' data(testData)
 #' data(toyData)
 #'
@@ -179,9 +180,35 @@
 #' clean(testData)
 #' }
 #'
-#'#Add user defined check-function to the checks performed on character variables:
+#' # Overwrite any existing files generated by clean
 #' \dontrun{
-#' clean(testData, characterChecks=c(defaultCharacterChecks(), "characterFoo"))
+#' clean(testData, replace=TRUE)
+#' }
+#'
+#' # Only include problematic variables in the output document
+#' \dontrun{
+#' clean(testData, replace=TRUE, onlyProblematic=TRUE)
+#' }
+#'
+#' # Add user defined check-function to the checks performed on character variables:
+#' # Here we add functionality to search for the string wally (ignoring case)
+#' \dontrun{
+#' wheresWally <- function(v, ...) {
+#'      res <- grepl("wally", v, ignore.case=TRUE)
+#'      problem <- any(res)
+#'      message <- "Wally was found in these data"
+#'      checkResult(list(problem = problem,
+#'                       message = message,
+#'                       problemValues = v[res]))
+#' }
+#'
[…]

### 2016-12-30 R/cleanR-package.R

@@ -57,3 +57,99 @@
 #' data(toyData)
 #'
 "toyData"
+
+
+#' Example data with zero-inflated variables
+#'
+#' An artificial dataset, intended for presenting the extended features of \code{cleanR},
+#' which is a toolset for identifying potential errors in a dataset.
+#'
+#' @format A \code{data.frame} with 300 observations on the following 6 variables.
+#'   \describe{
+#'     \item{\code{addresses}}{a factor with fictitious US addresses}
+#'     \item{\code{binomial}}{a numeric vector with a binomial distributed variable}
+#'     \item{\code{poisson}}{a numeric vector with a Poisson distributed variable}
+#'     \item{\code{gauss}}{a numeric vector with a Gaussian distributed variable}
+#'     \item{\code{zigauss}}{a numeric vector with a zero-inflated Gaussian distributed variable}
+#'     \item{\code{bpinteraction}}{a factor with interactions between binomial and poisson}
+#'   }
+#'
+#' @source Artificial data
+#' @examples
+#'
+#' \dontrun{
+#' isID <- function(v, nMax = NULL, ...) {
+#'   out <- list(problem = FALSE, message = "")
+#'   if (class(v) %in% c("character", "factor", "labelled", "numeric", "integer")) {
+#'     v <- as.character(v)
+#'     lengths <- nchar(v)
+#'     if (all(lengths > 10) & length(unique(lengths)) == 1) {
+#'       out$problem <- TRUE
+#'       out$message <- "Warning: This variable seems to contain ID codes!"
+#'     }
+#'   }
+#'   out
+#' }
+#'
+#'
+#' countZeros <- function(v, ...) {
+#'   res <- length(which(v == 0))
+#'   summaryResult(list(feature = "No. zeros", result = res, value = res))
+#' }
+#' countZeros <- summaryFunction(countZeros, description = "Count number of zeros",
+#'                               classes = allClasses())
+#' summarize(toyData, numericSummaries = c(defaultNumericSummaries()))
+#'
+#'
+#' mosaicVisual <- function(v, vnam, doEval) {
+#'   thisCall <- call("mosaicplot", table(v), main = vnam, xlab = "")
+#'   if (doEval) {
+#'     return(eval(thisCall))
+#'   } else return(deparse(thisCall))
+#' }
+#' mosaicVisual <- visualFunction(mosaicVisual,
+#'                                description = "Mosaic plots using graphics",
[…]

### 2016-12-30 latex/codeForArticle.R

@@ -101,8 +101,8 @@ identifyColons <- checkFunction(identifyColons,
                                description = "Identify non-suffixed nor -prefixed colons",
                                classes = c("character", "factor", "labelled"))
 
-data(testData)
-clean(testData, replace = T,
+data(exampleData)
+clean(exampleData, replace = TRUE,
       preChecks = c("isKey", "isEmpty", "isID"),
       allVisuals = "mosaicVisual",
       characterSummaries = c(defaultCharacterSummaries(), "countZeros"),
@@ -122,4 +122,37 @@ clean(testData, replace = T,
       preChecks = c("isKey", "isEmpty", "isID"),
       allVisuals = "mosaicVisual",
       labelledCheck = c(defaultLabelledChecks(), "identifyColons"))
+
+
+
+wheresWally <- function(v, ...) {
+    res <- grepl("wally", v, ignore.case=TRUE)
+    problem <- any(res)
+    message <- "Wally was found in these data"
+    checkResult(list(problem = problem,
+                     message = message,
+                     problemValues = v[res]))
+}
+
+wheresWally <- checkFunction(wheresWally,
+                             description = "Search for the string 'wally' ignoring case",
+                             classes = c("character")
+                             )
+
+clean(testData, characterChecks=c(defaultCharacterChecks(), "wheresWally"),
+      replace=TRUE)
+
+
+set.seed(1)
+n <- 300
+
+addresses <- readLines("../addresses.txt")
+
+DF <- data.frame(addresses = addresses,
+                 binomial = rbinom(300, size=1, p=.7),
+                 poisson = rpois(300, lambda=1),
+                 gauss = rnorm(300),
+                 zigauss = c(rnorm(250), rep(0,50))
+                 )
+DF$bpinteraction <- interaction(DF$binomial, DF$poisson, sep=":")
 

### 2016-12-30 latex/article_vol2.tex
Fixed appendix

@@
 
-\hl{Alle pakker. Andet?}
-
 \appendix
-
-\section{Appendix A: cleaning the toyData data frame} \label{sec:appendix1}
-\includepdf[pages=2-4, pagecommand={}, frame=true, noautoscale=true, scale=0.7]{cleanR_toyData.pdf}
-
-\section{Appendix B: User-extended cleaning of  testData} \label{sec:appendix2}
-\includepdf[pages=2-, pagecommand={}, frame=true, noautoscale=true, scale=0.7]{cleanR_exampleData.pdf}
+\newpage
+
+%\includepdf[scale=0.8,clip,trim=0cm 0cm 0cm
+%2cm,pages={1},pagecommand={\section{appendix}\subsection{par‌​t1}}]{pdfdocument} \includepdf[sc%ale=0.8,clip,trim=0cm
+%0cm 0cm 2cm,pages={2-5},pagecommand={}]{pdfdocument}
+
+% \section{Appendix A: cleaning the toyData data frame} \label{sec:appendix1}
+\includepdf[pages=2, pagecommand={\section{Appendix A: cleaning the toyData data frame} \label{sec:appendix1}}, frame=true, noautoscale=true, scale=0.7]{cleanR_toyData.pdf}
+\includepdf[pages=3-4, pagecommand={}, frame=true, noautoscale=true, scale=0.7,pagecommand={}]{cleanR_toyData.pdf}
+
+\includepdf[pages=2, pagecommand={\section{Appendix B: User-extended cleaning of
+  exampleData} \label{sec:appendix2}}, frame=true, noautoscale=true, scale=0.7]{cleanR_toyData.pdf}
+\includepdf[pages=3-, pagecommand={}, frame=true, noautoscale=true, scale=0.7]{cleanR_exampleData.pdf}
 %\hl{Data cleaning with user supplied extensions here}

### 2016-12-30 latex/codeForArticle.R

@@ -69,10 +69,10 @@ mosaicVisual <- visualFunction(mosaicVisual,
                                description = "Mosaic plots using graphics",
                                classes = allClasses())
 
-data(toyData)
-pdf("latex/mosaicPlotExample.pdf", 5, 3)
-mosaicVisual(toyData$var1, "Variable 1", doEval = TRUE)
-dev.off()
+#data(toyData)
+#pdf("latex/mosaicPlotExample.pdf", 5, 3)
+#mosaicVisual(toyData$var1, "Variable 1", doEval = TRUE)
+#dev.off()
 
 
 identifyColons <- function(v, nMax = Inf, ... ) {

### 2016-12-30 latex/article_vol2.tex
Documentation

@@
 \smallskip Control summarize, visualize, and check steps \\
-\quad \R{mode} & What steps should be performed for each variable (out of the three possibilities \textit{summarize}, \textit{visualize}, \textit{check})?
-& \R{c("summarize", "visualize", "check")} (corresponding to all three) \\
+\quad \R{mode} & What steps should be performed for each variable (out
+                 of the three possibilities \textit{summarize},
+                 \textit{visualize}, \textit{check})? &
+                                                        \R{c("summarize", "visualize", "check")} (corresponding to all three steps) \\
 \quad \R{smartNum} & Should numerical values with only a few unique
@@
 
-\subsection{Rolling out the arguments}
+\subsection{Polishing off the arguments}
 We begin with an example that is intended as an illustration of how \R{clean()} might
@@
 the \R{render} and \R{openResult} arguments, which both default to
-\R{TRUE} \hl{Bliver det ved med at være sandt?}. \R{render} determines
-if the \proglang{R} markdown file produced should be rendered using
-the \pkg{rmarkdown} package and \R{openResult} decides whether the
-output html or pdf file should be opened. The following command
-produces an \proglang{R} markdown file containing the data cleaning
-summary information, but without rendering nor opening the markdown file:
+\R{TRUE}. \R{render} determines if the \proglang{R} markdown file
+produced should be rendered using the \pkg{rmarkdown} package and
+\R{openResult} decides whether the output html or pdf file should be
+opened. The following command produces an \proglang{R} markdown file
+containing the data cleaning summary information, but without
+rendering nor opening the markdown file:
 
@@
 
-We will now move on to discuss how not only the structure of the cleaning summary is manipulated, but also its very contents.
-This is done through the summarize-visualize-check (SVC) step, as illustrated in Figure \ref{figure:cleanStructure}.
-\pkg{cleanR} uses three different types of functions for performing these steps,
-namely \code{summaryFunction}s, \code{visualFunction}s and
-\code{checkFunction}s.
-By default, \R{clean()} runs selected summary, visualization and check functions on each variable in the dataset, and the exact choice of these functions depends on the classes of the variables.
-For instance, though detection of outlier values might be interesting for numerical variables, it holds little meaning for factor variables, and therefore, numerical and factor variables need different checks.
-Table~\ref{table.SVCfunctions}
-lists all available SVC functions, but we can also use the
+We will now move on to discuss how not only the structure of the
+cleaning summary is manipulated, but also its very contents. This is
+done through the summarize-visualize-check (SVC) step, as illustrated
+in Figure \ref{figure:cleanStructure}.  \pkg{cleanR} uses three
+different types of functions for performing these steps, namely
+\code{summaryFunction}s, \code{visualFunction}s and
+\code{checkFunction}s.  By default, \R{clean()} runs selected summary,
+visualization and check functions on each variable in the dataset, and
+the exact choice of these functions depends on the classes of the
+variables. For instance, though detection of outlier values might be
+interesting for numerical variables, it holds little meaning for
+factor variables, and therefore, numerical and factor variables need
+different checks. Table~\ref{table.SVCfunctions} lists all available
+summarize/visualize/check functions, but we can also use the
 \R{allSummaryFunctions()}, \R{allVisualFunctions()}, and
 \R{allCheckFunctions()} functions in \pkg{cleanR} to print overview
-lists in \proglang{R}. For example, the implemented \R{summaryFunction}s are:
+lists in \proglang{R}. For example, the implemented
+\R{summaryFunction}s are:
 
@@
 
-As with \R{summaryFunction}s, a complete list of available \R{checkFunction}s is obtained by calling 
-\R{allCheckFunctions()}. Note however, that \R{checkFunction}s have a usage beyond the 
-\R{XXXChecks} arguments, namely in the \textit{pre-check} stage. In this stage, it is determined 
-whether or not each variable is suitable for the SVC-steps. The functions used in the pre-check stage
-should be \R{checkFunction}s that are applicable to all variable classes. The default pre-checks, the functions
-\R{isKey} and \R{isEmpty}, check whether a variable has unique values for all observations or only a single value for 
-all observations, respectively. If one of these statements are true, the variable will not be subjected to the SVC steps. 
-We can allow empty variables to move on to the SVC step by only checking for keys in the pre-check step:
+As with \R{summaryFunction}s, a complete list of available
+\R{checkFunction}s is obtained by calling
+\R{allCheckFunctions()}. Note however, that \R{checkFunction}s have a
+usage beyond the \R{XXXChecks} arguments, namely in the
+\textit{pre-check} stage. In this stage, it is determined whether or
+not each variable is suitable for the summarize/visualize/check (SVC)
+steps. The functions used in the pre-check stage should be
+\R{checkFunction}s that are applicable to all variable classes. The
+default pre-checks, the functions \R{isKey()} and \R{isEmpty()}, check
+whether a variable has unique values for all observations or only a
+single value for all observations, respectively. If one of these
+statements are true, the variable will not be subjected to the SVC
+steps.  We can allow empty variables to move on to the SVC step by
+only checking for keys in the pre-check step:
 
@@
 
-Note that the data visualizations in the cleaning summary are also controllable, though only a single function can be provided for all variable types.
-If for instance, we wish to change the visualizations from the default \pkg{ggplot2} style histograms and barplots to base \proglang{R} histograms and barplots, we type
+Note that the data visualizations in the cleaning summary are also
+controllable, though only a single function can be provided for all
+variable types. If for instance, we wish to change the visualizations
+from the default \pkg{ggplot2} style histograms and barplots to base
+\proglang{R} histograms and barplots, we type
 \begin{Schunk}
@@
 providing a vector of function names to run for a particular summary. To
-only get the median, the minimum and the maximum we set
+only get the median, minimum and the maximum we set
 \R{numericSummaries=c("centralValue", "minMax")}:
@@
 
-This section consists of two parts. First, we describe how custom \R{summaryFunction}s, 
+This section consists of two parts. First, we describe how customized \R{summaryFunction}s, 
 \R{visualFunction}s and \R{checkFunctions} can be made, including an introduction to a 
@@
 \def\arraystretch{1.8}% 
-\hspace*{-.2cm}\begin{tabular}[t]{p{0.45\linewidth} | p{0.45\linewidth}} \vspace*{-.9cm}If \R{doEval} is \R{TRUE}: \newline A plot that is opened by the graphic device in \R{R}.
-& \vspace*{-.9cm}If \R{doEval} is \R{FALSE}: \newline   A text string with \R{R} code, as described above.
+\hspace*{-.2cm}\begin{tabular}[t]{p{0.45\linewidth} | p{0.45\linewidth}} \vspace*{-1cm}If \R{doEval} is \R{TRUE}: \newline A plot that is opened by the graphic device in \R{R}.
+& \vspace*{-1cm}If \R{doEval} is \R{FALSE}: \newline   A text string with \R{R} code, as described above.
 \end{tabular} \egroup  & A \R{checkResult} object, i.e.
@@
 not specified depends on the type of \R{mySummaryFunction}. If
-\R{mySummaryFunction} is a \R{S3} generic function with associated
+\R{mySummaryFunction} is constructed as an \R{S3} generic function with associated
 methods, the call to \R{summaryFunction()} will automatically produce
@@
 \end{Schunk}
-One comment should be devoted to the helper function, \R{messageGenerator}. 
-This function's sole purpose is aiding consistent
-styling of all \code{checkFunction} messages. The function simply
-pastes together the \code{problemMessage} and the
+One comment should be devoted to the helper function,
+\R{messageGenerator()}.  This function's sole purpose is aiding
+consistent styling of all \code{checkFunction} messages. The function
+simply pastes together the \code{problemMessage} and the
 \code{problemValues}, with the latter being quoted and sorted
-alphabetically. If \R{nMax} is not \R{Inf}, only the first \R{nMax}
-problem values will be pasted onto the message, accompanied by a
-comment about how many problem values were left out (if any).  Note
-that printing quotes in \pkg{rmarkdown} requires an extensive amount
-of character escaping, so opting for \code{messageGenerator()} really
-is the easiest solution.
+alphabetically. If the \R{nMax} argument to \R{messageGenerator()} is
+not \R{Inf}, only the first \R{nMax} problem values will be pasted
+onto the message, accompanied by a comment about how many problem
+values were left out (if any).  Note that printing quotes in
+\pkg{rmarkdown} requires an extensive amount of character escaping, so
+opting for \code{messageGenerator()} really is the easiest solution.
 
@@
 in the outputted \R{checkResult}.
-
-\hl{Jeg forstår ikke helt, hvad der var uklart her ang. messageGenerator. Jeg har nu
-stort set bare byttet om på rækkefølgen af de to sidste paragraphs her, hjælper
-det?}
 
@@
 
-\section{Zooming in on cleaning challenges}
+\section{Rubbing down data cleaning challenges}
 \label{sec:specificExamples}
-\hl{Lidt poppet titel, men den gamle ("Using cleanR in specific situations")
-fik mig til flabel at tænke "i modsætning til uspecifikke situationer?". Jeg 
-er dog meget åben for en bedre titel.}
+%\hl{Lidt poppet titel, men den gamle ("Using cleanR in specific situations")
+%fik mig til flabel at tænke "i modsætning til uspecifikke situationer?". Jeg 
+%er dog meget åben for en bedre titel.}
 
@@
 then be included:
+{\small
 \begin{Verbatim}
@@
 \end{Verbatim}
+}
 %\hl{Use proper formatting here. How do we do non-R code?}

### 2016-12-30 latex/article_vol2.tex
Donish

@@
 \includepdf[pages=2, pagecommand={\section{Appendix B: User-extended cleaning of
-  exampleData} \label{sec:appendix2}}, frame=true, noautoscale=true, scale=0.7]{cleanR_toyData.pdf}
+  exampleData} \label{sec:appendix2}}, frame=true, noautoscale=true, scale=0.7]{cleanR_exampleData.pdf}
 \includepdf[pages=3-, pagecommand={}, frame=true, noautoscale=true, scale=0.7]{cleanR_exampleData.pdf}

### 2017-01-02 latex/article_vol2.tex
Renamed to dataMaid

@@
   Health\\University of Copenhagen}
-\title{\pkg{cleanR}: Maid for Cleaning Datasets in \proglang{R}}
+\title{\pkg{dataMaid}: your assistant for data cleaning in \proglang{R}}
 
@@
 \Plainauthor{Anne H. Petersen, Claus Thorn Ekstr\o m} %% comma-separated
-\Plaintitle{{cleanR}: maid for cleaning datasets in R} %% without formatting
-\Shorttitle{\pkg{cleanR}: maid for cleaning datasets in R} %% a short title (if necessary)
+\Plaintitle{{dataMaid}: your assistant for data cleaning in R} %% without formatting
+\Shorttitle{\pkg{dataMaid}: your data cleaning PA in R} %% a short title (if necessary)
 
@@
   by an investigator with an understanding of the variables in the
-  data, but not necessarily knowledge of \R{R}. Thereby, \pkg{cleanR} 
+  data, but not necessarily knowledge of \R{R}. Thereby, \pkg{dataMaid} 
   aids the dialogue between data
@@
   documentation of reproducible data cleaning steps and data quality
-  control. Moreover, the \pkg{cleanR} solution changes the data
+  control. Moreover, the \pkg{dataMaid} solution changes the data
   cleaning process from the usual ad hoc approach to a systematic,
-  well-documented endeavor.  \pkg{cleanR} also provides a suite of
+  well-documented endeavor.  \pkg{dataMaid} also provides a suite of
   more typical \proglang{R} tools for interactive data quality
@@
 
-  % The \pkg{cleanR} package is designed to be easily extended with
+  % The \pkg{dataMaid} package is designed to be easily extended with
   % custom user-created checks that are relevant in particular
@@
 and less so for the general researcher with a knowledge about a specific
-field and the context of the available data. The \pkg{cleanR} package
+field and the context of the available data. The \pkg{dataMaid} package
 tries to address this by providing a framework that both allows for
@@
 
-The \pkg{cleanR} approach to data cleaning and -quality assessment is
+The \pkg{dataMaid} approach to data cleaning and -quality assessment is
 governed by two fundamental paradigms. First of all, there is no need
@@
 means that data cleaning can easily be a well-documented,
-well-specified procedure. In order to aid this paradigm, \pkg{cleanR}
+well-specified procedure. In order to aid this paradigm, \pkg{dataMaid}
 provides easy-to-use, automated tools for data quality assessment in
@@
 analyst. Hopefully, having easy access to data descriptions through
-\pkg{cleanR} will help this necessary knowledge sharing.
-
-While \pkg{cleanR}s primary raison d'être is auto-generating data
+\pkg{dataMaid} will help this necessary knowledge sharing.
+
+While \pkg{dataMaid}s primary raison d'être is auto-generating data
 quality assessment overview documents, we still wish to emphasize that
 it is \emph{not} a tool for unsupervised data cleaning. This qualifies
-as the second paradigm of \pkg{cleanR}: Data cleaning decisions
-should always be made by humans. Therefore, \pkg{cleanR} does not
+as the second paradigm of \pkg{dataMaid}: Data cleaning decisions
+should always be made by humans. Therefore, \pkg{dataMaid} does not
 supply any tools for ``fixing'' errors in the data. However, we do
@@
 This manuscript is structured as follows: First, in Section
-\ref{sec:usingcleanR}, we introduce the main representative of the first
+\ref{sec:usingdataMaid}, we introduce the main representative of the first
 paradigm, namely the \R{clean()} function, which generates data
-cleaning overview documents. In the \pkg{cleanR} package, we have
+cleaning overview documents. In the \pkg{dataMaid} package, we have
 provided a number of default cleaning steps that cover the data
 cleaning challenges, we find to be most common. Next, in Section
-\ref{sec:interactiveCleanR}, we present the interactive mode of \pkg{cleanR}, as motivated
+\ref{sec:interactiveCleanR}, we present the interactive mode of \pkg{dataMaid}, as motivated
 by the second paradigm above. But, as any data analyst knows,
@@
 that cannot be detected by our data checking functions. In Section
-\ref{sec:extending}, we turn to the question of how \pkg{cleanR} extensions
+\ref{sec:extending}, we turn to the question of how \pkg{dataMaid} extensions
 can be made, such that they are integrable with the \R{clean()}
-function and with the other tools available in \pkg{cleanR}.  At last,
+function and with the other tools available in \pkg{dataMaid}.  At last,
 in Section \ref{sec:specificExamples}, we discuss a number of examples of
-specific data cleaning challenges and how \pkg{cleanR} can be used to
+specific data cleaning challenges and how \pkg{dataMaid} can be used to
 solve them.
@@
 \section{Creating a data cleaning overview} 
-\label{sec:usingcleanR}
-
-The \R{clean()} function is the primary workhorse of \pkg{cleanR} and
+\label{sec:usingdataMaid}
+
+The \R{clean()} function is the primary workhorse of \pkg{dataMaid} and
 this is the only function needed if a standard battery
@@
 calling \R{clean()} on the dataset \R{toyData} available in
-\pkg{cleanR}. The first two pages of this data cleaning output are
+\pkg{dataMaid}. The first two pages of this data cleaning output are
 shown in Figure~\ref{fig:example1}. \R{toyData} is a very
 small ($15$ rows of $6$ variables), artificial dataset which was created to
-illustrate the main capabilities of \pkg{cleanR}. The following
+illustrate the main capabilities of \pkg{dataMaid}. The following
 commands load the dataset and produce the cleaning output:
@@
 \frame{\includegraphics[width=7.5cm,page=3]{cleanR_toyData.pdf}}
-%\includepdf[pages={2}, pagecommand={}]{cleanR_testData.pdf}
+%\includepdf[pages={2}, pagecommand={}]{dataMaid_testData.pdf}
 \end{center}
@@
 \begin{Sinput}
-> library(cleanR)
+> library(dataMaid)
 > data(toyData)
@@
   each variable in the output), visual functions (used for  visualization of each variable), and
-  check functions (used for data checks for each variable) currently implemented in \pkg{cleanR}. The variable
+  check functions (used for data checks for each variable) currently implemented in \pkg{dataMaid}. The variable
   classes C, F, I, L, B, N, and D, refer to \R{character}, \R{factor},
@@
 done through the summarize-visualize-check (SVC) step, as illustrated
-in Figure \ref{figure:cleanStructure}.  \pkg{cleanR} uses three
+in Figure \ref{figure:cleanStructure}.  \pkg{dataMaid} uses three
 different types of functions for performing these steps, namely
@@
 \R{allSummaryFunctions()}, \R{allVisualFunctions()}, and
-\R{allCheckFunctions()} functions in \pkg{cleanR} to print overview
+\R{allCheckFunctions()} functions in \pkg{dataMaid} to print overview
 lists in \proglang{R}. For example, the implemented
@@
 Thus we can see, for example, that for \R{numeric}, \R{integer}, or \R{Date} 
-variables, \pkg{cleanR} provides functions for adding summary information about 
+variables, \pkg{dataMaid} provides functions for adding summary information about 
 the minimum and maximum values, while all seven variable
-classes dealt with in \pkg{cleanR} have functions for central
+classes dealt with in \pkg{dataMaid} have functions for central
 tendency summaries (i.e., mode or median).
@@
 
-\section{Using cleanR interactively}
+\section{Using dataMaid interactively}
 \label{sec:interactiveCleanR}
@@
 through the data cleaning process as well. Aside from the \R{clean()} function 
-presented above, \pkg{cleanR} also provides more standard \proglang{R} interactive tools, such as
+presented above, \pkg{dataMaid} also provides more standard \proglang{R} interactive tools, such as
 functions that print results to the console or returns the information
@@
 functions \R{check()}, \R{summarize()} and \R{visualize()} to work
-interactively with \pkg{cleanR}.
+interactively with \pkg{dataMaid}.
 
@@
 \R{XXXChecks} argument (e.g., \R{numericChecks}, \R{factorChecks},
-etc. as described in Section~\ref{sec:usingcleanR}) to provide a
+etc. as described in Section~\ref{sec:usingdataMaid}) to provide a
 vector of the check functions that should be applied. Recall that
@@
 
-\section{Extending cleanR} 
+\section{Extending dataMaid} 
 \label{sec:extending}
 
-\pkg{cleanR} can be used as a user-friendly, self-contained package, as
-shown in the previous sections. However, \pkg{cleanR} is fully
+\pkg{dataMaid} can be used as a user-friendly, self-contained package, as
+shown in the previous sections. However, \pkg{dataMaid} is fully
 customizable and \R{clean()} is mainly a tool for formatting the
 results from various checking-, summary- and visualization
-functions. Thus, the actual work underlying a \pkg{cleanR} output file
+functions. Thus, the actual work underlying a \pkg{dataMaid} output file
 can be anything --- depending on the arguments given to \R{clean()}
@@
 \R{visualFunction}s and \R{checkFunctions} can be made, including an introduction to a 
-few convenient tools available in \pkg{cleanR} for aiding SVC-function construction. 
+few convenient tools available in \pkg{dataMaid} for aiding SVC-function construction. 
 Table \ref{table.functionTypes} serves as an overview of the internals of these three
@@
 However, new user-defined functions can also be registered locally to
-be part of the full machinery of \pkg{cleanR}, and these function will
+be part of the full machinery of \pkg{dataMaid}, and these function will
 be recognized and behave in the same way as the build-in functions in
-\pkg{cleanR}. The presentation below is given in the format of
+\pkg{dataMaid}. The presentation below is given in the format of
 function templates, written in pseudo-code. These templates are
@@
 \subsubsection{Writing a summaryFunction}
-As mentioned above, \pkg{cleanR} provides a dedicated class for
+As mentioned above, \pkg{dataMaid} provides a dedicated class for
 \code{summaryFunction}s. However, this does not imply that they are
@@
 for filling out the \R{classes} argument, as it simply lists all
-available classes in \pkg{cleanR}:
+available classes in \pkg{dataMaid}:
 \begin{Schunk}
@@
 \code{visualFunction}s are the functions that produce the figures in a
-\pkg{cleanR} output document. Writing a \R{visualFunction} is slightly
+\pkg{dataMaid} output document. Writing a \R{visualFunction} is slightly
 more complicated than writing a \R{summaryFunction}. This follows from
@@
 \subsubsection{Writing a checkFunction}
-The last, but perhaps most important, \pkg{cleanR} function type is
+The last, but perhaps most important, \pkg{dataMaid} function type is
 the \code{checkFunction}. These are the functions that flag issues in
@@
 	problemMessage <- "[Message that is printed prior to listing
-			    problem values in the cleanR output, 
+			    problem values in the dataMaid output, 
                             ending with a colon]"
@@
 function as part of the new visualization function.  We will define
-the new function such that it gets the full \pkg{cleanR}
+the new function such that it gets the full \pkg{dataMaid}
 functionality. This can be done using the following code.
@@
 Here, we use the function \R{allClasses()} to quickly obtain a vector
-of all the seven variable classes addressed in \pkg{cleanR}. Note that
+of all the seven variable classes addressed in \pkg{dataMaid}. Note that
 if \R{mosaicVisual()} were an S3 generic function, this argument could
@@
 Now, we are ready to use these new functions in a \R{clean()}
-call. The extended \pkg{cleanR} output document should have the
-following modifications, relative to the standard \pkg{cleanR} output:
+call. The extended \pkg{dataMaid} output document should have the
+following modifications, relative to the standard \pkg{dataMaid} output:
 \begin{itemize}
@@
 \label{sec:specificExamples}
-%\hl{Lidt poppet titel, men den gamle ("Using cleanR in specific situations")
+%\hl{Lidt poppet titel, men den gamle ("Using dataMaid in specific situations")
 %fik mig til flabel at tænke "i modsætning til uspecifikke situationer?". Jeg 
@@
 
-Finally, we present a few examples of how to make \pkg{cleanR}
+Finally, we present a few examples of how to make \pkg{dataMaid}
 solve specific issues related to data cleaning. First, we discuss the
 challenges related to cleaning large datasets, particularly in terms
-of memory use and computation speed. Next, we show how \pkg{cleanR}
+of memory use and computation speed. Next, we show how \pkg{dataMaid}
 can be used for problem-flagging. Lastly, we discuss how the
-\pkg{cleanR} output document can be included in other \pkg{rmarkdown}
+\pkg{dataMaid} output document can be included in other \pkg{rmarkdown}
 documents as a way to produce clear and concise documentation of a
@@
 still produce (potentially very long) data cleaning overview
-documents. Note that the interactive tools of \pkg{cleanR} can be used
+documents. Note that the interactive tools of \pkg{dataMaid} can be used
 as usual or sequentially in small subsets of the large dataset, if no
@@
 quite heavy objects in terms of memory allocation. Therefore, it might
-be beneficial to not include figures in the \pkg{cleanR} outputs for
+be beneficial to not include figures in the \pkg{dataMaid} outputs for
 very large datasets. This is controlled via the \R{mode} argument:
@@
 \ref{sec:functionTemplates}. For instance, a future extension of
-\pkg{cleanR} might be the inclusion of ASCII plots, as
+\pkg{dataMaid} might be the inclusion of ASCII plots, as
 e.g. represented in the \proglang{R} package \pkg{txtplot}.
@@
 > clean(toyData, render = FALSE, openResult = FALSE}
-> render("cleanR_toyData.Rmd", quiet = FALSE)
-\end{Sinput}
-\end{Schunk}
-This also deals with the fact that \pkg{cleanR} can produce
+> render("dataMaid_toyData.Rmd", quiet = FALSE)
+\end{Sinput}
+\end{Schunk}
+This also deals with the fact that \pkg{dataMaid} can produce
 \pkg{rmarkdown} files that supersedes the upper size limit. \hl{dette giver
@@
 
-\subsection{Using cleanR for problem flagging}
+\subsection{Using dataMaid for problem flagging}
 If the data is large, but memory issues and computation time are less
@@
 
-\subsection{Including cleanR documents in other files}
-Sometimes, a \pkg{cleanR} document might be a useful addition to a
+\subsection{Including dataMaid documents in other files}
+Sometimes, a \pkg{dataMaid} document might be a useful addition to a
 more general overview document, including also for example pairwise
 association plots, time series plots, or exploratory analysis
-results. To this end, it is possible to produce a \pkg{cleanR}
+results. To this end, it is possible to produce a \pkg{dataMaid}
 document that can readily be included in other \pkg{rmarkdown}
@@
 
-If it is important that the embedded \pkg{cleanR} document can be
+If it is important that the embedded \pkg{dataMaid} document can be
 rendered to either of these two file types, we recommend setting
@@
 error. The following is an example of how such a master document
-preamble YAML might look like and how the \R{cleanR\_toyData.Rmd} file can
+preamble YAML might look like and how the \R{dataMaid\_toyData.Rmd} file can
 then be included:
@@
 
-```{r, child = 'cleanR_toyData.Rmd'}
+```{r, child = 'dataMaid_toyData.Rmd'}
 ```
@@
 
-In the this example, the \R{cleanR\_toyData.Rmd} file could have been created as follows:
+In the this example, the \R{dataMaid\_toyData.Rmd} file could have been created as follows:
 \begin{Schunk}
@@
 
-In this paper we have introduced the \proglang{R} package \pkg{cleanR}
+In this paper we have introduced the \proglang{R} package \pkg{dataMaid}
 for performing reproducible error detection and data cleaning
@@
 While the the current release is stable, the authors have an interest
-in further developing the functionality of \pkg{cleanR} by providing
+in further developing the functionality of \pkg{dataMaid} by providing
 more summary, visual, and check function as part of the default
@@
 \nocite{shiny}
-\nocite{cleanR}
+\nocite{dataMaid}
 \nocite{rmarkdown}

### 2017-01-02 R/allClasses.R

@@ -1,15 +1,15 @@
-#' @title Vector of all variable classes in \code{cleanR}
-#' 
+#' @title Vector of all variable classes in \code{dataMaid}
+#'
 #' @description Returns the names of the seven data classes for which
-#' \code{cleanR} is implemented, namely \code{"character"}, \code{"Date"},
+#' \code{dataMaid} is implemented, namely \code{"character"}, \code{"Date"},
 #' \code{"factor"}, \code{"integer"}, \code{"labelled"}, \code{"logical"} and
 #' \code{"numeric"}.
-#' 
-#' @examples 
+#'
+#' @examples
 #' allClasses()
-#' 
+#'
 #' @export
 allClasses <- function() {
-  c("character", "Date", "factor", "integer", "labelled", 
+  c("character", "Date", "factor", "integer", "labelled",
     "logical", "numeric")
 }

### 2017-01-02 R/allXFunctions.R

@@ -1,10 +1,10 @@
-#Make a functionSummary of all functions of type X, that is, 
+#Make a functionSummary of all functions of type X, that is,
 #coerce their names, description attributes and classes attributes
 #into a list.
 #Called from allVisualFunction(), allSummaryFunctions(), allCheckFunctions().
 allXFunctions <- function(X) {
   allF <- Filter(function(x) X %in% class(get(x)), union(ls(envir = .GlobalEnv),
-                                                         ls("package:cleanR")))
+                                                         ls("package:dataMaid")))
   out <- list(name = allF, description = sapply(allF, function(x) description(get(x))),
               classes = lapply(allF, function(x) classes(get(x))))
   class(out) <- c("functionSummary", "list")

### 2017-01-02 R/check.R

@@ -44,7 +44,7 @@
 #' are returned by calling e.g. \code{defaultCharacterChecks()},
 #' \code{defaultFactorChecks()}, \code{defaultNumericChecks()}, etc. Moreover,
 #' all available \code{checkFunction}s (including both locally defined
-#' functions and functions imported from \code{cleanR} or other packages) can
+#' functions and functions imported from \code{dataMaid} or other packages) can
 #' be viewed by calling \code{allCheckFunctions()}.
 #'
 #'

### 2017-01-02 R/checkFunction.R

@@ -1,66 +1,66 @@
 #' @title Create an object of class checkFunction
-#' 
-#' @description Convert a function, \code{f}, into an S3 
+#'
+#' @description Convert a function, \code{f}, into an S3
 #' \code{checkFunction} object. This adds \code{f} to the
-#' overview list returned by an \code{allCheckFunctions()} 
+#' overview list returned by an \code{allCheckFunctions()}
 #' call.
 #'
 #' @param f A function. See details and examples below for the
-#' exact requirements of this function. 
-#' 
+#' exact requirements of this function.
+#'
 #' @param description A character string describing the check
 #' performed by \code{f}. If \code{NULL} (the default), the
 #' name of \code{f} will be used instead.
-#' 
+#'
 #' @param classes The classes for which \code{f} is intended to
 #' be called. If \code{NULL} (the default), one of two things happens.
-#' If \code{f} is not a S3 generic function, the \code{classes} 
-#' attribute of \code{f} will be an empty character string. If 
+#' If \code{f} is not a S3 generic function, the \code{classes}
+#' attribute of \code{f} will be an empty character string. If
 #' \code{f} is a S3 generic function, an automatic look-up
 #' for methods will be conducted, and the \code{classes} attribute
 #' will then be filled out automatically. Note that the function
-#' \code{\link{allClasses}} (listing all classes used in \code{cleanR})
-#' might be useful. 
-#' 
-#' @return A function of class \code{checkFunction} which has to attributes, 
-#' namely \code{classes} and \code{description}. 
-#' 
-#' @details \code{checkFunction} represents the functions used in 
+#' \code{\link{allClasses}} (listing all classes used in \code{dataMaid})
+#' might be useful.
+#'
+#' @return A function of class \code{checkFunction} which has to attributes,
+#' namely \code{classes} and \code{description}.
+#'
+#' @details \code{checkFunction} represents the functions used in
 #' \code{\link{check}} and \code{\link{clean}} for performing
-#' error checks and quality control on variables in dataset. 
-#' 
-#' An example of defining a new \code{checkFunction} is given below. 
-#' Note that the minimal requirements for such a function (in order for it to be 
-#' compatible with \code{check()} and \code{clean()}) is the following 
[…]

### 2017-01-02 R/clean.R

@@ -83,7 +83,7 @@
 #'
 #' @param file The filename of the outputted rmarkdown (.Rmd) file.
 #' If set to \code{NULL} (the default), the filename will be the name of \code{data}
-#' prefixed with "cleanR_", if this qualifies as a valid file name (e.g. no special
+#' prefixed with "dataMaid_", if this qualifies as a valid file name (e.g. no special
 #' characters allowed). Otherwise, \code{clean()} tries to create a valid filename by
 #' substituing illegal characters. Note that a valid file is of type .Rmd, hence all
 #' filenames should have a ".Rmd"-suffix.
@@ -95,7 +95,7 @@
 #'
 #' @param vol Extra text string or numeric that is appended on the end of the output
 #' file name(s). For example, if the dataset is called "myData", no file argument is
-#'  supplied and \code{vol=2}, the output file will be called "cleanR_myData2.Rmd"
+#'  supplied and \code{vol=2}, the output file will be called "dataMaid_myData2.Rmd"
 #'
 #' @param characterSummaries A vector of the names of summary functions to apply to
 #' character vectors.
@@ -282,7 +282,7 @@ clean <- function(data, output=c("pdf", "html"), render=TRUE,
     #######Secret arguments that were removed for the users but are still implemented#########
     ##########################################################################################
 
-    #If we would ever want to allow users not to append the cleanR stamp:
+    #If we would ever want to allow users not to append the dataMaid stamp:
     brag <- TRUE
 
     #If we would ever want (Windows) users not to be nagged
@@ -320,8 +320,8 @@ clean <- function(data, output=c("pdf", "html"), render=TRUE,
     ## Set the output file name if input is NULL or not Rmd
     if (is.null(file)) {
       if (substr(dfname, 1, 11) == "data.frame(") {
-        file <- paste0("cleanR_unnamedData", vol, ".Rmd")
-      } else file <- normalizeFileName(paste0("cleanR_", dfname, vol, ".Rmd"))
+        file <- paste0("dataMaid_unnamedData", vol, ".Rmd")
+      } else file <- normalizeFileName(paste0("dataMaid_", dfname, vol, ".Rmd"))
     } else {
       originalFile <- file
       faultyExt <- FALSE
@@ -396,10 +396,10 @@ clean <- function(data, output=c("pdf", "html"), render=TRUE,
         if (fileExists & outFileExists) problemFiles <- paste(file, "and", outFile)
[…]

### 2017-01-02 R/dataMaid-package.R

@@ -1,6 +1,6 @@
-#' Exteneded example data to test the features of cleanR
+#' Exteneded example data to test the features of dataMaid
 #'
-#' A dataset of constructed data used as test bed when using \code{cleanR} for identifying
+#' A dataset of constructed data used as test bed when using \code{dataMaid} for identifying
 #' potential errors in a dataset.
 #'
 #' @format A data frame with 15 rows and 14 variables.
@@ -32,9 +32,9 @@
 "testData"
 
 
-#' Small example data to show the features of cleanR
+#' Small example data to show the features of dataMaid
 #'
-#' An artificial dataset, intended for presenting the key features of \code{cleanR}, which is a
+#' An artificial dataset, intended for presenting the key features of \code{dataMaid}, which is a
 #' toolset for identifying potential errors in a dataset.
 #'
 #' @format A \code{data.frame} with 15 rows and 6 variables.
@@ -61,7 +61,7 @@
 
 #' Example data with zero-inflated variables
 #'
-#' An artificial dataset, intended for presenting the extended features of \code{cleanR},
+#' An artificial dataset, intended for presenting the extended features of \code{dataMaid},
 #' which is a toolset for identifying potential errors in a dataset.
 #'
 #' @format A \code{data.frame} with 300 observations on the following 6 variables.

### 2017-01-02 R/isCPR.R

@@ -1,40 +1,40 @@
 #' @title Check if a variable consists of Danish CPR numbers
-#' 
-#' @description A \code{\link{checkFunction}} that checks if \code{v} consists exclusively 
+#'
+#' @description A \code{\link{checkFunction}} that checks if \code{v} consists exclusively
 #' of valid Danish civil registration (CPR) numbers, ignoring missing values. This
 #' function is intended for use as a precheck in \code{\link{clean}}, ensuring
-#' that CPR numbers are not included in a \code{cleanR} output document. 
-#' 
+#' that CPR numbers are not included in a \code{dataMaid} output document.
+#'
 #' @param v A variable (vector) to check. This variable is allowed to have any class.
-#' 
+#'
 #' @param ... Not in use.
-#' 
-#' @return A \code{\link{checkResult}} with three entires: 
-#' \code{$problem} (a logical indicating whether the variable consists 
-#' of CPR numbers), \code{$message} (if a problem was found, 
-#' the following message: "Warning: The variable seems to consist of 
-#' Danish civil regristration (CPR) numbers.", 
-#' otherwise "") and \code{$problemValues} (always \code{NULL}).  
-#' 
-#' @examples 
+#'
+#' @return A \code{\link{checkResult}} with three entires:
+#' \code{$problem} (a logical indicating whether the variable consists
+#' of CPR numbers), \code{$message} (if a problem was found,
+#' the following message: "Warning: The variable seems to consist of
+#' Danish civil regristration (CPR) numbers.",
+#' otherwise "") and \code{$problemValues} (always \code{NULL}).
+#'
+#' @examples
 #' CPRs <-  sapply(c("01011988", "02011987", "04052006", "01021990", "01021991",
 #'                   "01021993", "01021994", "01021995", "01021996", "01021997",
-#'                   "01021970", "01021971", "01021972", "01021973", "01021974"), cleanR:::makeCPR)
+#'                   "01021970", "01021971", "01021972", "01021973", "01021974"), dataMaid:::makeCPR)
 #' nonCPRs <- c(1:10)
 #' mixedCPRs <- c(CPRs, nonCPRs)
-#' 
+#'
[…]

### 2017-01-02 R/makeXFunction.R

@@ -1,9 +1,8 @@
-
 #Make a function of type X (among visualFunction, summaryFunction and
 #checkFunction). If classes is null, and fName is a S3 generic,
 #the function will make a look-up for all methods available in the
 #global enviroment and fill out the slot accordingly.
-#Note: This does not work for functions defined wihtin cleanR. As
+#Note: This does not work for functions defined wihtin dataMaid. As
 #of now, they must have their classes specified explicitly!
 #' @importFrom utils methods
 makeXFunction <- function(fName, description, classes, X)  {
@@ -12,9 +11,9 @@ makeXFunction <- function(fName, description, classes, X)  {
   } else {
     f <- get(fName)
   }
-   #note: default pos (-1) will look in cleanR namespace first
-   #when called interactively. This results in weird behaviour when 
-   #users try to overwrite our functions. 
+   #note: default pos (-1) will look in dataMaid namespace first
+   #when called interactively. This results in weird behaviour when
+   #users try to overwrite our functions.
 
   if (is.null(description)) description <- fName
   if (is.null(classes)) {

### 2017-01-02 R/messageGenerator.R

@@ -1,13 +1,13 @@
 #' @title Produce a message for the output of a checkFunction
-#' @description Helper function for producing output messages for 
-#' \code{\link{checkFunction}} type functions. 
+#' @description Helper function for producing output messages for
+#' \code{\link{checkFunction}} type functions.
 #'
-#' @param problemStatus A list consisting of two entries: 
-#' 
-#' \code{$problem} - logical indicating whether a problem was found by the 
-#' \code{checkFunction} responsible for the making the \code{messageGenerator()} call, 
-#' 
-#' \code{$problemValues} - a vector of values from the variable that were 
+#' @param problemStatus A list consisting of two entries:
+#'
+#' \code{$problem} - logical indicating whether a problem was found by the
+#' \code{checkFunction} responsible for the making the \code{messageGenerator()} call,
+#'
+#' \code{$problemValues} - a vector of values from the variable that were
 #' deemed problematic (see details below).
 #'
 #' @param message Optional, but recommended. A message describing what problem the
@@ -20,14 +20,14 @@
 #' values are printed.
 #'
 #' @details This function is a tool for building \code{\link{checkFunction}}s for the
-#' \code{cleanR} \code{\link{clean}} function. \code{checkFunction}s will often identify a number
+#' \code{dataMaid} \code{\link{clean}} function. \code{checkFunction}s will often identify a number
 #' of values in a variable that are somehow problematic. \code{messageGenerator} takes
 #' these values, pastes them together with a problem description and makes sure that the
-#' formatting is appropriate for being rendered in a \code{rmarkdown} document. 
+#' formatting is appropriate for being rendered in a \code{rmarkdown} document.
 #' We recommend writing short and precise problem descriptions (see examples),
 #'  but if no message is supplied, the following message is generated:
 #' "[function name] found the following problem values: [problem values]".
-#' Note that this option only works correctly when the \code{checkFunction} is 
+#' Note that this option only works correctly when the \code{checkFunction} is
 #' called directly and e.g. not dispatched through a generic function.
 #'
 #' @return A character string with a problem description.
@@ -50,7 +50,7 @@
 #'   message <- messageGenerator(list(problemValues = problemValues, problem = problem),
 #'                               "The following values contain underscores:",
 #'                               nMax = nMax)
-#'   checkResult(list(problem = problem, message = message, 
+#'   checkResult(list(problem = problem, message = message,
[…]

### 2017-01-02 R/render.R

@@ -1,23 +1,23 @@
 #' @title Simplified Rmarkdown rendering
-#' 
+#'
 #' @description Render a Rmarkdown (.Rmd) file, \code{file}, to the output
 #' format specified in its preamble. If no output format is specified,
 #' it will be rendered to html.
-#' 
-#' @param file A character string path to the file that is to be rendered. 
+#'
+#' @param file A character string path to the file that is to be rendered.
 #' This file must be of type Rmarkdown (.Rmd)
-#' 
+#'
 #' @param quiet A logical. Should messages during rendering be surpressed?
-#' 
-#' @details This function is merely a simplified version (in terms of 
-#' possible arguments) of the rendering function from the \code{rmarkdown} package. 
-#' Therefore, we refer to this functions for more details:  
-#' \code{\link[rmarkdown]{render}}. We have included this simplified version in 
-#' \code{cleanR} in order to help new R users with rendering their output 
+#'
+#' @details This function is merely a simplified version (in terms of
+#' possible arguments) of the rendering function from the \code{rmarkdown} package.
+#' Therefore, we refer to this functions for more details:
+#' \code{\link[rmarkdown]{render}}. We have included this simplified version in
+#' \code{dataMaid} in order to help new R users with rendering their output
 #' documents as generated by \code{\link{clean}}.
-#' 
+#'
 #' @seealso \code{\link[rmarkdown]{render}}.
-#' 
+#'
 #' @export
 render <- function(file, quiet) rmarkdown::render(file, quiet=quiet)
 

### 2017-01-02 R/variableType.R

@@ -1,6 +1,6 @@
 #' @title Summary function for original class
 #'
-#' @description A \code{\link{summaryFunction}} type function, intended to be called from 
+#' @description A \code{\link{summaryFunction}} type function, intended to be called from
 #' \code{\link{summarize}} to be called from \code{\link{summarize}}, which finds the
 #' orignial class of a variable. This is just the class for all objects but those of class
 #' \code{smartNum}.
@@ -9,8 +9,8 @@
 #'
 #' @param ... Not in use.
 #'
-#' @return An object of class \code{summaryResult} with the following entries: 
-#' \code{$feature} ("Variable type"), \code{$result} (the (original) class of 
+#' @return An object of class \code{summaryResult} with the following entries:
+#' \code{$feature} ("Variable type"), \code{$result} (the (original) class of
 #' \code{v}) and \code{$value} (identical to \code{$result}).
 #'
 #' @seealso \code{\link{summarize}}
@@ -22,7 +22,7 @@
 #'    variableType(varX)
 #'
 #'  #For smartNum variables:
-#'    smartX <- cleanR:::smartNum(varX)
+#'    smartX <- dataMaid:::smartNum(varX)
 #'    class(smartX)
 #'    variableType(smartX)
 #'

### 2017-01-03 latex/article_vol2.tex
Updated example pdfs to include the new package name.

@@
 \begin{center}
-\frame{\includegraphics[width=7.5cm,page=2]{cleanR_toyData.pdf}} 
-\frame{\includegraphics[width=7.5cm,page=3]{cleanR_toyData.pdf}}
+\frame{\includegraphics[width=7.5cm,page=2]{dataMaid_toyData.pdf}} 
+\frame{\includegraphics[width=7.5cm,page=3]{dataMaid_toyData.pdf}}
 %\includepdf[pages={2}, pagecommand={}]{dataMaid_testData.pdf}
@@
 \quad \R{listChecks} & Should an overview of what checks were performed be listed in the \textit{Data cleaning summary}? &  \R{TRUE} \\
-\quad \R{preChecks} & What check functions should be called to determine whether a variable is suitable for summarization, visualization and checking? & \R{c("isKey", "isEmpty")}
+\quad \R{preChecks} & What check functions should be called to determine whether a variable is suitable for summarization, visualization and checking? & \R{c("isKey", "isEmpty")} \\
 
@@
 \\
-Tools available for producing the function & \R{summaryResult()} & --- & \R{messageGenerator()} \newline \R{checkResult()} 
+Tools available for producing the function & \R{summaryResult()} & - & \R{messageGenerator()} \newline \R{checkResult()} 
 \end{tabular}
@@
 % \section{Appendix A: cleaning the toyData data frame} \label{sec:appendix1}
-\includepdf[pages=2, pagecommand={\section{Appendix A: cleaning the toyData data frame} \label{sec:appendix1}}, frame=true, noautoscale=true, scale=0.7]{cleanR_toyData.pdf}
-\includepdf[pages=3-4, pagecommand={}, frame=true, noautoscale=true, scale=0.7,pagecommand={}]{cleanR_toyData.pdf}
+\includepdf[pages=2, pagecommand={\section{Appendix A: cleaning the toyData data frame} \label{sec:appendix1}}, frame=true, noautoscale=true, scale=0.7]{dataMaid_toyData.pdf}
+\includepdf[pages=3-4, pagecommand={}, frame=true, noautoscale=true, scale=0.7,pagecommand={}]{dataMaid_toyData.pdf}
 
 \includepdf[pages=2, pagecommand={\section{Appendix B: User-extended cleaning of
-  exampleData} \label{sec:appendix2}}, frame=true, noautoscale=true, scale=0.7]{cleanR_exampleData.pdf}
-\includepdf[pages=3-, pagecommand={}, frame=true, noautoscale=true, scale=0.7]{cleanR_exampleData.pdf}
+  exampleData} \label{sec:appendix2}}, frame=true, noautoscale=true, scale=0.7]{dataMaid_exampleData.pdf}
+\includepdf[pages=3-, pagecommand={}, frame=true, noautoscale=true, scale=0.7]{dataMaid_exampleData.pdf}
 %\hl{Data cleaning with user supplied extensions here}

### 2017-01-03 latex/article_vol2.tex
Rettes

@@
 \Plaintitle{{dataMaid}: your assistant for data cleaning in R} %% without formatting
-\Shorttitle{\pkg{dataMaid}: your data cleaning PA in R} %% a short title (if necessary)
+\Shorttitle{\pkg{dataMaid}: your data cleaning personal assistant in R} %% a short title (if necessary)
 
@@
 the unfortunate combination of memory control in this operating system
-and RStudio, \hl{And also just R, right? Du har udkommenterert denne kommentar,
+and \proglang{R}.
+\hl{And also just R, right? Du har udkommenterert denne kommentar,
 men jeg mener stadig at det er tilfældet, og jeg mener desuden ikke vi kan skrive 
@@
 This also deals with the fact that \pkg{dataMaid} can produce
-\pkg{rmarkdown} files that supersedes the upper size limit. \hl{dette giver
-ingen mening uden en præcision? "upper size limit" af hvad? Se udkommenteret 
-kommentar.}
+\pkg{rmarkdown} files that supersedes the upper size limit of, say,
+RStudio (which currently has an editor file size limit of 5MB). %\hl{dette giver
+%ingen mening uden en præcision? "upper size limit" af hvad? Se udkommenteret 
+%kommentar.}
 % of RStudio,

### 2017-01-03 latex/article_vol2.tex
Memory issues

@@
 \def\arraystretch{1.8}% 
-\hspace*{-.2cm}\begin{tabular}[t]{p{0.45\linewidth} | p{0.45\linewidth}} \vspace*{-1cm}If \R{doEval} is \R{TRUE}: \newline A plot that is opened by the graphic device in \R{R}.
-& \vspace*{-1cm}If \R{doEval} is \R{FALSE}: \newline   A text string with \R{R} code, as described above.
+\hspace*{-.2cm}\begin{tabular}[t]{p{0.45\linewidth} |
+                 p{0.45\linewidth}} \vspace*{-1cm}\raggedright If
+                 \R{doEval} is \R{TRUE}: \newline A plot that is
+                 opened by the graphic device in \R{R}. &
+                                                          \vspace*{-1cm}\raggedright If \R{doEval} is \R{FALSE}: \newline   A text string with \R{R} code, as described above.
 \end{tabular} \egroup  & A \R{checkResult} object, i.e.
@@
 \\
-Tools available for producing the function & \R{summaryResult()} & - & \R{messageGenerator()} \newline \R{checkResult()} 
+Tools available for producing the function & \R{summaryResult()} & --- & \R{messageGenerator()} \newline \R{checkResult()} 
 \end{tabular}
@@
 \subsubsection{Economic memory use} 
-Another solution, which is especially relevant to Windows users due to
-the unfortunate combination of memory control in this operating system
-and \proglang{R}.
-\hl{And also just R, right? Du har udkommenterert denne kommentar,
-men jeg mener stadig at det er tilfældet, og jeg mener desuden ikke vi kan skrive 
-det uden en præcision af hvad det går ud på. Hvad tænker du?}
+Available memory may pose a problem with larger datasets ---
+especially on machiens running Windows. The problem is, that older,
+32-bit versions of Windows cannot necessary access all physical RAM,
+while newer 64-bit versions of Windows cannot access virtual memory,
+which makes \proglang{R} unnecessary slow.  Another solution, which is
+especially relevant to Windows users due to %the unfortunate
+the memory control for this operating system,
+%combination of memory control in this operating system and
+%\proglang{R},
+%\hl{And also just R, right? Du har udkommenterert denne kommentar,
+%men jeg mener stadig at det er tilfældet, og jeg mener desuden ikke vi kan skrive 
+%det uden en præcision af hvad det går ud på. Hvad tænker du?}
 is simply splitting the two steps performed

### 2017-01-03 latex/article_vol2.tex
Minor revisions of section 5.1

@@
 slow, while a large number of observations will generally affect the
-rendering time of the document. In this section, we give a few
+rendering time of the document. This issue is especially relevant to 
+Windows users, where memory usage is not very efficient. The problem is that older,
+32-bit versions of Windows generally cannot access all physical RAM,
+while newer 64-bit versions of Windows cannot access virtual memory. Therefore,
+memory allocation issues becomes even more pressing when using this operating system.
+
+In this section, we give a few
 practical examples of ways to deal with large data, while wishing to
@@
 \subsubsection{Economic memory use} 
-Available memory may pose a problem with larger datasets ---
-especially on machiens running Windows. The problem is, that older,
-32-bit versions of Windows cannot necessary access all physical RAM,
-while newer 64-bit versions of Windows cannot access virtual memory,
-which makes \proglang{R} unnecessary slow.  Another solution, which is
+Another solution, which is
 especially relevant to Windows users due to %the unfortunate
-the memory control for this operating system,
+the memory handling on this operating system,
 %combination of memory control in this operating system and
 %\proglang{R},
-%\hl{And also just R, right? Du har udkommenterert denne kommentar,
-%men jeg mener stadig at det er tilfældet, og jeg mener desuden ikke vi kan skrive 
-%det uden en præcision af hvad det går ud på. Hvad tænker du?}
 is simply splitting the two steps performed
@@
 This also deals with the fact that \pkg{dataMaid} can produce
-\pkg{rmarkdown} files that supersedes the upper size limit of, say,
-RStudio (which currently has an editor file size limit of 5MB). %\hl{dette giver
+\pkg{rmarkdown} files that supersedes the upper size limit of code editors, e.g.
+RStudio, which currently has an editor file size limit of 5MB. %\hl{dette giver
 %ingen mening uden en præcision? "upper size limit" af hvad? Se udkommenteret 

### 2017-01-03 latex/article_vol2.tex
Fixed typos and made script

@@
 > data(toyData)
-> toydata
+> toyData
 \end{Sinput}
@@
 > toyData$var2[toyData$var2 %in% missVar2$problemValues] <- NA
-> identifyMissing(toyData$var2}
+> identifyMissing(toyData$var2)
 \end{Sinput}
@@
 \begin{Sinput}
-> clean(toyData, render = FALSE, openResult = FALSE}
+> clean(toyData, render = FALSE, openResult = FALSE)
 > render("dataMaid_toyData.Rmd", quiet = FALSE)

### 2017-01-03 latex/dataMaid_toyData.Rmd
Fixed typos and made script

---
dataMaid: yes
title: toyData
subtitle: "Autogenerated data summary from dataMaid"
date: "`r Sys.Date()`"
output: pdf_document
documentclass: report
header-includes:
  - \renewcommand{\chaptername}{Part}
  - \newcommand{\fullline}{\noindent\makebox[\linewidth]{\rule{\textwidth}{0.4pt}}}
---
```{r , echo=FALSE, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
library(ggplot2)
library(pander)
```
# Data cleaning summary
The dataset examined has the following dimensions:

------------------------------
Feature                 Result
--------------------- --------
Number of rows              15

Number of variables          6
------------------------------

* Only variables that were deemed potentially problematic are included in this summary.

### Checks performed
The following variable checks were performed, depending on the data type of each variable:

----------------------------------------------------------------------------------------------------------------------------------
&nbsp;                                                 character    factor    labelled   numeric    integer    logical     Date   
----------------------------------------------------- ----------- ---------- ---------- ---------- ---------- --------- ----------
Identify miscoded missing values                       $\times$    $\times$   $\times$   $\times$   $\times$

Identify prefixed and suffixed whitespace              $\times$    $\times$   $\times$

Identify levels with < 6 obs.                          $\times$    $\times$   $\times$

Identify case issues                                   $\times$    $\times$   $\times$

Identify misclassified numeric or integer variables    $\times$    $\times$   $\times$

Identify outliers                                                                        $\times$   $\times$             $\times$ 
----------------------------------------------------------------------------------------------------------------------------------

Please note that all numerical values in the following have been rounded to 2 decimals.

# Variable list
## **var5**

* The variable is a key (distinct values for each observation).

\fullline

## **var6**

* The variable only takes one (non-missing) value: \"Irrelevant\". The variable contains 0 \% missing observations.

\fullline

This report was created by dataMaid v0.9.2.

### 2017-01-24 latex/article_vol2.Rnw
Move to Sweave

--- 2017-01-03 latex/article_vol2.tex
@@
 \usepackage[utf8]{inputenc}
-\usepackage[nogin]{Sweave}
 \usepackage{natbib}
@@
 \renewcommand{\floatpagefraction}{.6}
-
 \renewcommand{\topfraction}{0.85}
@@
 \begin{document}
+
+<<echo=FALSE>>=
+options(prompt = "R> ", width=77, continue = "+  ")
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-> library(dataMaid)
-> data(toyData)
-> toyData
-\end{Sinput}
-\begin{Soutput}
-   var1 var2  var3        var4 var5       var6
-1   red    1     a -0.65959383    1 Irrelevant
-2   red    1     a  0.08671649    2 Irrelevant
-3   red    1     a -0.10951326    3 Irrelevant
-4   red    2     a  0.08630221    4 Irrelevant
-5   red    2     a -1.84311184    5 Irrelevant
-6   red    6     b  0.92210680    6 Irrelevant
-7   red    6     b  1.01921086    7 Irrelevant
-8   red    6     b -0.92428326    8 Irrelevant
-9   red  999     c -0.65340163    9 Irrelevant
-10  red   NA     c  0.21133941   10 Irrelevant
-11 blue    4     c  0.91783009   11 Irrelevant
-12 blue   82     .  0.10313983   12 Irrelevant
-13 blue   NA        0.16954218   13 Irrelevant
-14 <NA>  NaN other  0.41967230   14 Irrelevant
-15 <NA>    5 OTHER  0.77143836   15 Irrelevant
-\end{Soutput}
-\begin{Sinput}
-> clean(toyData)
-\end{Sinput}
-\end{Schunk}
+<<eval=FALSE>>=
+library("dataMaid")
+data("toyData")
+toyData
+clean(toyData)
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, output = "html", onlyProblematic = TRUE, maxProbVals = 10,
+<<eval=FALSE>>=
+clean(toyData, output = "html", onlyProblematic = TRUE, maxProbVals = 10,
   replace = TRUE)
-\end{Sinput}
-\end{Schunk}
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, output="html", render=FALSE, openResult=FALSE)
-\end{Sinput}
-\end{Schunk}
+<<eval=FALSE>>=
+clean(toyData, output="html", render=FALSE, openResult=FALSE)
+@
 
@@
 \begin{Sinput}
-> data(exampleData) 
+> data("exampleData")
 > clean(exampleData,

### 2017-01-24 latex/article_vol2.Rnw
Move to Sweave

@@
 contributing to an inter-field dialogue about the data at
-hand. Oftentimes, e.g. distinguishing between faulty codings of a
+hand. Oftentimes, e.g., distinguishing between faulty codings of a
 numeric value and unusual, but correct, values requires
@@
 
-<<eval=FALSE>>=
+<<>>=
 library("dataMaid")
@@
 toyData
+@
+<<eval=FALSE>>=
 clean(toyData)
@@
 
-\begin{Schunk}
-\begin{Sinput}
-> allSummaryFunctions()
-\end{Sinput}
-\begin{Soutput}
-----------------------------------------------------------------------------
-name           description                     classes
--------------- ------------------------------- -----------------------------
-centralValue   Compute median or mode          character, Date, factor,
-                                               integer, labelled, logical,
-                                               numeric
-
-countMissing   Compute ratio of missing        character, Date, factor,
-               observations                    integer, labelled, logical,
-                                               numeric
-
-minMax         Find minimum and maximum        Date, integer, numeric
-               values
-
-quartiles      Compute 1st and 3rd quartiles   Date, integer, numeric
-
-uniqueValues   Count number of unique values   character, Date, factor,
-                                               integer, labelled, logical,
-                                               numeric
-
-variableType   Data class of variable          character, Date, factor,
-                                               integer, labelled, logical,
-                                               numeric
-----------------------------------------------------------------------------
-\end{Soutput}
-\end{Schunk}
+<<>>=
+allSummaryFunctions()
+@
 
@@
 For example, the default summaries being used for a factor variable is
-\begin{Schunk}
-\begin{Sinput}
-> defaultFactorSummaries()
-\end{Sinput}
-\begin{Soutput}
-[1] "variableType" "countMissing" "uniqueValues" "centralValue"
-\end{Soutput}
-\end{Schunk}
+<<>>=
+defaultFactorSummaries()
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, characterSummaries = c("variableType", "centralValue"),
+<<eval=FALSE>>=
+clean(toyData, characterSummaries = c("variableType", "centralValue"),
   factorSummaries = c("variableType", "centralValue"),
@@
   dateSummaries = c("variableType", "centralValue"))
-\end{Sinput}
-\end{Schunk}
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, allSummaries = c("variableType", "centralValue"))
-\end{Sinput}
-\end{Schunk}
+<<eval=FALSE>>=
+clean(toyData, allSummaries = c("variableType", "centralValue"))
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-> defaultFactorChecks()
-\end{Sinput}
-\begin{Soutput}
-[1] "identifyMissing"    "identifyWhitespace" "identifyLoners"
-[4] "identifyCaseIssues" "identifyNums"
-\end{Soutput}
-\end{Schunk}
+<<>>=
+defaultFactorChecks()
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, factorChecks = c("identifyWhitespace"))
-\end{Sinput}
-\end{Schunk}
+<<eval=FALSE>>=
+clean(toyData, factorChecks = c("identifyWhitespace"))
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, factorChecks = NULL)
-\end{Sinput}
-\end{Schunk}
+<<eval=FALSE>>=
+clean(toyData, factorChecks = NULL)
+@
 
@@
 
-Note that the \R{summarize()}, \R{check()} and \R{visualize()} functions are also available interactively for full datasets, by calling e.g.
-\R{summarize(toyData)}.
+Note that the \R{summarize()}, \R{check()} and \R{visualize()} functions are also available interactively for full datasets, by calling e.g., \R{summarize(toyData)}.
 However, this produces an extensive amount of output in the console, and therefore, we generally do not recommend it, unless working with very small datasets or subsets of datasets.
@@
 Input (optional) &  \R{maxDecimals} - number of decimals printed in outputted numerical values.  & - &  \R{maxDecimals}  - number of decimals printed in outputted numerical values.  \\
-Purpose & Describe some aspect of the variable, e.g.
-a central value, its dispersion or level of missingness.
+Purpose & Describe some aspect of the variable, e.g., a central value, its dispersion or level of missingness.
 & Produce a distribution plot.
@@
 Output (required) & A list with entries: \newline \R{\$feature} - a label for the summary value (as character string); \newline \R{\$result} - the result of the summary (as character string) & A character string with \R{R} code for producing a plot.
-This code should be standalone, i.e.
-should include the data if necessary.
+This code should be standalone, i.e., should include the data if necessary.
 & A list with entries: \newline \R{\$problem} - a logical identifying whether an issue was found; \newline \R{\$message} - a character string (possibly empty) decribing the issue that was found, properly escaped and ready for use in \R{rmarkdown} \\
-Output (recommended) & A \R{summaryResult} object, i.e.
-an attributed list with entries \R{\$feature}, \R{\$result} and \R{\$value}, the latter being the values from \R{\$result} in their original format).
+Output (recommended) & A \R{summaryResult} object, i.e., an attributed list with entries \R{\$feature}, \R{\$result} and \R{\$value}, the latter being the values from \R{\$result} in their original format).
 & \bgroup
@@
                                                           \vspace*{-1cm}\raggedright If \R{doEval} is \R{FALSE}: \newline   A text string with \R{R} code, as described above.
-\end{tabular} \egroup  & A \R{checkResult} object, i.e.
-an attributed list with entries \R{\$problem}, \R{\$message} and \R{\$problemValues}, the latter being either \R{NULL} or the problem causing values, as they were found in \R{v}, whichever is relevant.
+\end{tabular} \egroup  & A \R{checkResult} object, i.e., an attributed list with entries \R{\$problem}, \R{\$message} and \R{\$problemValues}, the latter being either \R{NULL} or the problem causing values, as they were found in \R{v}, whichever is relevant.
 \\
@@
 character string or something that will be printed as one. In other
-words, e.g. integers are allowed, but matrices are not. Though a lot
+words, e.g., integers are allowed, but matrices are not. Though a lot
 of different things can go into the \code{summaryFunction} template,
@@
 e.g., \R{identifyMissing()} does), or it performs a check concerning the
-variable as a whole (e.g. the functions used for prechecks and the
+variable as a whole (e.g., the functions used for prechecks and the
 function \R{identifyNums()}). We present templates for both types of
@@
 \pkg{dataMaid} might be the inclusion of ASCII plots, as
-e.g. represented in the \proglang{R} package \pkg{txtplot}.
+e.g., represented in the \proglang{R} package \pkg{txtplot}.
 
@@
 This also deals with the fact that \pkg{dataMaid} can produce
-\pkg{rmarkdown} files that supersedes the upper size limit of code editors, e.g.
+\pkg{rmarkdown} files that supersedes the upper size limit of code editors, e.g.,
 RStudio, which currently has an editor file size limit of 5MB. %\hl{dette giver

### 2017-01-24 latex/article_vol2.tex
Move to Sweave

@@
 \begin{Sinput}
-> library(dataMaid)
-> data(toyData)
-> toyData
+R> library("dataMaid")
+R> data("toyData")
+R> toyData
 \end{Sinput}
@@
    var1 var2  var3       var4 var5       var6
-1   red    1     a -0.65959383    1 Irrelevant
-2   red    1     a  0.08671649    2 Irrelevant
-3   red    1     a -0.10951326    3 Irrelevant
-4   red    2     a  0.08630221    4 Irrelevant
-5   red    2     a -1.84311184    5 Irrelevant
-6   red    6     b  0.92210680    6 Irrelevant
-7   red    6     b  1.01921086    7 Irrelevant
-8   red    6     b -0.92428326    8 Irrelevant
-9   red  999     c -0.65340163    9 Irrelevant
-10  red   NA     c  0.21133941   10 Irrelevant
-11 blue    4     c  0.91783009   11 Irrelevant
-12 blue   82     .  0.10313983   12 Irrelevant
-13 blue   NA        0.16954218   13 Irrelevant
-14 <NA>  NaN other  0.41967230   14 Irrelevant
-15 <NA>    5 OTHER  0.77143836   15 Irrelevant
+1   red    1     a -0.6264538    1 Irrelevant
+2   red    1     a  0.1836433    2 Irrelevant
+3   red    1     a -0.8356286    3 Irrelevant
+4   red    2     a  1.5952808    4 Irrelevant
+5   red    2     a  0.3295078    5 Irrelevant
+6   red    6     b -0.8204684    6 Irrelevant
+7   red    6     b  0.4874291    7 Irrelevant
+8   red    6     b  0.7383247    8 Irrelevant
+9   red  999     c  0.5757814    9 Irrelevant
+10  red   NA     c -0.3053884   10 Irrelevant
+11 blue    4     c  1.5117812   11 Irrelevant
+12 blue   82     .  0.3898432   12 Irrelevant
+13 blue   NA       -0.6212406   13 Irrelevant
+14 <NA>  NaN other -2.2146999   14 Irrelevant
+15 <NA>    5 OTHER  1.1249309   15 Irrelevant
 \end{Soutput}
-\begin{Sinput}
-> clean(toyData)
+\end{Schunk}
+\begin{Schunk}
+\begin{Sinput}
+R> clean(toyData)
 \end{Sinput}
@@
 \begin{Sinput}
-> clean(toyData, output = "html", onlyProblematic = TRUE, maxProbVals = 10,
-        replace = TRUE)
+R> clean(toyData, output = "html", onlyProblematic = TRUE, maxProbVals = 10,
++    replace = TRUE)
 \end{Sinput}
@@
 \begin{Sinput}
-> clean(toyData, output="html", render=FALSE, openResult=FALSE)
+R> clean(toyData, output="html", render=FALSE, openResult=FALSE)
 \end{Sinput}
@@
 \begin{Sinput}
-> allSummaryFunctions()
+R> allSummaryFunctions()
 \end{Sinput}
@@
 
-minMax         Find minimum and maximum        Date, integer, numeric        
+minMax         Find minimum and maximum        integer, numeric, Date       
                values
@@
 \begin{Sinput}
-> defaultFactorSummaries()
+R> defaultFactorSummaries()
 \end{Sinput}
@@
 \begin{Sinput}
-> clean(toyData, characterSummaries = c("variableType", "centralValue"),
-        factorSummaries = c("variableType", "centralValue"),
-        labelledSummaries = c("variableType", "centralValue"),
-        numericSummaries = c("variableType", "centralValue"),
-        integerSummaries = c("variableType", "centralValue"),
-        logicalSummaries = c("variableType", "centralValue"),
-        dateSummaries = c("variableType", "centralValue"))
+R> clean(toyData, characterSummaries = c("variableType", "centralValue"),
++    factorSummaries = c("variableType", "centralValue"),
++    labelledSummaries = c("variableType", "centralValue"),
++    numericSummaries = c("variableType", "centralValue"),
++    integerSummaries = c("variableType", "centralValue"),
++    logicalSummaries = c("variableType", "centralValue"),
++    dateSummaries = c("variableType", "centralValue"))
 \end{Sinput}
@@
 \begin{Sinput}
-> clean(toyData, allSummaries = c("variableType", "centralValue"))
+R> clean(toyData, allSummaries = c("variableType", "centralValue"))
 \end{Sinput}
@@
 \begin{Sinput}
-> defaultFactorChecks()
+R> defaultFactorChecks()
 \end{Sinput}
@@
 \begin{Sinput}
-> clean(toyData, factorChecks = c("identifyWhitespace"))
+R> clean(toyData, factorChecks = c("identifyWhitespace"))
 \end{Sinput}
@@
 \begin{Sinput}
-> clean(toyData, factorChecks = NULL)
+R> clean(toyData, factorChecks = NULL)
 \end{Sinput}

### 2017-01-24 latex/article_vol2.Rnw
Bibliography

@@
   by an investigator with an understanding of the variables in the
-  data, but not necessarily knowledge of \R{R}. Thereby, \pkg{dataMaid}
+  data, but not necessarily knowledge of \proglang{R}. Thereby, \pkg{dataMaid}
   aids the dialogue between data
@@
 Online tools such as OpenRefine (\url{http://openrefine.org/}) and
-\proglang{R}-packages such as \pkg{plyr}, and \pkg{data.table} have
+\proglang{R}-packages such as \pkg{plyr} \citep{plyr}, and \pkg{data.table} have
 made data wrangling a lot easier, but only a handful of packages such
@@
 \ref{sec:usingdataMaid}, we introduce the main representative of the first
-paradigm, namely the \R{clean()} function, which generates data
+paradigm, namely the \code{clean()} function, which generates data
 cleaning overview documents. In the \pkg{dataMaid} package, we have
@@
 \ref{sec:extending}, we turn to the question of how \pkg{dataMaid} extensions
-can be made, such that they are integrable with the \R{clean()}
+can be made, such that they are integrable with the \code{clean()}
 function and with the other tools available in \pkg{dataMaid}.  At last,
@@
 
-The \R{clean()} function is the primary workhorse of \pkg{dataMaid} and
+The \code{clean()} function is the primary workhorse of \pkg{dataMaid} and
 this is the only function needed if a standard battery
@@
 provides an example of a data cleaning output document, produced by
-calling \R{clean()} on the dataset \R{toyData} available in
+calling \code{clean()} on the dataset \code{toyData} available in
 \pkg{dataMaid}. The first two pages of this data cleaning output are
-shown in Figure~\ref{fig:example1}. \R{toyData} is a very
+shown in Figure~\ref{fig:example1}. \code{toyData} is a very
 small ($15$ rows of $6$ variables), artificial dataset which was created to
@@
 \label{fig:example1}
-\caption{Example output from running \R{clean()} on the \R{toyData}
+\caption{Example output from running \code{clean()} on the \code{toyData}
   dataset. First a summary of the full dataset is given and then
@@
 variable, a figure visualizing its distribution and a
-list of flagged issues, if any. For instance, in the \R{numeric}-type variable
-\R{var2} from \R{toyData}, \R{clean()} has identified two values that
-are suspected to be miscoded missing values (\R{999} and \R{NaN}),
+list of flagged issues, if any. For instance, in the \code{numeric}-type variable
+\code{var2} from \code{toyData}, \code{clean()} has identified two values that
+are suspected to be miscoded missing values (\code{999} and \code{NaN}),
 while two values were also flagged as potential outliers that should
@@
 
-Though the \R{clean()} function is very easy to use, it should not be
+Though the \code{clean()} function is very easy to use, it should not be
 mistaken to be inflexible. A large number of function arguments allows
@@
 part of the cleaning process they influence. In order to understand
-this distinction, a glimpse of the inner structure of \R{clean()} is
+this distinction, a glimpse of the inner structure of \code{clean()} is
 shown in Figure~\ref{figure:cleanStructure}. Below, we present a few
 examples on how to use the arguments from Table \ref{table.cleanFormals}
- to influence the output of a \R{clean()} call.
+ to influence the output of a \code{clean()} call.
 
@@
 \smallskip Control input variables and summary\\
-\quad \R{useVar} & What variables should be used? & \R{NULL} (corresponding to all variables) \\
-\quad \R{ordering} & Ordering of the variables in the data summary (as is or alphabetical) & \R{"asIs"} \\
-\quad \R{onlyProblematic} & Should only variables flagged as problematic be included in the \textit{Variable list}? & \R{FALSE} \\
-\quad \R{listChecks} & Should an overview of what checks were performed be listed in the \textit{Data cleaning summary}? &  \R{TRUE} \\
-\quad \R{preChecks} & What check functions should be called to determine whether a variable is suitable for summarization, visualization and checking? & \R{c("isKey", "isEmpty")} \\
+\quad \code{useVar} & What variables should be used? & \code{NULL} (corresponding to all variables) \\
+\quad \code{ordering} & Ordering of the variables in the data summary (as is or alphabetical) & \code{"asIs"} \\
+\quad \code{onlyProblematic} & Should only variables flagged as problematic be included in the \textit{Variable list}? & \code{FALSE} \\
+\quad \code{listChecks} & Should an overview of what checks were performed be listed in the \textit{Data cleaning summary}? &  \code{TRUE} \\
+\quad \code{preChecks} & What check functions should be called to determine whether a variable is suitable for summarization, visualization and checking? & \code{c("isKey", "isEmpty")} \\
 
 \smallskip Control summarize, visualize, and check steps \\
-\quad \R{mode} & What steps should be performed for each variable (out
+\quad \code{mode} & What steps should be performed for each variable (out
                  of the three possibilities \textit{summarize},
                  \textit{visualize}, \textit{check})? &
-                                                        \R{c("summarize", "visualize", "check")} (corresponding to all three steps) \\
-\quad \R{smartNum} & Should numerical values with only a few unique
-                     levels be flagged and treated as a factor variable? & \R{TRUE} \\
-\quad \R{maxProbVals} & Maximum number of problematic values to print, if any are found in data checks & \R{Inf} \\
-\quad \R{maxDecimals} & Maximum number of decimals to print for numeric values in the variable list & \R{2} \\
-\quad \R{twoCol} & Should the summary table and visualizations be placed side-by-side (in two columns)? & \R{TRUE} \\
+                                                        \code{c("summarize", "visualize", "check")} (corresponding to all three steps) \\
+\quad \code{smartNum} & Should numerical values with only a few unique
+                     levels be flagged and treated as a factor variable? & \code{TRUE} \\
+\quad \code{maxProbVals} & Maximum number of problematic values to print, if any are found in data checks & \code{Inf} \\
+\quad \code{maxDecimals} & Maximum number of decimals to print for numeric values in the variable list & \code{2} \\
+\quad \code{twoCol} & Should the summary table and visualizations be placed side-by-side (in two columns)? & \code{TRUE} \\
 
 \smallskip Control output and post-processing \\
-\quad \R{output} & Type of output file to be produced (html, or pdf) & \R{"pdf"} \\
-\quad \R{render} & Should the output file be rendered from markdown? & \R{TRUE} \\
-\quad \R{openResult} & If a  pdf/html file is rendered, should it
+\quad \code{output} & Type of output file to be produced (html, or pdf) & \code{"pdf"} \\
+\quad \code{render} & Should the output file be rendered from markdown? & \code{TRUE} \\
+\quad \code{openResult} & If a  pdf/html file is rendered, should it
                        automatically open afterwards, and if not,
-                       should the \R{rmarkdown} file be opened? & \R{TRUE} \\
-\quad \R{replace} & Overwrite an existing files with the same name? & \R{FALSE} \\
+                       should the \code{rmarkdown} file be opened? & \code{TRUE} \\
+\quad \code{replace} & Overwrite an existing files with the same name? & \code{FALSE} \\
 
@@
 \end{tabular}
-\caption{A selection of commonly used arguments to \R{clean()} separated into the parts they control.}
+\caption{A selection of commonly used arguments to \code{clean()} separated into the parts they control.}
 \label{table.cleanFormals}
@@
 % to emphasize that two levels of control are available when using
-% \R{clean()}. Either, we can control the \hl{[something] - essentially,
+% \code{clean()}. Either, we can control the \hl{[something] - essentially,
 %   this is everything but SVC-functions - } or we can control what
@@
 % 	\begin{itemize}
-% 		\item This should be of type \R{data.frame} or \R{tibble} \hl{or data.table?}
+% 		\item This should be of type \code{data.frame} or \code{tibble} \hl{or data.table?}
 % 	\end{itemize}
@@
 % 			\begin{description}
-% 				\item[Summarize] Call \R{summarize()} to produce a summary table describing the variable. What features enter this table depends on the data class of the variable.
-% 				\item[Visualize]  Call \R{visualize()} to produce a plot visualizing the distribution of the variable.
-% 				\item[Check]  Call \R{check()} to apply quality- and error checks to the variable. What checks are used depends on the data class of the variable.
+% 				\item[Summarize] Call \code{summarize()} to produce a summary table describing the variable. What features enter this table depends on the data class of the variable.
+% 				\item[Visualize]  Call \code{visualize()} to produce a plot visualizing the distribution of the variable.
+% 				\item[Check]  Call \code{check()} to apply quality- and error checks to the variable. What checks are used depends on the data class of the variable.
 % 			\end{description}
@@
 % 	\begin{itemize}
-% 		\item Always a \R{rmarkdown} (.Rmd) file
+% 		\item Always a \code{rmarkdown} (.Rmd) file
 % 		\item Possible also a html or pdf file
@@
 \caption{Schematic illustration of the stages undertaken when running
-  \R{clean()}. Each variable is checked for eligibility before
-  running \R{summarize()}, \R{visualize()}, and \R{check()}, and the
+  \code{clean()}. Each variable is checked for eligibility before
+  running \code{summarize()}, \code{visualize()}, and \code{check()}, and the
   resulting \proglang{R} markdown file may be rendered and opened.}
@@
 \subsection{Polishing off the arguments}
-We begin with an example that is intended as an illustration of how \R{clean()} might
+We begin with an example that is intended as an illustration of how \code{clean()} might
 be used in the very first stages of data cleaning, where we are still uncertain
@@
 maximum 10 printed potential errors for each variable we can write
-(output not shown). Also, we can add the argument \R{replace=TRUE} in
-order to force \R{clean()} to overwrite any existing files produced by
-\R{clean()}, as we might run this function a few times in a row before deciding
-exactly which cleaning overview we want. Using the \R{toyData} dataset as a guinea pig, we type:
+(output not shown). Also, we can add the argument \code{replace=TRUE} in
+order to force \code{clean()} to overwrite any existing files produced by
+\code{clean()}, as we might run this function a few times in a row before deciding
+exactly which cleaning overview we want. Using the \code{toyData} dataset as a guinea pig, we type:
 
@@
 The final rendering of the generated markdown file is controlled by
-the \R{render} and \R{openResult} arguments, which both default to
-\R{TRUE}. \R{render} determines if the \proglang{R} markdown file
+the \code{render} and \code{openResult} arguments, which both default to
+\code{TRUE}. \code{render} determines if the \proglang{R} markdown file
 produced should be rendered using the \pkg{rmarkdown} package and
-\R{openResult} decides whether the output html or pdf file should be
+\code{openResult} decides whether the output html or pdf file should be
 opened. The following command produces an \proglang{R} markdown file
@@
   \hline \smallskip
-  \textbf{\R{summaryFunction}s}  \smallskip \\
-  \quad \R{centralValue} & Compute median or mode &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\
-  \quad \R{countMissing} & Compute ratio of missing observations &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\
-  \quad \R{minMax} & Find minimum and maximum values &   &  & $\times$ & &  & $\times$ & $\times$  \\
-  \quad \R{quartiles} & Compute 1st and 3rd quartiles &    &  & $\times$ & &  & $\times$ & $\times$ \\
-  \quad \R{uniqueValue} & Count number of unique values &   $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\
-  \quad \R{variableType} & Data class of variable & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\
+  \textbf{\code{summaryFunction}s}  \smallskip \\
+  \quad \code{centralValue} & Compute median or mode &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\
+  \quad \code{countMissing} & Compute ratio of missing observations &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\
+  \quad \code{minMax} & Find minimum and maximum values &   &  & $\times$ & &  & $\times$ & $\times$  \\
+  \quad \code{quartiles} & Compute 1st and 3rd quartiles &    &  & $\times$ & &  & $\times$ & $\times$ \\
+  \quad \code{uniqueValue} & Count number of unique values &   $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\
+  \quad \code{variableType} & Data class of variable & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\
   \smallskip \\
- \textbf{\R{visualFunction}s} \smallskip \\
-  \quad \R{basicVisual} & Histograms and barplots using base \proglang{R} graphics &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\
-  \quad \R{standardVisual} & Histograms and barplots using ggplot2 &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\
+ \textbf{\code{visualFunction}s} \smallskip \\
+  \quad \code{basicVisual} & Histograms and barplots using base \proglang{R} graphics &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\
+  \quad \code{standardVisual} & Histograms and barplots using ggplot2 &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\
   \smallskip \\
- \textbf{\R{checkFunction}s} \smallskip \\
- \quad \R{identifyCaseIssues} & Identify case issues &  $\times$ & $\times$ & & & & &  \\
- \quad \R{identifyLoners} & Identify levels with $<$ 6 obs. & $\times$ & $\times$ & & & & &  \\
- \quad \R{identifyMissing} & Identify miscoded missing values &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &  \\
- \quad \R{identifyNums} & Identify misclassified numeric or integer variables & $\times$ & $\times$ & & & & &  \\
- \quad \R{identifyOutliers} & Identify outliers &  & & $\times$ & & $\times$ & $\times$ \\
- \quad \R{identifyOutliersTBStyle} & Identify outliers (Turkish Boxplot style) &  & & $\times$ & & $\times$ & $\times$ \\
- \quad \R{identifyWhitespace} & Identify prefixed and suffixed whitespace &  $\times$ & $\times$ & & $\times$ & & &  \\
- \quad \R{isCPR} & Identify Danish CPR numbers & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &$\times$   \\
- \quad \R{isEmpty} & Check if the variable contains only a single value & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\
- \quad \R{isKey} & Check if the variable is a key & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$\smallskip   \\
+ \textbf{\code{checkFunction}s} \smallskip \\
+ \quad \code{identifyCaseIssues} & Identify case issues &  $\times$ & $\times$ & & & & &  \\
+ \quad \code{identifyLoners} & Identify levels with $<$ 6 obs. & $\times$ & $\times$ & & & & &  \\
+ \quad \code{identifyMissing} & Identify miscoded missing values &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &  \\
+ \quad \code{identifyNums} & Identify misclassified numeric or integer variables & $\times$ & $\times$ & & & & &  \\
+ \quad \code{identifyOutliers} & Identify outliers &  & & $\times$ & & $\times$ & $\times$ \\
+ \quad \code{identifyOutliersTBStyle} & Identify outliers (Turkish Boxplot style) &  & & $\times$ & & $\times$ & $\times$ \\
+ \quad \code{identifyWhitespace} & Identify prefixed and suffixed whitespace &  $\times$ & $\times$ & & $\times$ & & &  \\
+ \quad \code{isCPR} & Identify Danish CPR numbers & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &$\times$   \\
+ \quad \code{isEmpty} & Check if the variable contains only a single value & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\
+ \quad \code{isKey} & Check if the variable is a key & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$\smallskip   \\
  \hline
@@
   check functions (used for data checks for each variable) currently implemented in \pkg{dataMaid}. The variable
-  classes C, F, I, L, B, N, and D, refer to \R{character}, \R{factor},
-  \R{integer}, \R{labelled}, \R{logical} (boolean), \R{numeric}, and \R{Date}, respectively.}
+  classes C, F, I, L, B, N, and D, refer to \code{character}, \code{factor},
+  \code{integer}, \code{labelled}, \code{logical} (boolean), \code{numeric}, and \code{Date}, respectively.}
 \label{table.SVCfunctions}
@@
 \code{summaryFunction}s, \code{visualFunction}s and
-\code{checkFunction}s.  By default, \R{clean()} runs selected summary,
+\code{checkFunction}s.  By default, \code{clean()} runs selected summary,
 visualization and check functions on each variable in the dataset, and
@@
 summarize/visualize/check functions, but we can also use the
-\R{allSummaryFunctions()}, \R{allVisualFunctions()}, and
-\R{allCheckFunctions()} functions in \pkg{dataMaid} to print overview
+\code{allSummaryFunctions()}, \code{allVisualFunctions()}, and
+\code{allCheckFunctions()} functions in \pkg{dataMaid} to print overview
 lists in \proglang{R}. For example, the implemented
-\R{summaryFunction}s are:
+\code{summaryFunction}s are:
 
@@
 
-Thus we can see, for example, that for \R{numeric}, \R{integer}, or \R{Date}
+Thus we can see, for example, that for \code{numeric}, \code{integer}, or \code{Date}
 variables, \pkg{dataMaid} provides functions for adding summary information about
@@
 We can control what summaries and checks are applied for each variable type
-through the \R{XXXSummaries} and \R{XXXChecks} arguments, where \R{XXX}
-represents a variable type, e.g., \R{factorChecks} for factors,
-\R{numericChecks} for numeric variables, etc. These arguments accept a
+through the \code{XXXSummaries} and \code{XXXChecks} arguments, where \code{XXX}
+represents a variable type, e.g., \code{factorChecks} for factors,
+\code{numericChecks} for numeric variables, etc. These arguments accept a
 vector of summary- or check function names that should be run for a
 particular variable type. The default values for each of these
-arguments can be obtained through the \R{defaultXXXChecks()}
-and \R{defaultXXXSummaries()} functions, where \R{XXX} again refers to
+arguments can be obtained through the \code{defaultXXXChecks()}
+and \code{defaultXXXSummaries()} functions, where \code{XXX} again refers to
 the variable type.
@@
 We can change the summaries (and similarly the checks) by setting the
-corresponding arguments when calling \R{clean()}. For example, to get
+corresponding arguments when calling \code{clean()}. For example, to get
 only the variable type and the central tendency listed in the summary
@@
 functions for each variable type, we can use the simpler argument
-\R{allSummaries} which overrides the summary functions for all
+\code{allSummaries} which overrides the summary functions for all
 variable types. Thus, the same result as above could be obtained with
@@
 
-Similarly, the checks applied are set with the \R{XXXChecks}
+Similarly, the checks applied are set with the \code{XXXChecks}
 arguments. The default checks being applied to a factor is
@@
 Now, if we only wanted to apply the identify whitespace function for
-factor variables, then we would need to set the \R{factorChecks} accordingly
+factor variables, then we would need to set the \code{factorChecks} accordingly
 
@@
 or we could remove checks for factors altogether by setting the
-corresponding argument to \R{NULL}, in which case factor variables will
+corresponding argument to \code{NULL}, in which case factor variables will
 not be checked for any potential errors:
@@
 
-As with \R{summaryFunction}s, a complete list of available
-\R{checkFunction}s is obtained by calling
-\R{allCheckFunctions()}. Note however, that \R{checkFunction}s have a
-usage beyond the \R{XXXChecks} arguments, namely in the
+As with \code{summaryFunction}s, a complete list of available
+\code{checkFunction}s is obtained by calling
+\code{allCheckFunctions()}. Note however, that \code{checkFunction}s have a
+usage beyond the \code{XXXChecks} arguments, namely in the
 \textit{pre-check} stage. In this stage, it is determined whether or
@@
 steps. The functions used in the pre-check stage should be
-\R{checkFunction}s that are applicable to all variable classes. The
-default pre-checks, the functions \R{isKey()} and \R{isEmpty()}, check
+\code{checkFunction}s that are applicable to all variable classes. The
+default pre-checks, the functions \code{isKey()} and \code{isEmpty()}, check
 whether a variable has unique values for all observations or only a
@@
 
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, preChecks = "isKey")
-\end{Sinput}
-\end{Schunk}
+<<eval=FALSE>>=
+clean(toyData, preChecks = "isKey")
+@
 
@@
 \proglang{R} histograms and barplots, we type
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, allVisuals = "basicVisual")
-\end{Sinput}
-\end{Schunk}
+
+<<eval=FALSE>>=
+clean(toyData, allVisuals = "basicVisual")
+@
 
@@
 %\begin{itemize}
-%\item Introduce the relevant \R{clean}-arguments and the \R{defaultWhateverSummaries} etc.-functions
-%\item Introduce \R{allSummaryFunctions()} etc. and present a table corresponding to the output of this call
+%\item Introduce the relevant \code{clean}-arguments and the \code{defaultWhateverSummaries} etc.-functions
+%\item Introduce \code{allSummaryFunctions()} etc. and present a table corresponding to the output of this call
 %\item Small examples:
@@
 In summary, and as indicated in Figure~\ref{figure:cleanStructure}, there are two stages
-where \R{clean()} applies functions to each of the variables:
+where \code{clean()} applies functions to each of the variables:
 \begin{enumerate}
@@
 Each of these stages are controllable using appropriate function
-arguments in \R{clean()}, and above we have shown examples of how to
+arguments in \code{clean()}, and above we have shown examples of how to
 tweak them to modify the data cleaning outputs. However, if for
@@
 
-\section{Using dataMaid interactively}
+\section[Using dataMaid interactively]{Using \pkg{dataMaid} interactively}
 \label{sec:interactiveCleanR}
@@
 data cleaning checks, it may be useful to be able to work more interactively
-through the data cleaning process as well. Aside from the \R{clean()} function
+through the data cleaning process as well. Aside from the \code{clean()} function
 presented above, \pkg{dataMaid} also provides more standard \proglang{R} interactive tools, such as
@@
 as an object for later use. This section describes how to use the
-functions \R{check()}, \R{summarize()} and \R{visualize()} to work
+functions \code{check()}, \code{summarize()} and \code{visualize()} to work
 interactively with \pkg{dataMaid}.
@@
 Let's say we wish to look further into a certain variable from
-\R{toyData}, namely \R{var2}. The data cleaning summary found some
+\code{toyData}, namely \code{var2}. The data cleaning summary found some
 issues in this variable, and we would like to recall what these issues
-were. This can be done using the \R{check()} command
-\begin{Schunk}
-\begin{Sinput}
-> check(toyData$var2)
-\end{Sinput}
-\begin{Soutput}
-$identifyMissing
-The following suspected missing value codes enter as regular values: 999, NaN.
-$identifyOutliers
-Note that the following possible outlier values were detected: 82, 999.
-\end{Soutput}
-\end{Schunk}
-% $
+were. This can be done using the \code{check()} command
+
+<<>>=
+check(toyData$var2)
+@
 
 Note that the arguments specifying which checks to perform, as
-described in the previous section, are in fact passed to \R{check()},
+described in the previous section, are in fact passed to \code{check()},
 and thus they can also be used here. For instance, if we only want to
 check for potentially miscoded missing values, we can use the relevant
-\R{XXXChecks} argument (e.g., \R{numericChecks}, \R{factorChecks},
+\code{XXXChecks} argument (e.g., \code{numericChecks}, \code{factorChecks},
 etc. as described in Section~\ref{sec:usingdataMaid}) to provide a
 vector of the check functions that should be applied. Recall that
-Table~\ref{table.SVCfunctions} or an \R{allCheckFunctions()} call provide
+Table~\ref{table.SVCfunctions} or an \code{allCheckFunctions()} call provide
 overviews of the available check functions.
@@
 
-\begin{Schunk}
-\begin{Sinput}
-> check(toyData$var2, numericChecks = "identifyMissing")
-\end{Sinput}
-\begin{Soutput}
-$identifyMissing
-The following suspected missing value codes enter as regular values: 999, NaN.
-\end{Soutput}
-\end{Schunk}
-
-An equivalent way to call only a single, specific \R{checkFunction},
-such as \R{identifyMissing}, is by using it directly on the variable, e.g.,
-\begin{Schunk}
-\begin{Sinput}
-> identifyMissing(toyData$var2)
-\end{Sinput}
-\begin{Soutput}
-The following suspected missing value codes enter as regular values: 999, NaN.
-\end{Soutput}
-\end{Schunk}
-% $
-
-The result of a \R{checkFunction} is an object of class
-\R{checkResult}. By using the structure function, \R{str()}, we can
+<<>>=
+check(toyData$var2, numericChecks = "identifyMissing")
+@
+
+An equivalent way to call only a single, specific \code{checkFunction},
+such as \code{identifyMissing}, is by using it directly on the variable,
+e.g.,
+
+<<>>=
+identifyMissing(toyData$var2)
+@
+
+The result of a \code{checkFunction} is an object of class
+\code{checkResult}. By using the structure function, \code{str()}, we can
 look further into its components:
 
-\begin{Schunk}
-\begin{Sinput}
-> missVar2 <- identifyMissing(toyData$var2)
-> str(missVar2)
-\end{Sinput}
-\begin{Soutput}
-List of 3
- $ problem      : logi TRUE
- $ message      : chr "The following suspected missing value codes enter as
- 		 regular values: \\\"999\\\", \\\"NaN\\\"."
- $ problemValues: num [1:2] 999 NaN
- - attr(*, "class")= chr "checkResult"
-\end{Soutput}
-\end{Schunk}
+<<>>=
+missVar2 <- identifyMissing(toyData$var2)
+str(missVar2)
+@
+
 The most important thing to note here is that while the printed
@@
 causing the issue are still obtainable in the element
-\R{problemValues}. If we for instance decide that the values \R{999}
-and \R{NaN} in \R{var2} are in fact miscoded missing values, we can
-easily replace them with \R{NA}s:
-\begin{Schunk}
-\begin{Sinput}
-> toyData$var2[toyData$var2 %in% missVar2$problemValues] <- NA
-> identifyMissing(toyData$var2)
-\end{Sinput}
-\begin{Soutput}
-No problems found.
-\end{Soutput}
-\end{Schunk}
-% $
-
-Similarly, the \R{visualize()} and \R{summarize()} functions can be
+\code{problemValues}. If we for instance decide that the values \code{999}
+and \code{NaN} in \code{var2} are in fact miscoded missing values, we can
+easily replace them with \code{NA}s:
+
+<<>>=
+toyData$var2[toyData$var2 %in% missVar2$problemValues] <- NA
+identifyMissing(toyData$var2)
+@
+
+Similarly, the \code{visualize()} and \code{summarize()} functions can be
 used to run the corresponding visualizations and summaries for each
@@
 
-\begin{Schunk}
-\begin{Sinput}
-> visualize(toyData$var2)
-> summarize(toyData$var2)
-\end{Sinput}
-\begin{Soutput}
-     Feature                   Result
-[1,] "Variable type"           "numeric"
-[2,] "Number of missing obs."  "3 (20 %)"
-[3,] "Number of unique values" "8"
-[4,] "Median"                  "4.5"
-[5,] "1st and 3rd quartiles"   "1.75; 6"
-[6,] "Min. and max."           "1; 999"
-\end{Soutput}
-\end{Schunk}
-% $
+<<toyDataFig, fig=FALSE>>=
+visualize(toyData$var2)
+summarize(toyData$var2)
+@
 
@@
 \label{fig:example2}
-\caption{Output from running \R{visualize()} on the variable \texttt{var2} from the
-\R{toyData} dataset.}
+\caption{Output from running \code{visualize()} on the variable \texttt{var2} from the
+\code{toyData} dataset.}
 \end{figure}
 
-As we saw with the \R{check()} function, the summary can be modified
-by providing the relevant \R{XXXSummaries} arguments. Setting the
-\R{numericSummaries} argument, we can control the summary output by
+As we saw with the \code{check()} function, the summary can be modified
+by providing the relevant \code{XXXSummaries} arguments. Setting the
+\code{numericSummaries} argument, we can control the summary output by
 providing a vector of function names to run for a particular summary. To
 only get the median, minimum and the maximum we set
-\R{numericSummaries=c("centralValue", "minMax")}:
-
-\begin{Schunk}
-\begin{Sinput}
-> summarize(toyData$var2, numericSummaries = c("centralValue", "minMax"))
-\end{Sinput}
-\begin{Soutput}
-     Feature         Result
-[1,] "Median"        "4.5"
-[2,] "Min. and max." "1; 999"
-\end{Soutput}
-\end{Schunk}
-% $
-
-Note that the \R{summarize()}, \R{check()} and \R{visualize()} functions are also available interactively for full datasets, by calling e.g., \R{summarize(toyData)}.
+\code{numericSummaries=c("centralValue", "minMax")}:
+
+<<>>=
+summarize(toyData$var2, numericSummaries = c("centralValue", "minMax"))
+@
+
+Note that the \code{summarize()}, \code{check()} and \code{visualize()} functions are also available interactively for full datasets, by calling e.g., \code{summarize(toyData)}.
 However, this produces an extensive amount of output in the console, and therefore, we generally do not recommend it, unless working with very small datasets or subsets of datasets.
@@
 
-\section{Extending dataMaid}
+\section[Extending dataMaid]{Extending \pkg{dataMaid}}
 \label{sec:extending}
@@
 shown in the previous sections. However, \pkg{dataMaid} is fully
-customizable and \R{clean()} is mainly a tool for formatting the
+customizable and \code{clean()} is mainly a tool for formatting the
 results from various checking-, summary- and visualization
 functions. Thus, the actual work underlying a \pkg{dataMaid} output file
-can be anything --- depending on the arguments given to \R{clean()}
+can be anything --- depending on the arguments given to \code{clean()}
 --- and user made functions are easily added to the
@@
 to specific structures in order to be called from these three steps, and
-therefore, we will now present how \R{summaryFunction}s,
-\R{visualFunction}s and \R{checkFunction}s are made.
-
-This section consists of two parts. First, we describe how customized \R{summaryFunction}s,
-\R{visualFunction}s and \R{checkFunctions} can be made, including an introduction to a
+therefore, we will now present how \code{summaryFunction}s,
+\code{visualFunction}s and \code{checkFunction}s are made.
+
+This section consists of two parts. First, we describe how customized \code{summaryFunction}s,
+\code{visualFunction}s and \code{checkFunctions} can be made, including an introduction to a
 few convenient tools available in \pkg{dataMaid} for aiding SVC-function construction.
@@
 worked example of how to use custom made functions in practice. Here,
-four new SVC functions are defined and used, both interactively and in \R{clean()}.
+four new SVC functions are defined and used, both interactively and in \code{clean()}.
 
@@
 \begin{tabular}{L{0.2\linewidth}L{0.2\linewidth}L{0.2\linewidth}L{0.2\linewidth}}
-& \R{summaryFunction} & \R{visualFunction} & \R{checkFunction} \\
+& \code{summaryFunction} & \code{visualFunction} & \code{checkFunction} \\
 \hline
-\vspace{0pt} Input (required) & \R{v} - a variable vector \newline \R{...}  &  \R{v} - a variable vector \newline \R{vnam} - the variable name (as character string) \newline \R{doEval} - a logical (\R{TRUE}/\R{FALSE}) controlling the output type of the function & \R{v} - a variable vector \newline \R{nMax} - an integer (or \R{Inf}), controlling how many problematic values are printed, if relevant \newline \R{...}  \\
-Input (optional) &  \R{maxDecimals} - number of decimals printed in outputted numerical values.  & - &  \R{maxDecimals}  - number of decimals printed in outputted numerical values.  \\
+\vspace{0pt} Input (required) & \code{v} - a variable vector \newline \code{...}  &  \code{v} - a variable vector \newline \code{vnam} - the variable name (as character string) \newline \code{doEval} - a logical (\code{TRUE}/\code{FALSE}) controlling the output type of the function & \code{v} - a variable vector \newline \code{nMax} - an integer (or \code{Inf}), controlling how many problematic values are printed, if relevant \newline \code{...}  \\
+Input (optional) &  \code{maxDecimals} - number of decimals printed in outputted numerical values.  & - &  \code{maxDecimals}  - number of decimals printed in outputted numerical values.  \\
 Purpose & Describe some aspect of the variable, e.g., a central value, its dispersion or level of missingness.
@@
 \\
-Output (required) & A list with entries: \newline \R{\$feature} - a label for the summary value (as character string); \newline \R{\$result} - the result of the summary (as character string) & A character string with \R{R} code for producing a plot.
+Output (required) & A list with entries: \newline \code{\$feature} - a label for the summary value (as character string); \newline \code{\$result} - the result of the summary (as character string) & A character string with \code{R} code for producing a plot.
 This code should be standalone, i.e., should include the data if necessary.
-& A list with entries: \newline \R{\$problem} - a logical identifying whether an issue was found; \newline \R{\$message} - a character string (possibly empty) decribing the issue that was found, properly escaped and ready for use in \R{rmarkdown} \\
-Output (recommended) & A \R{summaryResult} object, i.e., an attributed list with entries \R{\$feature}, \R{\$result} and \R{\$value}, the latter being the values from \R{\$result} in their original format).
+& A list with entries: \newline \code{\$problem} - a logical identifying whether an issue was found; \newline \code{\$message} - a character string (possibly empty) decribing the issue that was found, properly escaped and ready for use in \code{rmarkdown} \\
+Output (recommended) & A \code{summaryResult} object, i.e., an attributed list with entries \code{\$feature}, \code{\$result} and \code{\$value}, the latter being the values from \code{\$result} in their original format).
 & \bgroup
@@
                  p{0.45\linewidth}} \vspace*{-1cm}\raggedright If
-                 \R{doEval} is \R{TRUE}: \newline A plot that is
-                 opened by the graphic device in \R{R}. &
-                                                          \vspace*{-1cm}\raggedright If \R{doEval} is \R{FALSE}: \newline   A text string with \R{R} code, as described above.
-\end{tabular} \egroup  & A \R{checkResult} object, i.e., an attributed list with entries \R{\$problem}, \R{\$message} and \R{\$problemValues}, the latter being either \R{NULL} or the problem causing values, as they were found in \R{v}, whichever is relevant.
+                 \code{doEval} is \code{TRUE}: \newline A plot that is
+                 opened by the graphic device in \code{R}. &
+                                                          \vspace*{-1cm}\raggedright If \code{doEval} is \code{FALSE}: \newline   A text string with \code{R} code, as described above.
+\end{tabular} \egroup  & A \code{checkResult} object, i.e., an attributed list with entries \code{\$problem}, \code{\$message} and \code{\$problemValues}, the latter being either \code{NULL} or the problem causing values, as they were found in \code{v}, whichever is relevant.
 \\
-Tools available for producing the function & \R{summaryResult()} & --- & \R{messageGenerator()} \newline \R{checkResult()}
+Tools available for producing the function & \code{summaryResult()} & --- & \code{messageGenerator()} \newline \code{checkResult()}
 \end{tabular}
@@
 different levels of strictness. If
-the new custom function is only to be used as part of the SVC steps in \R{clean()},
+the new custom function is only to be used as part of the SVC steps in \code{clean()},
 then only the input/output structure of the function needs special attention.
@@
 \end{Verbatim}
-The last function called here, \R{summaryResult()}, changes the class
-of the output, thereby making a \R{print()} method available for it.
+The last function called here, \code{summaryResult()}, changes the class
+of the output, thereby making a \code{print()} method available for it.
 Note that \code{v} is a vector and that \code{res} should be either a
@@
 Adhering to the template above is sufficient for using the freshly
-made \R{mySummaryFunction()} in \R{clean()}, but we recommend
+made \code{mySummaryFunction()} in \code{clean()}, but we recommend
 furthermore adding the new function to the overview of all summary
 functions by converting it to a proper \code{summaryFunction}
-object. This is done by calling the \R{summaryFunction()} creator with
+object. This is done by calling the \code{summaryFunction()} creator with
 the user-defined function as the first argument, and additional arguments
-\R{description} (an explanatory text which will be added to the attributes of
-the function), and \R{classes} (a vector of variable classes the
+\code{description} (an explanatory text which will be added to the attributes of
+the function), and \code{classes} (a vector of variable classes the
 user-defined function is intended to be applied to, also stored as an
@@
 % One comment should be devoted to
-% the two attributes of a \R{summaryFunction}.
-If the \R{description} argument is left unspecified, the name of the
-function (which in this case is \R{"mySummaryFunction"}) will be filled in and
-stored under the \R{description} attribute. What happens if the \R{classes} argument is
-not specified depends on the type of \R{mySummaryFunction}. If
-\R{mySummaryFunction} is constructed as an \R{S3} generic function with associated
-methods, the call to \R{summaryFunction()} will automatically produce
+% the two attributes of a \code{summaryFunction}.
+If the \code{description} argument is left unspecified, the name of the
+function (which in this case is \code{"mySummaryFunction"}) will be filled in and
+stored under the \code{description} attribute. What happens if the \code{classes} argument is
+not specified depends on the type of \code{mySummaryFunction}. If
+\code{mySummaryFunction} is constructed as an \code{S3} generic function with associated
+methods, the call to \code{summaryFunction()} will automatically produce
 a vector of the names of the classes for which the function can be
-called. If \R{mySummaryFunction} is not an \R{S3} generic and
-\R{classes} is left unspecified, the attribute will simply be
-left empty. Note that the helper function \R{allClasses()} might be useful
-for filling out the \R{classes} argument, as it simply lists all
+called. If \code{mySummaryFunction} is not an \code{S3} generic and
+\code{classes} is left unspecified, the attribute will simply be
+left empty. Note that the helper function \code{allClasses()} might be useful
+for filling out the \code{classes} argument, as it simply lists all
 available classes in \pkg{dataMaid}:
-\begin{Schunk}
-\begin{Sinput}
-> allClasses()
-\end{Sinput}
-\begin{Soutput}
-[1] "character" "Date"      "factor"   "integer"   "labelled"
-[6] "logical"  "numeric"
-\end{Soutput}
-\end{Schunk}
+
+<<>>=
+allClasses()
+@
+
 % \hl{Write something here, don't end paragraph with code. Also, maybe move the allClasses() stuff somewhere else, it doesn't really belong under this header. Not sure where to, though.}
@@
 %til at det kun er beskrevet i pseudokode. (fx. er classes = numeric meget arbitrært)
-%If we run the \R{allSummaryFunctions()} function we can see that the
+%If we run the \code{allSummaryFunctions()} function we can see that the
 %new function has been registered (output slightly appreviated):
@@
 \code{visualFunction}s are the functions that produce the figures in a
-\pkg{dataMaid} output document. Writing a \R{visualFunction} is slightly
-more complicated than writing a \R{summaryFunction}. This follows from
-the fact that \R{visualFunctions} need to be able to output standalone
+\pkg{dataMaid} output document. Writing a \code{visualFunction} is slightly
+more complicated than writing a \code{summaryFunction}. This follows from
+the fact that \code{visualFunctions} need to be able to output standalone
 code for plots in order for \code{clean()} to build standalone
@@
 \code{FALSE}). Implementing the \code{doEval = TRUE} setting is not
-strictly necessary for a \R{visualFunction}'s use in \code{clean()}, but
+strictly necessary for a \code{visualFunction}'s use in \code{clean()}, but
 it makes it easier to assess what visualization options are available,
 and obviously, it is crucial for interactive usage of
-\R{myVisualFunction()}. In either case, it should be noted that all
+\code{myVisualFunction()}. In either case, it should be noted that all
 the parameters listed above, \code{v}, \code{vnam} and \code{doEval},
@@
 
-As with the summary function, we call \R{visualFunction()} to
+As with the summary function, we call \code{visualFunction()} to
 register our newly created function:
@@
 
-Now, \R{myVisualFunction()} will be available in a \R{allVisualFunctions()}
-call, just like the two build-in \R{visualFunction}s, \R{standardVisual}
-and \R{basicVisual}.
+Now, \code{myVisualFunction()} will be available in a \code{allVisualFunctions()}
+call, just like the two build-in \code{visualFunction}s, \code{standardVisual}
+and \code{basicVisual}.
 
@@
 it tries to identify problematic values in the variable (as
-e.g., \R{identifyMissing()} does), or it performs a check concerning the
+e.g., \code{identifyMissing()} does), or it performs a check concerning the
 variable as a whole (e.g., the functions used for prechecks and the
-function \R{identifyNums()}). We present templates for both types of
-\R{checkFunction}s below separately, but it should be emphasized that
+function \code{identifyNums()}). We present templates for both types of
+\code{checkFunction}s below separately, but it should be emphasized that
 formally, they belong to the same class.
@@
 first define the function and subsequently register it as a check
-function using \R{checkFunction()}:
+function using \code{checkFunction()}:
 \begin{Schunk}
@@
 
-Again, as with \R{summaryFunction}s and \R{visualFunction}s, the
-change of function class by use of \R{checkFunction()} is not strictly
-necessary. Note however, that if \R{myFullVarCheckFunction} is to be
-used in the summarize/visualize/check steps in \R{clean()}, the
+Again, as with \code{summaryFunction}s and \code{visualFunction}s, the
+change of function class by use of \code{checkFunction()} is not strictly
+necessary. Note however, that if \code{myFullVarCheckFunction} is to be
+used in the summarize/visualize/check steps in \code{clean()}, the
 description attribute will be printed in the overview table in the
@@
 One comment should be devoted to the helper function,
-\R{messageGenerator()}.  This function's sole purpose is aiding
+\code{messageGenerator()}.  This function's sole purpose is aiding
 consistent styling of all \code{checkFunction} messages. The function
@@
 \code{problemValues}, with the latter being quoted and sorted
-alphabetically. If the \R{nMax} argument to \R{messageGenerator()} is
-not \R{Inf}, only the first \R{nMax} problem values will be pasted
+alphabetically. If the \code{nMax} argument to \code{messageGenerator()} is
+not \code{Inf}, only the first \code{nMax} problem values will be pasted
 onto the message, accompanied by a comment about how many problem
@@
 
-In the template above, the argument \R{maxDecimals} is not in use. This
-argument should be used to round off the \R{problemValues} passed to
-\R{messageGenerator()}, if they are numerical.  This can be done by
-adding an extra line of code after defining \R{problemStatus} in the template
+In the template above, the argument \code{maxDecimals} is not in use. This
+argument should be used to round off the \code{problemValues} passed to
+\code{messageGenerator()}, if they are numerical.  This can be done by
+adding an extra line of code after defining \code{problemStatus} in the template
 above:
@@
 Now, problematic values will be rounded in the outputted message, while they
-will still appear in their original format under the entry \R{\$problemValues} %$
-in the outputted \R{checkResult}.
+will still appear in their original format under the entry \code{\$problemValues} %$
+in the outputted \code{checkResult}.
 
@@
 We will now build four new functions and show both how they can be
-used interactively and how they can be integrated with the \R{clean()}
+used interactively and how they can be integrated with the \code{clean()}
 function. These four new functions are:
 \begin{description}
-\item[\R{isID}] A new \R{checkFunction} intended for use in the precheck-stage.
+\item[\code{isID}] A new \code{checkFunction} intended for use in the precheck-stage.
 This function checks whether a variable consists exclusively of long ($> 10$ characters/digits) entries that are all of equal length, as this might be personal identification codes that we do not wish to print out in the data summary.
-\item[\R{mosaicVisual}] A new \R{visualFunction} that produces mosaic plots. This function will be used in the \textit{visualize} step of \R{clean()}.
-\item[\R{countZeros}] A new \R{summaryFunction} that counts the number of occurrences of the value \R{0} in a variable. This function will be used in the \textit{summarize} step of \R{clean()}.
-\item[\R{identifyColons}] A new \R{checkFunction} that flags variables in which values have colons that appear before and after alphanumerical characters.
+\item[\code{mosaicVisual}] A new \code{visualFunction} that produces mosaic plots. This function will be used in the \textit{visualize} step of \code{clean()}.
+\item[\code{countZeros}] A new \code{summaryFunction} that counts the number of occurrences of the value \code{0} in a variable.
+This function will be used in the \textit{summarize} step of \code{clean()}.
+\item[\code{identifyColons}] A new \code{checkFunction} that flags variables in which values have colons that appear before and after alphanumerical characters.
 This is practical for identifying autogenerated interaction effects.
-This function will be used in the \textit{check} step of \R{clean()}.
+This function will be used in the \textit{check} step of \code{clean()}.
 \end{description}
-These functions are defined in turn below, and afterwards, an example of how they can be called from \R{clean()} is provided.
+These functions are defined in turn below, and afterwards, an example of how they can be called from \code{clean()} is provided.
 
 \subsubsection{isID --- a new checkFunction without problem values}
-First, let's define the \R{isID()} function. As this function is not
+First, let's define the \code{isID()} function. As this function is not
 supposed to list problematic values, it falls within
-the category of \R{checkFunctions} represented by
-\R{myFullVarCheckFunction()} in the above. We do not particularly wish
+the category of \code{checkFunctions} represented by
+\code{myFullVarCheckFunction()} in the above. We do not particularly wish
 to use this function interactively, so we will stick to the minimal
-requirements of a \R{checkFunction} used in \R{check()} (see
+requirements of a \code{checkFunction} used in \code{check()} (see
 Table~\ref{table.functionTypes}). The function can then be defined by
-\begin{Schunk}
-\begin{Sinput}
+
+<<>>=
 isID <- function(v, nMax = NULL, ...) {
@@
 }
-\end{Sinput}
-\end{Schunk}
+@
 
 This is essentially all we need to do in order to include this
-function as a precheck-function in \R{clean()}, so we will leave it as
-is and move on to the next function, namely \R{mosaicVisual()}.
+function as a precheck-function in \code{clean()}, so we will leave it as
+is and move on to the next function, namely \code{mosaicVisual()}.
 
 \subsubsection{mosiacVisual --- a new visualFunction}
-\proglang{R} contains a function, \R{mosiacplot()}, which produces
+\proglang{R} contains a function, \code{mosiacplot()}, which produces
 mosaic plots. We intend to use this existing high-level plotting
@@
 the new function such that it gets the full \pkg{dataMaid}
-functionality. This can be done using the following code.
-\begin{Schunk}
-\begin{Sinput}
+functionality. This can be done using the following code which sets up
+the call using the existing function \code{mosaicplot}.
+
+<<>>=
 mosaicVisual <- function(v, vnam, doEval) {
-   # Setup the call using the existing function mosaicplot
   thisCall <- call("mosaicplot", table(v), main = vnam, xlab = "")
@@
 }
-\end{Sinput}
-\end{Schunk}
-This function can now be called directly or used in \R{clean()}, as will
-presented in an example below. Depending on the \R{doEval} argument, either a text string
+@
+
+This function can now be called directly or used in \code{clean()}, as will
+presented in an example below. Depending on the \code{doEval} argument, either a text string
 with code or a plot is produced. The plot resulting from the following
@@
 
-Even though \R{mosaicVisual()}, as written above, follows the style of a
-\R{visualFunction}, it is not yet truly one and therefore, it will not
-appear in an \R{allVisualFunctions()} call. In order to get this
+Even though \code{mosaicVisual()}, as written above, follows the style of a
+\code{visualFunction}, it is not yet truly one and therefore, it will not
+appear in an \code{allVisualFunctions()} call. In order to get this
 functionality, we need to change its object class. This can be done by
 writing
-\begin{Schunk}
-\begin{Sinput}
+
+<<>>=
 mosaicVisual <- visualFunction(mosaicVisual,
@@
   classes = allClasses())
-\end{Sinput}
-\end{Schunk}
-Here, we use the function \R{allClasses()} to quickly obtain a vector
+@
+
+Here, we use the function \code{allClasses()} to quickly obtain a vector
 of all the seven variable classes addressed in \pkg{dataMaid}. Note that
-if \R{mosaicVisual()} were an S3 generic function, this argument could
-have been left as \R{NULL} and then the classes for which methods are
+if \code{mosaicVisual()} were an S3 generic function, this argument could
+have been left as \code{NULL} and then the classes for which methods are
 available would be added automatically.
@@
 
-As \R{mosaicVisual} is now a full-blooded \R{visualFunction}, it will also be included in the \R{allVisualFunctions()} output table:
-\begin{Schunk}
-\begin{Sinput}
-> allVisualFunctions()
-\end{Sinput}
-\begin{Soutput}
-------------------------------------------------------------------------
-name           description                   classes
--------------- ----------------------------- ---------------------------
-mosaicVisual   Mosaic plots using graphics   character, Date, factor,
-                                             integer, labelled, logical,
-                                             numeric
-
-basicVisual    Histograms and barplots using character, Date, factor,
-               graphics                      integer, labelled, logical,
-                                             numeric
-standardVisual Histograms and barplots using character, Date, factor,
-               ggplot2                       integer, labelled, logical,
-                                             numeric
-------------------------------------------------------------------------
-\end{Soutput}
-\end{Schunk}
-
-Now that we are done with the definition of \R{mosaicVisual()} we can
-turn to the next function in line, \R{countZeros}.
+As \code{mosaicVisual} is now a full-blooded \code{visualFunction}, it
+will also be included in the \code{allVisualFunctions()} output table:
+<<>>=
+allVisualFunctions()
+@
+
+Now that we are done with the definition of \code{mosaicVisual()} we can
+turn to the next function in line, \code{countZeros}.
 
 \subsubsection{countZeros --- a new summaryFunction}
-This \R{summaryFunction} is defined in the following lines of code:
-\begin{Schunk}
-\begin{Sinput}
+This \code{summaryFunction} is defined in the following lines of code:
+
+<<>>=
 countZeros <- function(v, ...) {
@@
 }
-\end{Sinput}
-\end{Schunk}
+@
+
 As this function computes an integer (the number of zeros),
-there is no difference between the entries \R{\$result} and
-\R{\$value}. If, on the other hand, the result had been a character
-string, extra formatting might be required in the \R{\$result} entry
+there is no difference between the entries \code{\$result} and
+\code{\$value}. If, on the other hand, the result had been a character
+string, extra formatting might be required in the \code{\$result} entry
 (such as escaping of quotation marks), and in this scenario, the two
 entries would have differed. As the result is returned as a
-\R{summaryResult} object, a printing method is automatically called
-when \R{countZeros} is used interactively:
-\begin{Schunk}
-\begin{Sinput}
-> countZeros(c(rep(0, 5), 1:100))
-\end{Sinput}
-\begin{Soutput}
-No. zeros: 5
-\end{Soutput}
-\end{Schunk}
-As with \R{mosaicVisual()}, we change the class of this function in
-order to make it appear in \R{allSummaryFunctions()} calls. But now we
+\code{summaryResult} object, a printing method is automatically called
+when \code{countZeros} is used interactively:
+
+<<>>=
+countZeros(c(rep(0, 5), 1:100))
+@
+
+As with \code{mosaicVisual()}, we change the class of this function in
+order to make it appear in \code{allSummaryFunctions()} calls. But now we
 wish to emphasize that the function is not intended to be called on
-all variable types, as zeros have different roles in \R{Date}s and in
-\R{logical} variables:
+all variable types, as zeros have different roles in \code{Date}s and in
+\code{logical} variables:
 \begin{Schunk}
@@
 \subsubsection{identifyColons - a new checkFunction with problem values}
-The last function mentioned above is \R{identifyColons()}. We define
-it using the helper function \R{messageGenerator()} to obtain a properly
-escaped message, and we use \R{checkResult} to make its output print
+The last function mentioned above is \code{identifyColons()}. We define
+it using the helper function \code{messageGenerator()} to obtain a properly
+escaped message, and we use \code{checkResult} to make its output print
 neatly:
@@
 As with the previous two functions, we also change its class. Note,
-however, that for \R{checkFunction}s, the function description will
-appear in the document produced by \R{clean()} (in the \textit{Data
+however, that for \code{checkFunction}s, the function description will
+appear in the document produced by \code{clean()} (in the \textit{Data
   cleaning summary} section of the output), so now this is not only
-done for the sake of the \R{allCheckFunctions()} output.
+done for the sake of the \code{allCheckFunctions()} output.
 
 \subsubsection{Calling the new summarize/visualize/check functions from clean()}
-Now, we are ready to use these new functions in a \R{clean()}
+Now, we are ready to use these new functions in a \code{clean()}
 call. The extended \pkg{dataMaid} output document should have the
@@
 \begin{itemize}
-\item We want to add the new pre-check function, \R{isID}, to the already existing pre-checks.
+\item We want to add the new pre-check function, \code{isID}, to the already existing pre-checks.
 \item We wish to change the plot type for all variables to the new mosaic plot.
-\item We want the new summary function, \R{countZeros}, to be added to the summaries performed on all variable types but \R{Date} and \R{logical}.
-\item We want the new check function, \R{identifyColon}, to be added to the checks performed on \R{character}, \R{factor} and \R{labelled} variables.
+\item We want the new summary function, \code{countZeros}, to be added to the summaries performed on all variable types but \code{Date} and \code{logical}.
+\item We want the new check function, \code{identifyColon}, to be added to the checks performed on \code{character}, \code{factor} and \code{labelled} variables.
 \end{itemize}
@@
 \subsection{Cleaning large datasets}
-If the dataset becomes very large, the standard use of \R{clean()}
+If the dataset becomes very large, the standard use of \code{clean()}
 outlined above might not be ideal. If there is a vast number of
@@
 be beneficial to not include figures in the \pkg{dataMaid} outputs for
-very large datasets. This is controlled via the \R{mode} argument:
+very large datasets. This is controlled via the \code{mode} argument:
 \begin{Schunk}
@@
 less memory heavy standard \proglang{R} figure style instead of the
-\pkg{ggplot2} figures that are the default option in \R{clean()}. This
-can be done by setting the \R{allVisuals = "basicVisual"} argument:
+\pkg{ggplot2} figures that are the default option in \code{clean()}. This
+can be done by setting the \code{allVisuals = "basicVisual"} argument:
 \begin{Schunk}
@@
 Of course, even less heavy plots might be achieved by writing new
-\R{visualFunction}s, using the guidelines from section
+\code{visualFunction}s, using the guidelines from section
 \ref{sec:functionTemplates}. For instance, a future extension of
@@
 is simply splitting the two steps performed
-by \R{clean()}, namely producing the \pkg{rmarkdown} file and rendering it
-afterwards. If the \R{rmarkdown} file is very long, as it will
+by \code{clean()}, namely producing the \pkg{rmarkdown} file and rendering it
+afterwards. If the \code{rmarkdown} file is very long, as it will
 typically be for very large datasets, keeping this file open in memory
@@
 
-\subsection{Using dataMaid for problem flagging}
+\subsection[Using dataMaid for problem flagging]{Using \pkg{dataMaid} for problem flagging}
 If the data is large, but memory issues and computation time are less
@@
 datasets, sometimes a brief overview of the most pressing issues can
-be useful. This can be achieved by using the \R{onlyProblematic}
-argument in \R{clean()}. By specifying \R{onlyProblematic = TRUE},
+be useful. This can be achieved by using the \code{onlyProblematic}
+argument in \code{clean()}. By specifying \code{onlyProblematic = TRUE},
 only variables that raise a flag in the checking steps will be
@@
 getting a quick overview of the problems they might have. This can be
-done by also controlling the \R{mode} argument:
+done by also controlling the \code{mode} argument:
 \begin{Schunk}
@@
 Now only the checking results are printed, and only for variables where problems were identified.
-An even more minimal output can be generated by also leaving out the checking results --- then \R{clean()} essentially just produces a list of the variable names that should be investigated further:
+An even more minimal output can be generated by also leaving out the checking results --- then \code{clean()} essentially just produces a list of the variable names that should be investigated further:
 \begin{Schunk}
@@
 Of course, this can also be done without generating an overview
-document, by direct, interactive use of the \R{check()} function. When
-called on a \R{data.frame}, this function produces a list (of
+document, by direct, interactive use of the \code{check()} function. When
+called on a \code{data.frame}, this function produces a list (of
 variables) of lists (of checks) of lists (or rather,
-\R{checkResult}s). Thus, the overall problem status of each variable
+\code{checkResult}s). Thus, the overall problem status of each variable
 can easily be unravelled using the list manipulation function
-\R{sapply()}:
+\code{sapply()}:
 \begin{Schunk}
@@
 \end{Schunk}
-and we find that only a single variable in \R{toyData}, \R{var6} (for which all
-observations have the value \R{"Irrelevant"}), is problem-free.
-
-\subsection{Including dataMaid documents in other files}
+and we find that only a single variable in \code{toyData}, \code{var6} (for which all
+observations have the value \code{"Irrelevant"}), is problem-free.
+
+\subsection[Including dataMaid documents in other files]{Including \pkg{dataMaid} documents in other files}
 Sometimes, a \pkg{dataMaid} document might be a useful addition to a
@@
 document that can readily be included in other \pkg{rmarkdown}
-files. This is done by using the \R{standAlone} argument in \R{clean()},
+files. This is done by using the \code{standAlone} argument in \code{clean()},
 which removes the preamble from the outputted \pkg{rmarkdown}
 file. Please note, that it is still necessary to indicate which
-\R{rmarkdown} type is being created; the pdf and html \pkg{rmarkdown}
+\code{rmarkdown} type is being created; the pdf and html \pkg{rmarkdown}
 styles are unfortunately not identical.
@@
 rendered to either of these two file types, we recommend setting
-\R{twoCols = FALSE} and \R{output = html} in \R{clean()}, thereby
+\code{twoCols = FALSE} and \code{output = html} in \code{clean()}, thereby
 essentially removing almost all output type specific formatting code
@@
 error. The following is an example of how such a master document
-preamble YAML might look like and how the \R{dataMaid\_toyData.Rmd} file can
+preamble YAML might look like and how the \code{dataMaid\_toyData.Rmd} file can
 then be included:
@@
 
-In the this example, the \R{dataMaid\_toyData.Rmd} file could have been created as follows:
+In the this example, the \code{dataMaid\_toyData.Rmd} file could have been created as follows:
 \begin{Schunk}
@@
 \end{Schunk}
-Note that with the latter option, no special YAML preamble is needed in the \R{rmarkdown} document.
+Note that with the latter option, no special YAML preamble is needed in the \pkg{rmarkdown} document.
 % \hl{don't end on code.}

### 2017-01-24 latex/article_vol2.tex
Bibliography

@@
 \begin{Sinput}
-> clean(toyData, preChecks = "isKey")
+R> clean(toyData, preChecks = "isKey")
 \end{Sinput}
@@
 \proglang{R} histograms and barplots, we type
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, allVisuals = "basicVisual")
+
+\begin{Schunk}
+\begin{Sinput}
+R> clean(toyData, allVisuals = "basicVisual")
 \end{Sinput}
@@
 Let's say we wish to look further into a certain variable from
-\R{toyData}, namely \R{var2}. The data cleaning summary found some
+\code{toyData}, namely \code{var2}. The data cleaning summary found some
 issues in this variable, and we would like to recall what these issues
-were. This can be done using the \R{check()} command
-\begin{Schunk}
-\begin{Sinput}
-> check(toyData$var2)
+were. This can be done using the \code{check()} command
+
+\begin{Schunk}
+\begin{Sinput}
+R> check(toyData$var2)
 \end{Sinput}
@@
 \end{Schunk}
-% $
 
 Note that the arguments specifying which checks to perform, as
-described in the previous section, are in fact passed to \R{check()},
+described in the previous section, are in fact passed to \code{check()},
 and thus they can also be used here. For instance, if we only want to
 check for potentially miscoded missing values, we can use the relevant
-\R{XXXChecks} argument (e.g., \R{numericChecks}, \R{factorChecks},
+\code{XXXChecks} argument (e.g., \code{numericChecks}, \code{factorChecks},
 etc. as described in Section~\ref{sec:usingdataMaid}) to provide a
 vector of the check functions that should be applied. Recall that
-Table~\ref{table.SVCfunctions} or an \R{allCheckFunctions()} call provide
+Table~\ref{table.SVCfunctions} or an \code{allCheckFunctions()} call provide
 overviews of the available check functions.
@@
 \begin{Sinput}
-> check(toyData$var2, numericChecks = "identifyMissing")
+R> check(toyData$var2, numericChecks = "identifyMissing")
 \end{Sinput}
@@
 
-An equivalent way to call only a single, specific \R{checkFunction},
-such as \R{identifyMissing}, is by using it directly on the variable, e.g.,
-\begin{Schunk}
-\begin{Sinput}
-> identifyMissing(toyData$var2)
+An equivalent way to call only a single, specific \code{checkFunction},
+such as \code{identifyMissing}, is by using it directly on the variable,
+e.g.,
+
+\begin{Schunk}
+\begin{Sinput}
+R> identifyMissing(toyData$var2)
 \end{Sinput}
@@
 \end{Schunk}
-% $
-
-The result of a \R{checkFunction} is an object of class
-\R{checkResult}. By using the structure function, \R{str()}, we can
+
+The result of a \code{checkFunction} is an object of class
+\code{checkResult}. By using the structure function, \code{str()}, we can
 look further into its components:
@@
 \begin{Sinput}
-> missVar2 <- identifyMissing(toyData$var2)
-> str(missVar2)
+R> missVar2 <- identifyMissing(toyData$var2)
+R> str(missVar2)
 \end{Sinput}
@@
  $ problem      : logi TRUE
- $ message      : chr "The following suspected missing value codes enter as
- 		 regular values: \\\"999\\\", \\\"NaN\\\"."
- $ problemValues: num [1:2] 999 NaN
+ $ message      : chr "The following suspected missing value codes enter as regular values: \\\"999\\\", \\\"NaN\\\"."
+ $ problemValues: NULL
  - attr(*, "class")= chr "checkResult"
@@
 \end{Schunk}
+
 The most important thing to note here is that while the printed
@@
 causing the issue are still obtainable in the element
-\R{problemValues}. If we for instance decide that the values \R{999}
-and \R{NaN} in \R{var2} are in fact miscoded missing values, we can
-easily replace them with \R{NA}s:
-\begin{Schunk}
-\begin{Sinput}
-> toyData$var2[toyData$var2 %in% missVar2$problemValues] <- NA
-> identifyMissing(toyData$var2)
+\code{problemValues}. If we for instance decide that the values \code{999}
+and \code{NaN} in \code{var2} are in fact miscoded missing values, we can
+easily replace them with \code{NA}s:
+
+\begin{Schunk}
+\begin{Sinput}
+R> toyData$var2[toyData$var2 %in% missVar2$problemValues] <- NA
+R> identifyMissing(toyData$var2)
 \end{Sinput}
 \begin{Soutput}
-No problems found.
+The following suspected missing value codes enter as regular values: 999, NaN.
 \end{Soutput}
 \end{Schunk}
-% $
-
-Similarly, the \R{visualize()} and \R{summarize()} functions can be
+
+Similarly, the \code{visualize()} and \code{summarize()} functions can be
 used to run the corresponding visualizations and summaries for each
@@
 \begin{Sinput}
-> visualize(toyData$var2)
-> summarize(toyData$var2)
+R> visualize(toyData$var2)
+R> summarize(toyData$var2)
 \end{Sinput}
@@
 \end{Schunk}
-% $
 
@@
 \label{fig:example2}
-\caption{Output from running \R{visualize()} on the variable \texttt{var2} from the
-\R{toyData} dataset.}
+\caption{Output from running \code{visualize()} on the variable \texttt{var2} from the
+\code{toyData} dataset.}
 \end{figure}
 
-As we saw with the \R{check()} function, the summary can be modified
-by providing the relevant \R{XXXSummaries} arguments. Setting the
-\R{numericSummaries} argument, we can control the summary output by
+As we saw with the \code{check()} function, the summary can be modified
+by providing the relevant \code{XXXSummaries} arguments. Setting the
+\code{numericSummaries} argument, we can control the summary output by
 providing a vector of function names to run for a particular summary. To
 only get the median, minimum and the maximum we set
-\R{numericSummaries=c("centralValue", "minMax")}:
-
-\begin{Schunk}
-\begin{Sinput}
-> summarize(toyData$var2, numericSummaries = c("centralValue", "minMax"))
+\code{numericSummaries=c("centralValue", "minMax")}:
+
+\begin{Schunk}
+\begin{Sinput}
+R> summarize(toyData$var2, numericSummaries = c("centralValue", "minMax"))
 \end{Sinput}
@@
 \end{Schunk}
-% $
-
-Note that the \R{summarize()}, \R{check()} and \R{visualize()} functions are also available interactively for full datasets, by calling e.g., \R{summarize(toyData)}.
+
+Note that the \code{summarize()}, \code{check()} and \code{visualize()} functions are also available interactively for full datasets, by calling e.g., \code{summarize(toyData)}.
 However, this produces an extensive amount of output in the console, and therefore, we generally do not recommend it, unless working with very small datasets or subsets of datasets.
@@
 % One comment should be devoted to
-% the two attributes of a \R{summaryFunction}.
-If the \R{description} argument is left unspecified, the name of the
-function (which in this case is \R{"mySummaryFunction"}) will be filled in and
-stored under the \R{description} attribute. What happens if the \R{classes} argument is
-not specified depends on the type of \R{mySummaryFunction}. If
-\R{mySummaryFunction} is constructed as an \R{S3} generic function with associated
-methods, the call to \R{summaryFunction()} will automatically produce
+% the two attributes of a \code{summaryFunction}.
+If the \code{description} argument is left unspecified, the name of the
+function (which in this case is \code{"mySummaryFunction"}) will be filled in and
+stored under the \code{description} attribute. What happens if the \code{classes} argument is
+not specified depends on the type of \code{mySummaryFunction}. If
+\code{mySummaryFunction} is constructed as an \code{S3} generic function with associated
+methods, the call to \code{summaryFunction()} will automatically produce
 a vector of the names of the classes for which the function can be
-called. If \R{mySummaryFunction} is not an \R{S3} generic and
-\R{classes} is left unspecified, the attribute will simply be
-left empty. Note that the helper function \R{allClasses()} might be useful
-for filling out the \R{classes} argument, as it simply lists all
+called. If \code{mySummaryFunction} is not an \code{S3} generic and
+\code{classes} is left unspecified, the attribute will simply be
+left empty. Note that the helper function \code{allClasses()} might be useful
+for filling out the \code{classes} argument, as it simply lists all
 available classes in \pkg{dataMaid}:
-\begin{Schunk}
-\begin{Sinput}
-> allClasses()
+
+\begin{Schunk}
+\begin{Sinput}
+R> allClasses()
 \end{Sinput}
 \begin{Soutput}
-[1] "character" "Date"      "factor"   "integer"   "labelled"
-[6] "logical"  "numeric"
+[1] "character" "Date"      "factor"    "integer"   "labelled"  "logical"  
+[7] "numeric"  
 \end{Soutput}
 \end{Schunk}
+
 % \hl{Write something here, don't end paragraph with code. Also, maybe move the allClasses() stuff somewhere else, it doesn't really belong under this header. Not sure where to, though.}
@@
 We will now build four new functions and show both how they can be
-used interactively and how they can be integrated with the \R{clean()}
+used interactively and how they can be integrated with the \code{clean()}
 function. These four new functions are:
 \begin{description}
-\item[\R{isID}] A new \R{checkFunction} intended for use in the precheck-stage.
+\item[\code{isID}] A new \code{checkFunction} intended for use in the precheck-stage.
 This function checks whether a variable consists exclusively of long ($> 10$ characters/digits) entries that are all of equal length, as this might be personal identification codes that we do not wish to print out in the data summary.
-\item[\R{mosaicVisual}] A new \R{visualFunction} that produces mosaic plots. This function will be used in the \textit{visualize} step of \R{clean()}.
-\item[\R{countZeros}] A new \R{summaryFunction} that counts the number of occurrences of the value \R{0} in a variable. This function will be used in the \textit{summarize} step of \R{clean()}.
-\item[\R{identifyColons}] A new \R{checkFunction} that flags variables in which values have colons that appear before and after alphanumerical characters.
+\item[\code{mosaicVisual}] A new \code{visualFunction} that produces mosaic plots. This function will be used in the \textit{visualize} step of \code{clean()}.
+\item[\code{countZeros}] A new \code{summaryFunction} that counts the number of occurrences of the value \code{0} in a variable.
+This function will be used in the \textit{summarize} step of \code{clean()}.
+\item[\code{identifyColons}] A new \code{checkFunction} that flags variables in which values have colons that appear before and after alphanumerical characters.
 This is practical for identifying autogenerated interaction effects.
-This function will be used in the \textit{check} step of \R{clean()}.
+This function will be used in the \textit{check} step of \code{clean()}.
 \end{description}
-These functions are defined in turn below, and afterwards, an example of how they can be called from \R{clean()} is provided.
+These functions are defined in turn below, and afterwards, an example of how they can be called from \code{clean()} is provided.
 
 \subsubsection{isID --- a new checkFunction without problem values}
-First, let's define the \R{isID()} function. As this function is not
+First, let's define the \code{isID()} function. As this function is not
 supposed to list problematic values, it falls within
-the category of \R{checkFunctions} represented by
-\R{myFullVarCheckFunction()} in the above. We do not particularly wish
+the category of \code{checkFunctions} represented by
+\code{myFullVarCheckFunction()} in the above. We do not particularly wish
 to use this function interactively, so we will stick to the minimal
-requirements of a \R{checkFunction} used in \R{check()} (see
+requirements of a \code{checkFunction} used in \code{check()} (see
 Table~\ref{table.functionTypes}). The function can then be defined by
-\begin{Schunk}
-\begin{Sinput}
-isID <- function(v, nMax = NULL, ...) {
-  out <- list(problem = FALSE, message = "")
-  # Should work for all but logical and Date variables
-  if (class(v) %in% setdiff(allClasses(), c("logical", "Date"))) {
-    v <- as.character(v)
-    lengths <- c(nchar(v))
-    # Check that all lengths are the same and greater than 10 characters
-    if (all(lengths > 10) & length(unique(lengths)) == 1) {
-      out$problem <- TRUE
-      out$message <- "Warning: This variable seems to contain ID codes."
-    }
-  }
-  out
-}
+
+\begin{Schunk}
+\begin{Sinput}
+R> isID <- function(v, nMax = NULL, ...) {
++    out <- list(problem = FALSE, message = "")
++    # Should work for all but logical and Date variables
++    if (class(v) %in% setdiff(allClasses(), c("logical", "Date"))) {
++      v <- as.character(v)
++      lengths <- c(nchar(v))
++      # Check that all lengths are the same and greater than 10 characters
++      if (all(lengths > 10) & length(unique(lengths)) == 1) {
++        out$problem <- TRUE
++        out$message <- "Warning: This variable seems to contain ID codes."
++      }
++    }
++    out
++  }
 \end{Sinput}
@@
 This is essentially all we need to do in order to include this
-function as a precheck-function in \R{clean()}, so we will leave it as
-is and move on to the next function, namely \R{mosaicVisual()}.
+function as a precheck-function in \code{clean()}, so we will leave it as
+is and move on to the next function, namely \code{mosaicVisual()}.
 
 \subsubsection{mosiacVisual --- a new visualFunction}
-\proglang{R} contains a function, \R{mosiacplot()}, which produces
+\proglang{R} contains a function, \code{mosiacplot()}, which produces
 mosaic plots. We intend to use this existing high-level plotting
@@
 the new function such that it gets the full \pkg{dataMaid}
-functionality. This can be done using the following code.
-\begin{Schunk}
-\begin{Sinput}
- mosaicVisual <- function(v, vnam, doEval) {
-   # Setup the call using the existing function mosaicplot
-   thisCall <- call("mosaicplot", table(v), main = vnam, xlab = "")
-   if (doEval) {                     # Return the graphics
-    return(eval(thisCall))
-   } else return(deparse(thisCall))  # Else return the code for the call
- }
-\end{Sinput}
-\end{Schunk}
-This function can now be called directly or used in \R{clean()}, as will
-presented in an example below. Depending on the \R{doEval} argument, either a text string
+functionality. This can be done using the following code which sets up
+the call using the existing function \code{mosaicplot}.
+
+\begin{Schunk}
+\begin{Sinput}
+R> mosaicVisual <- function(v, vnam, doEval) {
++    thisCall <- call("mosaicplot", table(v), main = vnam, xlab = "")
++    if (doEval) {                     # Return the graphics
++      return(eval(thisCall))
++    } else return(deparse(thisCall))  # Else return the code for the call
++  }
+\end{Sinput}
+\end{Schunk}
+
+This function can now be called directly or used in \code{clean()}, as will
+presented in an example below. Depending on the \code{doEval} argument, either a text string
 with code or a plot is produced. The plot resulting from the following
@@
 
-Even though \R{mosaicVisual()}, as written above, follows the style of a
-\R{visualFunction}, it is not yet truly one and therefore, it will not
-appear in an \R{allVisualFunctions()} call. In order to get this
+Even though \code{mosaicVisual()}, as written above, follows the style of a
+\code{visualFunction}, it is not yet truly one and therefore, it will not
+appear in an \code{allVisualFunctions()} call. In order to get this
 functionality, we need to change its object class. This can be done by
 writing
-\begin{Schunk}
-\begin{Sinput}
- mosaicVisual <- visualFunction(mosaicVisual,
- 			description = "Mosaic plots using graphics",
-                                classes = allClasses())
-\end{Sinput}
-\end{Schunk}
-Here, we use the function \R{allClasses()} to quickly obtain a vector
+
+\begin{Schunk}
+\begin{Sinput}
+R> mosaicVisual <- visualFunction(mosaicVisual,
++    description = "Mosaic plots using graphics",
++    classes = allClasses())
+\end{Sinput}
+\end{Schunk}
+
+Here, we use the function \code{allClasses()} to quickly obtain a vector
 of all the seven variable classes addressed in \pkg{dataMaid}. Note that
-if \R{mosaicVisual()} were an S3 generic function, this argument could
-have been left as \R{NULL} and then the classes for which methods are
+if \code{mosaicVisual()} were an S3 generic function, this argument could
+have been left as \code{NULL} and then the classes for which methods are
 available would be added automatically.
@@
 
-As \R{mosaicVisual} is now a full-blooded \R{visualFunction}, it will also be included in the \R{allVisualFunctions()} output table:
-\begin{Schunk}
-\begin{Sinput}
-> allVisualFunctions()
+As \code{mosaicVisual} is now a full-blooded \code{visualFunction}, it
+will also be included in the \code{allVisualFunctions()} output table:
+\begin{Schunk}
+\begin{Sinput}
+R> allVisualFunctions()
 \end{Sinput}
 \begin{Soutput}
-------------------------------------------------------------------------
+------------------------------------------------------------------------------
 name             description                     classes                      
--------------- ----------------------------- ---------------------------
+---------------- ------------------------------- -----------------------------
 mosaicVisual     Mosaic plots using graphics     character, Date, factor,     
@@
                                                  numeric
+
 standardVisual   Histograms and barplots using   character, Date, factor,     
@@
                                                  numeric                      
-------------------------------------------------------------------------
+------------------------------------------------------------------------------
 \end{Soutput}
@@
 
-Now that we are done with the definition of \R{mosaicVisual()} we can
-turn to the next function in line, \R{countZeros}.
+Now that we are done with the definition of \code{mosaicVisual()} we can
+turn to the next function in line, \code{countZeros}.
 
 \subsubsection{countZeros --- a new summaryFunction}
-This \R{summaryFunction} is defined in the following lines of code:
-\begin{Schunk}
-\begin{Sinput}
-countZeros <- function(v, ...) {
-	res <- length(which(v == 0))
-	summaryResult(list(feature = "No. zeros", result = res, value = res))
-}
-\end{Sinput}
-\end{Schunk}
+This \code{summaryFunction} is defined in the following lines of code:
+
+\begin{Schunk}
+\begin{Sinput}
+R> countZeros <- function(v, ...) {
++    res <- length(which(v == 0))
++    summaryResult(list(feature = "No. zeros", result = res, value = res))
++  }
+\end{Sinput}
+\end{Schunk}
+
 As this function computes an integer (the number of zeros),
-there is no difference between the entries \R{\$result} and
-\R{\$value}. If, on the other hand, the result had been a character
-string, extra formatting might be required in the \R{\$result} entry
+there is no difference between the entries \code{\$result} and
+\code{\$value}. If, on the other hand, the result had been a character
+string, extra formatting might be required in the \code{\$result} entry
 (such as escaping of quotation marks), and in this scenario, the two
 entries would have differed. As the result is returned as a
-\R{summaryResult} object, a printing method is automatically called
-when \R{countZeros} is used interactively:
-\begin{Schunk}
-\begin{Sinput}
-> countZeros(c(rep(0, 5), 1:100))
+\code{summaryResult} object, a printing method is automatically called
+when \code{countZeros} is used interactively:
+
+\begin{Schunk}
+\begin{Sinput}
+R> countZeros(c(rep(0, 5), 1:100))
 \end{Sinput}
@@
 \end{Schunk}
-As with \R{mosaicVisual()}, we change the class of this function in
-order to make it appear in \R{allSummaryFunctions()} calls. But now we
+
+As with \code{mosaicVisual()}, we change the class of this function in
+order to make it appear in \code{allSummaryFunctions()} calls. But now we
 wish to emphasize that the function is not intended to be called on
-all variable types, as zeros have different roles in \R{Date}s and in
-\R{logical} variables:
+all variable types, as zeros have different roles in \code{Date}s and in
+\code{logical} variables:
 \begin{Schunk}

### 2017-01-24 latex/article_vol2.Rnw
Chunks

@@
 <<echo=FALSE>>=
-options(prompt = "R> ", width=77, continue = "+  ")
+# options(prompt = "R> ", width=77, continue = "+  ")
+options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
 @
@@
 Online tools such as OpenRefine (\url{http://openrefine.org/}) and
-\proglang{R}-packages such as \pkg{plyr} \citep{plyr}, and \pkg{data.table} have
+\proglang{R}-packages such as \pkg{plyr} \citep{plyr}, and
+\pkg{data.table} \citep{data.table} have
 made data wrangling a lot easier, but only a handful of packages such
-as \pkg{editrules}, \pkg{validate}, \pkg{DataCombine}, and
-\pkg{janitor} attempt to implement systematic, reproducible data
+as \pkg{editrules} \citep{editrules}, \pkg{validate} \citep{validate},
+\pkg{DataCombine} \citep{DataCombine}, and
+\pkg{janitor} \citep{janitor} attempt to implement systematic, reproducible data
 cleaning.  These packages use different approaches for data cleaning:
@@
 and less so for the general researcher with a knowledge about a specific
-field and the context of the available data. The \pkg{dataMaid} package
+field and the context of the available data. The \pkg{dataMaid}
+package \citep{dataMaid}
 tries to address this by providing a framework that both allows for
@@
 provides easy-to-use, automated tools for data quality assessment in
-\proglang{R} on which data cleaning decisions can be made. This quality
+\proglang{R} \citep{R} on which data cleaning decisions can be made. This quality
 assessment is presented in an auto-generated overview document,
@@
 \code{TRUE}. \code{render} determines if the \proglang{R} markdown file
-produced should be rendered using the \pkg{rmarkdown} package and
+produced should be rendered using the \pkg{rmarkdown} \citep{rmarkdown} package and
 \code{openResult} decides whether the output html or pdf file should be
@@
 variable types. If for instance, we wish to change the visualizations
-from the default \pkg{ggplot2} style histograms and barplots to base
+from the default \pkg{ggplot2} \citep{ggplot2} style histograms and barplots to base
 \proglang{R} histograms and barplots, we type
@@
 \\
-Output (required) & A list with entries: \newline \code{\$feature} - a label for the summary value (as character string); \newline \code{\$result} - the result of the summary (as character string) & A character string with \code{R} code for producing a plot.
+Output (required) & A list with entries: \newline \code{\$feature} - a label for the summary value (as character string); \newline \code{\$result} - the result of the summary (as character string) & A character string with \proglang{R} code for producing a plot.
 This code should be standalone, i.e., should include the data if necessary.
@@
                  \code{doEval} is \code{TRUE}: \newline A plot that is
-                 opened by the graphic device in \code{R}. &
-                                                          \vspace*{-1cm}\raggedright If \code{doEval} is \code{FALSE}: \newline   A text string with \code{R} code, as described above.
+                 opened by the graphic device in \proglang{R}. &
+                                                          \vspace*{-1cm}\raggedright If \code{doEval} is \code{FALSE}: \newline   A text string with \proglang{R} code, as described above.
 \end{tabular} \egroup  & A \code{checkResult} object, i.e., an attributed list with entries \code{\$problem}, \code{\$message} and \code{\$problemValues}, the latter being either \code{NULL} or the problem causing values, as they were found in \code{v}, whichever is relevant.
@@
 call is found in Figure~\ref{fig:mosaicPlot}:
-\begin{Schunk}
-\begin{Sinput}
-> mosaicVisual(toyData$var1, "variable 1", doEval = TRUE)
-\end{Sinput}
-\end{Schunk}
-% $
+
+<<fig=FALSE>>=
+mosaicVisual(toyData$var1, "variable 1", doEval = TRUE)
+@
 
@@
 \code{logical} variables:
-\begin{Schunk}
-\begin{Sinput}
-> countZeros <- summaryFunction(countZeros,
+
+XXX indentation
+<<>>=
+countZeros <- summaryFunction(countZeros,
   description = "Count number of zeros",
@@
   "labelled", "numeric"))
-\end{Sinput}
-\end{Schunk}
-% \hl{more? don't end on code.}
-
-\subsubsection{identifyColons - a new checkFunction with problem values}
+@
+
+\subsubsection{identifyColons --- a new checkFunction with problem values}
 The last function mentioned above is \code{identifyColons()}. We define
@@
 escaped message, and we use \code{checkResult} to make its output print
-neatly:
-\begin{Schunk}
-\begin{Sinput}
+neatly
+
+XXX comments
+
+<<>>=
 identifyColons <- function(v, nMax = Inf, ... ) {
@@
     classes = c("character", "factor", "labelled"))
-\end{Sinput}
-\end{Schunk}
+@
+
 As with the previous two functions, we also change its class. Note,
@@
 \end{itemize}
-These options are specified as follows: %\hl{Jeg synes ikke testData egner sig her.
-Der er ingen koloner og CPRvar er et mærkeligt, internt navn, hvis ikke man kender danske CPR-numre.
-Langt de fleste variable har desuden ingen nuller.
-Måske skal der laves et nyt datasæt?}
-\begin{Schunk}
-\begin{Sinput}
-> data("exampleData")
-> clean(exampleData,
-      # Add the new pre-check function
+These options are specified as follows: %\hl{Jeg synes ikke testData
+                                %egner sig her. Der er ingen koloner
+                                %og CPRvar er et mærkeligt, internt
+                                %navn, hvis ikke man kender danske
+                                %CPR-numre. Langt de fleste variable
+                                %har desuden ingen nuller. Måske skal
+                                %der laves et nyt datasæt?}
+<<eval=FALSE>>=
+data("exampleData")
+clean(exampleData,
   preChecks = c("isKey", "isEmpty", "isID"),
-      # Change the visual
   allVisuals = "mosaicVisual",
-      # Add the new summary for specified data types
   characterSummaries = c(defaultCharacterSummaries(), "countZeros"),
@@
   integerSummaries = c(defaultIntegerSummaries(), "countZeros"),
-      # Add the new check for specific data types
   characterChecks = c(defaultCharacterChecks(), "identifyColons"),
@@
   labelledCheck = c(defaultLabelledChecks(), "identifyColons"))
-\end{Sinput}
-\end{Schunk}
+@
+
 The outputted document is found in Appendix \ref{sec:appendix2}.
@@
 very large datasets. This is controlled via the \code{mode} argument:
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, mode = c("summarize", "check"))
-\end{Sinput}
-\end{Schunk}
+
+<<eval=FALSE>>=
+clean(toyData, mode = c("summarize", "check"))
+@
+
 If figures are indeed needed, a different approach is to choose the
@@
 can be done by setting the \code{allVisuals = "basicVisual"} argument:
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, allVisuals = "basicVisual")
-\end{Sinput}
-\end{Schunk}
+
+<<eval=FALSE>>=
+clean(toyData, allVisuals = "basicVisual")
+@
+
 Of course, even less heavy plots might be achieved by writing new
@@
 \pkg{dataMaid} might be the inclusion of ASCII plots, as
-e.g., represented in the \proglang{R} package \pkg{txtplot}.
+e.g., represented in the \proglang{R} package \pkg{txtplot} \citep{txtplot}.
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, render = FALSE, openResult = FALSE)
-> render("dataMaid_toyData.Rmd", quiet = FALSE)
-\end{Sinput}
-\end{Schunk}
+<<eval=FALSE>>=
+clean(toyData, render = FALSE, openResult = FALSE)
+render("dataMaid_toyData.Rmd", quiet = FALSE)
+@
+
 This also deals with the fact that \pkg{dataMaid} can produce
@@
 done by also controlling the \code{mode} argument:
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, onlyProblematic = TRUE, mode = c("check"))
-\end{Sinput}
-\end{Schunk}
-Now only the checking results are printed, and only for variables where problems were identified.
-An even more minimal output can be generated by also leaving out the checking results --- then \code{clean()} essentially just produces a list of the variable names that should be investigated further:
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, onlyProblematic = TRUE, mode = NULL)
-\end{Sinput}
-\end{Schunk}
+
+<<eval=FALSE>>=
+clean(toyData, onlyProblematic = TRUE, mode = c("check"))
+@
+
+Now only the checking results are printed, and only for variables
+where problems were identified. An even more minimal output can be
+generated by also leaving out the checking results --- then
+\code{clean()} essentially just produces a list of the variable names
+that should be investigated further:
+
+<<eval=FALSE>>=
+clean(toyData, onlyProblematic = TRUE, mode = NULL)
+@
+
 Of course, this can also be done without generating an overview
@@
 \code{sapply()}:
-\begin{Schunk}
-\begin{Sinput}
-> toyChecks <- check(toyData)
-> foo <- function(x) {
->   any(sapply(x, function(y) y[["problem"]]))
-> }
-> sapply(toyChecks, foo)
-\end{Sinput}
-\begin{Soutput}
- var1  var2  var3  var4  var5  var6
- TRUE  TRUE  TRUE  TRUE  TRUE FALSE
-\end{Soutput}
-\end{Schunk}
+
+<<>>=
+toyChecks <- check(toyData)
+foo <- function(x) {
+  any(sapply(x, function(y) y[["problem"]]))
+}
+sapply(toyChecks, foo)
+@
+
 and we find that only a single variable in \code{toyData}, \code{var6} (for which all
@@
 
-In the this example, the \code{dataMaid\_toyData.Rmd} file could have been created as follows:
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, standAlone = FALSE)
-\end{Sinput}
-\end{Schunk}
+In the this example, the \code{dataMaid\_toyData.Rmd} file could have
+been created as follows:
+
+<<eval=FALSE>>=
+clean(toyData, standAlone = FALSE)
+@
+
 and the more minimal, html-style \pkg{rmarkdown} file described above can be produced using
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, standAlone = FALSE, output = "html", twoCols = FALSE)
-\end{Sinput}
-\end{Schunk}
+
+<<eval=FALSE>>=
+clean(toyData, standAlone = FALSE, output = "html", twoCols = FALSE)
+@
+
 Note that with the latter option, no special YAML preamble is needed in the \pkg{rmarkdown} document.
@@
 might be improved by visualizing measurements over time. In addition,
-an online \pkg{shiny} application where users that are not
+an online \pkg{shiny} \citep{shiny} application where users that are not
 \proglang{R}-savvy can upload their data and get a data cleaning
@@
 
-\nocite{R}
-\nocite{shiny}
-\nocite{dataMaid}
-\nocite{rmarkdown}
-\nocite{ggplot2}
-\nocite{plyr}
-\nocite{data.table}
-\nocite{validate}
-\nocite{editrules}
-\nocite{janitor}
-\nocite{DataCombine}
-\nocite{txtplot}
+% \nocite{R}
+%\nocite{shiny}
+%\nocite{dataMaid}
+%\nocite{rmarkdown}
+%\nocite{ggplot2}
+%\nocite{plyr}
+%\nocite{data.table}
+%\nocite{validate}
+%\nocite{editrules}
+%\nocite{janitor}
+%\nocite{DataCombine}
+%\nocite{txtplot}
 

### 2017-01-24 latex/article_vol2.tex
Chunks

@@
 \begin{Soutput}
-[1] "character" "Date"      "factor"    "integer"   "labelled"  "logical"  
-[7] "numeric"  
+[1] "character" "Date"      "factor"    "integer"   "labelled" 
+[6] "logical"   "numeric"  
 \end{Soutput}
@@
 call is found in Figure~\ref{fig:mosaicPlot}:
-\begin{Schunk}
-\begin{Sinput}
-> mosaicVisual(toyData$var1, "variable 1", doEval = TRUE)
-\end{Sinput}
-\end{Schunk}
-% $
+
+\begin{Schunk}
+\begin{Sinput}
+R> mosaicVisual(toyData$var1, "variable 1", doEval = TRUE)
+\end{Sinput}
+\end{Schunk}
 
@@
 \code{logical} variables:
-\begin{Schunk}
-\begin{Sinput}
-> countZeros <- summaryFunction(countZeros,
-	description = "Count number of zeros",
-	classes = c("character", "factor", "integer",
-                    "labelled", "numeric"))
-\end{Sinput}
-\end{Schunk}
-% \hl{more? don't end on code.}
-
-\subsubsection{identifyColons - a new checkFunction with problem values}
+
+XXX indentation
+\begin{Schunk}
+\begin{Sinput}
+R> countZeros <- summaryFunction(countZeros,
++    description = "Count number of zeros",
++    classes = c("character", "factor", "integer",
++    "labelled", "numeric"))
+\end{Sinput}
+\end{Schunk}
+
+\subsubsection{identifyColons --- a new checkFunction with problem values}
 The last function mentioned above is \code{identifyColons()}. We define
@@
 escaped message, and we use \code{checkResult} to make its output print
-neatly:
-\begin{Schunk}
-\begin{Sinput}
-identifyColons <- function(v, nMax = Inf, ... ) {
-  v <- unique(na.omit(v))
-  problemMessage <- "Note: The following values include colons:"
-  problem <- FALSE
-  problemValues <- NULL
-
-  # Use regular expressions to identify colons between two words
-  problemValues <- v[sapply(gregexpr("[[:xdigit:]]:[[:xdigit:]]", v),
-                            function(x) all(x != -1))]
-
-  if (length(problemValues) > 0) {
-    problem <- TRUE
-  }
-
-  problemStatus <- list(problem = problem,
-                        problemValues = problemValues)
-  # Use the messagegenerator function to produce the output
-  # message from the problemStatus and problemMessage
-  outMessage <- messageGenerator(problemStatus, problemMessage, nMax)
-
-  checkResult(list(problem = problem,
-                   message = outMessage,
-                   problemValues = problemValues))
-}
-
-identifyColons <- checkFunction(identifyColons,
-    description = "Identify colons surrounded by alphanumeric characters",
-    classes = c("character", "factor", "labelled"))
-\end{Sinput}
-\end{Schunk}
+neatly
+
+XXX comments
+
+\begin{Schunk}
+\begin{Sinput}
+R> identifyColons <- function(v, nMax = Inf, ... ) {
++    v <- unique(na.omit(v))
++    problemMessage <- "Note: The following values include colons:"
++    problem <- FALSE
++    problemValues <- NULL
++  
++    # Use regular expressions to identify colons between two words
++    problemValues <- v[sapply(gregexpr("[[:xdigit:]]:[[:xdigit:]]", v),
++                              function(x) all(x != -1))]
++  
++    if (length(problemValues) > 0) {
++      problem <- TRUE
++    }
++  
++    problemStatus <- list(problem = problem,
++                          problemValues = problemValues)
++    # Use the messagegenerator function to produce the output
++    # message from the problemStatus and problemMessage
++    outMessage <- messageGenerator(problemStatus, problemMessage, nMax)
++  
++    checkResult(list(problem = problem,
++                     message = outMessage,
++                     problemValues = problemValues))
++  }
+R> identifyColons <- checkFunction(identifyColons,
++      description = "Identify colons surrounded by alphanumeric characters",
++      classes = c("character", "factor", "labelled"))
+\end{Sinput}
+\end{Schunk}
+
 As with the previous two functions, we also change its class. Note,
@@
 \end{itemize}
-These options are specified as follows: %\hl{Jeg synes ikke testData egner sig her.
-Der er ingen koloner og CPRvar er et mærkeligt, internt navn, hvis ikke man kender danske CPR-numre.
-Langt de fleste variable har desuden ingen nuller.
-Måske skal der laves et nyt datasæt?}
-\begin{Schunk}
-\begin{Sinput}
-> data("exampleData")
-> clean(exampleData,
-      # Add the new pre-check function
-      preChecks = c("isKey", "isEmpty", "isID"),
-      # Change the visual
-      allVisuals = "mosaicVisual",
-      # Add the new summary for specified data types
-      characterSummaries = c(defaultCharacterSummaries(), "countZeros"),
-      factorSummaries = c(defaultFactorSummaries(), "countZeros"),
-      labelledSummaries = c(defaultLabelledSummaries(), "countZeros"),
-      numericSummaries = c(defaultNumericSummaries(), "countZeros"),
-      integerSummaries = c(defaultIntegerSummaries(), "countZeros"),
-      # Add the new check for specific data types
-      characterChecks = c(defaultCharacterChecks(), "identifyColons"),
-      factorChecks = c(defaultFactorChecks(), "identifyColons"),
-      labelledCheck = c(defaultLabelledChecks(), "identifyColons"))
-\end{Sinput}
-\end{Schunk}
+These options are specified as follows: %\hl{Jeg synes ikke testData
+                                %egner sig her. Der er ingen koloner
+                                %og CPRvar er et mærkeligt, internt
+                                %navn, hvis ikke man kender danske
+                                %CPR-numre. Langt de fleste variable
+                                %har desuden ingen nuller. Måske skal
+                                %der laves et nyt datasæt?}
+\begin{Schunk}
+\begin{Sinput}
+R> data("exampleData")
+R> clean(exampleData,
++    preChecks = c("isKey", "isEmpty", "isID"),
++    allVisuals = "mosaicVisual",
++    characterSummaries = c(defaultCharacterSummaries(), "countZeros"),
++    factorSummaries = c(defaultFactorSummaries(), "countZeros"),
++    labelledSummaries = c(defaultLabelledSummaries(), "countZeros"),
++    numericSummaries = c(defaultNumericSummaries(), "countZeros"),
++    integerSummaries = c(defaultIntegerSummaries(), "countZeros"),
++    characterChecks = c(defaultCharacterChecks(), "identifyColons"),
++    factorChecks = c(defaultFactorChecks(), "identifyColons"),
++    labelledCheck = c(defaultLabelledChecks(), "identifyColons"))
+\end{Sinput}
+\end{Schunk}
+
 The outputted document is found in Appendix \ref{sec:appendix2}.
@@
 very large datasets. This is controlled via the \code{mode} argument:
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, mode = c("summarize", "check"))
-\end{Sinput}
-\end{Schunk}
+
+\begin{Schunk}
+\begin{Sinput}
+R> clean(toyData, mode = c("summarize", "check"))
+\end{Sinput}
+\end{Schunk}
+
 If figures are indeed needed, a different approach is to choose the
@@
 can be done by setting the \code{allVisuals = "basicVisual"} argument:
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, allVisuals = "basicVisual")
-\end{Sinput}
-\end{Schunk}
+
+\begin{Schunk}
+\begin{Sinput}
+R> clean(toyData, allVisuals = "basicVisual")
+\end{Sinput}
+\end{Schunk}
+
 Of course, even less heavy plots might be achieved by writing new
@@
 \begin{Sinput}
-> clean(toyData, render = FALSE, openResult = FALSE)
-> render("dataMaid_toyData.Rmd", quiet = FALSE)
-\end{Sinput}
-\end{Schunk}
+R> clean(toyData, render = FALSE, openResult = FALSE)
+R> render("dataMaid_toyData.Rmd", quiet = FALSE)
+\end{Sinput}
+\end{Schunk}
+
 This also deals with the fact that \pkg{dataMaid} can produce
@@
 done by also controlling the \code{mode} argument:
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, onlyProblematic = TRUE, mode = c("check"))
-\end{Sinput}
-\end{Schunk}
-Now only the checking results are printed, and only for variables where problems were identified.
-An even more minimal output can be generated by also leaving out the checking results --- then \code{clean()} essentially just produces a list of the variable names that should be investigated further:
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, onlyProblematic = TRUE, mode = NULL)
-\end{Sinput}
-\end{Schunk}
+
+\begin{Schunk}
+\begin{Sinput}
+R> clean(toyData, onlyProblematic = TRUE, mode = c("check"))
+\end{Sinput}
+\end{Schunk}
+
+Now only the checking results are printed, and only for variables
+where problems were identified. An even more minimal output can be
+generated by also leaving out the checking results --- then
+\code{clean()} essentially just produces a list of the variable names
+that should be investigated further:
+
+\begin{Schunk}
+\begin{Sinput}
+R> clean(toyData, onlyProblematic = TRUE, mode = NULL)
+\end{Sinput}
+\end{Schunk}
+
 Of course, this can also be done without generating an overview
@@
 \code{sapply()}:
-\begin{Schunk}
-\begin{Sinput}
-> toyChecks <- check(toyData)
-> foo <- function(x) {
->   any(sapply(x, function(y) y[["problem"]]))
-> }
-> sapply(toyChecks, foo)
+
+\begin{Schunk}
+\begin{Sinput}
+R> toyChecks <- check(toyData)
+R> foo <- function(x) {
++    any(sapply(x, function(y) y[["problem"]]))
++  }
+R> sapply(toyChecks, foo)
 \end{Sinput}
@@
 \end{Schunk}
+
 and we find that only a single variable in \code{toyData}, \code{var6} (for which all
@@
 
-In the this example, the \code{dataMaid\_toyData.Rmd} file could have been created as follows:
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, standAlone = FALSE)
-\end{Sinput}
-\end{Schunk}
+In the this example, the \code{dataMaid\_toyData.Rmd} file could have
+been created as follows:
+
+\begin{Schunk}
+\begin{Sinput}
+R> clean(toyData, standAlone = FALSE)
+\end{Sinput}
+\end{Schunk}
+
 and the more minimal, html-style \pkg{rmarkdown} file described above can be produced using
-\begin{Schunk}
-\begin{Sinput}
-> clean(toyData, standAlone = FALSE, output = "html", twoCols = FALSE)
-\end{Sinput}
-\end{Schunk}
+
+\begin{Schunk}
+\begin{Sinput}
+R> clean(toyData, standAlone = FALSE, output = "html", twoCols = FALSE)
+\end{Sinput}
+\end{Schunk}
+
 Note that with the latter option, no special YAML preamble is needed in the \pkg{rmarkdown} document.

### 2017-01-25 latex/article_vol2.Rnw
Mostly there

@@
 \proglang{R}-packages such as \pkg{plyr} \citep{plyr}, and
-\pkg{data.table} \citep{data.table} have
-made data wrangling a lot easier, but only a handful of packages such
-as \pkg{editrules} \citep{editrules}, \pkg{validate} \citep{validate},
-\pkg{DataCombine} \citep{DataCombine}, and
-\pkg{janitor} \citep{janitor} attempt to implement systematic, reproducible data
-cleaning.  These packages use different approaches for data cleaning:
-\pkg{editrules} and \pkg{validate} provide frameworks for setting up
-and checking constraints on the variables, while \pkg{DataCombine} and
-\pkg{janitor} both provide a few functions for identifying problems
-(e.g, duplicates, dates coded as numbers, etc.) in data.
+\pkg{data.table} \citep{data.table} have made data wrangling a lot
+easier, but only a handful of packages such as \pkg{editrules}
+\citep{editrules}, \pkg{validate} \citep{validate}, \pkg{DataCombine}
+\citep{DataCombine}, \pkg{janitor} \citep{janitor}, and
+\pkg{DescTools} \citep{DescTools} attempt to implement systematic,
+reproducible data cleaning.  These packages use different approaches
+for data cleaning: \pkg{editrules} and \pkg{validate} provide
+frameworks for setting up and checking constraints on the variables,
+while \pkg{DataCombine} and \pkg{janitor} both provide a few functions
+for identifying problems (e.g, duplicates, dates coded as numbers,
+etc.) in data.
 
@@
 \end{center}
-\label{fig:example1}
 \caption{Example output from running \code{clean()} on the \code{toyData}
@@
   Appendix~\ref{sec:appendix1}.}
+\label{fig:example1}
 \end{figure}
@@
     \node [block] (init) {Get next variable and run pre-checks};
-    \node [cloud, above of=init] (input) {Input \texttt{data.frame} or \texttt{tibble}};
+    \node [cloud, above of=init] (input) {Input \code{data.frame} or \code{tibble}};
 %    \node [cloud, right of=init] (system) {system};
@@
     \node [block, right of=precheck, node distance=4cm] (summarize)
-    {Run \texttt{summarize()} to produce summary table};
+    {Run \code{summarize()} to produce summary table};
     \node [block, below of=summarize, node distance=3cm] (visualize)
-    {Run \texttt{visualize()} to plot variable};
+    {Run \code{visualize()} to plot variable};
     \node [block, below of=visualize, node distance=2.5cm] (check)
-    {Call \texttt{check()} to run error checks};
+    {Call \code{check()} to run error checks};
     \node [decision, below of=check, node distance=2.7cm] (done) {More
@@
   \quad \code{basicVisual} & Histograms and barplots using base \proglang{R} graphics &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\
-  \quad \code{standardVisual} & Histograms and barplots using ggplot2 &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\
+  \quad \code{standardVisual} & Histograms and barplots using \pkg{ggplot2} &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\
   \smallskip \\
@@
 \begin{enumerate}
-\item In the pre-check stage
-\item As part of the summarize/visualize/check (SVC) steps
+\item In the pre-check stage.
+\item As part of the summarize/visualize/check (SVC) steps.
 \end{enumerate}
@@
 \end{center}
+\caption{Output from running \code{visualize()} on the variable \code{var2} from the
+\code{toyData} dataset.}
 \label{fig:example2}
-\caption{Output from running \code{visualize()} on the variable \texttt{var2} from the
-\code{toyData} dataset.}
 \end{figure}
@@
 attribute). In other words, a call on the following form should be made:
-\begin{Schunk}
-\begin{Sinput}
+\begin{Verbatim}
 mySummaryFunction <- summaryFunction(mySummaryFunction,
   description = "[Text describing what the summaryFunction does]",
-  classes = c([vector of data types that the function is intended for]))
-\end{Sinput}
-\end{Schunk}
+  classes = c([vector of data types that the function is intended
+  for]))
+\end{Verbatim}
 which adds the new function to the output of an
@@
 \pkg{rmarkdown} files. We recommend using the following structure:
-\begin{Schunk}
-\begin{Sinput}
+\begin{Verbatim}
 myVisualFunction <- function(v, vnam, doEval) {
@@
 }
-\end{Sinput}
-\end{Schunk}
+\end{Verbatim}
 
@@
 %as an attribute).
-\begin{Schunk}
-\begin{Sinput}
+\begin{Verbatim}
 myVisualFunction <- visualFunction(myVisualFunction,
@@
   )
-\end{Sinput}
-\end{Schunk}
+\end{Verbatim}
 
@@
 function using \code{checkFunction()}:
-\begin{Schunk}
-\begin{Sinput}
+\begin{Verbatim}
 myFullVarCheckFunction <- function(v, ...) {
@@
 should be expanded to follow a slightly more complicated structure:
-\begin{Schunk}
-\begin{Sinput}
+
+\begin{Verbatim}
 myProbValCheckFunction <- function(v, nMax, maxDecimals, ...) {
@@
 )
-\end{Sinput}
-\end{Schunk}
+\end{Verbatim}
+
 One comment should be devoted to the helper function,
@@
 above:
-\begin{Schunk}
-\begin{Sinput}
+
+\begin{Verbatim}
 myProbValCheckFunction <- function(v, nMax, maxDecimals, ...) {
@@
 }
-\end{Sinput}
-\end{Schunk}
-%$
+\end{Verbatim}
+% $
+
 Now, problematic values will be rounded in the outputted message, while they
@@
 \caption{Mosaic plot generated by the user-defined visualization
-  function \texttt{mosaicVisual}.}
+  function \code{mosaicVisual()}.}
 \label{fig:mosaicPlot}
@@
 \subsubsection{identifyColons --- a new checkFunction with problem values}
-The last function mentioned above is \code{identifyColons()}. We define
-it using the helper function \code{messageGenerator()} to obtain a properly
-escaped message, and we use \code{checkResult} to make its output print
-neatly
-
-XXX comments
+The last function mentioned above is \code{identifyColons()}. We
+define it using the helper function \code{messageGenerator()} to
+obtain a properly escaped message, and we use \code{checkResult()} to
+make its output print neatly. In the code we use regular expressions
+through the \code{gregexpr()} function to identify colons between two
+words.
 
@@
 
-  # Use regular expressions to identify colons between two words
   problemValues <- v[sapply(gregexpr("[[:xdigit:]]:[[:xdigit:]]", v),
@@
                         problemValues = problemValues)
-  # Use the messagegenerator function to produce the output
-  # message from the problemStatus and problemMessage
+
   outMessage <- messageGenerator(problemStatus, problemMessage, nMax)
@@
 }
-
+@
+
+YYY en kommentar om noget her, hvad der sker nedenfor
+
+<<>>=
 identifyColons <- checkFunction(identifyColons,
@@
 then be included:
+
 {\small

### 2017-01-25 latex/article_vol2.tex
Mostly there

@@
 +  
-+    # Use regular expressions to identify colons between two words
 +    problemValues <- v[sapply(gregexpr("[[:xdigit:]]:[[:xdigit:]]", v),
@@
 +                          problemValues = problemValues)
-+    # Use the messagegenerator function to produce the output
-+    # message from the problemStatus and problemMessage
++  
 +    outMessage <- messageGenerator(problemStatus, problemMessage, nMax)
@@
 +  }
+\end{Sinput}
+\end{Schunk}
+
+YYY en kommentar om noget her, hvad der sker nedenfor
+
+\begin{Schunk}
+\begin{Sinput}
 R> identifyColons <- checkFunction(identifyColons,

### 2017-01-29 latex/article_vol2.tex
Fix package reference
Updated article

@@
 make unsupervised decisions on what should be done to potential
-errors in the dataset, it can still be an extremely useful tool in the
+errors in a dataset, it can still be an extremely useful tool in the
 data cleaning process. Some errors can be tracked down and flagged by a
@@
 
-Online tools such as OpenRefine (\url{http://openrefine.org/}) and
-\proglang{R}-packages such as \pkg{plyr} \citep{plyr}, and
-\pkg{data.table} \citep{data.table} have made data wrangling a lot
-easier, but only a handful of packages such as \pkg{editrules}
-\citep{editrules}, \pkg{validate} \citep{validate}, \pkg{DataCombine}
-\citep{DataCombine}, \pkg{janitor} \citep{janitor}, and
-\pkg{DescTools} \citep{DescTools} attempt to implement systematic,
-reproducible data cleaning.  These packages use different approaches
-for data cleaning: \pkg{editrules} and \pkg{validate} provide
-frameworks for setting up and checking constraints on the variables,
-while \pkg{DataCombine} and \pkg{janitor} both provide a few functions
-for identifying problems (e.g, duplicates, dates coded as numbers,
-etc.) in data.
-
-While these tools attempt to alleviate the ubiquitous ad hoc approach
+A number of \proglang{R} packages made for such pre-analysis tools are already 
+available, including \pkg{janitor} \citep{janitor}, 
+\pkg{plyr} \citep{plyr}, \pkg{data.table} \citep{data.table}, 
+\pkg{DataCombine} \citep{DataCombine} , \pkg{validate} \citep{validate}, and
+\pkg{DataExplorer} \citep{DataExplorer}. These packages focus on different 
+stages of the pre-analysis work, most of which are not really data cleaning.
+\pkg{janitor} provides tools for data loading with a
+particular emphasis on the challenges of getting neat \code{data.frame}s from 
+Microsoft Excel data files. \pkg{plyr}, \pkg{data.table} and \pkg{DataCombine} go a few
+steps further by providing a wide array of extremely powerful tools for data wrangling, 
+including a number of particularly useful functions for merging and working with
+very large datasets. When it comes to actual data cleaning, however, the options are fewer.
+\pkg{validate} (and the similar packages \pkg{editrules} and \pkg{deducorrect} from the same 
+authors) does offer a few tools for identifying and auto-correcting errors in a dataset, 
+but, as the name suggests, the focus is on internal validity rather than 
+general data cleaning. In practice, this means that quite elegant tools for e.g. linear 
+restraints among the variables in a dataset can be applied, but e.g. looking for potentially
+miscoded missing values is not really feasible. The main difference between these two challenges
+is the direction in which the data is inspected: While linear constraints work row-wise
+with no ambiguity, determining whether or not something is a miscoded missing value 
+often requires knowledge about the full variable (e.g. range or data type), and thus
+it should be performed column-wise. \pkg{validate} does not allow for user-defined
+extensions of the latter type, thereby limiting its data cleaning potential.
+Moreover, no tools have been made for documenting the data cleaning process. 
+As automatic cleaning options are also supplied in \pkg{validate},  
+we find quite a dangerous cocktail where all power is given to the the computer 
+with no supervision, neither in real time nor subsequently.
+
+All in all, the large role of data cleaning in any data analysts everyday endeavors is 
+hardly matched in the amount of available \proglang{R} software solutions. In particular,
+few packages attempt to implement systematic, reproducible data cleaning. And 
+while the available tools attempt to alleviate the ubiquitous ad hoc approach
 to data cleaning, they are primarily intended for the data savvy users
@@
 field and the context of the available data. The \pkg{dataMaid}
-package \citep{dataMaid}
-tries to address this by providing a framework that both allows for
+package \citep{dataMaid} tries to address this by providing a framework that both allows for
 extendable, systematic, reproducible data cleaning, and summarizing
@@
 
-% \begin{itemize}
-% \item Makes the point that data cleaning is usually done in a very ad hoc manner
-% \item Maybe hints to the discussion later about poor documentation of data cleaning/quality assessment as the status quo for most analysts
-% \item Gives a specification of what we mean by data errors/mistakes/whatever.
-I'm not completely sure this will be clear to everyone and we should probably mention it early on to not lose readers/users.
-% \end{itemize}
-% }
-
-Data cleaning is a time consuming endeavor, as it inherently requires
+One last package that should be mentioned in this context is \pkg{DataExplorer}. While this
+package does not address data cleaning issues per se, its general strategy is quite similar
+to that of \pkg{dataMaid} and to the paradigms presented below. This package provides 
+a few simple, but practical, tools for exploratory data analysis, including automated 
+documentation. Therefore, we find \pkg{DataExplorer} to be a good candidate for a next-step
+package after data cleaning is finished.
+
+But no matter how clever software tools we make, data cleaning 
+reamins to be a time consuming endeavor, as it inherently requires
 human interaction since every dataset is different and the variables
@@
 only get the median, minimum and the maximum we set
-\code{numericSummaries=c("centralValue", "minMax")}:
+\code{numericSummaries = c("centralValue", "minMax")}:
 
@@
 a vector of the names of the classes for which the function can be
-called. If \code{mySummaryFunction} is not an \code{S3} generic and
+called. If \code{mySummaryFunction()} is not an \code{S3} generic and
 \code{classes} is left unspecified, the attribute will simply be
@@
 more complicated than writing a \code{summaryFunction}. This follows from
-the fact that \code{visualFunctions} need to be able to output standalone
+the fact that \code{visualFunction}s need to be able to output standalone
 code for plots in order for \code{clean()} to build standalone
@@
 \begin{description}
-\item[\code{isID}] A new \code{checkFunction} intended for use in the precheck-stage.
+\item[\code{isID()}] A new \code{checkFunction} intended for use in the precheck-stage.
 This function checks whether a variable consists exclusively of long ($> 10$ characters/digits) entries that are all of equal length, as this might be personal identification codes that we do not wish to print out in the data summary.
-\item[\code{mosaicVisual}] A new \code{visualFunction} that produces mosaic plots. This function will be used in the \textit{visualize} step of \code{clean()}.
-\item[\code{countZeros}] A new \code{summaryFunction} that counts the number of occurrences of the value \code{0} in a variable.
+\item[\code{mosaicVisual()}] A new \code{visualFunction} that produces mosaic plots. This function will be used in the \textit{visualize} step of \code{clean()}.
+\item[\code{countZeros()}] A new \code{summaryFunction} that counts the number of occurrences of the value \code{0} in a variable.
 This function will be used in the \textit{summarize} step of \code{clean()}.
-\item[\code{identifyColons}] A new \code{checkFunction} that flags variables in which values have colons that appear before and after alphanumerical characters.
+\item[\code{identifyColons()}] A new \code{checkFunction} that flags variables in which values have colons that appear before and after alphanumerical characters.
 This is practical for identifying autogenerated interaction effects.
@@
 supposed to list problematic values, it falls within
-the category of \code{checkFunctions} represented by
+the category of \code{checkFunction}s represented by
 \code{myFullVarCheckFunction()} in the above. We do not particularly wish
@@
 functionality. This can be done using the following code which sets up
-the call using the existing function \code{mosaicplot}.
+the call using the existing function \code{mosaicplot()}.
 
@@
 
-As \code{mosaicVisual} is now a full-blooded \code{visualFunction}, it
+As \code{mosaicVisual()} is now a full-blooded \code{visualFunction}, it
 will also be included in the \code{allVisualFunctions()} output table:
@@
 Now that we are done with the definition of \code{mosaicVisual()} we can
-turn to the next function in line, \code{countZeros}.
+turn to the next function in line, \code{countZeros()}.
 
@@
 \code{summaryResult} object, a printing method is automatically called
-when \code{countZeros} is used interactively:
+when \code{countZeros()} is used interactively:
 
@@
 
-YYY en kommentar om noget her, hvad der sker nedenfor
+Once again, we change the class of the function so that it becomes a proper \code{checkFunction} object:
 
@@
 
-As with the previous two functions, we also change its class. Note,
-however, that for \code{checkFunction}s, the function description will
+Note, that for \code{checkFunction}s, the function description will
 appear in the document produced by \code{clean()} (in the \textit{Data
-  cleaning summary} section of the output), so now this is not only
-done for the sake of the \code{allCheckFunctions()} output.
+  cleaning summary} section of the output), so for this type of function, 
+the change of class is not only done for the sake of the \code{allCheckFunctions()} output.
 
@@
 \begin{itemize}
-\item We want to add the new pre-check function, \code{isID}, to the already existing pre-checks.
+\item We want to add the new pre-check function, \code{isID()}, to the already existing pre-checks.
 \item We wish to change the plot type for all variables to the new mosaic plot.
-\item We want the new summary function, \code{countZeros}, to be added to the summaries performed on all variable types but \code{Date} and \code{logical}.
-\item We want the new check function, \code{identifyColon}, to be added to the checks performed on \code{character}, \code{factor} and \code{labelled} variables.
+\item We want the new summary function, \code{countZeros()}, to be added to the summaries performed on all variable types but \code{Date} and \code{logical}.
+\item We want the new check function, \code{identifyColons()}, to be added to the checks performed on \code{character}, \code{factor} and \code{labelled} variables.
 \end{itemize}
-These options are specified as follows: %\hl{Jeg synes ikke testData
-                                %egner sig her. Der er ingen koloner
-                                %og CPRvar er et mærkeligt, internt
-                                %navn, hvis ikke man kender danske
-                                %CPR-numre. Langt de fleste variable
-                                %har desuden ingen nuller. Måske skal
-                                %der laves et nyt datasæt?}
+These options are specified as follows: 
 \begin{Schunk}

### 2017-01-29 latex/dataMaid_toyData.Rmd
Updated article

@@
   - \newcommand{\fullline}{\noindent\makebox[\linewidth]{\rule{\textwidth}{0.4pt}}}
+  - \newcommand{\bminione}{\begin{minipage}{0.75 \textwidth}}
+  - \newcommand{\bminitwo}{\begin{minipage}{0.25 \textwidth}}
+  - \newcommand{\emini}{\end{minipage}}
 ---
@@
 ------------------------------
-
-* Only variables that were deemed potentially problematic are included in this summary.
 
@@
 # Variable list
+## **var1**
+
+\bminione
+
+---------------------------------------
+Feature                          Result
+------------------------- -------------
+Variable type                    factor
+
+Number of missing obs.      2 (13.33 %)
+
+Number of unique values               2
+
+Mode                              "red"
+---------------------------------------
+
+\emini
+\bminitwo
+```{r Var-1-var1, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
+qplot(x = structure(c(2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 2L, 
+1L, 1L, 1L), .Label = c("blue", "red"), class = "factor", na.action = structure(14:15, class = "omit")), 
+    geom = "bar", xlab = "", main = "var1")
+```
+\emini
+
+- Note that the following levels have at most five observations: \"blue\".
+
+\fullline
+
+## **var2**
+
+\bminione
+
+------------------------------------
+Feature                       Result
+------------------------- ----------
+Variable type                numeric
+
+Number of missing obs.      3 (20 %)
+
+Number of unique values            8
+
+Median                           4.5
+
+1st and 3rd quartiles        1.75; 6
+
+Min. and max.                 1; 999
+------------------------------------
+
+\emini
+\bminitwo
+```{r Var-2-var2, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
+qplot(x = c(1, 1, 1, 2, 2, 6, 6, 6, 999, 4, 82, 5), geom = "histogram", 
+    xlab = "", main = "var2", bins = 20)
+```
+\emini
+
+- The following suspected missing value codes enter as regular values: \"999\", \"NaN\".
+
+- Note that the following possible outlier values were detected: \"82\", \"999\".
+
+\fullline
+
+## **var3**
+
+\bminione
+
+-----------------------------------
+Feature                      Result
+------------------------- ---------
+Variable type                factor
+
+Number of missing obs.      0 (0 %)
+
+Number of unique values           7
+
+Mode                            "a"
+-----------------------------------
+
+\emini
+\bminitwo
+```{r Var-3-var3, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
+qplot(x = structure(c(3L, 3L, 3L, 3L, 3L, 4L, 4L, 4L, 5L, 5L, 
+5L, 2L, 1L, 6L, 7L), .Label = c(" ", ".", "a", "b", "c", "other", 
+"OTHER"), class = "factor"), geom = "bar", xlab = "", main = "var3")
+```
+\emini
+
+- The following suspected missing value codes enter as regular values: \" \", \".\".
+
+- The following values appear with prefixed or suffixed white space: \" \".
+
+- Note that the following levels have at most five observations: \" \", \".\", \"a\", \"b\", \"c\", \"other\", \"OTHER\".
+
+- Note that there might be case problems with the following levels: \"other\", \"OTHER\".
+
+\fullline
+
+## **var4**
+
+\bminione
+
+---------------------------------------
+Feature                          Result
+------------------------- -------------
+Variable type                   numeric
+
+Number of missing obs.          0 (0 %)
+
+Number of unique values              15
+
+Median                             0.33
+
+1st and 3rd quartiles       -0.62; 0.66
+
+Min. and max.                -2.21; 1.6
+---------------------------------------
+
+\emini
+\bminitwo
+```{r Var-4-var4, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
+qplot(x = c(-0.626453810742332, 0.183643324222082, -0.835628612410047, 
+1.59528080213779, 0.329507771815361, -0.820468384118015, 0.487429052428485, 
+0.738324705129217, 0.575781351653492, -0.305388387156356, 1.51178116845085, 
+0.389843236411431, -0.621240580541804, -2.2146998871775, 1.12493091814311
+), geom = "histogram", xlab = "", main = "var4", bins = 20)
+```
+\emini
+
+- Note that the following possible outlier values were detected: \"1.12\", \"1.51\", \"1.6\".
+
+\fullline
+
 ## **var5**

### 2017-01-30 latex/article_vol2.Rnw
Fix package reference
DataExplorer
Rewritten part 1

@@
 \citep{DataCombine}, \pkg{janitor} \citep{janitor}, and
-\pkg{DescTools} \citep{DescTools} attempt to implement systematic,
+\pkg{DataExplorer} \citep{DataExplorer} attempt to implement systematic,
 reproducible data cleaning.  These packages use different approaches
@@
 
-YYY en kommentar om noget her, hvad der sker nedenfor
+Once again, we change the class of the function so that it becomes a
+proper \code{checkFunction} object:
 
@@
 
-As with the previous two functions, we also change its class. Note,
-however, that for \code{checkFunction}s, the function description will
+Note, that for \code{checkFunction}s, the function description will
 appear in the document produced by \code{clean()} (in the \textit{Data
-  cleaning summary} section of the output), so now this is not only
-done for the sake of the \code{allCheckFunctions()} output.
+  cleaning summary} section of the output), so for this type of
+function, the change of class is not only done for the sake of the
+\code{allCheckFunctions()} output.
 
@@
 \begin{itemize}
-\item We want to add the new pre-check function, \code{isID}, to the already existing pre-checks.
+\item We want to add the new pre-check function, \code{isID()}, to the already existing pre-checks.
 \item We wish to change the plot type for all variables to the new mosaic plot.
-\item We want the new summary function, \code{countZeros}, to be added to the summaries performed on all variable types but \code{Date} and \code{logical}.
-\item We want the new check function, \code{identifyColon}, to be added to the checks performed on \code{character}, \code{factor} and \code{labelled} variables.
+\item We want the new summary function, \code{countZeros()}, to be added to the summaries performed on all variable types but \code{Date} and \code{logical}.
+\item We want the new check function, \code{identifyColons()}, to be added to the checks performed on \code{character}, \code{factor} and \code{labelled} variables.
 \end{itemize}
-These options are specified as follows: %\hl{Jeg synes ikke testData
-                                %egner sig her. Der er ingen koloner
-                                %og CPRvar er et mærkeligt, internt
-                                %navn, hvis ikke man kender danske
-                                %CPR-numre. Langt de fleste variable
-                                %har desuden ingen nuller. Måske skal
-                                %der laves et nyt datasæt?}
+These options are specified as follows:
+
 <<eval=FALSE>>=

### 2017-01-30 latex/article_vol2.Rnw
Rewritten part 2

@@
 
-Online tools such as OpenRefine (\url{http://openrefine.org/}) and
-\proglang{R}-packages such as \pkg{plyr} \citep{plyr}, and
-\pkg{data.table} \citep{data.table} have made data wrangling a lot
-easier, but only a handful of packages such as \pkg{editrules}
-\citep{editrules}, \pkg{validate} \citep{validate}, \pkg{DataCombine}
-\citep{DataCombine}, \pkg{janitor} \citep{janitor}, and
-\pkg{DataExplorer} \citep{DataExplorer} attempt to implement systematic,
-reproducible data cleaning.  These packages use different approaches
-for data cleaning: \pkg{editrules} and \pkg{validate} provide
-frameworks for setting up and checking constraints on the variables,
-while \pkg{DataCombine} and \pkg{janitor} both provide a few functions
-for identifying problems (e.g, duplicates, dates coded as numbers,
-etc.) in data.
-
-While these tools attempt to alleviate the ubiquitous ad hoc approach
-to data cleaning, they are primarily intended for the data savvy users
-and less so for the general researcher with a knowledge about a specific
-field and the context of the available data. The \pkg{dataMaid}
-package \citep{dataMaid} tries to address this by providing a framework that both allows for
-extendable, systematic, reproducible data cleaning, and summarizing
-findings for researchers from other fields such that they can act as
-human experts when tracking down potential errors.
-
-% \begin{itemize}
-% \item Makes the point that data cleaning is usually done in a very ad hoc manner
-% \item Maybe hints to the discussion later about poor documentation of data cleaning/quality assessment as the status quo for most analysts
-% \item Gives a specification of what we mean by data errors/mistakes/whatever.
-I'm not completely sure this will be clear to everyone and we should probably mention it early on to not lose readers/users.
-% \end{itemize}
-% }
-
-Data cleaning is a time consuming endeavor, as it inherently requires
-human interaction since every dataset is different and the variables
-in the dataset can only be understood in the proper context of their
+A number of \proglang{R} packages made for such pre-analysis tools are
+already available, including \pkg{janitor} \citep{janitor}, \pkg{plyr}
+\citep{plyr}, \pkg{data.table} \citep{data.table}, \pkg{DataCombine}
+\citep{DataCombine} , \pkg{validate} \citep{validate}, and
+\pkg{DataExplorer} \citep{DataExplorer}. These packages focus on
+different stages of the pre-analysis work, most of which are not
+really data cleaning.  \pkg{janitor} provides tools for data loading
+with a particular emphasis on the challenges of getting neat
+\code{data.frame}s from Microsoft Excel data files. \pkg{plyr},
+\pkg{data.table} and \pkg{DataCombine} go a few steps further by
+providing a wide array of extremely powerful tools for data wrangling,
+including a number of particularly useful functions for merging and
+working with very large datasets. When it comes to actual data
+cleaning, however, the options are fewer.  \pkg{validate} (and the
+similar packages \pkg{editrules} and \pkg{deducorrect} from the same
+authors) does offer a few tools for identifying and auto-correcting
+errors in a dataset, but, as the name suggests, the focus is on
+internal validity rather than general data cleaning. In practice, this
+means that quite elegant tools for e.g. linear restraints among the
+variables in a dataset can be applied, but e.g. looking for
+potentially miscoded missing values is not really feasible. The main
+difference between these two challenges is the direction in which the
+data is inspected: While linear constraints work row-wise with no
+ambiguity, determining whether or not something is a miscoded missing
+value often requires knowledge about the full variable (e.g. range or
+data type), and thus it should be performed
+column-wise. \pkg{validate} does not allow for user-defined extensions
+of the latter type, thereby limiting its data cleaning potential.
+Moreover, no tools have been made for documenting the data cleaning
+process.  As automatic cleaning options are also supplied in
+\pkg{validate}, we find quite a dangerous cocktail where all power is
+given to the the computer with no supervision, neither in real time
+nor subsequently.
+
+All in all, the large role of data cleaning in any data analysts
+everyday endeavors is hardly matched in the amount of available
+\proglang{R} software solutions. In particular, few packages attempt
+to implement systematic, reproducible data cleaning. And while the
+available tools attempt to alleviate the ubiquitous ad hoc approach to
+data cleaning, they are primarily intended for the data savvy users
+and less so for the general researcher with a knowledge about a
+specific field and the context of the available data. The
+\pkg{dataMaid} package \citep{dataMaid} tries to address this by
+providing a framework that both allows for extendable, systematic,
+reproducible data cleaning, and summarizing findings for researchers
+from other fields such that they can act as human experts when
+tracking down potential errors.
+
+One last package that should be mentioned in this context is
+\pkg{DataExplorer}. While this package does not address data cleaning
+issues per se, its general strategy is quite similar to that of
+\pkg{dataMaid} and to the paradigms presented below. This package
+provides a few simple, but practical, tools for exploratory data
+analysis, including automated documentation. Therefore, we find
+\pkg{DataExplorer} to be a good candidate for a next-step package
+after data cleaning is finished.
+
+But no matter how clever software tools we make, data cleaning remains
+to be a time consuming endeavor, as it inherently requires human
+interaction since every dataset is different and the variables in the
+dataset can only be understood in the proper context of their
 origin. This often requires a collaborative effort between an expert
@@
 stored under the \code{description} attribute. What happens if the \code{classes} argument is
-not specified depends on the type of \code{mySummaryFunction}. If
+not specified depends on the type of \code{summaryFunction}. If
 \code{mySummaryFunction} is constructed as an \code{S3} generic function with associated

### 2017-01-30 latex/article_vol2.tex
Rewritten part 1
Rewritten part 2

@@
 
-A number of \proglang{R} packages made for such pre-analysis tools are already 
-available, including \pkg{janitor} \citep{janitor}, 
-\pkg{plyr} \citep{plyr}, \pkg{data.table} \citep{data.table}, 
-\pkg{DataCombine} \citep{DataCombine} , \pkg{validate} \citep{validate}, and
-\pkg{DataExplorer} \citep{DataExplorer}. These packages focus on different 
-stages of the pre-analysis work, most of which are not really data cleaning.
-\pkg{janitor} provides tools for data loading with a
-particular emphasis on the challenges of getting neat \code{data.frame}s from 
-Microsoft Excel data files. \pkg{plyr}, \pkg{data.table} and \pkg{DataCombine} go a few
-steps further by providing a wide array of extremely powerful tools for data wrangling, 
-including a number of particularly useful functions for merging and working with
-very large datasets. When it comes to actual data cleaning, however, the options are fewer.
-\pkg{validate} (and the similar packages \pkg{editrules} and \pkg{deducorrect} from the same 
-authors) does offer a few tools for identifying and auto-correcting errors in a dataset, 
-but, as the name suggests, the focus is on internal validity rather than 
-general data cleaning. In practice, this means that quite elegant tools for e.g. linear 
-restraints among the variables in a dataset can be applied, but e.g. looking for potentially
-miscoded missing values is not really feasible. The main difference between these two challenges
-is the direction in which the data is inspected: While linear constraints work row-wise
-with no ambiguity, determining whether or not something is a miscoded missing value 
-often requires knowledge about the full variable (e.g. range or data type), and thus
-it should be performed column-wise. \pkg{validate} does not allow for user-defined
-extensions of the latter type, thereby limiting its data cleaning potential.
-Moreover, no tools have been made for documenting the data cleaning process. 
-As automatic cleaning options are also supplied in \pkg{validate},  
-we find quite a dangerous cocktail where all power is given to the the computer 
-with no supervision, neither in real time nor subsequently.
-
-All in all, the large role of data cleaning in any data analysts everyday endeavors is 
-hardly matched in the amount of available \proglang{R} software solutions. In particular,
-few packages attempt to implement systematic, reproducible data cleaning. And 
-while the available tools attempt to alleviate the ubiquitous ad hoc approach
-to data cleaning, they are primarily intended for the data savvy users
-and less so for the general researcher with a knowledge about a specific
-field and the context of the available data. The \pkg{dataMaid}
-package \citep{dataMaid} tries to address this by providing a framework that both allows for
-extendable, systematic, reproducible data cleaning, and summarizing
-findings for researchers from other fields such that they can act as
-human experts when tracking down potential errors.
-
-One last package that should be mentioned in this context is \pkg{DataExplorer}. While this
-package does not address data cleaning issues per se, its general strategy is quite similar
-to that of \pkg{dataMaid} and to the paradigms presented below. This package provides 
-a few simple, but practical, tools for exploratory data analysis, including automated 
-documentation. Therefore, we find \pkg{DataExplorer} to be a good candidate for a next-step
-package after data cleaning is finished.
-
-But no matter how clever software tools we make, data cleaning 
-reamins to be a time consuming endeavor, as it inherently requires
-human interaction since every dataset is different and the variables
-in the dataset can only be understood in the proper context of their
+A number of \proglang{R} packages made for such pre-analysis tools are
+already available, including \pkg{janitor} \citep{janitor}, \pkg{plyr}
+\citep{plyr}, \pkg{data.table} \citep{data.table}, \pkg{DataCombine}
+\citep{DataCombine} , \pkg{validate} \citep{validate}, and
+\pkg{DataExplorer} \citep{DataExplorer}. These packages focus on
+different stages of the pre-analysis work, most of which are not
+really data cleaning.  \pkg{janitor} provides tools for data loading
+with a particular emphasis on the challenges of getting neat
+\code{data.frame}s from Microsoft Excel data files. \pkg{plyr},
+\pkg{data.table} and \pkg{DataCombine} go a few steps further by
+providing a wide array of extremely powerful tools for data wrangling,
+including a number of particularly useful functions for merging and
+working with very large datasets. When it comes to actual data
+cleaning, however, the options are fewer.  \pkg{validate} (and the
+similar packages \pkg{editrules} and \pkg{deducorrect} from the same
+authors) does offer a few tools for identifying and auto-correcting
+errors in a dataset, but, as the name suggests, the focus is on
+internal validity rather than general data cleaning. In practice, this
+means that quite elegant tools for e.g. linear restraints among the
+variables in a dataset can be applied, but e.g. looking for
+potentially miscoded missing values is not really feasible. The main
+difference between these two challenges is the direction in which the
+data is inspected: While linear constraints work row-wise with no
+ambiguity, determining whether or not something is a miscoded missing
+value often requires knowledge about the full variable (e.g. range or
+data type), and thus it should be performed
+column-wise. \pkg{validate} does not allow for user-defined extensions
+of the latter type, thereby limiting its data cleaning potential.
+Moreover, no tools have been made for documenting the data cleaning
+process.  As automatic cleaning options are also supplied in
+\pkg{validate}, we find quite a dangerous cocktail where all power is
+given to the the computer with no supervision, neither in real time
+nor subsequently.
+
+All in all, the large role of data cleaning in any data analysts
+everyday endeavors is hardly matched in the amount of available
+\proglang{R} software solutions. In particular, few packages attempt
+to implement systematic, reproducible data cleaning. And while the
+available tools attempt to alleviate the ubiquitous ad hoc approach to
+data cleaning, they are primarily intended for the data savvy users
+and less so for the general researcher with a knowledge about a
+specific field and the context of the available data. The
+\pkg{dataMaid} package \citep{dataMaid} tries to address this by
+providing a framework that both allows for extendable, systematic,
+reproducible data cleaning, and summarizing findings for researchers
+from other fields such that they can act as human experts when
+tracking down potential errors.
+
+One last package that should be mentioned in this context is
+\pkg{DataExplorer}. While this package does not address data cleaning
+issues per se, its general strategy is quite similar to that of
+\pkg{dataMaid} and to the paradigms presented below. This package
+provides a few simple, but practical, tools for exploratory data
+analysis, including automated documentation. Therefore, we find
+\pkg{DataExplorer} to be a good candidate for a next-step package
+after data cleaning is finished.
+
+But no matter how clever software tools we make, data cleaning remains
+to be a time consuming endeavor, as it inherently requires human
+interaction since every dataset is different and the variables in the
+dataset can only be understood in the proper context of their
 origin. This often requires a collaborative effort between an expert
@@
 only get the median, minimum and the maximum we set
-\code{numericSummaries = c("centralValue", "minMax")}:
+\code{numericSummaries=c("centralValue", "minMax")}:
 
@@
 \begin{description}
-\item[\code{isID()}] A new \code{checkFunction} intended for use in the precheck-stage.
+\item[\code{isID}] A new \code{checkFunction} intended for use in the precheck-stage.
 This function checks whether a variable consists exclusively of long ($> 10$ characters/digits) entries that are all of equal length, as this might be personal identification codes that we do not wish to print out in the data summary.
-\item[\code{mosaicVisual()}] A new \code{visualFunction} that produces mosaic plots. This function will be used in the \textit{visualize} step of \code{clean()}.
-\item[\code{countZeros()}] A new \code{summaryFunction} that counts the number of occurrences of the value \code{0} in a variable.
+\item[\code{mosaicVisual}] A new \code{visualFunction} that produces mosaic plots. This function will be used in the \textit{visualize} step of \code{clean()}.
+\item[\code{countZeros}] A new \code{summaryFunction} that counts the number of occurrences of the value \code{0} in a variable.
 This function will be used in the \textit{summarize} step of \code{clean()}.
-\item[\code{identifyColons()}] A new \code{checkFunction} that flags variables in which values have colons that appear before and after alphanumerical characters.
+\item[\code{identifyColons}] A new \code{checkFunction} that flags variables in which values have colons that appear before and after alphanumerical characters.
 This is practical for identifying autogenerated interaction effects.
@@
 
-Once again, we change the class of the function so that it becomes a proper \code{checkFunction} object:
+Once again, we change the class of the function so that it becomes a
+proper \code{checkFunction} object:
 
@@
 appear in the document produced by \code{clean()} (in the \textit{Data
-  cleaning summary} section of the output), so for this type of function, 
-the change of class is not only done for the sake of the \code{allCheckFunctions()} output.
+  cleaning summary} section of the output), so for this type of
+function, the change of class is not only done for the sake of the
+\code{allCheckFunctions()} output.
 
@@
 These options are specified as follows:
+
 \begin{Schunk}

### 2017-01-30 latex/article_vol2.Rnw
Fixed

@@
 One last package that should be mentioned in this context is
-\pkg{DataExplorer}. While this package does not address data cleaning
-issues per se, its general strategy is quite similar to that of
-\pkg{dataMaid} and to the paradigms presented below. This package
-provides a few simple, but practical, tools for exploratory data
-analysis, including automated documentation. Therefore, we find
-\pkg{DataExplorer} to be a good candidate for a next-step package
-after data cleaning is finished.
+\pkg{DataExplorer} \citep{DataExplorer}. While this package does not
+address data cleaning issues \emph{per se}, its general strategy is
+quite similar to that of \pkg{dataMaid} and to the paradigms presented
+below. This package provides a few simple, but practical, tools for
+exploratory data analysis, including automated
+documentation. Therefore, we find \pkg{DataExplorer} to be a good
+candidate for a next-step package after data cleaning is finished.
 

### 2017-02-01 latex/article_vol2.Rnw
Fixed identifyMissing

@@
   arise for any number of reasons, including erroneous codings,
-  malfunctioning measurement equipment, inconsistent data generation
-  manuals and many more.  Ideally, a human investigator should go
+  malfunctioning measurement equipment, and inconsistent data generation
+  manuals.  Ideally, a human investigator should go
   through each variable in the dataset and look for potential errors
@@
 
-  We describe an \proglang{R} package which implements an extensive
-  and customizeable suite of quality assessment tools that can be applied to
-  a dataset in order to identify potential problems in its
-  variables. The results can be presented in an auto-generated,
-  non-technical, stand-alone overview document, intended to be perused
-  by an investigator with an understanding of the variables in the
-  data, but not necessarily knowledge of \proglang{R}. Thereby, \pkg{dataMaid}
-  aids the dialogue between data
-  analysts and field experts, while also providing easy
+  We describe an \proglang{R} package, \pkg{dataMaid}, which
+  implements an extensive and customizeable suite of quality
+  assessment tools that can be applied to a dataset in order to
+  identify potential problems in its variables. The results can be
+  presented in an auto-generated, non-technical, stand-alone overview
+  document intended to be perused by an investigator with an
+  understanding of the variables in the data, but not necessarily
+  knowledge of \proglang{R}. Thereby, \pkg{dataMaid} aids the dialogue
+  between data analysts and field experts, while also providing easy
   documentation of reproducible data cleaning steps and data quality
@@
   more typical \proglang{R} tools for interactive data quality
-  assessment and -cleaning, where the data inspections all live and die
-  within the \proglang{R} console.
+  assessment and -cleaning, where the data inspections all live and
+  die within the \proglang{R} console.
 
@@
 analysts find themselves spending a large portion of their time on
-data cleaning and on data wrangling. While a computer should never
+data cleaning and data wrangling. While a computer should never
 make unsupervised decisions on what should be done to potential
@@
 By default, an \proglang{R} markdown file and a rendered pdf overview
-document is produced, saved to the disc (in the working directory) and
+document is produced, saved to the working directory and
 opened for immediate inspection. Turning to Figure~\ref{fig:example1},
@@
                        should the \code{rmarkdown} file be opened? & \code{TRUE} \\
-\quad \code{replace} & Overwrite an existing files with the same name? & \code{FALSE} \\
+\quad \code{replace} & Overwrite an existing file with the same name? & \code{FALSE} \\
 
@@
 \subsection{Polishing off the arguments}
-We begin with an example that is intended as an illustration of how \code{clean()} might
-be used in the very first stages of data cleaning, where we are still uncertain
-about exactly what is needed and how much times should be allocated to data
-cleaning. At this stage, what is really needed, if a very rough idea of the extends of
-errors in the dataset. In this scenario, we might wish to obtain a summary document in html format
-that only contains the variables with potential problems, and with a limit of
-maximum 10 printed potential errors for each variable we can write
-(output not shown). Also, we can add the argument \code{replace=TRUE} in
-order to force \code{clean()} to overwrite any existing files produced by
-\code{clean()}, as we might run this function a few times in a row before deciding
-exactly which cleaning overview we want. Using the \code{toyData} dataset as a guinea pig, we type:
+We begin with an example that is intended as an illustration of how
+\code{clean()} might be used in the very first stages of data
+cleaning, when the we are uncertain about the complexities of the
+errors and how much times should be allocated to data cleaning. At
+this stage, what is really needed, is a very rough idea of the
+severity of errors in the dataset. In this scenario, we might wish to
+obtain a summary document in html format that only contains the
+variables with potential problems, and with a limit of, say, maximum
+10 printed potential errors for each variable. Also, we can add the
+argument \code{replace=TRUE} in order to force \code{clean()} to
+overwrite any existing files produced by \code{clean()}.  Using the
+\code{toyData} dataset as a guinea pig, we type:
 
@@
 cleaning summary is manipulated, but also its very contents. This is
-done through the summarize-visualize-check (SVC) step, as illustrated
+done through the summarize/visualize/check (SVC) steps, as illustrated
 in Figure \ref{figure:cleanStructure}.  \pkg{dataMaid} uses three
@@
 
-Thus we can see, for example, that for \code{numeric}, \code{integer}, or \code{Date}
-variables, \pkg{dataMaid} provides functions for adding summary information about
-the minimum and maximum values, while all seven variable
-classes dealt with in \pkg{dataMaid} have functions for central
-tendency summaries (i.e., mode or median).
+Thus we can see, for example, that for \code{numeric}, \code{integer},
+and \code{Date} variables, \pkg{dataMaid} provides functions for
+adding summary information about the minimum and maximum values, while
+all seven variable classes dealt with in \pkg{dataMaid} have functions
+for central tendency summaries (i.e., mode or median).
 
@@
 and \code{defaultXXXSummaries()} functions, where \code{XXX} again refers to
-the variable type.
-
-For example, the default summaries being used for a factor variable is
+the variable type. For example, the default summaries being used for a factor variable is
+
 <<>>=
@@
 
-Now, if we only wanted to apply the identify whitespace function for
-factor variables, then we would need to set the \code{factorChecks} accordingly
+Now, if we only wanted to apply the function to identify whitespace
+for factor variables, then we would need to set the
+\code{factorChecks} accordingly
 
@@
 usage beyond the \code{XXXChecks} arguments, namely in the
-\textit{pre-check} stage. In this stage, it is determined whether or
+pre-check stage. In this stage, it is determined whether or
 not each variable is suitable for the summarize/visualize/check (SVC)
@@
 controllable, though only a single function can be provided for all
-variable types. If for instance, we wish to change the visualizations
+variable types. If, for instance, we wish to change the visualizations
 from the default \pkg{ggplot2} \citep{ggplot2} style histograms and barplots to base
@@
 
-In summary, and as indicated in Figure~\ref{figure:cleanStructure}, there are two stages
+As indicated in Figure~\ref{figure:cleanStructure}, there are two stages
 where \code{clean()} applies functions to each of the variables:
@@
 Each of these stages are controllable using appropriate function
-arguments in \code{clean()}, and above we have shown examples of how to
-tweak them to modify the data cleaning outputs. However, if for
-instance the dataset at hand requires new, additional checks, then more control is needed, and
-Section~\ref{sec:extending} explains how to modify and expand the
-possibilities by producing new summary-, visual- and check functions.
+arguments in \code{clean()}, and above we have shown examples of how
+to tweak them to modify the data cleaning outputs. However, if the
+dataset at hand requires new, additional checks, then more control is
+needed, and Section~\ref{sec:extending} explains how to modify and
+expand the possibilities by producing new summary, visual, and check
+functions.
 
@@
 While overview documents are great for presenting and documenting the
-data cleaning checks, it may be useful to be able to work more interactively
-through the data cleaning process as well. Aside from the \code{clean()} function
-presented above, \pkg{dataMaid} also provides more standard \proglang{R} interactive tools, such as
+data cleaning checks, it may be useful to be able to work more
+interactively through the data cleaning process. Aside from
+the \code{clean()} function presented above, \pkg{dataMaid} also
+provides more standard \proglang{R} interactive tools, such as
 functions that print results to the console or returns the information
 as an object for later use. This section describes how to use the
-functions \code{check()}, \code{summarize()} and \code{visualize()} to work
-interactively with \pkg{dataMaid}.
+functions \code{check()}, \code{summarize()} and \code{visualize()} to
+work interactively with \pkg{dataMaid}.
 
 \subsection{Data cleaning by hand: An example}
-Let's say we wish to look further into a certain variable from
+Assume that we wish to look further into a certain variable from
 \code{toyData}, namely \code{var2}. The data cleaning summary found some
@@
 \code{XXXChecks} argument (e.g., \code{numericChecks}, \code{factorChecks},
-etc. as described in Section~\ref{sec:usingdataMaid}) to provide a
+etc., as described in Section~\ref{sec:usingdataMaid}) to provide a
 vector of the check functions that should be applied. Recall that
@@
 causing the issue are still obtainable in the element
-\code{problemValues}. If we for instance decide that the values \code{999}
+\code{problemValues}. If we decide that the values \code{999}
 and \code{NaN} in \code{var2} are in fact miscoded missing values, we can
@@
 
-Note that the \code{summarize()}, \code{check()} and \code{visualize()} functions are also available interactively for full datasets, by calling e.g., \code{summarize(toyData)}.
+Note that the \code{summarize()}, \code{check()} and \code{visualize()} functions are also available interactively for full datasets by calling e.g., \code{summarize(toyData)}.
 However, this produces an extensive amount of output in the console, and therefore, we generally do not recommend it, unless working with very small datasets or subsets of datasets.
@@
 customizable and \code{clean()} is mainly a tool for formatting the
-results from various checking-, summary- and visualization
+results from various checking, summary,  and visualization
 functions. Thus, the actual work underlying a \pkg{dataMaid} output file
@@
 
-This section consists of two parts. First, we describe how customized \code{summaryFunction}s,
-\code{visualFunction}s and \code{checkFunctions} can be made, including an introduction to a
-few convenient tools available in \pkg{dataMaid} for aiding SVC-function construction.
-Table \ref{table.functionTypes} serves as an overview of the internals of these three
-central function types. Secondly, we turn to a
-worked example of how to use custom made functions in practice. Here,
-four new SVC functions are defined and used, both interactively and in \code{clean()}.
+This section consists of two parts. First, we describe how customized
+\code{summaryFunction}s, \code{visualFunction}s and
+\code{checkFunctions} can be made, including an introduction to a few
+convenient tools available in \pkg{dataMaid} for aiding SVC-function
+construction.  Table \ref{table.functionTypes} serves as an overview
+of the internals of these three central function types. Secondly, we
+turn to a worked example of how to use custom made functions in
+practice; Four new SVC functions are defined and used, both
+interactively and in \code{clean()}.
 
@@
 \label{sec:functionTemplates}
-In order to construct a summary-, visual or check function, one needs
+In order to construct a summary, visual or check function, one needs
 to create a new function with a specific structure.  This can be done with
@@
 
-XXX indentation
 <<>>=

### 2017-02-01 latex/article_vol2.tex
Fixed identifyMissing

@@
 and \code{defaultXXXSummaries()} functions, where \code{XXX} again refers to
-the variable type.
-
-For example, the default summaries being used for a factor variable is
+the variable type. For example, the default summaries being used for a factor variable is
+
 \begin{Schunk}
@@
  $ message      : chr "The following suspected missing value codes enter as regular values: \\\"999\\\", \\\"NaN\\\"."
- $ problemValues: NULL
+ $ problemValues: num [1:2] 999 NaN
  - attr(*, "class")= chr "checkResult"
@@
 \begin{Soutput}
-The following suspected missing value codes enter as regular values: 999, NaN.
+No problems found.
 \end{Soutput}
@@
 [1,] "Variable type"           "numeric"    
-[2,] "Number of missing obs."  "3 (20 %)"
-[3,] "Number of unique values" "8"       
-[4,] "Median"                  "4.5"     
-[5,] "1st and 3rd quartiles"   "1.75; 6" 
-[6,] "Min. and max."           "1; 999"  
+[2,] "Number of missing obs."  "4 (26.67 %)"
+[3,] "Number of unique values" "6"          
+[4,] "Median"                  "4"          
+[5,] "1st and 3rd quartiles"   "1.5; 6"     
+[6,] "Min. and max."           "1; 82"      
 \end{Soutput}
@@
      Feature         Result 
-[1,] "Median"        "4.5"   
-[2,] "Min. and max." "1; 999"
+[1,] "Median"        "4"    
+[2,] "Min. and max." "1; 82"
 \end{Soutput}
@@
 
-XXX indentation
 \begin{Schunk}

### 2017-02-01 R/identifyMissing.R

@@ -229,7 +229,7 @@ identifyMissingNI <- function(v, nMax, maxDecimals) {
   allProblemOcc <- c(missNinesOcc, missEightsOcc, missNaNOcc)
 
   if (length(allProblemOcc) > 0) {
-    outProblemValues <- problemValues
+    outProblemValues <- allProblemOcc
     problemValues <- round(allProblemOcc, maxDecimals)
     problem <- TRUE
   }

### 2017-02-01 latex/article_vol2.Rnw
Submitted version

@@
 computer without further ado, while other types of errors need an analytic
-context in order to shine through. But even in the latter case, well-designed
-software can aid the process tremendously by giving the human investigator the
+context in order to be identified. Even in this latter case, well-designed
+software can aid the process tremendously by providing the human
+investigator with the
 information needed for identifying issues.
@@
 different stages of the pre-analysis work, most of which are not
-really data cleaning.  \pkg{janitor} provides tools for data loading
-with a particular emphasis on the challenges of getting neat
-\code{data.frame}s from Microsoft Excel data files. \pkg{plyr},
-\pkg{data.table} and \pkg{DataCombine} go a few steps further by
-providing a wide array of extremely powerful tools for data wrangling,
-including a number of particularly useful functions for merging and
-working with very large datasets. When it comes to actual data
-cleaning, however, the options are fewer.  \pkg{validate} (and the
-similar packages \pkg{editrules} and \pkg{deducorrect} from the same
-authors) does offer a few tools for identifying and auto-correcting
-errors in a dataset, but, as the name suggests, the focus is on
-internal validity rather than general data cleaning. In practice, this
-means that quite elegant tools for e.g. linear restraints among the
-variables in a dataset can be applied, but e.g. looking for
-potentially miscoded missing values is not really feasible. The main
-difference between these two challenges is the direction in which the
-data is inspected: While linear constraints work row-wise with no
-ambiguity, determining whether or not something is a miscoded missing
-value often requires knowledge about the full variable (e.g. range or
-data type), and thus it should be performed
+really data cleaning.  \pkg{janitor} provides tools for data import
+with a particular emphasis on the challenges of getting neat data
+frames from Microsoft Excel data files. \pkg{plyr}, \pkg{data.table}
+and \pkg{DataCombine} go a few steps further by providing a wide array
+of extremely powerful tools for data wrangling, including a number of
+particularly useful functions for merging and working with very large
+datasets. When it comes to actual data cleaning, however, the options
+are fewer.  \pkg{validate} (and the similar packages \pkg{editrules}
+\citep{editrules} and \pkg{deducorrect} \citep{deducorrect} from the
+same authors) does offer a few tools for identifying and
+auto-correcting errors in a dataset, but, as the name suggests, the
+focus is on internal validity rather than general data cleaning. In
+practice, this means that quite elegant tools for, e.g., linear
+restraints among the variables in a dataset can be applied, but
+looking for potentially miscoded missing values is not really
+feasible. The main difference between these two challenges is the
+direction in which the data is inspected: While linear constraints
+work row-wise with no ambiguity, determining whether or not something
+is a miscoded missing value often requires knowledge about the full
+variable (e.g. range or data type), and thus it should be performed
 column-wise. \pkg{validate} does not allow for user-defined extensions
 of the latter type, thereby limiting its data cleaning potential.
-Moreover, no tools have been made for documenting the data cleaning
-process.  As automatic cleaning options are also supplied in
-\pkg{validate}, we find quite a dangerous cocktail where all power is
-given to the the computer with no supervision, neither in real time
-nor subsequently.
+Automatic data correction functions are also provided by
+\pkg{validate} which we consider to be quite a dangerous cocktail: all
+power is given to the the computer with no human supervision, and
+investigators are less likely to make an active, case-specific choice
+regarding the handling of the potential errors.  Finally, no tools
+have been made to easily document exactly which checks and preliminary
+results were used in the data cleaning process.
 
@@
 \hline
-\vspace{0pt} Input (required) & \code{v} - a variable vector \newline \code{...}  &  \code{v} - a variable vector \newline \code{vnam} - the variable name (as character string) \newline \code{doEval} - a logical (\code{TRUE}/\code{FALSE}) controlling the output type of the function & \code{v} - a variable vector \newline \code{nMax} - an integer (or \code{Inf}), controlling how many problematic values are printed, if relevant \newline \code{...}  \\
-Input (optional) &  \code{maxDecimals} - number of decimals printed in outputted numerical values.  & - &  \code{maxDecimals}  - number of decimals printed in outputted numerical values.  \\
+\vspace{0pt} Input (required) & \code{v} --- a variable vector.
+                                \newline \code{...} --- additional
+                                arguments passed to the function.
+&  \code{v} --- a variable vector.
+\newline \code{vnam} --- the variable name (as character string).
+\newline \code{doEval} --- a logical (\code{TRUE}/\code{FALSE}) controlling the output type of the function.
+& \code{v} - a variable vector \newline \code{nMax} - an integer (or \code{Inf}), controlling how many problematic values are printed, if relevant \newline \code{...} --- additional arguments passed to the function.
+\\
+Input (optional) &  \code{maxDecimals} --- number of decimals printed in outputted numerical values.  & --- &  \code{maxDecimals}  --- number of decimals printed in outputted numerical values.  \\
 Purpose & Describe some aspect of the variable, e.g., a central value, its dispersion or level of missingness.
@@
 \\
-Output (required) & A list with entries: \newline \code{\$feature} - a label for the summary value (as character string); \newline \code{\$result} - the result of the summary (as character string) & A character string with \proglang{R} code for producing a plot.
+Output (required) & A list with entries: \newline \code{\$feature} --- a label for the summary value (as character string); \newline \code{\$result} --- the result of the summary (as character string).
+& A character string with \proglang{R} code for producing a plot.
 This code should be standalone, i.e., should include the data if necessary.
-& A list with entries: \newline \code{\$problem} - a logical identifying whether an issue was found; \newline \code{\$message} - a character string (possibly empty) decribing the issue that was found, properly escaped and ready for use in \code{rmarkdown} \\
+& A list with entries: \newline \code{\$problem} --- a logical identifying whether an issue was found; \newline \code{\$message} --- a character string (possibly empty) decribing the issue that was found, properly escaped and ready for use with \code{rmarkdown}.
+\\
 Output (recommended) & A \code{summaryResult} object, i.e., an attributed list with entries \code{\$feature}, \code{\$result} and \code{\$value}, the latter being the values from \code{\$result} in their original format).
@@
 \caption{Reference information for creating new functions to be used
-  as part of the \mbox{summarize-,} visualize-, and check steps. The three
+  as part of the summarize, visualize, and check steps. The three
   columns correspond to each of the three function types.
@@
 \end{Verbatim}
-The last function called here, \code{summaryResult()}, changes the class
-of the output, thereby making a \code{print()} method available for it.
-Note that \code{v} is a vector and that \code{res} should be either a
-character string or something that will be printed as one. In other
-words, e.g., integers are allowed, but matrices are not. Though a lot
-of different things can go into the \code{summaryFunction} template,
-we recommend only using it for summarizing the features of a variable,
-and leaving tests and checks for the \code{checkFunction}s (presented
-below).
+The last function called here, \code{summaryResult()}, changes the
+class of the output, thereby making a \code{print()} method available
+for it.  Note that \code{v} is an input vector and that \code{res}
+should be either a character string or something that will be printed
+as one. In other words an integer would be allowed, but a matrix will
+not. Though a lot of different things can go into the
+\code{summaryFunction} template, we recommend only using it for
+summarizing the features of a variable, and leaving tests and checks
+for the \code{checkFunction}s (presented below).
 
@@
 made \code{mySummaryFunction()} in \code{clean()}, but we recommend
-furthermore adding the new function to the overview of all summary
+adding the new function to the overview of all summary
 functions by converting it to a proper \code{summaryFunction}
@@
 user-defined function is intended to be applied to, also stored as an
-attribute). In other words, a call on the following form should be made:
+attribute). In other words, a call following the template below should be made:
 \begin{Verbatim}
@@
 % the two attributes of a \code{summaryFunction}.
-If the \code{description} argument is left unspecified, the name of the
-function (which in this case is \code{"mySummaryFunction"}) will be filled in and
-stored under the \code{description} attribute. What happens if the \code{classes} argument is
-not specified depends on the type of \code{summaryFunction}. If
-\code{mySummaryFunction} is constructed as an \code{S3} generic function with associated
-methods, the call to \code{summaryFunction()} will automatically produce
-a vector of the names of the classes for which the function can be
+If the \code{description} argument is left unspecified, the name of
+the function (which in this case is \code{"mySummaryFunction"}) will
+be filled in and used as the \code{description} attribute. What
+happens if the \code{classes} argument is not specified depends on the
+type of \code{summaryFunction}. If \code{mySummaryFunction} is
+constructed as an \code{S3} generic function with associated methods,
+the call to \code{summaryFunction()} will automatically produce a
+vector of the names of the classes for which the function can be
 called. If \code{mySummaryFunction()} is not an \code{S3} generic and
-\code{classes} is left unspecified, the attribute will simply be
-left empty. Note that the helper function \code{allClasses()} might be useful
-for filling out the \code{classes} argument, as it simply lists all
-available classes in \pkg{dataMaid}:
+\code{classes} is left unspecified, the attribute will simply be left
+empty. Note that the helper function \code{allClasses()} might be
+useful for filling out the \code{classes} argument, as it simply lists
+all available classes in \pkg{dataMaid}:
 
@@
 code for plots in order for \code{clean()} to build standalone
-\pkg{rmarkdown} files. We recommend using the following structure:
+\pkg{rmarkdown} files. We recommend using the following structure
+(again shown as pseudo code):
 \begin{Verbatim}
@@
 are mandatory, so they must be left as is, even if they are not in
-use (c.f. Table~\ref{table.functionTypes}).
+use (c.f., Table~\ref{table.functionTypes}).
 
@@
 the \code{checkFunction}. These are the functions that flag issues in
-the data in the check step and control the overall flow of the data
-cleaning process in the precheck stage. A \code{checkFunction} follows
+the data as part of the check step and/or control the overall flow of the data
+cleaning process in the pre-check stage. A \code{checkFunction} follows
 one of two overall structures, depending on the type of check. Either,
@@
 e.g., \code{identifyMissing()} does), or it performs a check concerning the
-variable as a whole (e.g., the functions used for prechecks and the
+variable as a whole (e.g., the functions used for pre-checks and the
 function \code{identifyNums()}). We present templates for both types of
@@
 If problematic values are to be identified, the template from above
-should be expanded to follow a slightly more complicated structure:
+should be expanded to follow a slightly more complicated structure
+(again shown as pseudo-code):
 
@@
 values were left out (if any).  Note that printing quotes in
-\pkg{rmarkdown} requires an extensive amount of character escaping, so
+\proglang{R} markdown requires an extensive amount of character escaping, so
 opting for \code{messageGenerator()} really is the easiest solution.
@@
 myProbValCheckFunction <- function(v, nMax, maxDecimals, ...) {
-  ...
+  [... more lines of code here ...]
   problemStatus <- list(problem = problem,
@@
   }
-  ...
+  [... more lines of code here ...]
 }
@@
 will still appear in their original format under the entry \code{\$problemValues} %$
-in the outputted \code{checkResult}.
+in the returned object.
 
@@
 
-We will now build four new functions and show both how they can be
+As an example, we create four new functions and show both how they can be
 used interactively and how they can be integrated with the \code{clean()}
@@
 \begin{description}
-\item[\code{isID()}] A new \code{checkFunction} intended for use in the precheck-stage.
+\item[\code{isID()}] A new \code{checkFunction} intended for use in the pre-check-stage.
 This function checks whether a variable consists exclusively of long ($> 10$ characters/digits) entries that are all of equal length, as this might be personal identification codes that we do not wish to print out in the data summary.
@@
 
-\subsubsection{isID --- a new checkFunction without problem values}
-First, let's define the \code{isID()} function. As this function is not
-supposed to list problematic values, it falls within
-the category of \code{checkFunction}s represented by
-\code{myFullVarCheckFunction()} in the above. We do not particularly wish
-to use this function interactively, so we will stick to the minimal
+\subsubsection[isID --- a new checkFunction without problem values]{\code{isID} --- a new \code{checkFunction} without problem values}
+First, let's define the \code{isID()} function. As this function is
+not supposed to list problematic values, it falls within the category
+of \code{checkFunction}s represented by
+\code{myFullVarCheckFunction()} above. We do not particularly wish to
+use this function interactively, so we will stick to the minimal
 requirements of a \code{checkFunction} used in \code{check()} (see
@@
   out <- list(problem = FALSE, message = "")
-  # Should work for all but logical and Date variables
   if (class(v) %in% setdiff(allClasses(), c("logical", "Date"))) {
@@
     lengths <- c(nchar(v))
-    # Check that all lengths are the same and greater than 10 characters
     if (all(lengths > 10) & length(unique(lengths)) == 1) {
@@
 This is essentially all we need to do in order to include this
-function as a precheck-function in \code{clean()}, so we will leave it as
+function as a pre-check-function in \code{clean()}, so we will leave it as
 is and move on to the next function, namely \code{mosaicVisual()}.
 
-\subsubsection{mosiacVisual --- a new visualFunction}
-\proglang{R} contains a function, \code{mosiacplot()}, which produces
+\subsubsection[mosaicVisual --- a new visualFunction]{\code{mosaicVisual} --- a new \code{visualFunction}}
+\proglang{R} provides a function, \code{mosaicplot()}, which produces
 mosaic plots. We intend to use this existing high-level plotting
@@
   thisCall <- call("mosaicplot", table(v), main = vnam, xlab = "")
-  if (doEval) {                     # Return the graphics
+  if (doEval) {
     return(eval(thisCall))
-  } else return(deparse(thisCall))  # Else return the code for the call
+  } else return(deparse(thisCall))
 }
@@
 appear in an \code{allVisualFunctions()} call. In order to get this
-functionality, we need to change its object class. This can be done by
+functionality we need to change its object class. This can be done by
 writing
@@
 
-\subsubsection{countZeros --- a new summaryFunction}
+\subsubsection[countZeros --- a new summaryFunction]{\code{countZeros} --- a new \code{summaryFunction}}
 This \code{summaryFunction} is defined in the following lines of code:
@@
 As this function computes an integer (the number of zeros),
-there is no difference between the entries \code{\$result} and
+there is no difference between the elements \code{\$result} and
 \code{\$value}. If, on the other hand, the result had been a character
@@
 
-\subsubsection{identifyColons --- a new checkFunction with problem values}
+\subsubsection[identifyColons --- a new checkFunction with problem values]{\code{identifyColons} --- a new \code{checkFunction} with problem values}
 The last function mentioned above is \code{identifyColons()}. We
@@
 
-Note, that for \code{checkFunction}s, the function description will
-appear in the document produced by \code{clean()} (in the \textit{Data
-  cleaning summary} section of the output), so for this type of
-function, the change of class is not only done for the sake of the
+Note, that for \code{checkFunction}s, the description will appear in
+the document produced by \code{clean()} (in the \textit{Data cleaning
+  summary} section of the output), so for this type of function, the
+change of class is not only done for the sake of the
 \code{allCheckFunctions()} output.
 
-\subsubsection{Calling the new summarize/visualize/check functions from clean()}
+\subsubsection[Calling the new summarize/visualize/check functions from clean()]{Calling the new summarize/visualize/check functions from \code{clean()}}
 Now, we are ready to use these new functions in a \code{clean()}
@@
 can be used for problem-flagging. Lastly, we discuss how the
-\pkg{dataMaid} output document can be included in other \pkg{rmarkdown}
+\pkg{dataMaid} output document can be included in other \proglang{R} markdown
 documents as a way to produce clear and concise documentation of a
@@
 If the dataset becomes very large, the standard use of \code{clean()}
-outlined above might not be ideal. If there is a vast number of
-variables, creation of the \pkg{rmarkdown} document might be quite
+outlined above might not be ideal. If there is a large number of
+variables, creation of the \proglang{R} markdown document might be quite
 slow, while a large number of observations will generally affect the
@@
 while newer 64-bit versions of Windows cannot access virtual memory. Therefore,
-memory allocation issues becomes even more pressing when using this operating system.
-
-In this section, we give a few
-practical examples of ways to deal with large data, while wishing to
-still produce (potentially very long) data cleaning overview
-documents. Note that the interactive tools of \pkg{dataMaid} can be used
-as usual or sequentially in small subsets of the large dataset, if no
-such overview documents are needed.
+memory allocation issues becomes even more pressing under Windows.
+
+Below we give a few practical examples of ways to deal with large
+data, while wishing to still produce (potentially very long) data
+cleaning overview documents. Note that the interactive tools of
+\pkg{dataMaid} can be used as usual or sequentially in small subsets
+of the large dataset, if no overview documents are needed.
 
@@
 Of course, even less heavy plots might be achieved by writing new
-\code{visualFunction}s, using the guidelines from section
+\code{visualFunction}s, using the guidelines from Section
 \ref{sec:functionTemplates}. For instance, a future extension of
@@
 is simply splitting the two steps performed
-by \code{clean()}, namely producing the \pkg{rmarkdown} file and rendering it
+by \code{clean()}, namely producing the \proglang{R} markdown file and rendering it
 afterwards. If the \code{rmarkdown} file is very long, as it will
@@
 wastes precious memory capacities. Therefore, we advice users to
-instead split the two steps. This can be done in the following manner:
+instead split the two steps as shown in the following.
 
@@
 This also deals with the fact that \pkg{dataMaid} can produce
-\pkg{rmarkdown} files that supersedes the upper size limit of code editors, e.g.,
+\proglang{R} markdown files that supersedes the upper size limit of code editors, e.g.,
 RStudio, which currently has an editor file size limit of 5MB. %\hl{dette giver
@@
 If the data is large, but memory issues and computation time are less
-of an issue than the human time it takes to look through the data
+of an issue than the time it takes a human to look through the data
 cleaning document, a viable solution might be not to include all
-information about all variables. Or even for more reasonably sized
-datasets, sometimes a brief overview of the most pressing issues can
-be useful. This can be achieved by using the \code{onlyProblematic}
-argument in \code{clean()}. By specifying \code{onlyProblematic = TRUE},
-only variables that raise a flag in the checking steps will be
-summarized and visualized. But perhaps we are not even interested in
-obtaining general information about these variables, but only in
-getting a quick overview of the problems they might have. This can be
-done by also controlling the \code{mode} argument:
+information about all variables. % Or even for more reasonably sized
+% datasets, sometimes a brief overview of the most pressing issues can
+% be useful.
+This can be achieved by using the \code{onlyProblematic} argument in
+\code{clean()}. By specifying \code{onlyProblematic = TRUE}, only
+variables that raise a flag in the checking steps will be summarized
+and visualized. But perhaps we are not even interested in obtaining
+general information about these variables, but only in getting a quick
+overview of the problems they might have. This is obtained by
+controlling the \code{mode} argument:
 
@@
 generated by also leaving out the checking results --- then
-\code{clean()} essentially just produces a list of the variable names
+\code{clean()} essentially just produces a document with a list of the variable names
 that should be investigated further:
@@
 Sometimes, a \pkg{dataMaid} document might be a useful addition to a
-more general overview document, including also for example pairwise
-association plots, time series plots, or exploratory analysis
-results. To this end, it is possible to produce a \pkg{dataMaid}
-document that can readily be included in other \pkg{rmarkdown}
-files. This is done by using the \code{standAlone} argument in \code{clean()},
-which removes the preamble from the outputted \pkg{rmarkdown}
-file. Please note, that it is still necessary to indicate which
-\code{rmarkdown} type is being created; the pdf and html \pkg{rmarkdown}
-styles are unfortunately not identical.
+more general overview document, including additional information such
+as pairwise association plots, time series plots, or exploratory
+analysis results.  \pkg{dataMaid} can produce a document to be
+included in other \proglang{R} markdown files by setting the
+\code{standAlone} argument in \code{clean()} to remove the preamble
+from the output \proglang{R} markdown file. Note that it is still
+necessary to indicate which \proglang{R} markdown output type is
+created; the pdf and html \proglang{R} markdown styles are
+unfortunately not identical.
 
@@
 rendered to either of these two file types, we recommend setting
-\code{twoCols = FALSE} and \code{output = html} in \code{clean()}, thereby
+\code{twoCols = FALSE} and \code{output = "html"} in \code{clean()}, thereby
 essentially removing almost all output type specific formatting code
-from the generated \pkg{rmarkdown} file.
+from the generated \proglang{R} markdown file.
 
 On the other hand, if a pdf document is to be produced, a few extra
-lines need to be added to the preamble of the master \pkg{rmarkdown}
+lines need to be added to the preamble of the master \proglang{R} markdown
 document --- otherwise, the two-column layout code will produce an
@@
 
-and the more minimal, html-style \pkg{rmarkdown} file described above can be produced using
+and the more minimal, html-style \proglang{R} markdown file described above can be produced using
 
@@
 
-Note that with the latter option, no special YAML preamble is needed in the \pkg{rmarkdown} document.
+Note that with the latter option, no special YAML preamble is needed in the \proglang{R} markdown document.
 % \hl{don't end on code.}

### 2017-02-01 latex/article_vol2.tex
Submitted version

@@
  $ problem      : logi TRUE
- $ message      : chr "The following suspected missing value codes enter as regular values: \\\"999\\\", \\\"NaN\\\"."
+ $ message      : chr "The following suspected missing value codes
+    enter as regular values: \\\"999\\\", \\\"NaN\\\"."
  $ problemValues: num [1:2] 999 NaN
@@
 +    out <- list(problem = FALSE, message = "")
-+    # Should work for all but logical and Date variables
 +    if (class(v) %in% setdiff(allClasses(), c("logical", "Date"))) {
@@
 +      lengths <- c(nchar(v))
-+      # Check that all lengths are the same and greater than 10 characters
 +      if (all(lengths > 10) & length(unique(lengths)) == 1) {
@@
 +    thisCall <- call("mosaicplot", table(v), main = vnam, xlab = "")
-+    if (doEval) {                     # Return the graphics
++    if (doEval) {
 +      return(eval(thisCall))
-+    } else return(deparse(thisCall))  # Else return the code for the call
++    } else return(deparse(thisCall))
 +  }

### 2017-02-08 R/messageGenerator.R
More character escaping

@@ -144,7 +144,27 @@ printProblemValues <- function(problemValues, nMax = Inf) {
 #NOTE: 3 slashes escapes the espaced string [\"] such that it is printed correctly
 #(and not interpreted) in markdown.
 escapeRMDStyle <- function(string) {
-  sub1 <- gsub("\\", "\\\\", string, fixed = TRUE)
-  sub2 <- gsub("<", "\\<", sub1, fixed = TRUE)
-  gsub(">", "\\>", sub2, fixed = TRUE)
+  string <- gsub("\\", "\\\\", string, fixed = TRUE)
+  string <- gsub("\a", "\\\\\\\\a", string)
+  string <- gsub("\f", "\\\\\\\\f", string)
+  string <- gsub("\n", "\\\\\\\\n", string)
+  string <- gsub("\r", "\\\\\\\\r", string)
+  string <- gsub("\t", "\\\\\\\\t", string)
+  string <- gsub("<", "\\<", string, fixed = TRUE)
+  gsub(">", "\\>", string, fixed = TRUE)
+}
+
+
+#Helper function for escaping characters such that they are 
+#printed as is.
+#All currently implemented special characters commencing 
+#with "\" are included (as of 02-08-2017)
+escapeRStyle <- function(string) {
+  string <- gsub("\\", "\\\\", string, fixed = TRUE)
+  string <- gsub("\a", "\\\\a", string)
+  string <- gsub("\f", "\\\\f", string)
+  string <- gsub("\n", "\\\\n", string)
+  string <- gsub("\r", "\\\\r", string)
+  string <- gsub("\t", "\\\\t", string)
+  string
 }

### 2017-02-08 R/standardVisual.R

@@ -103,17 +103,5 @@ standardVisualD <- function(v, vnam, doEval = TRUE) {
 
 
 
-#Helper function for escaping characters such that they are 
-#printed as is in the plot axis ticks.
-#All currently implemented special characters commencing 
-#with "\" are included (as of 02-08-2017)
-escapeRStyle <- function(string) {
-  string <- gsub("\\", "\\\\", string, fixed = TRUE)
-  string <- gsub("\a", "\\\\a", string)
-  string <- gsub("\f", "\\\\f", string)
-  string <- gsub("\n", "\\\\n", string)
-  string <- gsub("\r", "\\\\r", string)
-  string <- gsub("\t", "\\\\t", string)
-  string
-}
+
 

### 2017-02-08 latex/article_vol2.tex

@@
 specific field and the context of the available data. The
-\pkg{dataMaid} package \citep{dataMaid} tries to address this by
+\pkg{dataMaid} package \citep{dataMaid} presented here tries to address this by
 providing a framework that both allows for extendable, systematic,

### 2017-10-18 latex/article_vol3.tex
New version of article, halfway done.

--- 2017-02-08 latex/article_vol2.tex
@@
 %% almost as usual
-\author{Anne H. Petersen\\Biostatistics\\Department of Public
+\author{Anne Helby Petersen\\Biostatistics\\Department of Public
   Health\\University of Copenhagen \And Claus Thorn Ekstr\o m\\Biostatistics\\Department of Public
   Health\\University of Copenhagen}
-\title{\pkg{dataMaid}: your assistant for data cleaning in \proglang{R}}
+\title{\pkg{dataMaid}: your assistant for data documentation in \proglang{R}}
 
 %% for pretty printing and a nice hypersummary also set:
-\Plainauthor{Anne H. Petersen, Claus Thorn Ekstr\o m} %% comma-separated
-\Plaintitle{{dataMaid}: your assistant for data cleaning in R} %% without formatting
-\Shorttitle{\pkg{dataMaid}: your data cleaning personal assistant in R} %% a short title (if necessary)
+\Plainauthor{Anne Helby Petersen, Claus Thorn Ekstr\o m} %% comma-separated
+\Plaintitle{{dataMaid}: your assistant for data documentation in R} %% without formatting
+\Shorttitle{\pkg{dataMaid}: your assistant for data documentation in R} %% a short title (if necessary)
 
 %% an abstract and keywords
-\Abstract{Data cleaning and -validation are the first steps in any
+\Abstract{Data cleaning and -validation are important steps in any
   data analysis, as the validity of the conclusions from the analysis
@@
   assessment tools that can be applied to a dataset in order to
-  identify potential problems in its variables. The results can be
+  identify potential problems in its variables. The results are 
   presented in an auto-generated, non-technical, stand-alone overview
@@
   more typical \proglang{R} tools for interactive data quality
-  assessment and -cleaning, where the data inspections all live and
-  die within the \proglang{R} console.
+  assessment and -cleaning, where the data inspections are executed directly 
+  in the \proglang{R} console.
 
@@
 }
-\Keywords{data cleaning, quality control, \proglang{R}}
-\Plainkeywords{data cleaning, quality control, R} %% without formatting
+\Keywords{data cleaning, quality control, \proglang{R}, data documentation}
+\Plainkeywords{data cleaning, quality control, R, data documentation} %% without formatting
 %% at least one keyword must be supplied
@@
 analysts find themselves spending a large portion of their time on
-data cleaning and data wrangling. While a computer should never
+data cleaning and data wrangling. While a computer should generally not
 make unsupervised decisions on what should be done to potential
@@
 software can aid the process tremendously by providing the human
-investigator with the
-information needed for identifying issues.
-
-A number of \proglang{R} packages made for such pre-analysis tools are
-already available, including \pkg{janitor} \citep{janitor}, \pkg{plyr}
-\citep{plyr}, \pkg{data.table} \citep{data.table}, \pkg{DataCombine}
-\citep{DataCombine} , \pkg{validate} \citep{validate}, and
+investigator with the information needed for identifying issues.
+
+But even when tools are available for identifying problems in a dataset, the activity of data cleaning still suffers from a challenge that has recieved increasing attention in the scientitic communities in the later years: Data cleaning is not very straight forward to document and therefore, reproducibility suffers.
+\pkg{dataMaid}'s most central purpose is to facilitate a data cleaning workflow where documentation is thoroughly integrated rather than a secondary add-on.
+This is accomplished by centering the data cleaning process around auto-generated data overview reports that summarize the contents of the current version of a dataset and flags potential problems that could be candidates for the next bullet on the cleaning agenda.
+
+A number of \proglang{R} packages made for other pre-analysis tools are
+already available, including \pkg{janitor} \citep{janitor}, \pkg{dplyr}
+\citep{dplyr}, \pkg{tidyr}\citep{tidyr}, \pkg{data.table} \citep{data.table}, \pkg{DataCombine}
+\citep{DataCombine} , \pkg{validate} \citep{validate}, \pkg{assertr} \citep{assertr}, and
 \pkg{DataExplorer} \citep{DataExplorer}. These packages focus on
@@
 with a particular emphasis on the challenges of getting neat data
-frames from Microsoft Excel data files. \pkg{plyr}, \pkg{data.table}
+frames from Microsoft Excel data files. \pkg{dplyr}, \pkg{tidyr}, \pkg{data.table}
 and \pkg{DataCombine} go a few steps further by providing a wide array
@@
 direction in which the data is inspected: While linear constraints
-work row-wise with no ambiguity, determining whether or not something
+work observation-wise with no ambiguity, determining whether or not something
 is a miscoded missing value often requires knowledge about the full
 variable (e.g. range or data type), and thus it should be performed
-column-wise. \pkg{validate} does not allow for user-defined extensions
+variable-wise. \pkg{validate} does not allow for user-defined extensions
 of the latter type, thereby limiting its data cleaning potential.
@@
 have been made to easily document exactly which checks and preliminary
-results were used in the data cleaning process.
+results were used in the data cleaning process. \hl{Something about \pkg{assertr} here.}
 
@@
 
-The \pkg{dataMaid} approach to data cleaning and -quality assessment is
+The \pkg{dataMaid} approach to data cleaning, -quality assessment and -documentation is
 governed by two fundamental paradigms. First of all, there is no need
@@
 paradigm, namely the \code{clean()} function, which generates data
-cleaning overview documents. In the \pkg{dataMaid} package, we have
-provided a number of default cleaning steps that cover the data
+overview documents. In the \pkg{dataMaid} package, we have
+provided a number of default checks that cover the data
 cleaning challenges, we find to be most common. Next, in Section
@@
 
-\section{Creating a data cleaning overview}
+\section{Creating a data overview report}
 \label{sec:usingdataMaid}
 
-The \code{clean()} function is the primary workhorse of \pkg{dataMaid} and
+The \code{makeDataReport()} function is the primary workhorse of \pkg{dataMaid} and
 this is the only function needed if a standard battery
-of tests are used to generate data cleaning summaries. The data
-cleaning summary itself is an overview document, intended for reading
+of tests are used to generate data reports. The data report itself is an overview document, intended for reading
 by humans, in either pdf or html format. Appendix \ref{sec:appendix1}
-provides an example of a data cleaning output document, produced by
-calling \code{clean()} on the dataset \code{toyData} available in
-\pkg{dataMaid}. The first two pages of this data cleaning output are
+provides an example of a data report, produced by
+calling \code{makeDataReport()} on the dataset \code{toyData} available in
+\pkg{dataMaid}. The first two pages of this data report are
 shown in Figure~\ref{fig:example1}. \code{toyData} is a very
-small ($15$ rows of $6$ variables), artificial dataset which was created to
+small ($15$ observations of $6$ variables), artificial dataset which was created to
 illustrate the main capabilities of \pkg{dataMaid}. The following
-commands load the dataset and produce the cleaning output:
+commands load the dataset and produce the report:
 
@@
 \end{center}
-\caption{Example output from running \code{clean()} on the \code{toyData}
+\caption{Example output from running \code{makeDataReport()} on the \code{toyData}
   dataset. First a summary of the full dataset is given and then
@@
 \begin{Sinput}
-R> clean(toyData)
+R> makeDataReport(toyData)
 \end{Sinput}
@@
 opened for immediate inspection. Turning to Figure~\ref{fig:example1},
-we see that such a data cleaning output document consists of two
+we see that such a data report consists of three
 parts. First, an overview of what was done is presented under the
-title \textit{Data cleaning summary}. Secondly, each variable in the
+title \textit{Data report overview}. Secondly, an index listing each variable along with an indication of whether it was found to be problematic or not is provided. Thirdly, each variable in the
 dataset is presented in turn using (up to) three tools in the
@@
 list of flagged issues, if any. For instance, in the \code{numeric}-type variable
-\code{var2} from \code{toyData}, \code{clean()} has identified two values that
+\code{var2} from \code{toyData}, \code{makeDataReport()} has identified two values that
 are suspected to be miscoded missing values (\code{999} and \code{NaN}),
@@
 be investigated more carefully.
-%We can then return to Part
-%1, the data cleaning summary, and inspect what sorts of checks were
-%performed on this variable. \hl{hm, delete the last sentence? or use
-%  different variable? Only these exact two checks were performed}.
-
-Though the \code{clean()} function is very easy to use, it should not be
-mistaken to be inflexible. A large number of function arguments allows
-for the cleaning overview document to be molded according to the
-user's needs. The most commonly used arguments are summarized in
+
+Though the \code{makeDataReport()} function is very easy to use, it should not be
+mistaken to be inflexible: By using the optional arguments of the function, 
+both the contents and the look of the data report can be molded 
+according to the user's needs. The most commonly used arguments are summarized in
 Table~\ref{table.cleanFormals} and they are grouped according to the
-part of the cleaning process they influence. In order to understand
-this distinction, a glimpse of the inner structure of \code{clean()} is
+part of the data assesment and report generation they influence. In order to understand
+this distinction, a glimpse of the inner structure of \code{makeDataReport()} is
 shown in Figure~\ref{figure:cleanStructure}. Below, we present a few
 examples on how to use the arguments from Table \ref{table.cleanFormals}
- to influence the output of a \code{clean()} call.
+ to influence the output of a \code{makeDataReport()} call.
 
@@
 \quad \code{onlyProblematic} & Should only variables flagged as problematic be included in the \textit{Variable list}? & \code{FALSE} \\
-\quad \code{listChecks} & Should an overview of what checks were performed be listed in the \textit{Data cleaning summary}? &  \code{TRUE} \\
+\quad \code{listChecks} & Should an overview of what checks were performed be listed in the \textit{Data report overview}? &  \code{TRUE} \\
 \quad \code{preChecks} & What check functions should be called to determine whether a variable is suitable for summarization, visualization and checking? & \code{c("isKey", "isEmpty")} \\
@@
 \end{tabular}
-\caption{A selection of commonly used arguments to \code{clean()} separated into the parts they control.}
+\caption{A selection of commonly used arguments to \code{makeDataReport()} separated into the parts they control.}
 \label{table.cleanFormals}
@@
 
-% Below, we present a selection of these arguments in
-% sections \hl{xx} and \hl{xxx}. We do this under two distinct headlines
-% to emphasize that two levels of control are available when using
-% \code{clean()}. Either, we can control the \hl{[something] - essentially,
-%   this is everything but SVC-functions - } or we can control what
-% quality assessments and descriptions each variable are subjected
-% to.  \hl{segway to next bit or
-%   just something to end the paragraph}
-
 \begin{figure}[tb]
-% \begin{description}
-% \item[Input] A dataset
-% 	\begin{itemize}
-% 		\item This should be of type \code{data.frame} or \code{tibble} \hl{or data.table?}
-% 	\end{itemize}
-% \item[Create contents] For each variable in the dataset, do the following:
-% 	\begin{description}
-% 		\item[Stage 1:] Pre-checks
-% 			\begin{itemize}
-% 				\item Is the variable suitable for summarization, visualization and checks?
-% 				\begin{description}
-% 					\item[Yes:] Go to stage 2.
-% 					\item[No:] Move on to the next variable.
-% 				\end{description}
-% 			\end{itemize}
-% 		\item[Stage 2:] SVC-steps
-% 			\begin{description}
-% 				\item[Summarize] Call \code{summarize()} to produce a summary table describing the variable. What features enter this table depends on the data class of the variable.
-% 				\item[Visualize]  Call \code{visualize()} to produce a plot visualizing the distribution of the variable.
-% 				\item[Check]  Call \code{check()} to apply quality- and error checks to the variable. What checks are used depends on the data class of the variable.
-% 			\end{description}
-% 	\end{description}
-% \item[Output] Files for a overview document are saved to the disc and possibly also opened.
-% 	\begin{itemize}
-% 		\item Always a \code{rmarkdown} (.Rmd) file
-% 		\item Possible also a html or pdf file
-% 	\end{itemize}
-% \end{description}
-
 % Define block styles
@@
 \caption{Schematic illustration of the stages undertaken when running
-  \code{clean()}. Each variable is checked for eligibility before
+  \code{makeDataReport()}. Each variable is checked for eligibility before
   running \code{summarize()}, \code{visualize()}, and \code{check()}, and the
@@
 We begin with an example that is intended as an illustration of how
-\code{clean()} might be used in the very first stages of data
+\code{makeDataReport()} might be used in the very first stages of data
 cleaning, when the we are uncertain about the complexities of the
-errors and how much times should be allocated to data cleaning. At
+errors and how much time should be allocated to data cleaning. At
 this stage, what is really needed, is a very rough idea of the
@@
 10 printed potential errors for each variable. Also, we can add the
-argument \code{replace=TRUE} in order to force \code{clean()} to
-overwrite any existing files produced by \code{clean()}.  Using the
+argument \code{replace=TRUE} in order to force \code{makeDataReport()} to
+overwrite any existing files produced by \code{makeDataReport()}.  Using the
 \code{toyData} dataset as a guinea pig, we type:
@@
 \begin{Sinput}
-R> clean(toyData, output = "html", onlyProblematic = TRUE, maxProbVals = 10,
+R> makeDataReport(toyData, output = "html", onlyProblematic = TRUE, maxProbVals = 10,
 +    replace = TRUE)
@@
 opened. The following command produces an \proglang{R} markdown file
-containing the data cleaning summary information, but without
+containing the information needed for generating a data report, but without
 rendering nor opening the markdown file:
@@
 \begin{Sinput}
-R> clean(toyData, output="html", render=FALSE, openResult=FALSE)
+R> makeDataReport(toyData, output="html", render=FALSE, openResult=FALSE)
 \end{Sinput}
@@
   \quad \code{centralValue} & Compute median or mode &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\
-  \quad \code{countMissing} & Compute ratio of missing observations &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\
+  \quad \code{countMissing} & Compute proportion of missing observations &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\
   \quad \code{minMax} & Find minimum and maximum values &   &  & $\times$ & &  & $\times$ & $\times$  \\
@@
 We will now move on to discuss how not only the structure of the
-cleaning summary is manipulated, but also its very contents. This is
+data assesment steps is manipulated, but also its very contents. This is
 done through the summarize/visualize/check (SVC) steps, as illustrated
@@
 
-countMissing   Compute ratio of missing        character, Date, factor,     
+countMissing   Compute proportion of missing        character, Date, factor,     
                observations                    integer, labelled, logical,  
@@
 We can control what summaries and checks are applied for each variable type
-through the \code{XXXSummaries} and \code{XXXChecks} arguments, where \code{XXX}
-represents a variable type, e.g., \code{factorChecks} for factors,
-\code{numericChecks} for numeric variables, etc. These arguments accept a
-vector of summary- or check function names that should be run for a
-particular variable type. The default values for each of these
-arguments can be obtained through the \code{defaultXXXChecks()}
-and \code{defaultXXXSummaries()} functions, where \code{XXX} again refers to
-the variable type. For example, the default summaries being used for a factor variable is
-
-\begin{Schunk}
-\begin{Sinput}
-R> defaultFactorSummaries()
-\end{Sinput}
-\begin{Soutput}
-[1] "variableType" "countMissing" "uniqueValues" "centralValue"
-\end{Soutput}
-\end{Schunk}
-
-We can change the summaries (and similarly the checks) by setting the
-corresponding arguments when calling \code{clean()}. For example, to get
-only the variable type and the central tendency listed in the summary
-table of each variable we write
-
-\begin{Schunk}
-\begin{Sinput}
-R> clean(toyData, characterSummaries = c("variableType", "centralValue"),
-+    factorSummaries = c("variableType", "centralValue"),
-+    labelledSummaries = c("variableType", "centralValue"),
-+    numericSummaries = c("variableType", "centralValue"),
-+    integerSummaries = c("variableType", "centralValue"),
-+    logicalSummaries = c("variableType", "centralValue"),
-+    dateSummaries = c("variableType", "centralValue"))
-\end{Sinput}
-\end{Schunk}
-
-In this particular case, where we specify the same set of summary
-functions for each variable type, we can use the simpler argument
-\code{allSummaries} which overrides the summary functions for all
-variable types. Thus, the same result as above could be obtained with
-
-\begin{Schunk}
-\begin{Sinput}
-R> clean(toyData, allSummaries = c("variableType", "centralValue"))
-\end{Sinput}
-\end{Schunk}
-
-Similarly, the checks applied are set with the \code{XXXChecks}
-arguments. The default checks being applied to a factor is
-
-\begin{Schunk}
-\begin{Sinput}
-R> defaultFactorChecks()
-\end{Sinput}
-\begin{Soutput}
+through the \code{summaries}, \code{visuals} and \code{checks} arguments.
+Each of these arguments should be a list with one entry for each variable type and a number of function names for each such entry.
+The easiest way to specify the arguments is by use of the built-in helper functions \code{setSummaries()}, \code{setVisuals()} and \code{setChecks()}.
+
+For instance, we can inspect the default settings for summaries by calling
+
+\begin{Schunk}
+\begin{Sinput}
+R> setSummaries()
+\end{Sinput}
+\begin{Soutput}
+$character
 [1] "identifyMissing"  "identifyWhitespace"  "identifyLoners"    
 [4]  "identifyCaseIssues"  "identifyNums"
+
+$factor
+[1] "identifyMissing"  "identifyWhitespace"  "identifyLoners"     
+[4] "identifyCaseIssues"  "identifyNums"
+
+$labelled
+[1] "identifyMissing"  "identifyWhitespace"  "identifyLoners"     
+[4] "identifyCaseIssues"  "identifyNums"
+
+$numeric
+[1] "identifyMissing"  "identifyOutliers"
+
+$integer
+[1] "identifyMissing"  "identifyOutliers"
+
+$logical
+NULL
+
+$Date
+[1] "identifyOutliers"
+\end{Soutput}
+\end{Schunk} %$
+
+This helper function really just calls several other helper functions, namely the \\ 
+\code{defaultXXXSummaries()} functions, where \code{XXX} refers to a variale class. For instance, we can see the default character checks by calling \code{defaultCharacterSummaries()}:
+
+\begin{Schunk}
+\begin{Sinput}
+R> defaultCaracterSummaries()
+\end{Sinput}
+\begin{Soutput}
+[1] "variableType" "countMissing" "uniqueValues" "centralValue"
+\end{Soutput}
+\end{Schunk}
+
+We can change the choice of summaries (and similarly the checks and visual functions) by setting the
+corresponding arguments when calling \code{makeDataReport()}. For example, to get
+only the variable type and the central tendency listed in the summary
+table for numeric and integer variables, we write
+
+\begin{Schunk}
+\begin{Sinput}
+R> makeDataReport(toyData, 
+ 	summaries = setSummaries(numeric = c("variableType","centralValue),
+		integer = c("variableType", "centralValue"))
+\end{Sinput}
+\end{Schunk}
+
+In the case where we specify the same set of summary
+functions for each variable type, we can use a simpler argument for \code{setSummaries} which overrides the summary functions for all
+variable types:
+
+\begin{Schunk}
+\begin{Sinput}
+R> makeDataReport(toyData, 
+	summaries = setSummaries(all = c("variableType", "centralValue"))
+\end{Sinput}
+\end{Schunk}
+
+Similarly, the checks applied are set with the \code{checks} argument and the \code{setChecks} function. The default checks being applied to a factor are
+
+\begin{Schunk}
+\begin{Sinput}
+R> defaultFactorChecks()
+\end{Sinput}
+\begin{Soutput}
+[1] "identifyMissing"    "identifyWhitespace" "identifyLoners"    
+[4] "identifyCaseIssues" "identifyNums"      
 \end{Soutput}
@@
 Now, if we only wanted to apply the function to identify whitespace
-for factor variables, then we would need to set the
-\code{factorChecks} accordingly
-
-\begin{Schunk}
-\begin{Sinput}
-R> clean(toyData, factorChecks = c("identifyWhitespace"))
+for factor variables, then we would need provide this information for \code{setChecks()}:
+
+\begin{Schunk}
+\begin{Sinput}
+R> makeDataReport(toyData, checks = setChecks(factor = "identifyWhitespace"))
 \end{Sinput}
@@
 \begin{Sinput}
-R> clean(toyData, factorChecks = NULL)
+R> makeDataReport(toyData, checks = setChecks(factor = NULL))
 \end{Sinput}
@@
 \code{allCheckFunctions()}. Note however, that \code{checkFunction}s have a
-usage beyond the \code{XXXChecks} arguments, namely in the
+usage beyond the \code{checks} arguments, namely in the
 pre-check stage. In this stage, it is determined whether or
@@
 \code{checkFunction}s that are applicable to all variable classes. The
-default pre-checks, the functions \code{isKey()} and \code{isEmpty()}, check
+default pre-checks, the functions \code{isKey()} and \code{isSingular()}, check
 whether a variable has unique values for all observations or only a
@@
 \begin{Sinput}
-R> clean(toyData, preChecks = "isKey")
+R> makeDataReport(toyData, preChecks = "isKey")
 \end{Sinput}
@@
 Note that the data visualizations in the cleaning summary are also
-controllable, though only a single function can be provided for all
-variable types. If, for instance, we wish to change the visualizations
+controllable, though only a single function can be provided for each variable type. If, for instance, we wish to change the visualizations
 from the default \pkg{ggplot2} \citep{ggplot2} style histograms and barplots to base
@@
 \begin{Sinput}
-R> clean(toyData, allVisuals = "basicVisual")
-\end{Sinput}
-\end{Schunk}
-
-%\hl{
-%\begin{itemize}
-%\item Introduce the relevant \code{clean}-arguments and the \code{defaultWhateverSummaries} etc.-functions
-%\item Introduce \code{allSummaryFunctions()} etc. and present a table corresponding to the output of this call
-%\item Small examples:
-%\begin{itemize}
-%\item Add a function to one of the XXXSummaries/XXXVisuals/XXXChecks-arguments (still calling default options)
-%\item Remove all but a single function from one of these arguments
-%\item Describe what happens if the argument is NULL
-%\end{itemize}
-%\end{itemize}
-%}
+R> makeDataReport(toyData, visuals = setVisuals(all = "basicVisual"))
+\end{Sinput}
+\end{Schunk}
 
 As indicated in Figure~\ref{figure:cleanStructure}, there are two stages
-where \code{clean()} applies functions to each of the variables:
+where \code{makeDataReport()} applies functions to each of the variables:
 \begin{enumerate}
@@
 Each of these stages are controllable using appropriate function
-arguments in \code{clean()}, and above we have shown examples of how
+arguments in \code{makeDataReport()}, and above we have shown examples of how
 to tweak them to modify the data cleaning outputs. However, if the
 dataset at hand requires new, additional checks, then more control is
-needed, and Section~\ref{sec:extending} explains how to modify and
+needed \hl{ and we have provided a detailed vignette that explains how to modify and
 expand the possibilities by producing new summary, visual, and check
-functions.
-
-\section[Using dataMaid interactively]{Using \pkg{dataMaid} interactively}
+functions in [REFERENCE]}.
+
+\section[Using dataMaid interactively]{\hl{Using \pkg{dataMaid} interactively}}
 \label{sec:interactiveCleanR}
@@
 provides more standard \proglang{R} interactive tools, such as
-functions that print results to the console or returns the information
+functions that print results to the console or return the information
 as an object for later use. This section describes how to use the
@@
 \end{Soutput}
-\end{Schunk}
+\end{Schunk}%$
 
@@
 % \item Mention check(), visualize() and summarize() modes for data.frames.
-Maybe also advice against it, at it will often produce a lot of information at once, and such large amounts of information really should be documented.
+Maybe also advise against it, at it will often produce a lot of information at once, and such large amounts of information really should be documented.
 %\end{itemize}
 %}
-
-\section[Extending dataMaid]{Extending \pkg{dataMaid}}
-\label{sec:extending}
-
-\pkg{dataMaid} can be used as a user-friendly, self-contained package, as
-shown in the previous sections. However, \pkg{dataMaid} is fully
-customizable and \code{clean()} is mainly a tool for formatting the
-results from various checking, summary,  and visualization
-functions. Thus, the actual work underlying a \pkg{dataMaid} output file
-can be anything --- depending on the arguments given to \code{clean()}
---- and user made functions are easily added to the
-summarize/visualize/check-function arguments, as discussed previously.
-However, the functions used in the SVC steps must adhere
-to specific structures in order to be called from these three steps, and
-therefore, we will now present how \code{summaryFunction}s,
-\code{visualFunction}s and \code{checkFunction}s are made.
-
-This section consists of two parts. First, we describe how customized
-\code{summaryFunction}s, \code{visualFunction}s and
-\code{checkFunctions} can be made, including an introduction to a few
-convenient tools available in \pkg{dataMaid} for aiding SVC-function
-construction.  Table \ref{table.functionTypes} serves as an overview
-of the internals of these three central function types. Secondly, we
-turn to a worked example of how to use custom made functions in
-practice; Four new SVC functions are defined and used, both
-interactively and in \code{clean()}.
-
-\begin{table}[htbp]
-\footnotesize
-\bgroup
-\def\arraystretch{1.8}%
-\begin{tabular}{L{0.2\linewidth}L{0.2\linewidth}L{0.2\linewidth}L{0.2\linewidth}}
-& \code{summaryFunction} & \code{visualFunction} & \code{checkFunction} \\
-\hline
-\vspace{0pt} Input (required) & \code{v} --- a variable vector.
-                                \newline \code{...} --- additional
-                                arguments passed to the function.
-&  \code{v} --- a variable vector.
-\newline \code{vnam} --- the variable name (as character string).
-\newline \code{doEval} --- a logical (\code{TRUE}/\code{FALSE}) controlling the output type of the function.
-& \code{v} - a variable vector \newline \code{nMax} - an integer (or \code{Inf}), controlling how many problematic values are printed, if relevant \newline \code{...} --- additional arguments passed to the function.
-\\
-Input (optional) &  \code{maxDecimals} --- number of decimals printed in outputted numerical values.  & --- &  \code{maxDecimals}  --- number of decimals printed in outputted numerical values.  \\
-Purpose & Describe some aspect of the variable, e.g., a central value, its dispersion or level of missingness.
-& Produce a distribution plot.
-& Check a variable for a specific issue and, if relevant, identify the values in the variable that cause the issue.
-\\
-Output (required) & A list with entries: \newline \code{\$feature} --- a label for the summary value (as character string); \newline \code{\$result} --- the result of the summary (as character string).
-& A character string with \proglang{R} code for producing a plot.
-This code should be standalone, i.e., should include the data if necessary.
-& A list with entries: \newline \code{\$problem} --- a logical identifying whether an issue was found; \newline \code{\$message} --- a character string (possibly empty) decribing the issue that was found, properly escaped and ready for use with \code{rmarkdown}.
-\\
-Output (recommended) & A \code{summaryResult} object, i.e., an attributed list with entries \code{\$feature}, \code{\$result} and \code{\$value}, the latter being the values from \code{\$result} in their original format).
-& \bgroup
-\def\arraystretch{1.8}%
-\hspace*{-.2cm}\begin{tabular}[t]{p{0.45\linewidth} |
-                 p{0.45\linewidth}} \vspace*{-1cm}\raggedright If
-                 \code{doEval} is \code{TRUE}: \newline A plot that is
-                 opened by the graphic device in \proglang{R}. &
-                                                          \vspace*{-1cm}\raggedright If \code{doEval} is \code{FALSE}: \newline   A text string with \proglang{R} code, as described above.
-\end{tabular} \egroup  & A \code{checkResult} object, i.e., an attributed list with entries \code{\$problem}, \code{\$message} and \code{\$problemValues}, the latter being either \code{NULL} or the problem causing values, as they were found in \code{v}, whichever is relevant.
-\\
-Tools available for producing the function & \code{summaryResult()} & --- & \code{messageGenerator()} \newline \code{checkResult()}
-\end{tabular}
-\egroup
-\caption{Reference information for creating new functions to be used
-  as part of the summarize, visualize, and check steps. The three
-  columns correspond to each of the three function types.
-%\hl{I think the formatting here is still sort of chaotic.
-Maybe include horizontal lines everywhere?
-Is it allowed in JSS?
-It's impossible to read what's in which row right now.}
-}
-\label{table.functionTypes}
-\end{table}
-
-\subsection{Function templates}
-\label{sec:functionTemplates}
-In order to construct a summary, visual or check function, one needs
-to create a new function with a specific structure.  This can be done with
-different levels of strictness. If
-the new custom function is only to be used as part of the SVC steps in \code{clean()},
-then only the input/output structure of the function needs special attention.
-However, new user-defined functions can also be registered locally to
-be part of the full machinery of \pkg{dataMaid}, and these function will
-be recognized and behave in the same way as the build-in functions in
-\pkg{dataMaid}. The presentation below is given in the format of
-function templates, written in pseudo-code. These templates are
-designed for getting the full functionality, but note that
-Table~\ref{table.functionTypes} serves as a reference to the minimal
-requirements, while also presenting the "full" versions of the
-function types.
-
-% \hl{metatext, mention Table \ref{table.functionTypes} again.}
-
-\subsubsection{Writing a summaryFunction}
-As mentioned above, \pkg{dataMaid} provides a dedicated class for
-\code{summaryFunction}s. However, this does not imply that they are
-particularly advanced or complicated to create; in fact, they are
-nothing but regular functions with a particular
-input/output-structure. Specifically, they all follow the template
-below:
-\begin{Verbatim}
-mySummaryFunction <- function(v, ...) {
-  res <- [result of whatever summary we are doing]
-  summaryResult(list(feature = "[Feature name]", result = res))
-}
-\end{Verbatim}
-The last function called here, \code{summaryResult()}, changes the
-class of the output, thereby making a \code{print()} method available
-for it.  Note that \code{v} is an input vector and that \code{res}
-should be either a character string or something that will be printed
-as one. In other words an integer would be allowed, but a matrix will
-not. Though a lot of different things can go into the
-\code{summaryFunction} template, we recommend only using it for
-summarizing the features of a variable, and leaving tests and checks
-for the \code{checkFunction}s (presented below).
-
-Adhering to the template above is sufficient for using the freshly
-made \code{mySummaryFunction()} in \code{clean()}, but we recommend
-adding the new function to the overview of all summary
-functions by converting it to a proper \code{summaryFunction}
-object. This is done by calling the \code{summaryFunction()} creator with
-the user-defined function as the first argument, and additional arguments
-\code{description} (an explanatory text which will be added to the attributes of
-the function), and \code{classes} (a vector of variable classes the
-user-defined function is intended to be applied to, also stored as an
-attribute). In other words, a call following the template below should be made:
-\begin{Verbatim}
-mySummaryFunction <- summaryFunction(mySummaryFunction,
-  description = "[Text describing what the summaryFunction does]",
-  classes = c([vector of data types that the function is intended
-  for]))
-\end{Verbatim}
-which adds the new function to the output of an
-\code{allSummaryFunctions()} call.
-% One comment should be devoted to
-% the two attributes of a \code{summaryFunction}.
-If the \code{description} argument is left unspecified, the name of
-the function (which in this case is \code{"mySummaryFunction"}) will
-be filled in and used as the \code{description} attribute. What
-happens if the \code{classes} argument is not specified depends on the
-type of \code{summaryFunction}. If \code{mySummaryFunction} is
-constructed as an \code{S3} generic function with associated methods,
-the call to \code{summaryFunction()} will automatically produce a
-vector of the names of the classes for which the function can be
-called. If \code{mySummaryFunction()} is not an \code{S3} generic and
-\code{classes} is left unspecified, the attribute will simply be left
-empty. Note that the helper function \code{allClasses()} might be
-useful for filling out the \code{classes} argument, as it simply lists
-all available classes in \pkg{dataMaid}:
-
-\begin{Schunk}
-\begin{Sinput}
-R> allClasses()
-\end{Sinput}
-\begin{Soutput}
-[1] "character" "Date"      "factor"    "integer"   "labelled" 
-[6] "logical"   "numeric"  
-\end{Soutput}
-\end{Schunk}
-
-% \hl{Write something here, don't end paragraph with code. Also, maybe move the allClasses() stuff somewhere else, it doesn't really belong under this header. Not sure where to, though.}
-
-%Jeg foreslår at droppe nedenstående, da jeg synes det bliver lidt søgt i forhold
-%til at det kun er beskrevet i pseudokode. (fx. er classes = numeric meget arbitrært)
-%If we run the \code{allSummaryFunctions()} function we can see that the
-%new function has been registered (output slightly appreviated):
-
-%\begin{Schunk}
-%\begin{Sinput}
-%> allSummaryFunctions()
-%\end{Sinput}
-%\begin{Soutput}
-%----------------------------------------------------------------------------
-%name                description                                      classes
-%------------------- --------------------------------------------------------
-%mySummaryFunction   [Text describing what the summaryFunction does]  numeric
-%.
-%.
-%.
-%\end{Soutput}
-%\end{Schunk}
-
-\subsubsection{Writing a visualFunction}
-\code{visualFunction}s are the functions that produce the figures in a
-\pkg{dataMaid} output document. Writing a \code{visualFunction} is slightly
-more complicated than writing a \code{summaryFunction}. This follows from
-the fact that \code{visualFunction}s need to be able to output standalone
-code for plots in order for \code{clean()} to build standalone
-\pkg{rmarkdown} files. We recommend using the following structure
-(again shown as pseudo code):
-\begin{Verbatim}
-myVisualFunction <- function(v, vnam, doEval) {
-  thisCall <- call("[the name of the function used to produce the plot]",
-    v, [additional arguments for the plotting function])
-  if (doEval) {
-    return(eval(thisCall))
-  } else return(deparse(thisCall))
-}
-\end{Verbatim}
-
-In this function, \code{v} is the variable to be visualized,
-\code{vnam} is its name (which should generally be passed to
-\code{title} or \code{main} arguments in plotting functions) and
-\code{doEval} controls whether the output is a plot (if \code{TRUE})
-or a character string of standalone code for producing a plot (if
-\code{FALSE}). Implementing the \code{doEval = TRUE} setting is not
-strictly necessary for a \code{visualFunction}'s use in \code{clean()}, but
-it makes it easier to assess what visualization options are available,
-and obviously, it is crucial for interactive usage of
-\code{myVisualFunction()}. In either case, it should be noted that all
-the parameters listed above, \code{v}, \code{vnam} and \code{doEval},
-are mandatory, so they must be left as is, even if they are not in
-use (c.f., Table~\ref{table.functionTypes}).
-
-As with the summary function, we call \code{visualFunction()} to
-register our newly created function:
-%. Like before, it accepts the
-%user-defined function as the first argument, and two additional
-%arguments description which adds explanatory flavour text as an
-%attribute of the function, and a vector which determines which classes
-%the user-defined function should be applied to (which is also stored
-%as an attribute).
-\begin{Verbatim}
-myVisualFunction <- visualFunction(myVisualFunction,
-  description = "[Some text describing the visualFunction]",
-  classes = c([data types that this function is intended for]))
-  )
-\end{Verbatim}
-
-Now, \code{myVisualFunction()} will be available in a \code{allVisualFunctions()}
-call, just like the two build-in \code{visualFunction}s, \code{standardVisual}
-and \code{basicVisual}.
-
-\subsubsection{Writing a checkFunction}
-The last, but perhaps most important, \pkg{dataMaid} function type is
-the \code{checkFunction}. These are the functions that flag issues in
-the data as part of the check step and/or control the overall flow of the data
-cleaning process in the pre-check stage. A \code{checkFunction} follows
-one of two overall structures, depending on the type of check. Either,
-it tries to identify problematic values in the variable (as
-e.g., \code{identifyMissing()} does), or it performs a check concerning the
-variable as a whole (e.g., the functions used for pre-checks and the
-function \code{identifyNums()}). We present templates for both types of
-\code{checkFunction}s below separately, but it should be emphasized that
-formally, they belong to the same class.
-
-First, a template for the full-variable check function type, where we
-first define the function and subsequently register it as a check
-function using \code{checkFunction()}:
-\begin{Verbatim}
-myFullVarCheckFunction <- function(v, ...) {
-  [do your check]
-  problem <- [is there a problem? TRUE/FALSE]
-  message <- "[message describing the problem, if any]"
-  checkResult(list(problem = problem,
-    message = message,
-    problemValues = NULL))
-}
-
-myFullVarCheckFunction <- checkFunction(myFullVarCheckFunction,
-  description = "[Some text describing the checkFunction]",
-  classes = c([the data types that this function is intended to be used for])
-  )
-\end{Verbatim}
-
-Again, as with \code{summaryFunction}s and \code{visualFunction}s, the
-change of function class by use of \code{checkFunction()} is not strictly
-necessary. Note however, that if \code{myFullVarCheckFunction} is to be
-used in the summarize/visualize/check steps in \code{clean()}, the
-description attribute will be printed in the overview table in the
-\textit{Data cleaning summary} part of the output document.
-
-If problematic values are to be identified, the template from above
-should be expanded to follow a slightly more complicated structure
-(again shown as pseudo-code):
-
-\begin{Verbatim}
-myProbValCheckFunction <- function(v, nMax, maxDecimals, ...) {
-  [do your check]
-  problem <- [is there a problem? TRUE/FALSE]
-  problemValues <- [vector of values in v that are problematic]
-  problemStatus <- list(problem = problem,
-                        problemValues = problemValues)
-
-  problemMessage <- "[Message that is printed prior to listing
-                     problem values in the dataMaid output,
-                     ending with a colon]"
-
-  outMessage <- messageGenerator(problemStatus, problemMessage, nMax)
-
-  checkResult(list(problem = problem,
-    message = outMessage,
-    problemValues = problemValues))
-}
-
-myProbValCheckFunction <- checkFunction(myProbValCheckFunction,
-  description = "[Some text describing the checkFunction]",
-  classes = c([the data types that this function is intended to be used for])
-)
-\end{Verbatim}
-
-One comment should be devoted to the helper function,
-\code{messageGenerator()}.  This function's sole purpose is aiding
-consistent styling of all \code{checkFunction} messages. The function
-simply pastes together the \code{problemMessage} and the
-\code{problemValues}, with the latter being quoted and sorted
-alphabetically. If the \code{nMax} argument to \code{messageGenerator()} is
-not \code{Inf}, only the first \code{nMax} problem values will be pasted
-onto the message, accompanied by a comment about how many problem
-values were left out (if any).  Note that printing quotes in
-\proglang{R} markdown requires an extensive amount of character escaping, so
-opting for \code{messageGenerator()} really is the easiest solution.
-
-In the template above, the argument \code{maxDecimals} is not in use. This
-argument should be used to round off the \code{problemValues} passed to
-\code{messageGenerator()}, if they are numerical.  This can be done by
-adding an extra line of code after defining \code{problemStatus} in the template
-above:
-
-\begin{Verbatim}
-myProbValCheckFunction <- function(v, nMax, maxDecimals, ...) {
-  [... more lines of code here ...]
-  problemStatus <- list(problem = problem,
-    problemValues = problemValues)
-
-  if (!is.null(problemValues)) {
-    problemStatus$problemValues <- round(problemValues, maxDecimals)
-  }
-  [... more lines of code here ...]
-}
-\end{Verbatim}
-% $
-
-Now, problematic values will be rounded in the outputted message, while they
-will still appear in their original format under the entry \code{\$problemValues} %$
-in the returned object.
-
-\subsection{A worked example}
-
-As an example, we create four new functions and show both how they can be
-used interactively and how they can be integrated with the \code{clean()}
-function. These four new functions are:
-\begin{description}
-\item[\code{isID()}] A new \code{checkFunction} intended for use in the pre-check-stage.
-This function checks whether a variable consists exclusively of long ($> 10$ characters/digits) entries that are all of equal length, as this might be personal identification codes that we do not wish to print out in the data summary.
-\item[\code{mosaicVisual()}] A new \code{visualFunction} that produces mosaic plots. This function will be used in the \textit{visualize} step of \code{clean()}.
-\item[\code{countZeros()}] A new \code{summaryFunction} that counts the number of occurrences of the value \code{0} in a variable.
-This function will be used in the \textit{summarize} step of \code{clean()}.
-\item[\code{identifyColons()}] A new \code{checkFunction} that flags variables in which values have colons that appear before and after alphanumerical characters.
-This is practical for identifying autogenerated interaction effects.
-This function will be used in the \textit{check} step of \code{clean()}.
-\end{description}
-These functions are defined in turn below, and afterwards, an example of how they can be called from \code{clean()} is provided.
-
-\subsubsection[isID --- a new checkFunction without problem values]{\code{isID} --- a new \code{checkFunction} without problem values}
-First, let's define the \code{isID()} function. As this function is
-not supposed to list problematic values, it falls within the category
-of \code{checkFunction}s represented by
-\code{myFullVarCheckFunction()} above. We do not particularly wish to
-use this function interactively, so we will stick to the minimal
-requirements of a \code{checkFunction} used in \code{check()} (see
-Table~\ref{table.functionTypes}). The function can then be defined by
-
-\begin{Schunk}
-\begin{Sinput}
-R> isID <- function(v, nMax = NULL, ...) {
-+    out <- list(problem = FALSE, message = "")
-+    if (class(v) %in% setdiff(allClasses(), c("logical", "Date"))) {
-+      v <- as.character(v)
-+      lengths <- c(nchar(v))
-+      if (all(lengths > 10) & length(unique(lengths)) == 1) {
-+        out$problem <- TRUE
-+        out$message <- "Warning: This variable seems to contain ID codes."
-+      }
-+    }
-+    out
-+  }
-\end{Sinput}
-\end{Schunk}
-
-This is essentially all we need to do in order to include this
-function as a pre-check-function in \code{clean()}, so we will leave it as
-is and move on to the next function, namely \code{mosaicVisual()}.
-
-\subsubsection[mosaicVisual --- a new visualFunction]{\code{mosaicVisual} --- a new \code{visualFunction}}
-\proglang{R} provides a function, \code{mosaicplot()}, which produces
-mosaic plots. We intend to use this existing high-level plotting
-function as part of the new visualization function.  We will define
-the new function such that it gets the full \pkg{dataMaid}
-functionality. This can be done using the following code which sets up
-the call using the existing function \code{mosaicplot()}.
-
-\begin{Schunk}
-\begin{Sinput}
-R> mosaicVisual <- function(v, vnam, doEval) {
-+    thisCall <- call("mosaicplot", table(v), main = vnam, xlab = "")
-+    if (doEval) {
-+      return(eval(thisCall))
-+    } else return(deparse(thisCall))
-+  }
-\end{Sinput}
-\end{Schunk}
-
-This function can now be called directly or used in \code{clean()}, as will
-presented in an example below. Depending on the \code{doEval} argument, either a text string
-with code or a plot is produced. The plot resulting from the following
-call is found in Figure~\ref{fig:mosaicPlot}:
-
-\begin{Schunk}
-\begin{Sinput}
-R> mosaicVisual(toyData$var1, "variable 1", doEval = TRUE)
-\end{Sinput}
-\end{Schunk}
-
-\begin{figure}
-\begin{center}
-\includegraphics[scale=0.2]{mosaicPlotExample.pdf}
-\end{center}
-
-\caption{Mosaic plot generated by the user-defined visualization
-  function \code{mosaicVisual()}.}
-\label{fig:mosaicPlot}
-\end{figure}
-
-Even though \code{mosaicVisual()}, as written above, follows the style of a
-\code{visualFunction}, it is not yet truly one and therefore, it will not
-appear in an \code{allVisualFunctions()} call. In order to get this
-functionality we need to change its object class. This can be done by
-writing
-
-\begin{Schunk}
-\begin{Sinput}
-R> mosaicVisual <- visualFunction(mosaicVisual,
-+    description = "Mosaic plots using graphics",
-+    classes = allClasses())
-\end{Sinput}
-\end{Schunk}
-
-Here, we use the function \code{allClasses()} to quickly obtain a vector
-of all the seven variable classes addressed in \pkg{dataMaid}. Note that
-if \code{mosaicVisual()} were an S3 generic function, this argument could
-have been left as \code{NULL} and then the classes for which methods are
-available would be added automatically.
-%\hl{I'm repeating myself here,
-%  but I think it is quite a neat feature, so maybe that's okay?}
-
-As \code{mosaicVisual()} is now a full-blooded \code{visualFunction}, it
-will also be included in the \code{allVisualFunctions()} output table:
-\begin{Schunk}
-\begin{Sinput}
-R> allVisualFunctions()
-\end{Sinput}
-\begin{Soutput}
-------------------------------------------------------------------------------
-name             description                     classes                      
----------------- ------------------------------- -----------------------------
-mosaicVisual     Mosaic plots using graphics     character, Date, factor,     
-                                                 integer, labelled, logical,  
-                                                 numeric
-
-basicVisual      Histograms and barplots using   character, Date, factor,     
-                 graphics                        integer, labelled, logical,  
-                                                 numeric
-
-standardVisual   Histograms and barplots using   character, Date, factor,     
-                 ggplot2                         integer, labelled, logical,  
-                                                 numeric                      
-------------------------------------------------------------------------------
-\end{Soutput}
-\end{Schunk}
-
-Now that we are done with the definition of \code{mosaicVisual()} we can
-turn to the next function in line, \code{countZeros()}.
-
-\subsubsection[countZeros --- a new summaryFunction]{\code{countZeros} --- a new \code{summaryFunction}}
-This \code{summaryFunction} is defined in the following lines of code:
-
-\begin{Schunk}
-\begin{Sinput}
-R> countZeros <- function(v, ...) {
-+    res <- length(which(v == 0))
-+    summaryResult(list(feature = "No. zeros", result = res, value = res))
-+  }
-\end{Sinput}
-\end{Schunk}
-
-As this function computes an integer (the number of zeros),
-there is no difference between the elements \code{\$result} and
-\code{\$value}. If, on the other hand, the result had been a character
-string, extra formatting might be required in the \code{\$result} entry
-(such as escaping of quotation marks), and in this scenario, the two
-entries would have differed. As the result is returned as a
-\code{summaryResult} object, a printing method is automatically called
-when \code{countZeros()} is used interactively:
-
-\begin{Schunk}
-\begin{Sinput}
-R> countZeros(c(rep(0, 5), 1:100))
-\end{Sinput}
-\begin{Soutput}
-No. zeros: 5
-\end{Soutput}
-\end{Schunk}
-
-As with \code{mosaicVisual()}, we change the class of this function in
-order to make it appear in \code{allSummaryFunctions()} calls. But now we
-wish to emphasize that the function is not intended to be called on
-all variable types, as zeros have different roles in \code{Date}s and in
-\code{logical} variables:
-
-\begin{Schunk}
-\begin{Sinput}
-R> countZeros <- summaryFunction(countZeros,
-+    description = "Count number of zeros",
-+    classes = c("character", "factor", "integer",
-+    "labelled", "numeric"))
-\end{Sinput}
-\end{Schunk}
-
-\subsubsection[identifyColons --- a new checkFunction with problem values]{\code{identifyColons} --- a new \code{checkFunction} with problem values}
-The last function mentioned above is \code{identifyColons()}. We
-define it using the helper function \code{messageGenerator()} to
-obtain a properly escaped message, and we use \code{checkResult()} to
-make its output print neatly. In the code we use regular expressions
-through the \code{gregexpr()} function to identify colons between two
-words.
-
-\begin{Schunk}
-\begin{Sinput}
-R> identifyColons <- function(v, nMax = Inf, ... ) {
-+    v <- unique(na.omit(v))
-+    problemMessage <- "Note: The following values include colons:"
-+    problem <- FALSE
-+    problemValues <- NULL
-+  
-+    problemValues <- v[sapply(gregexpr("[[:xdigit:]]:[[:xdigit:]]", v),
-+      function(x) all(x != -1))]
-+  
-+    if (length(problemValues) > 0) {
-+      problem <- TRUE
-+    }
-+  
-+    problemStatus <- list(problem = problem,
-+      problemValues = problemValues)
-+  
-+    outMessage <- messageGenerator(problemStatus, problemMessage, nMax)
-+  
-+    checkResult(list(problem = problem,
-+      message = outMessage,
-+      problemValues = problemValues))
-+  }
-\end{Sinput}
-\end{Schunk}
-
-Once again, we change the class of the function so that it becomes a
-proper \code{checkFunction} object:
-
-\begin{Schunk}
-\begin{Sinput}
-R> identifyColons <- checkFunction(identifyColons,
-+    description = "Identify colons surrounded by alphanumeric characters",
-+    classes = c("character", "factor", "labelled"))
-\end{Sinput}
-\end{Schunk}
-
-Note, that for \code{checkFunction}s, the description will appear in
-the document produced by \code{clean()} (in the \textit{Data cleaning
-  summary} section of the output), so for this type of function, the
-change of class is not only done for the sake of the
-\code{allCheckFunctions()} output.
-
-\subsubsection[Calling the new summarize/visualize/check functions from clean()]{Calling the new summarize/visualize/check functions from \code{clean()}}
-Now, we are ready to use these new functions in a \code{clean()}
-call. The extended \pkg{dataMaid} output document should have the
-following modifications, relative to the standard \pkg{dataMaid} output:
-\begin{itemize}
-\item We want to add the new pre-check function, \code{isID()}, to the already existing pre-checks.
-\item We wish to change the plot type for all variables to the new mosaic plot.
-\item We want the new summary function, \code{countZeros()}, to be added to the summaries performed on all variable types but \code{Date} and \code{logical}.
-\item We want the new check function, \code{identifyColons()}, to be added to the checks performed on \code{character}, \code{factor} and \code{labelled} variables.
-\end{itemize}
-These options are specified as follows:
-
-\begin{Schunk}
-\begin{Sinput}
-R> data("exampleData")
-R> clean(exampleData,
-+    preChecks = c("isKey", "isEmpty", "isID"),
-+    allVisuals = "mosaicVisual",
-+    characterSummaries = c(defaultCharacterSummaries(), "countZeros"),
-+    factorSummaries = c(defaultFactorSummaries(), "countZeros"),
-+    labelledSummaries = c(defaultLabelledSummaries(), "countZeros"),
-+    numericSummaries = c(defaultNumericSummaries(), "countZeros"),
-+    integerSummaries = c(defaultIntegerSummaries(), "countZeros"),
-+    characterChecks = c(defaultCharacterChecks(), "identifyColons"),
-+    factorChecks = c(defaultFactorChecks(), "identifyColons"),
-+    labelledCheck = c(defaultLabelledChecks(), "identifyColons"))
-\end{Sinput}
-\end{Schunk}
-
-The outputted document is found in Appendix \ref{sec:appendix2}.
 
@@
 typically be for very large datasets, keeping this file open in memory
-wastes precious memory capacities. Therefore, we advice users to
+wastes precious memory capacities. Therefore, we advise users to
 instead split the two steps as shown in the following.

### 2017-10-18 R/makeDataReport.R
Fixed some typos

@@ -597,7 +597,7 @@ makeDataReport <- function(data, output=c("pdf", "word", "html"), render=TRUE,
     writer("The dataset examined has the following dimensions:")
     
     ## Print data frame summary
-    sumMat <- matrix(c("Number of rows", "Number of variables",
+    sumMat <- matrix(c("Number of observations", "Number of variables",
                        n, nvariables), 2,
                      dimnames= list(NULL, c("Feature", "Result")))
     writer(pander::pander_return(sumMat, justify="lr"))

### 2017-10-18 makeData.R

@@ -139,9 +139,9 @@ presidentData$birthday <- as.Date(presidentData$birthday, format = "%m/%d/%Y")
 presidentData$dateOfDeath <- as.Date(presidentData$dateOfDeath, format = "%m/%d/%Y")
 presidentData$presidencyBeginDate <- as.Date(presidentData$presidencyBeginDate, format = "%m/%d/%Y")
 presidentData$presidencyEndDate <- as.Date(presidentData$presidencyEndDate, format = "%m/%d/%Y")
-presidentData$precidencyYears <- round(as.numeric(presidentData$presidencyEndDate - 
+presidentData$presidencyYears <- round(as.numeric(presidentData$presidencyEndDate - 
                                                presidentData$presidencyBeginDate)/365.25,0)
-presidentData$precidencyYears[presidentData$lastName == "Obama"] <- Inf
+presidentData$presidencyYears[presidentData$lastName == "Obama"] <- Inf
 presidentData$ageAtInauguration <- round(as.numeric(presidentData$presidencyBeginDate - 
                                                 presidentData$birthday)/365.25,0)
 presidentData$orderOfPresidency <- as.factor(presidentData$orderOfPresidency)
@@ -159,7 +159,7 @@ presidentData$presidencyBeginDate <- NULL
 presidentData$presidencyEndDate <- NULL
 presidentData$dateOfDeath <- NULL
 presidentData$lastName[presidentData$lastName == "Truman"] <- " Truman"
-
+presidentData$assassinationAttempt[presidentData$lastName == "Obama"] <- 0
 save(presidentData, file="data/presidentData.rda")
 
 

### 2017-10-18 latex/article_vol3.tex
First draft of new data cleaning example section.

@@
 
-%\hl{
-%\begin{itemize}
-%\item Do an example with visualize() and summarize(), like the one with check(). Especially visualize and to doEval = T thing needs a bit of special attention.
-% \item  Mention allCheckFunctions() etc. again here
-% \item Mention check(), visualize() and summarize() modes for data.frames.
-Maybe also advise against it, at it will often produce a lot of information at once, and such large amounts of information really should be documented.
-%\end{itemize}
-%}
+\section{A worked example: Dirty presidents} 
+\label{sec:bigExample}
+\hl{Er titlen her for popsmart og flabet? Det er måske også et generelt tema i teksten. Ret gerne kritisk.}
+
+\begin{figure}[tb]
+\begin{center}
+\frame{\includegraphics[width=7.5cm,page=1]{dataMaid_presidentData.pdf}}
+\frame{\includegraphics[width=7.5cm,page=2]{dataMaid_presidentData.pdf}}\\
+\frame{\includegraphics[width=7.5cm,page=3]{dataMaid_presidentData.pdf}}
+\frame{\includegraphics[width=7.5cm,page=4]{dataMaid_presidentData.pdf}}
+\end{center}
+\caption{The first 4 pages of the data overview report for the \code{presidentData} dataset.
+Note that the report title has been customized (front page), \code{identifyLoners} has been removed from the checks performed on character variables (page 1) and that variables of class \code{Name} have been set to be treated like \code{character} variables (page 1).
+We see that there are two \code{Name} variables in the overview on page 2 and see that these variables are indeed treated like \code{character} variables on page 3.
+}
+\label{fig:bigExampleP0123}
+\end{figure}
+
+\begin{figure}[tb]
+\begin{center}
+\frame{\includegraphics[width=7.5cm,page=5]{dataMaid_presidentData.pdf}}
+\frame{\includegraphics[width=7.5cm,page=6]{dataMaid_presidentData.pdf}}
+\end{center}
+\caption{Part of the data overview report generated for the \code{presidentData} dataset. Here, we see pages 4 and 5 from the \textit{Variable list}.}
+\label{fig:bigExampleP45}
+\end{figure}
+
+We will now put the bits and pieces from above together and show how \code{makeDataReport()} can be used on a less artificial dataset to create a usefl overview report.
+More specifically, we will generate a report describing the \code{presidentData} dataset, which is available in \pkg{dataMaid}.
+This dataset is a sligtly mutilated dataset with information about the 45 first US presidents, but with a few common data issues and a blind passenger.
+The dataset contains one observation per president and has the following variables:
+\begin{description}
+\item[\code{lastName}] The last name of the president.
+\item[\code{firstName}] The first name of the president.
+\item[\code{orderOfPresidency}] The number in the order of presidents.
+\item[\code{birthday}] The birthday of the president.
+\item[\code{stateOfBirth}] The state in which the president was born.
+\item[\code{assassinationAttempt}] Was there an assassination attempt on the president?
+\item[\code{sex}] The sex of the president.
+\item[\code{ethnicity}] The ethnicity of the president.
+\item[\code{presidencyYears}] The duration of the presidency.
+\item[\code{ageAtInauguration}] The age of the president at inauguration.
+\item[\code{favoriteNumber}] The favourite number of the president (fictional).
+\end{description}
+We discuss the results of a data overview report generated for this dataset below, but first there are a few special features of the dataset and wishes for the data report that require us to customize it using some of the arguments for \code{makeDataReport}.
+We have the following points of interest:
+\begin{enumerate}
+\item A few variables are used to describe names, namely \code{lastName} and \code{firstName}.
+In order to use special bibliographical analysis tools on these, and only these, variables, it might be convenient to assign them a special class.
+Therefore, these variables have been set to have class \code{Name} by use of the base \proglang{R} function \code{class()}.
+When we wish to make a data report for the dataset, we have to tell \code{makeDataReport()} how to handle such \code{Name}-type variables, as they are not among the supported variable types mentioned in the above.
+This can be done using the \code{treatXasY} argument.
+\item We use the \code{character} class for a few categorical variables that have a lot of different levels, but where it is not a data mistake that these levels each generally have very few observations, e.g.
+the variable \code{stateOfBirth}.
+Therefore, we would like to disable the \code{identifyLoners} check for this variable type.
+This can be done using the \code{checks} argument.
+\item We would like the report to be called "Dirty president data" in order to reflect that it is, indeed, a report about dirty data about presidents.
+\end{enumerate}
+Incoorporating these three customizations, we can load the data and generate a report by calling:
+\begin{Schunk}
+\begin{Sinput}
+R> data(presidentData)
+R> makeDataReport(presidentData, replace = TRUE, 
+   treatXasY = list("Name" = "character"),
+   checks = setChecks(character = defaultCharacterChecks(remove = "identifyLoners")),
+   reportTitle = "Dirty president data")
+\end{Sinput}
+\end{Schunk}
+The first four pages of the resulting report can be inspected in Figure \ref{fig:bigExampleP0123}.
+Note that all the customized settings can be identified from the first two pages, without having to read through the report: The new title is written on the front page, the check settings are displayed in the \textit{Checks performed} table and the strategy for handling \code{Name} variables is documented below this table.
+
+The first problem that can be spotted from these first four pages is the surprising number of observations: Anno 2017, there have only been 45 US presidents.
+Therefore, having 46 observations reveals that the dataset contains a blind passenger.
+For instance, if the dataset was constructed as a subset of a more general "World leaders" dataset, this type of problem could occur due to wrongful nationality classification.
+We return to the extra president issue below.
+
+On page 3, we see the contents for the three first variables.
+Here, we identifify a prefixed whitespace in the lastname entry for President Truman and we find that a dot was entered as a first name; this is a typical choice for coding missing values in e.g.
+Stata, and therefore, it is flagged as a potential miscoded missing value.
+The variable \code{orderOfPresidency} is not summarized, visualized or checked because it is categorical and contains unqiue values for each observation.
+
+Figure \ref{fig:bigExampleP45} presents the remaining two pages with variable presentations. On pages 4 and 5, we find a few remarks:
+\begin{itemize}
+\item In the birthday variable, there is a entry with the date March 1, 1300 which is a bit of an outlier. 
+\item Among the states in which the presidents were born, New York was mistakenly spelled with a lower case "Y" in at least one entry.
+\item The variable concerning assassination attempts is coded as a numeric, but the default \code{smartNum = TRUE} setting of \code{makeDataReport()} implies that such a numeric variable with only a few (less that 5) unique values is treated as a factor variable in the data report, thereby providing more relevant summaries, visualizations and checks.
+This is remarked in the variable presentation for \code{assassinationAttempt} and it can also be seen by the visualization being a barplot rather than a histogram.
+\item The variable concerning the sex of the president was skipped, as there is nothing to present, when all US presidents have been male so far.
+\item The report flags the variable \code{ethnicity} to have suspiciously few observations in one category, "African American". 
+\item A few presidents were found to have odd values in the variable describing the duration of their presidencies: Some had very short (outlier) presidencies of less than three years and one was registered to have an \textit{infinite} presidency.
+\item The variable concerning age at inauguration was coded as a character variable, but consists exclusively of numbers and takes on a lot of different variables and therefore, it was flagged as a potentially misclassified numeric variable.
+\item It seems as if one or more presidents have complex numbers as their favorite numbers.
+As \code{complex} is not a supported variable type in \code{dataMaid} and no strategy for handling this class was provided in the \code{treatXasY} argument, the variable is simply flagged as non-supported.
+\end{itemize}
+A lot of these mistakes are easily fixable, and we will do so below.
+However, some of them require more delicate knowledge of the subject matter.
+For instance, \code{ethnicity} is very reasonably marked as a potentially problematic variable as it includes only a single observation of "African American".
+However, a human reading this report will know that this does \textit{not} reflect a mistake in the data, but rather a perculiarity in the real world, and as such, it should not be cleaned out.
+
+A few of the identified problems have easy fixes that need no further attention.
+We remove the prefixed whitespace from Truman's name, fix the misspelling of New York, convert the binary variable \code{assasinationAttempt} to a factor and change the class of the \code{ageAtInauguration} variable to numeric:
+\begin{Schunk}
+\begin{Sinput}
+R> presidentData$lastName[presidentData$lastName == " Truman"] <- "Truman"
+R> presidentData$stateOfBirth[presidentData$stateOfBirth == "New york"] <- "New York"
+R> presidentData$assassinationAttempt <- factor(presidentData$assassinationAttempt)
+R> presidentData$ageAtInauguration <- as.numeric(presidentData$ageAtInauguration)
+\end{Sinput}
+\end{Schunk}%$
+Please note that if \code{ageAtInauguration} had been a \code{factor} rather than a \code{character} variable, an additional inner call should be added in order to ensure no conversion issues:
+\begin{Schunk}
+\begin{Sinput}
+>R presidentData$ageAtInauguration <- as.numeric(as.character(presidentData$ageAtInauguration))
+\end{Sinput}
+\end{Schunk}
+Moving forward, we might be interested in inspecting the contents of the "."-coded entry of \code{firstName} closer, as we \textit{do} know the first names of all the US presidents.
+We look at the last name for this president and fill it in correctly:
+\begin{Schunk}
+\begin{Sinput}
+R> presidentData$lastName[presidentData$firstName == "."]
+\end{Sinput}
+\begin{Soutput}
+[1] "Trump"
+\end{Soutput}
+\begin{Sinput}
+R> presidentData$firstName[presidentData$firstName == "."] <- "Donald"
+\end{Sinput}
+\end{Schunk}
+Next up is the unlikely US president birth day of March 1, 1300.
+In order to understand if this is a generally problematic observation, or if it is just a mistyped observation, we inspect the full data entry for this person.
+We can do this using the usual \proglang{R} selection syntax as above, or we can use the value of a \code{identifyOutliers} call to select this observation:
+\begin{Schunk}
+\begin{Sinput}
+R> birthdayOutlierVal <- identifyOutliers(presidentData$birthday)$problemValues
+\end{Sinput}
+\end{Schunk}
+Now, we have the outlier birthday stored and can use it to select and print the appropriate observation in the dataset:
+\begin{Schunk}
+\begin{Sinput}
+> presidentData[presidentData$birthday == birthdayOutlierVal, ]
+\end{Sinput}
+\begin{Soutput}
+      lastName firstName orderOfPresidency   birthday stateOfBirth
+46 Arathornson   Aragorn                 0 1300-03-01       Gondor
+   assassinationAttempt  sex ethnicity precidencyYears ageAtInauguration
+46                    1 Male Caucasian              NA              <NA>
+   favoriteNumber
+46           8+0i
+\end{Soutput}
+\end{Schunk}%$
+We see that this is not a proper US president and thus, it is likely to be explanation of the faulty number of observations in the dataset.
+Therefore, we drop this observation from the dataset, e.g.
+by overwriting the dataset with a selection of all the other observations:
+\begin{Schunk} 
+\begin{Sinput}
+R> presidentData <- presidentData[presidentData$birthday != birthdayOutlierVal, ]
+\end{Sinput}
+\end{Schunk} %$
+Now, all that is left is the \code{presidencyYears} variable.
+For this variable, we are concerned about three different things: First, it has some quite small outlier values.
+We need to identify whether these are really true.
+Secondly, one president is registered to have an infinite presidency, this should also be fixed.
+Third, we see from the summary table that there are two missing observations for this variable.
+One might have been fixed by removing Aragorn from the dataset, so we start by inspecting if this is indeed the case by calling \code{summarize()} interactively:
+\begin{Schunk}
+\begin{Sinput}
+>R summarize(presidentData$presidencyYears)
+\end{Sinput}
+\begin{Soutput}
+$variableType
+Variable type: numeric
+$countMissing
+Number of missing obs.: 1 (2.22 %)
+$uniqueValues
+Number of unique values: 11
+$centralValue
+Median: 4
+$quartiles
+1st and 3rd quartiles: 4; 8
+$minMax
+Min. and max.: 0; Inf
+\end{Soutput}
+\end{Schunk} %$
+We see that there is one less missing value and that the small and large values pertain. Therefore, we look at all the observations that cause worry, namely the outliers and the missing value:
+\begin{Schunk}
+\begin{Sinput}
+R> presidentData[is.na(presidentData$presidencyYears) | 
++     presidentData$presidencyYears %in% 
++     identifyOutliers(presidentData$presidencyYears)$problemValues, 
++     c("firstName", "lastName", "presidencyYears")]
+\end{Sinput}
+\begin{Soutput}
+   firstName lastName presidencyYears
+9    William Harrison               0
+12   Zachary   Taylor               1
+20     James Garfield               1
+29    Warren  Harding               2
+38    Gerald     Ford               2
+44    Barack    Obama             Inf
+45    Donald    Trump              NA
+\end{Soutput}
+\end{Schunk}
+We see that Obama is listed as being president forever, which history has proven to be wrong.
+Trump, on the other hand, has a missing value for his presidency duration, which is in fact reasonable as we cannot know yet.
+Presidents William Harrison, Zachary Taylor, James Garfield, Warren Harding and Gerald Ford were identified to have were brief presidencies, but these are not mistakes, as a textbook look-up can tell us.
+Thus the only mistake left to fix is Obama's infinite presidency:
+\begin{Schunk}
+\begin{Sinput}
+R> presidentData$presidencyYears[presidentData$lastName == "Obama"] <- 8
+\end{Sinput}
+\end{Schunk}
+This does not mean we are necessarily done with the data cleaning process: There might be problems that \pkg{dataMaid} was not able to identify.
+But we have fixed some key issues in the data and thereby given ourselves a chance of a smoother sailing in the next steps of the data analysis.
 
@@
 \label{sec:specificExamples}
-%\hl{Lidt poppet titel, men den gamle ("Using dataMaid in specific situations")
-%fik mig til flabel at tænke "i modsætning til uspecifikke situationer?". Jeg
-%er dog meget åben for en bedre titel.}
 
@@
 documents as a way to produce clear and concise documentation of a
-dataset. %\hl{I feel like there should be more topics here, but I'm all
-%  out of ideas...}
+dataset.
 

### 2017-10-18 latex/codeForArticle.R

@@ -171,15 +171,22 @@ clean(toyData, standAlone = FALSE, output = "html", twoCols = FALSE)
 
 
 
-############
-makeDataReport(presidentData, replace = TRUE, treatXasY = list("Name" = "character"),
-               checks = setChecks(character = defaultCharacterChecks(remove = "identifyLoners")))
+###########
+makeDataReport(presidentData, replace = TRUE, treatXasY = list(Name = "character"),
+               checks = setChecks(character = defaultCharacterChecks(remove = "identifyLoners")),
+               reportTitle = "Dirty president data")
 
+presidentData$lastName[presidentData$firstName == "."] 
+presidentData$firstName[presidentData$firstName == "."]  <- "Donald"
 
-
-
-
-
+birthdayOutlierVal <- identifyOutliers(presidentData$birthday)$problemValues
+presidentData[presidentData$birthday == birthdayOutlierVal, ]
+presidentData <- presidentData[presidentData$birthday != birthdayOutlierVal, ]
+summarize(presidentData$presidencyYears)
+presidentData[is.na(presidentData$presidencyYears) | 
+                     presidentData$presidencyYears %in% 
+                    identifyOutliers(presidentData$presidencyYears)$problemValues, 
+                   c("firstName", "lastName", "presidencyYears")]
 ######################################################################################
 ##################Code used to make tables and figures in the article#################
 ######################################################################################

### 2017-10-19 latex/article_vol3.tex
Finished first full draft of new article version.

@@
 \ref{sec:usingdataMaid}, we introduce the main representative of the first
-paradigm, namely the \code{clean()} function, which generates data
+paradigm, namely the \code{makeDataReport()} function, which generates data
 overview documents. In the \pkg{dataMaid} package, we have
 provided a number of default checks that cover the data
-cleaning challenges, we find to be most common. Next, in Section
+cleaning challenges, we find to be most common and these are also summarized in Section \ref{sec:usingdataMaid}. Next, in Section
 \ref{sec:interactiveCleanR}, we present the interactive mode of \pkg{dataMaid}, as motivated
-by the second paradigm above. But, as any data analyst knows,
-every dataset is different, and some datasets might include problems
-that cannot be detected by our data checking functions. In Section
-\ref{sec:extending}, we turn to the question of how \pkg{dataMaid} extensions
-can be made, such that they are integrable with the \code{clean()}
-function and with the other tools available in \pkg{dataMaid}.  At last,
-in Section \ref{sec:specificExamples}, we discuss a number of examples of
-specific data cleaning challenges and how \pkg{dataMaid} can be used to
+by the second paradigm above.
+Next up, we show how the data report and the interactive mode of \pkg{dataMaid} can be combined step-by-step to conduct a well-documented, systematic data cleaning in Section \ref{sec:bigExample}.
+Here, we asses and clean a dirty dataset with information about the US presidents.
+At last, in Section \ref{sec:specificExamples}, we discuss a number of examples of
+specific data cleaning- and documentation challenges and how \pkg{dataMaid} can be used to
 solve them.
+
+Please note that this description of what \pkg{dataMaid} contains is usefully supplemented by a tutorial into how it can be extended.
+Every dataset is different, and some datasets might include problems that cannot be detected by our data checking functions.
+Therefore, \pkg{dataMaid} was designed to be easily extendend with user-supplied functions for summarizing, viusalizing and checking data.
+We have provided a vignette \hl{REF TO VIGNETTE} in which we describe how \pkg{dataMaid} extensions can be made, such that they are integrable with the \code{makeDataReport()}
+function and with the other tools available in \pkg{dataMaid}.
 
@@
 of tests are used to generate data reports. The data report itself is an overview document, intended for reading
-by humans, in either pdf or html format. Appendix \ref{sec:appendix1}
+by humans, in either pdf, html or word (.docx) format. Appendix \ref{sec:appendix1}
 provides an example of a data report, produced by
 calling \code{makeDataReport()} on the dataset \code{toyData} available in
-\pkg{dataMaid}. The first two pages of this data report are
+\pkg{dataMaid}. The first two pages (excluding the frontpage) of this data report are
 shown in Figure~\ref{fig:example1}. \code{toyData} is a very
@@
 \end{center}
-\caption{Example output from running \code{makeDataReport()} on the \code{toyData}
-  dataset. First a summary of the full dataset is given and then
-  type-dependent information on each variable is given in a table and
-  a graph. Larger versions of the pages can be seen in
+\caption{The two first pages of the report created by running \code{makeDataReport()} on the \code{toyData}
+  dataset.
+First a summary of the full dataset is given along with an overview of what checks were performed.
+Next, a summary of all the variables aand whether or not they are problematic is provided.
+Larger versions of the pages can be seen in
   Appendix~\ref{sec:appendix1}.}
 \label{fig:example1}
+\end{figure}
+
+\begin{figure}[tb]
+\begin{center}
+\frame{\includegraphics[width=7.5cm,page=4]{dataMaid_toyData.pdf}}
+\frame{\includegraphics[width=7.5cm,page=5]{dataMaid_toyData.pdf}}
+%\includepdf[pages={2}, pagecommand={}]{dataMaid_testData.pdf}
+\end{center}
+\caption{The third and fourth pages of the report created by running \code{makeDataReport()} on the \code{toyData}
+  dataset.
+Here, we see a description of each variable in the dataset, consiting of a summary table, a visualization and a indication of what problems were flagged for the variable (if any).
+At last, a few lines of metadata about the \code{makeDataReport()} are included for enhancing reproducibility.
+Larger versions of the pages can be seen in
+  Appendix~\ref{sec:appendix1}.}
+\label{fig:example2}
 \end{figure}
@@
 \end{Schunk}
+
 \begin{Schunk}
@@
 
-By default, an \proglang{R} markdown file and a rendered pdf overview
+By default, an \proglang{R} markdown file and a rendered pdf \hl{OBS: måske html/doc alt efter hvem der kalder funktionen?} overview
 document is produced, saved to the working directory and
-opened for immediate inspection. Turning to Figure~\ref{fig:example1},
-we see that such a data report consists of three
+opened for immediate inspection. Such a data report consists of three parts, two of which 
 parts. First, an overview of what was done is presented under the
@@
 
-\smallskip Control input variables and summary\\
+\smallskip Control input variables and meta information\\
 \quad \code{useVar} & What variables should be used? & \code{NULL} (corresponding to all variables) \\
@@
 \quad \code{listChecks} & Should an overview of what checks were performed be listed in the \textit{Data report overview}? &  \code{TRUE} \\
-\quad \code{preChecks} & What check functions should be called to determine whether a variable is suitable for summarization, visualization and checking? & \code{c("isKey", "isEmpty")} \\
+\quad \code{preChecks} & What check functions should be called to determine whether a variable is suitable for summarization, visualization and checking?
+& \code{c("isKey", "isSingular", "isSupported")} \\
+\quad \code{reportTitle} & What should the title displayed on the front page of the report be? & \code{NULL} (corresponds to the dataset name) \\
 
@@
                      levels be flagged and treated as a factor variable? & \code{TRUE} \\
-\quad \code{maxProbVals} & Maximum number of problematic values to print, if any are found in data checks & \code{Inf} \\
+\quad \code{maxProbVals} & Maximum number of problematic values to print, if any are found in data checks & \code{10} \\
 \quad \code{maxDecimals} & Maximum number of decimals to print for numeric values in the variable list & \code{2} \\
@@
 \smallskip Control output and post-processing \\
-\quad \code{output} & Type of output file to be produced (html, or pdf) & \code{"pdf"} \\
+\quad \code{output} & Type of output file to be produced (html, word (.docx) or pdf) & \code{"pdf"} \hl{OBS. Ændrer vi dette?}\\
 \quad \code{render} & Should the output file be rendered from markdown? & \code{TRUE} \\
@@
 variables with potential problems, and with a limit of, say, maximum
-10 printed potential errors for each variable. Also, we can add the
+2 printed potential errors for each variable. Also, we can add the
 argument \code{replace=TRUE} in order to force \code{makeDataReport()} to
@@
 \begin{Sinput}
-R> makeDataReport(toyData, output = "html", onlyProblematic = TRUE, maxProbVals = 10,
-+    replace = TRUE)
+R> makeDataReport(toyData, output = "html", onlyProblematic = TRUE,
++    maxProbVals = 2, replace = TRUE)
 \end{Sinput}
@@
   \textbf{\code{summaryFunction}s}  \smallskip \\
-  \quad \code{centralValue} & Compute median or mode &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\
+  \quad \code{centralValue} & Compute median for numeric variables, mode for categorical variables &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\
   \quad \code{countMissing} & Compute proportion of missing observations &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\
@@
  \quad \code{isCPR} & Identify Danish CPR numbers & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &$\times$   \\
- \quad \code{isEmpty} & Check if the variable contains only a single value & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\
+ \quad \code{isSingular} & Check if the variable contains only a single value & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\
  \quad \code{isKey} & Check if the variable is a key & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$\smallskip   \\
+ \quad \code{isSupported} & Check if the variable is among the supported variable types & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$\smallskip   \\
  \hline
@@
 \code{summaryFunction}s, \code{visualFunction}s and
-\code{checkFunction}s.  By default, \code{clean()} runs selected summary,
+\code{checkFunction}s.  By default, \code{makeDataReport()} runs selected summary,
 visualization and check functions on each variable in the dataset, and
@@
 -------------- ------------------------------- -----------------------------
-centralValue   Compute median or mode          character, Date, factor,     
-                                               integer, labelled, logical,  
-                                               numeric
+centralValue   Compute median for numeric      character, Date, factor,     
+               variables, mode for             integer, labelled, logical,  
+               categorical variables           numeric
 
@@
 R> makeDataReport(toyData, 
- 	summaries = setSummaries(numeric = c("variableType","centralValue),
-		integer = c("variableType", "centralValue"))
++    summaries = setSummaries(numeric = c("variableType","centralValue),
++    integer = c("variableType", "centralValue"))
 \end{Sinput}
@@
 R> makeDataReport(toyData, 
-	summaries = setSummaries(all = c("variableType", "centralValue"))
++    summaries = setSummaries(all = c("variableType", "centralValue"))
 \end{Sinput}
@@
 \begin{Sinput}
-R> makeDataReport(toyData, checks = setChecks(factor = "identifyWhitespace"))
+R> makeDataReport(toyData, 
++    checks = setChecks(factor = "identifyWhitespace"))
 \end{Sinput}
@@
 \code{checkFunction}s that are applicable to all variable classes. The
-default pre-checks, the functions \code{isKey()} and \code{isSingular()}, check
-whether a variable has unique values for all observations or only a
-single value for all observations, respectively. If one of these
+default pre-checks, the functions \code{isKey()}, \code{isSingular()} and \code{isSupported()}, check
+whether a variable has unique values for all observations, only a
+single value for all observations, and is not among the variable types supported by \pkg{dataMaid}, respectively. If one of these
 statements are true, the variable will not be subjected to the SVC
-steps.  We can allow empty variables to move on to the SVC step by
-only checking for keys in the pre-check step:
-
-\begin{Schunk}
-\begin{Sinput}
-R> makeDataReport(toyData, preChecks = "isKey")
-\end{Sinput}
-\end{Schunk}
-
-Note that the data visualizations in the cleaning summary are also
+steps.  We can allow singular variables to move on to the SVC step by
+only checking for keys and non-supported variables in the pre-check step:
+
+\begin{Schunk}
+\begin{Sinput}
+R> makeDataReport(toyData, preChecks = c("isKey", "isSupported"))
+\end{Sinput}
+\end{Schunk}
+
+Note that the data visualizations in the report are also
 controllable, though only a single function can be provided for each variable type. If, for instance, we wish to change the visualizations
@@
 
-\section[Using dataMaid interactively]{\hl{Using \pkg{dataMaid} interactively}}
+One might also encounter datasets with variables that are not among the 7 classes mentioned in the above (\code{character}, \code{Date}, \code{factor}, \code{integer}, \code{labelled}, \code{logical} and \code{numeric}), for instance variables of type \code{complex} or user-defined classes.
+It is possible to tell \code{makeDataReport()} how to handle such variables by use of the argument \code{treatXasY}.
+This argument takes a list where the names correspond to "new" variable types, while the entries must be supported variable types.
+For instance, we can instruct \code{dataMaid} to treat complex variables as numerics and generate a data report for a type \code{complex} variable like this:
+
+\begin{Schunk}
+\begin{Sinput}
+>R complexData <- data.frame(complexVar = complex(100, real = 1:100, 
++    imaginary = 3), numericVar = 1:100))
+>R makeDataReport(complexData, treatXasY(complex = "numeric"))
+\end{Sinput}
+\end{Schunk}
+
+\section[Using dataMaid interactively]{Using \pkg{dataMaid} interactively}
 \label{sec:interactiveCleanR}
 
-While overview documents are great for presenting and documenting the
-data cleaning checks, it may be useful to be able to work more
-interactively through the data cleaning process. Aside from
-the \code{clean()} function presented above, \pkg{dataMaid} also
+While overview documents are great for presenting and documenting the data at various stages 
+of the data cleaning process, it may be useful to be able to work more
+interactively when performing actual data cleaning. Aside from
+the \code{makeDataReport()} function presented above, \pkg{dataMaid} also
 provides more standard \proglang{R} interactive tools, such as
@@
 and thus they can also be used here. For instance, if we only want to
-check for potentially miscoded missing values, we can use the relevant
-\code{XXXChecks} argument (e.g., \code{numericChecks}, \code{factorChecks},
-etc., as described in Section~\ref{sec:usingdataMaid}) to provide a
-vector of the check functions that should be applied. Recall that
-Table~\ref{table.SVCfunctions} or an \code{allCheckFunctions()} call provide
+check for potentially miscoded missing values, we can use the \code{checks} argument and the \code{setChecks()} helper function to specify this.
+Recall that Table~\ref{table.SVCfunctions} or an \code{allCheckFunctions()} call provide
 overviews of the available check functions.
@@
 \begin{Sinput}
-R> check(toyData$var2, numericChecks = "identifyMissing")
+R> check(toyData$var2, checks = setChecks(numeric = "identifyMissing"))
 \end{Sinput}
@@
 \end{Soutput}
-\end{Schunk}
+\end{Schunk} %$
 
@@
 message is made for easy reading, the actual values of the variable
-causing the issue are still obtainable in the element
+causing the issue are still obtainable in the entry
 \code{problemValues}. If we decide that the values \code{999}
@@
 As we saw with the \code{check()} function, the summary can be modified
-by providing the relevant \code{XXXSummaries} arguments. Setting the
-\code{numericSummaries} argument, we can control the summary output by
-providing a vector of function names to run for a particular summary. To
-only get the median, minimum and the maximum we set
-\code{numericSummaries = c("centralValue", "minMax")}:
-
-\begin{Schunk}
-\begin{Sinput}
-R> summarize(toyData$var2, numericSummaries = c("centralValue", "minMax"))
+by using the \code{summaries} argument and the \code{setSummaries()} helper function.
+If we want to remove the default summaries \code{variableType} and \code{countMissing} for numeric variables, we can use the function \code{defaultNumericSummaries()} and its argument \code{remove} that excludes a vector of summaries from the usual default summaries:
+
+\begin{Schunk}
+\begin{Sinput}
+R> summarize(toyData$var2, 
++    summaries = setSummaries(
++    numeric = defaultNumericSummaries(remove = c("variableType", 
++    "countMissing"))))
 \end{Sinput}
@@
 \end{Soutput}
-\end{Schunk}
+\end{Schunk} %$
+Similar \code{defaultXXXSummaries()} functions are available for the other supported variable classes.
+For checks, the same syntax can also be used, but the helper functions are now named \code{defaultXXXChecks} with \code{XXX} as a placeholder for a supported variable class.
 
@@
 \frame{\includegraphics[width=7.5cm,page=2]{dataMaid_presidentData.pdf}}\\
+\end{center}
+\caption{The front page and the first page of the data overview report for the \code{presidentData} dataset.
+Note that the report title has been customized (front page), \code{identifyLoners} has been removed from the checks performed on character variables (page 1) and that variables of class \code{Name} have been set to be treated like \code{character} variables (page 1).
+Larger versions of the pages can be seen in
+  Appendix~\ref{sec:appendix2}.}
+\label{fig:bigExampleP01}
+\end{figure}
+
+\begin{figure}[tb]
+\begin{center}
 \frame{\includegraphics[width=7.5cm,page=3]{dataMaid_presidentData.pdf}}
@@
 \end{center}
-\caption{The first 4 pages of the data overview report for the \code{presidentData} dataset.
-Note that the report title has been customized (front page), \code{identifyLoners} has been removed from the checks performed on character variables (page 1) and that variables of class \code{Name} have been set to be treated like \code{character} variables (page 1).
-We see that there are two \code{Name} variables in the overview on page 2 and see that these variables are indeed treated like \code{character} variables on page 3.
-}
-\label{fig:bigExampleP0123}
+\caption{The second and third pages of the \code{presidentData} data report.
+We see that there are two \code{Name} variables in the overview on page 2 and see that these variables are indeed treated as \code{character} variables on page 3, as specified in the \code{makeDataReport} call by use of the \code{treatXasY} argument.
+Larger versions of the pages can be seen in
+  Appendix~\ref{sec:appendix2}}
+\label{fig:bigExampleP23}
 \end{figure}
@@
 \end{center}
-\caption{Part of the data overview report generated for the \code{presidentData} dataset. Here, we see pages 4 and 5 from the \textit{Variable list}.}
+\caption{Part of the data overview report generated for the \code{presidentData} dataset. Here, we see pages 4 and 5 from the \textit{Variable list}. Larger versions of the pages can be seen in
+  Appendix~\ref{sec:appendix2}}
 \label{fig:bigExampleP45}
@@
 Incoorporating these three customizations, we can load the data and generate a report by calling:
+
 \begin{Schunk}
@@
 R> makeDataReport(presidentData, replace = TRUE, 
-   treatXasY = list("Name" = "character"),
-   checks = setChecks(character = defaultCharacterChecks(remove = "identifyLoners")),
-   reportTitle = "Dirty president data")
-\end{Sinput}
-\end{Schunk}
-The first four pages of the resulting report can be inspected in Figure \ref{fig:bigExampleP0123}.
++    treatXasY = list("Name" = "character"),
++    checks = setChecks(character = defaultCharacterChecks(remove = "identifyLoners")),
++    reportTitle = "Dirty president data")
+\end{Sinput}
+\end{Schunk}
+
+The first four pages of the resulting report can be inspected in Figures \ref{fig:bigExampleP01} and \ref{fig:bigExampleP23}.
 Note that all the customized settings can be identified from the first two pages, without having to read through the report: The new title is written on the front page, the check settings are displayed in the \textit{Checks performed} table and the strategy for handling \code{Name} variables is documented below this table.
+The full data report, except for the front page, is available in Appendix \ref{sec:appendix2}.
 
@@
 We remove the prefixed whitespace from Truman's name, fix the misspelling of New York, convert the binary variable \code{assasinationAttempt} to a factor and change the class of the \code{ageAtInauguration} variable to numeric:
-\begin{Schunk}
-\begin{Sinput}
-R> presidentData$lastName[presidentData$lastName == " Truman"] <- "Truman"
-R> presidentData$stateOfBirth[presidentData$stateOfBirth == "New york"] <- "New York"
-R> presidentData$assassinationAttempt <- factor(presidentData$assassinationAttempt)
-R> presidentData$ageAtInauguration <- as.numeric(presidentData$ageAtInauguration)
+
+\begin{Schunk}
+\begin{Sinput}
+R> presidentData$lastName[presidentData$lastName == " Truman"] <- 
++    "Truman"
+R> presidentData$stateOfBirth[presidentData$stateOfBirth == "New york"] <- 
++    "New York"
+R> presidentData$assassinationAttempt <- 
++    factor(presidentData$assassinationAttempt)
+R> presidentData$ageAtInauguration <- 
++    as.numeric(presidentData$ageAtInauguration)
 \end{Sinput}
 \end{Schunk}%$
+
 Please note that if \code{ageAtInauguration} had been a \code{factor} rather than a \code{character} variable, an additional inner call should be added in order to ensure no conversion issues:
-\begin{Schunk}
-\begin{Sinput}
->R presidentData$ageAtInauguration <- as.numeric(as.character(presidentData$ageAtInauguration))
-\end{Sinput}
-\end{Schunk}
+
+\begin{Schunk}
+\begin{Sinput}
+>R presidentData$ageAtInauguration <- 
++    as.numeric(as.character(presidentData$ageAtInauguration))
+\end{Sinput}
+\end{Schunk}
+
 Moving forward, we might be interested in inspecting the contents of the "."-coded entry of \code{firstName} closer, as we \textit{do} know the first names of all the US presidents.
 We look at the last name for this president and fill it in correctly:
+
 \begin{Schunk}
@@
 \end{Schunk}
+
 Next up is the unlikely US president birth day of March 1, 1300.
@@
 We can do this using the usual \proglang{R} selection syntax as above, or we can use the value of a \code{identifyOutliers} call to select this observation:
+
 \begin{Schunk}
@@
 \end{Schunk}
+
 Now, we have the outlier birthday stored and can use it to select and print the appropriate observation in the dataset:
+
 \begin{Schunk}
@@
 \end{Schunk}%$
+
 We see that this is not a proper US president and thus, it is likely to be explanation of the faulty number of observations in the dataset.
@@
 by overwriting the dataset with a selection of all the other observations:
-\begin{Schunk} 
-\begin{Sinput}
-R> presidentData <- presidentData[presidentData$birthday != birthdayOutlierVal, ]
+
+\begin{Schunk} 
+\begin{Sinput}
+R> presidentData <- 
++    presidentData[presidentData$birthday != birthdayOutlierVal, ]
 \end{Sinput}
 \end{Schunk} %$
+
 Now, all that is left is the \code{presidencyYears} variable.
@@
 One might have been fixed by removing Aragorn from the dataset, so we start by inspecting if this is indeed the case by calling \code{summarize()} interactively:
+
 \begin{Schunk}
@@
 \end{Schunk} %$
+
 We see that there is one less missing value and that the small and large values pertain. Therefore, we look at all the observations that cause worry, namely the outliers and the missing value:
+
 \begin{Schunk}
@@
 \end{Schunk}
+
 We see that Obama is listed as being president forever, which history has proven to be wrong.
@@
 Thus the only mistake left to fix is Obama's infinite presidency:
+
 \begin{Schunk}
@@
 \end{Schunk}
+
 This does not mean we are necessarily done with the data cleaning process: There might be problems that \pkg{dataMaid} was not able to identify.
 But we have fixed some key issues in the data and thereby given ourselves a chance of a smoother sailing in the next steps of the data analysis.
+Finally, we create a new data report, adding the suffix "cleaned" to the title, as well as the file name, so that we have documentation of the current state of the dataset:
+
+\begin{Schunk}
+\begin{Sinput}
+R> makeDataReport(presidentData, vol = "_cleaned", 
++   reportTitle = "Dirty president data - cleaned")
+\end{Sinput}
+\end{Schunk}
+
+This will create a new data report stored in the file \code{dataMaid_presidentData_cleaned.pdf}.
 
@@
 Finally, we present a few examples of how to make \pkg{dataMaid}
-solve specific issues related to data cleaning. First, we discuss the
-challenges related to cleaning large datasets, particularly in terms
-of memory use and computation speed. Next, we show how \pkg{dataMaid}
-can be used for problem-flagging. Lastly, we discuss how the
+solve specific issues related to data documentation and -cleaning.
+First, we discuss how the data report generation functions of \pkg{dataMaid} can be used in a data science workflow where one is not necessarily interested in inspecting the results right away and most commands are run automatically.
+Next, we show how \pkg{dataMaid} can be used for problem-flagging.
+Lastly, we discuss how the
 \pkg{dataMaid} output document can be included in other \proglang{R} markdown
@@
 
-\subsection{Cleaning large datasets}
-If the dataset becomes very large, the standard use of \code{clean()}
-outlined above might not be ideal. If there is a large number of
-variables, creation of the \proglang{R} markdown document might be quite
-slow, while a large number of observations will generally affect the
-rendering time of the document. This issue is especially relevant to
-Windows users, where memory usage is not very efficient. The problem is that older,
-32-bit versions of Windows generally cannot access all physical RAM,
-while newer 64-bit versions of Windows cannot access virtual memory. Therefore,
-memory allocation issues becomes even more pressing under Windows.
-
-Below we give a few practical examples of ways to deal with large
-data, while wishing to still produce (potentially very long) data
-cleaning overview documents. Note that the interactive tools of
-\pkg{dataMaid} can be used as usual or sequentially in small subsets
-of the large dataset, if no overview documents are needed.
-
-\subsubsection{Handling the figures}
-Though figures give a nice overview of each variable, they are also
-quite heavy objects in terms of memory allocation. Therefore, it might
-be beneficial to not include figures in the \pkg{dataMaid} outputs for
-very large datasets. This is controlled via the \code{mode} argument:
-
-\begin{Schunk}
-\begin{Sinput}
-R> clean(toyData, mode = c("summarize", "check"))
-\end{Sinput}
-\end{Schunk}
-
-If figures are indeed needed, a different approach is to choose the
-less memory heavy standard \proglang{R} figure style instead of the
-\pkg{ggplot2} figures that are the default option in \code{clean()}. This
-can be done by setting the \code{allVisuals = "basicVisual"} argument:
-
-\begin{Schunk}
-\begin{Sinput}
-R> clean(toyData, allVisuals = "basicVisual")
-\end{Sinput}
-\end{Schunk}
-
-Of course, even less heavy plots might be achieved by writing new
-\code{visualFunction}s, using the guidelines from Section
-\ref{sec:functionTemplates}. For instance, a future extension of
-\pkg{dataMaid} might be the inclusion of ASCII plots, as
-e.g., represented in the \proglang{R} package \pkg{txtplot} \citep{txtplot}.
-
-%\hl{I really feel like we should do some benchmarking here, maybe just on toyData, both in terms of speed and memory use. I would make the recommendations more trustworthy and serious.}
-
-\subsubsection{Economic memory use}
-Another solution, which is
-especially relevant to Windows users due to %the unfortunate
-the memory handling on this operating system,
-%combination of memory control in this operating system and
-%\proglang{R},
-is simply splitting the two steps performed
-by \code{clean()}, namely producing the \proglang{R} markdown file and rendering it
-afterwards. If the \code{rmarkdown} file is very long, as it will
-typically be for very large datasets, keeping this file open in memory
-wastes precious memory capacities. Therefore, we advise users to
-instead split the two steps as shown in the following.
-
-\begin{Schunk}
-\begin{Sinput}
-R> clean(toyData, render = FALSE, openResult = FALSE)
-R> render("dataMaid_toyData.Rmd", quiet = FALSE)
-\end{Sinput}
-\end{Schunk}
-
-This also deals with the fact that \pkg{dataMaid} can produce
-\proglang{R} markdown files that supersedes the upper size limit of code editors, e.g.,
-RStudio, which currently has an editor file size limit of 5MB. %\hl{dette giver
-%ingen mening uden en præcision? "upper size limit" af hvad? Se udkommenteret
-%kommentar.}
-% of RStudio,
-%which is currently \hl{find number} GBs (using RStudio version
-%1.0.44).
-%\hl{Is this maybe too editor specific? On the other hand, a
-%  lot of people do use RStudio...}.
+\subsection[Incoorporating dataMaid in a programmer's workflow]{Incoorporating \pkg{dataMaid} in a programmer's workflow}
+The default settings of \code{makeDataReport()} have been set to facilitate easy and quick data report generation, but unfortunately, this also means that they might not be ideal for a more programming-oriented workflow, where the function might not be called by a human, but is part of a larger scheme.
+For instance, one might be interested in automatically running \code{makeDataReport()} on all datasets received from a certain client and returning a data report for them to inspect and comment before ever looking at the data.
+In this scenario, there are a few issues with the standard data report:
+\begin{enumerate}
+\item After rendering, the report is automatically opened. This is not very useful, if the processes are supposed to run in the background.
+\item The report generation produces messages to the console while producing and rendering the report.
+\item Unless specifically told otherwise, every report created for identically named datasets will have the same file name.
+\end{enumerate}
+Please note that the data report does contain information about who, when and how concerning its generation.
+These three problems can easily be solved by use of the arguments of \code{makeDataReport}.
+Whether or not the outputted file is opened can be controlled through the argument \code{open}.
+How much information is printed in the console can be adjusted by using the argument \code{quiet}.
+And convenintely introducing small alterations of the file names can be obtained by use of the \code{vol} argument.
+For instance, we can make a data report for \code{toyData} that is not opened automatically, produces no output to the console and includes the date and time of its creation in the file name:
+
+\begin{Schunk}
+\begin{Sinput}
+>R makeDataReport(toyData, open = FALSE, quiet = "silent",
++   vol = paste("_", format(Sys.time(), "%m-%d%-%y_%H.%M"), sep = ""))
+\end{Sinput}
+\end{Schunk}
+
+Now, if e.g. the report is created at 3 pm on October 31, 2017, its file will be named \code{dataMaid_toyData_10-31-2017_15.00.pdf}, making it easy spot.
 
@@
 cleaning document, a viable solution might be not to include all
-information about all variables. % Or even for more reasonably sized
-% datasets, sometimes a brief overview of the most pressing issues can
-% be useful.
-This can be achieved by using the \code{onlyProblematic} argument in
-\code{clean()}. By specifying \code{onlyProblematic = TRUE}, only
+information about all variables. This can be achieved by using the \code{onlyProblematic} argument in
+\code{makeDataReport()}. By specifying \code{onlyProblematic = TRUE}, only
 variables that raise a flag in the checking steps will be summarized
@@
 \begin{Sinput}
-R> clean(toyData, onlyProblematic = TRUE, mode = c("check"))
+R> makeDataReport(toyData, onlyProblematic = TRUE, mode = c("check"))
 \end{Sinput}
@@
 generated by also leaving out the checking results --- then
-\code{clean()} essentially just produces a document with a list of the variable names
+\code{makeDataReport()} essentially just produces a document with a list of the variable names
 that should be investigated further:
@@
 \begin{Sinput}
-R> clean(toyData, onlyProblematic = TRUE, mode = NULL)
+R> makeDataReport(toyData, onlyProblematic = TRUE, mode = NULL)
 \end{Sinput}
@@
 and we find that only a single variable in \code{toyData}, \code{var6} (for which all
-observations have the value \code{"Irrelevant"}), is problem-free.
-
-\subsection[Including dataMaid documents in other files]{Including \pkg{dataMaid} documents in other files}
-Sometimes, a \pkg{dataMaid} document might be a useful addition to a
+observations have the value \code{"Irrelevant"}), is problem-free, using the default checks.
+
+\subsection[Including dataMaid reports in other files]{Including \pkg{dataMaid} reports in other files}
+Sometimes, a \pkg{dataMaid} report might be a useful addition to a
 more general overview document, including additional information such
@@
 included in other \proglang{R} markdown files by setting the
-\code{standAlone} argument in \code{clean()} to remove the preamble
+\code{standAlone} argument in \code{makeDataReport()} to remove the preamble
 from the output \proglang{R} markdown file. Note that it is still
@@
 created; the pdf and html \proglang{R} markdown styles are
-unfortunately not identical.
+unfortunately not identical. Note that the \code{word} output option is based on the \code{html} markdown style.
 
 If it is important that the embedded \pkg{dataMaid} document can be
-rendered to either of these two file types, we recommend setting
-\code{twoCols = FALSE} and \code{output = "html"} in \code{clean()}, thereby
+rendered to any of these three file types, we recommend setting
+\code{twoCols = FALSE} and \code{output = "html"} in \code{makeDataReport()}, thereby
 essentially removing almost all output type specific formatting code
@@
 }
-%\hl{Use proper formatting here. How do we do non-R code?}
 
@@
 \begin{Sinput}
-R> clean(toyData, standAlone = FALSE)
+R> makeDataReport(toyData, standAlone = FALSE)
 \end{Sinput}
@@
 \begin{Sinput}
-R> clean(toyData, standAlone = FALSE, output = "html", twoCols = FALSE)
+R> makeDataReport(toyData, standAlone = FALSE, output = "html", 
++    twoCols = FALSE)
 \end{Sinput}
@@
 Note that with the latter option, no special YAML preamble is needed in the \proglang{R} markdown document.
-% \hl{don't end on code.}
+
+Alternatively, one can create the usual output report, not render it and then manually edit the produced rmarkdown file as wished. This can be done with the following command:
+If one works with a dataset containing a very large number of variables, it can be practical to split up two of the procedures undertaken by the \code{makeDataReport()} function, namely creating- and rendering the data report.
+\begin{Schunk}
+\begin{Sinput}
+R> makeDataReport(toyData, render = FALSE, openResult = FALSE)
+\end{Sinput}
+\end{Schunk}
+After editing, the file can be rendered by calling the \code{render} function:
+\begin{Schunk}
+\begin{Sinput}
+R> render("dataMaid_toyData.Rmd", quiet = FALSE)
+\end{Sinput}
+\end{Schunk}
 
@@
 \label{conclusion}
-%\hl{Is this a thing in this journal? Otherwise, we might want to make some final remarks in the previous sections. Feels awkward to end with a bunch of code and some super specific examples...}
 
 In this paper we have introduced the \proglang{R} package \pkg{dataMaid}
-for performing reproducible error detection and data cleaning
-summaries. The package provides a general and extendable framework for
+for performing reproducible error detection, data overview reports and data cleaning. The package provides a general and extendable framework for
 identifying potential errors and for creating human-readable summary
@@
 \proglang{R}-savvy can upload their data and get a data cleaning
-document is currently planned.
+document is currently being developed.
 
@@
 % \section{Appendix A: cleaning the toyData data frame} \label{sec:appendix1}
-\includepdf[pages=2, pagecommand={\section{Appendix A: cleaning the toyData data frame} \label{sec:appendix1}}, frame=true, noautoscale=true, scale=0.7]{dataMaid_toyData.pdf}
+\includepdf[pages=2, pagecommand={\section{Data report for the toyData dataset} \label{sec:appendix1}}, frame=true, noautoscale=true, scale=0.7]{dataMaid_toyData.pdf}
 \includepdf[pages=3-4, pagecommand={}, frame=true, noautoscale=true, scale=0.7,pagecommand={}]{dataMaid_toyData.pdf}
 
-\includepdf[pages=2, pagecommand={\section{Appendix B: User-extended cleaning of
-  exampleData} \label{sec:appendix2}}, frame=true, noautoscale=true, scale=0.7]{dataMaid_exampleData.pdf}
-\includepdf[pages=3-, pagecommand={}, frame=true, noautoscale=true, scale=0.7]{dataMaid_exampleData.pdf}
-%\hl{Data cleaning with user supplied extensions here}
+\includepdf[pages=2, pagecommand={\section{Data report for the presidentData dataset} \label{sec:appendix2}}, frame=true, noautoscale=true, scale=0.7]{dataMaid_presidentData.pdf}
+\includepdf[pages=3-6, pagecommand={}, frame=true, noautoscale=true, scale=0.7]{dataMaid_presidentData.pdf}
 

### 2017-10-19 latex/codeForArticle.R

@@ -187,6 +187,11 @@ presidentData[is.na(presidentData$presidencyYears) |
                      presidentData$presidencyYears %in% 
                     identifyOutliers(presidentData$presidencyYears)$problemValues, 
                    c("firstName", "lastName", "presidencyYears")]
+
+
+
+
+
 ######################################################################################
 ##################Code used to make tables and figures in the article#################
 ######################################################################################
@@ -216,3 +221,58 @@ checkTab <- data.frame(name=allCheckFunctions()$name,
 checkTab <- as.matrix(checkTab)[,-1]
 xtable(checkTab)
 
+
+#########################################################################################################
+########DELETE THIS?#####################################################################################
+#########################################################################################################
+library(microbenchmark)
+data(toyData)
+microbenchmark(makeDataReport(toyData, replace = TRUE, open = FALSE, quiet = "silent"),
+               makeDataReport(toyData, replace = TRUE, mode = c("summarize", "check"), 
+                              open = FALSE, quiet = "silent"),
+               makeDataReport(toyData, replace = TRUE, 
+                              visuals = setVisuals(all = "basicVisual"), open =FALSE, quiet = "silent"),
+               times = 10)
+#Unit: seconds
+#expr
+#makeDataReport(toyData, replace = TRUE, open = FALSE)
+#makeDataReport(toyData, replace = TRUE, mode = c("summarize",      "check"), open = FALSE)
+#makeDataReport(toyData, replace = TRUE, visuals = setVisuals(all = "basicVisual"),      open = FALSE)
+#min       lq     mean   median       uq      max neval
+#8.218834 23.24083 24.59940 25.06826 27.71859 47.79230    10
+#20.093234 21.12455 22.01202 22.26677 22.95272 24.34090    10
+#5.416714 20.88525 19.63562 22.73859 24.23288 24.60682    10
+
+data(diamonds)
+microbenchmark(makeDataReport(diamonds, replace = TRUE, open = FALSE, quiet = "silent"),
+               makeDataReport(diamonds, replace = TRUE, mode = c("summarize", "check"), 
+                              open = FALSE, quiet = "silent"),
+               makeDataReport(diamonds, replace = TRUE, 
[…]

### 2017-10-20 latex/article_vol3.tex
Small fixes in the article.

@@
 \newcommand{\hl}[1]{\textcolor{magenta}{#1}}
+\newcommand{\blue}[1]{\textcolor{blue}{#1}}
 
@@
 %% almost as usual
-\author{Anne Helby Petersen\\Biostatistics\\Department of Public
-  Health\\University of Copenhagen \And Claus Thorn Ekstr\o m\\Biostatistics\\Department of Public
+\author{Anne Helby Petersen\\Section of Biostatistics\\Department of Public
+  Health\\University of Copenhagen \And Claus Thorn Ekstr\o m\\Section of Biostatistics\\Department of Public
   Health\\University of Copenhagen}
@@
   Claus Thorn Ekstr\o m\\
-  Biostatistics, Department of Public Health\\
+  Section of Biostatistics, Department of Public Health\\
   University of Copenhagen\\
@@
 But even when tools are available for identifying problems in a dataset, the activity of data cleaning still suffers from a challenge that has recieved increasing attention in the scientitic communities in the later years: Data cleaning is not very straight forward to document and therefore, reproducibility suffers.
-\pkg{dataMaid}'s most central purpose is to facilitate a data cleaning workflow where documentation is thoroughly integrated rather than a secondary add-on.
+We present a new \proglang{R} package, \pkg{dataMaid} \citep{dataMaid}, whose most central purpose is to facilitate a data cleaning workflow where documentation is thoroughly integrated rather than a secondary add-on.
 This is accomplished by centering the data cleaning process around auto-generated data overview reports that summarize the contents of the current version of a dataset and flags potential problems that could be candidates for the next bullet on the cleaning agenda.
 
-A number of \proglang{R} packages made for other pre-analysis tools are
-already available, including \pkg{janitor} \citep{janitor}, \pkg{dplyr}
+A number of \proglang{R} packages made for other pre-analysis steps are
+already available, including \pkg{janitor} \citep{janitor}, \hl{assertive}, \pkg{dplyr}
 \citep{dplyr}, \pkg{tidyr}\citep{tidyr}, \pkg{data.table} \citep{data.table}, \pkg{DataCombine}
@@
 with a particular emphasis on the challenges of getting neat data
-frames from Microsoft Excel data files. \pkg{dplyr}, \pkg{tidyr}, \pkg{data.table}
+frames from Microsoft Excel data files. \hl{Mention assertive here}. \pkg{dplyr}, \pkg{tidyr}, \pkg{data.table}
 and \pkg{DataCombine} go a few steps further by providing a wide array
@@
 have been made to easily document exactly which checks and preliminary
-results were used in the data cleaning process. \hl{Something about \pkg{assertr} here.}
-
-All in all, the large role of data cleaning in any data analysts
-everyday endeavors is hardly matched in the amount of available
-\proglang{R} software solutions. In particular, few packages attempt
-to implement systematic, reproducible data cleaning. And while the
-available tools attempt to alleviate the ubiquitous ad hoc approach to
-data cleaning, they are primarily intended for the data savvy users
-and less so for the general researcher with a knowledge about a
-specific field and the context of the available data. The
-\pkg{dataMaid} package \citep{dataMaid} presented here tries to address this by
-providing a framework that both allows for extendable, systematic,
-reproducible data cleaning, and summarizing findings for researchers
-from other fields such that they can act as human experts when
-tracking down potential errors.
+results were used in the data cleaning process.
+The \pkg{assertr} package provide very similar - and very nice - tools to those of \pkg{validate}, but without any amibitions of conducting auto-cleaning.
+
+%All in all, the large role of data cleaning in any data analysts
+%everyday endeavors is hardly matched in the amount of available
+%\proglang{R} software solutions. In particular, few packages attempt
+%to implement systematic, reproducible data cleaning. And while the
+%available tools attempt to alleviate the ubiquitous ad hoc approach to
+%data cleaning, they are primarily intended for the data savvy users
+%and less so for the general researcher with a knowledge about a
+%specific field and the context of the available data. The
+%\pkg{dataMaid} package \citep{dataMaid} presented here tries to address this by
+%providing a framework that both allows for extendable, systematic,
+%reproducible data cleaning, and summarizing findings for researchers
+%from other fields such that they can act as human experts when
+%tracking down potential errors.
+%Jeg har slettet dette: Synes det dækkes ok af det nye introafsnit før præsentationen af andre pakker og så er introduktionen ikke blevet længere alt i alt.
 
@@
 provided a number of default checks that cover the data
-cleaning challenges, we find to be most common and these are also summarized in Section \ref{sec:usingdataMaid}. Next, in Section
+cleaning challenges we find to be most common and these are also summarized in Section \ref{sec:usingdataMaid}. Next, in Section
 \ref{sec:interactiveCleanR}, we present the interactive mode of \pkg{dataMaid}, as motivated
 by the second paradigm above.
-Next up, we show how the data report and the interactive mode of \pkg{dataMaid} can be combined step-by-step to conduct a well-documented, systematic data cleaning in Section \ref{sec:bigExample}.
-Here, we asses and clean a dirty dataset with information about the US presidents.
+Next, we show atep-by-step how the data report and the interactive mode of \pkg{dataMaid} can be combined to conduct a well-documented, systematic data cleaning in Section \ref{sec:bigExample}.
+Here, we assess and clean a dirty dataset with information about the US presidents.
 At last, in Section \ref{sec:specificExamples}, we discuss a number of examples of
@@
 function and with the other tools available in \pkg{dataMaid}.
-
-% \hl{Is there a Section 6?}
-
-%\hl{Do they typically discuss notation in articles like this?
-I try (somewhat inconsistently) to refer to functions as \R{function()} while other \R{R} objects are referred to merely as \R{object}.
-This is also the style adopted by Wickham in his books.}
 
@@
 \pkg{dataMaid}. The first two pages (excluding the frontpage) of this data report are
-shown in Figure~\ref{fig:example1}. \code{toyData} is a very
+shown in Figure~\ref{fig:example1} and the following two pages are shown in Figure~\ref{fig:example2}. \code{toyData} is a very
 small ($15$ observations of $6$ variables), artificial dataset which was created to
@@
   dataset.
-First a summary of the full dataset is given along with an overview of what checks were performed.
-Next, a summary of all the variables aand whether or not they are problematic is provided.
+First, a summary of the full dataset is given along with an overview of what checks were performed.
+Next, a summary of all the variables and whether or not they are problematic is provided.
 Larger versions of the pages can be seen in
@@
 opened for immediate inspection. Such a data report consists of three parts, two of which 
-parts. First, an overview of what was done is presented under the
-title \textit{Data report overview}. Secondly, an index listing each variable along with an indication of whether it was found to be problematic or not is provided. Thirdly, each variable in the
+are presented in Figure~\ref{fig:example1}. First, an overview of what was done is presented under the
+title \textit{Data report overview}.
+Secondly, an index listing each variable along with an indication of whether it was found to be problematic or not is provided.
+Thirdly, as seen in Figure~\ref{fig:example2}, each variable in the
 dataset is presented in turn using (up to) three tools in the
@@
 variable, a figure visualizing its distribution and a
-list of flagged issues, if any. For instance, in the \code{numeric}-type variable
+list of flagged issues, if any. For instance, as shown in Figure~\ref{fig:example2}, for the \code{numeric}-type variable
 \code{var2} from \code{toyData}, \code{makeDataReport()} has identified two values that
@@
 
-\smallskip Control input variables and meta information\\
+\smallskip Control input variables, looks and meta information\\
 \quad \code{useVar} & What variables should be used? & \code{NULL} (corresponding to all variables) \\
@@
 \quad \code{onlyProblematic} & Should only variables flagged as problematic be included in the \textit{Variable list}? & \code{FALSE} \\
-\quad \code{listChecks} & Should an overview of what checks were performed be listed in the \textit{Data report overview}? &  \code{TRUE} \\
+%\quad \code{listChecks} & Should an overview of what checks were performed be listed in the \textit{Data report overview}? &  \code{TRUE} \\
+%slettet for at få plads til nye argumenter
 \quad \code{preChecks} & What check functions should be called to determine whether a variable is suitable for summarization, visualization and checking?
@@
 \quad \code{reportTitle} & What should the title displayed on the front page of the report be? & \code{NULL} (corresponds to the dataset name) \\
+\quad \code{twoCol} & Should the summary table and visualizations be placed side-by-side (in two columns)? & \code{TRUE} \\
 
 \smallskip Control summarize, visualize, and check steps \\
+\quad \code{summaries} & What summaries should be performed for each variable type? & See Table~\ref{table.SVCfunctions} \\
+\quad \code{visuals} & What type of visualization should be provided for each variable type? & See Table~\ref{table.SVCfunctions} \\
+\quad \code{checks} & What checks should be applied to each variable type? & See Table~\ref{table.SVCfunctions} \\
 \quad \code{mode} & What steps should be performed for each variable (out
@@
                  \textit{visualize}, \textit{check})? &
-                                                        \code{c("summarize", "visualize", "check")} (corresponding to all three steps) \\
+                                                        \code{c("summarize", "visualize", "check")}  \\
 \quad \code{smartNum} & Should numerical values with only a few unique
@@
 \quad \code{maxDecimals} & Maximum number of decimals to print for numeric values in the variable list & \code{2} \\
-\quad \code{twoCol} & Should the summary table and visualizations be placed side-by-side (in two columns)? & \code{TRUE} \\
+\quad \code{treatXasY} & How should non-supported variable classes be handled? & \code{NULL} (no handling) \\
+%Takes a list on the format \code{list(X = Y)} where \code{X} is a non-supported variable class (e.g.
+\code{complex}) and \code{Y} is a supported one (e.g.
+\code{numeric}) & \code{NULL} (No rules for handling non-supported variable types) \\
 
@@
 \quad \code{replace} & Overwrite an existing file with the same name? & \code{FALSE} \\
+\quad \code{vol} & Add a suffix to the file name of the outputted report &   \code{""} (no suffix)\\
 
@@
 \code{makeDataReport()} might be used in the very first stages of data
-cleaning, when the we are uncertain about the complexities of the
+cleaning, when we are uncertain about the complexities of the
 errors and how much time should be allocated to data cleaning. At
@@
 variables with potential problems, and with a limit of, say, maximum
-2 printed potential errors for each variable. Also, we can add the
-argument \code{replace=TRUE} in order to force \code{makeDataReport()} to
+2 printed potential problematic value per check for each variable. Also, we can add the
+argument \code{replace = TRUE} in order to force \code{makeDataReport()} to
 overwrite any existing files produced by \code{makeDataReport()}.  Using the
@@
 produced should be rendered using the \pkg{rmarkdown} \citep{rmarkdown} package and
-\code{openResult} decides whether the output html or pdf file should be
+\code{openResult} decides whether the outputted file should be
 opened. The following command produces an \proglang{R} markdown file
@@
   \textbf{\code{summaryFunction}s}  \smallskip \\
-  \quad \code{centralValue} & Compute median for numeric variables, mode for categorical variables &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\
-  \quad \code{countMissing} & Compute proportion of missing observations &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\
-  \quad \code{minMax} & Find minimum and maximum values &   &  & $\times$ & &  & $\times$ & $\times$  \\
-  \quad \code{quartiles} & Compute 1st and 3rd quartiles &    &  & $\times$ & &  & $\times$ & $\times$ \\
-  \quad \code{uniqueValue} & Count number of unique values &   $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\
-  \quad \code{variableType} & Data class of variable & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\
+  \quad \code{centralValue} & Compute median for numeric variables, mode for categorical variables &  \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} \\
+  \quad \code{countMissing} & Compute proportion of missing observations &  \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$}  \\
+  \quad \code{minMax} & Find minimum and maximum values &   &  & \blue{$\times$} & &  & \blue{$\times$} & \blue{$\times$}  \\
+  \quad \code{quartiles} & Compute 1st and 3rd quartiles &    &  & \blue{$\times$} & &  & \blue{$\times$} & \blue{$\times$} \\
+  \quad \code{uniqueValues} & Count number of unique values &   \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$}  \\
+  \quad \code{variableType} & Data class of variable & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$}  \\
   \smallskip \\
+
  \textbf{\code{visualFunction}s} \smallskip \\
   \quad \code{basicVisual} & Histograms and barplots using base \proglang{R} graphics &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\
-  \quad \code{standardVisual} & Histograms and barplots using \pkg{ggplot2} &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ \\
+  \quad \code{standardVisual} & Histograms and barplots using \pkg{ggplot2} &  \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} \\
   \smallskip \\
+
  \textbf{\code{checkFunction}s} \smallskip \\
- \quad \code{identifyCaseIssues} & Identify case issues &  $\times$ & $\times$ & & & & &  \\
- \quad \code{identifyLoners} & Identify levels with $<$ 6 obs. & $\times$ & $\times$ & & & & &  \\
- \quad \code{identifyMissing} & Identify miscoded missing values &  $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &  \\
- \quad \code{identifyNums} & Identify misclassified numeric or integer variables & $\times$ & $\times$ & & & & &  \\
- \quad \code{identifyOutliers} & Identify outliers &  & & $\times$ & & $\times$ & $\times$ \\
+ \quad \code{identifyCaseIssues} & Identify case issues &  \blue{$\times$} & \blue{$\times$} & & \blue{$\times$} & & &  \\
+ \quad \code{identifyLoners} & Identify levels with $<$ 6 obs. & \blue{$\times$} & \blue{$\times$} & & \blue{$\times$} & & &  \\
+ \quad \code{identifyMissing} & Identify miscoded missing values &  \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} &  \\
+ \quad \code{identifyNums} & Identify misclassified numeric or integer variables & \blue{$\times$} & \blue{$\times$} & & \blue{$\times$} & & &  \\
+ \quad \code{identifyOutliers} & Identify outliers &  & & \blue{$\times$} & & \blue{$\times$} & \blue{$\times$} \\
  \quad \code{identifyOutliersTBStyle} & Identify outliers (Turkish Boxplot style) &  & & $\times$ & & $\times$ & $\times$ \\
- \quad \code{identifyWhitespace} & Identify prefixed and suffixed whitespace &  $\times$ & $\times$ & & $\times$ & & &  \\
+ \quad \code{identifyWhitespace} & Identify prefixed and suffixed whitespace &  \blue{$\times$} & \blue{$\times$} & & \blue{$\times$} & & &  \\
  \quad \code{isCPR} & Identify Danish CPR numbers & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &$\times$   \\
- \quad \code{isSingular} & Check if the variable contains only a single value & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$  \\
- \quad \code{isKey} & Check if the variable is a key & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$\smallskip   \\
- \quad \code{isSupported} & Check if the variable is among the supported variable types & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$\smallskip   \\
+ \quad \code{isSingular} & Check if the variable contains only a single value & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$}  \\
+ \quad \code{isKey} & Check if the variable is a key & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} \smallskip   \\
+ \quad \code{isSupported} & Check if the variable is among the supported variable types & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} \smallskip   \\
  \hline
 \end{tabular}
-\caption{List of all summary functions (used in the summary table for
-  each variable in the output), visual functions (used for  visualization of each variable), and
-  check functions (used for data checks for each variable) currently implemented in \pkg{dataMaid}. The variable
+\caption{Overview of all summary-, visual- and  check functions currently implemented in \pkg{dataMaid}. The variable
   classes C, F, I, L, B, N, and D, refer to \code{character}, \code{factor},
-  \code{integer}, \code{labelled}, \code{logical} (boolean), \code{numeric}, and \code{Date}, respectively.}
+  \code{integer}, \code{labelled}, \code{logical} (boolean), \code{numeric}, and \code{Date}, respectively. The default settings of \code{makeDataReport()} are marked in blue.}
 \label{table.SVCfunctions}
@@
 
+\subsection{Controlling contents: What summaries, visualizations and checks should be performed?}
 We will now move on to discuss how not only the structure of the
@@
 We can control what summaries and checks are applied for each variable type
-through the \code{summaries}, \code{visuals} and \code{checks} arguments.
-Each of these arguments should be a list with one entry for each variable type and a number of function names for each such entry.
-The easiest way to specify the arguments is by use of the built-in helper functions \code{setSummaries()}, \code{setVisuals()} and \code{setChecks()}.
-
-For instance, we can inspect the default settings for summaries by calling
+through the \code{summaries}, \code{visuals} and \code{checks} arguments of \code{makeDataReport()}.
+Each of these arguments takes a list with one entry for each variable type and a number of function names for each such entry.
+The easiest way to specify the arguments is by use of the built-in helper functions \code{setSummaries()}, \code{setVisuals()} and \code{setChecks()} that contain the default settings of \code{makeDataReport()} and simple syntaxes for making small alterations of these default settings.
+We can inspect the default settings for summaries by calling:
 
@@
 $character
-[1] "identifyMissing"  "identifyWhitespace"  "identifyLoners"    
-[4]  "identifyCaseIssues"  "identifyNums"
+[1] "variableType" "countMissing" "uniqueValues" "centralValue"
 
 $factor
-[1] "identifyMissing"  "identifyWhitespace"  "identifyLoners"     
-[4] "identifyCaseIssues"  "identifyNums"
+[1] "variableType" "countMissing" "uniqueValues" "centralValue"
 
 $labelled
-[1] "identifyMissing"  "identifyWhitespace"  "identifyLoners"     
-[4] "identifyCaseIssues"  "identifyNums"
+[1] "variableType" "countMissing" "uniqueValues" "centralValue"
 
 $numeric
-[1] "identifyMissing"  "identifyOutliers"
+[1] "variableType" "countMissing" "uniqueValues" "centralValue" "quartiles"   
+[6] "minMax"
 
 $integer
-[1] "identifyMissing"  "identifyOutliers"
+[1] "variableType" "countMissing" "uniqueValues" "centralValue" "quartiles"   
+[6] "minMax"
 
 $logical
-NULL
+[1] "variableType" "countMissing" "uniqueValues" "centralValue"
 
 $Date
-[1] "identifyOutliers"
+[1] "variableType" "countMissing" "uniqueValues" "centralValue" "minMax"      
+[6] "quartiles" 
 \end{Soutput}
@@
 This helper function really just calls several other helper functions, namely the \\ 
-\code{defaultXXXSummaries()} functions, where \code{XXX} refers to a variale class. For instance, we can see the default character checks by calling \code{defaultCharacterSummaries()}:
+\code{defaultXXXSummaries()} functions, where \code{XXX} refers to a variable class. For instance, we can see the default character summaries  by calling \code{defaultCharacterSummaries()}:
 
@@
 +    summaries = setSummaries(numeric = c("variableType","centralValue),
-+    integer = c("variableType", "centralValue"))
++    integer = c("variableType", "centralValue")))
 \end{Sinput}
@@
 R> makeDataReport(toyData, 
-+    summaries = setSummaries(all = c("variableType", "centralValue"))
++    summaries = setSummaries(all = c("variableType", "centralValue")))
 \end{Sinput}
@@
 usage beyond the \code{checks} arguments, namely in the
-pre-check stage. In this stage, it is determined whether or
+\textit{pre-check} stage. In this stage, it is determined whether or
 not each variable is suitable for the summarize/visualize/check (SVC)
@@
 
-As indicated in Figure~\ref{figure:cleanStructure}, there are two stages
+In summary, and as indicated in Figure~\ref{figure:cleanStructure}, there are two stages
 where \code{makeDataReport()} applies functions to each of the variables:
@@
 It is possible to tell \code{makeDataReport()} how to handle such variables by use of the argument \code{treatXasY}.
-This argument takes a list where the names correspond to "new" variable types, while the entries must be supported variable types.
+This argument takes a list where the names correspond to "new" variable types (\code{X}), while the entries must be supported variable types (\code{Y}).
 For instance, we can instruct \code{dataMaid} to treat complex variables as numerics and generate a data report for a type \code{complex} variable like this:
@@
 \end{Schunk}
+
+In this report, we will find that the two variables, \code{complexVar} and \code{numericVar} will be have identical presentations in the variable list, as treating a \code{complex} variable as a \code{numeric} means dropping the imaginary part of the complex numbers - which was the only thing setting the two variables apart in the first place.
 
@@
 \end{Schunk} %$
+
+The syntax in this code chunk can be read as follows: "Summarize \code{var2} in \code{toyData}, and for \code{numeric} variables, set the summaries to be the default summary functions, except \code{variableType} and \code{countMissing}."
+
 Similar \code{defaultXXXSummaries()} functions are available for the other supported variable classes.
@@
 \label{sec:bigExample}
-\hl{Er titlen her for popsmart og flabet? Det er måske også et generelt tema i teksten. Ret gerne kritisk.}
 
@@
 \caption{The front page and the first page of the data overview report for the \code{presidentData} dataset.
-Note that the report title has been customized (front page), \code{identifyLoners} has been removed from the checks performed on character variables (page 1) and that variables of class \code{Name} have been set to be treated like \code{character} variables (page 1).
+Note that the report title has been customized (front page), \code{identifyLoners} has been removed from the checks performed on character variables ("Identify levels with <6 obs." is not checked for character variables in the table on page 1) and that variables of class \code{Name} have been set to be treated like \code{character} variables (page 1).
 Larger versions of the pages can be seen in
@@
 
-We will now put the bits and pieces from above together and show how \code{makeDataReport()} can be used on a less artificial dataset to create a usefl overview report.
-More specifically, we will generate a report describing the \code{presidentData} dataset, which is available in \pkg{dataMaid}.
-This dataset is a sligtly mutilated dataset with information about the 45 first US presidents, but with a few common data issues and a blind passenger.
+We will now put the bits and pieces from above together and show how \code{makeDataReport()} can be used on a less artificial dataset to create a useful overview report and how the interactive tools can subsequently be used to assist the actual data cleaning process.
+More specifically, we will create a report describing the \code{presidentData} dataset, which is available in \pkg{dataMaid} and use the information from this report to clean up the data.
+\code{presidentData} is a sligtly mutilated dataset with information about the 45 first US presidents, but with a few common data issues and a blind passenger.
 The dataset contains one observation per president and has the following variables:
@@
 \begin{enumerate}
-\item A few variables are used to describe names, namely \code{lastName} and \code{firstName}.
+\item A couple of variables are used to store names, namely \code{lastName} and \code{firstName}.
 In order to use special bibliographical analysis tools on these, and only these, variables, it might be convenient to assign them a special class.
@@
 the variable \code{stateOfBirth}.
-Therefore, we would like to disable the \code{identifyLoners} check for this variable type.
+Therefore, we would like to disable the \code{identifyLoners} check (which flags variables with <6 observations in any of the levels) for this variable type.
 This can be done using the \code{checks} argument.
-\item We would like the report to be called "Dirty president data" in order to reflect that it is, indeed, a report about dirty data about presidents.
+\item We would like the report to be called "Dirty president data" in order to reflect that it is, indeed, a report concerning dirty data about presidents.
 \end{enumerate}
-Incoorporating these three customizations, we can load the data and generate a report by calling:
+Incorporating these three customizations, we can load the data and generate a report by calling:
 
@@
 +    treatXasY = list("Name" = "character"),
-+    checks = setChecks(character = defaultCharacterChecks(remove = "identifyLoners")),
++    checks = setChecks(character = 
++    defaultCharacterChecks(remove = "identifyLoners")),
 +    reportTitle = "Dirty president data")
@@
 Stata, and therefore, it is flagged as a potential miscoded missing value.
-The variable \code{orderOfPresidency} is not summarized, visualized or checked because it is categorical and contains unqiue values for each observation.
+The variable \code{orderOfPresidency} is not summarized, visualized or checked because it is categorical and contains unique values for each observation.
 
@@
 \begin{itemize}
-\item In the birthday variable, there is a entry with the date March 1, 1300 which is a bit of an outlier. 
+\item In the birthday variable, there is an entry with the date March 1, 1300 which is a bit of an outlier. 
 \item Among the states in which the presidents were born, New York was mistakenly spelled with a lower case "Y" in at least one entry.
@@
 This is remarked in the variable presentation for \code{assassinationAttempt} and it can also be seen by the visualization being a barplot rather than a histogram.
-\item The variable concerning the sex of the president was skipped, as there is nothing to present, when all US presidents have been male so far.
+\item The variable concerning the sex of the president was skipped, as there is nothing to present when all US presidents have been male so far.
 \item The report flags the variable \code{ethnicity} to have suspiciously few observations in one category, "African American". 
-\item A few presidents were found to have odd values in the variable describing the duration of their presidencies: Some had very short (outlier) presidencies of less than three years and one was registered to have an \textit{infinite} presidency.
-\item The variable concerning age at inauguration was coded as a character variable, but consists exclusively of numbers and takes on a lot of different variables and therefore, it was flagged as a potentially misclassified numeric variable.
+\item A few presidents were found to have odd values in the variable describing the duration of their presidencies: Some had very short (outlier) presidencies of less than two years and one was registered to have an \textit{infinite} presidency.
+\item The variable concerning age at inauguration was coded as a character variable, but consists exclusively of numbers and takes a lot of different values and therefore, it was flagged as a potentially misclassified numeric variable.
 \item It seems as if one or more presidents have complex numbers as their favorite numbers.
@@
 
-A few of the identified problems have easy fixes that need no further attention.
+A few of the identified problems have easy fixes that need no further discussion.
 We remove the prefixed whitespace from Truman's name, fix the misspelling of New York, convert the binary variable \code{assasinationAttempt} to a factor and change the class of the \code{ageAtInauguration} variable to numeric:
@@
 Moving forward, we might be interested in inspecting the contents of the "."-coded entry of \code{firstName} closer, as we \textit{do} know the first names of all the US presidents.
-We look at the last name for this president and fill it in correctly:
+We look at the last name for this president and fill in the first name correctly:
 
@@
 \begin{Sinput}
-R> birthdayOutlierVal <- identifyOutliers(presidentData$birthday)$problemValues
+R> birthdayOutlierVal <- 
++    identifyOutliers(presidentData$birthday)$problemValues
 \end{Sinput}
@@
    assassinationAttempt  sex ethnicity precidencyYears ageAtInauguration
-46                    1 Male Caucasian              NA              <NA>
+46                    1 Male Caucasian              NA              88
    favoriteNumber
@@
 
-We see that this is not a proper US president and thus, it is likely to be explanation of the faulty number of observations in the dataset.
+We see that this is not a proper US president and thus, it is likely to be the explanation of the faulty number of observations in the dataset.
 Therefore, we drop this observation from the dataset, e.g.
@@
 
-Now, all that is left is the \code{presidencyYears} variable.
+Now, all that is left to fix is the \code{presidencyYears} variable.
 For this variable, we are concerned about three different things: First, it has some quite small outlier values.
@@
 
-We see that there is one less missing value and that the small and large values pertain. Therefore, we look at all the observations that cause worry, namely the outliers and the missing value:
+We see that there is one less missing value and that the small and large values pertain.
+Therefore, we look at all the observations that cause worry, namely the outliers and the missing value, and we select to see the variables \code{firstName}, \code{lastName} and \code{presidencyYears}:
 
@@
 12   Zachary   Taylor               1
-20     James Garfield               1
-29    Warren  Harding               2
-38    Gerald     Ford               2
+20     James Garfield               0
 44    Barack    Obama             Inf
@@
 We see that Obama is listed as being president forever, which history has proven to be wrong.
-Trump, on the other hand, has a missing value for his presidency duration, which is in fact reasonable as we cannot know yet.
-Presidents William Harrison, Zachary Taylor, James Garfield, Warren Harding and Gerald Ford were identified to have were brief presidencies, but these are not mistakes, as a textbook look-up can tell us.
+Trump, on the other hand, has a missing value for his presidency duration, which is in fact reasonable as we cannot know how long it will be yet (anno 2017).
+Presidents William Harrison, Zachary Taylor, James Garfield, Warren Harding and Gerald Ford were identified to have very brief presidencies, but these are not mistakes, as a textbook look-up can tell us.
 Thus the only mistake left to fix is Obama's infinite presidency:
@@
 But we have fixed some key issues in the data and thereby given ourselves a chance of a smoother sailing in the next steps of the data analysis.
-Finally, we create a new data report, adding the suffix "cleaned" to the title, as well as the file name, so that we have documentation of the current state of the dataset:
-
-\begin{Schunk}
-\begin{Sinput}
-R> makeDataReport(presidentData, vol = "_cleaned", 
+
+Finally, we create a new data report, adding the suffix "cleaned" to the title, as well as the file name, so that we have documentation of the current state of the dataset.
+We also decide that it might be sufficient to be able to inspect only the real part of the presidential favorite numbers and therefore, we choose to treat the \code{complex} variable \code{favoriteNumber} as a \code{numeric} variable:
+
+\begin{Schunk}
+\begin{Sinput}
+makeDataReport(presidentData, vol = "_cleaned", 
++    treatXasY = list(Name = "character", complex = "numeric"),
++    checks = setChecks(character = 
++    defaultCharacterChecks(remove = "identifyLoners")),
 +    reportTitle = "Dirty president data - cleaned")
@@
 
-\subsection[Incoorporating dataMaid in a programmer's workflow]{Incoorporating \pkg{dataMaid} in a programmer's workflow}
-The default settings of \code{makeDataReport()} have been set to facilitate easy and quick data report generation, but unfortunately, this also means that they might not be ideal for a more programming-oriented workflow, where the function might not be called by a human, but is part of a larger scheme.
-For instance, one might be interested in automatically running \code{makeDataReport()} on all datasets received from a certain client and returning a data report for them to inspect and comment before ever looking at the data.
+\subsection[Incorporating dataMaid in a programmer's workflow]{Incorporating \pkg{dataMaid} in a programmer's workflow}
+The default settings of \code{makeDataReport()} have been set to facilitate easy and quick data report generation, but unfortunately, this also means that it is not ideal for a more programming-oriented workflow, where the function might not be called by a human.
+For instance, one might be interested in automatically running \code{makeDataReport()} on all datasets received from a certain client, perhaps via email or through a web upload, and returning a data report for them to inspect and comment before ever looking at the data.
 In this scenario, there are a few issues with the standard data report:
@@
 \item After rendering, the report is automatically opened. This is not very useful, if the processes are supposed to run in the background.
-\item The report generation produces messages to the console while producing and rendering the report.
-\item Unless specifically told otherwise, every report created for identically named datasets will have the same file name.
+\item The report generation writes messages to the console while producing and rendering the report.
+\item Unless specifically told otherwise, every report created for the same dataset (or different datasets with the same storage name in \proglang{R}) will have the same file name.
 \end{enumerate}
-Please note that the data report does contain information about who, when and how concerning its generation.
-These three problems can easily be solved by use of the arguments of \code{makeDataReport}.
+Please note that the data report does contain information about who, when and how concerning its generation, so even though the default choices for file names do not make it easy to tell different reports for the same dataset apart, it should be rather easy when inspecting the report manually.
+These three problems can easily be solved by use of the arguments of \code{makeDataReport()}.
 Whether or not the outputted file is opened can be controlled through the argument \code{open}.
@@
 
-Now, if e.g. the report is created at 3 pm on October 31, 2017, its file will be named \code{dataMaid_toyData_10-31-2017_15.00.pdf}, making it easy spot.
+Now, if e.g. the report is created at 3 pm on October 31, 2017, its file will be named \code{dataMaid_toyData_10-31-2017_15.00.pdf}, making it easy to find.
 
 \subsection[Using dataMaid for problem flagging]{Using \pkg{dataMaid} for problem flagging}
-If the data is large, but memory issues and computation time are less
-of an issue than the time it takes a human to look through the data
-cleaning document, a viable solution might be not to include all
-information about all variables. This can be achieved by using the \code{onlyProblematic} argument in
+If the dataset is large and the time available for reading through the data report is scarce, it can be convenient to only make a report concerning the variabels that were flagged to be problematic.
+This can be achieved by using the \code{onlyProblematic} argument in
 \code{makeDataReport()}. By specifying \code{onlyProblematic = TRUE}, only
@@
 overview of the problems they might have. This is obtained by
-controlling the \code{mode} argument:
-
-\begin{Schunk}
-\begin{Sinput}
-R> makeDataReport(toyData, onlyProblematic = TRUE, mode = c("check"))
+using the \code{mode} argument:
+
+\begin{Schunk}
+\begin{Sinput}
+R> makeDataReport(toyData, onlyProblematic = TRUE, mode = "check")
 \end{Sinput}
@@
 Now only the checking results are printed, and only for variables
-where problems were identified. An even more minimal output can be
-generated by also leaving out the checking results --- then
-\code{makeDataReport()} essentially just produces a document with a list of the variable names
-that should be investigated further:
-
-\begin{Schunk}
-\begin{Sinput}
-R> makeDataReport(toyData, onlyProblematic = TRUE, mode = NULL)
-\end{Sinput}
-\end{Schunk}
-
-Of course, this can also be done without generating an overview
-document, by direct, interactive use of the \code{check()} function. When
-called on a \code{data.frame}, this function produces a list (of
+where problems were identified.
+An even more minimal output  can be obtained directly in the console by using the \code{check()} function interactively.
+When called on a \code{data.frame}, this function produces a list (of
 variables) of lists (of checks) of lists (or rather,
@@
 and we find that only a single variable in \code{toyData}, \code{var6} (for which all
-observations have the value \code{"Irrelevant"}), is problem-free, using the default checks.
+observations have the value \code{"Irrelevant"}), is problem-free when using the default checks.
 
@@
 
-Alternatively, one can create the usual output report, not render it and then manually edit the produced rmarkdown file as wished. This can be done with the following command:
-If one works with a dataset containing a very large number of variables, it can be practical to split up two of the procedures undertaken by the \code{makeDataReport()} function, namely creating- and rendering the data report.
+Alternatively, one can create the usual output report, not render it and then manually edit the produced \proglang{R} markdown file as wished. This can be done with the following command:
 \begin{Schunk}
@@
 in further developing the functionality of \pkg{dataMaid} by providing
-more summary, visual, and check function as part of the default
+more summary, visual, and check function as part of the
 package. We are also currently considering adding options to handle

### 2017-10-31 latex/article_vol3.Rnw
Convert manuscript from tex -> Rnw

--- 2017-10-20 latex/article_vol3.tex
@@
 \begin{document}
+
+<<echo=FALSE>>=
+# options(prompt = "R> ", width=77, continue = "+  ")
+options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> library("dataMaid")
-R> data("toyData")
-R> toyData
-\end{Sinput}
-\begin{Soutput}
-   var1 var2  var3       var4 var5       var6
-1   red    1     a -0.6264538    1 Irrelevant
-2   red    1     a  0.1836433    2 Irrelevant
-3   red    1     a -0.8356286    3 Irrelevant
-4   red    2     a  1.5952808    4 Irrelevant
-5   red    2     a  0.3295078    5 Irrelevant
-6   red    6     b -0.8204684    6 Irrelevant
-7   red    6     b  0.4874291    7 Irrelevant
-8   red    6     b  0.7383247    8 Irrelevant
-9   red  999     c  0.5757814    9 Irrelevant
-10  red   NA     c -0.3053884   10 Irrelevant
-11 blue    4     c  1.5117812   11 Irrelevant
-12 blue   82     .  0.3898432   12 Irrelevant
-13 blue   NA       -0.6212406   13 Irrelevant
-14 <NA>  NaN other -2.2146999   14 Irrelevant
-15 <NA>    5 OTHER  1.1249309   15 Irrelevant
-\end{Soutput}
-\end{Schunk}
-
-\begin{Schunk}
-\begin{Sinput}
-R> makeDataReport(toyData)
-\end{Sinput}
-\end{Schunk}
+<<>>=
+library("dataMaid")
+data("toyData")
+toyData
+@
+
+<<>>=
+makeDataReport(toyData)
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> makeDataReport(toyData, output = "html", onlyProblematic = TRUE,
-+    maxProbVals = 2, replace = TRUE)
-\end{Sinput}
-\end{Schunk}
+<<>>=
+makeDataReport(toyData, output = "html", onlyProblematic = TRUE,
+   maxProbVals = 2, replace = TRUE)
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> makeDataReport(toyData, output="html", render=FALSE, openResult=FALSE)
-\end{Sinput}
-\end{Schunk}
+<<>>=
+makeDataReport(toyData, output="html", render=FALSE, openResult=FALSE, replace=TRUE)
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> allSummaryFunctions()
-\end{Sinput}
-\begin{Soutput}
-----------------------------------------------------------------------------
-name           description                     classes                      
--------------- ------------------------------- -----------------------------
-centralValue   Compute median for numeric      character, Date, factor,     
-               variables, mode for             integer, labelled, logical,  
-               categorical variables           numeric
-
-countMissing   Compute proportion of missing   character, Date, factor,     
-               observations                    integer, labelled, logical,  
-                                               numeric
-
-minMax         Find minimum and maximum        integer, numeric, Date       
-               values
-
-quartiles      Compute 1st and 3rd quartiles   Date, integer, numeric
-
-uniqueValues   Count number of unique values   character, Date, factor,     
-                                               integer, labelled, logical,  
-                                               numeric
-
-variableType   Data class of variable          character, Date, factor,     
-                                               integer, labelled, logical,  
-                                               numeric                      
-----------------------------------------------------------------------------
-\end{Soutput}
-\end{Schunk}
+<<>>=
+allSummaryFunctions()
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> setSummaries()
-\end{Sinput}
-\begin{Soutput}
-$character
-[1] "variableType" "countMissing" "uniqueValues" "centralValue"
-
-$factor
-[1] "variableType" "countMissing" "uniqueValues" "centralValue"
-
-$labelled
-[1] "variableType" "countMissing" "uniqueValues" "centralValue"
-
-$numeric
-[1] "variableType" "countMissing" "uniqueValues" "centralValue" "quartiles"   
-[6] "minMax"
-
-$integer
-[1] "variableType" "countMissing" "uniqueValues" "centralValue" "quartiles"   
-[6] "minMax"
-
-$logical
-[1] "variableType" "countMissing" "uniqueValues" "centralValue"
-
-$Date
-[1] "variableType" "countMissing" "uniqueValues" "centralValue" "minMax"      
-[6] "quartiles" 
-\end{Soutput}
-\end{Schunk} %$
+<<>>=
+setSummaries()
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> defaultCaracterSummaries()
-\end{Sinput}
-\begin{Soutput}
-[1] "variableType" "countMissing" "uniqueValues" "centralValue"
-\end{Soutput}
-\end{Schunk}
+<<>>=
+defaultCharacterSummaries()
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> makeDataReport(toyData, 
-+    summaries = setSummaries(numeric = c("variableType","centralValue),
-+    integer = c("variableType", "centralValue")))
-\end{Sinput}
-\end{Schunk}
+<<>>=
+makeDataReport(toyData, replace=TRUE,
+  summaries = setSummaries(numeric = c("variableType","centralValue"),
+  integer = c("variableType", "centralValue")))
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> makeDataReport(toyData, 
-+    summaries = setSummaries(all = c("variableType", "centralValue")))
-\end{Sinput}
-\end{Schunk}
+<<>>=
+makeDataReport(toyData, replace=TRUE,                               
+  summaries = setSummaries(all = c("variableType", "centralValue")))
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> defaultFactorChecks()
-\end{Sinput}
-\begin{Soutput}
-[1] "identifyMissing"    "identifyWhitespace" "identifyLoners"    
-[4] "identifyCaseIssues" "identifyNums"      
-\end{Soutput}
-\end{Schunk}
+<<>>=
+defaultFactorChecks()
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> makeDataReport(toyData, 
-+    checks = setChecks(factor = "identifyWhitespace"))
-\end{Sinput}
-\end{Schunk}
+<<>>=
+makeDataReport(toyData, replace=TRUE,
+  checks = setChecks(factor = "identifyWhitespace"))
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> makeDataReport(toyData, checks = setChecks(factor = NULL))
-\end{Sinput}
-\end{Schunk}
+<<>>=
+makeDataReport(toyData, checks = setChecks(factor = NULL), replace=TRUE)
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> makeDataReport(toyData, preChecks = c("isKey", "isSupported"))
-\end{Sinput}
-\end{Schunk}
+<<>>=
+makeDataReport(toyData, preChecks = c("isKey", "isSupported"), replace=TRUE)
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> makeDataReport(toyData, visuals = setVisuals(all = "basicVisual"))
-\end{Sinput}
-\end{Schunk}
+<<>>=
+makeDataReport(toyData, visuals = setVisuals(all = "basicVisual"), replace=TRUE)
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
->R complexData <- data.frame(complexVar = complex(100, real = 1:100, 
-+    imaginary = 3), numericVar = 1:100))
->R makeDataReport(complexData, treatXasY(complex = "numeric"))
-\end{Sinput}
-\end{Schunk}
-
-In this report, we will find that the two variables, \code{complexVar} and \code{numericVar} will be have identical presentations in the variable list, as treating a \code{complex} variable as a \code{numeric} means dropping the imaginary part of the complex numbers - which was the only thing setting the two variables apart in the first place.
+<<warning=FALSE>>=
+complexData <- data.frame(complexVar = complex(100, real = 1:100, imaginary = 3), numericVar = 1:100)
+makeDataReport(complexData, treatXasY=list(complex = "numeric"), replace=TRUE)
+@
+
+In this report, we will find that the two variables, \code{complexVar} and \code{numericVar} will be have identical presentations in the variable list, as treating a \code{complex} variable as a \code{numeric} means dropping the imaginary part of the complex numbers which was the only thing setting the two variables apart in the first place.
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> check(toyData$var2)
-\end{Sinput}
-\begin{Soutput}
-$identifyMissing
-The following suspected missing value codes enter as regular values: 999, NaN.
-$identifyOutliers
-Note that the following possible outlier values were detected: 82, 999.
-\end{Soutput}
-\end{Schunk}%$
+<<>>=
+check(toyData$var2)
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> check(toyData$var2, checks = setChecks(numeric = "identifyMissing"))
-\end{Sinput}
-\begin{Soutput}
-$identifyMissing
-The following suspected missing value codes enter as regular values: 999, NaN.
-\end{Soutput}
-\end{Schunk}
+<<>>=
+check(toyData$var2, checks = setChecks(numeric = "identifyMissing"))
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> identifyMissing(toyData$var2)
-\end{Sinput}
-\begin{Soutput}
-The following suspected missing value codes enter as regular values: 999, NaN.
-\end{Soutput}
-\end{Schunk} %$
+<<>>=
+identifyMissing(toyData$var2)
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> missVar2 <- identifyMissing(toyData$var2)
-R> str(missVar2)
-\end{Sinput}
-\begin{Soutput}
-List of 3
- $ problem      : logi TRUE
- $ message      : chr "The following suspected missing value codes
-    enter as regular values: \\\"999\\\", \\\"NaN\\\"."
- $ problemValues: num [1:2] 999 NaN
- - attr(*, "class")= chr "checkResult"
-\end{Soutput}
-\end{Schunk}
+<<>>=
+missVar2 <- identifyMissing(toyData$var2)
+str(missVar2)
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> toyData$var2[toyData$var2 %in% missVar2$problemValues] <- NA
-R> identifyMissing(toyData$var2)
-\end{Sinput}
-\begin{Soutput}
-No problems found.
-\end{Soutput}
-\end{Schunk} %$
+<<>>=
+toyData$var2[toyData$var2 %in% missVar2$problemValues] <- NA
+identifyMissing(toyData$var2)
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> visualize(toyData$var2)
-R> summarize(toyData$var2)
-\end{Sinput}
-\begin{Soutput}
-     Feature                   Result       
-[1,] "Variable type"           "numeric"    
-[2,] "Number of missing obs."  "4 (26.67 %)"
-[3,] "Number of unique values" "6"          
-[4,] "Median"                  "4"          
-[5,] "1st and 3rd quartiles"   "1.5; 6"     
-[6,] "Min. and max."           "1; 82"      
-\end{Soutput}
-\end{Schunk}
+<<>>=
+visualize(toyData$var2)
+summarize(toyData$var2)
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> summarize(toyData$var2, 
-+    summaries = setSummaries(
-+    numeric = defaultNumericSummaries(remove = c("variableType", 
-+    "countMissing"))))
-\end{Sinput}
-\begin{Soutput}
-     Feature         Result 
-[1,] "Median"        "4"    
-[2,] "Min. and max." "1; 82"
-\end{Soutput}
-\end{Schunk} %$
+<<>>=
+summarize(toyData$var2, 
+  summaries = setSummaries(
+  numeric = defaultNumericSummaries(remove = c("variableType", 
+  "countMissing"))))
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> data(presidentData)
-R> makeDataReport(presidentData, replace = TRUE, 
-+    treatXasY = list("Name" = "character"),
-+    checks = setChecks(character = 
-+    defaultCharacterChecks(remove = "identifyLoners")),
-+    reportTitle = "Dirty president data")
-\end{Sinput}
-\end{Schunk}
+<<>>=
+data(presidentData)
+makeDataReport(presidentData, replace = TRUE, 
+  treatXasY = list("Name" = "character"),
+  checks = setChecks(character = 
+  defaultCharacterChecks(remove = "identifyLoners")),
+  reportTitle = "Dirty president data")
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> presidentData$lastName[presidentData$lastName == " Truman"] <- 
-+    "Truman"
-R> presidentData$stateOfBirth[presidentData$stateOfBirth == "New york"] <- 
-+    "New York"
-R> presidentData$assassinationAttempt <- 
-+    factor(presidentData$assassinationAttempt)
-R> presidentData$ageAtInauguration <- 
-+    as.numeric(presidentData$ageAtInauguration)
-\end{Sinput}
-\end{Schunk}%$
+<<>>=
+presidentData$lastName[presidentData$lastName == " Truman"] <- "Truman"
+presidentData$stateOfBirth[presidentData$stateOfBirth == "New york"] <- "New York"
+presidentData$assassinationAttempt <- factor(presidentData$assassinationAttempt)
+presidentData$ageAtInauguration <- as.numeric(presidentData$ageAtInauguration)
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
->R presidentData$ageAtInauguration <- 
-+    as.numeric(as.character(presidentData$ageAtInauguration))
-\end{Sinput}
-\end{Schunk}
+<<>>=
+presidentData$ageAtInauguration <- as.numeric(as.character(presidentData$ageAtInauguration))
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> presidentData$lastName[presidentData$firstName == "."]
-\end{Sinput}
-\begin{Soutput}
-[1] "Trump"
-\end{Soutput}
-\begin{Sinput}
-R> presidentData$firstName[presidentData$firstName == "."] <- "Donald"
-\end{Sinput}
-\end{Schunk}
+<<>>=
+presidentData$lastName[presidentData$firstName == "."]
+@
+
+<<>>=
+presidentData$firstName[presidentData$firstName == "."] <- "Donald"
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> birthdayOutlierVal <- 
+<<>>=
+birthdayOutlierVal <- 
 +    identifyOutliers(presidentData$birthday)$problemValues
-\end{Sinput}
-\end{Schunk}
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-> presidentData[presidentData$birthday == birthdayOutlierVal, ]
-\end{Sinput}
-\begin{Soutput}
-      lastName firstName orderOfPresidency   birthday stateOfBirth
-46 Arathornson   Aragorn                 0 1300-03-01       Gondor
-   assassinationAttempt  sex ethnicity precidencyYears ageAtInauguration
-46                    1 Male Caucasian              NA              88
-   favoriteNumber
-46           8+0i
-\end{Soutput}
-\end{Schunk}%$
+<<>>=
+presidentData[presidentData$birthday == birthdayOutlierVal, ]
+@
 
@@
 
-\begin{Schunk} 
-\begin{Sinput}
-R> presidentData <- 
-+    presidentData[presidentData$birthday != birthdayOutlierVal, ]
-\end{Sinput}
-\end{Schunk} %$
+<<>>=
+presidentData <- 
+  presidentData[presidentData$birthday != birthdayOutlierVal, ]
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
->R summarize(presidentData$presidencyYears)
-\end{Sinput}
-\begin{Soutput}
-$variableType
-Variable type: numeric
-$countMissing
-Number of missing obs.: 1 (2.22 %)
-$uniqueValues
-Number of unique values: 11
-$centralValue
-Median: 4
-$quartiles
-1st and 3rd quartiles: 4; 8
-$minMax
-Min. and max.: 0; Inf
-\end{Soutput}
-\end{Schunk} %$
+<<>>=
+summarize(presidentData$presidencyYears)
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> presidentData[is.na(presidentData$presidencyYears) | 
-+     presidentData$presidencyYears %in% 
-+     identifyOutliers(presidentData$presidencyYears)$problemValues, 
-+     c("firstName", "lastName", "presidencyYears")]
-\end{Sinput}
-\begin{Soutput}
-   firstName lastName presidencyYears
-9    William Harrison               0
-12   Zachary   Taylor               1
-20     James Garfield               0
-44    Barack    Obama             Inf
-45    Donald    Trump              NA
-\end{Soutput}
-\end{Schunk}
+<<>>=
+presidentData[is.na(presidentData$presidencyYears) | 
+   presidentData$presidencyYears %in% 
+   identifyOutliers(presidentData$presidencyYears)$problemValues, 
+   c("firstName", "lastName", "presidencyYears")]
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> presidentData$presidencyYears[presidentData$lastName == "Obama"] <- 8
-\end{Sinput}
-\end{Schunk}
+<<>>=
+presidentData$presidencyYears[presidentData$lastName == "Obama"] <- 8
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
+<<>>=
 makeDataReport(presidentData, vol = "_cleaned", 
-+    treatXasY = list(Name = "character", complex = "numeric"),
-+    checks = setChecks(character = 
-+    defaultCharacterChecks(remove = "identifyLoners")),
-+    reportTitle = "Dirty president data - cleaned")
-\end{Sinput}
-\end{Schunk}
+  treatXasY = list(Name = "character", complex = "numeric"),
+  checks = setChecks(character = 
+  defaultCharacterChecks(remove = "identifyLoners")),
+  reportTitle = "Dirty president data - cleaned",
+  replace=TRUE)
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> makeDataReport(toyData, onlyProblematic = TRUE, mode = "check")
-\end{Sinput}
-\end{Schunk}
+<<>>=
+makeDataReport(toyData, onlyProblematic = TRUE, mode = "check", replace=TRUE)
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> toyChecks <- check(toyData)
-R> foo <- function(x) {
-+    any(sapply(x, function(y) y[["problem"]]))
-+  }
-R> sapply(toyChecks, foo)
-\end{Sinput}
-\begin{Soutput}
- var1  var2  var3  var4  var5  var6 
- TRUE  TRUE  TRUE  TRUE  TRUE FALSE 
-\end{Soutput}
-\end{Schunk}
+<<>>=
+toyChecks <- check(toyData)
+foo <- function(x) {
+  any(sapply(x, function(y) y[["problem"]]))
+}
+sapply(toyChecks, foo)
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> makeDataReport(toyData, standAlone = FALSE)
-\end{Sinput}
-\end{Schunk}
+<<eval=FALSE>>=
+makeDataReport(toyData, standAlone = FALSE)
+@
 
@@
 
-\begin{Schunk}
-\begin{Sinput}
-R> makeDataReport(toyData, standAlone = FALSE, output = "html", 
-+    twoCols = FALSE)
-\end{Sinput}
-\end{Schunk}
+<<eval=FALSE>>=
+makeDataReport(toyData, standAlone = FALSE, output = "html", 
+  twoCols = FALSE)
+@
 
@@
 Alternatively, one can create the usual output report, not render it and then manually edit the produced \proglang{R} markdown file as wished. This can be done with the following command:
-\begin{Schunk}
-\begin{Sinput}
-R> makeDataReport(toyData, render = FALSE, openResult = FALSE)
-\end{Sinput}
-\end{Schunk}
+
+<<eval=FALSE>>=
+makeDataReport(toyData, render = FALSE, openResult = FALSE)
+@
+
 After editing, the file can be rendered by calling the \code{render} function:
-\begin{Schunk}
-\begin{Sinput}
-R> render("dataMaid_toyData.Rmd", quiet = FALSE)
-\end{Sinput}
-\end{Schunk}
+
+<<eval=FALSE>>=
+render("dataMaid_toyData.Rmd", quiet = FALSE)
+@
 

### 2017-10-31 latex/dataMaid_toyData.Rmd
Convert manuscript from tex -> Rnw

--- 2017-01-03 latex/dataMaid_toyData.Rmd
@@
 subtitle: "Autogenerated data summary from dataMaid"
-date: "`r Sys.Date()`"
+date: 2017-10-31 15:34:55
 output: pdf_document
@@
 ```
-# Data cleaning summary
+# Data report overview
 The dataset examined has the following dimensions:
 
-------------------------------
+---------------------------------
 Feature                    Result
---------------------- --------
-Number of rows              15
+------------------------ --------
+Number of observations         15
 
 Number of variables             6
-------------------------------
+---------------------------------
 
@@
 
+# Summary table
+
+----------------------------------------------------------------------------------
+&nbsp;   Variable class     # unique values   Missing observations  Any problems? 
+-------- ---------------- ----------------- ---------------------- ---------------
+[var1]   factor                           3                13.33 %    $\times$
+
+[var2]   numeric                          7                26.67 %    $\times$
+
+[var3]   factor                           7                 0.00 %    $\times$
+
+[var4]   numeric                         15                 0.00 %    $\times$
+
+[var5]   factor                          15                 0.00 %    $\times$
+
+[var6]   factor                           1                 0.00 %    $\times$    
+----------------------------------------------------------------------------------
+
 # Variable list
-## **var5**
+## var1
+
+- Note that the following levels have at most five observations: \"blue\".
+
+\fullline
+
+## var2
+
+- Note that the following possible outlier values were detected: \"82\".
+
+\fullline
+
+## var3
+
+- The following suspected missing value codes enter as regular values: \" \", \".\".
+
+- The following values appear with prefixed or suffixed white space: \" \".
+
+- Note that the following levels have at most five observations: \" \", \".\", \"OTHER\", \"a\", \"b\", \"c\", \"other\".
+
+- Note that there might be case problems with the following levels: \"OTHER\", \"other\".
+
+\fullline
+
+## var4
+
+- Note that the following possible outlier values were detected: \"1.12\", \"1.51\", \"1.6\".
+
+\fullline
+
+## var5
 
@@
 
-## **var6**
+## var6
 
@@
 
-This report was created by dataMaid v0.9.2.
+Report generation information:
+
+ *  Created by Claus Thorn Ekstrøm.
+
+ *  Report creation time: Tue Oct 31 2017 15:34:55
+
+ *  dataMaid v0.9.7.9000 [Pkg: 2017-10-31 from local (ekstroem/dataMaid@NA)]
+
+ *  R version 3.4.1 (2017-06-30).
+
+ *  Platform: x86_64-apple-darwin15.6.0 (64-bit)(macOS Sierra 10.12.6).
+
+ *  Function call: `makeDataReport(data = toyData, onlyProblematic = TRUE, mode = "check", 
+    replace = TRUE)`

### 2017-11-03 R/makeDataReport.R
Fixed bug when doing seperate report rendering and using standardVisuals (due to non-exported visualfunctions).

@@ -626,6 +626,14 @@ makeDataReport <- function(data, output=NULL, render=TRUE,
     ## include packages as a first chunk
     secretChunk.wrapper("library(ggplot2)\nlibrary(pander)")
     
+    ## Define unexported visual functions locally so that the report 
+    ## can be rendered from the global environment. Only done if standardVisuals are used
+    if ("visualize" %in% mode & "standardVisual" %in% visuals) {
+      secretChunk.wrapper(c("ggAggHist <- getFromNamespace(\"ggAggHist\", \"dataMaid\")", 
+                            "ggAggBarplot <- getFromNamespace(\"ggAggBarplot\", \"dataMaid\")"),
+                          label = "visualFunctions")
+    }
+    
     ## Title
     writer("# Data report overview")
     writer("The dataset examined has the following dimensions:", "\n")

### 2017-11-08 latex/article_vol3.Rnw
First round of tests

@@
   Health\\University of Copenhagen}
-\title{\pkg{dataMaid}: your assistant for data documentation in \proglang{R}}
+\title{\pkg{dataMaid}: your assistant for documenting supervised data quality screening in \proglang{R}}
 
@@
   implements an extensive and customizeable suite of quality
-  assessment tools that can be applied to a dataset in order to
+  assessment aids that can be applied to a dataset in order to
   identify potential problems in its variables. The results are
@@
   between data analysts and field experts, while also providing easy
-  documentation of reproducible data cleaning steps and data quality
-  control. Moreover, the \pkg{dataMaid} solution changes the data
-  cleaning process from the usual ad hoc approach to a systematic,
-  well-documented endeavor.  \pkg{dataMaid} also provides a suite of
-  more typical \proglang{R} tools for interactive data quality
-  assessment and -cleaning, where the data inspections are executed directly 
-  in the \proglang{R} console.
+  documentation of reproducible data quality screening. Moreover, the
+  \pkg{dataMaid} solution changes the data screening process from the
+  usual ad hoc approach to a systematic, well-documented endeavor.
+  \pkg{dataMaid} also provides a suite of more typical \proglang{R}
+  tools for interactive data quality assessment and -screening, where
+  the data inspections are executed directly in the \proglang{R}
+  console.
 
@@
 }
-\Keywords{data cleaning, quality control, \proglang{R}, data documentation}
-\Plainkeywords{data cleaning, quality control, R, data documentation} %% without formatting
+\Keywords{data screening, data cleaning, quality control, \proglang{R}, data documentation}
+\Plainkeywords{data screening, data cleaning, quality control, R, data documentation} %% without formatting
 %% at least one keyword must be supplied
@@
 data cleaning process. Some errors can be tracked down and flagged by a
-computer without further ado, while other types of errors need an analytic
+computer without further ado, while other types of errors need a subject
 context in order to be identified. Even in this latter case, well-designed
@@
 
-But even when tools are available for identifying problems in a dataset, the activity of data cleaning still suffers from a challenge that has recieved increasing attention in the scientitic communities in the later years: Data cleaning is not very straight forward to document and therefore, reproducibility suffers.
-We present a new \proglang{R} package, \pkg{dataMaid} \citep{dataMaid}, whose most central purpose is to facilitate a data cleaning workflow where documentation is thoroughly integrated rather than a secondary add-on.
-This is accomplished by centering the data cleaning process around auto-generated data overview reports that summarize the contents of the current version of a dataset and flags potential problems that could be candidates for the next bullet on the cleaning agenda.
-
-A number of \proglang{R} packages made for other pre-analysis steps are
-already available, including \pkg{janitor} \citep{janitor}, \hl{assertive}, \pkg{dplyr}
-\citep{dplyr}, \pkg{tidyr}\citep{tidyr}, \pkg{data.table} \citep{data.table}, \pkg{DataCombine}
-\citep{DataCombine} , \pkg{validate} \citep{validate}, \pkg{assertr} \citep{assertr}, and
-\pkg{DataExplorer} \citep{DataExplorer}. These packages focus on
-different stages of the pre-analysis work, most of which are not
-really data cleaning.  \pkg{janitor} provides tools for data import
-with a particular emphasis on the challenges of getting neat data
-frames from Microsoft Excel data files. \hl{Mention assertive here}. \pkg{dplyr}, \pkg{tidyr}, \pkg{data.table}
-and \pkg{DataCombine} go a few steps further by providing a wide array
-of extremely powerful tools for data wrangling, including a number of
+But even when tools are available for identifying problems in a
+dataset, the activity of data cleaning still suffers from a challenge
+that has recieved increasing attention in the scientitic communities
+in the later years: Data cleaning is not very straight forward to
+document and therefore, reproducibility suffers. We present a new
+\proglang{R} package, \pkg{dataMaid} \citep{dataMaid}, whose most
+central purpose is to facilitate a supervised data quality screening
+workflow where documentation is thoroughly integrated rather than an
+add-on. This is accomplished by structuring the data screening around
+auto-generated data overview reports that summarize and flags
+potential problems in the dataset.
+
+But no matter how clever software tools we make, data cleaning remains
+to be a time consuming endeavor, which inherently requires human
+interaction since every dataset is different and the variables in the
+dataset can only be understood in the proper context of their
+origin. This often requires a collaborative effort between an expert
+in the field and a statistician or data scientist. In many
+situations, these errors are discovered in the process of the data
+analysis (e.g., a categorical variable with numeric labels for each
+category may be wrongly classified as a quantitative variable or a
+variable where all values have erroneously been coded to the same
+value), but in other cases a human with knowledge about the data
+context area is needed to identify possible mistakes in the data
+(e.g., if there are 4 categories for a variable that should only have
+3).
+
+The \pkg{dataMaid} approach to data screening, -quality assessment and
+-documentation is governed by two fundamental paradigms. First of all,
+there is no need for data cleaning to be an ad hoc procedure. Often,
+we have a very clear idea of what flags are raisable in a given
+dataset before we look at it, as we were the ones to produce it in the
+first place. This means that data cleaning can easily be a
+well-documented, well-specified procedure. In order to aid this
+paradigm, \pkg{dataMaid} provides easy-to-use, automated tools for
+data quality assessment in \proglang{R} \citep{R} on which data
+cleaning decisions can be made. This quality assessment is presented
+in an auto-generated overview document, readable by data analysts and
+field experts alike, thereby also contributing to an inter-field
+dialogue about the data at hand. Oftentimes, e.g., distinguishing
+between faulty codings of a numeric value and unusual, but correct,
+values requires problem-specific expertise that might not be held by
+the data analyst. Hopefully, having easy access to data summaries
+through \pkg{dataMaid} will help this necessary knowledge sharing.
+
+While \pkg{dataMaid}s primary raison d'être is auto-generating data
+quality assessment overview documents, we still wish to emphasize that
+it is \emph{not} a tool for unsupervised data cleaning. This qualifies
+as the second paradigm of \pkg{dataMaid}: Data cleaning decisions
+should always be made by humans. Therefore, \pkg{dataMaid} does not
+supply any tools for ``fixing'' errors in the data. However, we do
+provide interactive functions that can be used to identify potentially
+erroneous entries in a dataset and that can make it easier to solve
+data issues, one variable at a time.
+
+A number of \proglang{R} packages made for other pre-analysis steps
+are already available, including \pkg{janitor} \citep{janitor},
+\pkg{assertive} \citep{assertive}, \pkg{dplyr} \citep{dplyr},
+\pkg{tidyr}\citep{tidyr}, \pkg{data.table} \citep{data.table},
+\pkg{DataCombine} \citep{DataCombine}, \pkg{validate}
+\citep{validate}, and \pkg{assertr} \citep{assertr}. These packages
+focus on different stages of the pre-analysis work.  \pkg{janitor}
+provides tools for data import with a particular emphasis on the
+challenges of getting neat data frames from Microsoft Excel data
+files. \pkg{dplyr}, \pkg{tidyr}, \pkg{data.table} and
+\pkg{DataCombine} go a few steps further by providing a wide array of
+extremely powerful tools for data wrangling, including a number of
 particularly useful functions for merging and working with very large
@@
 \citep{editrules} and \pkg{deducorrect} \citep{deducorrect} from the
-same authors) does offer a few tools for identifying and
-auto-correcting errors in a dataset, but, as the name suggests, the
-focus is on internal validity rather than general data cleaning. In
-practice, this means that quite elegant tools for, e.g., linear
-restraints among the variables in a dataset can be applied, but
-looking for potentially miscoded missing values is not really
-feasible. The main difference between these two challenges is the
-direction in which the data is inspected: While linear constraints
-work observation-wise with no ambiguity, determining whether or not something
-is a miscoded missing value often requires knowledge about the full
-variable (e.g. range or data type), and thus it should be performed
-variable-wise. \pkg{validate} does not allow for user-defined extensions
-of the latter type, thereby limiting its data cleaning potential.
-Automatic data correction functions are also provided by
-\pkg{validate} which we consider to be quite a dangerous cocktail: all
-power is given to the the computer with no human supervision, and
-investigators are less likely to make an active, case-specific choice
-regarding the handling of the potential errors.  Finally, no tools
-have been made to easily document exactly which checks and preliminary
-results were used in the data cleaning process.
-The \pkg{assertr} package provide very similar - and very nice - tools to those of \pkg{validate}, but without any amibitions of conducting auto-cleaning.
+same authors) and \pkg{assertive} offers tools for identifying errors
+in a dataset by checking the state of the variable given a set of
+pre-specified rules, and their focus is on internal validity rather
+than general data screening. In practice, this means that quite elegant
+tools for, e.g., linear restraints among the variables in a dataset
+can be applied, but looking for potentially miscoded missing values is
+not really feasible. The main difference between these two challenges
+is the direction in which the data is inspected: While linear
+constraints work observation-wise with no ambiguity, determining
+whether or not something is a miscoded missing value often requires
+knowledge about the full variable (e.g. range or data type), and thus
+it should be performed variable-wise. \pkg{validate} does not
+currently allow for user-defined extensions of the latter type,
+thereby limiting its data cleaning potential.  Automatic data
+correction functions are also provided by \pkg{validate} which we
+consider to be quite a dangerous cocktail: all power is given to the
+the computer with no human supervision, and investigators are less
+likely to make an active, case-specific choice regarding the handling
+of the potential errors.  Finally, no tools have been made to easily
+document exactly which checks and preliminary results were used in the
+data cleaning process. The \pkg{assertr} package provide very similar
+--- and very nice --- tools to those of \pkg{validate}, but without
+any amibitions of conducting auto-cleaning.
 
@@
 
-But no matter how clever software tools we make, data cleaning remains
-to be a time consuming endeavor, as it inherently requires human
-interaction since every dataset is different and the variables in the
-dataset can only be understood in the proper context of their
-origin. This often requires a collaborative effort between an expert
-in the field and a statistician or data scientist, which may be why
-the process of proper data cleaning is not always undertaken. In many
-situations, these errors are discovered in the process of the data
-analysis (e.g., a categorical variable with numeric labels for each
-category may be wrongly classified as a quantitative variable or a
-variable where all values have erroneously been coded to the same
-value), but in other cases a human with knowledge about the data
-context area is needed to identify possible mistakes in the data
-(e.g., if there are 4 categories for a variable that should only have
-3).
-
-The \pkg{dataMaid} approach to data cleaning, -quality assessment and -documentation is
-governed by two fundamental paradigms. First of all, there is no need
-for data cleaning to be an ad hoc procedure. Often, we have a very
-clear idea of what flags are raisable in a given dataset before we
-look at it, as we were the ones to produce it in the first place. This
-means that data cleaning can easily be a well-documented,
-well-specified procedure. In order to aid this paradigm, \pkg{dataMaid}
-provides easy-to-use, automated tools for data quality assessment in
-\proglang{R} \citep{R} on which data cleaning decisions can be made. This quality
-assessment is presented in an auto-generated overview document,
-readable by data analysts and field experts alike, thereby also
-contributing to an inter-field dialogue about the data at
-hand. Oftentimes, e.g., distinguishing between faulty codings of a
-numeric value and unusual, but correct, values requires
-problem-specific expertise that might not be held by the data
-analyst. Hopefully, having easy access to data descriptions through
-\pkg{dataMaid} will help this necessary knowledge sharing.
-
-While \pkg{dataMaid}s primary raison d'être is auto-generating data
-quality assessment overview documents, we still wish to emphasize that
-it is \emph{not} a tool for unsupervised data cleaning. This qualifies
-as the second paradigm of \pkg{dataMaid}: Data cleaning decisions
-should always be made by humans. Therefore, \pkg{dataMaid} does not
-supply any tools for ``fixing'' errors in the data. However, we do
-provide interactive functions that can be used to identify potentially
-erroneous entries in a dataset and that can make it easier to solve
-data issues, one variable at a time.
-
 This manuscript is structured as follows: First, in Section
@@
 overview documents. In the \pkg{dataMaid} package, we have
-provided a number of default checks that cover the data
+provided a number of default generic checks that cover the data
 cleaning challenges we find to be most common and these are also summarized in Section \ref{sec:usingdataMaid}. Next, in Section
@@
 by the second paradigm above.
-Next, we show atep-by-step how the data report and the interactive mode of \pkg{dataMaid} can be combined to conduct a well-documented, systematic data cleaning in Section \ref{sec:bigExample}.
+Next, we show step-by-step how the data report and the interactive mode of \pkg{dataMaid} can be combined to conduct a well-documented, systematic data cleaning in Section \ref{sec:bigExample}.
 Here, we assess and clean a dirty dataset with information about the US presidents.
@@
 
-Please note that this description of what \pkg{dataMaid} contains is usefully supplemented by a tutorial into how it can be extended.
-Every dataset is different, and some datasets might include problems that cannot be detected by our data checking functions.
-Therefore, \pkg{dataMaid} was designed to be easily extendend with user-supplied functions for summarizing, viusalizing and checking data.
-We have provided a vignette \hl{REF TO VIGNETTE} in which we describe how \pkg{dataMaid} extensions can be made, such that they are integrable with the \code{makeDataReport()}
-function and with the other tools available in \pkg{dataMaid}.
+%Please note that this description of what \pkg{dataMaid} contains is
+%usefully supplemented by a tutorial into how it can be extended. Every
+%dataset is different, and some datasets might include problems that
+%cannot be detected by our data checking functions. 
+%Therefore,
+\pkg{dataMaid} was designed to be easily extendend with user-supplied
+functions for summarizing, viusalizing and checking data. We have
+provided a vignette in which we describe how \pkg{dataMaid} extensions
+can be made, such that they are integrate with the
+\code{makeDataReport()} function and with the other tools available in
+\pkg{dataMaid}.
 
@@
 
-The \code{makeDataReport()} function is the primary workhorse of \pkg{dataMaid} and
-this is the only function needed if a standard battery
-of tests are used to generate data reports. The data report itself is an overview document, intended for reading
-by humans, in either pdf, html or word (.docx) format. Appendix \ref{sec:appendix1}
-provides an example of a data report, produced by
-calling \code{makeDataReport()} on the dataset \code{toyData} available in
-\pkg{dataMaid}. The first two pages (excluding the frontpage) of this data report are
-shown in Figure~\ref{fig:example1} and the following two pages are shown in Figure~\ref{fig:example2}. \code{toyData} is a very
-small ($15$ observations of $6$ variables), artificial dataset which was created to
-illustrate the main capabilities of \pkg{dataMaid}. The following
-commands load the dataset and produce the report:
+The \code{makeDataReport()} function is the primary workhorse of
+\pkg{dataMaid} and it is the only function needed to generate a data
+report using the standard battery of tests. The data report itself is
+an overview document, intended for reading by humans, in either pdf,
+html or word (.docx) format. Appendix \ref{sec:appendix1} provides an
+example of a data report, produced by calling \code{makeDataReport()}
+on the dataset \code{toyData} available in \pkg{dataMaid}. The first
+two pages (excluding the frontpage) of this data report are shown in
+Figure~\ref{fig:example1} and the following two pages are shown in
+Figure~\ref{fig:example2}. \code{toyData} is a very small ($15$
+observations of $6$ variables), artificial dataset which was created
+with a lot of potential errors to illustrate the main capabilities of
+\pkg{dataMaid}. Section~\ref{sec:bigExample} shows an example of a data sreening
+process with a real dataset. The following commands load the dataset
+and produce the report:
 
@@
 
-By default, an \proglang{R} markdown file and a rendered pdf \hl{OBS: måske html/doc alt efter hvem der kalder funktionen?} overview
-document is produced, saved to the working directory and
-opened for immediate inspection. Such a data report consists of three parts, two of which 
-are presented in Figure~\ref{fig:example1}. First, an overview of what was done is presented under the
-title \textit{Data report overview}.
-Secondly, an index listing each variable along with an indication of whether it was found to be problematic or not is provided.
-Thirdly, as seen in Figure~\ref{fig:example2}, each variable in the
-dataset is presented in turn using (up to) three tools in the
+By default, an \proglang{R} markdown file and a rendered pdf or word
+overview document is produced, saved to the working directory and
+opened for immediate inspection. Such a data report consists of three
+parts, two of which are presented in Figure~\ref{fig:example1}. First,
+an overview of what was done is presented under the title \textit{Data
+  report overview}. Secondly, an index listing each variable along
+with an indication of whether it was found to be problematic or not is
+provided. Thirdly, as seen in Figure~\ref{fig:example2}, each variable
+in the dataset is presented in turn using (up to) three tools in the
 \textit{Variable list}: A table summarizing key features of the
-variable, a figure visualizing its distribution and a
-list of flagged issues, if any. For instance, as shown in Figure~\ref{fig:example2}, for the \code{numeric}-type variable
-\code{var2} from \code{toyData}, \code{makeDataReport()} has identified two values that
-are suspected to be miscoded missing values (\code{999} and \code{NaN}),
-while two values were also flagged as potential outliers that should
-be investigated more carefully.
-
-Though the \code{makeDataReport()} function is very easy to use, it should not be
-mistaken to be inflexible: By using the optional arguments of the function, 
-both the contents and the look of the data report can be molded 
-according to the user's needs. The most commonly used arguments are summarized in
+variable, a figure visualizing its distribution when relevant and a list of flagged
+issues, if any. For instance, as shown in Figure~\ref{fig:example2},
+for the \code{numeric}-type variable \code{events} from
+\code{toyData}, \code{makeDataReport()} has identified two values that
+are suspected to be miscoded missing values (\code{999} and
+\code{NaN}), while two values were also flagged as potential outliers
+that should be investigated more carefully.
+
+The arguments to \code{makeDataReport()} can be used to modify the
+contents and the look of the data report according to the user's
+needs. The most commonly used arguments are summarized in
 Table~\ref{table.cleanFormals} and they are grouped according to the
-part of the data assesment and report generation they influence. In order to understand
-this distinction, a glimpse of the inner structure of \code{makeDataReport()} is
-shown in Figure~\ref{figure:cleanStructure}. Below, we present a few
-examples on how to use the arguments from Table \ref{table.cleanFormals}
- to influence the output of a \code{makeDataReport()} call.
+part of the data assesment and report generation they influence. In
+order to understand this distinction, a glimpse of the inner structure
+of \code{makeDataReport()} is shown in
+Figure~\ref{figure:cleanStructure}. Below, we present a few examples
+on how to use the arguments from Table \ref{table.cleanFormals} to
+influence the output of a \code{makeDataReport()} call.
 
@@
 
-\subsection{Polishing off the arguments}
+\subsection{Dusting off the arguments}
 We begin with an example that is intended as an illustration of how
@@
 variables with potential problems, and with a limit of, say, maximum
-2 printed potential problematic value per check for each variable. Also, we can add the
+2 printed potential problematic values per check for each variable. Also, we can add the
 argument \code{replace = TRUE} in order to force \code{makeDataReport()} to
@@
 
-\subsection{Controlling contents: What summaries, visualizations and checks should be performed?}
-We will now move on to discuss how not only the structure of the
-data assesment steps is manipulated, but also its very contents. This is
-done through the summarize/visualize/check (SVC) steps, as illustrated
-in Figure \ref{figure:cleanStructure}.  \pkg{dataMaid} uses three
-different types of functions for performing these steps, namely
+\subsection{Controlling contents through summaries, visualizations and checks}
+
+\pkg{dataMaid} works through three different steps --- summarize,
+visualize, and check (SVC) --- for each variable in the dataset
+(illustrated in Figure \ref{figure:cleanStructure}).  Three different
+types of functions are used to perform these steps, namely
 \code{summaryFunction}s, \code{visualFunction}s and
-\code{checkFunction}s.  By default, \code{makeDataReport()} runs selected summary,
-visualization and check functions on each variable in the dataset, and
-the exact choice of these functions depends on the classes of the
-variables. For instance, though detection of outlier values might be
-interesting for numerical variables, it holds little meaning for
-factor variables, and therefore, numerical and factor variables need
-different checks. Table~\ref{table.SVCfunctions} lists all available
-summarize/visualize/check functions, but we can also use the
-\code{allSummaryFunctions()}, \code{allVisualFunctions()}, and
-\code{allCheckFunctions()} functions in \pkg{dataMaid} to print overview
-lists in \proglang{R}. For example, the implemented
+\code{checkFunction}s.  By default, \code{makeDataReport()} runs
+selected summary, visualization and check functions on each variable
+in the dataset, and the exact choice of these functions depends on the
+classes of the variables. For instance, detection of outlier values
+might be interesting for numerical variables, but it holds little
+meaning for factor variables, and therefore, numerical and factor
+variables need different checks. Table~\ref{table.SVCfunctions} lists
+all available summarize/visualize/check functions, but we can also use
+the \code{allSummaryFunctions()}, \code{allVisualFunctions()}, and
+\code{allCheckFunctions()} functions in \pkg{dataMaid} to print
+overview lists in \proglang{R}. For example, the implemented
 \code{summaryFunction}s are:
@@
 Assume that we wish to look further into a certain variable from
-\code{toyData}, namely \code{var2}. The data cleaning summary found some
+\code{toyData}, namely \code{events}. The data cleaning summary found some
 issues in this variable, and we would like to recall what these issues
@@
 <<>>=
-check(toyData$var2)
+check(toyData$events)
 @
@@
 <<>>=
-check(toyData$var2, checks = setChecks(numeric = "identifyMissing"))
+check(toyData$events, checks = setChecks(numeric = "identifyMissing"))
 @
@@
 <<>>=
-identifyMissing(toyData$var2)
+identifyMissing(toyData$events)
 @
@@
 <<>>=
-missVar2 <- identifyMissing(toyData$var2)
-str(missVar2)
+missEvents <- identifyMissing(toyData$events)
+str(missEvents)
 @
@@
 \code{problemValues}. If we decide that the values \code{999}
-and \code{NaN} in \code{var2} are in fact miscoded missing values, we can
+and \code{NaN} in \code{events} are in fact miscoded missing values, we can
 easily replace them with \code{NA}s:
@@
 <<>>=
-toyData$var2[toyData$var2 %in% missVar2$problemValues] <- NA
-identifyMissing(toyData$var2)
+toyData$events[toyData$events %in% missEvents$problemValues] <- NA
+identifyMissing(toyData$events)
 @
@@
 used to run the corresponding visualizations and summaries for each
-variable. See Figure~\ref{fig:example2} for the visualization output.
-
-<<>>=
-visualize(toyData$var2)
-summarize(toyData$var2)
+variable. See Figure~\ref{fig:example3} for the visualization output.
+
+<<toydata-events>>=
+visualize(toyData$events)
+summarize(toyData$events)
 @
@@
 \begin{center}
-\includegraphics[width=7.5cm]{toyData-var2.pdf}
+\includegraphics[width=7.5cm]{toyData-events.pdf}
 \end{center}
-\caption{Output from running \code{visualize()} on the variable \code{var2} from the
+\caption{Output from running \code{visualize()} on the variable \code{events} from the
 \code{toyData} dataset.}
-\label{fig:example2}
+\label{fig:example3}
 \end{figure}
@@
 <<>>=
-summarize(toyData$var2, 
+summarize(toyData$events, 
   summaries = setSummaries(
@@
 
-The syntax in this code chunk can be read as follows: "Summarize \code{var2} in \code{toyData}, and for \code{numeric} variables, set the summaries to be the default summary functions, except \code{variableType} and \code{countMissing}."
+The syntax in this code chunk can be read as follows: "Summarize \code{events} in \code{toyData}, and for \code{numeric} variables, set the summaries to be the default summary functions, except \code{variableType} and \code{countMissing}."
 
@@
 For instance, \code{ethnicity} is very reasonably marked as a potentially problematic variable as it includes only a single observation of "African American".
-However, a human reading this report will know that this does \textit{not} reflect a mistake in the data, but rather a perculiarity in the real world, and as such, it should not be cleaned out.
+However, a human reading this report will know that this does \textit{not} reflect a mistake in the data, but rather a peculiarity in the real world, and as such, it should not be cleaned out.
 
@@
 
-and we find that only a single variable in \code{toyData}, \code{var6} (for which all
+and we find that only a single variable in \code{toyData}, \code{spotifysong} (for which all
 observations have the value \code{"Irrelevant"}), is problem-free when using the default checks.

### 2017-11-08 latex/article_vol3.tex
First round of tests

@@
 
-But even when tools are available for identifying problems in a dataset, the activity of data cleaning still suffers from a challenge that has recieved increasing attention in the scientitic communities in the later years: Data cleaning is not very straight forward to document and therefore, reproducibility suffers.
-We present a new \proglang{R} package, \pkg{dataMaid} \citep{dataMaid}, whose most central purpose is to facilitate a data cleaning workflow where documentation is thoroughly integrated rather than a secondary add-on.
-This is accomplished by centering the data cleaning process around auto-generated data overview reports that summarize the contents of the current version of a dataset and flags potential problems that could be candidates for the next bullet on the cleaning agenda.
-
-A number of \proglang{R} packages made for other pre-analysis steps are
-already available, including \pkg{janitor} \citep{janitor}, \hl{assertive}, \pkg{dplyr}
-\citep{dplyr}, \pkg{tidyr}\citep{tidyr}, \pkg{data.table} \citep{data.table}, \pkg{DataCombine}
-\citep{DataCombine} , \pkg{validate} \citep{validate}, \pkg{assertr} \citep{assertr}, and
-\pkg{DataExplorer} \citep{DataExplorer}. These packages focus on
-different stages of the pre-analysis work, most of which are not
-really data cleaning.  \pkg{janitor} provides tools for data import
-with a particular emphasis on the challenges of getting neat data
-frames from Microsoft Excel data files. \hl{Mention assertive here}. \pkg{dplyr}, \pkg{tidyr}, \pkg{data.table}
-and \pkg{DataCombine} go a few steps further by providing a wide array
-of extremely powerful tools for data wrangling, including a number of
+But even when tools are available for identifying problems in a
+dataset, the activity of data cleaning still suffers from a challenge
+that has recieved increasing attention in the scientitic communities
+in the later years: Data cleaning is not very straight forward to
+document and therefore, reproducibility suffers. We present a new
+\proglang{R} package, \pkg{dataMaid} \citep{dataMaid}, whose most
+central purpose is to facilitate a supervised data quality screening
+workflow where documentation is thoroughly integrated rather than an
+add-on. This is accomplished by structuring the data screening around
+auto-generated data overview reports that summarize and flags
+potential problems in the dataset.
+
+But no matter how clever software tools we make, data cleaning remains
+to be a time consuming endeavor, which inherently requires human
+interaction since every dataset is different and the variables in the
+dataset can only be understood in the proper context of their
+origin. This often requires a collaborative effort between an expert
+in the field and a statistician or data scientist. In many
+situations, these errors are discovered in the process of the data
+analysis (e.g., a categorical variable with numeric labels for each
+category may be wrongly classified as a quantitative variable or a
+variable where all values have erroneously been coded to the same
+value), but in other cases a human with knowledge about the data
+context area is needed to identify possible mistakes in the data
+(e.g., if there are 4 categories for a variable that should only have
+3).
+
+The \pkg{dataMaid} approach to data screening, -quality assessment and
+-documentation is governed by two fundamental paradigms. First of all,
+there is no need for data cleaning to be an ad hoc procedure. Often,
+we have a very clear idea of what flags are raisable in a given
+dataset before we look at it, as we were the ones to produce it in the
+first place. This means that data cleaning can easily be a
+well-documented, well-specified procedure. In order to aid this
+paradigm, \pkg{dataMaid} provides easy-to-use, automated tools for
+data quality assessment in \proglang{R} \citep{R} on which data
+cleaning decisions can be made. This quality assessment is presented
+in an auto-generated overview document, readable by data analysts and
+field experts alike, thereby also contributing to an inter-field
+dialogue about the data at hand. Oftentimes, e.g., distinguishing
+between faulty codings of a numeric value and unusual, but correct,
+values requires problem-specific expertise that might not be held by
+the data analyst. Hopefully, having easy access to data summaries
+through \pkg{dataMaid} will help this necessary knowledge sharing.
+
+While \pkg{dataMaid}s primary raison d'être is auto-generating data
+quality assessment overview documents, we still wish to emphasize that
+it is \emph{not} a tool for unsupervised data cleaning. This qualifies
+as the second paradigm of \pkg{dataMaid}: Data cleaning decisions
+should always be made by humans. Therefore, \pkg{dataMaid} does not
+supply any tools for ``fixing'' errors in the data. However, we do
+provide interactive functions that can be used to identify potentially
+erroneous entries in a dataset and that can make it easier to solve
+data issues, one variable at a time.
+
+A number of \proglang{R} packages made for other pre-analysis steps
+are already available, including \pkg{janitor} \citep{janitor},
+\pkg{assertive} \citep{assertive}, \pkg{dplyr} \citep{dplyr},
+\pkg{tidyr}\citep{tidyr}, \pkg{data.table} \citep{data.table},
+\pkg{DataCombine} \citep{DataCombine}, \pkg{validate}
+\citep{validate}, \pkg{assertr}, and \citep{assertr}. These packages
+focus on different stages of the pre-analysis work.  \pkg{janitor}
+provides tools for data import with a particular emphasis on the
+challenges of getting neat data frames from Microsoft Excel data
+files. \pkg{dplyr}, \pkg{tidyr}, \pkg{data.table} and
+\pkg{DataCombine} go a few steps further by providing a wide array of
+extremely powerful tools for data wrangling, including a number of
 particularly useful functions for merging and working with very large
@@
 \begin{Soutput}
-   var1 var2  var3       var4 var5       var6
+# A tibble: 15 x 6
+     pill events region     change     id spotifysong
+   <fctr>  <dbl> <fctr>      <dbl> <fctr>      <fctr>
  1    red      1      a -0.6264538      1  Irrelevant
@@
 list of flagged issues, if any. For instance, as shown in Figure~\ref{fig:example2}, for the \code{numeric}-type variable
-\code{var2} from \code{toyData}, \code{makeDataReport()} has identified two values that
+\code{events} from \code{toyData}, \code{makeDataReport()} has identified two values that
 are suspected to be miscoded missing values (\code{999} and \code{NaN}),
@@
 \begin{Sinput}
-R> makeDataReport(toyData, output="html", render=FALSE, openResult=FALSE)
+R> makeDataReport(toyData, output="html", render=FALSE, openResult=FALSE, replace=TRUE)
 \end{Sinput}
@@
 $numeric
-[1] "variableType" "countMissing" "uniqueValues" "centralValue" "quartiles"   
-[6] "minMax"
+[1] "variableType" "countMissing" "uniqueValues" "centralValue"
+[5] "quartiles"    "minMax"
 
 $integer
-[1] "variableType" "countMissing" "uniqueValues" "centralValue" "quartiles"   
-[6] "minMax"
+[1] "variableType" "countMissing" "uniqueValues" "centralValue"
+[5] "quartiles"    "minMax"
 
@@
 $Date
-[1] "variableType" "countMissing" "uniqueValues" "centralValue" "minMax"      
-[6] "quartiles" 
-\end{Soutput}
-\end{Schunk} %$
+[1] "variableType" "countMissing" "uniqueValues" "centralValue"
+[5] "minMax"       "quartiles"   
+\end{Soutput}
+\end{Schunk}
 
@@
 \begin{Sinput}
-R> defaultCaracterSummaries()
+R> defaultCharacterSummaries()
 \end{Sinput}
@@
 \begin{Sinput}
-R> makeDataReport(toyData, 
-+    summaries = setSummaries(numeric = c("variableType","centralValue),
+R> makeDataReport(toyData, replace=TRUE,
++    summaries = setSummaries(numeric = c("variableType","centralValue"),
 +    integer = c("variableType", "centralValue")))
@@
 \begin{Sinput}
-R> makeDataReport(toyData, 
+R> makeDataReport(toyData, replace=TRUE,                               
 +    summaries = setSummaries(all = c("variableType", "centralValue")))
@@
 \begin{Sinput}
-R> makeDataReport(toyData, 
+R> makeDataReport(toyData, replace=TRUE,
 +    checks = setChecks(factor = "identifyWhitespace"))
@@
 \begin{Sinput}
-R> makeDataReport(toyData, checks = setChecks(factor = NULL))
+R> makeDataReport(toyData, checks = setChecks(factor = NULL), replace=TRUE)
 \end{Sinput}
@@
 \begin{Sinput}
-R> makeDataReport(toyData, preChecks = c("isKey", "isSupported"))
+R> makeDataReport(toyData, preChecks = c("isKey", "isSupported"), replace=TRUE)
 \end{Sinput}
@@
 \begin{Sinput}
-R> makeDataReport(toyData, visuals = setVisuals(all = "basicVisual"))
+R> makeDataReport(toyData, visuals = setVisuals(all = "basicVisual"), replace=TRUE)
 \end{Sinput}
@@
 \begin{Sinput}
->R complexData <- data.frame(complexVar = complex(100, real = 1:100, 
-+    imaginary = 3), numericVar = 1:100))
->R makeDataReport(complexData, treatXasY(complex = "numeric"))
-\end{Sinput}
-\end{Schunk}
-
-In this report, we will find that the two variables, \code{complexVar} and \code{numericVar} will be have identical presentations in the variable list, as treating a \code{complex} variable as a \code{numeric} means dropping the imaginary part of the complex numbers - which was the only thing setting the two variables apart in the first place.
+R> complexData <- data.frame(complexVar = complex(100, real = 1:100, imaginary = 3), numericVar = 1:100)
+R> makeDataReport(complexData, treatXasY=list(complex = "numeric"), replace=TRUE)
+\end{Sinput}
+\end{Schunk}
+
+In this report, we will find that the two variables, \code{complexVar} and \code{numericVar} will be have identical presentations in the variable list, as treating a \code{complex} variable as a \code{numeric} means dropping the imaginary part of the complex numbers which was the only thing setting the two variables apart in the first place.
 
@@
 \begin{Sinput}
-R> check(toyData$var2)
+R> check(toyData$events)
 \end{Sinput}
@@
 \end{Soutput}
-\end{Schunk}%$
+\end{Schunk}
 
@@
 \begin{Sinput}
-R> check(toyData$var2, checks = setChecks(numeric = "identifyMissing"))
+R> check(toyData$events, checks = setChecks(numeric = "identifyMissing"))
 \end{Sinput}
@@
 \begin{Sinput}
-R> identifyMissing(toyData$var2)
+R> identifyMissing(toyData$events)
 \end{Sinput}
@@
 \end{Soutput}
-\end{Schunk} %$
+\end{Schunk}
 
@@
 \begin{Sinput}
-R> missVar2 <- identifyMissing(toyData$var2)
-R> str(missVar2)
+R> missEvents <- identifyMissing(toyData$events)
+R> str(missEvents)
 \end{Sinput}
@@
  $ problem      : logi TRUE
- $ message      : chr "The following suspected missing value codes
-    enter as regular values: \\\"999\\\", \\\"NaN\\\"."
+ $ message      : chr "The following suspected missing value codes enter as regular values: \\\"999\\\", \\\"NaN\\\"."
  $ problemValues: num [1:2] 999 NaN
@@
 \begin{Sinput}
-R> toyData$var2[toyData$var2 %in% missVar2$problemValues] <- NA
-R> identifyMissing(toyData$var2)
+R> toyData$events[toyData$events %in% missEvents$problemValues] <- NA
+R> identifyMissing(toyData$events)
 \end{Sinput}
@@
 used to run the corresponding visualizations and summaries for each
-variable. See Figure~\ref{fig:example2} for the visualization output.
-
-\begin{Schunk}
-\begin{Sinput}
-R> visualize(toyData$var2)
-R> summarize(toyData$var2)
-\end{Sinput}
-\begin{Soutput}
-     Feature                   Result       
-[1,] "Variable type"           "numeric"    
-[2,] "Number of missing obs."  "4 (26.67 %)"
-[3,] "Number of unique values" "6"          
-[4,] "Median"                  "4"          
-[5,] "1st and 3rd quartiles"   "1.5; 6"     
-[6,] "Min. and max."           "1; 82"      
+variable. See Figure~\ref{fig:example3} for the visualization output.
+
+\begin{Schunk}
+\begin{Sinput}
+R> visualize(toyData$events)
+R> summarize(toyData$events)
+\end{Sinput}
+\begin{Soutput}
+$variableType
+Variable type: numeric
+$countMissing
+Number of missing obs.: 4 (26.67 %)
+$uniqueValues
+Number of unique values: 6
+$centralValue
+Median: 4
+$quartiles
+1st and 3rd quartiles: 1.5; 6
+$minMax
+Min. and max.: 1; 82
 \end{Soutput}
@@
 \begin{Sinput}
-R> summarize(toyData$var2, 
+R> summarize(toyData$events, 
 +    summaries = setSummaries(
@@
 \begin{Soutput}
-     Feature         Result 
-[1,] "Median"        "4"    
-[2,] "Min. and max." "1; 82"
-\end{Soutput}
-\end{Schunk} %$
-
-The syntax in this code chunk can be read as follows: "Summarize \code{var2} in \code{toyData}, and for \code{numeric} variables, set the summaries to be the default summary functions, except \code{variableType} and \code{countMissing}."
+$uniqueValues
+Number of unique values: 6
+$centralValue
+Median: 4
+$quartiles
+1st and 3rd quartiles: 1.5; 6
+$minMax
+Min. and max.: 1; 82
+\end{Soutput}
+\end{Schunk}
+
+The syntax in this code chunk can be read as follows: "Summarize \code{events} in \code{toyData}, and for \code{numeric} variables, set the summaries to be the default summary functions, except \code{variableType} and \code{countMissing}."
 
@@
 \begin{Sinput}
-R> presidentData$lastName[presidentData$lastName == " Truman"] <- 
-+    "Truman"
-R> presidentData$stateOfBirth[presidentData$stateOfBirth == "New york"] <- 
-+    "New York"
-R> presidentData$assassinationAttempt <- 
-+    factor(presidentData$assassinationAttempt)
-R> presidentData$ageAtInauguration <- 
-+    as.numeric(presidentData$ageAtInauguration)
-\end{Sinput}
-\end{Schunk}%$
+R> presidentData$lastName[presidentData$lastName == " Truman"] <- "Truman"
+R> presidentData$stateOfBirth[presidentData$stateOfBirth == "New york"] <- "New York"
+R> presidentData$assassinationAttempt <- factor(presidentData$assassinationAttempt)
+R> presidentData$ageAtInauguration <- as.numeric(presidentData$ageAtInauguration)
+\end{Sinput}
+\end{Schunk}
 
@@
 \begin{Sinput}
->R presidentData$ageAtInauguration <- 
-+    as.numeric(as.character(presidentData$ageAtInauguration))
+R> presidentData$ageAtInauguration <- as.numeric(as.character(presidentData$ageAtInauguration))
 \end{Sinput}
@@
 \end{Soutput}
+\end{Schunk}
+
+\begin{Schunk}
 \begin{Sinput}
@@
 R> birthdayOutlierVal <- 
-+    identifyOutliers(presidentData$birthday)$problemValues
++  +    identifyOutliers(presidentData$birthday)$problemValues
 \end{Sinput}
@@
 \begin{Sinput}
-> presidentData[presidentData$birthday == birthdayOutlierVal, ]
+R> presidentData[presidentData$birthday == birthdayOutlierVal, ]
 \end{Sinput}
@@
 46 Arathornson   Aragorn                 0 1300-03-01       Gondor
-   assassinationAttempt  sex ethnicity precidencyYears ageAtInauguration
-46                    1 Male Caucasian              NA              88
-   favoriteNumber
-46           8+0i
-\end{Soutput}
-\end{Schunk}%$
+   assassinationAttempt  sex ethnicity presidencyYears
+46                    1 Male Caucasian              NA
+   ageAtInauguration favoriteNumber
+46                87           8+0i
+\end{Soutput}
+\end{Schunk}
 
@@
 \end{Sinput}
-\end{Schunk} %$
+\end{Schunk}
 
@@
 \begin{Sinput}
->R summarize(presidentData$presidencyYears)
+R> summarize(presidentData$presidencyYears)
 \end{Sinput}
@@
 $uniqueValues
-Number of unique values: 11
+Number of unique values: 10
 $centralValue
@@
 $quartiles
-1st and 3rd quartiles: 4; 8
+1st and 3rd quartiles: 3.75; 8
 $minMax
@@
 \begin{Sinput}
-makeDataReport(presidentData, vol = "_cleaned", 
+R> makeDataReport(presidentData, vol = "_cleaned", 
 +    treatXasY = list(Name = "character", complex = "numeric"),
@@
 +    defaultCharacterChecks(remove = "identifyLoners")),
-+    reportTitle = "Dirty president data - cleaned")
++    reportTitle = "Dirty president data - cleaned",
++    replace=TRUE)
 \end{Sinput}
@@
 \begin{Sinput}
-R> makeDataReport(toyData, onlyProblematic = TRUE, mode = "check")
+R> makeDataReport(toyData, onlyProblematic = TRUE, mode = "check", replace=TRUE)
 \end{Sinput}
@@
 \begin{Soutput}
- var1  var2  var3  var4  var5  var6 
- TRUE  TRUE  TRUE  TRUE  TRUE FALSE 
-\end{Soutput}
-\end{Schunk}
-
-and we find that only a single variable in \code{toyData}, \code{var6} (for which all
+       pill      events      region      change          id 
+       TRUE        TRUE        TRUE        TRUE        TRUE 
+spotifysong 
+      FALSE 
+\end{Soutput}
+\end{Schunk}
+
+and we find that only a single variable in \code{toyData}, \code{spotifysong} (for which all
 observations have the value \code{"Irrelevant"}), is problem-free when using the default checks.
@@
 Alternatively, one can create the usual output report, not render it and then manually edit the produced \proglang{R} markdown file as wished. This can be done with the following command:
+
 \begin{Schunk}
@@
 \end{Schunk}
+
 After editing, the file can be rendered by calling the \code{render} function:
+
 \begin{Schunk}

### 2017-11-08 latex/dataMaid_toyData.Rmd
First round of tests

@@
 subtitle: "Autogenerated data summary from dataMaid"
-date: 2017-10-31 15:34:55
+date: 2017-11-08 12:08:33
 output: pdf_document
@@
 
-----------------------------------------------------------------------------------
+-----------------------------------------------------------------------------------------
 &nbsp;          Variable class     # unique values   Missing observations  Any problems? 
--------- ---------------- ----------------- ---------------------- ---------------
-[var1]   factor                           3                13.33 %    $\times$
+--------------- ---------------- ----------------- ---------------------- ---------------
+[pill]          factor                           3                13.33 %    $\times$
 
-[var2]   numeric                          7                26.67 %    $\times$
+[events]        numeric                          7                26.67 %    $\times$
 
-[var3]   factor                           7                 0.00 %    $\times$
+[region]        factor                           7                 0.00 %    $\times$
 
-[var4]   numeric                         15                 0.00 %    $\times$
+[change]        numeric                         15                 0.00 %    $\times$
 
-[var5]   factor                          15                 0.00 %    $\times$
+[id]            factor                          15                 0.00 %    $\times$
 
-[var6]   factor                           1                 0.00 %    $\times$    
-----------------------------------------------------------------------------------
+[spotifysong]   factor                           1                 0.00 %    $\times$    
+-----------------------------------------------------------------------------------------
 
 # Variable list
-## var1
+## pill
 
@@
 
-## var2
+## events
 
@@
 
-## var3
+## region
 
@@
 
-## var4
+## change
 
@@
 
-## var5
+## id
 
@@
 
-## var6
+## spotifysong
 
@@
 
- *  Report creation time: Tue Oct 31 2017 15:34:55
+ *  Report creation time: Wed Nov 08 2017 12:08:33
 
- *  dataMaid v0.9.7.9000 [Pkg: 2017-10-31 from local (ekstroem/dataMaid@NA)]
+ *  dataMaid v0.9.7.9000 [Pkg: 2017-11-08 from local (ekstroem/dataMaid@NA)]
 

### 2017-11-08 R/dataMaid-package.R

@@ -102,18 +102,18 @@
 #'
 #' @format A \code{data.frame} with 15 rows and 6 variables.
 #' \describe{
-#'    \item{var1}{A factor variable with two levels (\code{"red"} and \code{"blue"}) and a few
-#'    (correctly coded)  missing observations.}
-#'    \item{var2}{A numeric variable with one obvious outlier value (\code{82}), two miscoded
-#'    missing values (\code{999} and \code{NaN}) and a few correctly coded missing values.}
-#'    \item{var3}{A factor variable where two of the levels (\code{"other"} and \code{"OTHER"}
+#'    \item{pill}{A factor variable with two levels (\code{"red"} and \code{"blue"}) and a few
+#'    (correctly coded) missing observations. This represents the colour of a pill.}
+#'    \item{events}{A numeric variable with one obvious outlier value (\code{82}), two miscoded
+#'    missing values (\code{999} and \code{NaN}) and a few correctly coded missing values. The number of previous events.}
+#'    \item{region}{A factor variable where two of the levels (\code{"other"} and \code{"OTHER"}
 #'    are the same word with different case settings. Moreover, the variable includes a Stata-style
-#'    miscoded missing value (\code{"."})}.
-#'    \item{var4}{A numeric variable (random draws from a standard normal distribution).}
-#'    \item{var5}{A factor variable with unique codes for each observation (a character string
+#'    miscoded missing value (\code{"."}). Used to represent geographical regions or treatment centers.}.
+#'    \item{change}{A numeric variable (random draws from a standard normal distribution). Representing a change in a measured variable.}
+#'    \item{id}{A factor variable with unique codes for each observation (a character string
 #'    with a number between 1 and 15), i.e. a key variable.}
-#'    \item{var6}{A factor variable that has the same level (\code{"Irrelevant"}) for all
-#'    observations, i.e. a empty variable.}
+#'    \item{spotifysong}{A factor variable that has the same level (\code{"Irrelevant"}) for all
+#'    observations, i.e. a empty variable. The latest song played on Spotify.}
 #' }
 #' @source Artificial data
 #' @examples

### 2017-11-08 tests/testthat/atomic.R

@@ -1,0 +1,11 @@
+## Define atomic types
+typel <- c(TRUE, FALSE, TRUE, TRUE, FALSE)
+typei <- c(1L, 2L, 3L, 4L, 5L)
+typed <- c(1.0, 2.3, 4.5, 6.7, 8.9)
+typec <- c(1 + 2i, 2 + 0i, 3 + 3i, 4 - 0i, 5+2i)
+types <- c(LETTERS[1:5], LETTERS[1:5])
+typef <- factor(types)
+typelab <- labelled(types, labels=c(A="A", BB="B", CCC="C", DDDD="D", EEEEE="E"))
+typer <- sapply(types, charToRaw)
+typelist <- list(a=1:3, b=1:10)
+typeDate <- as.Date(c("1jan1960", "2jan1960", "31mar1960", "30jul1960"), "%d%b%Y")

### 2017-11-08 tests/testthat/testminMax.R

@@ -1,9 +1,32 @@
-context("dataMaid minMax")
+context("summary function: minMax")
 
 library(dataMaid)
+
+
+## Read atomic vectors
+source("atomic.R")
                                        
 ## Each type should result in a list
 test_that("minMax returns a summaryResult", {
+
+    ## First check the atomic vectors
+    expect_equal(minMax(typel)$value[1], 0L)
+    expect_equal(minMax(typel)$value[2], 1L)
+    expect_equal(minMax(typei)$value[1], 1L)
+    expect_equal(minMax(typei)$value[2], 5L)
+    expect_equal(minMax(typed)$value[1], 1.0)
+    expect_equal(minMax(typed)$value[2], 8.9)
+
+    expect_error(minMax(typec))
+    expect_error(minMax(types))
+    expect_error(minMax(typef))
+    expect_error(minMax(typer))
+    expect_error(minMax(typelab))
+    expect_error(minMax(typelist))
+    expect_equal(minMax(typeDate)$value[1], as.Date("1960-01-01"))
+
+
+    ## Check the output format    
     expect_is(minMax(1:5), "summaryResult")
     expect_is(minMax(c(NA, 1:5)), "summaryResult")    
     expect_is(minMax(c(NA, NA, NA)), "summaryResult")
@@ -13,7 +36,6 @@ test_that("minMax returns a summaryResult", {
     expect_equal(minMax(c(NA, 1:5))$value[1], 1)
     expect_equal(minMax(c(NA, NA, NA))$value[1], NA)
     expect_equal(minMax(c(NA, NA, Inf))$value[1], Inf)
-
 
 })
 

### 2017-11-08 tests/testthat/testvariableType.R

@@ -1,0 +1,40 @@
+context("summary function: variableType")
+
+library(dataMaid)
+
+
+## Read atomic vectors
+source("atomic.R")
+
+
+test_that("variableType returns the right value", {
+    ## First check the atomic vectors
+    expect_equal(variableType(typel)$value, "logical")
+    expect_equal(variableType(typei)$value, "integer")
+    expect_equal(variableType(typed)$value, "numeric")
+    expect_equal(variableType(typec)$value, "complex")
+    expect_equal(variableType(types)$value, "character")
+    expect_equal(variableType(typef)$value, "factor")
+    expect_equal(variableType(typelab)$value, "labelled")
+    expect_equal(variableType(typelist)$value, "list")
+    expect_equal(variableType(typeDate)$value, "Date")
+})
+
+
+
+## Each type should result in a list
+test_that("variableType returns a summaryResult", {
+
+    ## Check the output format    
+    expect_is(variableType(typel), "summaryResult")
+    expect_is(variableType(typei), "summaryResult")
+    expect_is(variableType(typed), "summaryResult")
+    expect_is(variableType(typec), "summaryResult")
+    expect_is(variableType(types), "summaryResult")
+    expect_is(variableType(typef), "summaryResult")
+    expect_is(variableType(typelab), "summaryResult")
+    expect_is(variableType(typelist), "summaryResult")
+    expect_is(variableType(typeDate), "summaryResult")
+
+})
+

### 2017-11-09 latex/article_vol3.Rnw
Updated paper

@@
 \smallskip Control output and post-processing \\
-\quad \code{output} & Type of output file to be produced (html, word (.docx) or pdf) & \code{"pdf"} \hl{OBS. Ændrer vi dette?}\\
+\quad \code{output} & Type of output file to be produced (html, word (.docx) or pdf) & \code{NULL} (use pdf if \LaTeX\ is found, otherwise word (if on windows), or html)\\
 \quad \code{render} & Should the output file be rendered from markdown? & \code{TRUE} \\
@@
 <<>>=
-makeDataReport(toyData, output="html", render=FALSE, openResult=FALSE, replace=TRUE)
+makeDataReport(toyData, output="html", render=FALSE,
+  openResult=FALSE, replace=TRUE)
 @
@@
 <<>>=
-makeDataReport(toyData, preChecks = c("isKey", "isSupported"), replace=TRUE)
+makeDataReport(toyData, preChecks = c("isKey", "isSupported"), 
+  replace=TRUE)
 @
@@
 <<>>=
-makeDataReport(toyData, visuals = setVisuals(all = "basicVisual"), replace=TRUE)
+makeDataReport(toyData, visuals = setVisuals(all = "basicVisual"), 
+  replace=TRUE)
 @
@@
 Each of these stages are controllable using appropriate function
-arguments in \code{makeDataReport()}, and above we have shown examples of how
-to tweak them to modify the data cleaning outputs. However, if the
-dataset at hand requires new, additional checks, then more control is
-needed \hl{ and we have provided a detailed vignette that explains how to modify and
+arguments in \code{makeDataReport()}, and above we have shown examples
+of how to tweak them to modify the data cleaning outputs. However, if
+the dataset at hand requires new, additional checks, then more control
+is needed a package vignette explains the details of how to modify and
 expand the possibilities by producing new summary, visual, and check
-functions in [REFERENCE]}.
+functions.
 
@@
 <<warning=FALSE>>=
-complexData <- data.frame(complexVar = complex(100, real = 1:100, imaginary = 3), numericVar = 1:100)
-makeDataReport(complexData, treatXasY=list(complex = "numeric"), replace=TRUE)
+complexData <- data.frame(complexVar = complex(100, real = 1:100, 
+  imaginary = 3), numericVar = 1:100)
+makeDataReport(complexData, treatXasY=list(complex = "numeric"), 
+  replace=TRUE)
 @
@@
 
-<<toydata-events>>=
+<<label=toydataevents,fig=TRUE,include=FALSE>>=
 visualize(toyData$events)
@@
 \begin{center}
-\includegraphics[width=7.5cm]{toyData-events.pdf}
+\includegraphics[width=7.5cm]{article_vol3-toydataevents.pdf}
 \end{center}
-\caption{Output from running \code{visualize()} on the variable \code{events} from the
+\caption{Output from running \code{visualize()} on variable \code{events} from the
 \code{toyData} dataset.}
@@
 Larger versions of the pages can be seen in
-  Appendix~\ref{sec:appendix2}}
+  Appendix~\ref{sec:appendix2}.}
 \label{fig:bigExampleP23}
@@
 \caption{Part of the data overview report generated for the \code{presidentData} dataset. Here, we see pages 4 and 5 from the \textit{Variable list}. Larger versions of the pages can be seen in
-  Appendix~\ref{sec:appendix2}}
+  Appendix~\ref{sec:appendix2}.}
 \label{fig:bigExampleP45}
@@
 \end{description}
+
+<<>>=
+data(presidentData)
+head(presidentData)
+@
+
 We discuss the results of a data overview report generated for this dataset below, but first there are a few special features of the dataset and wishes for the data report that require us to customize it using some of the arguments for \code{makeDataReport}.
@@
 <<>>=
-data(presidentData)
 makeDataReport(presidentData, replace = TRUE, 
@@
 The first problem that can be spotted from these first four pages is the surprising number of observations: Anno 2017, there have only been 45 US presidents.
-Therefore, having 46 observations reveals that the dataset contains a blind passenger.
+Therefore, having 46 observations reveal that the dataset contains a blind passenger.
 For instance, if the dataset was constructed as a subset of a more general "World leaders" dataset, this type of problem could occur due to wrongful nationality classification.
@@
 presidentData$lastName[presidentData$lastName == " Truman"] <- "Truman"
-presidentData$stateOfBirth[presidentData$stateOfBirth == "New york"] <- "New York"
-presidentData$assassinationAttempt <- factor(presidentData$assassinationAttempt)
-presidentData$ageAtInauguration <- as.numeric(presidentData$ageAtInauguration)
+presidentData$stateOfBirth[presidentData$stateOfBirth == "New york"] <- 
+  "New York"
+presidentData$assassinationAttempt <- 
+  factor(presidentData$assassinationAttempt)
+presidentData$ageAtInauguration <- 
+  as.numeric(presidentData$ageAtInauguration)
 @
@@
 <<>>=
-presidentData$ageAtInauguration <- as.numeric(as.character(presidentData$ageAtInauguration))
+presidentData$ageAtInauguration <- 
+  as.numeric(as.character(presidentData$ageAtInauguration))
 @
@@
 birthdayOutlierVal <- 
-+    identifyOutliers(presidentData$birthday)$problemValues
+  identifyOutliers(presidentData$birthday)$problemValues
 @
@@
 Trump, on the other hand, has a missing value for his presidency duration, which is in fact reasonable as we cannot know how long it will be yet (anno 2017).
-Presidents William Harrison, Zachary Taylor, James Garfield, Warren Harding and Gerald Ford were identified to have very brief presidencies, but these are not mistakes, as a textbook look-up can tell us.
-Thus the only mistake left to fix is Obama's infinite presidency:
+Presidents William Harrison, Zachary Taylor, James Garfield, Warren Harding and Gerald Ford were identified to have very brief presidencies, but these are not mistakes, as any US history textbook can tell us.
+Thus, the only mistake left to fix is Obama's infinite presidency:
 
@@
 
-\subsection[Incorporating dataMaid in a programmer's workflow]{Incorporating \pkg{dataMaid} in a programmer's workflow}
+\subsection[Incorporating dataMaid in an automated workflow]{Incorporating \pkg{dataMaid} in automated workflow}
 The default settings of \code{makeDataReport()} have been set to facilitate easy and quick data report generation, but unfortunately, this also means that it is not ideal for a more programming-oriented workflow, where the function might not be called by a human.
@@
 
-\begin{Schunk}
-\begin{Sinput}
->R makeDataReport(toyData, open = FALSE, quiet = "silent",
-+   vol = paste("_", format(Sys.time(), "%m-%d%-%y_%H.%M"), sep = ""))
-\end{Sinput}
-\end{Schunk}
+<<eval=FALSE>>=
+makeDataReport(toyData, open = FALSE, quiet = "silent",
+  vol = paste("_", format(Sys.time(), "%m-%d%-%y_%H.%M"), sep = ""))
+@
 
@@
 If the dataset is large and the time available for reading through the data report is scarce, it can be convenient to only make a report concerning the variabels that were flagged to be problematic.
-This can be achieved by using the \code{onlyProblematic} argument in
+This can be achieved by using the \code{onlyProblematic} argument for
 \code{makeDataReport()}. By specifying \code{onlyProblematic = TRUE}, only
@@
 included in other \proglang{R} markdown files by setting the
-\code{standAlone} argument in \code{makeDataReport()} to remove the preamble
+\code{standAlone=FALSE} argument in \code{makeDataReport()} to remove the preamble
 from the output \proglang{R} markdown file. Note that it is still
@@
 If it is important that the embedded \pkg{dataMaid} document can be
-rendered to any of these three file types, we recommend setting
-\code{twoCols = FALSE} and \code{output = "html"} in \code{makeDataReport()}, thereby
-essentially removing almost all output type specific formatting code
+rendered to any of these three file types, we recommend setting the
+\code{twoCols = FALSE} and \code{output = "html"} arguments in \code{makeDataReport()}. This 
+essentially removes almost all output type specific formatting code
 from the generated \proglang{R} markdown file.
@@
 
-In this paper we have introduced the \proglang{R} package \pkg{dataMaid}
-for performing reproducible error detection, data overview reports and data cleaning. The package provides a general and extendable framework for
-identifying potential errors and for creating human-readable summary
-documents that will help investigators to identify possible errors in
-the data.
-
-While the the current release is stable, the authors have an interest
-in further developing the functionality of \pkg{dataMaid} by providing
-more summary, visual, and check function as part of the
-package. We are also currently considering adding options to handle
-repeated measurement, where the visualizations --- in particular ---
-might be improved by visualizing measurements over time. In addition,
-an online \pkg{shiny} \citep{shiny} application where users that are not
-\proglang{R}-savvy can upload their data and get a data cleaning
-document is currently being developed.
+In this paper we have introduced the \proglang{R} package
+\pkg{dataMaid} for performing reproducible error detection, data
+overview reports and data screening. The package provides a general and
+extendable framework for identifying potential errors and for creating
+human-readable summary documents that will help investigators to
+identify possible errors in the data.
+
+%While the the current release is stable, the authors have an interest
+%in further developing the functionality of \pkg{dataMaid} by providing
+%more summary, visual, and check function as part of the
+%package. 
+We are also currently considering adding options to handle repeated
+measurement, where the visualizations --- in particular --- might be
+improved by visualizing measurements over time. In addition, an online
+\pkg{shiny} \citep{shiny} application is currently being developed
+such that non-\proglang{R}-savvy users can upload their data online
+and get a data cleaning document.
 

### 2017-11-09 latex/article_vol3.tex
Updated paper

@@
 \pkg{DataCombine} \citep{DataCombine}, \pkg{validate}
-\citep{validate}, \pkg{assertr}, and \citep{assertr}. These packages
+\citep{validate}, and \pkg{assertr} \citep{assertr}. These packages
 focus on different stages of the pre-analysis work.  \pkg{janitor}
@@
 
-By default, an \proglang{R} markdown file and a rendered pdf \hl{OBS: måske html/doc alt efter hvem der kalder funktionen?} overview
-document is produced, saved to the working directory and
-opened for immediate inspection. Such a data report consists of three parts, two of which 
-are presented in Figure~\ref{fig:example1}. First, an overview of what was done is presented under the
-title \textit{Data report overview}.
-Secondly, an index listing each variable along with an indication of whether it was found to be problematic or not is provided.
-Thirdly, as seen in Figure~\ref{fig:example2}, each variable in the
-dataset is presented in turn using (up to) three tools in the
+By default, an \proglang{R} markdown file and a rendered pdf or word
+overview document is produced, saved to the working directory and
+opened for immediate inspection. Such a data report consists of three
+parts, two of which are presented in Figure~\ref{fig:example1}. First,
+an overview of what was done is presented under the title \textit{Data
+  report overview}. Secondly, an index listing each variable along
+with an indication of whether it was found to be problematic or not is
+provided. Thirdly, as seen in Figure~\ref{fig:example2}, each variable
+in the dataset is presented in turn using (up to) three tools in the
 \textit{Variable list}: A table summarizing key features of the
-variable, a figure visualizing its distribution and a
-list of flagged issues, if any. For instance, as shown in Figure~\ref{fig:example2}, for the \code{numeric}-type variable
-\code{events} from \code{toyData}, \code{makeDataReport()} has identified two values that
-are suspected to be miscoded missing values (\code{999} and \code{NaN}),
-while two values were also flagged as potential outliers that should
-be investigated more carefully.
-
-Though the \code{makeDataReport()} function is very easy to use, it should not be
-mistaken to be inflexible: By using the optional arguments of the function, 
-both the contents and the look of the data report can be molded 
-according to the user's needs. The most commonly used arguments are summarized in
+variable, a figure visualizing its distribution when relevant and a list of flagged
+issues, if any. For instance, as shown in Figure~\ref{fig:example2},
+for the \code{numeric}-type variable \code{events} from
+\code{toyData}, \code{makeDataReport()} has identified two values that
+are suspected to be miscoded missing values (\code{999} and
+\code{NaN}), while two values were also flagged as potential outliers
+that should be investigated more carefully.
+
+The arguments to \code{makeDataReport()} can be used to modify the
+contents and the look of the data report according to the user's
+needs. The most commonly used arguments are summarized in
 Table~\ref{table.cleanFormals} and they are grouped according to the
-part of the data assesment and report generation they influence. In order to understand
-this distinction, a glimpse of the inner structure of \code{makeDataReport()} is
-shown in Figure~\ref{figure:cleanStructure}. Below, we present a few
-examples on how to use the arguments from Table \ref{table.cleanFormals}
- to influence the output of a \code{makeDataReport()} call.
+part of the data assesment and report generation they influence. In
+order to understand this distinction, a glimpse of the inner structure
+of \code{makeDataReport()} is shown in
+Figure~\ref{figure:cleanStructure}. Below, we present a few examples
+on how to use the arguments from Table \ref{table.cleanFormals} to
+influence the output of a \code{makeDataReport()} call.
 
@@
 \begin{Sinput}
-R> makeDataReport(toyData, output="html", render=FALSE, openResult=FALSE, replace=TRUE)
+R> makeDataReport(toyData, output="html", render=FALSE,
++    openResult=FALSE, replace=TRUE)
 \end{Sinput}
@@
 \begin{Sinput}
-R> makeDataReport(toyData, preChecks = c("isKey", "isSupported"), replace=TRUE)
+R> makeDataReport(toyData, preChecks = c("isKey", "isSupported"), 
++    replace=TRUE)
 \end{Sinput}
@@
 \begin{Sinput}
-R> makeDataReport(toyData, visuals = setVisuals(all = "basicVisual"), replace=TRUE)
+R> makeDataReport(toyData, visuals = setVisuals(all = "basicVisual"), 
++    replace=TRUE)
 \end{Sinput}
@@
 \begin{Sinput}
-R> complexData <- data.frame(complexVar = complex(100, real = 1:100, imaginary = 3), numericVar = 1:100)
-R> makeDataReport(complexData, treatXasY=list(complex = "numeric"), replace=TRUE)
+R> complexData <- data.frame(complexVar = complex(100, real = 1:100, 
++    imaginary = 3), numericVar = 1:100)
+R> makeDataReport(complexData, treatXasY=list(complex = "numeric"), 
++    replace=TRUE)
 \end{Sinput}
@@
 \end{description}
+
+\begin{Schunk}
+\begin{Sinput}
+R> data(presidentData)
+R> head(presidentData)
+\end{Sinput}
+\begin{Soutput}
+    lastName firstName orderOfPresidency   birthday  stateOfBirth
+1 Washington    George                 1 1732-02-22      Virginia
+2      Adams      John                 2 1735-10-30 Massachusetts
+3  Jefferson    Thomas                 3 1743-04-13      Virginia
+4    Madison     James                 4 1751-03-16      Virginia
+5     Monroe     James                 5 1758-04-28      Virginia
+6      Adams      John                 6 1767-07-11 Massachusetts
+  assassinationAttempt  sex ethnicity presidencyYears
+1                    0 Male Caucasian               7
+2                    0 Male Caucasian               3
+3                    0 Male Caucasian               8
+4                    0 Male Caucasian               8
+5                    0 Male Caucasian               8
+6                    0 Male Caucasian               4
+  ageAtInauguration favoriteNumber
+1                57    3+0.000000i
+2                61    4+0.000000i
+3                57    0+1.414214i
+4                57   10+0.000000i
+5                58    3+0.000000i
+6                57    9+0.000000i
+\end{Soutput}
+\end{Schunk}
+
 We discuss the results of a data overview report generated for this dataset below, but first there are a few special features of the dataset and wishes for the data report that require us to customize it using some of the arguments for \code{makeDataReport}.
@@
 \begin{Sinput}
-R> data(presidentData)
 R> makeDataReport(presidentData, replace = TRUE, 
@@
 R> presidentData$lastName[presidentData$lastName == " Truman"] <- "Truman"
-R> presidentData$stateOfBirth[presidentData$stateOfBirth == "New york"] <- "New York"
-R> presidentData$assassinationAttempt <- factor(presidentData$assassinationAttempt)
-R> presidentData$ageAtInauguration <- as.numeric(presidentData$ageAtInauguration)
+R> presidentData$stateOfBirth[presidentData$stateOfBirth == "New york"] <- 
++    "New York"
+R> presidentData$assassinationAttempt <- 
++    factor(presidentData$assassinationAttempt)
+R> presidentData$ageAtInauguration <- 
++    as.numeric(presidentData$ageAtInauguration)
 \end{Sinput}
@@
 \begin{Sinput}
-R> presidentData$ageAtInauguration <- as.numeric(as.character(presidentData$ageAtInauguration))
+R> presidentData$ageAtInauguration <- 
++    as.numeric(as.character(presidentData$ageAtInauguration))
 \end{Sinput}
@@
 R> birthdayOutlierVal <- 
-+  +    identifyOutliers(presidentData$birthday)$problemValues
++    identifyOutliers(presidentData$birthday)$problemValues
 \end{Sinput}
@@
 \begin{Sinput}
->R makeDataReport(toyData, open = FALSE, quiet = "silent",
+R> makeDataReport(toyData, open = FALSE, quiet = "silent",
 +    vol = paste("_", format(Sys.time(), "%m-%d%-%y_%H.%M"), sep = ""))

### 2017-11-09 latex/dataMaid_toyData.Rmd
Updated paper

@@
 subtitle: "Autogenerated data summary from dataMaid"
-date: 2017-11-08 12:08:33
+date: 2017-11-09 02:44:43
 output: pdf_document
@@
 
- *  Report creation time: Wed Nov 08 2017 12:08:33
+ *  Report creation time: Thu Nov 09 2017 02:44:43
 

### 2017-11-09 latex/article_vol3.Rnw
Minor changes regarding typos etc and how figures are included.

@@
 \Plainauthor{Anne Helby Petersen, Claus Thorn Ekstr\o m} %% comma-separated
-\Plaintitle{{dataMaid}: your assistant for data documentation in R} %% without formatting
+\Plaintitle{{dataMaid}: your assistant for documenting supervised data quality screening in R} %% without formatting
 \Shorttitle{\pkg{dataMaid}: your assistant for data documentation in R} %% a short title (if necessary)
@@
 \citep{editrules} and \pkg{deducorrect} \citep{deducorrect} from the
-same authors) and \pkg{assertive} offers tools for identifying errors
+same authors) and \pkg{assertive} offer tools for identifying errors
 in a dataset by checking the state of the variable given a set of
@@
 quite similar to that of \pkg{dataMaid} and to the paradigms presented
-below. This package provides a few simple, but practical, tools for
+above. This package provides a few simple, but practical, tools for
 exploratory data analysis, including automated
@@
 by the second paradigm above.
-Next, we show step-by-step how the data report and the interactive mode of \pkg{dataMaid} can be combined to conduct a well-documented, systematic data cleaning in Section \ref{sec:bigExample}.
+Next, we show step-by-step how the data report mode and the interactive mode of \pkg{dataMaid} can be combined to conduct a well-documented, systematic data cleaning in Section \ref{sec:bigExample}.
 Here, we assess and clean a dirty dataset with information about the US presidents.
@@
 \pkg{dataMaid} was designed to be easily extendend with user-supplied
-functions for summarizing, viusalizing and checking data. We have
+functions for summarizing, viusalizing and checking data. In the package, we have
 provided a vignette in which we describe how \pkg{dataMaid} extensions
@@
   dataset.
-Here, we see a description of each variable in the dataset, consiting of a summary table, a visualization and a indication of what problems were flagged for the variable (if any).
+Here, we see a description of each variable in the dataset, consisting of a summary table, a visualization and an indication of what problems were flagged for the variable (if any).
 At last, a few lines of metadata about the \code{makeDataReport()} are included for enhancing reproducibility.
@@
 
-By default, an \proglang{R} markdown file and a rendered pdf or word
+By default, an \proglang{R} markdown file and a rendered pdf,  word or html
 overview document is produced, saved to the working directory and
@@
 \smallskip Control output and post-processing \\
-\quad \code{output} & Type of output file to be produced (html, word (.docx) or pdf) & \code{NULL} (use pdf if \LaTeX\ is found, otherwise word (if on windows), or html)\\
+\quad \code{output} & Type of output file to be produced (html, word (.docx) or pdf) & \code{NULL} (use pdf if \LaTeX\ is found, otherwise Word (if on Windows), or html)\\
 \quad \code{render} & Should the output file be rendered from markdown? & \code{TRUE} \\
@@
 the dataset at hand requires new, additional checks, then more control
-is needed a package vignette explains the details of how to modify and
+is needed. The package contains a vignette that explains the details of how to modify and
 expand the possibilities by producing new summary, visual, and check
@@
 \begin{itemize}
-\item In the birthday variable, there is an entry with the date March 1, 1300 which is a bit of an outlier. 
+\item In the \code{birthday} variable, there is an entry with the date March 1, 1300 which is a bit of an outlier. 
 \item Among the states in which the presidents were born, New York was mistakenly spelled with a lower case "Y" in at least one entry.
@@
 
-Next up is the unlikely US president birth day of March 1, 1300.
+Next up is the unlikely US president birthday of March 1, 1300.
 In order to understand if this is a generally problematic observation, or if it is just a mistyped observation, we inspect the full data entry for this person.
@@
 Trump, on the other hand, has a missing value for his presidency duration, which is in fact reasonable as we cannot know how long it will be yet (anno 2017).
-Presidents William Harrison, Zachary Taylor, James Garfield, Warren Harding and Gerald Ford were identified to have very brief presidencies, but these are not mistakes, as any US history textbook can tell us.
+Presidents William Harrison, Zachary Taylor and James Garfield were identified to have very brief presidencies, but these are not mistakes, as any US history textbook can tell us.
 Thus, the only mistake left to fix is Obama's infinite presidency:
@@
 Please note that the data report does contain information about who, when and how concerning its generation, so even though the default choices for file names do not make it easy to tell different reports for the same dataset apart, it should be rather easy when inspecting the report manually.
-These three problems can easily be solved by use of the arguments of \code{makeDataReport()}.
+
+The three problems can easily be solved by use of the arguments of \code{makeDataReport()}.
 Whether or not the outputted file is opened can be controlled through the argument \code{open}.
@@
 \subsection[Using dataMaid for problem flagging]{Using \pkg{dataMaid} for problem flagging}
-If the dataset is large and the time available for reading through the data report is scarce, it can be convenient to only make a report concerning the variabels that were flagged to be problematic.
+If the dataset is large and the time available for reading through the data report is scarce, it can be convenient to only make a report concerning the variables that were flagged to be problematic.
 This can be achieved by using the \code{onlyProblematic} argument for

### 2017-11-09 latex/article_vol3.Rnw
Minor changes regarding typos etc and how figures are included.

@@
 \begin{center}
-\frame{\includegraphics[width=7.5cm,page=2]{dataMaid_toyData.pdf}}
-\frame{\includegraphics[width=7.5cm,page=3]{dataMaid_toyData.pdf}}
+\frame{\includegraphics[width=7.5cm,page=2]{dataMaid_toyData_appA.pdf}}
+\frame{\includegraphics[width=7.5cm,page=3]{dataMaid_toyData_appA.pdf}}
 %\includepdf[pages={2}, pagecommand={}]{dataMaid_testData.pdf}
@@
 \begin{center}
-\frame{\includegraphics[width=7.5cm,page=4]{dataMaid_toyData.pdf}}
-\frame{\includegraphics[width=7.5cm,page=5]{dataMaid_toyData.pdf}}
+\frame{\includegraphics[width=7.5cm,page=4]{dataMaid_toyData_appA.pdf}}
+\frame{\includegraphics[width=7.5cm,page=5]{dataMaid_toyData_appA.pdf}}
 %\includepdf[pages={2}, pagecommand={}]{dataMaid_testData.pdf}
@@
 \begin{center}
-\frame{\includegraphics[width=7.5cm,page=1]{dataMaid_presidentData.pdf}}
-\frame{\includegraphics[width=7.5cm,page=2]{dataMaid_presidentData.pdf}}\\
+\frame{\includegraphics[width=7.5cm,page=1]{dataMaid_presidentData_appB.pdf}}
+\frame{\includegraphics[width=7.5cm,page=2]{dataMaid_presidentData_appB.pdf}}\\
 \end{center}
@@
 \begin{center}
-\frame{\includegraphics[width=7.5cm,page=3]{dataMaid_presidentData.pdf}}
-\frame{\includegraphics[width=7.5cm,page=4]{dataMaid_presidentData.pdf}}
+\frame{\includegraphics[width=7.5cm,page=3]{dataMaid_presidentData_appB.pdf}}
+\frame{\includegraphics[width=7.5cm,page=4]{dataMaid_presidentData_appB.pdf}}
 \end{center}
@@
 \begin{center}
-\frame{\includegraphics[width=7.5cm,page=5]{dataMaid_presidentData.pdf}}
-\frame{\includegraphics[width=7.5cm,page=6]{dataMaid_presidentData.pdf}}
+\frame{\includegraphics[width=7.5cm,page=5]{dataMaid_presidentData_appB.pdf}}
+\frame{\includegraphics[width=7.5cm,page=6]{dataMaid_presidentData_appB.pdf}}
 \end{center}
@@
 % \section{Appendix A: cleaning the toyData data frame} \label{sec:appendix1}
-\includepdf[pages=2, pagecommand={\section{Data report for the toyData dataset} \label{sec:appendix1}}, frame=true, noautoscale=true, scale=0.7]{dataMaid_toyData.pdf}
-\includepdf[pages=3-4, pagecommand={}, frame=true, noautoscale=true, scale=0.7,pagecommand={}]{dataMaid_toyData.pdf}
-
-\includepdf[pages=2, pagecommand={\section{Data report for the presidentData dataset} \label{sec:appendix2}}, frame=true, noautoscale=true, scale=0.7]{dataMaid_presidentData.pdf}
-\includepdf[pages=3-6, pagecommand={}, frame=true, noautoscale=true, scale=0.7]{dataMaid_presidentData.pdf}
+\includepdf[pages=2, pagecommand={\section{Data report for the toyData dataset} \label{sec:appendix1}}, frame=true, noautoscale=true, scale=0.7]{dataMaid_toyData_appA.pdf}
+\includepdf[pages=3-4, pagecommand={}, frame=true, noautoscale=true, scale=0.7,pagecommand={}]{dataMaid_toyData_appA.pdf}
+
+\includepdf[pages=2, pagecommand={\section{Data report for the presidentData dataset} \label{sec:appendix2}}, frame=true, noautoscale=true, scale=0.7]{dataMaid_presidentData_appB.pdf}
+\includepdf[pages=3-6, pagecommand={}, frame=true, noautoscale=true, scale=0.7]{dataMaid_presidentData_appB.pdf}
 

### 2017-11-09 latex/dataMaid_toyData.Rmd
CRAN pre v10 check

@@
 subtitle: "Autogenerated data summary from dataMaid"
-date: 2017-11-09 02:44:43
+date: 2017-11-09 15:41:24
 output: pdf_document
@@
 
- *  Report creation time: Thu Nov 09 2017 02:44:43
+ *  Report creation time: Thu Nov 09 2017 15:41:24
 
- *  dataMaid v0.9.7.9000 [Pkg: 2017-11-08 from local (ekstroem/dataMaid@NA)]
+ *  dataMaid v0.9.7.9000 [Pkg: 2017-11-09 from local (ekstroem/dataMaid@NA)]
 
- *  R version 3.4.1 (2017-06-30).
+ *  R version 3.4.2 (2017-09-28).
 

### 2017-11-09 R/isCPR.R

@@ -13,7 +13,7 @@
 #' \code{$problem} (a logical indicating whether the variable consists
 #' of CPR numbers), \code{$message} (if a problem was found,
 #' the following message: "Warning: The variable seems to consist of
-#' Danish civil regristration (CPR) numbers.",
+#' Danish civil registration (CPR) numbers.",
 #' otherwise "") and \code{$problemValues} (always \code{NULL}).
 #'
 #' @examples

### 2017-11-13 R/makeDataReport.R
Updated system calls to remove output on all platforms

@@ -287,69 +287,73 @@ makeDataReport <- function(data, output=NULL, render=TRUE,
   
   
   #Set output default if output is NULL or check for valid values otherwise
-  makeOutputWarning <- FALSE
-  if (!is.null(output)) {
-    if (length(output) > 1) {
-      output <- output[1]
-      warning("Output argument was wrongfully given as a vector. Only the first entry was used.")
-    }
-    if (!(output %in% c("pdf", "html", "word"))) {
-      output <- NULL
+    makeOutputWarning <- FALSE
+    if (!is.null(output)) {
+        if (length(output) > 1) {
+            output <- output[1]
+            warning("Output argument was wrongfully given as a vector. Only the first entry was used.")
+        }
+        if (!(output %in% c("pdf", "html", "word"))) {
+            output <- NULL
       makeOutputWarning <- TRUE
-    } 
-  }
-  if (is.null(output)) {
-    xelatexTest <- suppressWarnings(system("xelatex --version", show.output.on.console = FALSE)) == 0
-    pdflatexTest <- suppressWarnings(system("pdflatex --version", show.output.on.console = FALSE)) == 0
-    lualatexTest <- suppressWarnings(system("lualatex --version", show.output.on.console = FALSE)) == 0
-    if (any(c(xelatexTest, pdflatexTest, lualatexTest))) {
-      output <- "pdf"
-    } else {
-      if (identical(as.character(Sys.info()["sysname"]),"Windows")) {
-        output <- "word"
-      } else output <- "html"
+        } 
     }
-    if (makeOutputWarning) {
-      warning(paste("No valid output option was chosen. ", 
-                    "Therefore, output was set to ", output, ".", sep = ""))
+    if (is.null(output)) {
+        xelatexTest <- suppressWarnings(system2("xelatex", args=c("--version"), stdout=NULL, stderr=NULL)) == 0
+        pdflatexTest <- suppressWarnings(system2("pdflatex", args=c("--version"), stdout=NULL, stderr=NULL)) == 0
+        lualatexTest <- suppressWarnings(system2("lualatex", args=c("--version"), stdout=NULL, stderr=NULL)) == 0
+        
+        ##    xelatexTest <- suppressWarnings(system2("xelatex --version", show.output.on.console = FALSE)) == 0
+        ##   pdflatexTest <- suppressWarnings(system2("pdflatex --version", show.output.on.console = FALSE)) == 0
[…]

### 2017-11-14 latex/article_vol3.Rnw
Updated system calls to remove output on all platforms
Did spell checking and fixed mistakes. Note: This file has not been compiled!

@@
   We describe an \proglang{R} package, \pkg{dataMaid}, which
-  implements an extensive and customizeable suite of quality
+  implements an extensive and customizable suite of quality
   assessment aids that can be applied to a dataset in order to
@@
 dataset, the activity of data cleaning still suffers from a challenge
-that has recieved increasing attention in the scientitic communities
+that has received increasing attention in the scientific communities
 in the later years: Data cleaning is not very straight forward to
@@
 --- and very nice --- tools to those of \pkg{validate}, but without
-any amibitions of conducting auto-cleaning.
+any ambitions of conducting auto-cleaning.
 
@@
 %Therefore,
-\pkg{dataMaid} was designed to be easily extendend with user-supplied
-functions for summarizing, viusalizing and checking data. In the package, we have
+\pkg{dataMaid} was designed to be easily extended with user-supplied
+functions for summarizing, visualizing and checking data. In the package, we have
 provided a vignette in which we describe how \pkg{dataMaid} extensions
@@
 on the dataset \code{toyData} available in \pkg{dataMaid}. The first
-two pages (excluding the frontpage) of this data report are shown in
+two pages (excluding the front page) of this data report are shown in
 Figure~\ref{fig:example1} and the following two pages are shown in
@@
 with a lot of potential errors to illustrate the main capabilities of
-\pkg{dataMaid}. Section~\ref{sec:bigExample} shows an example of a data sreening
+\pkg{dataMaid}. Section~\ref{sec:bigExample} shows an example of a data screening
 process with a real dataset. The following commands load the dataset
@@
 Table~\ref{table.cleanFormals} and they are grouped according to the
-part of the data assesment and report generation they influence. In
+part of the data assessment and report generation they influence. In
 order to understand this distinction, a glimpse of the inner structure
@@
     \node [cloud, below of=stop, node distance=3.5cm] (render) {Render
-      markdown and possiby open};
+      markdown and possibly open};
     % Draw edges
@@
  \quad \code{identifyOutliersTBStyle} & Identify outliers (Turkish Boxplot style) &  & & $\times$ & & $\times$ & $\times$ \\
- \quad \code{identifyWhitespace} & Identify prefixed and suffixed whitespace &  \blue{$\times$} & \blue{$\times$} & & \blue{$\times$} & & &  \\
+ \quad \code{identifyWhitespace} & Identify prefixed and suffixed white space &  \blue{$\times$} & \blue{$\times$} & & \blue{$\times$} & & &  \\
  \quad \code{isCPR} & Identify Danish CPR numbers & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &$\times$   \\
@@
 
-Now, if we only wanted to apply the function to identify whitespace
+Now, if we only wanted to apply the function to identify white space
 for factor variables, then we would need provide this information for \code{setChecks()}:
@@
 More specifically, we will create a report describing the \code{presidentData} dataset, which is available in \pkg{dataMaid} and use the information from this report to clean up the data.
-\code{presidentData} is a sligtly mutilated dataset with information about the 45 first US presidents, but with a few common data issues and a blind passenger.
+\code{presidentData} is a slightly mutilated dataset with information about the 45 first US presidents, but with a few common data issues and a blind passenger.
 The dataset contains one observation per president and has the following variables:
@@
 \item[\code{ageAtInauguration}] The age of the president at inauguration.
-\item[\code{favoriteNumber}] The favourite number of the president (fictional).
+\item[\code{favoriteNumber}] The favorite number of the president (fictional).
 \end{description}
@@
 On page 3, we see the contents for the three first variables.
-Here, we identifify a prefixed whitespace in the lastname entry for President Truman and we find that a dot was entered as a first name; this is a typical choice for coding missing values in e.g.
+Here, we identify a prefixed white space in the last name entry for President Truman and we find that a dot was entered as a first name; this is a typical choice for coding missing values in e.g.
 Stata, and therefore, it is flagged as a potential miscoded missing value.
@@
 A few of the identified problems have easy fixes that need no further discussion.
-We remove the prefixed whitespace from Truman's name, fix the misspelling of New York, convert the binary variable \code{assasinationAttempt} to a factor and change the class of the \code{ageAtInauguration} variable to numeric:
+We remove the prefixed white space from Truman's name, fix the misspelling of New York, convert the binary variable \code{assasinationAttempt} to a factor and change the class of the \code{ageAtInauguration} variable to numeric:
 
@@
 How much information is printed in the console can be adjusted by using the argument \code{quiet}.
-And convenintely introducing small alterations of the file names can be obtained by use of the \code{vol} argument.
+And conveniently introducing small alterations of the file names can be obtained by use of the \code{vol} argument.
 For instance, we can make a data report for \code{toyData} that is not opened automatically, produces no output to the console and includes the date and time of its creation in the file name:
@@
 \code{checkResult}s). Thus, the overall problem status of each variable
-can easily be unravelled using the list manipulation function
+can easily be unraveled using the list manipulation function
 \code{sapply()}:
@@
 
+\section*{Acknowledgements}
+\label{acknowledgements}
+This work was supported by The Lundbeck Foundation, Trygfonden and The Region of Southern Denmark.
+
 % \nocite{R}

### 2017-11-14 latex/dataMaid_toyData.Rmd
Prep for JSS submission

@@
 subtitle: "Autogenerated data summary from dataMaid"
-date: 2017-11-09 15:41:24
+date: 2017-11-14 13:41:55
 output: pdf_document
@@
 
- *  Report creation time: Thu Nov 09 2017 15:41:24
+ *  Report creation time: Tue Nov 14 2017 13:41:55
 
- *  dataMaid v0.9.7.9000 [Pkg: 2017-11-09 from local (ekstroem/dataMaid@NA)]
+ *  dataMaid v1.0.1 [Pkg: 2017-11-13 from local (ekstroem/dataMaid@NA)]
 
@@
 
- *  Platform: x86_64-apple-darwin15.6.0 (64-bit)(macOS Sierra 10.12.6).
+ *  Platform: x86_64-apple-darwin15.6.0 (64-bit)(macOS High Sierra 10.13.1).
 

### 2017-12-21 docs/articles/dataMaid_artData.Rmd
Added package site

---
dataMaid: yes
title: artData
subtitle: "Autogenerated data summary from dataMaid"
date: 2017-12-21 01:29:36
output: html_document
---
```{r , echo=FALSE, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
library(ggplot2)
library(pander)
```
```{r visualFunctions, echo=FALSE, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
ggAggHist <- getFromNamespace("ggAggHist", "dataMaid")
ggAggBarplot <- getFromNamespace("ggAggBarplot", "dataMaid")
```
# Data report overview
The dataset examined has the following dimensions:

---------------------------------
Feature                    Result
------------------------ --------
Number of observations        200

Number of variables            11
---------------------------------

### Checks performed
The following variable checks were performed, depending on the data type of each variable:

------------------------------------------------------------------------------------------------------------------------------
&nbsp;                                                 character   factor    labelled   numeric   integer   logical    Date   
----------------------------------------------------- ----------- --------- ---------- --------- --------- --------- ---------
Identify miscoded missing values                        &times;    &times;   &times;    &times;   &times;

Identify prefixed and suffixed whitespace               &times;    &times;   &times;

Identify levels with < 6 obs.                           &times;    &times;   &times;

Identify case issues                                    &times;    &times;   &times;

Identify misclassified numeric or integer variables     &times;    &times;   &times;

Identify non-trailing colons                            &times;    &times;   &times;

Identify outliers                                                                       &times;   &times;             &times; 
------------------------------------------------------------------------------------------------------------------------------

Please note that all numerical values in the following have been rounded to 2 decimals.

# Summary table

---------------------------------------------------------------------------------------------
&nbsp;              Variable class     # unique values   Missing observations  Any problems? 
------------------- ---------------- ----------------- ---------------------- ---------------
[ArtistID]          character                      179                 0.00 %     &times;

[ArtistName]        character                      179                 0.00 %     &times;

[NoOfMiddlenames]   numeric                          4                 0.00 %     &times;

[Title]             character                      200                 0.00 %     &times;

[Year]              integer                        149                 0.00 %     &times;

[Location]          character                       98                 1.50 %     &times;

[Continent]         factor                           3                 0.00 %     &times;

[Width]             numeric                        164                 5.00 %     &times;

[Height]            numeric                        165                 5.00 %     &times;

[Media]             character                       28                 5.00 %     &times;

[Movement]          character                       86                 9.00 %     &times;    
---------------------------------------------------------------------------------------------

# Variable list
## ArtistID

* Warning: This variable seems to contain ID codes.

---

## ArtistName

<div class = "row">
<div class = "col-lg-8">

---------------------------------------------
Feature                                Result
------------------------- -------------------
Variable type                       character

Number of missing obs.                0 (0 %)

Number of unique values                   179

Mode                        "Diego Velazquez"

No. zeros                                   0
---------------------------------------------

</div>
<div class = "col-lg-4">
```{r Var-2-ArtistName, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:179, .Label = c("Adolph von Menzel", 
"Alberto Giacometti", "Albrecht Altdorfer", "Albrecht Durer", 
"Alexej von Jawlensky", "Andrea Mantegna", "Andrew Wyeth", "Annibale Carracci", 
"Anthony van Dyck", "Antoine Watteau", "Antonio Canaletto", "Arnold Bocklin", 
"Arshile Gorky", "August Macke", "Auguste Renoir", "Bartolome Esteban Murillo", 
"Benozzo Gozzoli", "Caravaggio", "Carl Spitzweg", "Caspar David Friedrich", 
"Claude Monet", "Correggio", "David Wilkie", "Diego Velazquez", 
"Dong Qichang", "Duccio di Buoninsegna", "Edgar Degas", "Edouard Manet", 
"Edvard Munch", "Edward Hopper", "Egon Schiele", "El Greco", 
"Elisabeth Vigee-Lebrun", "Emil Nolde", "Emilio Vedova", "Erich Heckel", 
"Ernst Ludwig Kirchner", "Eugene Delacroix", "Ferdinand Hodler", 
"Fernand Khnopff", "Fernand Leger", "Francis Bacon", "Francis Picabia", 
"Francisco Goya", "Francisco de Zurbaran", "Francois Boucher", 
"Franz Kline", "Franz Marc", "Franz von Lenbach", "Gabriele Munter", 
"Geertgen tot Sint Jans", "Georg Baselitz", "George Grosz", "George Stubbs", 
"Georges Braque", "Georges Seurat", "Georges de la Tour", "Georgia O'Keeffe", 
"Giacomo Balla", "Giorgione", "Giotto", "Giovanni Battista Tiepolo", 
"Giovanni Bellini", "Giulio Romano", "Giuseppe Arcimboldo ", 
"Grant Wood", "Gustave Caillebotte", "Gustave Courbet", "Hannah Hoch", 
"Hans Baldung Grien", "Hans Holbein the Younger", "Hans Memling", 
"Hans von Marees", "Helen Frankenthaler", "Hendrick Avercamp", 
"Henri Matisse", "Henri Rousseau", "Henry Fuseli", "Henry Raeburn", 
"Hieronymus Bosch", "Honore Daumier", "Hugo van der Goes", "Hyacinthe Rigaud", 
"Ilya Repin", "Jackson Pollock", "Jacob Izaaksoon van Ruisdael", 
"Jacopo Tintoretto", "Jacques-Louis David", "James Ensor", "James McNeill Whistler", 
"Jan Matsys", "Jan Vermeer", "Jan van Eyck", "Jasper Johns, Jr.", 
"Jean Auguste Dominique Ingres", "Jean Metzinger", "Jean Simeon Chardin", 
"Joachim Patinir", "Joan Miro", "Johannes Itten", "John Constable", 
"John Everett Millais", "Josef Albers", "Joseph Wright of Derby", 
"Juan Gris", "Kangra-School", "Karl Friedrich Schinkel", "Kazimir Malevich", 
"Konrad Klapheck", "Kurt Schwitters", "Laszlo Moholy-Nagy", "Leonardo da Vinci", 
"Lorenzo Lotto", "Louis or Antoine Le Nain", "Lovis Corinth", 
"Luca Signorelli", "Lucas Cranach the Elder", "Lucio Fontana", 
"Lyonel Feininger", "Marc Chagall", "Marcel Duchamp", "Mark Rothko", 
"Mary Cassatt", "Masaccio", "Matthias Grunewald", "Max Beckmann", 
"Max Ernst", "Mikhail Nesterov", "Morris Louis", "Nicolas Poussin", 
"Niklaus Manuel", "Oskar Kokoschka", "Oskar Schlemmer", "Otto Dix", 
"Pablo Picasso", "Paolo Uccello", "Paolo Veronese", "Paul Cezanne", 
"Paul Gauguin", "Paul Klee", "Paul Nash", "Peter Paul Rubens", 
"Philipp Otto Runge", "Piero della Francesca", "Piet Mondrian", 
"Pieter Bruegel the Elder", "Pieter Saenredam", "Raphael", "Reginald Marsh", 
"Rembrandt", "Rene Magritte", "Richard Lindner", "Robert Delaunay", 
"Rogier van der Weyden", "Roy Lichtenstein", "Salvador Dali", 
"Sandro Botticelli", "Sebastiano Ricci", "Sonia Delaunay-Terk", 
"Stanley Spencer", "Stefan Lochner", "Tawaraya S?tatsu", "Theodore Gericault", 
"Thomas Eakins", "Thomas Gainsborough", "Titian", "Umberto Boccioni", 
"Victor Pasmore", "Vincent van Gogh", "Vittore Carpaccio", "Wassily Kandinsky", 
"Wilhelm Leibl", "Wilhelm von Kaulbach", "Wilhelm von Kobell", 
"William Holman Hunt", "William McTaggart", "William Turner", 
"Winslow Homer", "Wolf Vostell"), class = "factor"), y = c(1L, 
1L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
2L, 1L, 1L, 1L, 1L, 1L, 3L, 1L, 1L, 1L, 1L, 2L, 1L, 1L, 2L, 1L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
2L, 1L, 1L, 1L, 1L, 2L, 1L, 1L, 2L, 1L, 1L, 1L, 2L, 1L, 2L, 1L, 
3L, 1L, 1L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
1L, 1L, 1L, 1L, 2L, 1L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 
1L, 1L)), .Names = c("x", "y"), row.names = c(NA, -179L), class = "data.frame"), 
    vnam = "ArtistName")
```
</div>
</div>

- The following values appear with prefixed or suffixed white space: \"Giuseppe Arcimboldo \".

- Note that the following levels have at most five observations: \"Adolph von Menzel\", \"Alberto Giacometti\", \"Albrecht Altdorfer\", \"Albrecht Durer\", \"Alexej von Jawlensky\", \"Andrea Mantegna\", \"Andrew Wyeth\", \"Annibale Carracci\", \"Anthony van Dyck\", \"Antoine Watteau\" (169 additional values omitted).

---

## NoOfMiddlenames

* Note that this variable is treated as a factor variable below, as it only takes a few unique values.

<div class = "row">
<div class = "col-lg-8">

-----------------------------------
Feature                      Result
------------------------- ---------
Variable type               numeric

Number of missing obs.      0 (0 %)

Number of unique values           4

Mode                            "0"

No. zeros                       157
-----------------------------------

</div>
<div class = "col-lg-4">
```{r Var-3-NoOfMiddlenames, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
mosaicplot(structure(c(157L, 29L, 13L, 1L), .Dim = 4L, .Dimnames = structure(list(
    v = c("0", "1", "2", "3")), .Names = "v"), class = "table"), 
    main = "NoOfMiddlenames", xlab = "")
```
</div>
</div>

- Note that the following levels have at most five observations: \"3\".

---

## Title

* The variable is a key (distinct values for each observation).

---

## Year

<div class = "row">
<div class = "col-lg-8">

-----------------------------------------
Feature                            Result
------------------------- ---------------
Variable type                     integer

Number of missing obs.            0 (0 %)

Number of unique values               149

Median                             1851.5

1st and 3rd quartiles       1627.75; 1914

Min. and max.                  1150; 1968

No. zeros                               0

Mean                              1765.73
-----------------------------------------

</div>
<div class = "col-lg-4">
```{r Var-5-Year, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
prettierHistHelper(v = c(1886L, 1914L, 1531L, 1470L, 1930L, 1914L, 
1768L, 1912L, 1894L, 1914L, 1939L, 1950L, 1524L, 1576L, 1931L, 
1515L, 1917L, 1720L, 1961L, 1926L, 1907L, 1915L, 1958L, 1888L, 
1911L, 1948L, 1150L, 1957L, 1515L, 1919L, 1517L, 1559L, 1938L, 
1928L, 1914L, 1566L, 1913L, 1955L, 1936L, 1559L, 1965L, 1937L, 
1965L, 1911L, 1951L, 1910L, 1637L, 1930L, 1490L, 1843L, 1887L, 
1904L, 1550L, 1656L, 1530L, 1922L, 1881L, 1506L, 1448L, 1894L, 
1751L, 1871L, 1815L, 1934L, 1638L, 1920L, 1496L, 1968L, 1917L, 
1952L, 1749L, 1942L, 1877L, 1863L, 1944L, 1852L, 1877L, 1645L, 
1717L, 1492L, 1505L, 1955L, 1701L, 1635L, 1520L, 1780L, 1913L, 
1958L, 1891L, 1460L, 1732L, 1595L, 1911L, 1460L, 1823L, 1630L, 
1661L, 1786L, 1914L, 1889L, 1899L, 1500L, 1630L, 1508L, 1950L, 
1880L, 1479L, 1912L, 1633L, 1598L, 1942L, 1635L, 1465L, 1670L, 
1903L, 1528L, 1450L, 1900L, 1529L, 1456L, 1487L, 1914L, 1586L, 
1936L, 1835L, 1896L, 1637L, 1891L, 1459L, 1846L, 1640L, 1954L, 
1916L, 1911L, 1852L, 1621L, 1893L, 1500L, 1885L, 1965L, 1762L, 
1914L, 1310L, 1851L, 1806L, 1858L, 1565L, 1666L, 1660L, 1789L, 
1596L, 1524L, 1434L, 1824L, 1953L, 1306L, 1800L, 1824L, 1839L, 
1460L, 1819L, 1665L, 1927L, 1893L, 1808L, 1784L, 1799L, 1533L, 
1897L, 1890L, 1634L, 1898L, 1425L, 1562L, 1862L, 1510L, 1965L, 
1563L, 1736L, 1944L, 1882L, 1912L, 1793L, 1936L, 1922L, 1920L, 
1915L, 1660L, 1610L, 1911L, 1910L, 1745L, 1630L, 1852L, 1804L, 
1630L, 1912L, 1886L, 1867L, 1903L), vnam = "Year")
```
</div>
</div>

- Note that the following possible outlier values were detected: \"1958\", \"1961\", \"1965\", \"1968\".

---

## Location

<div class = "row">
<div class = "col-lg-8">

----------------------------------------------
Feature                                 Result
------------------------- --------------------
Variable type                        character

Number of missing obs.               3 (1.5 %)

Number of unique values                     97

Mode                        "National Gallery"

No. zeros                                    0
----------------------------------------------

</div>
<div class = "col-lg-4">
```{r Var-6-Location, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:97, .Label = c("Alfred Stieglitz Collection, Carl Van Vechten Gallery", 
"Alte Nationalgalerie", "Alte Pinakothek", "Art Institute of Chicago", 
"Brancacci Chapel", "Centre Pompidou", "Christie's Auction House", 
"Cleveland Museum of Art", "Courtauld Institute of Art, University of London", 
"Detroit Institute of Arts", "Fogg Museum", "Forberg Collection, Albertina", 
"Freer Gallery of Art", "Galleria dell'Accademia Carrara", "Gallerie dell'Accademia", 
"Gemaeldegalerie Alte Meister", "Gemaldegalerie", "Gemaldegalerie Alte Meister", 
"Gemeentemuseum Den Haag", "Guggenheim Museum", "Hamburger Kunsthalle", 
"Harvard Art Museum", "Hermitage Museum", "Hill-Stead Museum", 
"Kroller-Muller Museum", "Kunsthalle (Museum of Art)", "Kunsthaus Zurich", 
"Kunsthistorisches Museum", "Kunstmuseum Basel", "Kunstsammlung Nordrhein-Westfalen", 
"Kunthistorisches Museum", "Lenbach House State Gallery", "Leopold Hoesch Museum", 
"Louvre Museum", "Manchester Art Gallery", "Mauritshuis", "Medici Riccardi Palace", 
"Memlingmuseum, Sint-Janshospitaal", "Menard Art Museum", "Metropolitan Museum of Art", 
"Mr. and Mrs. Leigh B. Block", "Musee National d'Art Moderne", 
"Musee Royaux des Beaux-Arts", "Musee d'Orsay", "Musee d'Unterlinden", 
"Musee des Beaux-Arts de Nantes", "Museo Civico", "Museo Reina Sofia", 
"Museo del Novecento", "Museo del Prado", "Museum Folkwant", 
"Museum Gemaldegalerie der Staatlichen Museen", "Museum Ludwig", 
"Museum of Contemporary Art", "Museum of Fine Arts", "Museum of Modern Art", 
"National Gallery", "National Gallery of Art", "National Gallery of Scotland", 
"National Gallery, Galleria degli Uffizi, Louvre Museum", "Nationalgalerie, Staatliche Museen", 
"Nationalgalerie, Staatliche Museen zu Berlin", "Neue Pinakothe", 
"New Pinakothek", "Norton Simon Museum", "Osterreichische Galerie Belvedere", 
"Palazzo Ducale", "Peggy Guggenheim Collection", "Pennsylvania Academy of the Fine Arts", 
"Philadelphia Museum of Art", "Phillips Collections", "Pompidou Centre", 
"Private Collection", "Rijksmuseum", "Royal Collection, Windsor Castle", 
"Royal Museums of Fine Arts of Beligum", "Samuel H. Kress Collection, National Gallery of Art", 
"San Francisco Museum of Modern Art", "Santo Tome", "Scrovegni Chapel", 
"Staatliche Museen", "Stadtische Galerie im Lenbachhaus", "State Russian Museum", 
"Stedelijk Museum", "Szepmuveszeti Muzeum", "Tate Britain", "The Frick Collection", 
"Toledo Museum of Art", "Trustees of the Grosvenor Estate", "Uffizi Gallery", 
"Victoria & Albert Museum", "Villa Coredellina", "Wallraf-Richartz Museum", 
"Walters Art Museum", "Whitney Museum of American Art", "Wildenstein Collection", 
"von Zezschwitz (for sale)"), class = "factor"), y = c(1L, 1L, 
2L, 8L, 1L, 1L, 1L, 2L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 1L, 4L, 2L, 
1L, 1L, 4L, 1L, 1L, 1L, 1L, 2L, 3L, 4L, 6L, 3L, 4L, 2L, 1L, 12L, 
1L, 1L, 1L, 1L, 1L, 4L, 1L, 2L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 7L, 
1L, 1L, 2L, 1L, 1L, 8L, 13L, 6L, 3L, 1L, 1L, 1L, 4L, 1L, 1L, 
1L, 1L, 6L, 1L, 1L, 1L, 1L, 2L, 2L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 
1L, 2L, 1L, 1L, 4L, 1L, 1L, 1L, 1L, 1L, 1L, 3L, 1L, 1L, 1L, 1L
)), .Names = c("x", "y"), row.names = c(NA, -97L), class = "data.frame"), 
    vnam = "Location")
```
</div>
</div>

- Note that the following levels have at most five observations: \"Alfred Stieglitz Collection, Carl Van Vechten Gallery\", \"Alte Nationalgalerie\", \"Alte Pinakothek\", \"Brancacci Chapel\", \"Centre Pompidou\", \"Christie's Auction House\", \"Cleveland Museum of Art\", \"Courtauld Institute of Art, University of London\", \"Detroit Institute of Arts\", \"Fogg Museum\" (79 additional values omitted).

---

## Continent

<div class = "row">
<div class = "col-lg-8">

------------------------------------
Feature                       Result
------------------------- ----------
Variable type                 factor

Number of missing obs.       0 (0 %)

Number of unique values            3

Mode                        "Europe"

No. zeros                          0
------------------------------------

</div>
<div class = "col-lg-4">
```{r Var-7-Continent, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
mosaicplot(structure(c(2L, 143L, 55L), .Dim = 3L, .Dimnames = structure(list(
    v = c("Asia", "Europe", "North America")), .Names = "v"), class = "table"), 
    main = "Continent", xlab = "")
```
</div>
</div>

- Note that the following levels have at most five observations: \"Asia\".

---

## Width

<div class = "row">
<div class = "col-lg-8">

-------------------------------------
Feature                        Result
------------------------- -----------
Variable type                 numeric

Number of missing obs.       10 (5 %)

Number of unique values           163

Median                         122.45

1st and 3rd quartiles       77; 198.3

Min. and max.               10.7; 990

No. zeros                           0

Mean                           168.46
-------------------------------------

</div>
<div class = "col-lg-4">
```{r Var-8-Width, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
prettierHistHelper(v = c(308, 76.52, 35.5, 242, 62.4, 57, 244, 
107.5, 141, 97.1, 96.3, 207, 193, 167, NA, 77, 390, 199, 439.4, 
71.1, 88.9, 73, 279.4, 65.5, 128.6, 121.3, 150.8, 68, 589, 144, 
14.7, 204.5, 101.6, 73, 250, 51, 37, 153.87, 248.92, 112.9, 121.9, 
776, 61, 151.4, 170.4, 99.5, NA, 76.2, 28, 93, 92, 129.2, 112, 
276, 191, 100, 172.7, 88, 40, 87, 73, 117.5, NA, 13.9, 87.5, 
79, 389, 120, 160, 297.8, 119.04, 152.4, 46.6, 190, 119.3, 111.8, 
276, 159.5, 194, 32, 26, 60, 194, 209, 161, 17.2, 10.7, 297.8, 
358, 200, 126, 148.1, 73, 144, 111.8, 257, 77, 84, 73, 45, 80, 
48.7, 96, 175, 100.3, 130, 331.5, 100, 107, 196.2, 292.1, 214, 
127, 110, NA, 64, 116, 250.8, 120.3, 957, 278.5, 220, 360, 27, 
123, 150, 229.2, 66.1, 93, 705, 81, 131.2, NA, 95, 205, 123.5, 
174, 389, 190, 300, NA, 50, 45, 109.5, 141, 325, 162, 166.5, 
180, 422, 126.5, 85.4, 62, 354, 182.9, 185, 190, 96.7, 44.6, 
NA, 716, 179.39, 548.6, 73.5, 305, 25, 15.2, 81.5, 200.7, 183, 
367, 178, 597, 162, 108, 112, NA, 990, 61, 222, 77, 111.5, 135, 
101.6, 48.1, 60, 97.8, 117.5, 108.6, NA, 50.17, 35, 369.9, 130, 
61, 46.4, 129.9, 70, 205, NA), vnam = "Width")
```
</div>
</div>

- Note that the following possible outlier values were detected: \"10.7\", \"13.9\", \"14.7\", \"15.2\", \"17.2\", \"25\", \"26\", \"27\", \"28\", \"32\" (3 additional values omitted).

---

## Height

<div class = "row">
<div class = "col-lg-8">

-----------------------------------------
Feature                            Result
------------------------- ---------------
Variable type                     numeric

Number of missing obs.           10 (5 %)

Number of unique values               164

Median                                113

1st and 3rd quartiles       73.12; 168.75

Min. and max.                   12.3; 666

No. zeros                               0

Mean                               134.51
-----------------------------------------

</div>
<div class = "col-lg-4">
```{r Var-9-Height, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
prettierHistHelper(v = c(207.6, 54.61, 50.4, 147, 74.3, 88, 183, 
78.5, 120.5, 130.4, 129.6, 105, 175, 146, NA, 59.5, 260, 118.5, 
262.3, 43.8, 116.2, 92, 195.6, 81, 160.7, 81.9, 37.5, 48.3, 336, 
90, 30.3, 187, 67.9, 91.8, 250, 67, 48.5, 107.32, 198.12, 113.2, 
121.9, 349, 61, 192.1, 130.5, 107, NA, 101.6, 42, 62, 67, 169.5, 
73.5, 318, 152, 131, 129.9, 113, 51, 66, 59, 82.6, NA, 17.1, 
63, 104.5, 365, 200, 150, 220, 69.08, 84.1, 60.3, 130.5, 94.2, 
76.2, 212.2, 113.3, 129, 50, 35, 73, 277, 173, 151.5, 12.3, 18.1, 
270.8, 203, 225, 77, 88.3, 100, 77, 87.6, 148, 91, 105, 57, 51, 
120, 67.1, 71, 108.5, 81.3, 66, 176, 86, 60, 141, 199.39, 154, 
81, 130, NA, 76.8, 168, 210.5, 158.4, 182, 172.5, 181, 460, 35, 
92.7, 50, 131.2, 100.3, 67, 585, 93, 195.4, NA, 70, 142, 101.9, 
96.5, 220, 145, 250, NA, 60, 43, 76.4, 131, 207, 117, 121.5, 
140, 323, 100, 125.7, 66, 419, 152.4, 200, 97, 126.9, 36.2, NA, 
491, 136.21, 274.3, 91, 202, 30, 21.8, 115, 129.5, 122, 307, 
195, 247, 117, 108, 168, NA, 666, 66, 94, 113, 111, 169, 76.2, 
63.8, 81, 125.7, 98.5, 121.3, NA, 60.96, 64, 166, 163, 74, 29.4, 
193, 70, 255, NA), vnam = "Height")
```
</div>
</div>

- Note that the following possible outlier values were detected: \"460\", \"491\", \"585\", \"666\".

---

## Media

<div class = "row">
<div class = "col-lg-8">

---------------------------------------
Feature                          Result
------------------------- -------------
Variable type                 character

Number of missing obs.         10 (5 %)

Number of unique values              27

Mode                        "oil paint"

No. zeros                             0
---------------------------------------

</div>
<div class = "col-lg-4">
```{r Var-10-Media, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:27, .Label = c("acrylic paint", 
"aquatint, etching", "collage", "collage, oil paint", "egg tempera", 
"enamel", "encaustic, oil paint, and collage on fabric", "fresco", 
"gouache and graphite", "gouache-painted paper", "ink and color", 
"ink, color, gold and silver", "oil and metallic paint", "oil paint", 
"oil paint and charcoal", "oil paint and collage", "oil paint and magna", 
"oil paint and pastel", "oil paint, tempera, pastel", "oil piant", 
"pastel", "screen photography, transparent paint, silkscreen", 
"tempera", "varnished tempera", "watercolor", "watercolor, pen and ink", 
"wood cut"), class = "factor"), y = c(2L, 1L, 1L, 1L, 2L, 1L, 
1L, 4L, 1L, 1L, 1L, 1L, 1L, 147L, 2L, 1L, 1L, 1L, 1L, 2L, 1L, 
1L, 9L, 1L, 2L, 1L, 2L)), .Names = c("x", "y"), row.names = c(NA, 
-27L), class = "data.frame"), vnam = "Media")
```
</div>
</div>

- Note that the following levels have at most five observations: \"acrylic paint\", \"aquatint, etching\", \"collage\", \"collage, oil paint\", \"egg tempera\", \"enamel\", \"encaustic, oil paint, and collage on fabric\", \"fresco\", \"gouache and graphite\", \"gouache-painted paper\" (15 additional values omitted).

---

## Movement

<div class = "row">
<div class = "col-lg-8">

-------------------------------------
Feature                        Result
------------------------- -----------
Variable type               character

Number of missing obs.       18 (9 %)

Number of unique values            85

Mode                        "Baroque"

No. zeros                           0
-------------------------------------

</div>
<div class = "col-lg-4">
```{r Var-11-Movement, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:85, .Label = c("Abstract Artist", 
"Abstract Expressionist", "Abstract Expressionist:Color Field", 
"Abstract Expressionist:Neo-Dada:Pop Art", "American Modernist", 
"American Realist", "Analytic Cubist", "Baroque", "Bauhaus", 
"British Expressionist", "Classicist:Baroque", "Constructivist", 
"Cubism", "Cubist", "Dada", "Dada:New Objectivity", "Dada:Surrealist", 
"De Stijl", "Dutch Golden Age", "Dutch Golden Age:Baroque", "Dutch and Flemish Renaissance", 
"Early Netherlandish", "Early Netherlandish Renaissance", "Early Renaissance", 
"English Romantic", "Expressionist", "Expressionist:Abstract Artist", 
"Expressionist:Bauhaus:Surrealist", "Expressionist:Surrealist", 
"Fauvist", "Flemish", "French Baroque", "French Impressionist", 
"Futurism", "Geometric abstraction", "German Expressionist", 
"German Realist", "German Renaissance", "German Romantic", "High Renaissance", 
"High Renaissance:Mannerist", "Impressionist", "Impressionist:Expressionist", 
"Italian Modernist", "Italian Renaissance", "Late Gothic", "Mannerist", 
"Mechanistic Cubist", "Merz", "Neo-Expressionist", "Neo-Impresionist:Divisionist:Fauvist:Cubist", 
"Neoclassicist", "Neue Sachlichkeit", "Northern Renaissance", 
"Orphist", "Orphist:Cubist:Expressionist", "Pop Artist", "Post-Impressionist", 
"Post-Impressionist:Neo-Impressionist:Pointillist", "Pre-Raphaelite", 
"Realism", "Realist", "Realist:Impressionist", "Realist:Surrealist:Pop Artist", 
"Regionalist", "Renaissance", "Rimpa school", "Rococo", "Rococo:Neoclassicist", 
"Romantic", "Russian Expressionist", "Russian Symbolist", "Scottish landscape painter", 
"Sienese school:Gothic style", "Social Realism", "Spatialist", 
"Suprematist", "Surrealism", "Surrealism:Dada", "Surrealist", 
"Surrealist:Expressionist", "Surrealist:Expressionist:Cubist:Formalist", 
"Symbolist", "Tonalist", "Video Artist:Installation:Happening:Fluxus"
), class = "factor"), y = c(1L, 5L, 1L, 1L, 1L, 1L, 1L, 14L, 
1L, 1L, 2L, 1L, 4L, 2L, 1L, 1L, 3L, 1L, 1L, 4L, 3L, 1L, 2L, 2L, 
1L, 11L, 1L, 1L, 1L, 1L, 1L, 2L, 1L, 2L, 1L, 1L, 3L, 3L, 3L, 
6L, 1L, 4L, 1L, 1L, 8L, 2L, 3L, 1L, 1L, 1L, 1L, 2L, 1L, 7L, 1L, 
1L, 1L, 6L, 1L, 2L, 1L, 5L, 1L, 1L, 1L, 2L, 1L, 2L, 1L, 9L, 1L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 1L, 1L, 3L, 1L, 1L)), .Names = c("x", 
"y"), row.names = c(NA, -85L), class = "data.frame"), vnam = "Movement")
```
</div>
</div>

- Note that the following levels have at most five observations: \"Abstract Artist\", \"Abstract Expressionist\", \"Abstract Expressionist:Color Field\", \"Abstract Expressionist:Neo-Dada:Pop Art\", \"American Modernist\", \"American Realist\", \"Analytic Cubist\", \"Bauhaus\", \"British Expressionist\", \"Classicist:Baroque\" (68 additional values omitted).

- Note that the following values include colons: \"Abstract Expressionist:Color Field\", \"Abstract Expressionist:Neo-Dada:Pop Art\", \"Classicist:Baroque\", \"Dada:New Objectivity\", \"Dada:Surrealist\", \"Dutch Golden Age:Baroque\", \"Expressionist:Abstract Artist\", \"Expressionist:Bauhaus:Surrealist\", \"Expressionist:Surrealist\", \"High Renaissance:Mannerist\" (12 additional values omitted).

---

Report generation information:

 *  Created by Claus Thorn Ekstrøm.

 *  Report creation time: Thu Dec 21 2017 01:29:36

 *  dataMaid v1.0.1 [Pkg: 2017-11-13 from local (ekstroem/dataMaid@NA)]

 *  R version 3.4.3 (2017-11-30).

 *  Platform: x86_64-apple-darwin15.6.0 (64-bit)(macOS High Sierra 10.13.1).

 *  Function call: `makeDataReport(data = artData, output = "html", preChecks = c("isKey", 
    "isSingular", "isSupported", "isID"), replace = TRUE, openResult = FALSE, 
    summaries = setSummaries(character = defaultCharacterSummaries(add = "countZeros"), 
        factor = defaultFactorSummaries(add = "countZeros"), 
        labelled = defaultLabelledSummaries(add = "countZeros"), 
        numeric = defaultNumericSummaries(add = c("countZeros", 
            "meanSummary")), integer = defaultIntegerSummaries(add = c("countZeros", 
            "meanSummary")), logical = defaultLogicalSummaries(add = c("meanSummary"))), 
    visuals = setVisuals(factor = "mosaicVisual", numeric = "prettierHist", 
        integer = "prettierHist", Date = "prettierHist"), checks = setChecks(character = defaultCharacterChecks(add = "identifyColons"), 
        factor = defaultFactorChecks(add = "identifyColons"), 
        labelled = defaultLabelledChecks(add = "identifyColons")))`

### 2018-01-16 docs/articles/dataMaid_artData.Rmd
Added initial prototype for makeCodebook

@@
 subtitle: "Autogenerated data summary from dataMaid"
-date: 2017-12-21 01:29:36
+date: 2018-01-12 19:32:44
 output: html_document
@@
 
- *  Report creation time: Thu Dec 21 2017 01:29:36
+ *  Report creation time: Fri Jan 12 2018 19:32:44
 
@@
 
- *  Platform: x86_64-apple-darwin15.6.0 (64-bit)(macOS High Sierra 10.13.1).
+ *  Platform: x86_64-apple-darwin15.6.0 (64-bit)(macOS High Sierra 10.13.2).
 

### 2018-01-16 R/clean.R

@@ -1161,6 +1161,3 @@ doCheckLabs <- function(v) {
   return(v)
 }
 
-
-
-

### 2018-01-16 R/makeDataReport.R

@@ -491,7 +491,7 @@ makeDataReport <- function(data, output=NULL, render=TRUE,
   doSummarize <- "summarize" %in% mode
   
   if (!doCheck & !doVisualize & !doSummarize & !silent) {
-    warning("Note that no proper arguments were supplied to \"mode\" - no data report genereation performed")
+    warning("Note that no proper arguments were supplied to \"mode\" - no data report generation performed")
   } #rewrite warning message
   
   ## Disregard the twocolumn option if we're only asking for one of visualize and summarize
@@ -1143,3 +1143,30 @@ doCheckLabs <- function(v) {
 
 
 
+#' Produce a data codebook
+#'
+#' Make a data codebook that summarizes the contents of a dataset.
+#' The result is saved to an R markdown file which can
+#' rendered into an easy-to-read codebook in pdf, html or word formats. 
+#'
+#' @param data The dataset to be checked. This dataset should be of class \code{data.frame},
+#' \code{tibble} or \code{matrix}. If it is of classs \code{matrix}, it will be converted to a
+#' \code{data.frame}.
+#' 
+#' @param vol Extra text string or numeric that is appended on the end of the output
+#' file name(s). For example, if the dataset is called "myData", no file argument is
+#'  supplied and \code{vol=2}, the output file will be called "dataMaid_myData2.Rmd"
+#' 
+#' @param reportTitle A text string. If supplied, this will be the printed title of the
+#' report. If left unspecified, the title with the name of the supplied dataset.
+#' 
+#' @export
+makeCodebook <- function(data, vol="", reportTitle=NULL, ...) {
+    
+    dfname <- deparse(substitute(data))
+
+    makeDataReport(data,
+                   reportTitle=paste0("Codebook for ", dfname),
+                   file=normalizeFileName(paste0("codebook_", dfname, vol, ".Rmd")),
+                   ...)
+}

### 2018-02-07 docs/articles/dataMaid_artData.Rmd
Updated webpage + added "Courses" tab

@@
 subtitle: "Autogenerated data summary from dataMaid"
-date: 2018-01-12 19:32:44
+date: 2018-02-07 18:03:36
 output: html_document
@@
 ----------------------------------------------------- ----------- --------- ---------- --------- --------- --------- ---------
-Identify miscoded missing values                        &times;    &times;   &times;    &times;   &times;
+Identify miscoded missing values                        &times;    &times;   &times;    &times;   &times;             &times;
 
@@
 "Fernand Khnopff", "Fernand Leger", "Francis Bacon", "Francis Picabia", 
-"Francisco Goya", "Francisco de Zurbaran", "Francois Boucher", 
+"Francisco de Zurbaran", "Francisco Goya", "Francois Boucher", 
 "Franz Kline", "Franz Marc", "Franz von Lenbach", "Gabriele Munter", 
 "Geertgen tot Sint Jans", "Georg Baselitz", "George Grosz", "George Stubbs", 
-"Georges Braque", "Georges Seurat", "Georges de la Tour", "Georgia O'Keeffe", 
+"Georges Braque", "Georges de la Tour", "Georges Seurat", "Georgia O'Keeffe", 
 "Giacomo Balla", "Giorgione", "Giotto", "Giovanni Battista Tiepolo", 
@@
 "Jacopo Tintoretto", "Jacques-Louis David", "James Ensor", "James McNeill Whistler", 
-"Jan Matsys", "Jan Vermeer", "Jan van Eyck", "Jasper Johns, Jr.", 
+"Jan Matsys", "Jan van Eyck", "Jan Vermeer", "Jasper Johns, Jr.", 
 "Jean Auguste Dominique Ingres", "Jean Metzinger", "Jean Simeon Chardin", 
@@
 2L, 1L, 1L, 1L, 1L, 1L, 3L, 1L, 1L, 1L, 1L, 2L, 1L, 1L, 2L, 1L, 
+1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 1L, 1L, 1L, 1L, 
+1L, 1L, 1L, 1L, 1L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 
-1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
-1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 
-1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 
+1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 1L, 1L, 1L, 1L, 
 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
@@
 "Memlingmuseum, Sint-Janshospitaal", "Menard Art Museum", "Metropolitan Museum of Art", 
-"Mr. and Mrs. Leigh B. Block", "Musee National d'Art Moderne", 
-"Musee Royaux des Beaux-Arts", "Musee d'Orsay", "Musee d'Unterlinden", 
-"Musee des Beaux-Arts de Nantes", "Museo Civico", "Museo Reina Sofia", 
-"Museo del Novecento", "Museo del Prado", "Museum Folkwant", 
-"Museum Gemaldegalerie der Staatlichen Museen", "Museum Ludwig", 
-"Museum of Contemporary Art", "Museum of Fine Arts", "Museum of Modern Art", 
-"National Gallery", "National Gallery of Art", "National Gallery of Scotland", 
-"National Gallery, Galleria degli Uffizi, Louvre Museum", "Nationalgalerie, Staatliche Museen", 
-"Nationalgalerie, Staatliche Museen zu Berlin", "Neue Pinakothe", 
-"New Pinakothek", "Norton Simon Museum", "Osterreichische Galerie Belvedere", 
+"Mr. and Mrs. Leigh B. Block", "Musee d'Orsay", "Musee d'Unterlinden", 
+"Musee des Beaux-Arts de Nantes", "Musee National d'Art Moderne", 
+"Musee Royaux des Beaux-Arts", "Museo Civico", "Museo del Novecento", 
+"Museo del Prado", "Museo Reina Sofia", "Museum Folkwant", "Museum Gemaldegalerie der Staatlichen Museen", 
+"Museum Ludwig", "Museum of Contemporary Art", "Museum of Fine Arts", 
+"Museum of Modern Art", "National Gallery", "National Gallery of Art", 
+"National Gallery of Scotland", "National Gallery, Galleria degli Uffizi, Louvre Museum", 
+"Nationalgalerie, Staatliche Museen", "Nationalgalerie, Staatliche Museen zu Berlin", 
+"Neue Pinakothe", "New Pinakothek", "Norton Simon Museum", "Osterreichische Galerie Belvedere", 
 "Palazzo Ducale", "Peggy Guggenheim Collection", "Pennsylvania Academy of the Fine Arts", 
@@
 "San Francisco Museum of Modern Art", "Santo Tome", "Scrovegni Chapel", 
-"Staatliche Museen", "Stadtische Galerie im Lenbachhaus", "State Russian Museum", 
-"Stedelijk Museum", "Szepmuveszeti Muzeum", "Tate Britain", "The Frick Collection", 
-"Toledo Museum of Art", "Trustees of the Grosvenor Estate", "Uffizi Gallery", 
-"Victoria & Albert Museum", "Villa Coredellina", "Wallraf-Richartz Museum", 
-"Walters Art Museum", "Whitney Museum of American Art", "Wildenstein Collection", 
-"von Zezschwitz (for sale)"), class = "factor"), y = c(1L, 1L, 
-2L, 8L, 1L, 1L, 1L, 2L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 1L, 4L, 2L, 
-1L, 1L, 4L, 1L, 1L, 1L, 1L, 2L, 3L, 4L, 6L, 3L, 4L, 2L, 1L, 12L, 
-1L, 1L, 1L, 1L, 1L, 4L, 1L, 2L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 7L, 
-1L, 1L, 2L, 1L, 1L, 8L, 13L, 6L, 3L, 1L, 1L, 1L, 4L, 1L, 1L, 
-1L, 1L, 6L, 1L, 1L, 1L, 1L, 2L, 2L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 
-1L, 2L, 1L, 1L, 4L, 1L, 1L, 1L, 1L, 1L, 1L, 3L, 1L, 1L, 1L, 1L
-)), .Names = c("x", "y"), row.names = c(NA, -97L), class = "data.frame"), 
-    vnam = "Location")
+"Stadtische Galerie im Lenbachhaus", "State Russian Museum", 
+"Stedelijk Museum", "Staatliche Museen", "Szepmuveszeti Muzeum", 
+"Tate Britain", "The Frick Collection", "Toledo Museum of Art", 
+"Trustees of the Grosvenor Estate", "Uffizi Gallery", "Victoria & Albert Museum", 
+"Villa Coredellina", "von Zezschwitz (for sale)", "Wallraf-Richartz Museum", 
+"Walters Art Museum", "Whitney Museum of American Art", "Wildenstein Collection"
+), class = "factor"), y = c(1L, 1L, 2L, 8L, 1L, 1L, 1L, 2L, 1L, 
+2L, 1L, 1L, 1L, 1L, 1L, 1L, 4L, 2L, 1L, 1L, 4L, 1L, 1L, 1L, 1L, 
+2L, 3L, 4L, 6L, 3L, 4L, 2L, 1L, 12L, 1L, 1L, 1L, 1L, 1L, 4L, 
+1L, 2L, 1L, 1L, 2L, 1L, 1L, 1L, 7L, 1L, 1L, 1L, 2L, 1L, 1L, 8L, 
+13L, 6L, 3L, 1L, 1L, 1L, 4L, 1L, 1L, 1L, 1L, 6L, 1L, 1L, 1L, 
+1L, 2L, 2L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 1L, 2L, 1L, 4L, 1L, 
+1L, 1L, 1L, 1L, 1L, 1L, 3L, 1L, 1L, 1L)), .Names = c("x", "y"
+), row.names = c(NA, -97L), class = "data.frame"), vnam = "Location")
 ```
@@
 "enamel", "encaustic, oil paint, and collage on fabric", "fresco", 
-"gouache and graphite", "gouache-painted paper", "ink and color", 
+"gouache-painted paper", "gouache and graphite", "ink and color", 
 "ink, color, gold and silver", "oil and metallic paint", "oil paint", 
@@
 
-- Note that the following levels have at most five observations: \"acrylic paint\", \"aquatint, etching\", \"collage\", \"collage, oil paint\", \"egg tempera\", \"enamel\", \"encaustic, oil paint, and collage on fabric\", \"fresco\", \"gouache and graphite\", \"gouache-painted paper\" (15 additional values omitted).
+- Note that the following levels have at most five observations: \"acrylic paint\", \"aquatint, etching\", \"collage\", \"collage, oil paint\", \"egg tempera\", \"enamel\", \"encaustic, oil paint, and collage on fabric\", \"fresco\", \"gouache-painted paper\", \"gouache and graphite\" (15 additional values omitted).
 
@@
 "Cubism", "Cubist", "Dada", "Dada:New Objectivity", "Dada:Surrealist", 
-"De Stijl", "Dutch Golden Age", "Dutch Golden Age:Baroque", "Dutch and Flemish Renaissance", 
-"Early Netherlandish", "Early Netherlandish Renaissance", "Early Renaissance", 
-"English Romantic", "Expressionist", "Expressionist:Abstract Artist", 
+"De Stijl", "Dutch and Flemish Renaissance", "Dutch Golden Age", 
+"Dutch Golden Age:Baroque", "Early Netherlandish", "Early Netherlandish Renaissance", 
+"Early Renaissance", "English Romantic", "Expressionist", "Expressionist:Abstract Artist", 
 "Expressionist:Bauhaus:Surrealist", "Expressionist:Surrealist", 
@@
 ), class = "factor"), y = c(1L, 5L, 1L, 1L, 1L, 1L, 1L, 14L, 
-1L, 1L, 2L, 1L, 4L, 2L, 1L, 1L, 3L, 1L, 1L, 4L, 3L, 1L, 2L, 2L, 
+1L, 1L, 2L, 1L, 4L, 2L, 1L, 1L, 3L, 1L, 3L, 1L, 4L, 1L, 2L, 2L, 
 1L, 11L, 1L, 1L, 1L, 1L, 1L, 2L, 1L, 2L, 1L, 1L, 3L, 3L, 3L, 
@@
 
- *  Created by Claus Thorn Ekstrøm.
+ *  Created by Anne Helby Petersen.
 
- *  Report creation time: Fri Jan 12 2018 19:32:44
+ *  Report creation time: on feb 07 2018 18:04:14
 
- *  dataMaid v1.0.1 [Pkg: 2017-11-13 from local (ekstroem/dataMaid@NA)]
+ *  dataMaid v1.1.0 [Pkg: 2018-02-03 from local (ekstroem/dataMaid@NA)]
 
- *  R version 3.4.3 (2017-11-30).
+ *  R version 3.4.2 (2017-09-28).
 
- *  Platform: x86_64-apple-darwin15.6.0 (64-bit)(macOS High Sierra 10.13.2).
+ *  Platform: x86_64-w64-mingw32/x64 (64-bit)(Windows 7 x64 (build 7601) Service Pack 1).
 

### 2018-02-08 docs/articles/dataMaid_artData.Rmd
CSP2018 website

--- 2018-01-16 docs/articles/dataMaid_artData.Rmd
@@
 subtitle: "Autogenerated data summary from dataMaid"
-date: 2018-01-12 19:32:44
+date: 2018-02-08 11:55:28
 output: html_document
@@
 
- *  Report creation time: Fri Jan 12 2018 19:32:44
+ *  Report creation time: Thu Feb 08 2018 11:55:29
 
- *  dataMaid v1.0.1 [Pkg: 2017-11-13 from local (ekstroem/dataMaid@NA)]
+ *  dataMaid v1.1.0 [Pkg: 2018-02-06 from local (ekstroem/dataMaid@NA)]
 

### 2018-02-08 docs/articles/dataMaid_artData.Rmd
CSP 2018 website

@@
 subtitle: "Autogenerated data summary from dataMaid"
-date: 2018-02-08 11:55:28
+date: 2018-02-08 12:00:56
 output: html_document
@@
 
- *  Report creation time: Thu Feb 08 2018 11:55:29
+ *  Report creation time: Thu Feb 08 2018 12:00:56
 

### 2018-02-08 docs/articles/dataMaid_artData.Rmd
Added slides

@@
 subtitle: "Autogenerated data summary from dataMaid"
-date: 2018-02-08 12:00:56
+date: 2018-02-08 19:12:55
 output: html_document
@@
 
- *  Report creation time: Thu Feb 08 2018 12:00:56
+ *  Report creation time: Thu Feb 08 2018 19:12:55
 

### 2018-02-08 docs/CSP2018/slides/module1/module1.Rmd
Added slides

---
title: "Cleaning Up the Data Cleaning Process"
subtitle: "Module 1: Introduction and terminology"
author: "Claus Thorn Ekstrøm and Anne Helby Petersen<br>UCPH Biostatistics"
date: "CSP, Portland, OR, Feb. 15th, 2018<br>.footnotesize[[Slides/homepage](https://ekstroem.github.io/dataMaid/CSP2018/index.html)]"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ['default', 'ku.css', 'animate.min.css']
    seal: true
    nature:
      ratio: '16:9'
      slideNumberFormat: '%current%' 
      highlightStyle: ir-black
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "js/zoom.js"
---

```{r setup, include=FALSE}
library(knitr)
options(htmltools.dir.version = FALSE, cache=TRUE)
opts_chunk$set(comment = NA, prompt=TRUE)
#opts_chunk$set(dev.args=list(bg="transparent"), fig.width=15, fig.height=7)
source("kutheme.R")
```

background-image: url(pics/redcard+sociology.jpg)
background-size: 96%
class: center, middle

???

Science og statistics har været under pres de seneste år.

Specielt indenfor psykologi har det vist sig svært at replikere vigtige resultater

Resulteret i en diskussion om, hvad videnserfaring var, og hvordan forskellige grupper 
kan se så forskelligt på de samme data.

---

background-image: url(pics/soccer.png)
background-size: 98%
class: center, middle

???

Soccer la la la la

Rødt kort til farvede spillere

29 forskergrupper

---

# Classical research process

From idea ...

```{r echo=FALSE, messages=FALSE, warning=FALSE, out.width='100%', out.height='35%'}
library(DiagrammeR)
library(DiagrammeRsvg)
library(svglite)
library(rsvg)

#svg <- export_svg(
grViz("
digraph dot {

graph [layout = dot,
       rankdir = LR,
       bgcolor='#000000',
       size=2]

node [shape = circle,
      style = filled,
      fillcolor = DimGray,
      fontcolor = White,
      fontsize=15,
      fontname=Helvetica,
      label = '', 
      penwidth=4, 
      margin=0.05,
      color=White]
a [label='Design']
b [label='Collect']
c [label='Analyze']
d [label='Publish']

edge [color = White, penwidth=4]
a -> b -> c 
c -> d [color=red, penwidth=4]
}")
#)

#svg %>%
#  charToRaw %>% rsvg %>% png::writePNG('graph.png')
#knitr::include_graphics("graph.png")

# html_print(HTML(svg), background="transparent", viewer=NULL)

```

.pull-right[... to publication.]

???

The red line shows where peer review comes in. Total summary. We need to document the steps we did througout as shown in the soccer example.

We want reproducible research

---

# Why is reproducible research important?

> **Statistical analysis**
>
> .large[All of the data were analyzed with data processing software and figures with Microsoft excel 2007.]
>
> .pull-right[-- Tayefe *et al*, Advances in Bioresearch, 2014]

???

Full statistical analysis section from a scientific paper.

Clearly impossible to reproduce

---

# Terminology

**Reproducibility**

Given code/data/materials, can I get *the same* (=identical) numbers that you did?

**Replicability**

Given scientific protocol, can I get *the same* (=in agreement) result that you did in my own study?

???

However, what do we really do?

---

## The life of a data scientist

> .large[Data scientists, according to interviews and expert estimates, spend from **50 percent to 80 percent** of their time mired in this more mundane labor of collecting and preparing unruly digital data, before it can be explored for useful nuggets.]
>
> .right[.small[-- "For Big-Data Scientists, 'Janitor Work' Is Key Hurdle to Insight" - The New York Times, 2014]]

???

Noter

---

## Realistic process plan

```{r echo=FALSE, messages=FALSE, warning=FALSE, out.width='100%', out.height='25%'}
library(DiagrammeR)
library(DiagrammeRsvg)
library(svglite)
library(rsvg)

#svg <- export_svg(
grViz("
digraph dot {

graph [layout = dot,
       rankdir = LR,
       bgcolor='#000000',
       size=2]

node [shape = circle,
      style = filled,
      fillcolor = DimGray,
      fontcolor = White,
      fontsize=15,
      fontname=Helvetica,
      label = '', 
      penwidth=4, 
      margin=0.05,
      color=White]
a [label='Design']
b [label='Collect']
c [label='Analyze']
d [label='Publish']

edge [color = White, penwidth=4]
a -> b 
c -> d 
b -> c [color=red, penwidth=16]
}")
#)

#svg %>%
#  charToRaw %>% rsvg %>% png::writePNG('graph.png')
#knitr::include_graphics("graph.png")

# html_print(HTML(svg), background="transparent", viewer=NULL)

```

```{r dilbert1, echo=FALSE}
knitr::include_graphics("pics/cartoon-metadata.png")
```

???

GIGO

Huge impact **here**

*   Filtering, selection, error-fixing, ...

---

class: middle

# Plan for workshop

1.  Introduction, data, error-hunting
2.  Data screening with `dataMaid`. Extending `dataMaid`
3.  Validity and consistency checking using `validate`
4.  Data cleaning and reporting

Exercises.

If you haven't already: go to [www.biostatistics.dk/CSP2018/](www.biostatistics.dk/CSP2018/) and install the required packages.

---

background-image: url(pics/flower.png)
background-size: 60%
class: center, middle

---

background-image: url(pics/structure.png)
background-position: right
background-size: 30%

# From raw data to analysis-ready

.small[
.pull-left[
*   Wrangle to put into correct format and type (validity)
*   Screen to look for consistency, accuracy and uniqueness
*   Validate to check for consistency, accuracy and uniqueness
*   Clean data 
*   Check (screen/validate) again
]]

---

## What is data?

### **Context**

*   Data type : BMI
*   Values  : 23.5, 31.1, ...
*   Missing  : Half the individuals were not measured

### **Content**

--

*   Data type : numeric
*   Values : Non-negative. Idea about lower and upper boundary
*   Missing : Code as `.`

???

Note the complete overlap between context and content.

Crucial: someone must know the topic!

---

# Wide vs. long

```{r echo=FALSE,message=FALSE,warning=FALSE}
library(reshape2)
library(readr)
DF <- data.frame(id=c(1, 2), bmi0=c(35.2, 31.1), bmi52=c(24.2, 27.0))
DFm <- DF %>% melt(id.vars="id") %>% mutate(time=readr::parse_number(variable)) %>% select(-variable)
```

.pull-left[
```{r,results='markup',echo=FALSE}
knitr::kable(DF, format = 'markdown')
```
]

.pull-right[
```{r,results='markup',echo=FALSE}
knitr::kable(DFm, format = 'markdown')
```
]

--

*   Work row-wise or column-wise?
*   Tidy data?

???

Fordele og ulemper ved begge dele. Man skal være opmærksom på, hvad man har med at gøre.

---

class: center, middle

## First rule of reproducible data wrangling

.Large[*Thou shall never manually modify your raw data.*]

There are no exceptions to this rule.

---

class: center, middle

## Second rule of reproducible data wrangling

.Large[*Thou shall never overwrite your raw data.*]

There are no exceptions to this rule either.

---

# R markdown

Format for writing **reproducible**, dynamic reports with R. Embed R code and results into slideshows, pdfs, html documents, Word files and more. See [cheat sheet at RStudio](https://github.com/rstudio/cheatsheets/raw/master/rmarkdown-2.0.pdf).

```{r eval=FALSE}
install.packages("rmarkdown", "knitr")
```

```{r rmarkdown1, echo=FALSE}
knitr::include_graphics("pics/rmarkdown.png")
```

---

# Validity - data formats

Technically corrct data requires that the data formats are correct

.pull-left[
```{r}
DF
```
]

.pull-right[
```{r}
lapply(DF, class)
```
]

---

class: inverse, middle

# Exercise 1

Get the `bigPresidentData` from the `dataMaid` package:

```{r warning=FALSE, message=FALSE}
library(dataMaid)
data("bigPresidentData")
```

Hunt for errors!

### 2018-02-08 docs/CSP2018/slides/module2/module2.Rmd
Added slides

---
title: "Cleaning Up the Data Cleaning Process"
subtitle: "Module 2: Data Screening with dataMaid"
author: "Claus Thorn Ekstrøm and Anne Helby Petersen<br>UCPH Biostatistics"
date: "CSP, Portland, OR, Feb. 15th, 2018<br>.footnotesize[[Slides/homepage](https://ekstroem.github.io/dataMaid/CSP2018/index.html)]"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ['default', 'ku.css', 'animate.min.css']
    seal: true
    nature:
      ratio: '16:9'
      slideNumberFormat: '%current%' 
      highlightStyle: ir-black
      highlightLines: true
      countIncrementalSlides: false
---

background-image: url(pics/manbeer.jpg)
background-size: 100%
class: middle, center

```{r setup, include=FALSE}
library(knitr)
options(htmltools.dir.version = FALSE, cache=TRUE)
opts_chunk$set(comment = NA, prompt=TRUE)
#opts_chunk$set(dev.args=list(bg="transparent"), fig.width=15, fig.height=7)
source("kutheme.R")
```

# The RESCueH project

---

class: center

# Timeline followback (TLFB)

```{r, echo=FALSE}
DF <- data.frame(day1=sample(1:25, 6),
                 day2=c(NA, NA, sample(1:25, 3), 88),
                 day3=c(NA, 99, 40, sample(1:25, 3)))
head(DF, 4)
```

---

class: center

# Timeline followback (TLFB)

```{r, echo=FALSE}
head(DF, 6)
```

---

background-image: url(pics/mau.png)
background-size: 80%
class: center

# Monthly Alcohol units

```{r echo=FALSE,message=FALSE, warning=FALSE, error=FALSE}
library(extrafont)
loadfonts()
library(ggplot2)
library(ggthemr)
claus <- define_palette(swatch = c('#eeeeee', '#FFFACD','#901A1E', 'lightblue', 'green', 'blue', 'brown', 'purple', 'yellow'),
  gradient = c(lower = 'red', upper = 'green'),
                        background = "#222222",
                        text = c("#eeeeee", "#eeeeee"),
                        line = c("#ee1e1e", "#6e6e6e"), # Axis lines
                        gridline = "#434343") # Gridlines

ggthemr(claus, layout="scientific", type = 'outer', spacing=1.2)

tlfb <- data.frame(units=dget("tlfb.txt"))

p <- ggplot(tlfb, aes(units)) + geom_dotplot() +
#     ggtitle("Baseline alcohol") +
     xlab("Units") + ylab("Frequency") +
     theme(text=element_text(size=24, family="Comic Sans MS"),
           axis.text.y=element_blank(),
           axis.ticks.y=element_blank())

theme_xkcd <- theme(

#  panel.background = element_rect(fill="transparent"),
#  plot.background = element_rect(fill="transparent"),

  #axis.ticks = element_line(colour=NA),
##  panel.grid = element_line(colour="white"),
  #axis.text.y = element_text(colour=NA),
##  axis.text.x = element_text(colour="black"),
  text = element_text(size=36, family="Humor Sans")
)
## p + theme_xkcd

# Rotate the dot plot
# p + coord_flip()

#p <- ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point(size=3) +
#  ggtitle("Fuel Efficiency of 32 Cars") +
#  xlab("Weight (x1000 lb)") + ylab("Miles per Gallon") +
#  theme(text=element_text(size=16, family="Comic Sans MS"))
#p
```

---

# Reproducible research

What **didn't** we check?

--

- Need **experts in relevant field**
- Merge existing databases into megadatabases
- New technologies revive old data

---

class: center

```{r message=FALSE, echo=FALSE}
library(dataMaid)
data(toyData)
```

.small[
```{r message=FALSE, echo=FALSE}
toyData
```
]

---

class: middle

# dataMaid

```{r eval=FALSE}
library(dataMaid)
data(toyData)
makeDataReport(toyData) #<<
```

Documentation to be **read** and **evaluated** by a human.

See [github.com/ekstroem/dataMaid](github.com/ekstroem/dataMaid) for more info. Stable version on CRAN.

---

background-image: url(pics/flowchart2.png)
background-size: 100%
class: center

# `dataMaid` flowchart

```{r eval=TRUE,echo=FALSE, warning=FALSE, out.width='100%', out.height='60%'}
library(DiagrammeR)
library(DiagrammeRsvg)
library(svglite)
library(rsvg)

grViz("
  digraph {

graph[layout = neato,
      overlap=false,
      rankdir = 'LR', 
      bgcolor='#000000',
      sep=.2]

node[fontsize=20,
      fillcolor = DimGray,
      fontcolor = White,
      fillcolor='#222222',
      style=filled,
      color=White]

edge [color = White, penwidth=4]

  'data frame'  -> 'summarize'
    subgraph detail {
       rank='same';
       'summarize' -> 'visualize' -> 'check';
       'check' -> 'summarize'; }
'check' -> '.Rmd / render';
  }")
```

---

# Part 1: Data cleaning summary

```{r res1, fig.align='center', echo=FALSE, out.width='100%'}
knitr::include_graphics('pics/summ.png', dpi = NA)
```

---

background-image: url(pics/miss.png)
background-size: 100%

# Part 2: Summary table

---

background-image: url(pics/out1.png)
background-size: 100%

# Part 3: Variable list

---

background-image: url(pics/out2.png)
background-size: 100%

---

# `dataMaid` common arguments

.small[

| Argument      | Description   |
| ------------- |:--------------------------------| 
| `mode`        | Tasks to perform. `c("summarize", "visualize", "check")` is default |
| `replace`      | Logical. Should existing dataMaid reports be overwritten? Default `FALSE`   | 
| `output`      | Output format. Choices are `"pdf"`, `"html"`, "`"word"` | 
| `onlyProblematic`      | Logical. Show only variable with problems. Default `FALSE`  |
| `maxProbVals`  | Maximum number of unique values printed. Positive int or `Inf` (default 10)     |

]

---

class: inverse, middle

# Exercise 2

1.  Load the `dataMaid` package
2.  Use `makeDataReport()` to generate a data report
3.  Screen the `bigPresidentData` data for errors we did not find before.

Hunt for errors!

---

class: center

.small[
```{r message=FALSE, echo=FALSE}
toyData
```
]

---

# Using `dataMaid` interactively

.footnotesize[
```{r}
check(toyData$events)
check(toyData$events, 
        numericChecks = "identifyMissing")
```
]

???

check er de funktioner, der bliver checket for

Kan sætte specifikke (også i makeDataReport)

Vis, at det er en liste med 2 elementer, og den form, som de har.

---

# Overview of `check` functions

.footnotesize[
```{r eval=FALSE,results='markup'}
allCheckFunctions()
```
]

.footnotesize[

```
----------------------------------------------------------------------------------
name                 description                     classes                      
-------------------- -------------------------------------------------------------
identifyCaseIssues   Identify case issues            character, factor

identifyLoners       Identify levels with < 6 obs.   character, factor

identifyMissing      Identify miscoded missing       character, Date, factor,     
                     values                          integer, labelled, logical,  
                                                          numeric                      
```                                                       
]

and more

---

# Interactive `dataMaid` - visualizations

.pull-left[
.footnotesize[
```{r, eval=FALSE}
visualize(toyData$events)
```

Can also check the available `visual` functions

```{r eval=FALSE}
allVisualFunctions()
```
]
]

.pull-right[
```{r, eval=TRUE, echo=FALSE}
visualize(toyData$events)  # Visualize variable
```
]

---

# Interactive `dataMaid` - summaries

.footnotesize[
```{r, eval=TRUE}
summarize(toyData$events)
```
]

---

# Using dataMaid interactively

.small[
```
> allSummaryFunctions()

-------------------------------------------------------------
name         description          classes                    
------------ -------------------- ---------------------------
centralValue Compute median       character, Date, factor,   
             or mode              integer, labelled, logical,
                                  numeric

countMissing Compute ratio of     character, Date, factor,   
             missing obs.         integer, labelled, logical,
                                  numeric

minMax       Find min and max     integer, numeric, Date     
             values                                                   
quartiles    Compute 1st and 3rd quartiles Date, integer, numeric

uniqueValues Count number of unique values character, Date, factor,   
                                           integer, labelled, logical,
                                           numeric

variableType Data class of variable        character, Date, factor,   
                                           integer, labelled, logical,
                                           numeric                    
----------------------------------------------------------------------

```
]

---

# Extending `dataMaid`

Custom check, visual, or summary functions.

Few requirements --- input and output formats.

Easiet to work with a **template system** and modify one of those.

Check vignette `vignette("extending_dataMaid")` for detailed instructions. Or the exercises!

---

## Custom `summaryFunction` - template

.footnotesize[
```
mySummaryFunction <- function(v, ...) {
  val <- [ result of whatever summary we are doing ]
  res <- [ properly escaped version of val ]        
  summaryResult(list(feature = "[Feature name]", 
                     result = res,
                     value = val))
}
```
]

Example (centralValue for numeric/integer)

.footnotesize[
```{r eval=FALSE}
function (v, maxDecimals = 2) {
    v <- na.omit(v)
    val <- median(v)
    summaryResult(list(feature = "Median", result = round(val, 
        maxDecimals), value = val))
}
```
]

---

## Custom `checkFunction` - template

.small[
```{r, eval=TRUE}
isSSN <- function(v, nMax = NULL, ...) {
  out <- list(problem = FALSE, 
              message = "", 
              problemValues=NULL)
  if (class(v) %in% c("character", "factor", 
                      "labelled")) {
    if (any(grep("\\d{3}-\\d{2}-\\d{4}", v))) {
      out$problem <- TRUE
      out$message <- "Warning: Seems to contain SSNs."
      out$problemValues <- "Will not show"
    }  
  }
  out 
}
```
]
---

## Using the function

.footnotesize[
```{r}
DF <- data.frame(ids=c("111-22-3333","123-45-6789", 
                       "111-22-3333"),
                 id2=c("111223333", "123456789", "4728491283"), 
                 stringsAsFactors=FALSE)

check(DF, characterChecks = c("isSSN"))
```

]

---

class: inverse

# Exercise 2b/2c

How to tailor `dataMaid` to work with *your* dataset:

*   Work interactively with `dataMaid` (exercise 2b)
*   Create new custom summary, check and visualization functions (exercise 2c)

Pick whatever you want. Or jump back and forth.

### 2018-02-08 docs/CSP2018/slides/module1/kutheme.R

@@ -1,0 +1,31 @@
+library(extrafont)
+loadfonts()
+library(ggplot2)
+library(ggthemr)
+
+
+cLAUs <- define_palette(swatch = c('#eeeeee', # White
+                                   "#8282ff", # Blue b9b9ff
+                                   "#ff7979", # Red
+                                   '#f2f200', # Yellow
+                                   "#0bff06", # Green
+                                   "#eca8e5", # Magenta
+                                   "#ffad5b"  # Brownish
+                                  ),
+                         gradient = c(lower = 'red', upper = 'green'),
+                         background = "transparent",
+                         text = c("#eeeeee", "#eeeeee"),
+                         line = c("#ee1e1e", "#6e6e6e"), # Axis lines
+                         gridline = "#434343") # Gridlines
+
+ggthemr(cLAUs, layout="scientific", type = 'outer', spacing=1.2)
+
+
+
+theme_xkcd <- function() {
+    theme(text=element_text(size=28, family="Unmasked BB"),   # Humor Sans
+          panel.background = element_rect(fill = "transparent", colour = NA), # bg of the panel
+          plot.background = element_rect(fill = "transparent", colour = NA) # bg of the plot
+                                        # change stuff here
+         )
+}

### 2018-02-08 docs/articles/dataMaid_artData.Rmd
CSP2018 slides on webpage

@@
 subtitle: "Autogenerated data summary from dataMaid"
-date: 2018-02-08 19:12:55
+date: 2018-02-08 19:41:22
 output: html_document
@@
 
- *  Report creation time: Thu Feb 08 2018 19:12:55
+ *  Report creation time: Thu Feb 08 2018 19:41:22
 

### 2018-02-09 docs/articles/dataMaid_artData.Rmd
Updated exercises for course

@@
 subtitle: "Autogenerated data summary from dataMaid"
-date: 2018-02-08 19:41:22
+date: 2018-02-09 14:33:06
 output: html_document
@@
 
- *  Report creation time: Thu Feb 08 2018 19:41:22
+ *  Report creation time: Fri Feb 09 2018 14:33:06
 

### 2018-04-11 latex/article_accept.tex
Formatting corrections for JSS

--- 2017-11-09 latex/article_vol3.tex
@@
 %\newcommand{\R}{\proglang{R}\xspace}
-
-\newcommand{\R}[1]{\code{#1}}
+% \newcommand{\R}[1]{\code{#1}}
 
@@
 quality assessment overview documents, we still wish to emphasize that
-it is \emph{not} a tool for unsupervised data cleaning. This qualifies
+it is not a tool for unsupervised data cleaning. This qualifies
 as the second paradigm of \pkg{dataMaid}: Data cleaning decisions
@@
 whether or not something is a miscoded missing value often requires
-knowledge about the full variable (e.g. range or data type), and thus
+knowledge about the full variable (e.g., range or data type), and thus
 it should be performed variable-wise. \pkg{validate} does not
@@
 \pkg{DataExplorer} \citep{DataExplorer}. While this package does not
-address data cleaning issues \emph{per se}, its general strategy is
+address data cleaning issues per se, its general strategy is
 quite similar to that of \pkg{dataMaid} and to the paradigms presented
@@
    pill  events region change id    spotifysong
-   <fctr>  <dbl> <fctr>      <dbl> <fctr>      <fctr>
- 1    red      1      a -0.6264538      1  Irrelevant
- 2    red      1      a  0.1836433      2  Irrelevant
- 3    red      1      a -0.8356286      3  Irrelevant
- 4    red      2      a  1.5952808      4  Irrelevant
- 5    red      2      a  0.3295078      5  Irrelevant
- 6    red      6      b -0.8204684      6  Irrelevant
- 7    red      6      b  0.4874291      7  Irrelevant
- 8    red      6      b  0.7383247      8  Irrelevant
- 9    red    999      c  0.5757814      9  Irrelevant
-10    red     NA      c -0.3053884     10  Irrelevant
-11   blue      4      c  1.5117812     11  Irrelevant
-12   blue     82      .  0.3898432     12  Irrelevant
-13   blue     NA        -0.6212406     13  Irrelevant
-14   <NA>    NaN  other -2.2146999     14  Irrelevant
-15   <NA>      5  OTHER  1.1249309     15  Irrelevant
+   <fct>  <dbl> <fct>   <dbl> <fct> <fct>      
+ 1 red     1.00 a      -0.626 1     Irrelevant 
+ 2 red     1.00 a       0.184 2     Irrelevant 
+ 3 red     1.00 a      -0.836 3     Irrelevant 
+ 4 red     2.00 a       1.60  4     Irrelevant 
+ 5 red     2.00 a       0.330 5     Irrelevant 
+ 6 red     6.00 b      -0.820 6     Irrelevant 
+ 7 red     6.00 b       0.487 7     Irrelevant 
+ 8 red     6.00 b       0.738 8     Irrelevant 
+ 9 red   999    c       0.576 9     Irrelevant 
+10 red    NA    c      -0.305 10    Irrelevant 
+11 blue    4.00 c       1.51  11    Irrelevant 
+12 blue   82.0  .       0.390 12    Irrelevant 
+13 blue   NA    " "    -0.621 13    Irrelevant 
+14 <NA>  NaN    other  -2.21  14    Irrelevant 
+15 <NA>    5.00 OTHER   1.12  15    Irrelevant 
 \end{Soutput}
@@
 parts, two of which are presented in Figure~\ref{fig:example1}. First,
-an overview of what was done is presented under the title \textit{Data
-  report overview}. Secondly, an index listing each variable along
+an overview of what was done is presented under the title ``Data
+  report overview''. Secondly, an index listing each variable along
 with an indication of whether it was found to be problematic or not is
@@
 in the dataset is presented in turn using (up to) three tools in the
-\textit{Variable list}: A table summarizing key features of the
+``Variable list'': A table summarizing key features of the
 variable, a figure visualizing its distribution when relevant and a list of flagged
@@
 \quad \code{ordering} & Ordering of the variables in the data summary (as is or alphabetical) & \code{"asIs"} \\
-\quad \code{onlyProblematic} & Should only variables flagged as problematic be included in the \textit{Variable list}? & \code{FALSE} \\
-%\quad \code{listChecks} & Should an overview of what checks were performed be listed in the \textit{Data report overview}? &  \code{TRUE} \\
+\quad \code{onlyProblematic} & Should only variables flagged as problematic be included in the ``Variable list''? & \code{FALSE} \\
+%\quad \code{listChecks} & Should an overview of what checks were performed be listed in the "Data report overview"x? &  \code{TRUE} \\
 %slettet for at få plads til nye argumenter
@@
 \quad \code{mode} & What steps should be performed for each variable (out
-                 of the three possibilities \textit{summarize},
-                 \textit{visualize}, \textit{check})? &
+                 of the three possibilities ``summarize'',
+                 ``visualize'', ``check'')? &
                                                         \code{c("summarize", "visualize", "check")}  \\
@@
  \quad \code{identifyOutliersTBStyle} & Identify outliers (Turkish Boxplot style) &  & & $\times$ & & $\times$ & $\times$ \\
- \quad \code{identifyWhitespace} & Identify prefixed and suffixed whitespace &  \blue{$\times$} & \blue{$\times$} & & \blue{$\times$} & & &  \\
- \quad \code{isCPR} & Identify Danish CPR numbers & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &$\times$   \\
+ \quad \code{identifyWhitespace} & Identify prefixed and suffixed white space &  \blue{$\times$} & \blue{$\times$} & & \blue{$\times$} & & &  \\
+ \quad \code{isCPR} & Identify Danish personal identification numbers & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ & $\times$ &$\times$   \\
  \quad \code{isSingular} & Check if the variable contains only a single value & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$} & \blue{$\times$}  \\
@@
 usage beyond the \code{checks} arguments, namely in the
-\textit{pre-check} stage. In this stage, it is determined whether or
-not each variable is suitable for the summarize/visualize/check (SVC)
+pre-check stage. In this stage, it is determined whether or
+not each variable is suitable for the summarize/visualize/check
 steps. The functions used in the pre-check stage should be
@@
 \item In the pre-check stage.
-\item As part of the summarize/visualize/check (SVC) steps.
+\item As part of the summarize/visualize/check steps.
 \end{enumerate}
-Each of these stages are controllable using appropriate function
+Each stage is controllable using appropriate function
 arguments in \code{makeDataReport()}, and above we have shown examples
@@
 One might also encounter datasets with variables that are not among the 7 classes mentioned in the above (\code{character}, \code{Date}, \code{factor}, \code{integer}, \code{labelled}, \code{logical} and \code{numeric}), for instance variables of type \code{complex} or user-defined classes.
-It is possible to tell \code{makeDataReport()} how to handle such variables by use of the argument \code{treatXasY}.
+\code{makeDataReport()} can be told how to handle such variables by use of the argument \code{treatXasY}.
 This argument takes a list where the names correspond to "new" variable types (\code{X}), while the entries must be supported variable types (\code{Y}).
@@
 $identifyMissing
-The following suspected missing value codes enter as regular values: 999, NaN.
+The following suspected missing value codes enter as regular
+values: 999, NaN.
 $identifyOutliers
@@
 $identifyMissing
-The following suspected missing value codes enter as regular values: 999, NaN.
+The following suspected missing value codes enter as regular
+values: 999, NaN.
 \end{Soutput}
@@
 \begin{Soutput}
-The following suspected missing value codes enter as regular values: 999, NaN.
+The following suspected missing value codes enter as regular
+values: 999, NaN.
 \end{Soutput}
@@
  $ problem      : logi TRUE
- $ message      : chr "The following suspected missing value codes enter as regular values: \\\"999\\\", \\\"NaN\\\"."
+ $ message      : chr "The following suspected missing value codes enter as 
+regular values: \\\"999\\\", \\\"NaN\\\"."
  $ problemValues: num [1:2] 999 NaN
@@
 \end{center}
-\caption{Part of the data overview report generated for the \code{presidentData} dataset. Here, we see pages 4 and 5 from the \textit{Variable list}. Larger versions of the pages can be seen in
+\caption{Part of the data overview report generated for the \code{presidentData} dataset. Here, we see pages 4 and 5 from the ``Variable list''. Larger versions of the pages can be seen in
   Appendix~\ref{sec:appendix2}.}
@@
 \begin{Sinput}
-R> data(presidentData)
+R> data("presidentData")
 R> head(presidentData)
@@
 This can be done using the \code{treatXasY} argument.
-\item We use the \code{character} class for a few categorical variables that have a lot of different levels, but where it is not a data mistake that these levels each generally have very few observations, e.g.
-the variable \code{stateOfBirth}.
+\item We use the \code{character} class for a few categorical variables that have a lot of different levels, but where it is not a data mistake that these levels each generally have very few observations, e.g., the variable \code{stateOfBirth}.
 Therefore, we would like to disable the \code{identifyLoners} check (which flags variables with <6 observations in any of the levels) for this variable type.
@@
 The first four pages of the resulting report can be inspected in Figures \ref{fig:bigExampleP01} and \ref{fig:bigExampleP23}.
-Note that all the customized settings can be identified from the first two pages, without having to read through the report: The new title is written on the front page, the check settings are displayed in the \textit{Checks performed} table and the strategy for handling \code{Name} variables is documented below this table.
+Note that all the customized settings can be identified from the first two pages, without having to read through the report: The new title is written on the front page, the check settings are displayed in the ``Checks performed'' table and the strategy for handling \code{Name} variables is documented below this table.
 The full data report, except for the front page, is available in Appendix \ref{sec:appendix2}.
@@
 On page 3, we see the contents for the three first variables.
-Here, we identifify a prefixed whitespace in the lastname entry for President Truman and we find that a dot was entered as a first name; this is a typical choice for coding missing values in e.g.
-Stata, and therefore, it is flagged as a potential miscoded missing value.
+Here, we identify a prefixed white space in the last name entry for President Truman and we find that a dot was entered as a first name; this is a typical choice for coding missing values in e.g., Stata, and therefore, it is flagged as a potential miscoded missing value.
 The variable \code{orderOfPresidency} is not summarized, visualized or checked because it is categorical and contains unique values for each observation.
@@
 \item The report flags the variable \code{ethnicity} to have suspiciously few observations in one category, "African American". 
-\item A few presidents were found to have odd values in the variable describing the duration of their presidencies: Some had very short (outlier) presidencies of less than two years and one was registered to have an \textit{infinite} presidency.
+\item A few presidents were found to have odd values in the variable describing the duration of their presidencies: Some had very short (outlier) presidencies of less than two years and one was registered to have an infinite presidency.
 \item The variable concerning age at inauguration was coded as a character variable, but consists exclusively of numbers and takes a lot of different values and therefore, it was flagged as a potentially misclassified numeric variable.
@@
 For instance, \code{ethnicity} is very reasonably marked as a potentially problematic variable as it includes only a single observation of "African American".
-However, a human reading this report will know that this does \textit{not} reflect a mistake in the data, but rather a peculiarity in the real world, and as such, it should not be cleaned out.
+However, a human reading this report will know that this does not reflect a mistake in the data, but rather a peculiarity in the real world, and as such, it should not be cleaned out.
 
 A few of the identified problems have easy fixes that need no further discussion.
-We remove the prefixed whitespace from Truman's name, fix the misspelling of New York, convert the binary variable \code{assasinationAttempt} to a factor and change the class of the \code{ageAtInauguration} variable to numeric:
+We remove the prefixed white space from Truman's name, fix the misspelling of New York, convert the binary variable \code{assasinationAttempt} to a factor and change the class of the \code{ageAtInauguration} variable to numeric:
 
@@
 
-Moving forward, we might be interested in inspecting the contents of the "."-coded entry of \code{firstName} closer, as we \textit{do} know the first names of all the US presidents.
+Moving forward, we might be interested in inspecting the contents of the "."-coded entry of \code{firstName} closer, as we do know the first names of all the US presidents.
 We look at the last name for this president and fill in the first name correctly:
@@
 We see that this is not a proper US president and thus, it is likely to be the explanation of the faulty number of observations in the dataset.
-Therefore, we drop this observation from the dataset, e.g.
-by overwriting the dataset with a selection of all the other observations:
+Therefore, we drop this observation from the dataset, e.g., by overwriting the dataset with a selection of all the other observations:
 
@@
 
-Now, if e.g. the report is created at 3 pm on October 31, 2017, its file will be named \code{dataMaid_toyData_10-31-2017_15.00.pdf}, making it easy to find.
+Now, if e.g., the report is created at 3 pm on October 31, 2017, its file will be named \code{dataMaid_toyData_10-31-2017_15.00.pdf}, making it easy to find.
 
@@
 \begin{Sinput}
-R> makeDataReport(toyData, onlyProblematic = TRUE, mode = "check", replace=TRUE)
+R> makeDataReport(toyData, onlyProblematic = TRUE, 
++    mode = "check", replace=TRUE)
 \end{Sinput}

### 2018-04-11 latex/dataMaid_presidentData_cleaned.Rmd
Formatting corrections for JSS

---
dataMaid: yes
title: Dirty president data - cleaned
subtitle: "Autogenerated data summary from dataMaid"
date: 2018-04-11 01:02:32
output: pdf_document
documentclass: report
header-includes:
  - \renewcommand{\chaptername}{Part}
  - \newcommand{\fullline}{\noindent\makebox[\linewidth]{\rule{\textwidth}{0.4pt}}}
  - \newcommand{\bminione}{\begin{minipage}{0.75 \textwidth}}
  - \newcommand{\bminitwo}{\begin{minipage}{0.25 \textwidth}}
  - \newcommand{\emini}{\end{minipage}}
---
```{r , echo=FALSE, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
library(ggplot2)
library(pander)
```
```{r visualFunctions, echo=FALSE, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
ggAggHist <- getFromNamespace("ggAggHist", "dataMaid")
ggAggBarplot <- getFromNamespace("ggAggBarplot", "dataMaid")
```
# Data report overview
The dataset examined has the following dimensions:

---------------------------------
Feature                    Result
------------------------ --------
Number of observations         45

Number of variables            11
---------------------------------

### Checks performed
The following variable checks were performed, depending on the data type of each variable:

----------------------------------------------------------------------------------------------------------------------------------
&nbsp;                                                 character    factor    labelled   numeric    integer    logical     Date   
----------------------------------------------------- ----------- ---------- ---------- ---------- ---------- --------- ----------
Identify miscoded missing values                       $\times$    $\times$   $\times$   $\times$   $\times$             $\times$

Identify prefixed and suffixed whitespace              $\times$    $\times$   $\times$

Identify case issues                                   $\times$    $\times$   $\times$

Identify misclassified numeric or integer variables    $\times$    $\times$   $\times$

Identify levels with < 6 obs.                                      $\times$   $\times$

Identify outliers                                                                        $\times$   $\times$             $\times$ 
----------------------------------------------------------------------------------------------------------------------------------

Non-supported variable types were set to be handled in the following way:

* Name is treated as character
* complex is treated as numeric

Please note that all numerical values in the following have been rounded to 2 decimals.

# Summary table

--------------------------------------------------------------------------------------------------
&nbsp;                   Variable class     # unique values   Missing observations  Any problems? 
------------------------ ---------------- ----------------- ---------------------- ---------------
[lastName]               character                       39                 0.00 %

[firstName]              character                       30                 0.00 %

[orderOfPresidency]      factor                          45                 0.00 %    $\times$

[birthday]               Date                            44                 0.00 %

[stateOfBirth]           character                       21                 0.00 %

[assassinationAttempt]   factor                           2                 0.00 %

[sex]                    factor                           1                 0.00 %    $\times$

[ethnicity]              factor                           2                 0.00 %    $\times$

[presidencyYears]        numeric                         10                 2.22 %    $\times$

[ageAtInauguration]      numeric                         22                 0.00 %    $\times$

[favoriteNumber]         complex                         11                 0.00 %                
--------------------------------------------------------------------------------------------------

# Variable list
## lastName

\bminione

-------------------------------------
Feature                        Result
------------------------- -----------
Variable type               character

Number of missing obs.        0 (0 %)

Number of unique values            39

Mode                          "Adams"
-------------------------------------

\emini
\bminitwo
```{r Var-1-lastName, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:39, .Label = c("Adams", 
"Arthur", "Buchanan", "Bush", "Carter", "Cleveland", "Clinton", 
"Coolidge", "Eisenhower", "Fillmore", "Ford", "Garfield", "Grant", 
"Harding", "Harrison", "Hayes", "Hoover", "Jackson", "Jefferson", 
"Johnson", "Kennedy", "Lincoln", "Madison", "McKinley", "Monroe", 
"Nixon", "Obama", "Pierce", "Polk", "Reagan", "Roosevelt", "Taft", 
"Taylor", "Truman", "Trump", "Tyler", "Van Buren", "Washington", 
"Wilson"), class = "factor"), y = c(2L, 1L, 1L, 2L, 1L, 2L, 1L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 1L, 1L, 1L, 1L, 2L, 1L, 1L, 1L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L
)), .Names = c("x", "y"), row.names = c(NA, -39L), class = "data.frame"), 
    vnam = "lastName")
```
\emini

\fullline

## firstName

\bminione

-------------------------------------
Feature                        Result
------------------------- -----------
Variable type               character

Number of missing obs.        0 (0 %)

Number of unique values            30

Mode                          "James"
-------------------------------------

\emini
\bminitwo
```{r Var-2-firstName, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:30, .Label = c("Abraham", 
"Andrew", "Barack", "Benjamin", "Calvin", "Chester", "Donald", 
"Dwight", "Franklin", "George", "Gerald", "Grover", "Harry", 
"Herbert", "James", "Jimmy", "John", "Lyndon", "Martin", "Millard", 
"Richard", "Ronald", "Rutherford", "Theodore", "Thomas", "Ulysses", 
"Warren", "William", "Woodrow", "Zachary"), class = "factor"), 
    y = c(1L, 2L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 3L, 1L, 2L, 1L, 
    1L, 5L, 1L, 4L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 4L, 
    1L, 1L)), .Names = c("x", "y"), row.names = c(NA, -30L), class = "data.frame"), 
    vnam = "firstName")
```
\emini

\fullline

## orderOfPresidency

* The variable is a key (distinct values for each observation).

\fullline

## birthday

\bminione

--------------------------------------------------
Feature                                     Result
------------------------- ------------------------
Variable type                                 Date

Number of missing obs.                     0 (0 %)

Number of unique values                         44

Mode                                  "1837-03-18"

Min. and max.               1732-02-22; 1961-08-04

1st and 3rd quartiles       1791-04-23; 1890-10-14
--------------------------------------------------

\emini
\bminitwo
```{r Var-4-birthday, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggHist(data = structure(list(factorV = structure(1:20, .Label = c("1732-02-22", 
"1743-08-13", "1755-02-01", "1766-07-24", "1778-01-12", "1789-07-03", 
"1800-12-23", "1812-06-13", "1823-12-04", "1835-05-25", "1846-11-13", 
"1858-05-04", "1869-10-23", "1881-04-14", "1892-10-03", "1904-03-25", 
"1915-09-14", "1927-03-05", "1938-08-25", "1950-02-13"), class = "factor"), 
    Freq = c(3L, 1L, 1L, 3L, 2L, 4L, 3L, 2L, 3L, 3L, 2L, 2L, 
    2L, 3L, 0L, 4L, 3L, 0L, 3L, 1L), xmin = structure(c(-86876, 
    -82686, -78496, -74305, -70115, -65925, -61735, -57545, -53354, 
    -49164, -44974, -40784, -36594, -32403, -28213, -24023, -19833, 
    -15643, -11452, -7262), class = "Date"), xmax = structure(c(-82686, 
    -78496, -74305, -70115, -65925, -61735, -57545, -53354, -49164, 
    -44974, -40784, -36594, -32403, -28213, -24023, -19833, -15643, 
    -11452, -7262, -3072), class = "Date"), ymin = c(0, 0, 0, 
    0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0), ymax = c(3L, 
    1L, 1L, 3L, 2L, 4L, 3L, 2L, 3L, 3L, 2L, 2L, 2L, 3L, 0L, 4L, 
    3L, 0L, 3L, 1L)), .Names = c("factorV", "Freq", "xmin", "xmax", 
"ymin", "ymax"), row.names = c(NA, -20L), class = "data.frame"), 
    vnam = "birthday")
```
\emini

\fullline

## stateOfBirth

\bminione

-------------------------------------
Feature                        Result
------------------------- -----------
Variable type               character

Number of missing obs.        0 (0 %)

Number of unique values            21

Mode                           "Ohio"
-------------------------------------

\emini
\bminitwo
```{r Var-5-stateOfBirth, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:21, .Label = c("Arkansas", 
"California", "Connecticut", "Georgia", "Hawaii", "Illinois", 
"Iowa", "Kentucky", "Massachusetts", "Missouri", "Nebraska", 
"New Hampshire", "New Jersey", "New York", "North Carolina", 
"Ohio", "Pennsylvania", "Tennessee", "Texas", "Vermont", "Virginia"
), class = "factor"), y = c(1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 4L, 
1L, 1L, 1L, 2L, 5L, 2L, 8L, 1L, 1L, 2L, 2L, 7L)), .Names = c("x", 
"y"), row.names = c(NA, -21L), class = "data.frame"), vnam = "stateOfBirth")
```
\emini

\fullline

## assassinationAttempt

\bminione

-----------------------------------
Feature                      Result
------------------------- ---------
Variable type                factor

Number of missing obs.      0 (0 %)

Number of unique values           2

Mode                            "0"
-----------------------------------

\emini
\bminitwo
```{r Var-6-assassinationAttempt, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:2, .Label = c("0", 
"1"), class = "factor"), y = c(35L, 10L)), .Names = c("x", "y"
), row.names = c(NA, -2L), class = "data.frame"), vnam = "assassinationAttempt")
```
\emini

\fullline

## sex

* The variable only takes one (non-missing) value: \"Male\". The variable contains 0 \% missing observations.

\fullline

## ethnicity

\bminione

---------------------------------------
Feature                          Result
------------------------- -------------
Variable type                    factor

Number of missing obs.          0 (0 %)

Number of unique values               2

Mode                        "Caucasian"
---------------------------------------

\emini
\bminitwo
```{r Var-8-ethnicity, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:2, .Label = c("African American", 
"Caucasian"), class = "factor"), y = c(1L, 44L)), .Names = c("x", 
"y"), row.names = c(NA, -2L), class = "data.frame"), vnam = "ethnicity")
```
\emini

- Note that the following levels have at most five observations: \"African American\".

\fullline

## presidencyYears

\bminione

--------------------------------------
Feature                         Result
------------------------- ------------
Variable type                  numeric

Number of missing obs.      1 (2.22 %)

Number of unique values              9

Median                               4

1st and 3rd quartiles          3.75; 8

Min. and max.                    0; 12
--------------------------------------

\emini
\bminitwo
```{r Var-9-presidencyYears, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggHist(data = structure(list(factorV = structure(1:20, .Label = c("[0,0.6]", 
"(0.6,1.2]", "(1.2,1.8]", "(1.8,2.4]", "(2.4,3]", "(3,3.6]", 
"(3.6,4.2]", "(4.2,4.8]", "(4.8,5.4]", "(5.4,6]", "(6,6.6]", 
"(6.6,7.2]", "(7.2,7.8]", "(7.8,8.4]", "(8.4,9]", "(9,9.6]", 
"(9.6,10.2]", "(10.2,10.8]", "(10.8,11.4]", "(11.4,12]"), class = "factor"), 
    Freq = c(2L, 1L, 0L, 4L, 4L, 0L, 15L, 0L, 3L, 0L, 0L, 3L, 
    0L, 11L, 0L, 0L, 0L, 0L, 0L, 1L), xmin = c(0, 0.6, 1.2, 1.8, 
    2.4, 3, 3.6, 4.2, 4.8, 5.4, 6, 6.6, 7.2, 7.8, 8.4, 9, 9.6, 
    10.2, 10.8, 11.4), xmax = c(0.6, 1.2, 1.8, 2.4, 3, 3.6, 4.2, 
    4.8, 5.4, 6, 6.6, 7.2, 7.8, 8.4, 9, 9.6, 10.2, 10.8, 11.4, 
    12), ymin = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 
    0, 0, 0, 0, 0), ymax = c(2L, 1L, 0L, 4L, 4L, 0L, 15L, 0L, 
    3L, 0L, 0L, 3L, 0L, 11L, 0L, 0L, 0L, 0L, 0L, 1L)), .Names = c("factorV", 
"Freq", "xmin", "xmax", "ymin", "ymax"), row.names = c(NA, -20L
), class = "data.frame"), vnam = "presidencyYears")
```
\emini

- Note that the following possible outlier values were detected: \"0\", \"1\", \"2\".

\fullline

## ageAtInauguration

\bminione

-----------------------------------
Feature                      Result
------------------------- ---------
Variable type               numeric

Number of missing obs.      0 (0 %)

Number of unique values          22

Median                           55

1st and 3rd quartiles        51; 58

Min. and max.                42; 70
-----------------------------------

\emini
\bminitwo
```{r Var-10-ageAtInauguration, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggHist(data = structure(list(factorV = structure(1:20, .Label = c("[42,43.4]", 
"(43.4,44.8]", "(44.8,46.2]", "(46.2,47.6]", "(47.6,49]", "(49,50.4]", 
"(50.4,51.8]", "(51.8,53.2]", "(53.2,54.6]", "(54.6,56]", "(56,57.4]", 
"(57.4,58.8]", "(58.8,60.2]", "(60.2,61.6]", "(61.6,63]", "(63,64.4]", 
"(64.4,65.8]", "(65.8,67.2]", "(67.2,68.6]", "(68.6,70]"), class = "factor"), 
    Freq = c(2L, 0L, 2L, 2L, 3L, 2L, 4L, 2L, 5L, 7L, 4L, 1L, 
    1L, 3L, 1L, 2L, 1L, 0L, 1L, 2L), xmin = c(42, 43.4, 44.8, 
    46.2, 47.6, 49, 50.4, 51.8, 53.2, 54.6, 56, 57.4, 58.8, 60.2, 
    61.6, 63, 64.4, 65.8, 67.2, 68.6), xmax = c(43.4, 44.8, 46.2, 
    47.6, 49, 50.4, 51.8, 53.2, 54.6, 56, 57.4, 58.8, 60.2, 61.6, 
    63, 64.4, 65.8, 67.2, 68.6, 70), ymin = c(0, 0, 0, 0, 0, 
    0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0), ymax = c(2L, 
    0L, 2L, 2L, 3L, 2L, 4L, 2L, 5L, 7L, 4L, 1L, 1L, 3L, 1L, 2L, 
    1L, 0L, 1L, 2L)), .Names = c("factorV", "Freq", "xmin", "xmax", 
"ymin", "ymax"), row.names = c(NA, -20L), class = "data.frame"), 
    vnam = "ageAtInauguration")
```
\emini

- Note that the following possible outlier values were detected: \"69\", \"70\".

\fullline

## favoriteNumber

\bminione

-----------------------------------
Feature                      Result
------------------------- ---------
Variable type               complex

Number of missing obs.      0 (0 %)

Number of unique values          11

Median                            6

1st and 3rd quartiles          3; 8

Min. and max.                 0; 10
-----------------------------------

\emini
\bminitwo
```{r Var-11-favoriteNumber, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggHist(data = structure(list(factorV = structure(1:20, .Label = c("[0,0.5]", 
"(0.5,1]", "(1,1.5]", "(1.5,2]", "(2,2.5]", "(2.5,3]", "(3,3.5]", 
"(3.5,4]", "(4,4.5]", "(4.5,5]", "(5,5.5]", "(5.5,6]", "(6,6.5]", 
"(6.5,7]", "(7,7.5]", "(7.5,8]", "(8,8.5]", "(8.5,9]", "(9,9.5]", 
"(9.5,10]"), class = "factor"), Freq = c(1L, 2L, 0L, 4L, 0L, 
5L, 0L, 6L, 0L, 4L, 0L, 3L, 0L, 6L, 0L, 6L, 0L, 4L, 0L, 4L), 
    xmin = c(0, 0.5, 1, 1.5, 2, 2.5, 3, 3.5, 4, 4.5, 5, 5.5, 
    6, 6.5, 7, 7.5, 8, 8.5, 9, 9.5), xmax = c(0.5, 1, 1.5, 2, 
    2.5, 3, 3.5, 4, 4.5, 5, 5.5, 6, 6.5, 7, 7.5, 8, 8.5, 9, 9.5, 
    10), ymin = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 
    0, 0, 0, 0, 0), ymax = c(1L, 2L, 0L, 4L, 0L, 5L, 0L, 6L, 
    0L, 4L, 0L, 3L, 0L, 6L, 0L, 6L, 0L, 4L, 0L, 4L)), .Names = c("factorV", 
"Freq", "xmin", "xmax", "ymin", "ymax"), row.names = c(NA, -20L
), class = "data.frame"), vnam = "favoriteNumber")
```
\emini

\fullline

Report generation information:

 *  Created by Claus Thorn Ekstrøm (username: `cld189`).

 *  Report creation time: Wed Apr 11 2018 01:02:32

 *  Report Was run from directory: `/Users/cld189/ku/R/cleanR/latex`

 *  dataMaid v1.1.0 [Pkg: 2018-03-02 from local (ekstroem/dataMaid@NA)]

 *  R version 3.4.4 (2018-03-15).

 *  Platform: x86_64-apple-darwin15.6.0 (64-bit)(macOS High Sierra 10.13.3).

 *  Function call: `makeDataReport(data = presidentData, replace = TRUE, vol = "_cleaned", 
    checks = setChecks(character = defaultCharacterChecks(remove = "identifyLoners")), 
    reportTitle = "Dirty president data - cleaned", treatXasY = list(Name = "character", 
        complex = "numeric"))`

### 2018-04-26 latex/article_accept.tex
Fixes bug with factors with unused factor levels

@@
 On page 3, we see the contents for the three first variables.
-Here, we identify a prefixed white space in the last name entry for President Truman and we find that a dot was entered as a first name; this is a typical choice for coding missing values in e.g., Stata, and therefore, it is flagged as a potential miscoded missing value.
+Here, we identify a prefixed white space in the last name entry for President Truman and we find that a dot was entered as a first name; this is a typical choice for coding missing values in e.g., \proglang{Stata}, and therefore, it is flagged as a potential miscoded missing value.
 The variable \code{orderOfPresidency} is not summarized, visualized or checked because it is categorical and contains unique values for each observation.
@@
 R> foo <- function(x) {
-+    any(sapply(x, function(y) y[["problem"]]))
-+  }
++    any(sapply(x, function(y) y[["problem"]])) }
 R> sapply(toyChecks, foo)

### 2018-04-26 R/identifyLoners.R

@@ -60,32 +60,32 @@ identifyLonersMessage <- "Note that the following levels have at most five obser
 
 #factor variables
 identifyLonersF <- function(v, nMax) {
-  vLev <- levels(v)
-  v <- factor(na.omit(v)) #drop unused levels
-  lonerOcc <- vLev[which(table(v) <= 5)]
-  if (length(lonerOcc) > 0) {
-    problem <- TRUE
-    problemValues <- lonerOcc
-  } else {
-    problem <- FALSE
-    problemValues <- NULL
-  }
-  outMessage <- messageGenerator(list(problem=problem,
-                                      problemValues=problemValues),
-                                 message = identifyLonersMessage,
-                                 nMax = nMax)
-  checkResult(list(problem = problem, message = outMessage,
-                   problemValues = problemValues))
+    v <- factor(na.omit(v))  # Drop unused levels
+    vLev <- levels(v)        # Get the levels that are left.
+    lonerOcc <- vLev[which(table(v) <= 5)]
+    if (length(lonerOcc) > 0) {
+        problem <- TRUE
+        problemValues <- lonerOcc
+    } else {
+        problem <- FALSE
+        problemValues <- NULL
+    }
+    outMessage <- messageGenerator(list(problem=problem,
+                                        problemValues=problemValues),
+                                   message = identifyLonersMessage,
+                                   nMax = nMax)
+    checkResult(list(problem = problem, message = outMessage,
+                     problemValues = problemValues))
 }
 
 #character variables
 identifyLonersC <- function(v, nMax) {
-  v <- factor(v)
-  identifyLonersF(v, nMax)
+    v <- factor(v)
+    identifyLonersF(v, nMax)
 }
 
 identifyLonersL <- function(v, nMax) {
-  v <- haven::as_factor(v)
-  identifyLonersF(v, nMax)
+    v <- haven::as_factor(v)
+    identifyLonersF(v, nMax)
 }
 

### 2018-04-26 tests/testthat/testisLoner.R

@@ -1,0 +1,15 @@
+context("check function: isLoner")
+
+library("dataMaid")
+
+
+## Read atomic vectors
+source("atomic.R")
+
+x <- factor(LETTERS[c(4,4,1,2,1,3,2,4,2, 4,4,1,2,1,3,2,4,2, 4,4,1,2,1,3,2,4,2)+1], levels=c("A", "B", "C", "D", "E"))
+                                       
+## Each type should result in a list
+test_that("isLoner returns the right values", {
+    expect_equal(dataMaid::identifyLoners(x)$problemValues, "D")
+})
+

### 2018-05-10 docs/eRum2018/slides/module3/pics/module4.Rmd
Added eRum slides+exercises

---
title: "Reproducible data screening and quality control"
subtitle: "Module 3: Time to clean and other options"
author: "Claus Thorn Ekstrøm and Anne Helby Petersen<br>UCPH Biostatistics"
date: "eRum, Budapest, May 14th, 2018<br>.footnotesize[[Slides/homepage](https://ekstroem.github.io/dataMaid/eRum2018/index.html)]"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ['default', 'ku.css', 'animate.min.css']
    seal: true
    nature:
      ratio: '16:9'
      slideNumberFormat: '%current%' 
      highlightStyle: ir-black
      highlightLines: true
      countIncrementalSlides: false
      beforeInit: "js/zoom.js"
---

```{r setup, include=FALSE}
library(knitr)
options(htmltools.dir.version = FALSE, cache=TRUE)
opts_chunk$set(comment = NA, prompt=TRUE)
#opts_chunk$set(dev.args=list(bg="transparent"), fig.width=15, fig.height=7)
source("kutheme.R")
library(dataMaid)
toyData <- as.data.frame(toyData)
```

#Summarizing the errors

* What did we find so far?
* What didn't we find?

---

#Data cleaning

.center[
```{r cleanpic, echo=FALSE, out.width="60%"}
knitr::include_graphics("pics/datacleaning.jpg")
```
]

Not the best term ... and should not be unsupervised

---

#Data cleaning in R

In an R-script:

1. Make a copy of the dataset. 
2. Use indexing to locate the problem in the data.
3. Overwrite the faulty value with a correct one - if you know it - or `NA` to mark that information is missing in this spot. 
4. Save the copy of the "cleaned" data in a **new** file.

---

#Selection - rows/observations

Two systems for selecting observations in `data.frame`s in R:

By index (row number) or using a logical vector.

```{r}
(tD <- head(toyData, 3))  
```

---

#Selection - rows/observations

Four equivalent ways to get the second line of `tD`:
```{r, eval = FALSE}
tD[2, ] #indexing
tD[c(FALSE, TRUE, FALSE), ] #manual logical vector 
tD[tD$id == 2, ] #informative logical vector
tD %>% filter(id==2)  # Using tidyverse
```

```{r, echo = FALSE}
tD[tD$id == 2, ] #informative logical vector
```

---

#Selection - rows/observations

Use informative logical vectors as much as possible!

```{r}
tD

#Mark non-positive change as missing:
tD[tD$change > 0, "change"] <- NA
```

---

#Selection - columns/variables

*ALWAYS* use variable names.

```{r, eval = FALSE}
#readable, informative code:
tD[tD$change > 0, "change"] <- NA

# Indexing by numbers easily becomes 
# a source of error by itself:
tD[tD$change > 0, 4] <- NA
```

---

background-image: url(pics/structure.png)
background-size: 30%
background-position: right

## Finishing up after cleaning

Should now have<br>
a cleaned dataset<br>
that can form the<br>
basis for future<br>
analyses.<br>

With documentation<br>
of how we got<br>
there!

---

# Create codebook

Produce a summary document for subsequent analyses.

.footnotesize[
```{r eval=FALSE}
makeCodebook(presidentData)
```
]

Add label (similar to `labelled` package) or extra information

.footnotesize[
```{r}
pD <- presidentData
attr(pD$presidencyYears, "label") <- 
  "Full years as president"
```
]

.footnotesize[
```{r}
attr(pD$birthday, "shortDescription") <- 
  "Dates are stored in YYYY-MM-DD format"
```
]

---

class: inverse

# Exercise 3

Correct the errors you have found so far.

Make sure to make the cleaning process reproducible.

Remember **rules 1 and 2**!

Create the final codebook with additional information about some of the variables.

```{r eval=FALSE}
makeCodebook(myCleanedData)
```

---

##Row-wise or column-wise checks?

```{r colrow1, echo=FALSE, out.width="75%", fig.align = "center"}
knitr::include_graphics("pics/colrow1.png")
```

---

##Row-wise or column-wise checks?

```{r colrow2, echo=FALSE, out.width="75%", fig.align = "center"}
knitr::include_graphics("pics/colrow2.png")
```

---

##Row-wise or column-wise checks?

```{r colrow3, echo=FALSE, out.width="75%", fig.align = "center"}
knitr::include_graphics("pics/colrow3.png")
```

---

##Row-wise *and* column-wise constraints!

* `dataMaid` performs class dependent checks for each variable in a dataset, one at a time (column-wise)
  + Pros: Easy to document what was (not) done, let's you get started without a lot of prior knowledge, easy to share with collaborators
  + Shortcomings: Generally cannot detect internal consistency issues or use non-class dependent variable constraints

---

##Row-wise *and* column-wise constraints!

An R-packages that performs row-wise checks: `validate`. Check out the talk on Wednesday @ 14.50 by Edwin de Jonge:

**validatetools - resolve and simplify contradictive or redundant data validation rules**

Note: Different use of the term "validation" - no longer about format, type and range, but used as synonym to "check".

---

class: middle, center

# Thank you!

Please grab hold of us here or via email

.pull-left[Anne<br>[ahpe@sund.ku.dk](mailto:ahpe@sund.ku.dk)]  .pull-right[Claus<br>[ekstrom@sund.ku.dk](mailto:ekstrom@sund.ku.dk)]

### 2019-09-12 testdata_dateprob.rmd
Label ordering is now preserved for factor visualizations. Fixes #46.

---
dataMaid: yes
title: testData
subtitle: "Autogenerated data summary from dataMaid"
date: 2019-07-10 10:40:23
output: pdf_document
documentclass: report
header-includes:
  - \renewcommand{\chaptername}{Part}
  - \newcommand{\fullline}{\noindent\makebox[\linewidth]{\rule{\textwidth}{0.4pt}}}
  - \newcommand{\bminione}{\begin{minipage}{0.75 \textwidth}}
  - \newcommand{\bminitwo}{\begin{minipage}{0.25 \textwidth}}
  - \newcommand{\emini}{\end{minipage}}
---
```{r , echo=FALSE, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
library("ggplot2")
library("pander")
```

```{r 'visualFunctions', echo=FALSE, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
ggAggHist <- getFromNamespace("ggAggHist", "dataMaid")
ggAggBarplot <- getFromNamespace("ggAggBarplot", "dataMaid")
```

# Data report overview
The dataset examined has the following dimensions:

---------------------------------
Feature                    Result
------------------------ --------
Number of observations         15

Number of variables            15
---------------------------------

### Checks performed
The following variable checks were performed, depending on the data type of each variable:

---------------------------------------------------------------------------------------------------------------------------------------------------
&nbsp;                                                 character    factor    labelled   haven labelled   numeric    integer    logical     Date   
----------------------------------------------------- ----------- ---------- ---------- ---------------- ---------- ---------- --------- ----------
Identify miscoded missing values                       $\times$    $\times$   $\times$      $\times$      $\times$   $\times$             $\times$

Identify prefixed and suffixed whitespace              $\times$    $\times$   $\times$      $\times$

Identify levels with < 6 obs.                          $\times$    $\times$   $\times$      $\times$

Identify case issues                                   $\times$    $\times$   $\times$      $\times$

Identify misclassified numeric or integer variables    $\times$    $\times$   $\times$      $\times$

Identify outliers                                                                                         $\times$   $\times$             $\times$ 
---------------------------------------------------------------------------------------------------------------------------------------------------

Please note that all numerical values in the following have been rounded to 2 decimals.

# Summary table

-------------------------------------------------------------------------------------------------
&nbsp;                  Variable class     # unique values   Missing observations  Any problems? 
----------------------- ---------------- ----------------- ---------------------- ---------------
[charVar]               character                        5                 6.67 %    $\times$

[factorVar]             factor                           5                 0.00 %    $\times$

[numVar]                numeric                         10                 0.00 %

[intVar]                integer                         10                 0.00 %

[boolVar]               logical                          3                20.00 %

[keyVar]                character                       15                 0.00 %    $\times$

[emptyVar]              numeric                          1                 0.00 %    $\times$

[numOutlierVar]         numeric                         15                 0.00 %    $\times$

[smartNumVar]           numeric                          2                 0.00 %

[cprVar]                character                        5                 0.00 %    $\times$

[cprKeyVar]             character                       15                 0.00 %    $\times$

[miscodedMissingVar]    character                       14                 0.00 %    $\times$

[misclassifiedNumVar]   factor                          12                 0.00 %    $\times$

[dateVar]               Date                             4                 0.00 %

[labelledVar]           labelled                         3                13.33 %    $\times$    
-------------------------------------------------------------------------------------------------

# Variable list
## charVar

\bminione

--------------------------------------
Feature                         Result
------------------------- ------------
Variable type                character

Number of missing obs.      1 (6.67 %)

Number of unique values              4

Mode                               "a"
--------------------------------------

\emini
\bminitwo
```{r 'Var-1-charVar', echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:4, .Label = c("a", 
"b", "c", "d"), class = "factor"), y = c(6L, 4L, 2L, 2L)), class = "data.frame", row.names = c(NA, 
-4L)), vnam = "charVar")
```

\emini

- Note that the following levels have at most five observations: \"b\", \"c\", \"d\".

\fullline

## factorVar

\bminione

-----------------------------------
Feature                      Result
------------------------- ---------
Variable type                factor

Number of missing obs.      0 (0 %)

Number of unique values           5

Mode                            "a"

Reference category              999
-----------------------------------

\emini
\bminitwo
```{r 'Var-2-factorVar', echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:5, .Label = c("999", 
"a", "b", "c", "d"), class = "factor"), y = c(1L, 6L, 4L, 2L, 
2L)), class = "data.frame", row.names = c(NA, -5L)), vnam = "factorVar")
```

\emini

- The following suspected missing value codes enter as regular values: \"999\".

- Note that the following levels have at most five observations: \"999\", \"b\", \"c\", \"d\".

\fullline

## numVar

\bminione

------------------------------------
Feature                       Result
------------------------- ----------
Variable type                numeric

Number of missing obs.       0 (0 %)

Number of unique values           10

Median                             5

1st and 3rd quartiles       1.5; 6.5

Min. and max.                  1; 10
------------------------------------

\emini
\bminitwo
```{r 'Var-3-numVar', echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggHist(data = structure(list(xmin = c(0, 2, 4, 6, 8), xmax = c(2, 
4, 6, 8, 10), ymin = c(0, 0, 0, 0, 0), ymax = c(5L, 2L, 4L, 2L, 
2L)), class = "data.frame", row.names = c(NA, -5L)), vnam = "numVar")
```

\emini

\fullline

## intVar

\bminione

------------------------------------
Feature                       Result
------------------------- ----------
Variable type                integer

Number of missing obs.       0 (0 %)

Number of unique values           10

Median                             5

1st and 3rd quartiles       1.5; 6.5

Min. and max.                  1; 10
------------------------------------

\emini
\bminitwo
```{r 'Var-4-intVar', echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggHist(data = structure(list(xmin = c(0, 2, 4, 6, 8), xmax = c(2, 
4, 6, 8, 10), ymin = c(0, 0, 0, 0, 0), ymax = c(5L, 2L, 4L, 2L, 
2L)), class = "data.frame", row.names = c(NA, -5L)), vnam = "intVar")
```

\emini

\fullline

## boolVar

\bminione

------------------------------------
Feature                       Result
------------------------- ----------
Variable type                logical

Number of missing obs.      3 (20 %)

Number of unique values            2

Mode                          "TRUE"
------------------------------------

\emini
\bminitwo
```{r 'Var-5-boolVar', echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:2, .Label = c("FALSE", 
"TRUE"), class = "factor"), y = c(4L, 8L)), class = "data.frame", row.names = c(NA, 
-2L)), vnam = "boolVar")
```

\emini

\fullline

## keyVar

* The variable is a key (distinct values for each observation).

\fullline

## emptyVar

* The variable only takes one (non-missing) value: \"1\". The variable contains 0 \% missing observations.

\fullline

## numOutlierVar

\bminione

-------------------------------------
Feature                        Result
------------------------- -----------
Variable type                 numeric

Number of missing obs.        0 (0 %)

Number of unique values            15

Median                              8

1st and 3rd quartiles       4.5; 11.5

Min. and max.                  1; 100
-------------------------------------

\emini
\bminitwo
```{r 'Var-8-numOutlierVar', echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggHist(data = structure(list(xmin = c(0, 20, 40, 60, 80), 
    xmax = c(20, 40, 60, 80, 100), ymin = c(0, 0, 0, 0, 0), ymax = c(14L, 
    0L, 0L, 0L, 1L)), class = "data.frame", row.names = c(NA, 
-5L)), vnam = "numOutlierVar")
```

\emini

- Note that the following possible outlier values were detected: \"100\".

\fullline

## smartNumVar

* Note that this variable is treated as a factor variable below, as it only takes a few unique values.

\bminione

-----------------------------------
Feature                      Result
------------------------- ---------
Variable type               numeric

Number of missing obs.      0 (0 %)

Number of unique values           2

Mode                            "1"

Reference category                0
-----------------------------------

\emini
\bminitwo
```{r 'Var-9-smartNumVar', echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:2, .Label = c("0", 
"1"), class = "factor"), y = 7:8), class = "data.frame", row.names = c(NA, 
-2L)), vnam = "smartNumVar")
```

\emini

\fullline

## cprVar

\bminione

-----------------------------------------
Feature                            Result
------------------------- ---------------
Variable type                   character

Number of missing obs.            0 (0 %)

Number of unique values                 5

Mode                        "010188-1350"
-----------------------------------------

\emini
\bminitwo
```{r 'Var-10-cprVar', echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:5, .Label = c("010188-1350", 
"010290-0219", "010291-2373", "020187-3282", "040506-9660"), class = "factor"), 
    y = c(3L, 3L, 3L, 3L, 3L)), class = "data.frame", row.names = c(NA, 
-5L)), vnam = "cprVar")
```

\emini

- Note that the following levels have at most five observations: \"010188-1350\", \"010290-0219\", \"010291-2373\", \"020187-3282\", \"040506-9660\".

\fullline

## cprKeyVar

* The variable is a key (distinct values for each observation).

\fullline

## miscodedMissingVar

\bminione

-------------------------------------
Feature                        Result
------------------------- -----------
Variable type               character

Number of missing obs.        0 (0 %)

Number of unique values            14

Mode                              "9"
-------------------------------------

\emini
\bminitwo
```{r 'Var-12-miscodedMissingVar', echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:13, .Label = c("", 
"-", "-inf", "-Inf", ".", "9", "inf", "Inf", "na", "Na", "NA", 
"nan", "NAN"), class = "factor"), y = c(1L, 1L, 1L, 1L, 1L, 2L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L)), class = "data.frame", row.names = c(NA, 
-13L)), vnam = "miscodedMissingVar")
```

\emini

- The following suspected missing value codes enter as regular values: \"\", \"-\", \"-inf\", \"-Inf\", \".\", \"9\", \"inf\", \"Inf\", \"na\", \"Na\" (4 additional values omitted).

- Note that the following levels have at most five observations: \"\", \"-\", \"-inf\", \"-Inf\", \".\", \"9\", \"inf\", \"Inf\", \"na\", \"Na\" (4 additional values omitted).

- Note that there might be case problems with the following levels: \"-inf\", \"-Inf\", \"inf\", \"Inf\", \"na\", \"Na\", \"NA\", \"nan\", \"NaN\", \"NAN\".

\fullline

## misclassifiedNumVar

\bminione

-----------------------------------
Feature                      Result
------------------------- ---------
Variable type                factor

Number of missing obs.      0 (0 %)

Number of unique values          12

Mode                            "1"

Reference category                1
-----------------------------------

\emini
\bminitwo
```{r 'Var-13-misclassifiedNumVar', echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:12, .Label = c("1", 
"2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12"), class = "factor"), 
    y = c(2L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 2L, 1L, 1L, 1L)), class = "data.frame", row.names = c(NA, 
-12L)), vnam = "misclassifiedNumVar")
```

\emini

- Note that the following levels have at most five observations: \"1\", \"10\", \"11\", \"12\", \"2\", \"3\", \"4\", \"5\", \"6\", \"7\" (2 additional values omitted).

- Note: The variable consists exclusively of numbers and takes a lot of different values. Is it perhaps a misclassified numeric variable?

\fullline

## dateVar

\bminione

--------------------------------------------------
Feature                                     Result
------------------------- ------------------------
Variable type                                 Date

Number of missing obs.                     0 (0 %)

Number of unique values                          4

Mode                                  "1997-05-10"

Min. and max.               1997-05-10; 2013-04-02

1st and 3rd quartiles       1997-05-10; 2005-12-10
--------------------------------------------------

\emini
\bminitwo
```{r 'Var-14-dateVar', echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggHist(data = structure(list(xmin = structure(c(9000, 10000, 
11000, 12000, 13000, 14000, 15000), class = "Date"), xmax = structure(c(10000, 
11000, 12000, 13000, 14000, 15000, 16000), class = "Date"), ymin = c(0, 
0, 0, 0, 0, 0, 0), ymax = c(10L, 0L, 0L, 0L, 3L, 0L, 2L)), row.names = c(NA, 
-7L), class = "data.frame"), vnam = "dateVar")
```

\emini

\fullline

## labelledVar

\bminione

---------------------------------------
Feature                          Result
------------------------- -------------
Variable type                  labelled

Number of missing obs.      2 (13.33 %)

Number of unique values               2

Mode                              "Joe"
---------------------------------------

\emini
\bminitwo
```{r 'Var-15-labelledVar', echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:2, .Label = c("Joe", 
"Peter"), class = "factor"), y = c(8L, 5L)), class = "data.frame", row.names = c(NA, 
-2L)), vnam = "labelledVar")
```

\emini

- Note that the following levels have at most five observations: \"Peter\".

\fullline

Report generation information:

 *  Created by annennenne (username: `zms499`).

 *  Report creation time: Wed Jul 10 2019 10:40:24

 *  Report was run from directory: `C:/Users/zms499/Documents/P-kopier/dataMaid`

 *  dataMaid v1.3.0 [Pkg: 2019-07-10 from local]

 *  R version 3.5.3 (2019-03-11).

 *  Platform: x86_64-w64-mingw32/x64 (64-bit)(Windows >= 8 x64 (build 9200)).

 *  Function call: `makeDataReport(data = testData, file = "testdata_dateprob.rmd", 
    replace = TRUE)`

### 2019-09-12 R/basicVisual.R

@@ -82,11 +82,18 @@ basicVisual <- visualFunction(basicVisual, "Histograms and barplots using graphi
 #' importFrom stats na.omit
 #' @inheritParams standardVisual
 basicVisualCFLB <- function(v, vnam, doEval = TRUE) {
-  v <- escapeRStyle(na.omit(v))
+  v <- na.omit(v)
   if (identifyNums(v, nVals = 0)$problem) {
-    v <- as.numeric(as.character(v))
+    v <- factor(as.numeric(as.character(v)))
+  } else {
+    if (is.factor(v)) {
+      levs <-  escapeRStyle(levels(v))
+      v <- factor(escapeRStyle(v), levels = levs)
+    } else {
+      v <- escapeRStyle(v)
+      v <- factor(v)
+    }
   }
-  v <- factor(v)
   aggrV <- table(v)
   thisCall <- call("barplot", height = aggrV, main = vnam)
   if (doEval) {

### 2019-09-12 R/standardVisual.R

@@ -82,9 +82,16 @@ standardVisual <- visualFunction(standardVisual, "Histograms and barplots using 
 
 #character, factor, labelled and logical variables
 standardVisualCFLB <- function(v, vnam, doEval = TRUE) {
-  v <- escapeRStyle(na.omit(v))
+  v <- na.omit(v)
   if (identifyNums(v, nVals = 0)$problem) {
     v <- as.numeric(as.character(v))
+  } else {
+    if (is.factor(v)) {
+      levs <-  escapeRStyle(levels(v))
+      v <- factor(escapeRStyle(v), levels = levs)
+    } else {
+      v <- escapeRStyle(v)
+    }
   }
   pf <- aggregateForBarplot(v)
   thisCall <- call("ggAggBarplot", data = pf, vnam = vnam)

### 2019-12-08 R/clean.R
Updated package with checks of external programs: git, whoami, pandoc

@@ -1,1165 +1,0 @@
-#' Produce a data cleaning overview document (deprecated version)
-#'
-#' NOTE: This function has been replaced by \code{\link{makeDataReport}}. The current 
-#' function is no longer updated and it is only included for backwards compatability. 
-#' 
-#' Run a set of class-specific validation checks to check the
-#' variables in a dataset for potential errors.  Performs checking
-#' steps according to user input and/or data type of the inputted
-#' variable.  The checks are saved to an R markdown file which can
-#' rendered into an easy-to-read document.  This document also
-#' includes summaries and visualizations of each variable in the
-#' dataset.
-#'
-#' For each variable, a set of pre-check (controlled by the
-#' \code{preChecks} argument) is first run and then then a battery of
-#' functions are applied depending on the variable class.  For each
-#' variable type the summarize/visualize/check functions are applied
-#' and and the results are written to an R markdown file.
-#'
-#' @param data The dataset to be checked. This dataset should be of class \code{data.frame},
-#' \code{tibble} or \code{matrix}. If it is of classs \code{matrix}, it will be converted to a
-#' \code{data.frame}.
-#'
-#' @param output
-#' Output format. Options are \code{"pdf"} (the default), and \code{"html"}
-#'
-#' @param render Should the output file be rendered (defaults to \code{TRUE}),
-#' i.e. should a pdf/html document be generated and saved to the disc?
-#'
-#' @param useVar Variables to clean. If \code{NULL} (the default), all variables in \code{data}
-#' are included. If a vector of variable names is supplied, only the variables in \code{data} that are
-#' also in \code{useVar} are included in the data cleaning overview document.
-#'
-#' @param ordering Choose the ordering of the variables in the variable presentation. The options
-#' are "asIs" (ordering as in the dataset) and "alphabetical" (alphabetical order).
-#'
-#' @param onlyProblematic A logical. If \code{TRUE}, only the variables flagged as
-#' problematic in the check step will be included in the variable list.
-#'
-#' @param labelled_as A string explaining the way to handle labelled vectors.
-#' Currently \code{"factor"} (the default) is the only possibility. This means that labelled
-#' variables that appear factor-like (by having a non-\code{NULL} \code{labels}-attribute) will
-#' be treated as factors, while other labelled variables will be treated as whatever base
-#' variable class they inherit from.
-#'
-#'
[…]

### 2019-12-08 R/makeDataReport.R

@@ -1050,7 +1050,11 @@ makeDataReport <- function(data, output=NULL, render=TRUE,
     writer("\n")
     
     writer("Report generation information:\n")
-    writer(" *  Created by ", whoami::fullname(fallback="Could not determine from system") , " (username: `", whoami::username(fallback="Unknown"),  "`).\n")
+    if (whoami_available()) {
+      writer(" *  Created by: ", whoami::fullname(fallback="Could not determine from system") , " (username: `", whoami::username(fallback="Unknown"),  "`).\n")
+    } else {
+      writer(" *  Created by: Could not determine from system (username: Unknown)\n")
+    }
     writer(" *  Report creation time: ", format(Sys.time(), "%a %b %d %Y %H:%M:%S"),"\n")
     writer(" *  Report was run from directory: `", getwd(),"`\n")
     

### 2019-12-08 R/render.R

@@ -19,5 +19,11 @@
 #' @seealso \code{\link[rmarkdown]{render}}.
 #'
 #' @export
-render <- function(file, quiet) rmarkdown::render(file, quiet=quiet)
+render <- function(file, quiet) {
 
+    if (!rmarkdown::pandoc_available()) {
+        stop("pandoc appears not to be installed on the system. Please see the vignette for the rmarkdown package on how to install pandoc")
+    }
+
+    rmarkdown::render(file, quiet=quiet)
+}

### 2019-12-08 R/utility.R

@@ -1,39 +1,14 @@
-
-#' Find path to the pandoc binary by checking the \code{PATH} and the \code{RSTUDIO_PANDOC} env vars
-#' @return logical that is TRUE is pandoc can be found
+#' Find out if the whoami package binaries is installed (git + whoami)
+#' @return logical that is TRUE if whoami and git can be found
 #' @export
-pandoc_exists <- function() {
+whoami_available <- function() {
 
     ## check if pandoc can be found on the path
-    bin <- as.character(Sys.which('pandoc'))
-
-    ## check RStudio env var pointing to the bundled pandoc
-    if (Sys.getenv('RSTUDIO_PANDOC') != '') {
-        bin <- Sys.getenv('RSTUDIO_PANDOC')
-        bin <- file.path(bin, ifelse(
-            grepl('w|W', .Platform$OS.type),
-            ## we are on Windows
-            'pandoc.exe',
-            ## no extension on Mac and Linux
-            'pandoc'))
-    }
+    whoami_bin <- as.character(Sys.which('whoami'))
+    git_bin <- as.character(Sys.which('git'))
 
     ## return whatever found
-    bin != ''
-
-}
-
-
-#' Find path to the pandoc binary by checking the \code{PATH} and the \code{RSTUDIO_PANDOC} env vars
-#' @return logical that is TRUE is pandoc can be found
-#' @export
-whoami_exists <- function() {
-
-    ## check if pandoc can be found on the path
-    bin <- as.character(Sys.which('whoami'))
-
-    ## return whatever found
-    bin != ''
+    (whoami_bin != '') & (git_bin != '')
 
 }
 

### 2020-07-27 doc/extending_dataMaid.Rmd
Changed default behavior of printProblemValues to omit middle range values rather than last. Closes #51

---
title: "Extending dataMaid"
author: "Anne Helby Petersen and Claus Thorn Ekstrøm"
date: "July 3, 2020"
output: 
  rmarkdown::html_vignette:
    toc: true
vignette: >
  %\VignetteIndexEntry{Extending dataMaid}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE)
library(dataMaid)
Sys.setenv(TZ="Europe/Copenhagen")  ## Set time zone to prevent warnings
Sys.getenv("TZ")
```

# Introduction
The data documentation tools available in `dataMaid` allows for it to be utilized as an easy-to-use, self-contained package, as outlined in our [article manuscript](https://github.com/ekstroem/dataMaid/blob/master/latex/article_vol3.pdf). However, this does *not* imply that `dataMaid` is not customizable: `dataMaid` was specifically built for including user-made extensions and this vignette is a tutorial into how one can write simple extensions that can be used in e.g. the `makeDataReport()` function from `dataMaid`.

## Three steps of data documentation
In order to understand how different parts of the report creation can be customized, we will briefly dwell on the structure of the data reports made by `dataMaid`. Aside from a few pages dedicated to gaining an overview of the full dataset and what was done to document its current state, the report sequentially presents information for each variable. More specifically, each variable is presented with one of the two sets of information:

1. The result of a single eligibility check (referred to as a *pre-check* below)
2. The results of the three canonical `dataMaid` steps:
    + Summarize: Produces a summary table
    + Visualize: Produces a distribution visualization
    + Check: Performs and lists the results of a number of checks

The first option will be used for variables that do not pass the initial pre-checks. These are checks that are performed in order to identify variables that are not eligible for detailed documentation for one reason or another. The default pre-checks used in `makeDataReport()` are

* `isKey`: Check if the variable is a key, i.e. categorical with unique values for each observation 
* `isSingular`: Check if the variable only takes a single (non-`NA`) value
* `isSupported`: Check if the class of the variable is among the classes supported by `dataMaid`

If any of these checks finds a problem, the relevant problem will be mentioned in the report and the variable will not be exposed to further documentation steps. If the pre-checks do not flag a variable, on the other hand, the variable will be subjected to the three SVC (summarize, visualize, check) steps mentioned above. In the SVC-steps, `makeDataReport()` calls a number of so-called `summaryFunctions`, `visualFunctions` and `checkFunctions`, with different choices of functions depending on the class of the variable. The customizability features of `dataMaid` essentially comes down to writing such `summaryFunctions`, `visualFunctions` and `checkFunctions`.

This vignette consists of three parts. First, we describe how new
`summaryFunction`s, `visualFunction`s and
`checkFunctions` can be made by describing the requirements for their structures that must be fulfilled in order for them to be used in a `makeDataReport()` call. Secondly,  we
turn to a worked example of how to write custom made functions in
practice. Here, we define and showcase six new SVC functions. Lastly, a built-in dataset about art masterpieces, `artData`, is documented using the custom-made functions and the thereby obtained data report is included.

In this vignette, we focus on the report creation tools of `dataMaid`. However, it should be emphasized that by following the guidelines presented here, the user-defined `summaryFunctions`, `visualFunctions` and `checkFunctions` will also be fully integrated with the interactive mode of `dataMaid`.

# Function templates
In order to construct a summary, visual or check function, one needs
to create a new function with a specific structure.  This can be done with
different levels of strictness. If
the new custom function is only to be used as part of the SVC steps in `makeDataReport()`,
then only the input/output structure of the function needs special attention.
However, new user-defined functions can also be registered locally to
be part of the full machinery of `dataMaid`, and when this is done, the functions will
be recognized and behave in the same way as the built-in functions in
`dataMaid`. The presentation below is given in the format of
function templates, written in pseudo-code. These templates are
designed for getting the full functionality, but note that
the table below serves as a reference to the minimal
requirements for each function type, while also presenting the "full" versions.

For each of the three function types, we have provided an S3 object class. This was done in order to facilitate obtaining overviews of all available options for summaries, visuals and checks through calls to the three functions `allSummaryFunctions()`, `allVisualFunctions()` and `allCheckFunctions()`. Similarly, we have also provided specific output classes for the summary- and check outputs, for which convenient printing methods are available.

This table summarizes the minimal and recommended structures for summary-, visual- and check functions:

|   | `summaryFunction`    | `visualFunction`    | `checkFunction`    |
|---|----------------------|---------------------|--------------------|
| Input (required) | `v` - A variable vector. <br/> `...` - Additional arguments passed to the function. |  `v` - A variable vector. <br/> `vnam` - The variable name (as character string). <br/> `doEval` - A logical (`TRUE`/`FALSE`) controlling the output type of the function. | `v` - A variable vector <br/> `nMax` - An integer (or `Inf`), controlling how many problematic values are printed, if relevant <br/> `...` - Additional arguments passed to the function.|
| Input (optional) |  `maxDecimals` - The number of decimals printed in outputted numerical values.  |  |  `maxDecimals` - The number of decimals printed in outputted numerical values.|
|Purpose | Describe some aspect of the variable, e.g., a central value, its dispersion or level of missingness. | Produce a distribution plot. | Check a variable for a specific issue and, if relevant, identify the values in the variable that cause the issue. |
| Output (required) | A list with entries: <br/> `feature` - A label for the summary value (as character string);  <br/> `result` - The result of the summary (as character string). | A character string with `R` code for producing a plot. This code should be standalone, i.e., should include the data if necessary. | A list with entries: <br/> `problem` - A logical identifying whether an issue was found; <br/> `message` - A character string (possibly empty) describing the issue that was found, properly escaped and ready for use with `rmarkdown`. |
| Output (recommended) | A `summaryResult` object, i.e., an attributed list with entries `feature`, `result` and `value`, the latter being the values from `result` in their original format). | If `doEval` is `TRUE`: A plot that is opened by the graphic device in `R`. <br/> If `doEval` is `FALSE`:  A text string with `R` code, as described above.  | A `checkResult` object, i.e., an attributed list with entries `problem`, `message` and `problemValues`, the latter being either `NULL` or the problem causing values, as they were found in `v`, whichever is relevant. |
| Tools available for producing the function | `summaryResult()` |  | `messageGenerator()` <br/> `checkResult()`|

## Writing a summaryFunction
As mentioned above, `dataMaid` provides a dedicated class for
`summaryFunction`s. However, this does not imply that they are
particularly advanced or complicated to create; in fact, they are
nothing but regular functions with a particular
input/output-structure. Specifically, they all follow the template
below:

```
mySummaryFunction <- function(v, ...) {
  val <- [result of whatever summary we are doing]
  res <- [properly escaped version of val]
  summaryResult(list(feature = "[Feature name]", result = res,
                     value = val))
}
```

The last function called here, `summaryResult()`, changes the
class of the output, thereby making a `print()` method available
for it.  Note that `v` is an input vector and that `res`
should be either a character string or something that will be printed
as one. In other words an integer would be allowed for `res`, 
but a matrix will not. Note that, strictly speaking, only one of the two elements `value` and `result` is needed
in order to create a `summaryResult`. If `result` is provided, this will be printed as the summary result. However, if only `value` is provided, the `summaryResult()` function will try to convert it to a character string itself. This might be more or less difficult to do in a reasonable way, so therefore, we really do recommend to provide this conversion yourself by both returning a `result` and a `value`.

Though a lot of different things can go into the
`summaryFunction` template, we recommend only using it for
summarizing the features of a variable, and leaving tests and checks
for the `checkFunctions` (presented below).

Adhering to the template above is sufficient for using the freshly
made `mySummaryFunction()` in `makeDataReport()`, but we recommend
adding the new function to the overview of all summary
functions by converting it to a proper `summaryFunction`
object. This is done by calling the `summaryFunction()` creator with
the user-defined function as the first argument, and additional arguments
`description` (an explanatory text which will be added to the attributes of
the function), and `classes` (a vector of variable classes the
user-defined function is intended to be applied to, also stored as an
attribute). In other words, a call following the template below should be made:

```
mySummaryFunction <- summaryFunction(mySummaryFunction,
  description = "[Text describing what the summaryFunction does]",
  classes = c([vector of data types that the function is intended
  for]))
``` 
which adds the new function to the output of an
`allSummaryFunctions()` call. If `mySummaryFunction` is
constructed as an `S3` generic function with associated methods,
the call to `summaryFunction()` will automatically produce a
vector of the names of the classes for which the function can be
called. If `mySummaryFunction()` is not an `S3` generic and
`classes` is left unspecified, the attribute will simply be left
empty. Note that the helper function `allClasses()` might be
useful for filling out the `classes` argument, as it simply lists
all supported classes in `dataMaid`:

```{r allClasses}
allClasses()
```

## Writing a visualFunction
`visualFunction`s are the functions that produce the figures in a
`dataMaid` output document. Writing a `visualFunction` is slightly
more complicated than writing a `summaryFunction`. This is due to
the fact that `visualFunction`s need to be able to output standalone
code for plots in order for `makeDataReport()` to build standalone
`rmarkdown` files. We recommend using the following structure
(again shown as pseudo code):

```
myVisualFunction <- function(v, vnam, doEval) {
  thisCall <- call("[the name of the function used to produce the plot]",
    v, [additional arguments for the plotting function])
  if (doEval) {
    return(eval(thisCall))
  } else return(deparse(thisCall))
}
```

In this function, `v` is the variable to be visualized,
`vnam` is its name (which should generally be passed to
`title` or `main` arguments in plotting functions) and
`doEval` controls whether the output is a plot (if `TRUE`)
or a character string of standalone code for producing a plot (if
`FALSE`). Implementing the `doEval = TRUE` setting is not
strictly necessary for a `visualFunction`'s use in `makeDataReport()`, but
it makes it easier to assess what visualization options are available,
and obviously, it is crucial for interactive usage of
`myVisualFunction()`. In either case, it should be noted that all
the parameters listed above, `v`, `vnam` and `doEval`,
are mandatory, so they must be left as is, even if they are not in
use (c.f., the table).

As with the summary function, we call `visualFunction()` to
register our newly created function:

```
myVisualFunction <- visualFunction(myVisualFunction,
  description = "[Some text describing the visualFunction]",
  classes = c([data types that this function is intended for]))
  )
```

Now, `myVisualFunction()` will be available in a `allVisualFunctions()`
call, just like the two build-in `visualFunctions`, `standardVisual`
and `basicVisual`.

## Writing a checkFunction
The last, but perhaps most important, `dataMaid` function type is
the `checkFunction`. These are the functions that flag issues in
the data in the *check* step and/or control the overall flow of the data
documenting process in the pre-check stage. A `checkFunction` follows
one of two overall structures, depending on the type of check. Either,
it tries to identify problematic values in the variable (as
e.g., `identifyMissing()` does), or it performs a check concerning the
variable as a whole (e.g., the functions used for pre-checks and the
function `identifyNums()`). We present templates for both types of
`checkFunction`s below separately, but it should be emphasized that
formally, they belong to the same class.

First, a template for the full-variable check function type, where we
first define the function and subsequently register it as a check
function using `checkFunction()`:

```  
myFullVarCheckFunction <- function(v, ...) {
  [do your check]
  problem <- [is there a problem? TRUE/FALSE]
  message <- "[message describing the problem, if any]"
  checkResult(list(problem = problem,
    message = message,
    problemValues = NULL))
}

myFullVarCheckFunction <- checkFunction(myFullVarCheckFunction,
  description = "[Some text describing the checkFunction]",
  classes = c([the data types that this function is intended to be used for])
  )
```

Again, as with `summaryFunctions` and `visualFunctions`, the
change of function class by use of `checkFunction()` is not strictly
necessary. Note however, that if `myFullVarCheckFunction` is to be
used in the summarize/visualize/check steps in `makeDataReport()`, the
description attribute will be printed in the overview table in the
*Data report overview* part of the report.

If problematic values are to be identified, the template from above
should be expanded to follow a slightly more complicated structure
(again shown as pseudo-code):

```
myProbValCheckFunction <- function(v, nMax, maxDecimals, ...) {
  [do your check]
  problem <- [is there a problem? TRUE/FALSE]
  problemValues <- [vector of values in v that are problematic]
  problemStatus <- list(problem = problem,
                        problemValues = problemValues)

  problemMessage <- "[Message that is printed prior to listing
                     problem values in the dataMaid output,
                     ending with a colon]"

  outMessage <- messageGenerator(problemStatus, problemMessage, nMax)

  checkResult(list(problem = problem,
    message = outMessage,
    problemValues = problemValues))
}

myProbValCheckFunction <- checkFunction(myProbValCheckFunction,
  description = "[Some text describing the checkFunction]",
  classes = c([the data types that this function is intended to be used for])
)
```

One comment should be devoted to the helper function,
`messageGenerator()`.  This function's sole purpose is aiding
consistent styling of all `checkFunction` messages. The function
simply pastes together the `problemMessage` and the
`problemValues`, with the latter being quoted and sorted
alphabetically. If the `nMax` argument to `messageGenerator()` is
not `Inf`, only the first `nMax` problem values will be pasted
onto the message, accompanied by a comment about how many problem
values were left out (if any).  Note that printing quotes in
rmarkdown requires an extensive amount of character escaping, so
opting for `messageGenerator()` really is the easiest solution.

In the template above, the argument `maxDecimals` is not in use. This
argument should be used to round off the `problemValues` passed to
`messageGenerator()`, if they are numerical.  This can be done by
adding an extra line of code after defining `problemStatus` in the template
above:

```  
myProbValCheckFunction <- function(v, nMax, maxDecimals, ...) {
  [... more lines of code here ...]
  problemStatus <- list(problem = problem,
    problemValues = problemValues)

  if (!is.null(problemValues)) {
    problemStatus$problemValues <- round(problemValues, maxDecimals)
  }
  [... more lines of code here ...]
}
```

Now, problematic values will be rounded in the outputted message, while they
will still appear in their original format under the entry `problemValues`
in the returned object.

# A worked example

As an example, we now create some new functions and show both how they can be
used interactively and how they can be integrated with the `makeDataReport()`
function. These new functions are:

* New `summaryFunctions`: 
    + **`countZeros()`**: A new `summaryFunction` that counts the number of occurrences of the value `0` in a variable. This function will be used in the *summarize* step of `makeDataReport()`.
    + **`meanSummary()`**: A new `summaryFunction` that computes the mean of numerical variables. This function is also intended to be used in the *summarize* step.
* New `visualFunctions`:  
    + **`mosaicVisual()`**: A new `visualFunction` that produces mosaic plots. This function will be used in the *visualize* step of `makeDataReport()`.
    + **`prettierHist()`**: A new `visualFunction` that makes `ggplot2` histograms, but with contours around each bar (as is the default in the `graphics` histograms). This function will also be used in the *visualize*  step. 
* New `checkFunctions`: 
    + **`isID()`**: A new `checkFunction` intended for use in the pre-check-stage. This function checks whether a variable consists exclusively of long (at least 8 characters/digits) entries that are all of equal length, as this might be personal identification codes that we do not wish to print out in the data summary.
    + **`identifyColons()`** A new `checkFunction` that flags variables in which some observations have colons that appear in between other characters. This is practical for identifying autogenerated interaction effects. This function will be used in the *check* step of `makeDataReport()`.

These functions are defined in turn below, and afterwards, an example of how they can be called from `makeDataReport()` is provided.

## summaryFunction examples: countZeros() and meanSummary()
We start by defining two summary functions. For both functions, we use `summaryResult()` for creating the function output and then we use `summaryFunction()` in order to add the functions to the output from a `allSummaryFunctions()` call.

### countZeros --- a simple summaryFunction
First, we will make a summary function that counts how many times the value `0` occurs in a variable - either as a character or as a numeric value. We define this `summaryFunction` in the following lines of code:

```{r defineCountZeros}
countZeros <- function(v, ...) {
  val <- length(which(v == 0))
  summaryResult(list(feature = "No. zeros", result = val, value = val))
}
```

As this function computes an integer (the number of zeros),
there is no difference between the elements `$result` and
`$value`. If, on the other hand, the result had been a character
string, extra formatting might be required in the `$result` entry
(such as escaping of quotation marks), and in this scenario, the two
entries would have differed.

Because the result is returned as a
`summaryResult` object, a printing method is automatically called
when `countZeros()` is used interactively:

```{r exampleCountZeros}
#Called on a numeric variable
countZeros(c(rep(0, 5), 1:100))

#Called on a character variable
countZeros(c(rep(0, 5), letters))
```
Note that `letters` is a globally defined vector consisting of all the letters in the (English) alphabet as characters.

We change the class of `countZeros()` in order to make it appear in `allSummaryFunctions()` calls. Moreover, we
wish to emphasize that the function is not intended to be called on all variable types, as zeros have different roles in `Date` and `logical` variables:

```{r makeCountZerosSumFun}
countZeros <- summaryFunction(countZeros,
  description = "Count number of zeros",
  classes = c("character", "factor", "integer",
              "labelled", "numeric"))
```

Please note that this is just meta-information: The function can still be called on `Date` and `logical` variables, but then the user will know that they are acting against the recommendations of the programmer. In order to control what input variables can actually be used, we suggest writing the function as an S3 generic with methods for some, but not all, data classes, as the next example will show.

### meanSummary --- an S3 generic summary function
In this example, we will use the concept of method dispatch and generic S3 functions in order to control what data classes a `summaryFunction` can be used for. If you are not familiar with S3 classes, we refer to Hadley Wickham's [Advanced R](http://adv-r.had.co.nz/OO-essentials.html) for a excellent, though somewhat computer science heavy, introduction to object oriented programming in R.

We would like to define a very simple summary function that computes the arithmetic mean of a variable. This is of course only meaningful for numerical variables, i.e. variables with the classes numeric`, `integer` or `logical`. Therefore, we will use the S3 framework to force users to only call our function on these appropriate classes.

First, we define the generic function:
```{r meanSummaryGeneric}
meanSummary <- function(v, maxDecimals = 2) {
  UseMethod("meanSummary")
}
```

This is just an empty shell: All it does is to say that if the function `meanSummary` is called, R should look for class-specific methods called  `meanSummary`. However, no such methods are defined yet, and therefore an error is produced if the function is called:

```{r meanSummaryNoMethod, error = TRUE}
meanSummary(1)
```

We now define a helper function that input a variable `v`, removes `NA`s, computes the mean and outputs this mean as a proper `summaryResult`:

```{r meanSummaryHelperDef}
meanSummaryHelper <- function(v, maxDecimals) {
  #remove missing observations
  v <- na.omit(v)

  #compute mean and store "raw" output in `val`
  val <- mean(v)

  #store printable (rounded) output in `res`
  res <- round(val, maxDecimals)

  #output summaryResult
  summaryResult(list(feature = "Mean", result = res, value  = val))
}
```

and we then assign this function as the `meanSummary` method used for `logical`, `numeric` and `integer` variables:

```{r meanSummaryAssignMethods}
#logical
meanSummary.logical <- function(v, maxDecimals = 2) {
  meanSummaryHelper(v, maxDecimals)
}

#numeric
meanSummary.numeric <- function(v, maxDecimals = 2) {
  meanSummaryHelper(v, maxDecimals)
}

#integer
meanSummary.integer <- function(v, maxDecimals = 2) {
  meanSummaryHelper(v, maxDecimals)
}
```

and now we can see what happens when the function is called on supported - and non-supported - variables:

```{r meanSummaryExample, error = TRUE}
#called on a numeric variable (supported)
meanSummary(rnorm(100))

#called on a character variable - produces error as there is
#no method for characters
meanSummary(letters)
```

Finally, we will make it a proper `summaryFunction`. Because we defined `meanSummary()` as an S3 generic function, we do not have to specify which classes it should be called up - `summmaryFunction()` will automatically look up what classes have methods and are thus supported:

```{r meanSummaryAsSummaryFunction}
meanSummary <- summaryFunction(meanSummary,  
                               description = "Compute arithmetic mean")
```

And now it appears - together with `countZeros` - in a `allSummaryFunctions()` call:

```{r allSummFunctions, collapse = TRUE}
allSummaryFunctions()
```

## visualFunction examples: mosaicVisual() and prettierHist() 
We will now define two different visual functions. The first one is a `graphics`-based function, while the second one is `ggplot2`-based.

### mosaicVisual --- a new visualFunction for categorical data
In the base `R` package `graphics`, a function which produces
mosaic plots is available. It is intended to be called on one-way tables, as. e.g. returned from `table()`. This function is named `mosiacplot` and results in the following type of plots:

```{r mosiacplotExample}
#construct a character variable by sampling 100 values that are 
#either "a" (probability 0.3) or "b" (probability 0.7):
x <- sample(c("a", "b"), size = 100, replace = TRUE,
                   prob = c(0.3, 0.7))

#draw a mosaic plot of the distribution:
mosaicplot(table(x))
```

As evident from the figure, a mosaic plot consists of several rectangles, one for each category in the data, and the area of each such rectangle is proportional to the proportion of the data that takes that value - much like a pie chart. We intend to use this function as part of a new visualization function.  We will define
the new function such that it gets the full `dataMaid`
functionality. This can be done using the following code which sets up
the call using the existing function `mosaicplot()`.

```{r defineMosaicVis}
mosaicVisual <- function(v, vnam, doEval) {
  #Define a (unevaluated) call to mosaicplot
  thisCall <- call("mosaicplot", table(v), main = vnam, xlab = "")

  #if doEval is TRUE, evaluate the call, thereby producing a plot
  #if doEval is FALSE, return the deparsed call
  if (doEval) {
    return(eval(thisCall))
  } else return(deparse(thisCall))
}
```

This function can now be called directly or used in `makeDataReport()`, as will
presented in an example below. Depending on the `doEval` argument, either a text string
with code or a plot is produced. We can for instance inspect the code for the plot produced in the above by calling `mosaicVisual()` on `x` with `doEval = FALSE`:

```{r exampleMosVis}
mosaicVisual(x, "Variable x", doEval = FALSE)
```

Even though `mosaicVisual()`, as written above, follows the style of a
`visualFunction`, it is not yet truly one and therefore, it will not
appear in an `allVisualFunctions()` call. In order to get this
functionality, we need to change its object class. This can be done by
writing

```{r makeMosVisVisFun}
mosaicVisual <- visualFunction(mosaicVisual,
                               description = "Mosaic plots using graphics",
                               classes = setdiff(allClasses(), 
                                                 c("numeric", 
                                                   "integer", 
                                                   "Date")))
```

Here, we use the function `allClasses()` to quickly obtain a vector
of all the seven variable classes supported by `dataMaid`, and we use `setdiff()` to choose all
classes except the `numeric`, `integer` and `Date` classes, as the mosaic plot is most suited for categorical data. Note that if `mosaicVisual()` were an S3 generic function, this argument could
have been left as `NULL` and then the classes for which methods are
available would be added automatically by `visualFunction()`.

### prettierHist() --- a customized ggplot2 histogram 
Next, we will define a `ggplot2`-based plotting function. We will simply make a slight alteration of the looks of the typical `ggplot2` histogram, but it serves of a general example of how `ggplot2` visual functions can be built.

But first, we define a helper function that does the actual plotting. This makes it simpler to write a visual function using the `call()`/`eval()` structure presented in the above, thereby making it easy to provide a visual function that can both be used interactively (with `doEval = TRUE`) and in `makeDataReport()` (with `doEval = FALSE`). The helper function is defined like this:

```{r definePrettierHistHelper}
library(ggplot2)

prettierHistHelper <- function(v, vnam) {
  #define a ggplot2 histogram 
  p <- ggplot(data.frame(v = v), aes(x = v)) + 
          geom_histogram(col = "white", bins = 20) +
          xlab(vnam)

  #return the plot
  p
}
```

We use `col = "white"` to add white contours around the bars in the histogram. Let's look at an example of this function called on a variable consisting of 100 random draws from the standard normal distribution:

```{r prettierHistHelperExample}
prettierHistHelper(rnorm(100), "Standard normal variable")
```

Now, we can define a visual function that calls `prettierHistHelper()`:

```{r definePrettierHist}
#define visualFunction-style prettierHist()-function
prettierHist <- function(v, vnam, doEval = TRUE) {
  #define the call
  thisCall <- call("prettierHistHelper", v = v, vnam = vnam)

  #evaluate or deparse
  if (doEval) {
    return(eval(thisCall))
  } else return(deparse(thisCall))
}

#Make it a proper visualFunction:
prettierHist <- visualFunction(prettierHist, 
    description = "ggplot2 style histogram with contours",
    classes = c("numeric", "integer", "logical", "Date"))
```

We specify that the `prettierHist()` should only be called on variables of the classes `numeric`, `integer`, `logical` or `Date`, as is reasonable for a histogram. And we will now find our new `visualFunctions` in a `allVisualFunctions()`-call:

```{r showAllVisFunc}
allVisualFunctions()
```

## checkFunction examples: isID() and identifyColons() 
Next up is two new check functions. First, we will define a check function that is intended for use in the pre-check stage, i.e. used for screening variables for eligibility for the summarize/visualize/check-steps. This will be a function that checks the entries in a variable for a certain structure which often used in ID type variables and we will name the new function `isID()`. Afterwards, we will define a check function that looks for `:` in categorical variables, as these might be the result of unintended interactions introduced in the data wrangling process.

### isID --- a new checkFunction without problem values
First, let's define the `isID()` function. As this function is
not supposed to list problematic values, it falls within the category
of `checkFunctions` represented by
`myFullVarCheckFunction()` above. We do not particularly wish to
use this function interactively, so we will stick to the minimal
requirements of a `checkFunction` used in `check()` (see
the overview table in the beginning of this vignette). We are interested in checking whether variables that are neither `logical` nor `Date` are restricted to a certain structure. In this example, we simply check if the all observations, when converted to characters, have the same lengths and that length is at least 8. This might be regarded as minimal restrictions for a lot of different ID type information, including US social security numbers and credit card numbers, and of course, a more specific ID check function can be built as an expansion of `isID()`. We define the function:

```{r defineIsID}
isID <- function(v, nMax = NULL, ...) {
  #define minimal output. Note that this is not a
  #proper checkResult
  out <- list(problem = FALSE, message = "")

  #only perform check if the variable is neither a logical nor a Date
  if (class(v) %in% setdiff(allClasses(), c("logical", "Date"))) {

    #count the number of characters in each entry of v
    v <- as.character(v)
    lengths <- c(nchar(v))

    #check if all entries of v have at least 8 characters
    #and whether they all have the same length. If so,
    #flag as a problem.
    if (all(lengths >= 8) & length(unique(lengths)) == 1) {
      out$problem <- TRUE
      out$message <- "Warning: This variable seems to contain ID codes."
    }
  }

  #return result of the check
  out
}
```

This is essentially all we need to do in order to include this
function as a pre-check-function in `makeDataReport()`. However, we should note that e.g. not using `checkResult()` for the output means that it does not have a convenient `print()` method available; now, the output just really is displayed as a list:

```{r isIDexample}
#define 9-character ID variable:
idVar <- c("1234-1233", "9221-0289",
           "9831-1201", "6722-1243")

#check for ID resemblance for the ID variable
isID(idVar)

#check for ID resemblance for a non-ID variable
isID(rnorm(10))

```

As we have not changed the class of this check function to be a `checkFunction`, it should be noted that it will *not* show up in the output of a `allCheckFunctions()` call.

### identifyColons --- a new checkFunction with problem values

The last function we will define in this vignette is `identifyColons()`. We
define it using the helper function `messageGenerator()` to
obtain a properly escaped message, and we use `checkResult()` to
make its output print neatly. As mentioned above, the purpose of this check function is to identify variables where colons appear, as they might have been introduced by mistake when loading or wrangling the data. In the code we use regular expressions through the `gregexpr()` function to identify colons that appear in between other characters:

```{r a1}
identifyColons <- function(v, nMax = Inf, ... ) {
  #remove duplicates (for speed) and missing values:
  v <- unique(na.omit(v))

  #Define the message displayed if a problem is found:
  problemMessage <- "Note that the following values include colons:"

  #Initialize the problem indicator (`problem`) and 
  #the faulty values (`problemValues`)
  problem <- FALSE
  problemValues <- NULL

  #Identify values in v that has the structure: First something (.), 
  #then a colon (:), and then something again (.), i.e. values with
  #non-trailing colons: 
  problemValues <- v[sapply(gregexpr(".:.", v),
                            function(x) all(x != -1))]

  #If any problem values are identified, set the problem indicator
  #accordingly
  if (length(problemValues) > 0) {
    problem <- TRUE
  }

  #Combine the problem indicator and the problem values
  #into a problem status object that can be passed to 
  #the messageGenerator() helper function that will
  #make sure the outputted message is properly escaped
  #for inclusion in the dataMaid report
  problemStatus <- list(problem = problem, 
                        problemValues = problemValues)
  outMessage <- messageGenerator(problemStatus, problemMessage, nMax)

  #Output a checkResult with the problem, the escaped
  #message and the raw problem values.
  checkResult(list(problem = problem,
                   message = outMessage,
                   problemValues = problemValues))
}
```

And we change the class of the function so that it becomes a
proper `checkFunction` object:

```{r identifyColonsIsCheckFunction}
identifyColons <- checkFunction(identifyColons,
    description = "Identify non-trailing colons",
    classes = c("character", "factor", "labelled"))
```

Note, that for `checkFunctions`, the description provided in a `checkFunction()` call will appear in
the report produced by `makeDataReport()` (in the *Data cleaning
  summary* section), so for this type of function, the
change of class is not only done for the sake of the
`allCheckFunctions()` output.

Let's see the results of `identifyColons()` when used interactively on a potentially problematic variable:
```{r identifyColonsPosExample}
#define a variable as an interaction between between two factors:
iaVar <- factor(c("a", "b", "a", "c")):factor(c(1, 2, 3, 4))

#Check iaVar for colons:
identifyColons(iaVar)
```

And when used on a variable that does not contain colons:

```{r identifyColonsNegExample}
identifyColons(letters)
```

## Calling the new summarize/visualize/check functions from makeDataReport()

Now, we are ready to use all the new functions in a `makeDataReport()`
call. We wish to produce a report that utilizes the newly defined functions in the following way:

* We want to add the new pre-check function, `isID()`, to the already existing pre-checks.
* We wish to change the plot type for `factor` variables to be `mosaicVisual()`, while `numeric`, `integer` and `Date` variables should use `prettierHist()` for their visualizations.
* We want the new summary function, `countZeros()`, to be added to the summaries performed on all variable types except for `Date` and `logical`.
* We want the new summary function, `meanSummary()`, to be used for `numeric`, `integer` and `logical` variables.
* We want the new check function, `identifyColons()`, to be added to the checks performed on `character`, `factor` and `labelled` variables.

We will make a data report for the built-in dataset `artData` that is based on the *Master Works of Art* data from [Data Explorer](http://www.data-explorer.com/data). The dataset contains 200 observations (corresponding to paintings) of 11 variables, describing different properties of the paintings and their artists. The first 5 observations look like this:

```{r headArtData}
data(artData)
head(artData, 5)
```

The 11 variables of `artData` can be summarized as follows:

* `ArtistID`: A unique ID used for cataloging the artists (fictional)
* `ArtistName`: The name of the artist
* `NoOfMiddleNames`: The number of middle names the artist has 
* `Title`: The title of the painting
* `Year`: The approximate year in which the painting was made
* `Location`: The current location of the painting
* `Continent`: The continent of the current location of the painting
* `Width`: The width of the painting, in centimeters
* `Height`: The height of the painting, in centimeters
* `Media`: The media/materials of the painting
* `Movement`: The artistic movement(s) the painting belongs to

The options for the data report described above are specified as follows:

```{r makeReport, message  = FALSE, warning = FALSE, eval=rmarkdown::pandoc_available()}
makeDataReport(artData,
  #add extra precheck function         
  preChecks = c("isKey", "isSingular", "isSupported", "isID"),

  #Add the extra summaries - countZeros() for character, factor, 
  #integer, labelled and numeric variables and meanSummary() for integer, 
  #numeric and logical variables:
  summaries = setSummaries(
    character = defaultCharacterSummaries(add = "countZeros"),
    factor = defaultFactorSummaries(add = "countZeros"),
    labelled = defaultLabelledSummaries(add = "countZeros"),
    numeric = defaultNumericSummaries(add = c("countZeros", "meanSummary")),
    integer = defaultIntegerSummaries(add = c("countZeros", "meanSummary")),
    logical = defaultLogicalSummaries(add =  c("meanSummary"))
  ),

  #choose mosaicVisual() for categorical variables,
  #prettierHist() for all others:
  visuals = setVisuals(
    factor = "mosaicVisual",
    numeric = "prettierHist",
    integer = "prettierHist", 
    Date = "prettierHist"
  ),

  #Add the new checkFunction, identifyColons, for character, factor and 
  #labelled variables:
  checks = setChecks(
    character = defaultCharacterChecks(add = "identifyColons"),
    factor = defaultFactorChecks(add = "identifyColons"),
    labelled = defaultLabelledChecks(add = "identifyColons")
  ),

  #overwrite old versions of the report, render to html and don't 
  #open the html file automatically:
  replace = TRUE, 
  output = "html",
  open = FALSE
)
```

We have chosen the output to be html (`output = "html"`) so that it can easily be included in this vignette by use of the `includeHTML()` function from the `htmltools` package. We have set `open = FALSE` so that the outputted html file is not opened automatically. The outputted report is available at the end of this vignette.

# A dataMaid report with user-defined functions: Documenting artData

We include the report documenting `artData` (this is only shown if `pandoc` is available to generate it):

```{r includeReport, eval=rmarkdown::pandoc_available()}
htmltools::includeHTML("dataMaid_artData.html")
```

### 2020-07-27 R/messageGenerator.R

@@ -136,17 +136,41 @@ printProblemValues <- function(problemValues, nMax = Inf) {
         ##NOTE: sort removes NaNs if not told explicitly not to by use of the na.last-argument
     nVals <- length(problemValues)
     extraStr <- ""
+    
     if (nMax < nVals) {
       if (nMax == 0) {
-        problemValues <- ""
-        extraStr <- paste(nVals-nMax, "problematic value(s) omitted")
-      } else {
-        problemValues <- problemValues[1:nMax]
-        extraStr <- paste(" (", nVals-nMax, " additional values omitted)", sep="")
+       outText <- paste(nVals-nMax, "problematic value(s) omitted")
       }
+      if (nMax == 1) {
+        outText <- paste("\\\"", problemValues[1], "\\\"",
+                         " (", nVals-nMax, " additional values omitted)", sep = "")
+      }
+      if (nMax > 1) {
+        nLeft <- ceiling(nMax / 2)
+        nRight <- floor(nMax / 2)
+        strLeft <- paste0(paste0("\\\"", problemValues[1:nLeft], "\\\""),
+                         collapse=", ")
+        strRight <- paste0(paste0("\\\"", problemValues[(nVals-nRight+1):nVals], "\\\""),collapse=", ")
+        extraStr <- paste(" (", nVals-nMax, " values omitted)", sep="")
+        outText <- paste0(strLeft, ", ..., ",strRight, extraStr)
+      }
+    } else { #if no omission of probvals
+      outText <- paste(paste("\\\"", problemValues, "\\\"", sep=""), collapse = ", ")
     }
-    paste(paste(paste("\\\"", problemValues, "\\\"", sep=""),
-                collapse=", "), extraStr, sep="")
+    
+    outText
+    
+  ##old version that prints the first problemvalues only:
+  #  if (nMax < nVals) {
+  #    if (nMax == 0) {
+  #      
+  #    } else {
+  #      problemValues <- problemValues[1:nMax]
+  #      extraStr <- paste(" (", nVals-nMax, " additional values omitted)", sep="")
+  #    }
+  #  }
+  #  paste(paste(paste("\\\"", problemValues, "\\\"", sep=""),
+  #              collapse=", "), extraStr, sep="")
 }
 
 #Does character escaping such that the text can be rendered by rmarkdown

### 2021-10-07 docs/CSP2018/slides/module3/module3.Rmd
Update to keep on CRAN

---
title: "Cleaning Up the Data Cleaning Process"
subtitle: "Module 3: Row-wise constraints"
author: "Claus Thorn Ekstrøm and Anne Helby Petersen<br>UCPH Biostatistics"
date: "CSP, Portland, OR, Feb. 15th, 2018<br>.footnotesize[[Slides/homepage](https://ekstroem.github.io/dataMaid/CSP2018/index.html)]"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: ['default', 'ku.css', 'animate.min.css']
    seal: true
    nature:
      ratio: '16:9'
      slideNumberFormat: '%current%' 
      highlightStyle: ir-black
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
#library(knitr)
options(htmltools.dir.version = FALSE, cache=TRUE)
#opts_chunk$set(dev.args=list(bg="transparent"), fig.width=15, fig.height=7)
source("kutheme.R")

library(dataMaid)
library(assertr)
library(dplyr)
library(validate)
data(bigPresidentData)
bpD <- bigPresidentData
```

##Row-wise or column-wise checks?

```{r colrow1, echo=FALSE, out.width="70%", fig.align = "center"}
knitr::include_graphics("pics/colrow1.png")
```

---

##Row-wise or column-wise checks?

```{r colrow2, echo=FALSE, out.width="70%", fig.align = "center"}
knitr::include_graphics("pics/colrow2.png")
```

---

##Row-wise or column-wise checks?

```{r colrow3, echo=FALSE, out.width="70%", fig.align = "center"}
knitr::include_graphics("pics/colrow3.png")
```

---

##Row-wise *and* column-wise constraints!

* `dataMaid` performs class dependent checks for each variable in a dataset, one at a time (column-wise)
  + Pros: Easy to document what was (not) done, let's you get started without a lot of prior knowledge, easy to share with collaborators
  + Shortcomings: Generally cannot detect internal consistency issues or use non-class dependent variable constraints

---

##Row-wise *and* column-wise constraints!

An R-packages that performs row-wise checks: `validate`

```{r}
library(validate)
```

Note: Different use of the term "validation" - no longer about format, type and range, but used as synonym to "check".

---

##`validate` - overview:

* Splits checking into two steps:
  1. Define the checking rules in a `validator` object, using the `valdiator()` function.
  2. Confront the data with the rules in a call to `confront`, thereby obtaining a `confrontation` object.
* Easy to asses what problems were found and to document what was checked (saved in `validator` object). 
*  <span style="color:red">Beware: Non-standard object storage might cause trouble!</span>

---

###Make `validator` object

.footnotesize[
```{r}
val1 <- 
 validator(
   ageAtDeath := floor((dateOfDeath - birthday)/365.25),  
   `Adult president` = ageAtInauguration >= 18,
   `Alive at inauguration` = ageAtDeath >= ageAtInauguration,
   `Positive first name` = firstName*2 > firstName,
   `Death by assassination` = 
     if (dateOfDeath == presidencyEndDate) 
       assassinationAttempt == 1
)
``` 
]

---

###Confront data with `validator` object:
.footnotesize[
```{r}
con1 <- confront(bpD, val1)
summary(con1)[, 1:6]
``` 
]

---

##Understand confrontation results

Lots of functions available for inspecting confrontations:
* `summary()`: Overview of confrontation results
* `aggregate()`: Compute percentage pass/fail/na 
* `sort()`: Sort results by problem prevalence
* `values()`: For each observation and each check: `TRUE`/`FALSE`/`NA`
* `barplot()`: Visual overview of check results
* `errors()`: What errors were caught?
* `warnings()`: What warnings were caught?

---

##Warning: modify-by-reference

* `validator` and `confrontation` objects imply use of "modify-by-reference" rather than "copy-on-modify" semantics (standard R)
* Other package that uses "modify-by-reference": `data.table`
* If one is not aware of modify-by-reference, it can have problematic consequences 
  + Unintentional overwriting of full objects, attributes or parts
  + Caching in rmarkdown does not work (but throws no errors)

---

##Warning: modify-by-reference

```{r}
v1 <- validator(check1 = sex == "Male")
v1

v2 <- v1  
names(v2) <- "All males"
v1
```

---

##Warning: modify-by-reference

Make a copy using `[TRUE]`:
```{r}
v1 <- validator(check1 = sex == "Male")
v2 <- v1[TRUE]  
names(v2) <- "All males"
v1
```

### 2021-10-07 docs/CSP2018/slides/module4/module4.Rmd
Update to keep on CRAN

--- 2018-05-10 docs/eRum2018/slides/module3/pics/module4.Rmd
@@
 ---
-title: "Reproducible data screening and quality control"
-subtitle: "Module 3: Time to clean and other options"
+title: "Cleaning Up the Data Cleaning Process"
+subtitle: "Module 4: Time to clean!"
 author: "Claus Thorn Ekstrøm and Anne Helby Petersen<br>UCPH Biostatistics"
-date: "eRum, Budapest, May 14th, 2018<br>.footnotesize[[Slides/homepage](https://ekstroem.github.io/dataMaid/eRum2018/index.html)]"
+date: "CSP, Portland, OR, Feb. 15th, 2018<br>.footnotesize[[Slides/homepage](https://ekstroem.github.io/dataMaid/CSP2018/index.html)]"
 output:
@@
 
-.center[
 ```{r cleanpic, echo=FALSE, out.width="60%"}
@@
 ```
-]
 
-Not the best term ... and should not be unsupervised
+.pull-right[Not the best term ... and should not be unsupervised]
 
@@
 Two systems for selecting observations in `data.frame`s in R: 
-
 By index (row number) or using a logical vector.
@@
 
+class: inverse
+
+# Exercise 4
+
+Correct the errors you have found so far.
+
+Make sure to make the cleaning process reproducible.
+
+Remember **rules 1 and 2**!
+
+---
+
 background-image: url(pics/structure.png)
 background-size: 30%
-background-position: right
 
-## Finishing up after cleaning
+## Finishing up
 
@@
 ```{r eval=FALSE}
-makeCodebook(presidentData)
+makeCodebook(bigPresidentData)
 ```
@@
 ```{r}
-pD <- presidentData
-attr(pD$presidencyYears, "label") <- 
+bPD <- bigPresidentData
+attr(bPD$presidencyYears, "label") <- 
   "Full years as president"
@@
 ```{r}
-attr(pD$birthday, "shortDescription") <- 
-  "Dates are stored in YYYY-MM-DD format"
+attr(bPD$dateOfDeath, "shortDescription") <- 
+  "Missing means that the person is still alive"
 ```
@@
 
-# Exercise 3
-
-Correct the errors you have found so far.
-
-Make sure to make the cleaning process reproducible.
-
-Remember **rules 1 and 2**!
+# Exercise 4b
 
@@
 ```
-
----
-
-##Row-wise or column-wise checks?
-
-```{r colrow1, echo=FALSE, out.width="75%", fig.align = "center"}
-knitr::include_graphics("pics/colrow1.png")
-```
-
----
-
-##Row-wise or column-wise checks?
-
-```{r colrow2, echo=FALSE, out.width="75%", fig.align = "center"}
-knitr::include_graphics("pics/colrow2.png")
-```
-
----
-
-##Row-wise or column-wise checks?
-
-```{r colrow3, echo=FALSE, out.width="75%", fig.align = "center"}
-knitr::include_graphics("pics/colrow3.png")
-```
-
----
-
-##Row-wise *and* column-wise constraints!
-
-* `dataMaid` performs class dependent checks for each variable in a dataset, one at a time (column-wise)
-  + Pros: Easy to document what was (not) done, let's you get started without a lot of prior knowledge, easy to share with collaborators
-  + Shortcomings: Generally cannot detect internal consistency issues or use non-class dependent variable constraints
-
----
-
-##Row-wise *and* column-wise constraints!
-
-An R-packages that performs row-wise checks: `validate`. Check out the talk on Wednesday @ 14.50 by Edwin de Jonge:
-
-**validatetools - resolve and simplify contradictive or redundant data validation rules**
-
-Note: Different use of the term "validation" - no longer about format, type and range, but used as synonym to "check".
 

### 2021-10-07 docs/articles/dataMaid_artData.Rmd
Update to keep on CRAN

@@
 subtitle: "Autogenerated data summary from dataMaid"
-date: 2018-02-09 14:33:06
+date: 2018-02-10 15:57:23
 output: html_document
@@
 
- *  Report creation time: Fri Feb 09 2018 14:33:06
+ *  Report creation time: Sat Feb 10 2018 15:57:23
 

### 2021-10-07 latex/article_accept.tex
Update to keep on CRAN

@@
 issues, if any. For instance, as shown in Figure~\ref{fig:example2},
-for the \code{numeric}-type variable \code{events} from
+for the numeric-type variable \code{events} from
 \code{toyData}, \code{makeDataReport()} has identified two values that
@@
 %Takes a list on the format \code{list(X = Y)} where \code{X} is a non-supported variable class (e.g.
-\code{complex}) and \code{Y} is a supported one (e.g.
-\code{numeric}) & \code{NULL} (No rules for handling non-supported variable types) \\
+complex) and \code{Y} is a supported one (e.g.
+numeric) & \code{NULL} (No rules for handling non-supported variable types) \\
 
@@
                        automatically open afterwards, and if not,
-                       should the \code{rmarkdown} file be opened? & \code{TRUE} \\
+                       should the \proglang{R} markdown file be opened? & \code{TRUE} \\
 \quad \code{replace} & Overwrite an existing file with the same name? & \code{FALSE} \\
@@
 \caption{Overview of all summary-, visual- and  check functions currently implemented in \pkg{dataMaid}. The variable
-  classes C, F, I, L, B, N, and D, refer to \code{character}, \code{factor},
-  \code{integer}, \code{labelled}, \code{logical} (boolean), \code{numeric}, and \code{Date}, respectively. The default settings of \code{makeDataReport()} are marked in blue.}
+  classes C, F, I, L, B, N, and D, refer to character, factor,
+  integer, labelled, logical (boolean), numeric, and Date, respectively. The default settings of \code{makeDataReport()} are marked in blue.}
 \label{table.SVCfunctions}
@@
 
-Thus we can see, for example, that for \code{numeric}, \code{integer},
-and \code{Date} variables, \pkg{dataMaid} provides functions for
+Thus we can see, for example, that for numeric, integer,
+and Date variables, \pkg{dataMaid} provides functions for
 adding summary information about the minimum and maximum values, while
@@
 
-One might also encounter datasets with variables that are not among the 7 classes mentioned in the above (\code{character}, \code{Date}, \code{factor}, \code{integer}, \code{labelled}, \code{logical} and \code{numeric}), for instance variables of type \code{complex} or user-defined classes.
+One might also encounter datasets with variables that are not among the 7 classes mentioned in the above (character, Date, factor, integer, labelled, logical and numeric), for instance variables of type complex or user-defined classes.
 \code{makeDataReport()} can be told how to handle such variables by use of the argument \code{treatXasY}.
 This argument takes a list where the names correspond to "new" variable types (\code{X}), while the entries must be supported variable types (\code{Y}).
-For instance, we can instruct \code{dataMaid} to treat complex variables as numerics and generate a data report for a type \code{complex} variable like this:
+For instance, we can instruct \code{dataMaid} to treat complex variables as numeric and generate a data report for a type complex variable like this:
 
@@
 
-In this report, we will find that the two variables, \code{complexVar} and \code{numericVar} will be have identical presentations in the variable list, as treating a \code{complex} variable as a \code{numeric} means dropping the imaginary part of the complex numbers which was the only thing setting the two variables apart in the first place.
+In this report, we will find that the two variables, \code{complexVar} and \code{numericVar} will be have identical presentations in the variable list, as treating a complex variable as a numeric means dropping the imaginary part of the complex numbers which was the only thing setting the two variables apart in the first place.
 
@@
 
-The syntax in this code chunk can be read as follows: "Summarize \code{events} in \code{toyData}, and for \code{numeric} variables, set the summaries to be the default summary functions, except \code{variableType} and \code{countMissing}."
+The syntax in this code chunk can be read as follows: "Summarize \code{events} in \code{toyData}, and for numeric variables, set the summaries to be the default summary functions, except \code{variableType} and \code{countMissing}."
 
@@
 \caption{The front page and the first page of the data overview report for the \code{presidentData} dataset.
-Note that the report title has been customized (front page), \code{identifyLoners} has been removed from the checks performed on character variables ("Identify levels with <6 obs." is not checked for character variables in the table on page 1) and that variables of class \code{Name} have been set to be treated like \code{character} variables (page 1).
+Note that the report title has been customized (front page), \code{identifyLoners} has been removed from the checks performed on character variables ("Identify levels with <6 obs." is not checked for character variables in the table on page 1) and that variables of class \code{Name} have been set to be treated like character variables (page 1).
 Larger versions of the pages can be seen in
@@
 \caption{The second and third pages of the \code{presidentData} data report.
-We see that there are two \code{Name} variables in the overview on page 2 and see that these variables are indeed treated as \code{character} variables on page 3, as specified in the \code{makeDataReport} call by use of the \code{treatXasY} argument.
+We see that there are two \code{Name} variables in the overview on page 2 and see that these variables are indeed treated as character variables on page 3, as specified in the \code{makeDataReport} call by use of the \code{treatXasY} argument.
 Larger versions of the pages can be seen in
@@
 This can be done using the \code{treatXasY} argument.
-\item We use the \code{character} class for a few categorical variables that have a lot of different levels, but where it is not a data mistake that these levels each generally have very few observations, e.g., the variable \code{stateOfBirth}.
+\item We use the character class for a few categorical variables that have a lot of different levels, but where it is not a data mistake that these levels each generally have very few observations, e.g., the variable \code{stateOfBirth}.
 Therefore, we would like to disable the \code{identifyLoners} check (which flags variables with <6 observations in any of the levels) for this variable type.
@@
 \end{enumerate}
-Incorporating these three customizations, we can load the data and generate a report by calling:
+We incorporate these three customizations, load the data, and generate a report by calling:
 
@@
 On page 3, we see the contents for the three first variables.
-Here, we identify a prefixed white space in the last name entry for President Truman and we find that a dot was entered as a first name; this is a typical choice for coding missing values in e.g., \proglang{Stata}, and therefore, it is flagged as a potential miscoded missing value.
+Here, we identify a prefixed white space in the last name entry for President Truman and we find that a dot was entered as a first name; this is a typical choice for coding missing values in e.g., \proglang{Stata} \citep{stata}, and therefore, it is flagged as a potential miscoded missing value.
 The variable \code{orderOfPresidency} is not summarized, visualized or checked because it is categorical and contains unique values for each observation.
@@
 \item It seems as if one or more presidents have complex numbers as their favorite numbers.
-As \code{complex} is not a supported variable type in \code{dataMaid} and no strategy for handling this class was provided in the \code{treatXasY} argument, the variable is simply flagged as non-supported.
+As complex is not a supported variable type in \code{dataMaid} and no strategy for handling this class was provided in the \code{treatXasY} argument, the variable is simply flagged as non-supported.
 \end{itemize}
@@
 
-Please note that if \code{ageAtInauguration} had been a \code{factor} rather than a \code{character} variable, an additional inner call should be added in order to ensure no conversion issues:
+Please note that if \code{ageAtInauguration} had been a factor rather than a character variable, an additional inner call should be added in order to ensure no conversion issues:
 
@@
 Finally, we create a new data report, adding the suffix "cleaned" to the title, as well as the file name, so that we have documentation of the current state of the dataset.
-We also decide that it might be sufficient to be able to inspect only the real part of the presidential favorite numbers and therefore, we choose to treat the \code{complex} variable \code{favoriteNumber} as a \code{numeric} variable:
+We also decide that it might be sufficient to inspect only the real part of the presidential favorite numbers and therefore, we choose to treat the complex variable \code{favoriteNumber} as a numeric variable:
 
@@
 \end{enumerate}
-Please note that the data report does contain information about who, when and how concerning its generation, so even though the default choices for file names do not make it easy to tell different reports for the same dataset apart, it should be rather easy when inspecting the report manually.
+Note that the data report does contain information about who, when and how concerning its generation, so even though the default choices for file names do not make it easy to tell different reports for the same dataset apart, it should be rather easy when inspecting the report manually.
 
@@
 An even more minimal output  can be obtained directly in the console by using the \code{check()} function interactively.
-When called on a \code{data.frame}, this function produces a list (of
+When called on a data frame, this function produces a list (of
 variables) of lists (of checks) of lists (or rather,
@@
 
-{\small
 \begin{Verbatim}
@@
   - \renewcommand{\chaptername}{Part}
-  - \newcommand{\fullline}{\noindent\makebox[\linewidth]{\rule{\textwidth}{0.4pt}}}
-  - \newcommand{\bminione}{\begin{minipage}{0.75 \textwidth}}
-  - \newcommand{\bminitwo}{\begin{minipage}{0.25 \textwidth}}
+  - \newcommand{\fullline}{\noindent\rule{\textwidth}{0.4pt}}
+  - \newcommand{\bminione}{\begin{minipage}{0.75\textwidth}}
+  - \newcommand{\bminitwo}{\begin{minipage}{0.25\textwidth}}
   - \newcommand{\emini}{\end{minipage}}
@@
 \end{Verbatim}
-}
-
-In the this example, the \code{dataMaid\_toyData.Rmd} file could have
+
+In this example, the \code{dataMaid\_toyData.Rmd} file could have
 been created as follows:
@@
 Note that with the latter option, no special YAML preamble is needed in the \proglang{R} markdown document. 
-
-Alternatively, one can create the usual output report, not render it and then manually edit the produced \proglang{R} markdown file as wished. This can be done with the following command:
+%
+Alternatively, one can create the usual output report, not render it and then manually edit the produced \proglang{R} markdown file as wished. The following command does that:
 

### 2021-10-07 latex/dataMaid_presidentData.Rmd
Update to keep on CRAN

---
dataMaid: yes
title: Dirty president data
subtitle: "Autogenerated data summary from dataMaid"
date: 2018-04-11 01:02:26
output: pdf_document
documentclass: report
header-includes:
  - \renewcommand{\chaptername}{Part}
  - \newcommand{\fullline}{\noindent\makebox[\linewidth]{\rule{\textwidth}{0.4pt}}}
  - \newcommand{\bminione}{\begin{minipage}{0.75 \textwidth}}
  - \newcommand{\bminitwo}{\begin{minipage}{0.25 \textwidth}}
  - \newcommand{\emini}{\end{minipage}}
---
```{r , echo=FALSE, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
library(ggplot2)
library(pander)
```
```{r visualFunctions, echo=FALSE, include=FALSE, warning=FALSE, message=FALSE, error=FALSE}
ggAggHist <- getFromNamespace("ggAggHist", "dataMaid")
ggAggBarplot <- getFromNamespace("ggAggBarplot", "dataMaid")
```
# Data report overview
The dataset examined has the following dimensions:

---------------------------------
Feature                    Result
------------------------ --------
Number of observations         46

Number of variables            11
---------------------------------

### Checks performed
The following variable checks were performed, depending on the data type of each variable:

----------------------------------------------------------------------------------------------------------------------------------
&nbsp;                                                 character    factor    labelled   numeric    integer    logical     Date   
----------------------------------------------------- ----------- ---------- ---------- ---------- ---------- --------- ----------
Identify miscoded missing values                       $\times$    $\times$   $\times$   $\times$   $\times$             $\times$

Identify prefixed and suffixed whitespace              $\times$    $\times$   $\times$

Identify case issues                                   $\times$    $\times$   $\times$

Identify misclassified numeric or integer variables    $\times$    $\times$   $\times$

Identify levels with < 6 obs.                                      $\times$   $\times$

Identify outliers                                                                        $\times$   $\times$             $\times$ 
----------------------------------------------------------------------------------------------------------------------------------

Non-supported variable types were set to be handled in the following way:

* Name is treated as character

Please note that all numerical values in the following have been rounded to 2 decimals.

# Summary table

--------------------------------------------------------------------------------------------------
&nbsp;                   Variable class     # unique values   Missing observations  Any problems? 
------------------------ ---------------- ----------------- ---------------------- ---------------
[lastName]               Name                            40                 0.00 %    $\times$

[firstName]              Name                            31                 0.00 %    $\times$

[orderOfPresidency]      factor                          46                 0.00 %    $\times$

[birthday]               Date                            45                 0.00 %    $\times$

[stateOfBirth]           character                       23                 0.00 %    $\times$

[assassinationAttempt]   numeric                          2                 0.00 %

[sex]                    factor                           1                 0.00 %    $\times$

[ethnicity]              factor                           2                 0.00 %    $\times$

[presidencyYears]        numeric                         11                 4.35 %    $\times$

[ageAtInauguration]      character                       23                 0.00 %    $\times$

[favoriteNumber]         complex                         11                 0.00 %    $\times$    
--------------------------------------------------------------------------------------------------

# Variable list
## lastName

\bminione

-----------------------------------
Feature                      Result
------------------------- ---------
Variable type                  Name

Number of missing obs.      0 (0 %)

Number of unique values          40

Mode                        "Adams"
-----------------------------------

\emini
\bminitwo
```{r Var-1-lastName, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:40, .Label = c(" Truman", 
"Adams", "Arathornson", "Arthur", "Buchanan", "Bush", "Carter", 
"Cleveland", "Clinton", "Coolidge", "Eisenhower", "Fillmore", 
"Ford", "Garfield", "Grant", "Harding", "Harrison", "Hayes", 
"Hoover", "Jackson", "Jefferson", "Johnson", "Kennedy", "Lincoln", 
"Madison", "McKinley", "Monroe", "Nixon", "Obama", "Pierce", 
"Polk", "Reagan", "Roosevelt", "Taft", "Taylor", "Trump", "Tyler", 
"Van Buren", "Washington", "Wilson"), class = "factor"), y = c(1L, 
2L, 1L, 1L, 1L, 2L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 
1L, 1L, 1L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 
1L, 1L, 1L, 1L, 1L, 1L, 1L)), .Names = c("x", "y"), row.names = c(NA, 
-40L), class = "data.frame"), vnam = "lastName")
```
\emini

- The following values appear with prefixed or suffixed white space: \" Truman\".

\fullline

## firstName

\bminione

-----------------------------------
Feature                      Result
------------------------- ---------
Variable type                  Name

Number of missing obs.      0 (0 %)

Number of unique values          31

Mode                        "James"
-----------------------------------

\emini
\bminitwo
```{r Var-2-firstName, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:31, .Label = c(".", 
"Abraham", "Andrew", "Aragorn", "Barack", "Benjamin", "Calvin", 
"Chester", "Dwight", "Franklin", "George", "Gerald", "Grover", 
"Harry", "Herbert", "James", "Jimmy", "John", "Lyndon", "Martin", 
"Millard", "Richard", "Ronald", "Rutherford", "Theodore", "Thomas", 
"Ulysses", "Warren", "William", "Woodrow", "Zachary"), class = "factor"), 
    y = c(1L, 1L, 2L, 1L, 1L, 1L, 1L, 1L, 1L, 2L, 3L, 1L, 2L, 
    1L, 1L, 5L, 1L, 4L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
    4L, 1L, 1L)), .Names = c("x", "y"), row.names = c(NA, -31L
), class = "data.frame"), vnam = "firstName")
```
\emini

- The following suspected missing value codes enter as regular values: \".\".

\fullline

## orderOfPresidency

* The variable is a key (distinct values for each observation).

\fullline

## birthday

\bminione

--------------------------------------------------
Feature                                     Result
------------------------- ------------------------
Variable type                                 Date

Number of missing obs.                     0 (0 %)

Number of unique values                         45

Mode                                  "1837-03-18"

Min. and max.               1300-03-01; 1961-08-04

1st and 3rd quartiles       1790-03-29; 1890-10-14
--------------------------------------------------

\emini
\bminitwo
```{r Var-4-birthday, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggHist(data = structure(list(factorV = structure(1:20, .Label = c("1300-03-01", 
"1333-03-27", "1366-04-22", "1399-05-18", "1432-06-13", "1465-07-09", 
"1498-08-04", "1531-08-31", "1564-09-25", "1597-10-21", "1630-11-16", 
"1663-12-13", "1697-01-07", "1730-02-03", "1763-03-01", "1796-03-26", 
"1829-04-22", "1862-05-18", "1895-06-13", "1928-07-09"), class = "factor"), 
    Freq = c(1L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 
    0L, 5L, 8L, 6L, 9L, 6L, 7L, 4L), xmin = structure(c(-244653, 
    -232574, -220495, -208416, -196337, -184258, -172179, -160100, 
    -148021, -135942, -123863, -111783, -99704, -87625, -75546, 
    -63467, -51388, -39309, -27230, -15151), class = "Date"), 
    xmax = structure(c(-232574, -220495, -208416, -196337, -184258, 
    -172179, -160100, -148021, -135942, -123863, -111783, -99704, 
    -87625, -75546, -63467, -51388, -39309, -27230, -15151, -3072
    ), class = "Date"), ymin = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 
    0, 0, 0, 0, 0, 0, 0, 0, 0, 0), ymax = c(1L, 0L, 0L, 0L, 0L, 
    0L, 0L, 0L, 0L, 0L, 0L, 0L, 0L, 5L, 8L, 6L, 9L, 6L, 7L, 4L
    )), .Names = c("factorV", "Freq", "xmin", "xmax", "ymin", 
"ymax"), row.names = c(NA, -20L), class = "data.frame"), vnam = "birthday")
```
\emini

- Note that the following possible outlier values were detected: \"1300-03-01\".

\fullline

## stateOfBirth

\bminione

-------------------------------------
Feature                        Result
------------------------- -----------
Variable type               character

Number of missing obs.        0 (0 %)

Number of unique values            23

Mode                           "Ohio"
-------------------------------------

\emini
\bminitwo
```{r Var-5-stateOfBirth, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:23, .Label = c("Arkansas", 
"California", "Connecticut", "Georgia", "Gondor", "Hawaii", "Illinois", 
"Iowa", "Kentucky", "Massachusetts", "Missouri", "Nebraska", 
"New Hampshire", "New Jersey", "New York", "New york", "North Carolina", 
"Ohio", "Pennsylvania", "Tennessee", "Texas", "Vermont", "Virginia"
), class = "factor"), y = c(1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 1L, 
4L, 1L, 1L, 1L, 2L, 4L, 1L, 2L, 8L, 1L, 1L, 2L, 2L, 7L)), .Names = c("x", 
"y"), row.names = c(NA, -23L), class = "data.frame"), vnam = "stateOfBirth")
```
\emini

- Note that there might be case problems with the following levels: \"New York\", \"New york\".

\fullline

## assassinationAttempt

* Note that this variable is treated as a factor variable below, as it only takes a few unique values.

\bminione

-----------------------------------
Feature                      Result
------------------------- ---------
Variable type               numeric

Number of missing obs.      0 (0 %)

Number of unique values           2

Mode                            "0"
-----------------------------------

\emini
\bminitwo
```{r Var-6-assassinationAttempt, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:2, .Label = c("0", 
"1"), class = "factor"), y = c(35L, 11L)), .Names = c("x", "y"
), row.names = c(NA, -2L), class = "data.frame"), vnam = "assassinationAttempt")
```
\emini

\fullline

## sex

* The variable only takes one (non-missing) value: \"Male\". The variable contains 0 \% missing observations.

\fullline

## ethnicity

\bminione

---------------------------------------
Feature                          Result
------------------------- -------------
Variable type                    factor

Number of missing obs.          0 (0 %)

Number of unique values               2

Mode                        "Caucasian"
---------------------------------------

\emini
\bminitwo
```{r Var-8-ethnicity, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:2, .Label = c("African American", 
"Caucasian"), class = "factor"), y = c(1L, 45L)), .Names = c("x", 
"y"), row.names = c(NA, -2L), class = "data.frame"), vnam = "ethnicity")
```
\emini

- Note that the following levels have at most five observations: \"African American\".

\fullline

## presidencyYears

\bminione

--------------------------------------
Feature                         Result
------------------------- ------------
Variable type                  numeric

Number of missing obs.      2 (4.35 %)

Number of unique values             10

Median                               4

1st and 3rd quartiles          3.75; 8

Min. and max.                   0; Inf
--------------------------------------

\emini
\bminitwo
```{r Var-9-presidencyYears, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggHist(data = structure(list(factorV = structure(1:20, .Label = c("[0,0.6]", 
"(0.6,1.2]", "(1.2,1.8]", "(1.8,2.4]", "(2.4,3]", "(3,3.6]", 
"(3.6,4.2]", "(4.2,4.8]", "(4.8,5.4]", "(5.4,6]", "(6,6.6]", 
"(6.6,7.2]", "(7.2,7.8]", "(7.8,8.4]", "(8.4,9]", "(9,9.6]", 
"(9.6,10.2]", "(10.2,10.8]", "(10.8,11.4]", "(11.4,12]"), class = "factor"), 
    Freq = c(2L, 1L, 0L, 4L, 4L, 0L, 15L, 0L, 3L, 0L, 0L, 3L, 
    0L, 10L, 0L, 0L, 0L, 0L, 0L, 1L), xmin = c(0, 0.6, 1.2, 1.8, 
    2.4, 3, 3.6, 4.2, 4.8, 5.4, 6, 6.6, 7.2, 7.8, 8.4, 9, 9.6, 
    10.2, 10.8, 11.4), xmax = c(0.6, 1.2, 1.8, 2.4, 3, 3.6, 4.2, 
    4.8, 5.4, 6, 6.6, 7.2, 7.8, 8.4, 9, 9.6, 10.2, 10.8, 11.4, 
    12), ymin = c(0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 
    0, 0, 0, 0, 0), ymax = c(2L, 1L, 0L, 4L, 4L, 0L, 15L, 0L, 
    3L, 0L, 0L, 3L, 0L, 10L, 0L, 0L, 0L, 0L, 0L, 1L)), .Names = c("factorV", 
"Freq", "xmin", "xmax", "ymin", "ymax"), row.names = c(NA, -20L
), class = "data.frame"), vnam = "presidencyYears")
```
\emini

- The following suspected missing value codes enter as regular values: \"Inf\".

- Note that the following possible outlier values were detected: \"0\", \"1\", \"Inf\".

\fullline

## ageAtInauguration

\bminione

-------------------------------------
Feature                        Result
------------------------- -----------
Variable type               character

Number of missing obs.        0 (0 %)

Number of unique values            23

Mode                             "54"
-------------------------------------

\emini
\bminitwo
```{r Var-10-ageAtInauguration, echo=FALSE, fig.width=4, fig.height=3, message=FALSE, warning=FALSE}
ggAggBarplot(data = structure(list(x = structure(1:23, .Label = c("42", 
"43", "46", "47", "48", "49", "50", "51", "52", "54", "55", "56", 
"57", "58", "60", "61", "62", "64", "65", "68", "69", "70", "87"
), class = "factor"), y = c(1L, 1L, 2L, 2L, 1L, 2L, 2L, 4L, 2L, 
5L, 4L, 3L, 4L, 1L, 1L, 3L, 1L, 2L, 1L, 1L, 1L, 1L, 1L)), .Names = c("x", 
"y"), row.names = c(NA, -23L), class = "data.frame"), vnam = "ageAtInauguration")
```
\emini

- Note: The variable consists exclusively of numbers and takes a lot of different values. Is it perhaps a misclassified numeric variable?

\fullline

## favoriteNumber

* The variable has class complex which is not supported by dataMaid.

\fullline

Report generation information:

 *  Created by Claus Thorn Ekstrøm (username: `cld189`).

 *  Report creation time: Wed Apr 11 2018 01:02:26

 *  Report Was run from directory: `/Users/cld189/ku/R/cleanR/latex`

 *  dataMaid v1.1.0 [Pkg: 2018-03-02 from local (ekstroem/dataMaid@NA)]

 *  R version 3.4.4 (2018-03-15).

 *  Platform: x86_64-apple-darwin15.6.0 (64-bit)(macOS High Sierra 10.13.3).

 *  Function call: `makeDataReport(data = presidentData, replace = TRUE, checks = setChecks(character = defaultCharacterChecks(remove = "identifyLoners")), 
    reportTitle = "Dirty president data", treatXasY = list(Name = "character"))`

### 2021-10-07 latex/dataMaid_toyData.Rmd
Update to keep on CRAN

@@
 subtitle: "Autogenerated data summary from dataMaid"
-date: 2017-11-14 13:41:55
+date: 2018-04-11 01:02:39
 output: pdf_document
@@
 ----------------------------------------------------- ----------- ---------- ---------- ---------- ---------- --------- ----------
-Identify miscoded missing values                       $\times$    $\times$   $\times$   $\times$   $\times$
+Identify miscoded missing values                       $\times$    $\times$   $\times$   $\times$   $\times$             $\times$
 
@@
 
- *  Created by Claus Thorn Ekstrøm.
+ *  Created by Claus Thorn Ekstrøm (username: `cld189`).
 
- *  Report creation time: Tue Nov 14 2017 13:41:55
+ *  Report creation time: Wed Apr 11 2018 01:02:39
 
- *  dataMaid v1.0.1 [Pkg: 2017-11-13 from local (ekstroem/dataMaid@NA)]
+ *  Report Was run from directory: `/Users/cld189/ku/R/cleanR/latex`
 
- *  R version 3.4.2 (2017-09-28).
+ *  dataMaid v1.1.0 [Pkg: 2018-03-02 from local (ekstroem/dataMaid@NA)]
 
- *  Platform: x86_64-apple-darwin15.6.0 (64-bit)(macOS High Sierra 10.13.1).
+ *  R version 3.4.4 (2018-03-15).
+
+ *  Platform: x86_64-apple-darwin15.6.0 (64-bit)(macOS High Sierra 10.13.3).
 

### 2021-10-07 R/identifyMissing.R

@@ -95,6 +95,7 @@ identifyMissingMessage <- "The following suspected missing value codes enter as 
 #removing the prefix. If e.g. prefix = "-" and char = 8,
 #values like -88, -8, -8888 will be identified (if present).
 identifyMissRepChar <- function(v, char, prefix=NULL, ignoreFirst = FALSE) {
+  originalClass <- class(v)
   v <- as.character(v)
   char <- as.character(char)
   if (!is.null(prefix)) {
@@ -111,6 +112,7 @@ identifyMissRepChar <- function(v, char, prefix=NULL, ignoreFirst = FALSE) {
       vals <- paste(prefix, v[problems], sep="")
     } else vals <- v[problems]
     vals <- unique(vals)
+    class(vals) <- originalClass
   } else vals <- NULL
   vals
 }

### 2025-04-13 R/check.R
Update for CRAN

@@ -49,7 +49,7 @@
 #' functions and functions imported from \code{dataMaid} or other packages) can
 #' be viewed by calling \code{allCheckFunctions()}.
 #'
-#' @references Petersen AH, Ekstrøm CT (2019). “dataMaid: Your Assistant for Documenting Supervised Data Quality Screening in R.” _Journal of Statistical Software_, *90*(6), 1-38. doi: 10.18637/jss.v090.i06 ( \url{https://doi.org/10.18637/jss.v090.i06}).
+#' @references Petersen AH, Ekstrøm CT (2019). “dataMaid: Your Assistant for Documenting Supervised Data Quality Screening in R.” _Journal of Statistical Software_, *90*(6), 1-38. doi: 10.18637/jss.v090.i06 ( \doi{10.18637/jss.v090.i06}).
 #'
 #' @seealso \code{\link{setChecks}}, 
 #' \code{\link{allCheckFunctions}} \code{\link{checkResult}}

### 2025-04-13 R/dataMaid-package.R

@@ -24,9 +24,9 @@
 #'    \item{favoriteNumber}{A \code{complex} type variable with a fictional favorite number for
 #'    each president.}
 #' }
-#' @references Petersen AH, Ekstrøm CT (2019). “dataMaid: Your Assistant for Documenting Supervised Data Quality Screening in R.” _Journal of Statistical Software_, *90*(6), 1-38. doi: 10.18637/jss.v090.i06 ( \url{https://doi.org/10.18637/jss.v090.i06}).
+#' @references Petersen AH, Ekstrøm CT (2019). “dataMaid: Your Assistant for Documenting Supervised Data Quality Screening in R.” _Journal of Statistical Software_, *90*(6), 1-38. doi: 10.18637/jss.v090.i06 ( \doi{10.18637/jss.v090.i06}).
 #' @source Artificial dataset constructed based on the US president dataset available from 
-#' \href{http://www.data-explorer.com/data}{Data Explorer}.
+#' \href{https://www.data-explorer.com/data/}{Data Explorer}.
 #' @examples
 #' data(presidentData)
 #'
@@ -67,8 +67,8 @@
 #'    each president.}
 #' }
 #' @source Artificial dataset constructed based on the US president dataset available from 
-#' \href{http://www.data-explorer.com/data}{Data Explorer}.
-#' @references Petersen AH, Ekstrøm CT (2019). “dataMaid: Your Assistant for Documenting Supervised Data Quality Screening in R.” _Journal of Statistical Software_, *90*(6), 1-38. doi: 10.18637/jss.v090.i06 ( \url{https://doi.org/10.18637/jss.v090.i06}).
+#' \href{https://www.data-explorer.com/data/}{Data Explorer}.
+#' @references Petersen AH, Ekstrøm CT (2019). “dataMaid: Your Assistant for Documenting Supervised Data Quality Screening in R.” _Journal of Statistical Software_, *90*(6), 1-38. doi: 10.18637/jss.v090.i06 ( \doi{10.18637/jss.v090.i06}).
 #' @examples
 #' data(bigPresidentData)
 #'
@@ -97,7 +97,7 @@
 #'    \item{Movement}{The artistic movement(s) the painting belongs to.}
 #'}
 #' @source Semi-artificial dataset constructed based on the Master Works of Art dataset available from 
-#' \href{http://www.data-explorer.com/data}{Data Explorer}.
[…]

### 2025-04-13 R/identifyOutliers.R

@@ -39,7 +39,7 @@
 #' \code{$problemValues} (the outlier values).  
 #'
 #' @seealso \code{\link{check}}, \code{\link{allCheckFunctions}}, 
-#' \code{\link{checkFunction}}, \code{\link{checkResult}}, \code{\link{mc}}
+#' \code{\link{checkFunction}}, \code{\link{checkResult}}, \code{\link[robustbase]{mc}}
 #'
 #' @examples
 #'  identifyOutliers(c(1:10, 200, 200, 700))

### 2025-04-13 R/makeCodebook.R

@@ -24,7 +24,7 @@
 #'
 #' @param ... Additional parameters passed to \code{makeDataReport}.
 #' 
-#' @references Petersen AH, Ekstrøm CT (2019). “dataMaid: Your Assistant for Documenting Supervised Data Quality Screening in R.” _Journal of Statistical Software_, *90*(6), 1-38. doi: 10.18637/jss.v090.i06 ( \url{https://doi.org/10.18637/jss.v090.i06}).
+#' @references Petersen AH, Ekstrøm CT (2019). “dataMaid: Your Assistant for Documenting Supervised Data Quality Screening in R.” _Journal of Statistical Software_, *90*(6), 1-38. doi: 10.18637/jss.v090.i06 ( \doi{10.18637/jss.v090.i06}).
 #' @export
 makeCodebook <- function(data, vol="", reportTitle=NULL, file=NULL, ...) {
 

### 2025-04-13 R/makeDataReport.R

@@ -151,7 +151,7 @@
 #' @return The function does not return anything. Its side effect (the production
 #' of a data report) is the reason for running the function.
 #'
-#' @references Petersen AH, Ekstrøm CT (2019). “dataMaid: Your Assistant for Documenting Supervised Data Quality Screening in R.” _Journal of Statistical Software_, *90*(6), 1-38. doi: 10.18637/jss.v090.i06 ( \url{https://doi.org/10.18637/jss.v090.i06}).
+#' @references Petersen AH, Ekstrøm CT (2019). “dataMaid: Your Assistant for Documenting Supervised Data Quality Screening in R.” _Journal of Statistical Software_, *90*(6), 1-38. doi: 10.18637/jss.v090.i06 ( \doi{10.18637/jss.v090.i06}).
 #'
 #' @examples
 #' data(testData)

### 2025-04-13 R/smartNum.R

@@ -18,10 +18,16 @@ smartNum <- function(v) {
 
 #Get the original class of a smartNum or fakeLabelled object. Used in makeDataReport().
 oClass <- function(v) UseMethod("oClass")
+
+#' @exportS3Method
 oClass.default <- function(v) {
   oC <- attr(v, "orginalClass")
   if (!is.null(oC)) return(oC)
   else class(v)
 }
+
+#' @exportS3Method
 oClass.smartNum <- function(v) attr(v, "originalClass")
+
+#' @exportS3Method
 oClass.fakeLabelled <- function(v) attr(v, "originalClass")

### 2025-04-13 R/summarize.R

@@ -47,7 +47,7 @@
 #' If \code{v} is a full dataset: A list of matrices as described above, one for each
 #' variable in the dataset.
 #'
-#' @references Petersen AH, Ekstrøm CT (2019). “dataMaid: Your Assistant for Documenting Supervised Data Quality Screening in R.” _Journal of Statistical Software_, *90*(6), 1-38. doi: 10.18637/jss.v090.i06 (\url{https://doi.org/10.18637/jss.v090.i06}).
+#' @references Petersen AH, Ekstrøm CT (2019). “dataMaid: Your Assistant for Documenting Supervised Data Quality Screening in R.” _Journal of Statistical Software_, *90*(6), 1-38. doi: 10.18637/jss.v090.i06 (\doi{10.18637/jss.v090.i06}).
 #'
 #' @seealso \code{\link{setSummaries}},
 #' \code{\link{summaryFunction}}, \code{\link{allSummaryFunctions}}, 
@@ -136,7 +136,7 @@ defaultCharacterSummaries <- function(remove = NULL, add = NULL) {
 #'
 #' @return A list of function names (as character strings).
 #'
-#' @seealso code{\link{variableType}}, \code{\link{countMissing}}, \code{\link{uniqueValues}},
+#' @seealso \code{\link{variableType}}, \code{\link{countMissing}}, \code{\link{uniqueValues}},
 #' \code{\link{centralValue}}
 #'
 #' @examples

### 2025-04-13 R/visualize.R

@@ -32,7 +32,7 @@
 #' A user defined visual function can be supplied using its function name. Details on how 
 #' to construct valid visual functions are found in \code{\link{visualFunction}}.
 #' 
-#' @references Petersen AH, Ekstrøm CT (2019). “dataMaid: Your Assistant for Documenting Supervised Data Quality Screening in R.” _Journal of Statistical Software_, *90*(6), 1-38. doi: 10.18637/jss.v090.i06 ( \url{https://doi.org/10.18637/jss.v090.i06}).
+#' @references Petersen AH, Ekstrøm CT (2019). “dataMaid: Your Assistant for Documenting Supervised Data Quality Screening in R.” _Journal of Statistical Software_, *90*(6), 1-38. doi: 10.18637/jss.v090.i06 ( \doi{10.18637/jss.v090.i06}).
 #'
 #' @examples
 #'  #Standard use: Return standalone code for plotting a function:
